### 2018-07-02 Manuscript/ManuscriptMarkdown.Rmd
Manuscript Update

---
title: "Working title: Regression models for Cylindrical data."
author: |
  | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
  | $^1$Department of Methodology and Statistics, Utrecht University
  | $^2$TNO
  | $^3$Department of Education, Utrecht University
  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University

bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
keywords: bla vla
documentclass: article
pandoc_args: --natbib
abstract: |
  This is the abstract
---
```{r inlude = FALSE, echo = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
```

\section{Introduction}\label{Introduction}

Cylindrical data are data that consist of a linear component and a directional 
component (a set of angles). In this paper, the directional component is 
circular meaning that it consists of a single angle instead of a set of angles. 
Figure \ref{} \textcolor{red}{NOG MAKEN} shows the difference between a circular variable (left) and a 
linear variable (right). The most obviuous difference is that on a circular 
scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in fact 
represent the same number while on a linear scale the two ends, $-\infty$ and 
$\infty$ are not connected. Cylindrical data occur in several fields of 
research, such as for instance in meteorology [@garcia2013exploring], ecology 
[@garcia2014test] or marine research [@lagona2015hidden]. Several types of data 
in psychological research are also of a cylindrical nature. For example, in 
research on human navigation in the field of cognitive psychology both distance,
a linear variable, and direction, a circular variable, of movement are of 
interest [@chrastil2017rotational]. In eye-tracking research, which is used for 
investigating cognitive processes we can also speak of cylindrical data. When 
measuring eye-movements we speak of the eye-movements themselves, the saccades, 
and fixations, the periods of time between movements when the eyes are still and
looking at one point. Of the saccades both the direction, a circular variable, 
and the duration, a linear variable, are of interest (for a review of 
eye-tracking research see @rayner200935th). Data from circumplex measurement
instruments, e.g. the interpersonal circumplex as used in personality
psychology, are also of a cylindrical nature (see Section \ref{Example} for a
more detailed explanation). In this paper we will discuss how a correct
statistical treatment of such cylindrical data can lead to new insights. In
particular we will show how cylindrical models pave the way for circular-linear
and linear-circular regression. We will do this for a motivating example from
the field of educational psychology. We will refer to this example data as the
teacher data. In this example, apart from modelling the relation between the
linear and circular component of a cylindrical variable we would also like to
predict the two components from a set of covariates in a regression model. The
teacher data will be further introduced in Sections \ref{Example} and
\ref{DataAnalysis}.

As is the case for circular data, the analysis of cylindrical data requires 
special methods. For an overview of methods for circular data see
@mardia2000directional and @jammalamadaka2001topics and for modern treatments
@ley2017modern. In the literature several methods have been put forward to
model the relation between the linear and circular component of a cylindrical 
variable. Some of these are based on regressing the linear component onto the 
circular component using the following type of relation: $$y = \beta_0 + 
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear
component and $\theta$ the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a 
different way, e.g. by specifying a multivariate model for several linear and 
circular variables and modelling their covariance matrix 
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. 
For example, @abe2017tractable introduce a cylindrical distribution  based on a 
Weibull distribution for the linear component and a sine-skewed von Mises 
distribution for the circular component and link these through their shape and 
concentration parameters respectively. However, none of the methods that have 
been proposed thus far include additional covariates on which both the circular 
and linear component are regressed.

Our aim in this paper is to fill this gap in the literature by  adapting four of
the existing methods for the analysis of cylindrical data in such a way that
they include a regression of both the linear and circular component of a
cylindrical variable onto a set of covariates. First however, we will introduce
the teacher data in Section \ref{Example}. In Section \ref{Models} we introduce 
the four modified models for cylindrical data that we use to analyze the data from
the motivating example. We also choose a model selection criterion to compare
the models. The teacher data will be analysed in Section \ref{DataAnalysis}. The
paper will be concluded with a discussion in Section \ref{Discussion}. The
Appendix contains technical details.

\section{Teacher data}\label{Example}

The motivating example for this article comes from the field of educational 
psychology and was collected for the studies of @pennings2017interpersonal,
@Claessens2016side and @vanderWant2015role. These studies concern research on 
classroom climate. An indicator of classroom climate is the students'
perceptions of their teachers' interpersonal behavior. These interpersonal
perceptions, both in educational psychology as well as in other areas of
psychology, can be measured using circumplex measurement instruments (see
@horowitz2010handbook for an overview of many such instruments).

The Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] is
one such circumplex measurement instrument that can be used to study student
perceptions of their teachers' interpersonal behavior. This instrument contains
items that load on two dimensions: Agency and Communion. Agency refers to the
degree of power or control a teacher exerts in interaction with his/her
students. Communion refers to the degree of friendliness or affiliation a
teacher conveys in interaction with his/her students. The loadings on the two 
dimensions of the QTI can be placed in a two-dimensional space formed by Agency
(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
this space are characterized by different teacher behavior, e.g. 'helpful' or
'uncertain'. We call the two-dimensional space the interpersonal circle (IPC).
The idea is that the IPC is ``a continuous order with no beginning or end''
[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
the IPC is therefore often called the interpersonal circumplex. The ordering
also implies that scores on the IPC could be viewed as a circular variable.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{IPC-T.png}
\caption{The interpersonal circle for teachers (IPC-T). The words presented in
the circumference of the circle are anchor words to describe the type of
behavior located in each part of the IPC.}
\label{QTI}
\end{figure}

@Cremers2018Assessing show how data from the IPC can be
considered circular data and analyzed as such using a regression model. The 
two-dimension scores Agency and Communion can be converted to a circular score 
using the two-argument arctangent function in Equation \eqref{PredVal}, where 
$A$ represents a score on the Agency dimension and $C$ represents a score on the
Communion dimension. The resulting circular variable $\theta$ can then be
modeled and for example be regressed onto covariates in a circular regression 
model. However, when two-dimensional data are converted to the circle we lose 
some information, namely the length of the two-dimensional vector $(A, C)$, 
\emph{i.e.}, its euclidean norm $\mid\mid (A, C) \mid\mid$. This length 
represents the strength of the type of interpersonal behavior a teacher shows 
towards his/her students. In a cylindrical model we are able to incorporate this
information, and model a circular variable $\theta$ together with a linear 
variable corresponding to $\mid\mid (A, C) \mid\mid$. This will lead to an 
improved analysis of data from the IPC.  In the next section we will introduce 
several models that we can use for a more accurate and more informative regression
analysis on the teacher data. More information on the teacher data will be given
in Section \ref{DataAnalysis}.

\begin{equation}\label{PredVal}
\theta          = \text{atan2}\left(A, \: C\right)  =
\left|{\begin{array}{lcl}
                                                                       \arctan\left(\frac{A}{C}\right) & \text{if}  \quad&C > 0 \\
\arctan\left(\frac{A}{C}\right) + \pi & \text{if}  \quad& C  <  0  \:\: \&\:\: A \geq 0\\
 \arctan\left(\frac{A}{C}\right) - \pi & \text{if}  \quad&C  <  0 \:\:  \&\:\:A  < 0\\
 +\frac{\pi}{2} & \text{if}  \quad& C  =  0  \:\: \&\:\:A > 0\\
 -\frac{\pi}{2} & \text{if}  \quad& C =  0  \:\: \&\:\:A < 0\\
 \text{undefined} & \text{if} \quad& C =  0   \:\: \&\:\:A = 0.
 \end{array}}
\right.
\end{equation}

```{r loaddata, return = FALSE, echo = FALSE}
Dat <- read_spss("Data Heleen/mergeddata.sav")

#only select the first measurement occasion
Dat <- subset(Dat, Time == 0)

#sort on class size
Dat <- Dat[order(-Dat$nklas), ]

#Now remove duplicates 
#(because dataframe is sorted on class size the smallest classes are removed)

duplicates <- duplicated(Dat[,c("docnr", "Time")])
Dat <- Dat[!duplicates,]

#center and remove missings for self-efficacy
Dat <- Dat %>% drop_na(Efficacy_SE)
Dat$SEc <- Dat$Efficacy_SE- mean(Dat$Efficacy_SE)

set.seed(50)

#Create holdout and test sample
Dat.hold <- sample_n(Dat, 15)
Dat.test <- Dat[-as.numeric(rownames(Dat.hold)),]

#Compute circular and linear outcome
theta.test <- atan2(Dat.test$lAgency, Dat.test$lcommunion)
y.test     <- sqrt(Dat.test$lAgency^2 + Dat.test$lcommunion^2)

theta.hold <- atan2(Dat.hold$lAgency, Dat.hold$lcommunion)
y.hold     <- sqrt(Dat.hold$lAgency^2 + Dat.hold$lcommunion^2)

```

\section{Four cylindrical regression models}\label{Models}

In this section we will introduce four cylindrical regression models. We will 
extend the models from @mastrantonio2018joint and @abe2017tractable to contain 
predictors for the linear and circular outcomes in the cylindrical model. 
Additionally we introduce two models where both variates $\Theta$ and $Y$ are
predicted by covariates and that model the relation between the linear and
circular outcome as follows (following @wang2015joint and
@mastrantonio2015bayesian):

\begin{equation}\label{circlinlink}
y = \gamma_0 + \gamma_{cos}*\cos(\theta)*r + \gamma_{sin}*\sin(\theta)*r + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}

where $r$ will be introduced in Section \ref{CL-(G)PN}, the error term $\epsilon
\sim N(0, \sigma)$, $\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, 
\gamma_q$ are the intercept and regression coefficients and   $x_1, \dots, x_q$
are the $q$ covariate values. In this model $Y$ follows a normal distribution
and $\Theta$ a projected normal (PN) or general projected normal (GPN)
distribution on the circle.

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

In both of these models the relation between the circular variable $\Theta \in
[0, 2\pi)$ and linear variable $Y \in (-\infty, + \infty)$ is specified as in
Equation \ref{circlinlink}. The distribution of $Y$ conditional on $\Theta$ is
then given as:

\begin{equation}\label{ycondtheta}
f(y \mid \theta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[\frac{c^2 + (y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))^{2} - 2c(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))}{2\sigma^2}\right],
\end{equation}

where $c = \begin{bmatrix} r \cos \theta \\  r\sin \theta \end{bmatrix}^t 
\begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}$, $r \geq 0$, 
$\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q$ are the
intercept and regression coefficients and $\sigma^2 \geq 0$ is the error
variance. The linear outcome thus has a normal distribution, conditional on
$\Theta$.

For the circular outcome we assume either a projected normal (PN) or a
general projected normal (GPN) distribution. These distributions arise from the
radial projection of a distribution defined on the plane onto the circle. The
relation between a bivariate variate $\boldsymbol{S}$ in the plane and the
circular outcome $\Theta$ is defined as follows:

\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos \Theta \\  R\sin \Theta \end{bmatrix},
\end{equation}

where $R = \mid\mid S \mid\mid$, the euclidean norm of the bivariate vector 
$\boldsymbol{S}$. In the PN distribution we assume $\boldsymbol{S} \sim 
N_2(\boldsymbol{\mu}, \boldsymbol{I})$ and in the GPN we assume $\boldsymbol{S} 
\sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})$ where $\boldsymbol{\Sigma} = 
\begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$, $\rho \in 
(-\infty, +\infty)$ and $\tau^2 \geq 0$ (as in @hernandez2016general).

Following @nunez2011bayesian, the joint density of $\Theta$ and $R$ for the PN
distribution in a regression set-up equals:

\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = [2\pi]^{-1} \exp\left[ \frac{-r^2 - \boldsymbol{\mu}^2\boldsymbol{\mu} + 2r\boldsymbol{u}^t\boldsymbol{\mu}}{2}\right],
\end{equation}

In a regression setup the outcome $\theta_i$ for each individual $i = 1,
\dots, n$, where $n$ is the sample size, is generated independently from
Equation (\ref{pnreg}). The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is 
defined as $\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and $\boldsymbol{B} = 
(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ contain the regression 
coefficients. Note however that the dimensions of $\boldsymbol{\beta}^{I}$ and 
$\boldsymbol{\beta }^{I}$ need not necessarily be the same and we are thus 
allowed to have a different set of predictor variables and vectors 
$\boldsymbol{z}_i^I$ and $\boldsymbol{z}_i^{II}$ for the two components of 
$\boldsymbol{\mu}_i$.

Following @wang2012directional and @hernandez2016general the joint density of
$r$ and $\theta$ for the GPN distribution equals:

\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = r(2\pi\tau)^{-1} \exp\left[ -0.5 \sigma^2(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})\right],
\end{equation}

where $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2  & \rho\\ \rho & 1 
\end{bmatrix}$, $\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta 
\end{bmatrix}$. In a regression setup the outcome $\theta_i$ for each individual
$i = 1, \dots, n$, where $n$ is the sample size, is generated independently
from Equation (\ref{gpnreg}). The mean vector $\boldsymbol{\mu}_i \in 
\mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
\boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. The vector 
$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and each 
$\boldsymbol{\beta}^{k}$ is a vector with two regression coefficients, one for 
each of the two components of $\boldsymbol{\mu}_i$. Note that for the CL-GPN 
model we do need to have the same predictors for both components of 
$\boldsymbol{\mu}_i$.

Both cylindrical models introduced in this section are estimated using MCMC 
methods. These methods are based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in
Appendices \ref{A1} and \ref{A2}.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced by
@abe2017tractable to the regression context. The density of the model for the
pair of the circular variable $\theta \in [0, 2\pi)$ and linear variable $y$ on
the positive real half-line $[0, + \infty)$ is:

\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha(\beta)^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-((\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu)))],
\end{equation}

In a regression setup the outcome vector $(\theta_i, y_i)^t$ for each individual
$i = 1, \dots, n$, where $n$ is the sample size, is generated independently from
Equation (\ref{WeiSSVMdensity}). The parameter $\alpha > 0$ is a linear shape 
parameter, $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ is a linear 
scale parameter, $\mu_i = \eta_0 + 
2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$ is a circular 
location parameter, $\kappa > 0$ is a cicular concentration parameter and 
$\lambda \in [-1, 1]$ is a circular skewness parameter. The parameter 
$\boldsymbol{\nu}$ is a vector of $q$ regression coefficients and intercept
$\nu_j \in (-\infty, +\infty)$ where $j = 0, \dots, q$, for the prediction of
$y$. The parameter $\eta_0 \in [0, 2\pi)$ is the intercept and
$\boldsymbol{\eta}$, is a vector of $p$ regression coefficients $\eta_j \in 
(-\infty, +\infty)$, where $j = 1, \dots, p$, for the prediction of $\theta$. 
The vector $\boldsymbol{x}_i$ is a vector of predictor values for the prediction
of $y$ and $\boldsymbol{z}_i$ is a vector of predictor values for the prediction
of $\theta$.

As in @abe2017tractable, the conditional distribution of $y$ given $\theta$ is a
Weibull distribution with scale $\alpha$ and shape
$\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}$ and the conditional 
distribution of $\theta$ given $y$ is a sine skewed von Mises distribution with 
location parameter $\mu$ and concentration parameter $(\beta 
y)^\alpha\tanh(\kappa)$.

The log-likelihood for this model equals:

\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \ln(\exp(\boldsymbol{x}_i^t\boldsymbol{\nu})) \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})))
\end{align}

We can use numerical optimization (Nelder-Mead) to find solutions for the
maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains $p$
circular outcomes and $q$ linear outcomes. The circular outcomes 
$\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ 
are modeled together by a multivariate GPN distribution and the linear outcomes 
$\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled 
together by a multivariate skew normal distribution [@sahu2003new]. Because the
GPN distribution is modelled using a so-called augmented representation (as in
(\ref{projection}) and (\ref{gpnreg})) it is convenient to use a similar tactic
for modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint we use the following representation of the linear
outcomes:

$$\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},$$
where where $\boldsymbol{\mu}_y$ is a mean vector for the linear outcome 
$\boldsymbol{Y}$, $\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})$ is a $q 
\times q$ diagonal matrix with diagonal elements $\boldsymbol{\lambda} = 
\lambda_1, \dots, \lambda_q$ (the parameters that indicate the skewness of the 
linear outcomes), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
\boldsymbol{I}_q)$, a q-dimensional half normal distribution
[@olmos2012extension] and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
\boldsymbol{\Sigma}_y)$. This means that conditional on the auxiliary data
$\boldsymbol{D}$, $\boldsymbol{Y}$ is normally distributed with mean
$\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}$ and covariance matrix
$\boldsymbol{\Sigma}_y$. The joint density for $(\boldsymbol{Y},
\boldsymbol{D})^t$ is defined as follows:

\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_q(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_1(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q).
\end{equation}

As in @mastrantonio2018joint dependence between the linear and circular outcome
is created by modelling the augmented representations of $\boldsymbol{\Theta}$,
and $\boldsymbol{Y}$ together in a $2 \times p + q$ dimensional normal
distribution. The joint density of the model is then represented by:

\begin{equation}\label{YDThetarjoint}
f(\boldsymbol{\theta}, \boldsymbol{r}, \boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_{2p+q}((\boldsymbol{s}, \boldsymbol{y})^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2p}, diag(\boldsymbol{\lambda})\boldsymbol{d})^t, \boldsymbol{\Sigma}) \phi_q(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q) \prod_{j = 1}^{p}r_j,
\end{equation}

where the mean vector $\boldsymbol{\mu} = (\boldsymbol{\mu}_s,
\boldsymbol{\mu}_y)^t$ and $\boldsymbol{\Sigma} = \left ( \begin{matrix}
\boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} &
\boldsymbol{\Sigma}_y \\ \end{matrix} \right )$. The matrix
$\boldsymbol{\Sigma}_s$ is the covariance matrix for the variances of and
covariances between the augmented representations of the circular outcome and
the matrix $\boldsymbol{\Sigma}_{sy}$ contains covariances between the augmented
representations of the circular outcome and the linear outcome.

In our regression extension we have $i = 1, \dots, n$ observations of $p$ 
circular outcomes, $q$ linear outcomes and $k$ covariates. The mean in the 
density in (\ref{YDThetarjoint}) then becomes $\boldsymbol{\mu}_i = 
\boldsymbol{x}_i\boldsymbol{B}$ where $\boldsymbol{B}$ is a $k \times (2 
\times p + q)$ matrix with regression coeffients. We estimate the model using 
MCMC methods. A detailed description of these methods is given in Appendix
\ref{A3}.

\subsection{Model fit criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample predictive 
performance to determine the fit of the model. To do so we split our data in a
training and holdout set (10 $\%$ of the sample). A proper criterion to compare
out-of-sample predictive performance is the the Predictive Log Scoring Loss
(PLSL) [@gneiting2007strictly]. The lower the value of this criterion, the
better the predictive performance of the model. Using ML estimates this
criterion can be computed as follows:

\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),
\end{equation}

where $l$ is the model likelihood, $M$ is the sample size of the holdout set, 
$x_i$ is the $i^{th}$ datapoint from the holdout set and 
$\hat{\boldsymbol{\vartheta}}$ are the ML estimates of the model parameters. 
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) as outlined in @BDA and can be computed as:

\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),
\end{equation}

where $B$ is the amount of posterior samples and $\boldsymbol{\vartheta}^{(j)}$
are the posterior estimates of the model parameters for the $j^{th}$ iteration.
Note that although we fit the CL-PN, CL-GPN and joint GPN-SSN models using
Bayesian statistics, we do not take prior information (only the likelihood) into
account when assessing model fit. According to @BDA this is not necessary since
we are assessing the fit of a model to data only \textcolor{red}{AFMAKEN!}. For 
each of the cylindrical models we can then compute a PLSL for the circular and 
linear outcome by using the conditional log-likelihoods of the respective 
outcome.

\section{Data Analysis}\label{DataAnalysis}

In this section we will analyze the teacher dataset with the help of
the four cylindrical models introduced in Section \ref{Models}. We will however
first give a more detailed description of the dataset.

```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling informative Prior CL-PN.R")

its <- 20000
set.seed(101)

ZI.PN.test  <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
ZII.PN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.PN.test   <- as.matrix(cbind(rep(1, length(theta.test)),
                               cos(theta.test),
                               sin(theta.test),
                               Dat.test$SEc))

ZI.PN.hold  <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
ZII.PN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.PN.hold   <- as.matrix(cbind(rep(1, length(theta.hold)),
                               cos(theta.hold),
                               sin(theta.hold),
                               Dat.hold$SEc))

res_CLPN <- CLPN(theta.test, y.test, X.PN.test, ZI.PN.test, ZII.PN.test, its,
                 theta.hold, y.hold, X.PN.hold, ZI.PN.hold, ZII.PN.hold)
```

```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling CL-GPN.R")

its <- 20000
set.seed(101)

Z.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)),
                              cos(theta.test),
                              sin(theta.test),
                              Dat.test$SEc))

Z.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)),
                              cos(theta.hold),
                              sin(theta.hold),
                              Dat.hold$SEc))

res_CLGPN <- CLGPN(theta.test, y.test, X.GPN.test, Z.GPN.test, its, p = 2, 
                   theta.hold, y.hold, X.GPN.hold, Z.GPN.hold)

```

```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling CL-GPN_multivariate.R")

its <- 20000
set.seed(101)

X.MGPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))

X.MGPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))

res_CLGPNM <- CLGPNM(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
                     theta.hold, y.hold, X.MGPN.hold)

```

```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("WeisSVM optimization.R")

set.seed(101)

Z.test     <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.test    <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))

Z.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))

colnames(X.test) <- c("ax", "bx")
colnames(Z.test) <- c("az", "bz")
colnames(X.hold) <- c("ax", "bx")
colnames(Z.hold) <- c("az", "bz")

dat.t <- as.data.frame(cbind(theta.test, y.test, X.test, Z.test))
dat.h <- as.data.frame(cbind(theta.hold, y.hold, X.hold, Z.hold))

ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
               c(-1, 0, 0, 0, 0, 0, 0),
               c(0, 0, 0, 0, 1, 0, 0),
               c(0, 0, 0, 0, 0, 1, 0),
               c(0, 0, 0, 0, 0, 0, 1),
               c(0, 0, 0, 0, 0, 0, -1))

ci.reg = c(-pi, -pi, 0, 0, -1, -1)

param <- c(1,1,1,1,1,1,0)*0.5

ui.reg %*% param - ci.reg

resAL <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
                     control = list(maxit = 1000000, fnscale = -1), data = dat.t)

```

\subsection{Data description}\label{DataDescriptives}

The teacher data was collected between 2010 and 2015 and contains several repeated
measures on the IPC of 161 teachers gathered for the studies of
@vanderWant2015role, @Claessens2016side and @pennings2017interpersonal. 
The measurements were obtained using the QTI and taken in different years and in
different classes. For this paper we only consider one measurement, namely the
one for the first measurement occasion (2010) and the one for the largest class
if data for multiple classes were available. In addition to a variable `IPC`
containing the circular outcome and the 'length' of the IPC, the linear outcome,
a teachers' self-efficacy (\verb|SE|) will be used as covariate in the analysis.
After the removal of missing cases we end up with a sample of 148 teachers Table
\ref{Tableteacherdescriptives} shows descriptives for the dataset.

```{r descripitves, echo = FALSE, results = FALSE}
#sample size
nrow(Dat)

#summary statistics
summary(Dat$Efficacy_SE)
summary(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
summary(as.circular(atan2(Dat$lAgency, Dat$lcommunion)%%(2*pi)))

sd(Dat$Efficacy_SE)
sd(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))

```

\begin{table}
\centering
\caption{Descriptives for the teacher dataset} 
\begin{tabular}{lrrrl}
  \noalign{\smallskip}\hline\noalign{\smallskip}
Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
IPC &33.23$^\circ$& 0.76 & - & Circular\\
length IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
SE & 4.72 & 0.87 & 2 - 6.5 & Linear\\
   \hline
\end{tabular}
\label{Tableteacherdescriptives}
\caption*{Note that $\hat{\rho}$ is an sample estimate for the circular concentration where a value of 0 means that the data is not concentrated at all, i.e. spread over the entire circle, and a value of 1 means that all data is concentrated at a single point on the circle. }
\end{table}

\subsection{Models}\label{DataModels}

The regression equations for the linear and the circular outcome of the four
cylindrical models fit to the teacher dataset are as follows:

\begin{itemize}
\item For the modified CL-PN and CL-GPN models:

$\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
  \mu_{i}^{I}  \\
\mu_{i}^{II}
 \end{pmatrix}=\begin{pmatrix}
  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
  \beta_0^{II} + \beta_1^{II}\text{SE}_i
 \end{pmatrix},$

$\hat{y}_i = \gamma_0 + \gamma_{cos}\cos\theta_ir_i + \gamma_{sin}\sin\theta_ir_i + \gamma_1\text{SE}_i.$

\item For the modified Abe-Ley model:

$\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i),$

$\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).$

\item For the modified joint projected and skew normal model:

$\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 + \boldsymbol{\beta}_1\text{SE}_i$, where $\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t$, $\boldsymbol{\beta}_0 = (\beta_{0_s^{I}}, \beta_{0_s^{II}},\beta_{0_y})$ and $\boldsymbol{\beta}_1 = (\beta_{1_s^{I}}, \beta_{1_s^{II}},\beta_{1_y})$.
\end{itemize}

The variable \verb|SE| is thus used as a predictor for the linear and circular 
outcomes in all four cylindrical models. We will use the loglikelihoods of the
following conditional densities \textcolor{red}{CHRISTOPHE, MAG HIER WEL $\sim$ GEBRUIKT WORDEN?}for the computation of the PLSL criterion to
evaluate the model fit:

\begin{itemize}
\item for the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{pnreg}).

\item for the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{gpnreg}).

\item for the modified Abe-Ley model:

$y_i \mid \boldsymbol{\theta}_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item for the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \boldsymbol{\theta}_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_w^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \boldsymbol{\Sigma}_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1} (y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.

\end{itemize}

\subsection{Results \& Analysis}\label{DataResults}

Before analysis there were a couple of settings that we had to specify for the 
cylindrical models. Starting values for the WeiSSVM model were the following 
$\eta_0 = 0.5, \eta_1 = 0.5, \nu_0 = 0.5, \nu_1 = 0.5, \kappa = 0.5, \alpha = 
0.5, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
was set to 2000. After we checked convergence via traceplots we concluded that 
some of the parameters of the multivariate GPN model did not converge. Therefore
we set the amount of iterations of the MCMC models to 20,000 and subtracted a 
burn-in of 5000. With this specification the MCMC chains for all parameters 
converge. Note that we choose the same amount of iterations for all three 
Bayesian cylindrical models to make their comparison via the PLSL as fair as 
possible. Lastly, the predictor \verb|SE| was centered before inclusion in the
analysis.

```{r estCLPN, echo = FALSE}

CLPN.Gamma.m   <- apply(res_CLPN$Gamma[5000:20000,], 2, mode_est)
CLPN.Gamma.hpd <- apply(res_CLPN$Gamma[5000:20000,], 2, hpd_est)

CLPN.BI.m   <- apply(res_CLPN$BI[5000:20000,], 2, mode_est)
CLPN.BI.hpd <- apply(res_CLPN$BI[5000:20000,], 2, hpd_est)

CLPN.BII.m <- apply(res_CLPN$BII[5000:20000,], 2, mode_est)
CLPN.BII.hpd <- apply(res_CLPN$BII[5000:20000,], 2, hpd_est)

CLPN.Sigma.m   <- mode_est(res_CLPN$Sigma[5000:20000])
CLPN.Sigma.hpd <- hpd_est(res_CLPN$Sigma[5000:20000])

modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)

CLPNres.tab <- cbind(modes.PN, hpd.LB.PN, hpd.UB.PN)

```

```{r estCLGPN, echo = FALSE}

CLGPN.Sig.m   <- mode_est(res_CLGPN$Sig)
CLGPN.Sig.hpd <- hpd_est(res_CLGPN$Sig)

CLGPN.Gamma.m   <- apply(res_CLGPN$Gamma[5000:20000,], 2, mode_est)
CLGPN.Gamma.hpd <- apply(res_CLGPN$Gamma[5000:20000,], 2, hpd_est)

CLGPN.BI.m   <- apply(res_CLGPN$BI[5000:20000,], 2, mode_est)
CLGPN.BI.hpd <- apply(res_CLGPN$BI[5000:20000,], 2, hpd_est)

CLGPN.BII.m   <- apply(res_CLGPN$BII[5000:20000,], 2, mode_est)
CLGPN.BII.hpd <- apply(res_CLGPN$BII[5000:20000,], 2, hpd_est)

CLGPN.Sigma11.m   <- mode_est(res_CLGPN$Sigma[1,1,5000:20000])
CLGPN.Sigma11.hpd <- hpd_est(res_CLGPN$Sigma[1,1,5000:20000])
CLGPN.Sigma22.m   <- mode_est(res_CLGPN$Sigma[2,2,5000:20000])
CLGPN.Sigma22.hpd <- hpd_est(res_CLGPN$Sigma[2,2,5000:20000])
CLGPN.Sigma12.m   <- mode_est(res_CLGPN$Sigma[1,2,5000:20000])
CLGPN.Sigma12.hpd <- hpd_est(res_CLGPN$Sigma[1,2,5000:20000])

modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
                CLGPN.Gamma.m, CLGPN.Sig.m,
                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])

CLGPNres.tab <- cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN)
CLPNGPNres.tab <- round(cbind(CLPNres.tab, CLGPNres.tab), 2)

rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_0^{I}$",
                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
                              "$\\gamma_0$", "$\\gamma_{cos}$",
                              "$\\gamma_{sin}$", "$\\gamma_1$",
                              "$\\sigma$", "$\\sum_{1,1}$",
                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
colnames(CLPNGPNres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE,
      caption = "Results for the modified CL-PN and CL-GPN model")%>%
add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))

```

```{r estSL, echo = FALSE}

ALres.tab <- matrix(round(c(resAL$par[1:4], resAL$par[6], resAL$par[5], resAL$par[7]), 2),7,1)
rownames(ALres.tab) <- c("$\\eta_0$", "$\\eta_1$",
                         "$\\nu_0$", "$\\nu_1$",
                         "$\\alpha$", "$\\kappa$", "$\\lambda$")

kable(ALres.tab, "latex", booktabs = T, escape = FALSE,
      caption = "Results for the modified Abe-Ley model")%>%
add_header_above(c("Parameter" = 1, "ML-estimate" = 1))

```

```{r estCLMGPN, echo = FALSE}

CLGPNM.lambda.m   <- mode_est(res_CLGPNM$lambda[5000:20000, ])
CLGPNM.lambda.hpd <- hpd_est(res_CLGPNM$lambda[5000:20000, ])

CLGPNM.Beta11.m <- mode_est(res_CLGPNM$Beta[1,1,5000:20000])
CLGPNM.Beta21.m <- mode_est(res_CLGPNM$Beta[2,1,5000:20000])
CLGPNM.Beta12.m <- mode_est(res_CLGPNM$Beta[1,2,5000:20000])
CLGPNM.Beta22.m <- mode_est(res_CLGPNM$Beta[2,2,5000:20000])
CLGPNM.Beta13.m <- mode_est(res_CLGPNM$Beta[1,3,5000:20000])
CLGPNM.Beta23.m <- mode_est(res_CLGPNM$Beta[2,3,5000:20000])

CLGPNM.Beta11.hpd <- hpd_est(res_CLGPNM$Beta[1,1,5000:20000])
CLGPNM.Beta21.hpd <- hpd_est(res_CLGPNM$Beta[2,1,5000:20000])
CLGPNM.Beta12.hpd <- hpd_est(res_CLGPNM$Beta[1,2,5000:20000])
CLGPNM.Beta22.hpd <- hpd_est(res_CLGPNM$Beta[2,2,5000:20000])
CLGPNM.Beta13.hpd <- hpd_est(res_CLGPNM$Beta[1,3,5000:20000])
CLGPNM.Beta23.hpd <- hpd_est(res_CLGPNM$Beta[2,3,5000:20000])

CLGPNM.Betacon11.m <- mode_est(res_CLGPNM$Betacon[1,1,5000:20000])
CLGPNM.Betacon21.m <- mode_est(res_CLGPNM$Betacon[2,1,5000:20000])
CLGPNM.Betacon12.m <- mode_est(res_CLGPNM$Betacon[1,2,5000:20000])
CLGPNM.Betacon22.m <- mode_est(res_CLGPNM$Betacon[2,2,5000:20000])
CLGPNM.Betacon13.m <- mode_est(res_CLGPNM$Betacon[1,3,5000:20000])
CLGPNM.Betacon23.m <- mode_est(res_CLGPNM$Betacon[2,3,5000:20000])

CLGPNM.Betacon11.hpd <- hpd_est(res_CLGPNM$Betacon[1,1,5000:20000])
CLGPNM.Betacon21.hpd <- hpd_est(res_CLGPNM$Betacon[2,1,5000:20000])
CLGPNM.Betacon12.hpd <- hpd_est(res_CLGPNM$Betacon[1,2,5000:20000])
CLGPNM.Betacon22.hpd <- hpd_est(res_CLGPNM$Betacon[2,2,5000:20000])
CLGPNM.Betacon13.hpd <- hpd_est(res_CLGPNM$Betacon[1,3,5000:20000])
CLGPNM.Betacon23.hpd <- hpd_est(res_CLGPNM$Betacon[2,3,5000:20000])

CLGPNM.Sigma11.m <- mode_est(res_CLGPNM$Sigma[1,1,5000:20000])
CLGPNM.Sigma12.m <- mode_est(res_CLGPNM$Sigma[1,2,5000:20000])
CLGPNM.Sigma13.m <- mode_est(res_CLGPNM$Sigma[1,3,5000:20000])
CLGPNM.Sigma22.m <- mode_est(res_CLGPNM$Sigma[2,2,5000:20000])
CLGPNM.Sigma23.m <- mode_est(res_CLGPNM$Sigma[2,3,5000:20000])
CLGPNM.Sigma33.m <- mode_est(res_CLGPNM$Sigma[3,3,5000:20000])

CLGPNM.Sigma11.hpd <- hpd_est(res_CLGPNM$Sigma[1,1,5000:20000])
CLGPNM.Sigma12.hpd <- hpd_est(res_CLGPNM$Sigma[1,2,5000:20000])
CLGPNM.Sigma13.hpd <- hpd_est(res_CLGPNM$Sigma[1,3,5000:20000])
CLGPNM.Sigma22.hpd <- hpd_est(res_CLGPNM$Sigma[2,2,5000:20000])
CLGPNM.Sigma23.hpd <- hpd_est(res_CLGPNM$Sigma[2,3,5000:20000])
CLGPNM.Sigma33.hpd <- hpd_est(res_CLGPNM$Sigma[3,3,5000:20000])

CLGPNM.Sigmacon11.m <- mode_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
CLGPNM.Sigmacon12.m <- mode_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
CLGPNM.Sigmacon13.m <- mode_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
CLGPNM.Sigmacon22.m <- mode_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
CLGPNM.Sigmacon23.m <- mode_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
CLGPNM.Sigmacon33.m <- mode_est(res_CLGPNM$Sigmacon[3,3,5000:20000])

CLGPNM.Sigmacon11.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
CLGPNM.Sigmacon12.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
CLGPNM.Sigmacon13.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
CLGPNM.Sigmacon22.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
CLGPNM.Sigmacon23.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
CLGPNM.Sigmacon33.hpd <- hpd_est(res_CLGPNM$Sigmacon[3,3,5000:20000])

modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
                CLGPNM.lambda.m)
hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
                CLGPNM.lambda.hpd[1])
hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
                CLGPNM.lambda.hpd[2])

modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
                   CLGPNM.lambda.m)
hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
                   CLGPNM.lambda.hpd[1])
hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
                   CLGPNM.lambda.hpd[2])

CLGPNMres.tab <- round(cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN, modescon.GPN, hpdcon.LB.GPN, hpdcon.UB.GPN),2)

rownames(CLGPNMres.tab) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
                             "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
                             "$\\beta_{0_s^{II}}$", "$\\beta_{1_y}$",
                             "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
                             "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
                             "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
                             "$\\lambda$")
colnames(CLGPNMres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

kable(CLGPNMres.tab, "latex", booktabs = TRUE, escape = FALSE, caption = "Results for the modified joint projected and skew normal model")%>%
add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))

```

Tables 2, 3 and 4 show results from the four models that were fit to the teacher
dataset. There are several parameters that we can compare between these results.
Firstly, we have the estimated mean of the circular outcome. In Table 5 we see 
the posterior (predictive) distributions for the estimated circular means of the
CL-PN, CL-GPN and joint GPN-SSN  model. For the CL-PN model we can actually also
compute this means from the estimates in Table 2 as follows: $\mu_{circ} = 
atan2(\beta_0^{II}, \beta_0^{I})$. This leads to an estimate of 31.53$^\circ$. 
The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are about equal and
correspond quite well to the actual data mean of 33.23$^\circ$ in Table
\ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model which is
0.37 radians or 21.20$^\circ$ is different. This difference could be caused by
the fact that the densities for the circular ouctome, projected normal or
sine-skewed von Mises, differ between the models.

```{r means, echo = FALSE}
means <- round(rbind(mean_circ(res_CLPN$theta_pred[,5000:20000]),
                     mean_circ(res_CLGPN$theta_pred[,5000:20000]),
                     mean_circ(res_CLGPNM$theta_pred[,5000:20000]))*(180/pi), 2)

rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
colnames(means) <- "Mean (degrees)"
kable(means, "latex", booktabs = T, escape = FALSE, caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models")%>%
   kable_styling() %>%
   add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", notation="alphabet")

```

We can also compare the effect of \verb|SE| on the circular outcome. For the 
CL-PN model and Abe-Ley models we can get estimates of a circular regression 
coefficient of this effect. For the Abe-Ley model this estimate is the parameter
$\nu_1 = -0.01$. This means that for each unit increate in self-efficacy, at the
inflection point of the circular regression line, the score of the teacher on 
the IPC decreases with $0.01*(180/\pi) = 0.57^\circ$. For the CL-PN model we 
can use methods from @CremersMulderKlugkist2017. The estimated posterior mode of
$b_c$, a parameter that is comparable to $\nu_1$ in the Abe-Ley model, equals 
$1.13$ and its 95\% HPD interval is $(-26.55; 25.46)$. This means that for each 
unit increate in self-efficacy, at the inflection point of the circular 
regression line, the score of the teacher on the IPC increases with 
$1.13*(180/\pi) = 74.74^\circ$. Values of $b_c$ and $\nu_1$ are quite 
different even though they should represent the same effect. However, the slope 
at the inflection point, which $b_c$ and $\nu_1$ describe is not necessarily 
representative of the effect of \verb|SE| in the data range. In Figure \ref{} we
see that the inflection point (square) for the CL-PN model lies just ouside of 
the data range and that for the Abe-Ley model even further (outside the x-axis 
range). The slope of both regression lines in the data range is much more alike
(even though one is positive and the other negative). The slope at the mean
Self-efficicay, which has a value of 0 because it was centered, $SAM$ is
estimated at $0.02 (-0.09; 0.15)$ The $SAM$ can be interpreted as the amount of
change on the circle for a unit increase at the mean of the predictor variable.
However, for the CL-PN model the 95\% HPD interval of $SAM$ for the effect of
\verb|SE| includes 0 meaning that the value is not different from 0. For the
Abe-Ley model standard errors of the parameters are not known so we cannot
formally test whether $\nu_1$ differs from 0. Neither is there a known way of
computing the effect at predictor values other than the one of the inflection
point.

For the CL-GPN and multivariate 
CL-GPN models we cannot compute circular regression coefficients. Instead, we 
will compute posterior predictive distributions for the predicted circular
outcome of individuals scoring 1 unit above and 1 unit below the average
self-efficacy.

```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("regline.tex", standAlone =TRUE, height = 4, width = 6, pointsize = 12, engine = "pdftex")

a1 <- mode_est(res_CLPN$BI[5000:20000, 1])
b1 <- mode_est(res_CLPN$BI[5000:20000, 2])
a2 <- mode_est(res_CLPN$BII[5000:20000, 1])
b2 <- mode_est(res_CLPN$BII[5000:20000, 2])

x <- seq(-5, 5, by = 0.1)

pred.AL <- 0.37 + 2*atan(-0.01*x)
pred.PN <- atan2(a2 + b2*x, a1 + b1*x)

plot(Dat.test$SEc, theta.test, bty = "n",
     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "Self-efficacy")
points(x, pred.AL, type = "l", lty = 1)
points(x, pred.PN, type = "l", lty = 2)

dev.off()
tools::texi2dvi("regline.tex", pdf=TRUE)

```

```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}

a1 <- res_CLPN$BI[5000:20000, 1]
b1 <- res_CLPN$BI[5000:20000, 2]
a2 <- res_CLPN$BII[5000:20000, 1]
b2 <- res_CLPN$BII[5000:20000, 2]

ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
ac <- atan2(a2 + b2*ax, a1 + b1*ax)
bc <- -tan(atan2(a2, a1)-ac)/ax

mode_est(ax)

mode_est(bc)
hpd_est(bc)

mode_est(bc/(1 + (bc*(-ax))^2))
hpd_est(bc/(1 + (bc*(-ax))^2))

```

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{regline.pdf}
\caption{Plot showing circular regresion lines as predicted by the Abe-Ley (solid line) and CL-PN model (dashed line).}
\label{regline}
\end{figure}

Next, we compare the intercepts and regression coefficients for the linear outcome.

\begin{itemize}
\item How to interpret weibull regression for Abe-Ley model?
\item Comment on different distributions for linear and circular outcome ($\lambda$ in multivariate CL-GPN), integrate with model fit does this correspond. 
\item Comment on different estimation methods, in CL-PN and CL-GPN the regression of $\theta$ is independent of $y$ but not vice-versa. In the Abe-Ley and Multivariate method dependence between $\theta$ and $y$ is modelled both ways. 
\end{itemize}

\subsubsection{Model fit}

```{r ModelFit, echo = FALSE}

CLGPNM.ll.circ <- -2*mean(res_CLGPNM$ll.circ[5000:20000])
CLGPNM.ll.lin  <- -2*mean(res_CLGPNM$ll.lin[5000:20000])

CLGPN.ll.circ <- -2*mean(res_CLGPN$ll.circ[5000:20000])
CLGPN.ll.lin  <- -2*mean(res_CLGPN$ll.lin[5000:20000])

CLPN.ll.circ <- -2*mean(res_CLPN$ll.circ[5000:20000])
CLPN.ll.lin  <- -2*mean(res_CLPN$ll.lin[5000:20000])

AL.ll.circ <- -2*func.reg.cond.circ(resAL$par, dat.h)
AL.ll.lin <- -2*func.reg.cond.lin(resAL$par, dat.h)

PLSL <- round(rbind(c(CLPN.ll.circ, CLGPN.ll.circ, AL.ll.circ, CLGPNM.ll.circ),
                    c(CLPN.ll.lin, CLGPN.ll.lin, AL.ll.lin, CLGPNM.ll.lin)), 2)
rownames(PLSL) <- c("circular", "linear")
colnames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "Joint GPN-SSN")

kable(PLSL, "latex", booktabs = T, escape = FALSE, caption = "PLSL criteria for the circular and linear outcome in the four cylindrical models")

```

```{r preddist, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

source("Sample Abe-Ley.R")

ALdy <- matrix(NA, length(y.test), 1000)
ALdth <- matrix(NA, length(theta.test), 1000)

ALdy.h <- matrix(NA, length(y.hold), 1000)
ALdth.h <- matrix(NA, length(theta.hold), 1000)

tikz("preddist.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")

layout(matrix(c(1,2,3, 4,5,5), 3, 2, byrow = TRUE),heights=c(2,2,1))

par(mar = c(5,5,1,1))

x.h <- Dat.hold$SEc
x <- Dat.test$SEc

for(i in 1:length(theta.test)){
  beta <- exp(resAL$par[3] + resAL$par[4]*x[i])
  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x[i])

  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
  ALdy[i,] <- dat$x
  ALdth[i,] <- dat$theta
}

for(i in 1:length(theta.hold)){
  beta <- exp(resAL$par[3] + resAL$par[4]*x.h[i])
  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x.h[i])

  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
  ALdy.h[i,] <- dat$x
  ALdth.h[i,] <- dat$theta

}

hist(theta.test, prob = TRUE, ylim = c(0, 1.25),
     main = "", ylab = "Circular", xlab = "$\\theta_{test}$")
lines(density(res_CLPN$theta_pred[,19000:20000]), lty = 1)
lines(density(res_CLGPN$theta_pred[,19000:20000]), lty = 2)
lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 3)
lines(density(ALdth), lty = 4)

hist(theta.hold, prob = TRUE, ylim = c(0, 1.25),
     main = "", ylab = "Circular", xlab = "$\\theta_{holdout}$")
lines(density(res_CLPN$theta_pred.hold[,19000:20000]), lty = 1)
lines(density(res_CLGPN$theta_pred.hold[,19000:20000]), lty = 2)
lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 3)
lines(density(ALdth.h), lty = 4)

hist(y.test, prob = TRUE, ylim = c(0, 3),
     main = "", ylab = "Linear", xlab = "$y_{test}$")
lines(density(res_CLPN$y_pred[,5000:20000]), lty = 1)
lines(density(res_CLGPN$y_pred[,5000:20000]), lty = 2)
lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 3)
lines(density(ALdy), lty = 4)

hist(y.hold, prob = TRUE, ylim = c(0, 4),
     main = "", ylab = "Linear", xlab = "$y_{holdout}$")
lines(density(res_CLPN$y_pred.hold[,5000:20000]), lty = 1)
lines(density(res_CLGPN$y_pred.hold[,5000:20000]), lty = 2)
lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 3)
lines(density(ALdy.h), lty = 4)

par(mar = c(2,2,1,1))

plot(1, type = "n", axes=FALSE, xlab="", ylab="")
legend(x = "top", inset = 0, legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"), lty = c(1,2,4,3), horiz = TRUE)

dev.off()

tools::texi2dvi("preddist.tex",pdf=TRUE)

```

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{preddist.pdf}
\caption{Histograms of $\boldsymbol{y}$ and $\boldsymbol{\theta}$ values of the test and holdout set of the teacher data plotted together with (posterior predictive) density estimates for the modified CL-PN, CL-GPN, Abe-Ley and joint GPN-SSN models.}
\label{preddist}
\end{figure}

\section{Discussion}\label{Discussion}

\begin{itemize}
\item Comment on difficulty of interpreting parameters in GPN model.
\end{itemize}

\newpage
\section*{References}
<div id="refs"></div>

\newpage
\begin{appendices}
\section{Appendix}\label{Appendix}

In this appendix we outline the MCMC procedures to fit the cylindrical regression models from Section \ref{Models}. R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.

\subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with  $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and  $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_2$ for $k \in I,II$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$ and $\boldsymbol{\beta^{k}} = (0,0)$ for $k \in I,II$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior with $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}^k_n, \boldsymbol{\Lambda}^{k}_n)$, with $\boldsymbol{\mu}^k_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)^{-1}(\boldsymbol{\Lambda}^k_0\boldsymbol{\mu}^k_0 + (\boldsymbol{Z}^k)^t\boldsymbol{s}^k)$ and $\boldsymbol{\Lambda}^{k}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)$.
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{(-\frac{1}{2}(r_i)^2 + b_ir_i)}$$ 
where $b_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$ and $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$.

We can sample from this posterior using a slice sampling technique (Cremers et al., 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}) \propto r_{i} \textbf{I}\left(0 < v_i < \exp\left\{ -\frac{1}{2}(r_{i} - b_{i})^2\right\}\right)\textbf{I}(r_i > 0).$$

\noindent The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is:

$$U\left(0, \exp\left\{-\frac{1}{2}(r_{i} -  b _{i})^2\right\}\right)$$

and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(b_{i} + \max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\} < r_{i} < b_{i} + \sqrt{-2\ln v_{i}}\right).$$

\noindent We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $ r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=b_{i} +\max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\}$ and $ r_{i_{2}}= b_{i} + \sqrt{-2\ln v_{i}}$.
\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified CL-GPN mode}\label{A2}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in 1, \dots, p$ where $p$ is the number of covariates in $\boldsymbol{Z}$.
\item Prior for $\rho$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
\item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$, $\rho = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$, $\rho$ and $\tau$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior where $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + 0.5(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in 1, \dots, p$ where p is the number of covariates in $\boldsymbol{Z}$.
\item Posterior for $\rho$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = \boldsymbol{z}_i\boldsymbol{\beta}^{I}$ and $\mu_i^{II} = \boldsymbol{z}_i\boldsymbol{\beta}^{II}$.
\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse-gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{II} + \rho(s^{II}_{i} - \mu^{II})\})^2 + \beta_0$
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$, $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$.

We can sample from this posterior using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$, $\rho$ and $\tau$ for the training dataset. Use the density $f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma})$ for the circular outcome.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure multivariate GPN model}\label{A3}

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort.

\item Define the prior parameters for the training set. Note that in this paper we have only one circular outcome and one linear outcome, so $p = 1$ and $q = 1$. In this paper we use the following priors:

\begin{itemize}
\item Prior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_0, \nu_0)$, an inverse-Wishart with $\boldsymbol{\Psi}_0 = 10^{-4}\boldsymbol{I}_{2p + q}$ and $\nu_0 = 1$.   
\item Prior for $\boldsymbol{\beta}$: $MN(\boldsymbol{\beta}_0, \boldsymbol{\Sigma}_0  \otimes \boldsymbol{\kappa}_0)$, where $\boldsymbol{\beta}$ is a vectorized $\boldsymbol{B}$, the matrix with regression coefficients, $\boldsymbol{\beta}_0 = \boldsymbol{0}_{k(2p + q)}$, $\boldsymbol{B}_0 = \boldsymbol{0}_{k \times (2p + q)}$ and $\boldsymbol{\kappa}_0 = 10^{-4}\boldsymbol{I}_k$.
\item Prior for $\lambda$: $N(\gamma_0, \omega_0)$, with $\gamma_0 = 0$ and $\omega_0 = 10000$.
\end{itemize}

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:

$$\tilde{y}_i = \lambda d_i $$

\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual i:

$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i,y_i)^t - (\boldsymbol{0}_{2p}, \lambda d_i)$$

\item Sample $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{B}$: $MN(\boldsymbol{B}_n, \boldsymbol{\kappa}_n, \boldsymbol{\Sigma}_n)$, with $\boldsymbol{B}_n = \boldsymbol{\kappa}_n^{-1}\boldsymbol{X}^t\boldsymbol{\eta} + \boldsymbol{\kappa}_0\boldsymbol{B}_0$ and $\boldsymbol{\kappa}_n = \boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\kappa}_0$.
\item Posterior for $\Sigma$: $IW(\boldsymbol{\Psi}_n, \nu_n)$, an inverse-Wishart with $\boldsymbol{\Psi}_n = \boldsymbol{\Psi}_0 + (\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B})^t(\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B}) + (\boldsymbol{B} - \boldsymbol{B}_0)^t\boldsymbol{\kappa}_0(\boldsymbol{B} - \boldsymbol{B}_0)$ and $\nu_n = \nu_0 + n$
\item Posterior for $\lambda$: $N(\gamma_n, \omega_n)$, with $\omega_n = \big(\sum_{i = 1}^{n}d_i^2\boldsymbol{\Sigma}_{y|s}^{-1} + \omega_0^{-1}\big)^{-1}$ and $\gamma_n = \omega_n \big(\sum_{i = 1}^{n}d_i\boldsymbol{\Sigma}_{y|s}^{-1}(y_i - \boldsymbol{\mu}_{y_i|s_i}) + \omega_0^{-1}\gamma_0 \big)$
\end{itemize}

\item Sample new $d_i$ for the training and holdout dataset from the following posterior:

$$f(d_i) \propto \phi_q(y_i|\boldsymbol{\mu}_{y_i|s_i} + \lambda d_i, \boldsymbol{\Sigma}_{y|s})\phi_q(d_i|0, 1),$$

where $\boldsymbol{\mu}_{y_i|s_i} = \boldsymbol{x}_i\boldsymbol{B}_{y_i|s_i}$. We can see $d_i$ as a positive regressor with $\lambda$ as covariate and $\phi_q(d_i|0, 1)$ as prior (Mastrantonio, 2018). The full conditional is then a q-dimensional truncated normal with support $\mathbb{R}^{+}$ as follows:

$$N_q(\boldsymbol{M}_{d_i}, \boldsymbol{V}_q),$$ \\where $\boldsymbol{V}_q = \big(\lambda^2\boldsymbol{\Sigma}_{y|s}^{-1} + 1\big)$ and $\boldsymbol{M}_{d_i} = \boldsymbol{V}_d\lambda\boldsymbol{\Sigma}_{y|s}^{-1}\big(y_i - \boldsymbol{\mu}_{y_i, s_i}\big)$.

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i|y_i}^{-1}\boldsymbol{\mu}_{s_i|y_i}$, $\boldsymbol{\mu}_{s_i|y_i} = \boldsymbol{x}_i\boldsymbol{B}_{s_i|y_i}$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i| y_i}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$. The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i, y_i)^t \sim N_{2p+q}(\boldsymbol{\mu} + (\boldsymbol{0}_{2p}, \lambda d_i)^t, \boldsymbol{\Sigma})$.

Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can simply define the $r_i$ as the euclidean norm of the bivariate datapoints. However, for didactic purposes continue with the explanation of the sampling procedure. We
can sample from posterior for $r_i$ using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{x}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditionals for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$ , is proportional to :
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.

\item Repeat steps 4 to 10 until the sampled parameter estimates have converged.

\item In the MCMC sampler we have estimated an unconstrained $\boldsymbol{\Sigma}$. However, for identification of the model we need to apply the following constraint to both $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$:

$$\boldsymbol{C} = \begin{bmatrix} \boldsymbol{C}_w & \boldsymbol{0}_{2p \times q} \\ \boldsymbol{0}_{2p \times q}^t & \boldsymbol{I}_q \end{bmatrix}$$
where $\boldsymbol{C}_w$ is a $2p \times 2p$ diagonal matrix with every $(2(j-1) + k)^{th}$ entry $> 0$ where $k \in 1, 2$ and $k = 1, \dots, p$ (Mastrantonio, 2018). The estimates $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$ can then be related to their constrained versions, $\tilde{\boldsymbol{\Sigma}}$ and $\tilde{\boldsymbol{\mu}}$ as follows:

$$\boldsymbol{\mu} = \boldsymbol{C}\tilde{\boldsymbol{\mu}}$$
$$\boldsymbol{\Sigma} = \boldsymbol{C}\tilde{\boldsymbol{\Sigma}}\boldsymbol{C}.$$

\end{enumerate}

\end{appendices}

### 2018-07-02 Manuscript/ManuscriptMarkdown.tex
Manuscript Update

--- 2018-07-02 Manuscript/ManuscriptMarkdown.Rmd
@@
----
-title: "Working title: Regression models for Cylindrical data."
-author: |
-  | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
-  | $^1$Department of Methodology and Statistics, Utrecht University
-  | $^2$TNO
-  | $^3$Department of Education, Utrecht University
-  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
-
-bibliography: CircularData.bib
-csl: apa.csl
-output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    number_sections: yes
-  html_document: default
-  word_document: default
-fontsize: 11pt
-geometry: margin=1in
-header-includes:
-  - \usepackage{multirow}
-  - \usepackage{appendix}
-  - \usepackage{color}
-  - \usepackage{hyperref}
-  - \usepackage{subcaption}
-keywords: bla vla
-documentclass: article
-pandoc_args: --natbib
-abstract: |
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Working title: Regression models for Cylindrical data.},
+            pdfkeywords={bla vla},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{Working title: Regression models for Cylindrical data.}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{Jolien Cremers\footnote{Corresponding author:
+  \href{mailto:j.cremers@uu.nl}{\nolinkurl{j.cremers@uu.nl}}} \(^1\),
+Helena J.M. Pennings\(^{2,3}\), Christophe Ley \(^{4}\)\\
+\(^1\)Department of Methodology and Statistics, Utrecht University\\
+\(^2\)TNO\\
+\(^3\)Department of Education, Utrecht University\\
+\(^4\)Department of Applied Mathematics, Computer Science and
+Statistics, Ghent University}
+  \preauthor{\centering\large\emph}
+  \postauthor{\par}
+  \date{}
+  \predate{}\postdate{}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+
+\usepackage{multirow}
+\usepackage{appendix}
+\usepackage{color}
+\usepackage{hyperref}
+\usepackage{subcaption}
+
+\begin{document}
+\maketitle
+\begin{abstract}
 This is the abstract
----
-```{r inlude = FALSE, echo = FALSE}
-
-knitr::opts_chunk$set(message = FALSE, warning = FALSE)
-```
-
-```{r, echo = FALSE}
-library(circular)
-library(bpnreg)
-library(haven)
-library(tikzDevice)
-library(plotrix)
-library(tidyr)
-library(dplyr) #sample_n()
-library(MASS) #mvrnorm()
-library(kableExtra) #latex tables
-```
+\end{abstract}
 
@@
 
-Cylindrical data are data that consist of a linear component and a directional 
-component (a set of angles). In this paper, the directional component is 
-circular meaning that it consists of a single angle instead of a set of angles. 
-Figure \ref{} \textcolor{red}{NOG MAKEN} shows the difference between a circular variable (left) and a 
-linear variable (right). The most obviuous difference is that on a circular 
-scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in fact 
-represent the same number while on a linear scale the two ends, $-\infty$ and 
-$\infty$ are not connected. Cylindrical data occur in several fields of 
-research, such as for instance in meteorology [@garcia2013exploring], ecology 
-[@garcia2014test] or marine research [@lagona2015hidden]. Several types of data 
-in psychological research are also of a cylindrical nature. For example, in 
-research on human navigation in the field of cognitive psychology both distance,
-a linear variable, and direction, a circular variable, of movement are of 
-interest [@chrastil2017rotational]. In eye-tracking research, which is used for 
-investigating cognitive processes we can also speak of cylindrical data. When 
-measuring eye-movements we speak of the eye-movements themselves, the saccades, 
-and fixations, the periods of time between movements when the eyes are still and
-looking at one point. Of the saccades both the direction, a circular variable, 
-and the duration, a linear variable, are of interest (for a review of 
-eye-tracking research see @rayner200935th). Data from circumplex measurement
-instruments, e.g. the interpersonal circumplex as used in personality
-psychology, are also of a cylindrical nature (see Section \ref{Example} for a
-more detailed explanation). In this paper we will discuss how a correct
-statistical treatment of such cylindrical data can lead to new insights. In
-particular we will show how cylindrical models pave the way for circular-linear
-and linear-circular regression. We will do this for a motivating example from
-the field of educational psychology. We will refer to this example data as the
-teacher data. In this example, apart from modelling the relation between the
-linear and circular component of a cylindrical variable we would also like to
-predict the two components from a set of covariates in a regression model. The
-teacher data will be further introduced in Sections \ref{Example} and
-\ref{DataAnalysis}.
-
-As is the case for circular data, the analysis of cylindrical data requires 
-special methods. For an overview of methods for circular data see
-@mardia2000directional and @jammalamadaka2001topics and for modern treatments
-@ley2017modern. In the literature several methods have been put forward to
-model the relation between the linear and circular component of a cylindrical 
-variable. Some of these are based on regressing the linear component onto the 
-circular component using the following type of relation: $$y = \beta_0 + 
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear
-component and $\theta$ the circular component [@mardia1978model;
-@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a 
-different way, e.g. by specifying a multivariate model for several linear and 
-circular variables and modelling their covariance matrix 
-[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. 
-For example, @abe2017tractable introduce a cylindrical distribution  based on a 
-Weibull distribution for the linear component and a sine-skewed von Mises 
-distribution for the circular component and link these through their shape and 
-concentration parameters respectively. However, none of the methods that have 
-been proposed thus far include additional covariates on which both the circular 
-and linear component are regressed.
-
-Our aim in this paper is to fill this gap in the literature by  adapting four of
-the existing methods for the analysis of cylindrical data in such a way that
-they include a regression of both the linear and circular component of a
-cylindrical variable onto a set of covariates. First however, we will introduce
-the teacher data in Section \ref{Example}. In Section \ref{Models} we introduce 
-the four modified models for cylindrical data that we use to analyze the data from
-the motivating example. We also choose a model selection criterion to compare
-the models. The teacher data will be analysed in Section \ref{DataAnalysis}. The
-paper will be concluded with a discussion in Section \ref{Discussion}. The
-Appendix contains technical details.
+Cylindrical data are data that consist of a linear component and a
+directional component (a set of angles). In this paper, the directional
+component is circular meaning that it consists of a single angle instead
+of a set of angles. Figure \ref{} \textcolor{red}{NOG MAKEN} shows the
+difference between a circular variable (left) and a linear variable
+(right). The most obviuous difference is that on a circular scale the
+datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in fact
+represent the same number while on a linear scale the two ends,
+\(-\infty\) and \(\infty\) are not connected. Cylindrical data occur in
+several fields of research, such as for instance in meteorology
+(García-Portugués, Crujeiras, \& González-Manteiga, 2013), ecology
+(García-Portugués, Barros, Crujeiras, González-Manteiga, \& Pereira,
+2014) or marine research (Lagona, Picone, Maruotti, \& Cosoli, 2015).
+Several types of data in psychological research are also of a
+cylindrical nature. For example, in research on human navigation in the
+field of cognitive psychology both distance, a linear variable, and
+direction, a circular variable, of movement are of interest (Chrastil \&
+Warren, 2017). In eye-tracking research, which is used for investigating
+cognitive processes we can also speak of cylindrical data. When
+measuring eye-movements we speak of the eye-movements themselves, the
+saccades, and fixations, the periods of time between movements when the
+eyes are still and looking at one point. Of the saccades both the
+direction, a circular variable, and the duration, a linear variable, are
+of interest (for a review of eye-tracking research see Rayner (2009)).
+Data from circumplex measurement instruments, e.g.~the interpersonal
+circumplex as used in personality psychology, are also of a cylindrical
+nature (see Section \ref{Example} for a more detailed explanation). In
+this paper we will discuss how a correct statistical treatment of such
+cylindrical data can lead to new insights. In particular we will show
+how cylindrical models pave the way for circular-linear and
+linear-circular regression. We will do this for a motivating example
+from the field of educational psychology. We will refer to this example
+data as the teacher data. In this example, apart from modelling the
+relation between the linear and circular component of a cylindrical
+variable we would also like to predict the two components from a set of
+covariates in a regression model. The teacher data will be further
+introduced in Sections \ref{Example} and \ref{DataAnalysis}.
+
+As is the case for circular data, the analysis of cylindrical data
+requires special methods. For an overview of methods for circular data
+see Mardia \& Jupp (2000) and Jammalamadaka \& Sengupta (2001) and for
+modern treatments Ley \& Verdebout (2017). In the literature several
+methods have been put forward to model the relation between the linear
+and circular component of a cylindrical variable. Some of these are
+based on regressing the linear component onto the circular component
+using the following type of relation: \[y = \beta_0 + 
+\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is
+the linear component and \(\theta\) the circular component (Johnson \&
+Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \&
+Jona-Lasinio, 2015). Others model the relation in a different way,
+e.g.~by specifying a multivariate model for several linear and circular
+variables and modelling their covariance matrix (Mastrantonio, 2018) or
+by proposing a joint cylindrical distribution. For example, Abe \& Ley
+(2017) introduce a cylindrical distribution based on a Weibull
+distribution for the linear component and a sine-skewed von Mises
+distribution for the circular component and link these through their
+shape and concentration parameters respectively. However, none of the
+methods that have been proposed thus far include additional covariates
+on which both the circular and linear component are regressed.
+
+Our aim in this paper is to fill this gap in the literature by adapting
+four of the existing methods for the analysis of cylindrical data in
+such a way that they include a regression of both the linear and
+circular component of a cylindrical variable onto a set of covariates.
+First however, we will introduce the teacher data in Section
+\ref{Example}. In Section \ref{Models} we introduce the four modified
+models for cylindrical data that we use to analyze the data from the
+motivating example. We also choose a model selection criterion to
+compare the models. The teacher data will be analysed in Section
+\ref{DataAnalysis}. The paper will be concluded with a discussion in
+Section \ref{Discussion}. The Appendix contains technical details.
 
@@
 
-The motivating example for this article comes from the field of educational 
-psychology and was collected for the studies of @pennings2017interpersonal,
-@Claessens2016side and @vanderWant2015role. These studies concern research on 
-classroom climate. An indicator of classroom climate is the students'
-perceptions of their teachers' interpersonal behavior. These interpersonal
-perceptions, both in educational psychology as well as in other areas of
-psychology, can be measured using circumplex measurement instruments (see
-@horowitz2010handbook for an overview of many such instruments).
-
-The Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] is
-one such circumplex measurement instrument that can be used to study student
-perceptions of their teachers' interpersonal behavior. This instrument contains
-items that load on two dimensions: Agency and Communion. Agency refers to the
-degree of power or control a teacher exerts in interaction with his/her
-students. Communion refers to the degree of friendliness or affiliation a
-teacher conveys in interaction with his/her students. The loadings on the two 
-dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
-this space are characterized by different teacher behavior, e.g. 'helpful' or
-'uncertain'. We call the two-dimensional space the interpersonal circle (IPC).
-The idea is that the IPC is ``a continuous order with no beginning or end''
-[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
-the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular variable.
+The motivating example for this article comes from the field of
+educational psychology and was collected for the studies of Pennings et
+al. (2017), Claessens (2016) and Want (2015). These studies concern
+research on classroom climate. An indicator of classroom climate is the
+students' perceptions of their teachers' interpersonal behavior. These
+interpersonal perceptions, both in educational psychology as well as in
+other areas of psychology, can be measured using circumplex measurement
+instruments (see Horowitz \& Strack (2011) for an overview of many such
+instruments).
+
+The Questionnaire on Teacher Interaction (QTI) (Wubbels, Brekelmans,
+Brok, \& Tartwijk, 2006) is one such circumplex measurement instrument
+that can be used to study student perceptions of their teachers'
+interpersonal behavior. This instrument contains items that load on two
+dimensions: Agency and Communion. Agency refers to the degree of power
+or control a teacher exerts in interaction with his/her students.
+Communion refers to the degree of friendliness or affiliation a teacher
+conveys in interaction with his/her students. The loadings on the two
+dimensions of the QTI can be placed in a two-dimensional space formed by
+Agency (vertical) and Communion (horizontal), see Figure \ref{QTI}.
+Different parts of this space are characterized by different teacher
+behavior, e.g. `helpful' or `uncertain'. We call the two-dimensional
+space the interpersonal circle (IPC). The idea is that the IPC is ``a
+continuous order with no beginning or end'' (Gurtman, 2009, p. 2). We
+call such ordering a circumplex ordering and the IPC is therefore often
+called the interpersonal circumplex. The ordering also implies that
+scores on the IPC could be viewed as a circular variable.
 
@@
 
-@Cremers2018Assessing show how data from the IPC can be
-considered circular data and analyzed as such using a regression model. The 
-two-dimension scores Agency and Communion can be converted to a circular score 
-using the two-argument arctangent function in Equation \eqref{PredVal}, where 
-$A$ represents a score on the Agency dimension and $C$ represents a score on the
-Communion dimension. The resulting circular variable $\theta$ can then be
-modeled and for example be regressed onto covariates in a circular regression 
-model. However, when two-dimensional data are converted to the circle we lose 
-some information, namely the length of the two-dimensional vector $(A, C)$, 
-\emph{i.e.}, its euclidean norm $\mid\mid (A, C) \mid\mid$. This length 
-represents the strength of the type of interpersonal behavior a teacher shows 
-towards his/her students. In a cylindrical model we are able to incorporate this
-information, and model a circular variable $\theta$ together with a linear 
-variable corresponding to $\mid\mid (A, C) \mid\mid$. This will lead to an 
-improved analysis of data from the IPC.  In the next section we will introduce 
-several models that we can use for a more accurate and more informative regression
-analysis on the teacher data. More information on the teacher data will be given
-in Section \ref{DataAnalysis}.
+Cremers et al. (2018a) show how data from the IPC can be considered
+circular data and analyzed as such using a regression model. The
+two-dimension scores Agency and Communion can be converted to a circular
+score using the two-argument arctangent function in Equation
+\eqref{PredVal}, where \(A\) represents a score on the Agency dimension
+and \(C\) represents a score on the Communion dimension. The resulting
+circular variable \(\theta\) can then be modeled and for example be
+regressed onto covariates in a circular regression model. However, when
+two-dimensional data are converted to the circle we lose some
+information, namely the length of the two-dimensional vector \((A, C)\),
+\emph{i.e.}, its euclidean norm \(\mid\mid (A, C) \mid\mid\). This
+length represents the strength of the type of interpersonal behavior a
+teacher shows towards his/her students. In a cylindrical model we are
+able to incorporate this information, and model a circular variable
+\(\theta\) together with a linear variable corresponding to
+\(\mid\mid (A, C) \mid\mid\). This will lead to an improved analysis of
+data from the IPC. In the next section we will introduce several models
+that we can use for a more accurate and more informative regression
+analysis on the teacher data. More information on the teacher data will
+be given in Section \ref{DataAnalysis}.
 
@@
 
-```{r loaddata, return = FALSE, echo = FALSE}
-Dat <- read_spss("Data Heleen/mergeddata.sav")
-
-#only select the first measurement occasion
-Dat <- subset(Dat, Time == 0)
-
-#sort on class size
-Dat <- Dat[order(-Dat$nklas), ]
-
-#Now remove duplicates 
-#(because dataframe is sorted on class size the smallest classes are removed)
-
-duplicates <- duplicated(Dat[,c("docnr", "Time")])
-Dat <- Dat[!duplicates,]
-
-#center and remove missings for self-efficacy
-Dat <- Dat %>% drop_na(Efficacy_SE)
-Dat$SEc <- Dat$Efficacy_SE- mean(Dat$Efficacy_SE)
-
-set.seed(50)
-
-#Create holdout and test sample
-Dat.hold <- sample_n(Dat, 15)
-Dat.test <- Dat[-as.numeric(rownames(Dat.hold)),]
-
-#Compute circular and linear outcome
-theta.test <- atan2(Dat.test$lAgency, Dat.test$lcommunion)
-y.test     <- sqrt(Dat.test$lAgency^2 + Dat.test$lcommunion^2)
-
-theta.hold <- atan2(Dat.hold$lAgency, Dat.hold$lcommunion)
-y.hold     <- sqrt(Dat.hold$lAgency^2 + Dat.hold$lcommunion^2)
-
-```
-
 \section{Four cylindrical regression models}\label{Models}
 
-In this section we will introduce four cylindrical regression models. We will 
-extend the models from @mastrantonio2018joint and @abe2017tractable to contain 
-predictors for the linear and circular outcomes in the cylindrical model. 
-Additionally we introduce two models where both variates $\Theta$ and $Y$ are
-predicted by covariates and that model the relation between the linear and
-circular outcome as follows (following @wang2015joint and
-@mastrantonio2015bayesian):
+In this section we will introduce four cylindrical regression models. We
+will extend the models from Mastrantonio (2018) and Abe \& Ley (2017) to
+contain predictors for the linear and circular outcomes in the
+cylindrical model. Additionally we introduce two models where both
+variates \(\Theta\) and \(Y\) are predicted by covariates and that model
+the relation between the linear and circular outcome as follows
+(following Wang, Gelfand, \& Jona-Lasinio (2015) and Mastrantonio et al.
+(2015)):
 
@@
 
-where $r$ will be introduced in Section \ref{CL-(G)PN}, the error term $\epsilon
-\sim N(0, \sigma)$, $\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, 
-\gamma_q$ are the intercept and regression coefficients and   $x_1, \dots, x_q$
-are the $q$ covariate values. In this model $Y$ follows a normal distribution
-and $\Theta$ a projected normal (PN) or general projected normal (GPN)
-distribution on the circle.
+where \(r\) will be introduced in Section \ref{CL-(G)PN}, the error term
+\(\epsilon \sim N(0, \sigma)\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
+the intercept and regression coefficients and \(x_1, \dots, x_q\) are
+the \(q\) covariate values. In this model \(Y\) follows a normal
+distribution and \(\Theta\) a projected normal (PN) or general projected
+normal (GPN) distribution on the circle.
 
@@
 
-In both of these models the relation between the circular variable $\Theta \in
-[0, 2\pi)$ and linear variable $Y \in (-\infty, + \infty)$ is specified as in
-Equation \ref{circlinlink}. The distribution of $Y$ conditional on $\Theta$ is
-then given as:
+In both of these models the relation between the circular variable
+\(\Theta \in [0, 2\pi)\) and linear variable
+\(Y \in (-\infty, + \infty)\) is specified as in Equation
+\ref{circlinlink}. The distribution of \(Y\) conditional on \(\Theta\)
+is then given as:
 
@@
 
-where $c = \begin{bmatrix} r \cos \theta \\  r\sin \theta \end{bmatrix}^t 
-\begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}$, $r \geq 0$, 
-$\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q$ are the
-intercept and regression coefficients and $\sigma^2 \geq 0$ is the error
-variance. The linear outcome thus has a normal distribution, conditional on
-$\Theta$.
+where
+\(c = \begin{bmatrix} r \cos \theta \\ r\sin \theta \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
+\(r \geq 0\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
+the intercept and regression coefficients and \(\sigma^2 \geq 0\) is the
+error variance. The linear outcome thus has a normal distribution,
+conditional on \(\Theta\).
 
 For the circular outcome we assume either a projected normal (PN) or a
-general projected normal (GPN) distribution. These distributions arise from the
-radial projection of a distribution defined on the plane onto the circle. The
-relation between a bivariate variate $\boldsymbol{S}$ in the plane and the
-circular outcome $\Theta$ is defined as follows:
+general projected normal (GPN) distribution. These distributions arise
+from the radial projection of a distribution defined on the plane onto
+the circle. The relation between a bivariate variate \(\boldsymbol{S}\)
+in the plane and the circular outcome \(\Theta\) is defined as follows:
 
@@
 
-where $R = \mid\mid S \mid\mid$, the euclidean norm of the bivariate vector 
-$\boldsymbol{S}$. In the PN distribution we assume $\boldsymbol{S} \sim 
-N_2(\boldsymbol{\mu}, \boldsymbol{I})$ and in the GPN we assume $\boldsymbol{S} 
-\sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})$ where $\boldsymbol{\Sigma} = 
-\begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$, $\rho \in 
-(-\infty, +\infty)$ and $\tau^2 \geq 0$ (as in @hernandez2016general).
-
-Following @nunez2011bayesian, the joint density of $\Theta$ and $R$ for the PN
-distribution in a regression set-up equals:
+where \(R = \mid\mid S \mid\mid\), the euclidean norm of the bivariate
+vector \(\boldsymbol{S}\). In the PN distribution we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the
+GPN we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
+\(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
+Hernandez-Stumpfhauser, Breidt, \& Woerd (2016)).
+
+Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
+density of \(\Theta\) and \(R\) for the PN distribution in a regression
+set-up equals:
 
@@
 
-In a regression setup the outcome $\theta_i$ for each individual $i = 1,
-\dots, n$, where $n$ is the sample size, is generated independently from
-Equation (\ref{pnreg}). The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is 
-defined as $\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
-$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and $\boldsymbol{B} = 
-(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ contain the regression 
-coefficients. Note however that the dimensions of $\boldsymbol{\beta}^{I}$ and 
-$\boldsymbol{\beta }^{I}$ need not necessarily be the same and we are thus 
-allowed to have a different set of predictor variables and vectors 
-$\boldsymbol{z}_i^I$ and $\boldsymbol{z}_i^{II}$ for the two components of 
-$\boldsymbol{\mu}_i$.
-
-Following @wang2012directional and @hernandez2016general the joint density of
-$r$ and $\theta$ for the GPN distribution equals:
+In a regression setup the outcome \(\theta_i\) for each individual
+\(i = 1, \dots, n\), where \(n\) is the sample size, is generated
+independently from Equation (\ref{pnreg}). The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}\). The vector
+\(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values and
+\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
+contain the regression coefficients. Note however that the dimensions of
+\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{I}\) need not
+necessarily be the same and we are thus allowed to have a different set
+of predictor variables and vectors \(\boldsymbol{z}_i^I\) and
+\(\boldsymbol{z}_i^{II}\) for the two components of
+\(\boldsymbol{\mu}_i\).
+
+Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
+(2016) the joint density of \(r\) and \(\theta\) for the GPN
+distribution equals:
 
@@
 
-where $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2  & \rho\\ \rho & 1 
-\end{bmatrix}$, $\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta 
-\end{bmatrix}$. In a regression setup the outcome $\theta_i$ for each individual
-$i = 1, \dots, n$, where $n$ is the sample size, is generated independently
-from Equation (\ref{gpnreg}). The mean vector $\boldsymbol{\mu}_i \in 
-\mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
-\boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. The vector 
-$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and each 
-$\boldsymbol{\beta}^{k}$ is a vector with two regression coefficients, one for 
-each of the two components of $\boldsymbol{\mu}_i$. Note that for the CL-GPN 
-model we do need to have the same predictors for both components of 
-$\boldsymbol{\mu}_i$.
-
-Both cylindrical models introduced in this section are estimated using MCMC 
-methods. These methods are based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in
-Appendices \ref{A1} and \ref{A2}.
+where
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
+\(\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta \end{bmatrix}\).
+In a regression setup the outcome \(\theta_i\) for each individual
+\(i = 1, \dots, n\), where \(n\) is the sample size, is generated
+independently from Equation (\ref{gpnreg}). The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\).
+The vector \(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values
+and each \(\boldsymbol{\beta}^{k}\) is a vector with two regression
+coefficients, one for each of the two components of
+\(\boldsymbol{\mu}_i\). Note that for the CL-GPN model we do need to
+have the same predictors for both components of \(\boldsymbol{\mu}_i\).
+
+Both cylindrical models introduced in this section are estimated using
+MCMC methods. These methods are based on Nuñez-Antonio et al. (2011),
+Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
+regression of the circular outcome. A detailed description of the
+Bayesian estimation and MCMC samplers can be found in Appendices
+\ref{A1} and \ref{A2}.
 
@@
 
-This model is an extension of the cylindrical model introduced by
-@abe2017tractable to the regression context. The density of the model for the
-pair of the circular variable $\theta \in [0, 2\pi)$ and linear variable $y$ on
-the positive real half-line $[0, + \infty)$ is:
+This model is an extension of the cylindrical model introduced by Abe \&
+Ley (2017) to the regression context. The density of the model for the
+pair of the circular variable \(\theta \in [0, 2\pi)\) and linear
+variable \(y\) on the positive real half-line \([0, + \infty)\) is:
 
@@
 
-In a regression setup the outcome vector $(\theta_i, y_i)^t$ for each individual
-$i = 1, \dots, n$, where $n$ is the sample size, is generated independently from
-Equation (\ref{WeiSSVMdensity}). The parameter $\alpha > 0$ is a linear shape 
-parameter, $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ is a linear 
-scale parameter, $\mu_i = \eta_0 + 
-2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$ is a circular 
-location parameter, $\kappa > 0$ is a cicular concentration parameter and 
-$\lambda \in [-1, 1]$ is a circular skewness parameter. The parameter 
-$\boldsymbol{\nu}$ is a vector of $q$ regression coefficients and intercept
-$\nu_j \in (-\infty, +\infty)$ where $j = 0, \dots, q$, for the prediction of
-$y$. The parameter $\eta_0 \in [0, 2\pi)$ is the intercept and
-$\boldsymbol{\eta}$, is a vector of $p$ regression coefficients $\eta_j \in 
-(-\infty, +\infty)$, where $j = 1, \dots, p$, for the prediction of $\theta$. 
-The vector $\boldsymbol{x}_i$ is a vector of predictor values for the prediction
-of $y$ and $\boldsymbol{z}_i$ is a vector of predictor values for the prediction
-of $\theta$.
-
-As in @abe2017tractable, the conditional distribution of $y$ given $\theta$ is a
-Weibull distribution with scale $\alpha$ and shape
-$\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}$ and the conditional 
-distribution of $\theta$ given $y$ is a sine skewed von Mises distribution with 
-location parameter $\mu$ and concentration parameter $(\beta 
-y)^\alpha\tanh(\kappa)$.
+In a regression setup the outcome vector \((\theta_i, y_i)^t\) for each
+individual \(i = 1, \dots, n\), where \(n\) is the sample size, is
+generated independently from Equation (\ref{WeiSSVMdensity}). The
+parameter \(\alpha > 0\) is a linear shape parameter,
+\(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) is a linear
+scale parameter,
+\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)\)
+is a circular location parameter, \(\kappa > 0\) is a cicular
+concentration parameter and \(\lambda \in [-1, 1]\) is a circular
+skewness parameter. The parameter \(\boldsymbol{\nu}\) is a vector of
+\(q\) regression coefficients and intercept
+\(\nu_j \in (-\infty, +\infty)\) where \(j = 0, \dots, q\), for the
+prediction of \(y\). The parameter \(\eta_0 \in [0, 2\pi)\) is the
+intercept and \(\boldsymbol{\eta}\), is a vector of \(p\) regression
+coefficients \(\eta_j \in (-\infty, +\infty)\), where
+\(j = 1, \dots, p\), for the prediction of \(\theta\). The vector
+\(\boldsymbol{x}_i\) is a vector of predictor values for the prediction
+of \(y\) and \(\boldsymbol{z}_i\) is a vector of predictor values for
+the prediction of \(\theta\).
+
+As in Abe \& Ley (2017), the conditional distribution of \(y\) given
+\(\theta\) is a Weibull distribution with scale \(\alpha\) and shape
+\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}\) and the
+conditional distribution of \(\theta\) given \(y\) is a sine skewed von
+Mises distribution with location parameter \(\mu\) and concentration
+parameter \((\beta y)^\alpha\tanh(\kappa)\).
 
@@
 
-We can use numerical optimization (Nelder-Mead) to find solutions for the
-maximum likelihood (ML) estimates for the parameters of the model.
+We can use numerical optimization (Nelder-Mead) to find solutions for
+the maximum likelihood (ML) estimates for the parameters of the model.
 
@@
 This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. The model contains $p$
-circular outcomes and $q$ linear outcomes. The circular outcomes 
-$\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ 
-are modeled together by a multivariate GPN distribution and the linear outcomes 
-$\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled 
-together by a multivariate skew normal distribution [@sahu2003new]. Because the
-GPN distribution is modelled using a so-called augmented representation (as in
-(\ref{projection}) and (\ref{gpnreg})) it is convenient to use a similar tactic
-for modelling the multivariate skew normal distribution. As in
-@mastrantonio2018joint we use the following representation of the linear
-outcomes:
-
-$$\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},$$
-where where $\boldsymbol{\mu}_y$ is a mean vector for the linear outcome 
-$\boldsymbol{Y}$, $\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})$ is a $q 
-\times q$ diagonal matrix with diagonal elements $\boldsymbol{\lambda} = 
-\lambda_1, \dots, \lambda_q$ (the parameters that indicate the skewness of the 
-linear outcomes), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
-\boldsymbol{I}_q)$, a q-dimensional half normal distribution
-[@olmos2012extension] and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
-\boldsymbol{\Sigma}_y)$. This means that conditional on the auxiliary data
-$\boldsymbol{D}$, $\boldsymbol{Y}$ is normally distributed with mean
-$\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}$ and covariance matrix
-$\boldsymbol{\Sigma}_y$. The joint density for $(\boldsymbol{Y},
-\boldsymbol{D})^t$ is defined as follows:
+Mastrantonio (2018) to the regression context. The model contains \(p\)
+circular outcomes and \(q\) linear outcomes. The circular outcomes
+\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)\)
+are modeled together by a multivariate GPN distribution and the linear
+outcomes
+\(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)\) are
+modeled together by a multivariate skew normal distribution (Sahu, Dey,
+\& Branco, 2003). Because the GPN distribution is modelled using a
+so-called augmented representation (as in (\ref{projection}) and
+(\ref{gpnreg})) it is convenient to use a similar tactic for modelling
+the multivariate skew normal distribution. As in Mastrantonio (2018) we
+use the following representation of the linear outcomes:
+
+\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
+where where \(\boldsymbol{\mu}_y\) is a mean vector for the linear
+outcome \(\boldsymbol{Y}\),
+\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
+\(q \times q\) diagonal matrix with diagonal elements
+\(\boldsymbol{\lambda} = \lambda_1, \dots, \lambda_q\) (the parameters
+that indicate the skewness of the linear outcomes),
+\(\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, \boldsymbol{I}_q)\), a
+q-dimensional half normal distribution (Olmos, Varela, Gómez, \&
+Bolfarine, 2012) and
+\(\boldsymbol{H} \sim N_q(\boldsymbol{0}_q, \boldsymbol{\Sigma}_y)\).
+This means that conditional on the auxiliary data \(\boldsymbol{D}\),
+\(\boldsymbol{Y}\) is normally distributed with mean
+\(\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and
+covariance matrix \(\boldsymbol{\Sigma}_y\). The joint density for
+\((\boldsymbol{Y}, \boldsymbol{D})^t\) is defined as follows:
 
@@
 
-As in @mastrantonio2018joint dependence between the linear and circular outcome
-is created by modelling the augmented representations of $\boldsymbol{\Theta}$,
-and $\boldsymbol{Y}$ together in a $2 \times p + q$ dimensional normal
-distribution. The joint density of the model is then represented by:
+As in Mastrantonio (2018) dependence between the linear and circular
+outcome is created by modelling the augmented representations of
+\(\boldsymbol{\Theta}\), and \(\boldsymbol{Y}\) together in a
+\(2 \times p + q\) dimensional normal distribution. The joint density of
+the model is then represented by:
 
@@
 
-where the mean vector $\boldsymbol{\mu} = (\boldsymbol{\mu}_s,
-\boldsymbol{\mu}_y)^t$ and $\boldsymbol{\Sigma} = \left ( \begin{matrix}
-\boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} &
-\boldsymbol{\Sigma}_y \\ \end{matrix} \right )$. The matrix
-$\boldsymbol{\Sigma}_s$ is the covariance matrix for the variances of and
-covariances between the augmented representations of the circular outcome and
-the matrix $\boldsymbol{\Sigma}_{sy}$ contains covariances between the augmented
-representations of the circular outcome and the linear outcome.
-
-In our regression extension we have $i = 1, \dots, n$ observations of $p$ 
-circular outcomes, $q$ linear outcomes and $k$ covariates. The mean in the 
-density in (\ref{YDThetarjoint}) then becomes $\boldsymbol{\mu}_i = 
-\boldsymbol{x}_i\boldsymbol{B}$ where $\boldsymbol{B}$ is a $k \times (2 
-\times p + q)$ matrix with regression coeffients. We estimate the model using 
-MCMC methods. A detailed description of these methods is given in Appendix
-\ref{A3}.
+where the mean vector
+\(\boldsymbol{\mu} = (\boldsymbol{\mu}_s, \boldsymbol{\mu}_y)^t\) and
+\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
+The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
+variances of and covariances between the augmented representations of
+the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
+contains covariances between the augmented representations of the
+circular outcome and the linear outcome.
+
+In our regression extension we have \(i = 1, \dots, n\) observations of
+\(p\) circular outcomes, \(q\) linear outcomes and \(k\) covariates. The
+mean in the density in (\ref{YDThetarjoint}) then becomes
+\(\boldsymbol{\mu}_i = \boldsymbol{x}_i\boldsymbol{B}\) where
+\(\boldsymbol{B}\) is a \(k \times (2 \times p + q)\) matrix with
+regression coeffients. We estimate the model using MCMC methods. A
+detailed description of these methods is given in Appendix \ref{A3}.
 
@@
 
-For the four cylindrical models we focus on their out-of-sample predictive 
-performance to determine the fit of the model. To do so we split our data in a
-training and holdout set (10 $\%$ of the sample). A proper criterion to compare
-out-of-sample predictive performance is the the Predictive Log Scoring Loss
-(PLSL) [@gneiting2007strictly]. The lower the value of this criterion, the
-better the predictive performance of the model. Using ML estimates this
-criterion can be computed as follows:
+For the four cylindrical models we focus on their out-of-sample
+predictive performance to determine the fit of the model. To do so we
+split our data in a training and holdout set (10 \(\%\) of the sample).
+A proper criterion to compare out-of-sample predictive performance is
+the the Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007).
+The lower the value of this criterion, the better the predictive
+performance of the model. Using ML estimates this criterion can be
+computed as follows:
 
@@
 
-where $l$ is the model likelihood, $M$ is the sample size of the holdout set, 
-$x_i$ is the $i^{th}$ datapoint from the holdout set and 
-$\hat{\boldsymbol{\vartheta}}$ are the ML estimates of the model parameters. 
-Using posterior samples the criterion is similar to the log pointwise predictive
-density (lppd) as outlined in @BDA and can be computed as:
+where \(l\) is the model likelihood, \(M\) is the sample size of the
+holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set
+and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model
+parameters. Using posterior samples the criterion is similar to the log
+pointwise predictive density (lppd) as outlined in Gelman et al. (2014)
+and can be computed as:
 
@@
 
-where $B$ is the amount of posterior samples and $\boldsymbol{\vartheta}^{(j)}$
-are the posterior estimates of the model parameters for the $j^{th}$ iteration.
-Note that although we fit the CL-PN, CL-GPN and joint GPN-SSN models using
-Bayesian statistics, we do not take prior information (only the likelihood) into
-account when assessing model fit. According to @BDA this is not necessary since
-we are assessing the fit of a model to data only \textcolor{red}{AFMAKEN!}. For 
-each of the cylindrical models we can then compute a PLSL for the circular and 
-linear outcome by using the conditional log-likelihoods of the respective 
-outcome.
+where \(B\) is the amount of posterior samples and
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
+model parameters for the \(j^{th}\) iteration. Note that although we fit
+the CL-PN, CL-GPN and joint GPN-SSN models using Bayesian statistics, we
+do not take prior information (only the likelihood) into account when
+assessing model fit. According to Gelman et al. (2014) this is not
+necessary since we are assessing the fit of a model to data only
+\textcolor{red}{AFMAKEN!}. For each of the cylindrical models we can
+then compute a PLSL for the circular and linear outcome by using the
+conditional log-likelihoods of the respective outcome.
 
@@
 
-In this section we will analyze the teacher dataset with the help of
-the four cylindrical models introduced in Section \ref{Models}. We will however
-first give a more detailed description of the dataset.
-
-```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("Posterior Sampling informative Prior CL-PN.R")
-
-its <- 20000
-set.seed(101)
-
-ZI.PN.test  <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-ZII.PN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.PN.test   <- as.matrix(cbind(rep(1, length(theta.test)),
-                               cos(theta.test),
-                               sin(theta.test),
-                               Dat.test$SEc))
-
-ZI.PN.hold  <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-ZII.PN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.PN.hold   <- as.matrix(cbind(rep(1, length(theta.hold)),
-                               cos(theta.hold),
-                               sin(theta.hold),
-                               Dat.hold$SEc))
-
-res_CLPN <- CLPN(theta.test, y.test, X.PN.test, ZI.PN.test, ZII.PN.test, its,
-                 theta.hold, y.hold, X.PN.hold, ZI.PN.hold, ZII.PN.hold)
-```
-
-```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("Posterior Sampling CL-GPN.R")
-
-its <- 20000
-set.seed(101)
-
-Z.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)),
-                              cos(theta.test),
-                              sin(theta.test),
-                              Dat.test$SEc))
-
-Z.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)),
-                              cos(theta.hold),
-                              sin(theta.hold),
-                              Dat.hold$SEc))
-
-res_CLGPN <- CLGPN(theta.test, y.test, X.GPN.test, Z.GPN.test, its, p = 2, 
-                   theta.hold, y.hold, X.GPN.hold, Z.GPN.hold)
-
-```
-
-```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("Posterior Sampling CL-GPN_multivariate.R")
-
-its <- 20000
-set.seed(101)
-
-X.MGPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-
-X.MGPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-
-res_CLGPNM <- CLGPNM(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
-                     theta.hold, y.hold, X.MGPN.hold)
-
-```
-
-```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("WeisSVM optimization.R")
-
-set.seed(101)
-
-Z.test     <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.test    <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-
-Z.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-
-colnames(X.test) <- c("ax", "bx")
-colnames(Z.test) <- c("az", "bz")
-colnames(X.hold) <- c("ax", "bx")
-colnames(Z.hold) <- c("az", "bz")
-
-dat.t <- as.data.frame(cbind(theta.test, y.test, X.test, Z.test))
-dat.h <- as.data.frame(cbind(theta.hold, y.hold, X.hold, Z.hold))
-
-ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
-               c(-1, 0, 0, 0, 0, 0, 0),
-               c(0, 0, 0, 0, 1, 0, 0),
-               c(0, 0, 0, 0, 0, 1, 0),
-               c(0, 0, 0, 0, 0, 0, 1),
-               c(0, 0, 0, 0, 0, 0, -1))
-
-ci.reg = c(-pi, -pi, 0, 0, -1, -1)
-
-param <- c(1,1,1,1,1,1,0)*0.5
-
-ui.reg %*% param - ci.reg
-
-resAL <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
-                     control = list(maxit = 1000000, fnscale = -1), data = dat.t)
-
-```
+In this section we will analyze the teacher dataset with the help of the
+four cylindrical models introduced in Section \ref{Models}. We will
+however first give a more detailed description of the dataset.
 
@@
 
-The teacher data was collected between 2010 and 2015 and contains several repeated
-measures on the IPC of 161 teachers gathered for the studies of
-@vanderWant2015role, @Claessens2016side and @pennings2017interpersonal. 
-The measurements were obtained using the QTI and taken in different years and in
-different classes. For this paper we only consider one measurement, namely the
-one for the first measurement occasion (2010) and the one for the largest class
-if data for multiple classes were available. In addition to a variable `IPC`
-containing the circular outcome and the 'length' of the IPC, the linear outcome,
-a teachers' self-efficacy (\verb|SE|) will be used as covariate in the analysis.
-After the removal of missing cases we end up with a sample of 148 teachers Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset.
-
-```{r descripitves, echo = FALSE, results = FALSE}
-#sample size
-nrow(Dat)
-
-#summary statistics
-summary(Dat$Efficacy_SE)
-summary(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
-summary(as.circular(atan2(Dat$lAgency, Dat$lcommunion)%%(2*pi)))
-
-sd(Dat$Efficacy_SE)
-sd(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
-
-```
+The teacher data was collected between 2010 and 2015 and contains
+several repeated measures on the IPC of 161 teachers gathered for the
+studies of Want (2015), Claessens (2016) and Pennings et al. (2017). The
+measurements were obtained using the QTI and taken in different years
+and in different classes. For this paper we only consider one
+measurement, namely the one for the first measurement occasion (2010)
+and the one for the largest class if data for multiple classes were
+available. In addition to a variable \texttt{IPC} containing the
+circular outcome and the `length' of the IPC, the linear outcome, a
+teachers' self-efficacy (\verb|SE|) will be used as covariate in the
+analysis. After the removal of missing cases we end up with a sample of
+148 teachers Table \ref{Tableteacherdescriptives} shows descriptives for
+the dataset.
 
@@
 
-The regression equations for the linear and the circular outcome of the four
-cylindrical models fit to the teacher dataset are as follows:
+The regression equations for the linear and the circular outcome of the
+four cylindrical models fit to the teacher dataset are as follows:
 
@@
 
-The variable \verb|SE| is thus used as a predictor for the linear and circular 
-outcomes in all four cylindrical models. We will use the loglikelihoods of the
-following conditional densities \textcolor{red}{CHRISTOPHE, MAG HIER WEL $\sim$ GEBRUIKT WORDEN?}for the computation of the PLSL criterion to
-evaluate the model fit:
+The variable \verb|SE| is thus used as a predictor for the linear and
+circular outcomes in all four cylindrical models. We will use the
+loglikelihoods of the following conditional densities
+\textcolor{red}{CHRISTOPHE, MAG HIER WEL $\sim$ GEBRUIKT WORDEN?}for the
+computation of the PLSL criterion to evaluate the model fit:
 
@@
 
-Before analysis there were a couple of settings that we had to specify for the 
-cylindrical models. Starting values for the WeiSSVM model were the following 
-$\eta_0 = 0.5, \eta_1 = 0.5, \nu_0 = 0.5, \nu_1 = 0.5, \kappa = 0.5, \alpha = 
-0.5, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
-was set to 2000. After we checked convergence via traceplots we concluded that 
-some of the parameters of the multivariate GPN model did not converge. Therefore
-we set the amount of iterations of the MCMC models to 20,000 and subtracted a 
-burn-in of 5000. With this specification the MCMC chains for all parameters 
-converge. Note that we choose the same amount of iterations for all three 
-Bayesian cylindrical models to make their comparison via the PLSL as fair as 
-possible. Lastly, the predictor \verb|SE| was centered before inclusion in the
-analysis.
-
-```{r estCLPN, echo = FALSE}
-
-CLPN.Gamma.m   <- apply(res_CLPN$Gamma[5000:20000,], 2, mode_est)
-CLPN.Gamma.hpd <- apply(res_CLPN$Gamma[5000:20000,], 2, hpd_est)
-
-CLPN.BI.m   <- apply(res_CLPN$BI[5000:20000,], 2, mode_est)
-CLPN.BI.hpd <- apply(res_CLPN$BI[5000:20000,], 2, hpd_est)
-
-CLPN.BII.m <- apply(res_CLPN$BII[5000:20000,], 2, mode_est)
-CLPN.BII.hpd <- apply(res_CLPN$BII[5000:20000,], 2, hpd_est)
-
-CLPN.Sigma.m   <- mode_est(res_CLPN$Sigma[5000:20000])
-CLPN.Sigma.hpd <- hpd_est(res_CLPN$Sigma[5000:20000])
-
-modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
-hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
-hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)
-
-CLPNres.tab <- cbind(modes.PN, hpd.LB.PN, hpd.UB.PN)
-
-```
-
-```{r estCLGPN, echo = FALSE}
-
-CLGPN.Sig.m   <- mode_est(res_CLGPN$Sig)
-CLGPN.Sig.hpd <- hpd_est(res_CLGPN$Sig)
-
-CLGPN.Gamma.m   <- apply(res_CLGPN$Gamma[5000:20000,], 2, mode_est)
-CLGPN.Gamma.hpd <- apply(res_CLGPN$Gamma[5000:20000,], 2, hpd_est)
-
-CLGPN.BI.m   <- apply(res_CLGPN$BI[5000:20000,], 2, mode_est)
-CLGPN.BI.hpd <- apply(res_CLGPN$BI[5000:20000,], 2, hpd_est)
-
-CLGPN.BII.m   <- apply(res_CLGPN$BII[5000:20000,], 2, mode_est)
-CLGPN.BII.hpd <- apply(res_CLGPN$BII[5000:20000,], 2, hpd_est)
-
-CLGPN.Sigma11.m   <- mode_est(res_CLGPN$Sigma[1,1,5000:20000])
-CLGPN.Sigma11.hpd <- hpd_est(res_CLGPN$Sigma[1,1,5000:20000])
-CLGPN.Sigma22.m   <- mode_est(res_CLGPN$Sigma[2,2,5000:20000])
-CLGPN.Sigma22.hpd <- hpd_est(res_CLGPN$Sigma[2,2,5000:20000])
-CLGPN.Sigma12.m   <- mode_est(res_CLGPN$Sigma[1,2,5000:20000])
-CLGPN.Sigma12.hpd <- hpd_est(res_CLGPN$Sigma[1,2,5000:20000])
-
-modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
-                CLGPN.Gamma.m, CLGPN.Sig.m,
-                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
-hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
-               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
-               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
-hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
-               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
-               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])
-
-CLGPNres.tab <- cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN)
-CLPNGPNres.tab <- round(cbind(CLPNres.tab, CLGPNres.tab), 2)
-
-rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_0^{I}$",
-                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
-                              "$\\gamma_0$", "$\\gamma_{cos}$",
-                              "$\\gamma_{sin}$", "$\\gamma_1$",
-                              "$\\sigma$", "$\\sum_{1,1}$",
-                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
-colnames(CLPNGPNres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")
-
-kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE,
-      caption = "Results for the modified CL-PN and CL-GPN model")%>%
-add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))
-
-```
-
-```{r estSL, echo = FALSE}
-
-ALres.tab <- matrix(round(c(resAL$par[1:4], resAL$par[6], resAL$par[5], resAL$par[7]), 2),7,1)
-rownames(ALres.tab) <- c("$\\eta_0$", "$\\eta_1$",
-                         "$\\nu_0$", "$\\nu_1$",
-                         "$\\alpha$", "$\\kappa$", "$\\lambda$")
-
-kable(ALres.tab, "latex", booktabs = T, escape = FALSE,
-      caption = "Results for the modified Abe-Ley model")%>%
-add_header_above(c("Parameter" = 1, "ML-estimate" = 1))
-
-```
-
-```{r estCLMGPN, echo = FALSE}
-
-CLGPNM.lambda.m   <- mode_est(res_CLGPNM$lambda[5000:20000, ])
-CLGPNM.lambda.hpd <- hpd_est(res_CLGPNM$lambda[5000:20000, ])
-
-CLGPNM.Beta11.m <- mode_est(res_CLGPNM$Beta[1,1,5000:20000])
-CLGPNM.Beta21.m <- mode_est(res_CLGPNM$Beta[2,1,5000:20000])
-CLGPNM.Beta12.m <- mode_est(res_CLGPNM$Beta[1,2,5000:20000])
-CLGPNM.Beta22.m <- mode_est(res_CLGPNM$Beta[2,2,5000:20000])
-CLGPNM.Beta13.m <- mode_est(res_CLGPNM$Beta[1,3,5000:20000])
-CLGPNM.Beta23.m <- mode_est(res_CLGPNM$Beta[2,3,5000:20000])
-
-CLGPNM.Beta11.hpd <- hpd_est(res_CLGPNM$Beta[1,1,5000:20000])
-CLGPNM.Beta21.hpd <- hpd_est(res_CLGPNM$Beta[2,1,5000:20000])
-CLGPNM.Beta12.hpd <- hpd_est(res_CLGPNM$Beta[1,2,5000:20000])
-CLGPNM.Beta22.hpd <- hpd_est(res_CLGPNM$Beta[2,2,5000:20000])
-CLGPNM.Beta13.hpd <- hpd_est(res_CLGPNM$Beta[1,3,5000:20000])
-CLGPNM.Beta23.hpd <- hpd_est(res_CLGPNM$Beta[2,3,5000:20000])
-
-CLGPNM.Betacon11.m <- mode_est(res_CLGPNM$Betacon[1,1,5000:20000])
-CLGPNM.Betacon21.m <- mode_est(res_CLGPNM$Betacon[2,1,5000:20000])
-CLGPNM.Betacon12.m <- mode_est(res_CLGPNM$Betacon[1,2,5000:20000])
-CLGPNM.Betacon22.m <- mode_est(res_CLGPNM$Betacon[2,2,5000:20000])
-CLGPNM.Betacon13.m <- mode_est(res_CLGPNM$Betacon[1,3,5000:20000])
-CLGPNM.Betacon23.m <- mode_est(res_CLGPNM$Betacon[2,3,5000:20000])
-
-CLGPNM.Betacon11.hpd <- hpd_est(res_CLGPNM$Betacon[1,1,5000:20000])
-CLGPNM.Betacon21.hpd <- hpd_est(res_CLGPNM$Betacon[2,1,5000:20000])
-CLGPNM.Betacon12.hpd <- hpd_est(res_CLGPNM$Betacon[1,2,5000:20000])
-CLGPNM.Betacon22.hpd <- hpd_est(res_CLGPNM$Betacon[2,2,5000:20000])
-CLGPNM.Betacon13.hpd <- hpd_est(res_CLGPNM$Betacon[1,3,5000:20000])
-CLGPNM.Betacon23.hpd <- hpd_est(res_CLGPNM$Betacon[2,3,5000:20000])
-
-CLGPNM.Sigma11.m <- mode_est(res_CLGPNM$Sigma[1,1,5000:20000])
-CLGPNM.Sigma12.m <- mode_est(res_CLGPNM$Sigma[1,2,5000:20000])
-CLGPNM.Sigma13.m <- mode_est(res_CLGPNM$Sigma[1,3,5000:20000])
-CLGPNM.Sigma22.m <- mode_est(res_CLGPNM$Sigma[2,2,5000:20000])
-CLGPNM.Sigma23.m <- mode_est(res_CLGPNM$Sigma[2,3,5000:20000])
-CLGPNM.Sigma33.m <- mode_est(res_CLGPNM$Sigma[3,3,5000:20000])
-
-CLGPNM.Sigma11.hpd <- hpd_est(res_CLGPNM$Sigma[1,1,5000:20000])
-CLGPNM.Sigma12.hpd <- hpd_est(res_CLGPNM$Sigma[1,2,5000:20000])
-CLGPNM.Sigma13.hpd <- hpd_est(res_CLGPNM$Sigma[1,3,5000:20000])
-CLGPNM.Sigma22.hpd <- hpd_est(res_CLGPNM$Sigma[2,2,5000:20000])
-CLGPNM.Sigma23.hpd <- hpd_est(res_CLGPNM$Sigma[2,3,5000:20000])
-CLGPNM.Sigma33.hpd <- hpd_est(res_CLGPNM$Sigma[3,3,5000:20000])
-
-CLGPNM.Sigmacon11.m <- mode_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
-CLGPNM.Sigmacon12.m <- mode_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
-CLGPNM.Sigmacon13.m <- mode_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
-CLGPNM.Sigmacon22.m <- mode_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
-CLGPNM.Sigmacon23.m <- mode_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
-CLGPNM.Sigmacon33.m <- mode_est(res_CLGPNM$Sigmacon[3,3,5000:20000])
-
-CLGPNM.Sigmacon11.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
-CLGPNM.Sigmacon12.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
-CLGPNM.Sigmacon13.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
-CLGPNM.Sigmacon22.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
-CLGPNM.Sigmacon23.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
-CLGPNM.Sigmacon33.hpd <- hpd_est(res_CLGPNM$Sigmacon[3,3,5000:20000])
-
-modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
-                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
-                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
-                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
-                CLGPNM.lambda.m)
-hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
-                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
-                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
-                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
-                CLGPNM.lambda.hpd[1])
-hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
-                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
-                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
-                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
-                CLGPNM.lambda.hpd[2])
-
-modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
-                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
-                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
-                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
-                   CLGPNM.lambda.m)
-hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
-                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
-                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
-                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
-                   CLGPNM.lambda.hpd[1])
-hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
-                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
-                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
-                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
-                   CLGPNM.lambda.hpd[2])
-
-CLGPNMres.tab <- round(cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN, modescon.GPN, hpdcon.LB.GPN, hpdcon.UB.GPN),2)
-
-rownames(CLGPNMres.tab) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
-                             "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
-                             "$\\beta_{0_s^{II}}$", "$\\beta_{1_y}$",
-                             "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
-                             "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
-                             "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
-                             "$\\lambda$")
-colnames(CLGPNMres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")
-
-kable(CLGPNMres.tab, "latex", booktabs = TRUE, escape = FALSE, caption = "Results for the modified joint projected and skew normal model")%>%
-add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))
-
-```
-
-Tables 2, 3 and 4 show results from the four models that were fit to the teacher
-dataset. There are several parameters that we can compare between these results.
-Firstly, we have the estimated mean of the circular outcome. In Table 5 we see 
-the posterior (predictive) distributions for the estimated circular means of the
-CL-PN, CL-GPN and joint GPN-SSN  model. For the CL-PN model we can actually also
-compute this means from the estimates in Table 2 as follows: $\mu_{circ} = 
-atan2(\beta_0^{II}, \beta_0^{I})$. This leads to an estimate of 31.53$^\circ$. 
-The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are about equal and
-correspond quite well to the actual data mean of 33.23$^\circ$ in Table
-\ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model which is
-0.37 radians or 21.20$^\circ$ is different. This difference could be caused by
-the fact that the densities for the circular ouctome, projected normal or
-sine-skewed von Mises, differ between the models.
-
-```{r means, echo = FALSE}
-means <- round(rbind(mean_circ(res_CLPN$theta_pred[,5000:20000]),
-                     mean_circ(res_CLGPN$theta_pred[,5000:20000]),
-                     mean_circ(res_CLGPNM$theta_pred[,5000:20000]))*(180/pi), 2)
-
-rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
-colnames(means) <- "Mean (degrees)"
-kable(means, "latex", booktabs = T, escape = FALSE, caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models")%>%
-   kable_styling() %>%
-   add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", notation="alphabet")
-
-```
-
-We can also compare the effect of \verb|SE| on the circular outcome. For the 
-CL-PN model and Abe-Ley models we can get estimates of a circular regression 
-coefficient of this effect. For the Abe-Ley model this estimate is the parameter
-$\nu_1 = -0.01$. This means that for each unit increate in self-efficacy, at the
-inflection point of the circular regression line, the score of the teacher on 
-the IPC decreases with $0.01*(180/\pi) = 0.57^\circ$. For the CL-PN model we 
-can use methods from @CremersMulderKlugkist2017. The estimated posterior mode of
-$b_c$, a parameter that is comparable to $\nu_1$ in the Abe-Ley model, equals 
-$1.13$ and its 95\% HPD interval is $(-26.55; 25.46)$. This means that for each 
+Before analysis there were a couple of settings that we had to specify
+for the cylindrical models. Starting values for the WeiSSVM model were
+the following
+\(\eta_0 = 0.5, \eta_1 = 0.5, \nu_0 = 0.5, \nu_1 = 0.5, \kappa = 0.5, \alpha = 0.5, \lambda = 0\).
+The initial amount of iterations for the three MCMC samplers was set to
+2000. After we checked convergence via traceplots we concluded that some
+of the parameters of the multivariate GPN model did not converge.
+Therefore we set the amount of iterations of the MCMC models to 20,000
+and subtracted a burn-in of 5000. With this specification the MCMC
+chains for all parameters converge. Note that we choose the same amount
+of iterations for all three Bayesian cylindrical models to make their
+comparison via the PLSL as fair as possible. Lastly, the predictor
+\verb|SE| was centered before inclusion in the analysis.
+
+\begin{table}
+
+\caption{\label{tab:estCLGPN}Results for the modified CL-PN and CL-GPN model}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{CL-PN} & \multicolumn{3}{c}{CL-GPN} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
+  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
+\midrule
+$\beta_0^{I}$ & 1.63 & 1.36 & 1.87 & 2.15 & 1.64 & 2.59\\
+$\beta_0^{I}$ & 0.52 & 0.23 & 0.80 & 0.65 & 0.14 & 1.06\\
+$\beta_0^{II}$ & 1.00 & 0.80 & 1.23 & 1.24 & 0.98 & 1.51\\
+$\beta_1^{II}$ & 0.39 & 0.16 & 0.64 & 0.46 & 0.18 & 0.72\\
+$\gamma_0$ & 0.35 & 0.29 & 0.41 & 0.35 & 0.30 & 0.40\\
+\addlinespace
+$\gamma_{cos}$ & 0.05 & 0.02 & 0.07 & 0.04 & 0.02 & 0.06\\
+$\gamma_{sin}$ & -0.01 & -0.04 & 0.03 & -0.01 & -0.04 & 0.03\\
+$\gamma_1$ & 0.03 & -0.01 & 0.06 & 0.03 & -0.01 & 0.06\\
+$\sigma$ & 0.14 & 0.13 & 0.16 & 0.14 & 0.12 & 0.16\\
+$\sum_{1,1}$ & NA & NA & NA & 2.67 & 1.77 & 4.12\\
+\addlinespace
+$\sum_{1,2}$ & NA & NA & NA & 0.81 & 0.53 & 1.07\\
+$\sum_{2,2}$ & NA & NA & NA & 1.00 & 1.00 & 1.00\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:estSL}Results for the modified Abe-Ley model}
+\centering
+\begin{tabular}[t]{lr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
+$\eta_0$ & 0.37\\
+$\eta_1$ & -0.01\\
+$\nu_0$ & 1.19\\
+$\nu_1$ & 0.00\\
+$\alpha$ & 3.82\\
+\addlinespace
+$\kappa$ & 1.57\\
+$\lambda$ & 0.66\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:estCLMGPN}Results for the modified joint projected and skew normal model}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
+  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
+\midrule
+$\beta_{0_s^{I}}$ & 0.30 & 0.26 & 0.34 & 1.94 & 1.63 & 2.33\\
+$\beta_{0_s^{II}}$ & 0.19 & 0.16 & 0.21 & 1.20 & 0.98 & 1.43\\
+$\beta_{0_y}$ & 0.33 & 0.30 & 0.37 & 0.33 & 0.30 & 0.37\\
+$\beta_{1_s^{I}}$ & 0.07 & 0.03 & 0.12 & 0.50 & 0.16 & 0.79\\
+$\beta_{0_s^{II}}$ & 0.04 & 0.01 & 0.08 & 0.27 & 0.09 & 0.50\\
+\addlinespace
+$\beta_{1_y}$ & 0.07 & 0.04 & 0.12 & 0.07 & 0.04 & 0.12\\
+$\sum_{s_{1,1}}$ & 0.05 & 0.04 & 0.07 & 2.16 & 1.51 & 2.99\\
+$\sum_{s_{2,2}}$ & 0.02 & 0.02 & 0.03 & 1.00 & 1.00 & 1.00\\
+$\sum_{y_{3,3}}$ & 0.03 & 0.02 & 0.05 & 0.03 & 0.02 & 0.05\\
+$\sum_{s_{1,2}}$ & 0.00 & 0.00 & 0.01 & 0.19 & -0.06 & 0.46\\
+\addlinespace
+$\sum_{sy_{1,3}}$ & 0.04 & 0.03 & 0.05 & 0.24 & 0.18 & 0.33\\
+$\sum_{sy_{2,3}}$ & 0.01 & 0.01 & 0.02 & 0.10 & 0.06 & 0.13\\
+$\lambda$ & 0.14 & 0.13 & 0.17 & 0.14 & 0.13 & 0.17\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Tables 2, 3 and 4 show results from the four models that were fit to the
+teacher dataset. There are several parameters that we can compare
+between these results. Firstly, we have the estimated mean of the
+circular outcome. In Table 5 we see the posterior (predictive)
+distributions for the estimated circular means of the CL-PN, CL-GPN and
+joint GPN-SSN model. For the CL-PN model we can actually also compute
+this means from the estimates in Table 2 as follows:
+\(\mu_{circ} = atan2(\beta_0^{II}, \beta_0^{I})\). This leads to an
+estimate of 31.53\(^\circ\). The estimates for the CL-PN, CL-GPN and
+joint GPN-SSN model are about equal and correspond quite well to the
+actual data mean of 33.23\(^\circ\) in Table
+\ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model
+which is 0.37 radians or 21.20\(^\circ\) is different. This difference
+could be caused by the fact that the densities for the circular ouctome,
+projected normal or sine-skewed von Mises, differ between the models.
+
+\begin{table}
+
+\caption{\label{tab:means}Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models}
+\centering
+\begin{tabular}[t]{lr}
+\toprule
+  & Mean (degrees)\\
+\midrule
+Cl-PN & 31.98\\
+CL-GPN & 32.98\\
+GPN-SSN & 34.50\\
+\bottomrule
+\multicolumn{2}{l}{\textsuperscript{a} Note that these means}\\
+\multicolumn{2}{l}{are based on the}\\
+\multicolumn{2}{l}{posterior predictive}\\
+\multicolumn{2}{l}{distribution for the}\\
+\multicolumn{2}{l}{intercepts following}\\
+\multicolumn{2}{l}{(Wang \& Gelfand,}\\
+\multicolumn{2}{l}{2013).}\\
+\end{tabular}
+\end{table}
+
+We can also compare the effect of \verb|SE| on the circular outcome. For
+the CL-PN model and Abe-Ley models we can get estimates of a circular
+regression coefficient of this effect. For the Abe-Ley model this
+estimate is the parameter \(\nu_1 = -0.01\). This means that for each
 unit increate in self-efficacy, at the inflection point of the circular
-regression line, the score of the teacher on the IPC increases with 
-$1.13*(180/\pi) = 74.74^\circ$. Values of $b_c$ and $\nu_1$ are quite 
-different even though they should represent the same effect. However, the slope 
-at the inflection point, which $b_c$ and $\nu_1$ describe is not necessarily 
-representative of the effect of \verb|SE| in the data range. In Figure \ref{} we
-see that the inflection point (square) for the CL-PN model lies just ouside of 
-the data range and that for the Abe-Ley model even further (outside the x-axis 
-range). The slope of both regression lines in the data range is much more alike
-(even though one is positive and the other negative). The slope at the mean
-Self-efficicay, which has a value of 0 because it was centered, $SAM$ is
-estimated at $0.02 (-0.09; 0.15)$ The $SAM$ can be interpreted as the amount of
-change on the circle for a unit increase at the mean of the predictor variable.
-However, for the CL-PN model the 95\% HPD interval of $SAM$ for the effect of
-\verb|SE| includes 0 meaning that the value is not different from 0. For the
-Abe-Ley model standard errors of the parameters are not known so we cannot
-formally test whether $\nu_1$ differs from 0. Neither is there a known way of
-computing the effect at predictor values other than the one of the inflection
-point.
-
-For the CL-GPN and multivariate 
-CL-GPN models we cannot compute circular regression coefficients. Instead, we 
-will compute posterior predictive distributions for the predicted circular
-outcome of individuals scoring 1 unit above and 1 unit below the average
-self-efficacy.
-
-```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("regline.tex", standAlone =TRUE, height = 4, width = 6, pointsize = 12, engine = "pdftex")
-
-a1 <- mode_est(res_CLPN$BI[5000:20000, 1])
-b1 <- mode_est(res_CLPN$BI[5000:20000, 2])
-a2 <- mode_est(res_CLPN$BII[5000:20000, 1])
-b2 <- mode_est(res_CLPN$BII[5000:20000, 2])
-
-x <- seq(-5, 5, by = 0.1)
-
-pred.AL <- 0.37 + 2*atan(-0.01*x)
-pred.PN <- atan2(a2 + b2*x, a1 + b1*x)
-
-plot(Dat.test$SEc, theta.test, bty = "n",
-     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "Self-efficacy")
-points(x, pred.AL, type = "l", lty = 1)
-points(x, pred.PN, type = "l", lty = 2)
-
-dev.off()
-tools::texi2dvi("regline.tex", pdf=TRUE)
-
-```
-
-```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}
-
-a1 <- res_CLPN$BI[5000:20000, 1]
-b1 <- res_CLPN$BI[5000:20000, 2]
-a2 <- res_CLPN$BII[5000:20000, 1]
-b2 <- res_CLPN$BII[5000:20000, 2]
-
-ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-bc <- -tan(atan2(a2, a1)-ac)/ax
-
-mode_est(ax)
-
-mode_est(bc)
-hpd_est(bc)
-
-mode_est(bc/(1 + (bc*(-ax))^2))
-hpd_est(bc/(1 + (bc*(-ax))^2))
-
-```
+regression line, the score of the teacher on the IPC decreases with
+\(0.01*(180/\pi) = 0.57^\circ\). For the CL-PN model we can use methods
+from Cremers et al. (2018b). The estimated posterior mode of \(b_c\), a
+parameter that is comparable to \(\nu_1\) in the Abe-Ley model, equals
+\(1.13\) and its 95\% HPD interval is \((-26.55; 25.46)\). This means
+that for each unit increate in self-efficacy, at the inflection point of
+the circular regression line, the score of the teacher on the IPC
+increases with \(1.13*(180/\pi) = 74.74^\circ\). Values of \(b_c\) and
+\(\nu_1\) are quite different even though they should represent the same
+effect. However, the slope at the inflection point, which \(b_c\) and
+\(\nu_1\) describe is not necessarily representative of the effect of
+\verb|SE| in the data range. In Figure \ref{} we see that the inflection
+point (square) for the CL-PN model lies just ouside of the data range
+and that for the Abe-Ley model even further (outside the x-axis range).
+The slope of both regression lines in the data range is much more alike
+(even though one is positive and the other negative). The slope at the
+mean Self-efficicay, which has a value of 0 because it was centered,
+\(SAM\) is estimated at \(0.02 (-0.09; 0.15)\) The \(SAM\) can be
+interpreted as the amount of change on the circle for a unit increase at
+the mean of the predictor variable. However, for the CL-PN model the
+95\% HPD interval of \(SAM\) for the effect of \verb|SE| includes 0
+meaning that the value is not different from 0. For the Abe-Ley model
+standard errors of the parameters are not known so we cannot formally
+test whether \(\nu_1\) differs from 0. Neither is there a known way of
+computing the effect at predictor values other than the one of the
+inflection point.
+
+For the CL-GPN and multivariate CL-GPN models we cannot compute circular
+regression coefficients. Instead, we will compute posterior predictive
+distributions for the predicted circular outcome of individuals scoring
+1 unit above and 1 unit below the average self-efficacy.
 
@@
 
-Next, we compare the intercepts and regression coefficients for the linear outcome.
+Next, we compare the intercepts and regression coefficients for the
+linear outcome.
 
@@
 
-```{r ModelFit, echo = FALSE}
-
-CLGPNM.ll.circ <- -2*mean(res_CLGPNM$ll.circ[5000:20000])
-CLGPNM.ll.lin  <- -2*mean(res_CLGPNM$ll.lin[5000:20000])
-
-CLGPN.ll.circ <- -2*mean(res_CLGPN$ll.circ[5000:20000])
-CLGPN.ll.lin  <- -2*mean(res_CLGPN$ll.lin[5000:20000])
-
-CLPN.ll.circ <- -2*mean(res_CLPN$ll.circ[5000:20000])
-CLPN.ll.lin  <- -2*mean(res_CLPN$ll.lin[5000:20000])
-
-AL.ll.circ <- -2*func.reg.cond.circ(resAL$par, dat.h)
-AL.ll.lin <- -2*func.reg.cond.lin(resAL$par, dat.h)
-
-PLSL <- round(rbind(c(CLPN.ll.circ, CLGPN.ll.circ, AL.ll.circ, CLGPNM.ll.circ),
-                    c(CLPN.ll.lin, CLGPN.ll.lin, AL.ll.lin, CLGPNM.ll.lin)), 2)
-rownames(PLSL) <- c("circular", "linear")
-colnames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "Joint GPN-SSN")
-
-kable(PLSL, "latex", booktabs = T, escape = FALSE, caption = "PLSL criteria for the circular and linear outcome in the four cylindrical models")
-
-```
-
-```{r preddist, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-source("Sample Abe-Ley.R")
-
-ALdy <- matrix(NA, length(y.test), 1000)
-ALdth <- matrix(NA, length(theta.test), 1000)
-
-ALdy.h <- matrix(NA, length(y.hold), 1000)
-ALdth.h <- matrix(NA, length(theta.hold), 1000)
-
-tikz("preddist.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")
-
-layout(matrix(c(1,2,3, 4,5,5), 3, 2, byrow = TRUE),heights=c(2,2,1))
-
-par(mar = c(5,5,1,1))
-
-x.h <- Dat.hold$SEc
-x <- Dat.test$SEc
-
-for(i in 1:length(theta.test)){
-  beta <- exp(resAL$par[3] + resAL$par[4]*x[i])
-  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x[i])
-
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
-  ALdy[i,] <- dat$x
-  ALdth[i,] <- dat$theta
-}
-
-for(i in 1:length(theta.hold)){
-  beta <- exp(resAL$par[3] + resAL$par[4]*x.h[i])
-  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x.h[i])
-
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
-  ALdy.h[i,] <- dat$x
-  ALdth.h[i,] <- dat$theta
-
-}
-
-hist(theta.test, prob = TRUE, ylim = c(0, 1.25),
-     main = "", ylab = "Circular", xlab = "$\\theta_{test}$")
-lines(density(res_CLPN$theta_pred[,19000:20000]), lty = 1)
-lines(density(res_CLGPN$theta_pred[,19000:20000]), lty = 2)
-lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 3)
-lines(density(ALdth), lty = 4)
-
-hist(theta.hold, prob = TRUE, ylim = c(0, 1.25),
-     main = "", ylab = "Circular", xlab = "$\\theta_{holdout}$")
-lines(density(res_CLPN$theta_pred.hold[,19000:20000]), lty = 1)
-lines(density(res_CLGPN$theta_pred.hold[,19000:20000]), lty = 2)
-lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 3)
-lines(density(ALdth.h), lty = 4)
-
-hist(y.test, prob = TRUE, ylim = c(0, 3),
-     main = "", ylab = "Linear", xlab = "$y_{test}$")
-lines(density(res_CLPN$y_pred[,5000:20000]), lty = 1)
-lines(density(res_CLGPN$y_pred[,5000:20000]), lty = 2)
-lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 3)
-lines(density(ALdy), lty = 4)
-
-hist(y.hold, prob = TRUE, ylim = c(0, 4),
-     main = "", ylab = "Linear", xlab = "$y_{holdout}$")
-lines(density(res_CLPN$y_pred.hold[,5000:20000]), lty = 1)
-lines(density(res_CLGPN$y_pred.hold[,5000:20000]), lty = 2)
-lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 3)
-lines(density(ALdy.h), lty = 4)
-
-par(mar = c(2,2,1,1))
-
-plot(1, type = "n", axes=FALSE, xlab="", ylab="")
-legend(x = "top", inset = 0, legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"), lty = c(1,2,4,3), horiz = TRUE)
-
-dev.off()
-
-tools::texi2dvi("preddist.tex",pdf=TRUE)
-
-```
+\begin{table}
+
+\caption{\label{tab:ModelFit}PLSL criteria for the circular and linear outcome in the four cylindrical models}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+  & CL-PN & CL-GPN & Abe-Ley & Joint GPN-SSN\\
+\midrule
+circular & 86.15 & 52.88 & 58.53 & 103.83\\
+linear & -20.17 & -20.32 & 51.83 & 11.98\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \section*{References}
-<div id="refs"></div>
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-abe2017tractable}{}%
+Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
+for cylindrical data, with applications. \emph{Econometrics and
+Statistics}, \emph{4}, 91--104.
+
+\leavevmode\hypertarget{ref-chrastil2017rotational}{}%
+Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
+integration: Encoding and execution errors in angle reproduction.
+\emph{Experimental Brain Research}, \emph{235}(6), 1885--1897.
+
+\leavevmode\hypertarget{ref-Claessens2016side}{}%
+Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
+Teachers' perceptions of teacher--student relationships} (PhD thesis).
+
+\leavevmode\hypertarget{ref-Cremers2018Assessing}{}%
+Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
+bayesian embedding approach to circular regression models.
+\emph{Methodology}.
+
+\leavevmode\hypertarget{ref-CremersMulderKlugkist2017}{}%
+Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
+interpretation of regression coefficients. \emph{British Journal of
+Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
+doi:\href{https://doi.org/10.1111/bmsp.12108}{10.1111/bmsp.12108}
+
+\leavevmode\hypertarget{ref-garcia2014test}{}%
+García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
+González-Manteiga, W., \& Pereira, J. (2014). A test for
+directional-linear independence, with applications to wildfire
+orientation and size. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{28}(5), 1261--1275.
+
+\leavevmode\hypertarget{ref-garcia2013exploring}{}%
+García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
+Exploring wind direction and so2 concentration by circular--linear
+density estimation. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{27}(5), 1055--1067.
+
+\leavevmode\hypertarget{ref-BDA}{}%
+Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
+(2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman
+\& Hall/CRC.
+
+\leavevmode\hypertarget{ref-gneiting2007strictly}{}%
+Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
+prediction, and estimation. \emph{Journal of the American Statistical
+Association}, \emph{102}(477), 359--378.
+
+\leavevmode\hypertarget{ref-gurtman2009exploring}{}%
+Gurtman, M. B. (2009). Exploring personality with the interpersonal
+circumplex. \emph{Social and Personality Psychology Compass},
+\emph{3}(4), 601--619.
+doi:\href{https://doi.org/10.1111/j.1751-9004.2009.00172.x}{10.1111/j.1751-9004.2009.00172.x}
+
+\leavevmode\hypertarget{ref-hernandez2016general}{}%
+Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
+(2016). The general projected normal distribution of arbitrary
+dimension: Modeling and bayesian inference. \emph{Bayesian Analysis},
+\emph{12}(1).
+doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}
+
+\leavevmode\hypertarget{ref-horowitz2010handbook}{}%
+Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
+psychology: Theory, research, assessment, and therapeutic
+interventions}. Hoboken, NJ: John Wiley \& Sons.
+
+\leavevmode\hypertarget{ref-jammalamadaka2001topics}{}%
+Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
+statistics} (Vol. 5). World Scientific.
+
+\leavevmode\hypertarget{ref-johnson1978some}{}%
+Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
+distributions and related regression models. \emph{Journal of the
+American Statistical Association}, \emph{73}(363), 602--606.
+
+\leavevmode\hypertarget{ref-lagona2015hidden}{}%
+Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
+markov approach to the analysis of space--time environmental data with
+linear and circular components. \emph{Stochastic Environmental Research
+and Risk Assessment}, \emph{29}(2), 397--409.
+
+\leavevmode\hypertarget{ref-ley2017modern}{}%
+Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
+CRC Press.
+
+\leavevmode\hypertarget{ref-mardia2000directional}{}%
+Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
+(Vol. 494). Chichester, England: Wiley.
+
+\leavevmode\hypertarget{ref-mardia1978model}{}%
+Mardia, K., \& Sutton, T. (1978). A model for cylindrical variables with
+applications. \emph{Journal of the Royal Statistical Society. Series B
+(Methodological)}, 229--233.
+
+\leavevmode\hypertarget{ref-mastrantonio2018joint}{}%
+Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
+distribution for poly-cylindrical data. \emph{Journal of Multivariate
+Analysis}, \emph{165}, 14--26.
+
+\leavevmode\hypertarget{ref-mastrantonio2015bayesian}{}%
+Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
+hidden markov modelling using circular-linear general projected normal
+distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
+
+\leavevmode\hypertarget{ref-nunez2011bayesian}{}%
+Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
+Bayesian regression model for circular data based on the projected
+normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
+185--201.
+doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
+
+\leavevmode\hypertarget{ref-olmos2012extension}{}%
+Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
+extension of the half-normal distribution. \emph{Statistical Papers},
+\emph{53}(4), 875--886.
+
+\leavevmode\hypertarget{ref-pennings2017interpersonal}{}%
+Pennings, H. J., Brekelmans, M., Sadler, P., Claessens, L. C., Want, A.
+C. van der, \& Tartwijk, J. van. (2017). Interpersonal adaptation in
+teacher-student interaction. \emph{Learning and Instruction}.
+
+\leavevmode\hypertarget{ref-rayner200935th}{}%
+Rayner, K. (2009). The 35th sir frederick bartlett lecture: Eye
+movements and attention in reading, scene perception, and visual search.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8),
+1457--1506.
+
+\leavevmode\hypertarget{ref-sahu2003new}{}%
+Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
+multivariate skew distributions with applications to bayesian regression
+models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
+
+\leavevmode\hypertarget{ref-wang2012directional}{}%
+Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
+general projected normal distribution. \emph{Statistical Methodology},
+\emph{10}(1, 1), 113--127.
+doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}
+
+\leavevmode\hypertarget{ref-wang2015joint}{}%
+Wang, F., Gelfand, A. E., \& Jona-Lasinio, G. (2015). Joint
+spatio-temporal analysis of a linear and a directional variable:
+Space-time modeling of wave heights and wave directions in the adriatic
+sea. \emph{Statistica Sinica}, 25--39.
+
+\leavevmode\hypertarget{ref-vanderWant2015role}{}%
+Want, A. C. van der. (2015). \emph{Teachers' interpersonal role
+identity.} (PhD thesis).
+
+\leavevmode\hypertarget{ref-wubbels2006interpersonal}{}%
+Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
+An interpersonal perspective on classroom management in secondary
+classrooms in the netherlands. In C. Evertson \& C. S. Weinstein (Eds.),
+\emph{Handbook of classroom management: Research, practice, and
+contemporary issues} (pp. 1161--1191). Malwah, NJ: Lawrence Erlbaum
+Associates.
 
@@
 \end{appendices}
+
+\end{document}

### 2018-07-02 Manuscript/ManuscriptMarkdownDifferentSE.Rmd
Manuscript Update

--- 2018-07-02 Manuscript/ManuscriptMarkdown.Rmd
@@
 ---
-title: "Working title: Regression models for Cylindrical data."
+title: "A Comparison of Regression models for Cylindrical data in Psychology."
 author: |
@@
   - \usepackage{subcaption}
+
 keywords: bla vla
@@
 abstract: |
-  This is the abstract
+  Cylindrical dats is multivariate data which consists of a directional, in this paper circular, component   and a linear component.
+Examples of cylindrical data in psychology include human navigation (direction     and distance of movement), eye-tracking research (direction and length of saccades) and data from an       interpersonal circumplex (type and strength of interpersonal behavior).
+In this paper we adapt four        models for cylindrical data to include a regression of the circular and linear component onto a set of     covariates.
+Subsequently, we illustrate how to fit these models and interpret their results on a dataset   on the interpersonal behavior of teachers.
 ---
@@
 library(kableExtra) #latex tables
+library(plotrix)
 ```
@@
 
-Cylindrical data are data that consist of a linear component and a directional 
-component (a set of angles). In this paper, the directional component is 
-circular meaning that it consists of a single angle instead of a set of angles. 
-Figure \ref{} \textcolor{red}{NOG MAKEN} shows the difference between a circular variable (left) and a 
-linear variable (right). The most obviuous difference is that on a circular 
-scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in fact 
-represent the same number while on a linear scale the two ends, $-\infty$ and 
-$\infty$ are not connected. Cylindrical data occur in several fields of 
-research, such as for instance in meteorology [@garcia2013exploring], ecology 
-[@garcia2014test] or marine research [@lagona2015hidden]. Several types of data 
-in psychological research are also of a cylindrical nature. For example, in 
-research on human navigation in the field of cognitive psychology both distance,
-a linear variable, and direction, a circular variable, of movement are of 
-interest [@chrastil2017rotational]. In eye-tracking research, which is used for 
-investigating cognitive processes we can also speak of cylindrical data. When 
+Cylindrical data are data that consist of a linear variable and a directional 
+variable. In this paper, the directional variable is circular meaning that it 
+consists of a single angle instead of a set of angles. A circular variable is 
+different from a linear variable in the sense that it is measured on a different
+scale. Figure \ref{circline} shows the difference between a circular scale 
+(right) and a linear scale (left). The most important difference is that on a 
+circular scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in 
+fact represent the same number while on a linear scale the two ends, $-\infty$ 
+and $\infty$ are not connected. This difference requires us to use different 
+statistical methods for circular variables (see e.g. @fisher1995statistical for 
+an introduction to circular data and @mardia2000directional,
+@jammalamadaka2001topics and @ley2017modern for a more elaborate overview).
+
+Cylindrical data occur in several fields of research, such as for instance in 
+meteorology [@garcia2013exploring], ecology [@garcia2014test] or marine research
+[@lagona2015hidden]. Several types of data in psychology are also of a 
+cylindrical nature. For example, in research on human navigation in the field of
+cognitive psychology both distance, a linear variable, and direction, a circular
+variable, of movement are of interest [@chrastil2017rotational]. In eye-tracking
+research, which can be used for investigating various cognitive processes (e.g. 
+those involved in reading a text), we can also speak of cylindrical data. When 
 measuring eye-movements we speak of the eye-movements themselves, the saccades, 
-and fixations, the periods of time between movements when the eyes are still and
-looking at one point. Of the saccades both the direction, a circular variable, 
-and the duration, a linear variable, are of interest (for a review of 
-eye-tracking research see @rayner200935th). Data from circumplex measurement
-instruments, e.g. the interpersonal circumplex as used in personality
-psychology, are also of a cylindrical nature (see Section \ref{Example} for a
-more detailed explanation). In this paper we will discuss how a correct
-statistical treatment of such cylindrical data can lead to new insights. In
-particular we will show how cylindrical models pave the way for circular-linear
-and linear-circular regression. We will do this for a motivating example from
-the field of educational psychology. We will refer to this example data as the
-teacher data. In this example, apart from modelling the relation between the
-linear and circular component of a cylindrical variable we would also like to
-predict the two components from a set of covariates in a regression model. The
-teacher data will be further introduced in Sections \ref{Example} and
-\ref{DataAnalysis}.
+and fixations, the periods of time between movements when the eyes are looking 
+at one point. Of the saccades both the direction, a circular variable, and the 
+duration, a linear variable, are of interest (for a review of eye-tracking 
+research see @rayner200935th). Data from circumplex measurement instruments,
+e.g. the interpersonal circumplex as used in personality psychology, are also of
+a cylindrical nature (see Section \ref{Example} for a more detailed
+explanation).
+
+In this paper we will discuss how a correct statistical treatment of such 
+cylindrical data can lead to new insights. In particular we will show how 
+cylindrical models pave the way for circular-linear and linear-circular 
+regression. We will do this for a motivating example, the teacher data,  from
+the field of educational psychology . In this example, apart from modelling the
+relation between the linear and circular component of a cylindrical variable we
+would also like to predict the two components from a set of covariates in a
+regression model. The teacher data will be further introduced in Sections
+\ref{Example} and \ref{DataAnalysis}.
+
+\begin{figure}
+\centering
+\includegraphics[width = 0.8\textwidth]{Plots/circline.pdf}
+\caption{The difference between a linear scale (left) and a circular scale (right).}
+\label{circline}
+\end{figure}
 
 As is the case for circular data, the analysis of cylindrical data requires 
-special methods. For an overview of methods for circular data see
-@mardia2000directional and @jammalamadaka2001topics and for modern treatments
-@ley2017modern. In the literature several methods have been put forward to
-model the relation between the linear and circular component of a cylindrical 
-variable. Some of these are based on regressing the linear component onto the 
-circular component using the following type of relation: $$y = \beta_0 + 
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear
-component and $\theta$ the circular component [@mardia1978model;
+special methods. Several methods have been put forward to model the relation 
+between the linear and circular component of a cylindrical variable. Some of 
+these are based on regressing the linear component onto the circular component 
+using the following type of relation: $$y = \beta_0 + \beta_1*\cos(\theta) + 
+\beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear component and $\theta$
+the circular component [@mardia1978model;
 @johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a 
@@
 circular variables and modelling their covariance matrix 
-[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. 
-For example, @abe2017tractable introduce a cylindrical distribution  based on a 
+[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For 
+example, @abe2017tractable introduce a cylindrical distribution based on a 
 Weibull distribution for the linear component and a sine-skewed von Mises 
@@
 concentration parameters respectively. However, none of the methods that have 
-been proposed thus far include additional covariates on which both the circular 
-and linear component are regressed.
-
-Our aim in this paper is to fill this gap in the literature by  adapting four of
-the existing methods for the analysis of cylindrical data in such a way that
-they include a regression of both the linear and circular component of a
-cylindrical variable onto a set of covariates. First however, we will introduce
-the teacher data in Section \ref{Example}. In Section \ref{Models} we introduce 
-the four modified models for cylindrical data that we use to analyze the data from
-the motivating example. We also choose a model selection criterion to compare
-the models. The teacher data will be analysed in Section \ref{DataAnalysis}. The
-paper will be concluded with a discussion in Section \ref{Discussion}. The
-Appendix contains technical details.
+been proposed thus far include additional covariates onto which both the
+circular and linear component are regressed.
+
+Our aim in this paper is to fill this gap in the literature by adapting four
+existing cylindrical methods in such a way that they include a regression of
+both the linear and circular component of a cylindrical variable onto a set of
+covariates. First however, we will introduce the teacher data in Section
+\ref{Example}. In Section \ref{Models} we introduce the four modified models for
+cylindrical data that we use to analyze the data from the motivating example. We
+also choose a model selection criterion to compare the models. The teacher data
+will be analysed in Section \ref{DataAnalysis}. The paper will be concluded with
+a discussion in Section \ref{Discussion}. The Appendix contains technical
+details.
 
@@
 The motivating example for this article comes from the field of educational 
-psychology and was collected for the studies of @pennings2017interpersonal,
-@Claessens2016side and @vanderWant2015role. These studies concern research on 
-classroom climate. An indicator of classroom climate is the students'
-perceptions of their teachers' interpersonal behavior. These interpersonal
-perceptions, both in educational psychology as well as in other areas of
-psychology, can be measured using circumplex measurement instruments (see
+psychology and was collected for the studies on classroom climate of
+@pennings2017interpersonal, @Claessens2016side and @vanderWant2015role. An 
+indicator of classroom climate is the students' perceptions of their teachers'
+interpersonal behavior. These interpersonal perceptions, both in educational
+psychology as well as in other areas of psychology, can be measured using
+circumplex measurement instruments (see
 @horowitz2010handbook for an overview of many such instruments).
@@
 The Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] is
-one such circumplex measurement instrument that can be used to study student
-perceptions of their teachers' interpersonal behavior. This instrument contains
+one such circumplex measurement instrument. It is used to study student
+perceptions of their teachers' interpersonal behavior and contains
 items that load on two dimensions: Agency and Communion. Agency refers to the
@@
 \centering
-\includegraphics[width = 0.8\textwidth]{IPC-T.png}
+\includegraphics[width = 0.8\textwidth]{Plots/IPC-T.png}
 \caption{The interpersonal circle for teachers (IPC-T). The words presented in
@@
 
-@Cremers2018Assessing show how data from the IPC can be
-considered circular data and analyzed as such using a regression model. The 
-two-dimension scores Agency and Communion can be converted to a circular score 
-using the two-argument arctangent function in Equation \eqref{PredVal}, where 
-$A$ represents a score on the Agency dimension and $C$ represents a score on the
-Communion dimension. The resulting circular variable $\theta$ can then be
-modeled and for example be regressed onto covariates in a circular regression 
-model. However, when two-dimensional data are converted to the circle we lose 
-some information, namely the length of the two-dimensional vector $(A, C)$, 
-\emph{i.e.}, its euclidean norm $\mid\mid (A, C) \mid\mid$. This length 
-represents the strength of the type of interpersonal behavior a teacher shows 
-towards his/her students. In a cylindrical model we are able to incorporate this
-information, and model a circular variable $\theta$ together with a linear 
-variable corresponding to $\mid\mid (A, C) \mid\mid$. This will lead to an 
-improved analysis of data from the IPC.  In the next section we will introduce 
-several models that we can use for a more accurate and more informative regression
-analysis on the teacher data. More information on the teacher data will be given
-in Section \ref{DataAnalysis}.
+@Cremers2018Assessing show how data from the IPC can be considered circular data
+and analyzed as such using a regression model. The two-dimension scores Agency
+and Communion can be converted to a circular score using the two-argument
+arctangent function in \eqref{PredVal}, where $A$ represents a score on
+the Agency dimension and $C$ represents a score on the Communion dimension. The
+resulting circular variable $\theta$ can then be modeled. However, when
+two-dimensional data are converted to the circle we lose some information, the
+length of the two-dimensional vector $(A, C)$, \emph{i.e.}, its euclidean norm
+$\mid\mid (A, C) \mid\mid$. This length represents the strength of the type of
+interpersonal behavior a teacher shows towards his/her students. In a
+cylindrical model we are able to incorporate this information, and model a
+circular variable $\theta$ together with a linear variable corresponding to
+$\mid\mid (A, C) \mid\mid$. This leads to an improved analysis of data from the
+IPC. In the next section we introduce several models that can be used for more
+accurate and informative regression analysis on the teacher data. Descriptives
+for the teacher data are given in Section \ref{DataAnalysis}.
 
@@
 #center and remove missings for self-efficacy
-Dat <- Dat %>% drop_na(Efficacy_SE)
-Dat$SEc <- Dat$Efficacy_SE- mean(Dat$Efficacy_SE)
+Dat <- Dat %>% drop_na(Efficacy_CM)
+Dat$SEc <- Dat$Efficacy_CM- mean(Dat$Efficacy_CM)
 
@@
 
-In this section we will introduce four cylindrical regression models. We will 
-extend the models from @mastrantonio2018joint and @abe2017tractable to contain 
-predictors for the linear and circular outcomes in the cylindrical model. 
-Additionally we introduce two models where both variates $\Theta$ and $Y$ are
-predicted by covariates and that model the relation between the linear and
-circular outcome as follows (following @wang2015joint and
+In this section we introduce four cylindrical regression models to contain 
+predictors for the linear and circular outcomes, $\Theta$ and $Y$. We extend the models from @mastrantonio2018joint and
+@abe2017tractable. Additionally, we introduce two models where the relation
+between $\Theta$ and $Y$ is modelled as follows (following
 @mastrantonio2015bayesian):
@@
 
-In both of these models the relation between the circular variable $\Theta \in
-[0, 2\pi)$ and linear variable $Y \in (-\infty, + \infty)$ is specified as in
-Equation \ref{circlinlink}. The distribution of $Y$ conditional on $\Theta$ is
-then given as:
+In both of these models the relation between $\Theta \in [0, 2\pi)$ and $Y\in 
+(-\infty, + \infty)$ is specified as in \eqref{circlinlink} with the following
+conditional distribution:
 
@@
 
-In a regression setup the outcome $\theta_i$ for each individual $i = 1,
-\dots, n$, where $n$ is the sample size, is generated independently from
-Equation (\ref{pnreg}). The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is 
-defined as $\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
+In a regression setup the outcome $\theta_i$ for each individual $i = 1, \dots,
+n$, where $n$ is the sample size, is generated independently from \eqref{pnreg}.
+The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is defined as
+$\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
 $\boldsymbol{z}_i$ is a a vector of $p$ covariate values and $\boldsymbol{B} = 
@@
 \end{bmatrix}$. In a regression setup the outcome $\theta_i$ for each individual
-$i = 1, \dots, n$, where $n$ is the sample size, is generated independently
-from Equation (\ref{gpnreg}). The mean vector $\boldsymbol{\mu}_i \in 
-\mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
+is generated independently from \eqref{gpnreg}. The mean vector 
+$\boldsymbol{\mu}_i \in \mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
 \boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. The vector 
@@
 
-Both cylindrical models introduced in this section are estimated using MCMC 
-methods. These methods are based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in
-Appendices \ref{A1} and \ref{A2}.
+Both cylindrical models introduced here are estimated using MCMC methods based
+on @nunez2011bayesian, @wang2012directional and @hernandez2016general for the 
+regression of the circular outcome. A detailed description of the Bayesian
+estimation and MCMC samplers can be found in Appendices \ref{A1} and \ref{A2}.
 
@@
 This model is an extension of the cylindrical model introduced by
-@abe2017tractable to the regression context. The density of the model for the
-pair of the circular variable $\theta \in [0, 2\pi)$ and linear variable $y$ on
-the positive real half-line $[0, + \infty)$ is:
+@abe2017tractable to the regression context. The joint density of $\Theta$ and
+$Y$, in this model defined on the positive real half-line $[0, + \infty)$, is:
 
@@
 In a regression setup the outcome vector $(\theta_i, y_i)^t$ for each individual
-$i = 1, \dots, n$, where $n$ is the sample size, is generated independently from
-Equation (\ref{WeiSSVMdensity}). The parameter $\alpha > 0$ is a linear shape 
-parameter, $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ is a linear 
-scale parameter, $\mu_i = \eta_0 + 
-2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$ is a circular 
-location parameter, $\kappa > 0$ is a cicular concentration parameter and 
-$\lambda \in [-1, 1]$ is a circular skewness parameter. The parameter 
-$\boldsymbol{\nu}$ is a vector of $q$ regression coefficients and intercept
-$\nu_j \in (-\infty, +\infty)$ where $j = 0, \dots, q$, for the prediction of
-$y$. The parameter $\eta_0 \in [0, 2\pi)$ is the intercept and
-$\boldsymbol{\eta}$, is a vector of $p$ regression coefficients $\eta_j \in 
-(-\infty, +\infty)$, where $j = 1, \dots, p$, for the prediction of $\theta$. 
-The vector $\boldsymbol{x}_i$ is a vector of predictor values for the prediction
-of $y$ and $\boldsymbol{z}_i$ is a vector of predictor values for the prediction
-of $\theta$.
+is generated independently from \eqref{WeiSSVMdensity}. The parameters $\alpha >
+0$ and $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ are linear shape
+and scale parameters, $\mu_i = \eta_0 + 
+2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$, $\kappa > 0$ and
+$\lambda \in [-1, 1]$ are circular location, concentration and skewness 
+parameters. The parameter $\boldsymbol{\nu}$ is a vector of $q$ regression 
+coefficients $\nu_j \in (-\infty, +\infty)$ for the prediction of $y$ where $j =
+0, \dots, q$ and $\nu_0$ is the intercept. The parameter $\eta_0 \in [0, 2\pi)$
+is the intercept and $\boldsymbol{\eta}$ is a vector of $p$ regression
+coefficients $\eta_j \in (-\infty, +\infty)$ for the prediction of $\theta$
+where $j = 1, \dots, p$. The vector $\boldsymbol{x}_i$ is a vector of predictor
+values for the prediction of $y$ and $\boldsymbol{z}_i$ is a vector of predictor
+values for the prediction of $\theta$.
 
@@
 This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. The model contains $p$
-circular outcomes and $q$ linear outcomes. The circular outcomes 
-$\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ 
-are modeled together by a multivariate GPN distribution and the linear outcomes 
-$\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled 
-together by a multivariate skew normal distribution [@sahu2003new]. Because the
-GPN distribution is modelled using a so-called augmented representation (as in
-(\ref{projection}) and (\ref{gpnreg})) it is convenient to use a similar tactic
-for modelling the multivariate skew normal distribution. As in
-@mastrantonio2018joint we use the following representation of the linear
-outcomes:
+@mastrantonio2018joint to the regression context. It contains $p$ circular
+outcomes and $q$ linear outcomes. The circular outcomes $\boldsymbol{\Theta} = 
+(\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ are modeled together by a
+multivariate GPN distribution and the linear outcomes $\boldsymbol{Y} = 
+(\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled together by a 
+multivariate skew normal distribution [@sahu2003new]. Because the GPN 
+distribution is modelled using a so-called augmented representation (as in 
+\eqref{projection} and \eqref{gpnreg}) it is convenient to use a similar tactic
+for modelling the multivariate skew normal distribution. Following
+@mastrantonio2018joint the linear outcomes are represented as:
 
@@
 \times q$ diagonal matrix with diagonal elements $\boldsymbol{\lambda} = 
-\lambda_1, \dots, \lambda_q$ (the parameters that indicate the skewness of the 
-linear outcomes), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
+\lambda_1, \dots, \lambda_q$ (skewness parameters), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
 \boldsymbol{I}_q)$, a q-dimensional half normal distribution
-[@olmos2012extension] and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
+[@olmos2012extension], and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
 \boldsymbol{\Sigma}_y)$. This means that conditional on the auxiliary data
@@
 circular outcomes, $q$ linear outcomes and $k$ covariates. The mean in the 
-density in (\ref{YDThetarjoint}) then becomes $\boldsymbol{\mu}_i = 
+density in \eqref{YDThetarjoint} then becomes $\boldsymbol{\mu}_i = 
 \boldsymbol{x}_i\boldsymbol{B}$ where $\boldsymbol{B}$ is a $k \times (2 
@@
 Note that although we fit the CL-PN, CL-GPN and joint GPN-SSN models using 
-Bayesian statistics, we do not take prior information (only the likelihood) into
-account when assessing model fit. According to @BDA this is not necessary since
-we are assessing the fit of a model to data only \textcolor{red}{AFMAKEN!}. For 
-each of the cylindrical models we can then compute a PLSL for the circular and 
-linear outcome by using the conditional log-likelihoods of the respective 
+Bayesian statistics, we do not take prior information into account when 
+assessing model fit. According to @BDA \textcolor{red}{(p. ???)} this is not 
+necessary since we are assessing the fit of a model to data, the holdout set,
+only. They argue that the prior in such case is only of interest for estimating
+the parameters of the model but not for determining the predictive accuracy. 
+For each of the cylindrical models we can then compute a PLSL for the circular
+and linear outcome by using the conditional log-likelihoods of the respective
 outcome.
@@
 
-In this section we will analyze the teacher dataset with the help of
-the four cylindrical models introduced in Section \ref{Models}. We will however
-first give a more detailed description of the dataset.
+In this section we analyze the teacher data with the help of the four
+cylindrical models from Section \ref{Models}. First however, we give a more
+detailed description of the dataset.
 
@@
 
-source("Posterior Sampling informative Prior CL-PN.R")
+source("R-code/Posterior Sampling CL-PN.R")
 
@@
 
-source("Posterior Sampling CL-GPN.R")
+source("R-code/Posterior Sampling CL-GPN.R")
 
@@
 
-source("Posterior Sampling CL-GPN_multivariate.R")
+source("R-code/Posterior Sampling Joint GPN-SSN.R")
 
@@
 
-res_CLGPNM <- CLGPNM(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
+res_CLGPNM <- JGPNSSN(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
                       theta.hold, y.hold, X.MGPN.hold)
@@
 
-source("WeisSVM optimization.R")
+source("R-code/Abe-Ley optimization.R")
 
@@
 
-param <- c(1,1,1,1,1,1,0)*0.5
+param <- c(1,1,1,1,1,1,0)*0.9
 
@@
 
-The teacher data was collected between 2010 and 2015 and contains several repeated
-measures on the IPC of 161 teachers gathered for the studies of
-@vanderWant2015role, @Claessens2016side and @pennings2017interpersonal. 
-The measurements were obtained using the QTI and taken in different years and in
-different classes. For this paper we only consider one measurement, namely the
-one for the first measurement occasion (2010) and the one for the largest class
-if data for multiple classes were available. In addition to a variable `IPC`
-containing the circular outcome and the 'length' of the IPC, the linear outcome,
-a teachers' self-efficacy (\verb|SE|) will be used as covariate in the analysis.
-After the removal of missing cases we end up with a sample of 148 teachers Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset.
+The teacher data was collected between 2010 and 2015 and contains several 
+repeated measures on the IPC of 161 teachers. Measurements were obtained using 
+the QTI and taken in different years and classes. For this paper we only 
+consider one measurement, the first occasion (2010) and largest class if data 
+for multiple classes were available. In addition to the score on the IPC, the 
+circular outcome, and the strength of the score on the IPC, the linear outcome, 
+a teachers' self-efficacy (\verb|SE|) concerning classroom management ia used as
+covariate in the analysis. After the removal of missings we have a sample of 148
+teachers of which 10\% (n = 15) is used as a holdout set for the computation of
+the PLSL. Table \ref{Tableteacherdescriptives} shows descriptives
+for the dataset. Figure \ref{dataplot} is a scatterplot showing the relation
+between the linear and circular outcome of the teacher data.
+
+```{r dataplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+
+tikz("Plots/dataplot.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+plot(y.test, theta.test*(180/pi), 
+     ylab = "IPC location", xlab = "IPC strength",
+     bty = "n", ylim = c(-180, 180))
+
+dev.off()
+tools::texi2dvi("Plots/dataplot.tex", pdf=TRUE)
+
+```
 
@@
 #summary statistics
-summary(Dat$Efficacy_SE)
+summary(Dat$Efficacy_CM)
 summary(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
@@
 
-sd(Dat$Efficacy_SE)
+sd(Dat$Efficacy_CM)
 sd(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
@@
 ```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/dataplot.pdf}
+\caption{Plot showing the relation between the linear and circular component of the teacher data.}
+\label{dataplot}
+\end{figure}
 
@@
 Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
-IPC &33.23$^\circ$& 0.76 & - & Circular\\
-length IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
-SE & 4.72 & 0.87 & 2 - 6.5 & Linear\\
+IPC &33.22$^\circ$& 0.76 & - & Circular\\
+strength IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
+SE & 5.04 & 1.00 & 1.5 - 7.0 & Linear\\
    \hline
@@
 
-The variable \verb|SE| is thus used as a predictor for the linear and circular 
-outcomes in all four cylindrical models. We will use the loglikelihoods of the
-following conditional densities \textcolor{red}{CHRISTOPHE, MAG HIER WEL $\sim$ GEBRUIKT WORDEN?}for the computation of the PLSL criterion to
-evaluate the model fit:
+We use the loglikelihoods of the following conditional densities for the
+computation of the PLSL:
 
@@
 
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{pnreg}).
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
 
@@
 
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{gpnreg}).
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
 
@@
 Before analysis there were a couple of settings that we had to specify for the 
-cylindrical models. Starting values for the WeiSSVM model were the following 
-$\eta_0 = 0.5, \eta_1 = 0.5, \nu_0 = 0.5, \nu_1 = 0.5, \kappa = 0.5, \alpha = 
-0.5, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
+cylindrical models. Starting values for the Abe-Ley model were the following 
+$\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 
+0.9, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
 was set to 2000. After we checked convergence via traceplots we concluded that 
-some of the parameters of the multivariate GPN model did not converge. Therefore
+some of the parameters of the joint GPN-SSN model did not converge. Therefore
 we set the amount of iterations of the MCMC models to 20,000 and subtracted a 
@@
 
-rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_0^{I}$",
+rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_1^{I}$",
                               "$\\beta_0^{II}$", "$\\beta_1^{II}$",
@@
                              "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
-                             "$\\beta_{0_s^{II}}$", "$\\beta_{1_y}$",
+                             "$\\beta_{1_s^{II}}$", "$\\beta_{1_y}$",
                              "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
@@
 Tables 2, 3 and 4 show results from the four models that were fit to the teacher
-dataset. There are several parameters that we can compare between these results.
-Firstly, we have the estimated mean of the circular outcome. In Table 5 we see 
-the posterior (predictive) distributions for the estimated circular means of the
-CL-PN, CL-GPN and joint GPN-SSN  model. For the CL-PN model we can actually also
-compute this means from the estimates in Table 2 as follows: $\mu_{circ} = 
-atan2(\beta_0^{II}, \beta_0^{I})$. This leads to an estimate of 31.53$^\circ$. 
-The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are about equal and
-correspond quite well to the actual data mean of 33.23$^\circ$ in Table
-\ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model which is
-0.37 radians or 21.20$^\circ$ is different. This difference could be caused by
-the fact that the densities for the circular ouctome, projected normal or
-sine-skewed von Mises, differ between the models.
+dataset. For the models in which Bayesian estimation was used we show both the 
+estimated posterior mode and the 95\% highest posterior density (HPD) interval 
+for each parameter. Before we turn to an evaluation of the fit of the models, we
+discuss results of and the interpretation of parameters from the four models. We
+will cover parameters concerning the circular outcome, linear outcome and the
+association between them separately.
+
+\subsubsection{Circular component}
+
+Firstly, we can compare the estimated mean of the circular outcome between the
+four models. In Table 5 we see the means of the posterior predictive
+distributions for the circular outcome of the CL-PN, CL-GPN and joint GPN-SSN 
+model. For the CL-PN model we can actually also compute this mean from the
+estimates in Table 2 as follows: $\mu_{circ} = atan2(\beta_0^{II}, \beta_0^{I})
+= 32.62^\circ$. The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are
+about equal and correspond quite well to the actual data mean of 33.22$^\circ$
+in Table \ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model
+which is 0.36 radians or 20.62$^\circ$ is different. This difference could be
+caused by the fact that the densities for the circular ouctome, projected normal
+or sine-skewed von Mises, differ between the models.
 
@@
 
-We can also compare the effect of \verb|SE| on the circular outcome. For the 
-CL-PN model and Abe-Ley models we can get estimates of a circular regression 
-coefficient of this effect. For the Abe-Ley model this estimate is the parameter
-$\nu_1 = -0.01$. This means that for each unit increate in self-efficacy, at the
-inflection point of the circular regression line, the score of the teacher on 
-the IPC decreases with $0.01*(180/\pi) = 0.57^\circ$. For the CL-PN model we 
-can use methods from @CremersMulderKlugkist2017. The estimated posterior mode of
-$b_c$, a parameter that is comparable to $\nu_1$ in the Abe-Ley model, equals 
-$1.13$ and its 95\% HPD interval is $(-26.55; 25.46)$. This means that for each 
-unit increate in self-efficacy, at the inflection point of the circular 
-regression line, the score of the teacher on the IPC increases with 
-$1.13*(180/\pi) = 74.74^\circ$. Values of $b_c$ and $\nu_1$ are quite 
-different even though they should represent the same effect. However, the slope 
-at the inflection point, which $b_c$ and $\nu_1$ describe is not necessarily 
-representative of the effect of \verb|SE| in the data range. In Figure \ref{} we
-see that the inflection point (square) for the CL-PN model lies just ouside of 
-the data range and that for the Abe-Ley model even further (outside the x-axis 
-range). The slope of both regression lines in the data range is much more alike
-(even though one is positive and the other negative). The slope at the mean
-Self-efficicay, which has a value of 0 because it was centered, $SAM$ is
-estimated at $0.02 (-0.09; 0.15)$ The $SAM$ can be interpreted as the amount of
-change on the circle for a unit increase at the mean of the predictor variable.
+We can also compare the effect of self-efficacy (\verb|SE|) on the circular
+outcome. For the CL-PN model and Abe-Ley models we can get estimates of a
+circular regression coefficient of this effect. For the Abe-Ley model this
+estimate is the parameter $\eta_1 = -0.02$. This means that for each unit
+increase in self-efficacy, at the inflection point of the circular regression
+line, the score of the teacher on the IPC decreases with $0.02*(180/\pi) =
+1.15^\circ$. \textcolor{red}{CHRISTOPHE: KLOPT DIT, OF MOET IK $\lambda$ HIER NOG MEENEMEN; NEGEER IK NU HET SINE-SKEWED GEDEELTE} The inflection point is the point at which the regression line
+starts flattening off, i.e. the steepness of the slope decreases. For the CL-PN
+model we can use methods from @CremersMulderKlugkist2017. The estimated
+posterior mode of $b_c$, a parameter that is comparable to $\eta_1$ in the
+Abe-Ley model, equals $1.53$ and its 95\% HPD interval is $(-17.37; 21.26)$.
+This means that for each unit increase in self-efficacy, at the inflection point
+of the circular regression line, the score of the teacher on the IPC increases
+with $1.53*(180/\pi) = 87.66^\circ$. The values of $b_c$ and $\eta_1$ are thus
+quite different even though they should represent the same effect. However, the
+slope at the inflection point that $b_c$ and $\eta_1$ describe is not
+necessarily representative of the effect of \verb|SE| in the data range. In
+Figure \ref{regline} we see that the inflection point (square) of the predicted
+circular regression line for the CL-PN model lies just ouside of the data range
+and that for the Abe-Ley model even further (outside the x-axis range). The
+slope of the regression lines in the data range is much more alike even though
+one is positive and the other negative. The slope at the mean self-efficicay,
+which has a value of 0 because it was centered, $SAM$ is estimated at $0.07
+(-0.03; 0.17)$. The $SAM$ can be interpreted as the amount of change on the
+circle (in radians) for a unit increase at the mean of the predictor variable.
 However, for the CL-PN model the 95\% HPD interval of $SAM$ for the effect of
-\verb|SE| includes 0 meaning that the value is not different from 0. For the
-Abe-Ley model standard errors of the parameters are not known so we cannot
-formally test whether $\nu_1$ differs from 0. Neither is there a known way of
-computing the effect at predictor values other than the one of the inflection
-point.
-
-For the CL-GPN and multivariate 
-CL-GPN models we cannot compute circular regression coefficients. Instead, we 
-will compute posterior predictive distributions for the predicted circular
-outcome of individuals scoring 1 unit above and 1 unit below the average
-self-efficacy.
+\verb|SE| includes 0 meaning that the value is not different from 0. THis
+explains in part how the slopes of the two regression lines are predicted to be
+different. For the Abe-Ley model standard errors of the parameters are not known
+so we cannot formally test whether $\eta_1$ differs from 0. Neither is there a
+known way of computing the effect at predictor values other than the one of the
+inflection point.
+
+For the CL-GPN and GPN-SSN we cannot compute circular regression coefficients. 
+Instead, we will compute posterior predictive distributions for the predicted 
+circular outcome of individuals scoring the minimum, maximum and average 
+self-efficacy. The modes and 95\% HPD intervals of these posterior predictive 
+distributions in the CL-GPN model are $\hat{\theta}_{SE_{min}} = 211.31^\circ 
+(144.41^\circ \: 43.43^\circ)$, $\hat{\theta}_{SE_{mean}} = 25.82^\circ 
+(335.55^\circ \: 145.11^\circ)$ and $\hat{\theta}_{SE_{max}} =30.96^\circ 
+(6.72^\circ \: 74.35^\circ)$ respectively. For the GPN-SSN model the modes and
+95\% HPD intervals of the posterior predictive distributions are
+$\hat{\theta}_{SE_{min}} = 352.06^\circ (121.82^\circ \: 77.75^\circ)$,
+$\hat{\theta}_{SE_{mean}} = 22.74^\circ (336.71^\circ \: 132.25^\circ)$ and
+$\hat{\theta}_{SE_{max}} = 30.87^\circ (358.80^\circ \: 82.13^\circ)$
+respectively. For both models we can thus conclude that as the self-efficacy
+increases, the score of the teacher on the IPC moves counterlockwise. We can
+however not reach any conclusions about the size of this effect or whether it is
+different from zero.
+
+```{r ppreg, echo = FALSE, results = FALSE}
+mode_est_circ(res_CLGPN$theta_pred.min)
+mode_est_circ(res_CLGPN$theta_pred.mean)
+mode_est_circ(res_CLGPN$theta_pred.max)
+
+hpd_est_circ(res_CLGPN$theta_pred.min)
+hpd_est_circ(res_CLGPN$theta_pred.mean)
+hpd_est_circ(res_CLGPN$theta_pred.max)
+
+mode_est_circ(res_CLGPNM$theta_pred.min)
+mode_est_circ(res_CLGPNM$theta_pred.mean)
+mode_est_circ(res_CLGPNM$theta_pred.max)
+
+hpd_est_circ(res_CLGPNM$theta_pred.min)
+hpd_est_circ(res_CLGPNM$theta_pred.mean)
+hpd_est_circ(res_CLGPNM$theta_pred.max)
+
+```
 
@@
 
-tikz("regline.tex", standAlone =TRUE, height = 4, width = 6, pointsize = 12, engine = "pdftex")
+tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
 
@@
 b2 <- mode_est(res_CLPN$BII[5000:20000, 2])
+ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
+ac <- atan2(a2 + b2*ax, a1 + b1*ax)
 
@@
 
-plot(Dat.test$SEc, theta.test, bty = "n",
-     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "Self-efficacy")
-points(x, pred.AL, type = "l", lty = 1)
-points(x, pred.PN, type = "l", lty = 2)
+plot(Dat.test$SEc, theta.test*(180/pi), bty = "n", yaxt = "n",
+     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "self-efficacy")
+axis(side = 2,  at = c(-160, -80, 0, 80, 160))
+points(x, pred.AL*(180/pi), type = "l", lty = 1)
+points(x, pred.PN*(180/pi), type = "l", lty = 2)
+points(ax, ac*(180/pi), pch = 15)
 
 dev.off()
-tools::texi2dvi("regline.tex", pdf=TRUE)
+tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)
 
@@
 \centering
-\includegraphics[width = 0.8\textwidth]{regline.pdf}
-\caption{Plot showing circular regresion lines as predicted by the Abe-Ley (solid line) and CL-PN model (dashed line).}
+\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
+\caption{Plot showing circular regresion lines for the effect of self-efficacy as predicted by the Abe-Ley (solid line) and CL-PN model (dashed line).
+The black square indicated the inflection point of the circular regression line for the CL-PN model.}
 \label{regline}
@@
 
-Next, we compare the intercepts and regression coefficients for the linear outcome.
-
-\begin{itemize}
-\item How to interpret weibull regression for Abe-Ley model?
-\item Comment on different distributions for linear and circular outcome ($\lambda$ in multivariate CL-GPN), integrate with model fit does this correspond. 
-\item Comment on different estimation methods, in CL-PN and CL-GPN the regression of $\theta$ is independent of $y$ but not vice-versa.
-In the Abe-Ley and Multivariate method dependence between $\theta$ and $y$ is modelled both ways.
-\end{itemize}
+\subsubsection{Linear component}
+
+For the linear outcome, we compare the regression coefficients of \verb|SE| for
+the linear outcome, the strength of the interpersonal behavior of a teacher. For
+the CL-PN, CL-GPN and joint GPN-SSN model this are the parameters $\gamma_1$ and
+$\beta_{1_y}$ respectively. Note that in the CL-PN and CL-GPN model the effect 
+of \verb|SE| is the effect controlled for $\cos\theta$ and $\sin\theta$ similar 
+to a 'normal' multiple regression model. For the GPN-SSN model the effect of
+\verb|SE| is also controlled for the circular outcome. However in this model the
+associaton is modelled in a way that is similar to modelling the association
+between several outcome variables in a multivariate regression or ANOVA model.
+Additionally, the distribution we assume for the linear outcome is different.
+For the CL-PN and CL-GPN models we assume a normal distribution while for the
+GPN-SSN model we assume a skewed normal distribution. These differences will
+lead to differences in the estimated effect of \verb|SE| on the linear outcome.
+The interpretation of the regression coefficients is however similar for all
+three models and is like the interpretation of a regression coefficient in any
+other standard linear regression model.
+
+For all three models there is an effect of \verb|SE| on the strength of the 
+interpersonal behavior of a teacher, the HPD interval does not include 0:
+$\gamma_1 = 0.03 \: (0.00; 0.07)$ (CL-PN), $\gamma_1 = 0.03 \: (0.00; 0.06)$
+(CL-GPN) and $\beta_{1_y} = 0.08 \: (0.05; 0.12)$ (GPN-SSN). This means that
+teachers with a higher self-efficacy for classroom management are stronger in
+their interpersonal behavior than teachers with a lower self-efficacy. In
+particular, for each unit increase in self-efficacy the strength of their
+interpersonal behavior is 0.03 or 0.08 (depending on the model) higher. This is
+an increase of 4.17 or 11.11 \% considering the range of the strength on the IPC
+in the data.
+
+For the Abe-Ley model we make use of the fact that the conditional distribution
+for the linear outcome is Weibull. This means that we can use methods from
+survival analsis in medical statistics to interpret the effect of self-efficacy.
+In survival analysis they make use of a 'survival' function in which the time is
+plotted against the probability of survival of subjects suffering from a
+specific medical condition. In our data however we plot the strength on the IPC
+(instead of time) against the probability of a teacher having such a strength.
+This probability can be computed using the so-called 'survival-function' which
+in our case is equal to $\exp(-\alpha y_i^{\beta})$ where $\beta = \exp(\nu_0 +
+\nu_1SE_i)$. In Figure \ref{reglineweib} we plot the survival function for the
+minimum, mean and maximum value of self-efficacy. From this Figure we conclude
+that the stronger interpersonal behaviors are less probable. We also see that
+the self-efficacy positively influences the strength on the IPC, the probability
+of having a stronger interpersonal behavior increases with increasing
+self-efficacy. This result is equivalent to the effect of self-efficacy found
+using the CL-PN, CL-GPN and GPN-SSN models.
+
+```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+summary(Dat.test$SEc)
+
+x <- c(-3.5, 0, 2)
+y <- seq(0, 1, by = 0.05)
+h <- matrix(NA, length(y), length(x))
+s <- matrix(NA, length(y), length(x))
+
+for(j in 1:length(x)){
+
+  for(i in 1:length(y)){
+    beta <- exp(resAL$par[3] + resAL$par[4]*x[j])
+    h[i,j] <- beta*resAL$par[6]*y[i]^(resAL$par[6]-1)
+    s[i,j] <- exp(-resAL$par[6]*y[i]^beta)
+  }
+
+}
+
+tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
+points(y, s[,2], type = "l", lty = 2)
+points(y, s[,3], type = "l", lty = 3)
+legend(0.7, 1, legend = c("min(SE)", "mean(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")
+
+dev.off()
+tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
+\caption{Plot showing the probability of having a particular strength of interersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
+\label{reglineweib}
+\end{figure}
+
+\subsubsection{Association between the linear and circular component}
+
+For all four models we can also investigate the association between the linear 
+and circular component. In the CL-PN and CL-GPN model we do this using the 
+parameters $\gamma_{\cos}$ and $\gamma_{sin}$, the coefficients of the
+regression of the linear outcome onto the sine and cosine of the circular
+outcome. In the joint SSN-GPN model we look at the covariances between the
+linear outcome and the sine and cosine of the circular outcome $\sum_{sy_{2,3}}$
+and $\sum_{sy_{1,3}}$. For the Abe-Ley model we can use the model parameters in
+Table 3 to compute a circular-linear correlation [@abe2017tractable].
+
+In both the CL-PN and CL-GPN model there is an effect of the cosine of the 
+circular outcome onto the linear outcome (HPD interval does not include 0). In 
+the teacher data the sine and cosine component have a substantive meaning. In 
+this case the Communion component of the IPC positively effects the strength of 
+a teachers' type of interpersonal behavior. Thus the teachers showing 
+interpersonal behavior types with higher communion scores (e.g. 'helpful' and 
+'understanding' in Figure \ref{QTI}) are stronger in their behavior. In the 
+joint SSN-GPN model we reach a sligtly different conclusion. In this model both 
+covariances, $\sum_{sy_{2,3}} = 0.09$ and $\sum_{sy_{1,3}} = 0.23$, are 
+different from 0, but the one between the cosine of the circular outcome and the
+linear outcome is larger. This means that both the Communion and Agency 
+component of the IPC positively effect the strength of a teachers' type of 
+interpersonal behavior but the effect of the Communon component is larger. Thus,
+teachers scoring high on Agency but slightly higher on Communion (the 'helpful' 
+category) are strongest in their behavior. Following @abe2017tractable the
+estimated correlation between the circular and linear outcome is equal to
+\textcolor{red}{CHRISTOPHE: KAN JIJ DEZE BEREKENEN, IN R KAN IK GEEN MANIER
+VINDEN OM LEGENDRE FUNCTIES MET NIET INTEGER DEGREE TE EVALUEREN EN ALS IK
+http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
+GEBRUIKT KRIJG IK IMAGINAIRE GETALLEN (voor m = 1 en m = 2), IK WEET NIET OF DAT
+DE BEDOELING IS..}. This correlation is not affected by $\mu$ or $\beta$ meaning
+that the covariate, self-efficacy, can not influence the correlation between the
+type and strength of interpersonal behavior.
+
+```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}
+
+require(pracma)
+
+alpha <- resAL$par[6]
+kappa <- resAL$par[5]
+lambda <- resAL$par[7]
+
+#R cannot compute legendre functions of non-integer degree
+#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
+#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
+#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]
+
+#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
+
+1/3.82
+cosh(1.58)
+
+```
 
 \subsubsection{Model fit}
+
+Instead of interpreting specific model parameters we can also look at the 
+overall fit of the models to the data. In Section \ref{Modelfit} we introduced 
+the criterion, PLSL, that we use to assess model fit in this paper. This
+criterion is focused on the out-of-sample predictive performance. Table 6 shows
+the values of this criterion for the linear and circular outcome of the four
+different models fit to the teacher data.
+
+The CL-PN and CL-GPN model have the best out-of-sample predictive performance 
+for the linear outcome. Note that they have roughly the same performance since 
+they model the linear outcome in the same way apart from the value of $r$ in 
+\eqref{circlinlink} which is a parameter computed in the estimation of the 
+parameters for the circular variable. This similarity in fit is also shown in 
+Figure \ref{preddist} which shows histograms of the test and holdout data
+together with the (posterior) predicitive distributions for all four cylindrical
+models. The posterior predictive distributions for the CL-PN and CL-GPN model
+are almost the same for the linear outcome.
+
+The CL-GPN model has the best out-of-sample predicitve performance for the 
+circular outcome. The Abe-Ley model has the second best performance. This means 
+that for the circular variable a slightly skewed distribution (see Figure 
+\ref{preddist}) fits best. The GPN-SSN model fits worst even though it assumes
+the same distribution for the circular outcome as the CL-GPN model. This could
+be due to the fact that the relation between the circular and linear outcome is 
+modeled differently in these two models. In the CL-GPN model the linear outcome 
+depends on the circular one following the relation in \eqref{circlinlink} but 
+the circular outcome is not dependent on the linear one. In the GPN-SSN model 
+both outcomes are modelled with a joint variance-covariance matrix governing 
+their dependence. In Figure \ref{preddist} we see however that the actual 
+predictive densities do not differ that much between the four models. It is 
+possible that the large difference in PLSL are caused by the fact that they are 
+computed for a relatively small holdout dataset (n = 15) 
+\textcolor{red}{CHRISTOPHE: IS DIT LOGISCH?}.
 
@@
 
-source("Sample Abe-Ley.R")
+source("R-code/Sample Abe-Ley.R")
 
@@
 
-tikz("preddist.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")
+tikz("Plots/preddistdiffSE.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")
 
@@
 
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
+  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu,
+                  kappa = resAL$par[5], lambda = resAL$par[7])
+
   ALdy[i,] <- dat$x
@@
 
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
+  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu,
+                  kappa = resAL$par[5], lambda = resAL$par[7])
+
   ALdy.h[i,]  <- dat$x
@@
 lines(density(res_CLGPN$theta_pred[,19000:20000]), lty = 2)
-lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 3)
-lines(density(ALdth), lty = 4)
+lines(density(ALdth), lty = 3)
+lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 4)
 
@@
 lines(density(res_CLGPN$theta_pred.hold[,19000:20000]), lty = 2)
-lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 3)
-lines(density(ALdth.h), lty = 4)
+lines(density(ALdth.h), lty = 3)
+lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 4)
 
@@
 lines(density(res_CLGPN$y_pred[,5000:20000]), lty = 2)
-lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 3)
-lines(density(ALdy), lty = 4)
+lines(density(ALdy), lty = 3)
+lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 4)
 
@@
 lines(density(res_CLGPN$y_pred.hold[,5000:20000]), lty = 2)
-lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 3)
-lines(density(ALdy.h), lty = 4)
+lines(density(ALdy.h), lty = 3)
+lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 4)
 
@@
 plot(1, type = "n", axes=FALSE, xlab="", ylab="")
-legend(x = "top", inset = 0, legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"), lty = c(1,2,4,3), horiz = TRUE)
+legend(x = "top", inset = 0,
+       legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"),
+       lty = c(1,2,3,4), horiz = TRUE)
 
@@
 
-tools::texi2dvi("preddist.tex",pdf=TRUE)
+tools::texi2dvi("Plots/preddistdiffSE.tex",pdf=TRUE)
 
@@
 \centering
-\includegraphics[width = 0.8\textwidth]{preddist.pdf}
-\caption{Histograms of $\boldsymbol{y}$ and $\boldsymbol{\theta}$ values of the test and holdout set of the teacher data plotted together with (posterior predictive) density estimates for the modified CL-PN, CL-GPN, Abe-Ley and joint GPN-SSN models.}
+\includegraphics[width = 0.8\textwidth]{Plots/preddistdiffSE.pdf}
+\caption{Histograms of linear and circular values of the test and holdout set of the teacher data plotted together with (posterior) predictive density estimates for the modified CL-PN, CL-GPN, Abe-Ley and joint GPN-SSN models based on $n \times 1000$ samples from the (posterior) predictive distribution.}
 \label{preddist}
@@
 
-\begin{itemize}
-\item Comment on difficulty of interpreting parameters in GPN model.
-\end{itemize}
+In this paper we have modified four models for cylindrical data such that they 
+include a regression of both the linear and circular components onto a set of 
+covariates. Subsequently we have used these four methods to analyze a dataset on
+the interpersonal behavior of teachers. Here we will comment on the differences
+between these models, the results from the analysis of the teacher data and how
+cylindrical models improve the analysis of this type of data; data from an
+interpersonal circumplex.
+
+In terms of interpretability, the CL-PN and Abe-Ley model score best. As shown 
+in the previous section it is relatively easy to interpret the parameters of the
+linear and the circular component for these models. In the CL-GPN and joint 
+GPN-SSN model the interpretation of the parameters of the circular component it 
+not easy, if at all possible. This is caused by the fact that in addition to the
+mean vector the covariance matrix of the GPN distribution affects the location 
+of the circular data, making it difficult to compute regression coefficients on 
+the circle. @wang2012directional state that we can use Monte Carlo integration 
+to compute a circular mean and variance for the GPN distribution. This might 
+also be a solution to computing coefficients on the circle and could be applied 
+to the methods of @CremersMulderKlugkist2017 such that they can be used for 
+circular regression coefficients in GPN models as well. \textcolor{red}{HIER OOK
+NOG EEN OPMERKING OVER HOE DE ASSOCIATION TUSSEN CIRCULAR AND LINEAR COMPONENT
+WORDT GEMODELLEERD}.
+
+In terms of flexibility the joint GPN-SSN scores best. In this model multiple 
+linear and circular outcomes can be included and we can thus apply it to 
+multivariate cylindrical data. In addition for both the linear anc circular 
+components this model and the CL-GPN and CL-PN models are extendable to a 
+mixed-effects model and can thus also be fit to longitudinal data. Extension to 
+mixed-effects, longitudinal (see @nunez2014bayesian and @hernandez2016general 
+for hierarchical/mixed-effects models for the PN and GPN distributions 
+respectively). For the Abe-Ley model this may also be possible and has been done
+for the conditional distribution of its linear component, the Weibull 
+distribution, but has not been done in previous literature for the conditional 
+distribution of its circular component (sine-skewed von 
+Mises)\textcolor{red}{CHRISTOPHE: KLOPT DIT?}. In addition, the joint GPN-SSN
+model allows for non-symmetrical shapes of the distributions of both the linear
+(sine-skewed normal) and the circular (general projected normal) components. The
+Abe-Ley model also allows for this and the CL-GPN model allows for a
+non-symmetrical shape of the circular component.
+
+To investigate model fit for the teacher data we asessed the fit to the linear
+and circular component separately using the PLSL, a criterion that is designed
+to assess the out-of-sample preditive performance. The CL-PN and CL-GPN model
+have the best fit for the linear component and the CL-GPN model has the best fit
+for the circular component. Differences in fit, in addition to being a result of
+the different distributions that were used to model the linear and circular
+component of the data, may also be caused by the way in which the relation
+between the linear and circular component is modeled. Whereas in both the
+Abe-Ley and joint GPN-SSN model the distribution of the linear component is
+conditional on the circular component and vice versa, the distribution of the
+circular component in the CL-PN and CL-GPN model is independent of the linear
+component. In these models the linear component are related through a regression
+structure where the circular component serves as a predictor and the linear
+component as the outcome. We also investigated the (posterior) predictive
+distributions of each model and its fit to the training data, the part of the
+teacher data the model was fit on, and the holdout data, the part of the teacher
+data that we used to assess the out-of-sample predictive performance. From these
+plots we saw that whereas the PLSL criteria seem to indicate a substantive
+difference in fit, the (posterior) predicitve distributions seem to be quite
+similar (especially for the circular component). This discrepancy might be
+caused by the fact that the holdout dataset, for which the PLSL criteria were
+computed, is quite small and thus that small deviations in fit could lead to
+large deviations in the criterion value.
+
+The four cylindrical models that were modified to the regression context in this
+paper are not the only cylindrical distributions available from the literature.
+Several other cylindrical distributions, amongst which are @kato2008dependent,
+@sugasawa2015flexible and @fernandez2007models have been introduced (for more
+references we refer to chapter 2 of @ley2017modern). In the present research we
+have decided not to include these models for reasons of space, complexity of the
+models and ease of implementing a regression structure. In further research
+however it would be interesting to investigate other types of cylindrical
+distributions as well.
+
+Even though some of the models have downsides regarding interpretation, e.g. the
+Abe-Ley model does not provide standard errors and parameters from the models 
+with a GPN distribution for the circular component are hard to interpret, 
+cylindrical models in general offer new insights into data of a cylindrical 
+nature in psychology. Concerning the particular example in this paper, the 
+teacher data that were measured using an interpersonal circumplex, we were able 
+to analyze all information in the data simultaneously. Before, the two 
+components from the interpersonal circumplex were analyzed separately. With such
+an approach we get information about the strength of a teachers' score on the 
+two axes of interpersonal behavior, in our case Agency and Communion. This 
+however does not allow us to distinguish between different types of behavior 
+that are specific combinations of Agency and Communion. A solution to that would
+be to treat circumplex data as circular [@Cremers2018Assessing]. This solution 
+however, does not retain the information about the strength of the specific type
+of interpersonal behavior a teacher displays. This is possible using the 
+cylindrical models in the present paper, we can simultaneosly model the 
+information about the type (a circular variable) and strength (a linear 
+variable) of interpersonal behavior that a teacher displays and the association
+between them.
 

### 2018-07-02 Manuscript/ManuscriptMarkdownDifferentSE.tex
Manuscript Update

--- 2018-07-02 Manuscript/ManuscriptMarkdownDifferentSE.Rmd
@@
----
-title: "A Comparison of Regression models for Cylindrical data in Psychology."
-author: |
-  | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
-  | $^1$Department of Methodology and Statistics, Utrecht University
-  | $^2$TNO
-  | $^3$Department of Education, Utrecht University
-  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
-
-bibliography: CircularData.bib
-csl: apa.csl
-output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    number_sections: yes
-  html_document: default
-  word_document: default
-fontsize: 11pt
-geometry: margin=1in
-header-includes:
-  - \usepackage{multirow}
-  - \usepackage{appendix}
-  - \usepackage{color}
-  - \usepackage{hyperref}
-  - \usepackage{subcaption}
-
-keywords: bla vla
-documentclass: article
-pandoc_args: --natbib
-abstract: |
-  Cylindrical dats is multivariate data which consists of a directional, in this paper circular, component   and a linear component.
-Examples of cylindrical data in psychology include human navigation (direction     and distance of movement), eye-tracking research (direction and length of saccades) and data from an       interpersonal circumplex (type and strength of interpersonal behavior).
-In this paper we adapt four        models for cylindrical data to include a regression of the circular and linear component onto a set of     covariates.
-Subsequently, we illustrate how to fit these models and interpret their results on a dataset   on the interpersonal behavior of teachers.
----
-```{r inlude = FALSE, echo = FALSE}
-
-knitr::opts_chunk$set(message = FALSE, warning = FALSE)
-```
-
-```{r, echo = FALSE}
-library(circular)
-library(bpnreg)
-library(haven)
-library(tikzDevice)
-library(plotrix)
-library(tidyr)
-library(dplyr) #sample_n()
-library(MASS) #mvrnorm()
-library(kableExtra) #latex tables
-library(plotrix)
-```
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={A Comparison of Regression models for Cylindrical data in Psychology.},
+            pdfkeywords={bla vla},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{A Comparison of Regression models for Cylindrical data in Psychology.}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Jolien Cremers\footnote{Corresponding author:
+  \href{mailto:j.cremers@uu.nl}{\nolinkurl{j.cremers@uu.nl}}} \(^1\),
+Helena J.M. Pennings\(^{2,3}\), Christophe Ley \(^{4}\)\\
+\(^1\)Department of Methodology and Statistics, Utrecht University\\
+\(^2\)TNO\\
+\(^3\)Department of Education, Utrecht University\\
+\(^4\)Department of Applied Mathematics, Computer Science and
+Statistics, Ghent University}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+    \date{}
+    \predate{}\postdate{}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+
+\usepackage{multirow}
+\usepackage{appendix}
+\usepackage{color}
+\usepackage{hyperref}
+\usepackage{subcaption}
+
+\begin{document}
+\maketitle
+\begin{abstract}
+Cylindrical dats is multivariate data which consists of a directional,
+in this paper circular, component and a linear component. Examples of
+cylindrical data in psychology include human navigation (direction and
+distance of movement), eye-tracking research (direction and length of
+saccades) and data from an interpersonal circumplex (type and strength
+of interpersonal behavior). In this paper we adapt four models for
+cylindrical data to include a regression of the circular and linear
+component onto a set of covariates. Subsequently, we illustrate how to
+fit these models and interpret their results on a dataset on the
+interpersonal behavior of teachers.
+\end{abstract}
 
@@
 
-Cylindrical data are data that consist of a linear variable and a directional 
-variable. In this paper, the directional variable is circular meaning that it 
-consists of a single angle instead of a set of angles. A circular variable is 
-different from a linear variable in the sense that it is measured on a different
-scale. Figure \ref{circline} shows the difference between a circular scale 
-(right) and a linear scale (left). The most important difference is that on a 
-circular scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in 
-fact represent the same number while on a linear scale the two ends, $-\infty$ 
-and $\infty$ are not connected. This difference requires us to use different 
-statistical methods for circular variables (see e.g. @fisher1995statistical for 
-an introduction to circular data and @mardia2000directional,
-@jammalamadaka2001topics and @ley2017modern for a more elaborate overview).
-
-Cylindrical data occur in several fields of research, such as for instance in 
-meteorology [@garcia2013exploring], ecology [@garcia2014test] or marine research
-[@lagona2015hidden]. Several types of data in psychology are also of a 
-cylindrical nature. For example, in research on human navigation in the field of
-cognitive psychology both distance, a linear variable, and direction, a circular
-variable, of movement are of interest [@chrastil2017rotational]. In eye-tracking
-research, which can be used for investigating various cognitive processes (e.g. 
-those involved in reading a text), we can also speak of cylindrical data. When 
-measuring eye-movements we speak of the eye-movements themselves, the saccades, 
-and fixations, the periods of time between movements when the eyes are looking 
-at one point. Of the saccades both the direction, a circular variable, and the 
-duration, a linear variable, are of interest (for a review of eye-tracking 
-research see @rayner200935th). Data from circumplex measurement instruments,
-e.g. the interpersonal circumplex as used in personality psychology, are also of
-a cylindrical nature (see Section \ref{Example} for a more detailed
-explanation).
-
-In this paper we will discuss how a correct statistical treatment of such 
-cylindrical data can lead to new insights. In particular we will show how 
-cylindrical models pave the way for circular-linear and linear-circular 
-regression. We will do this for a motivating example, the teacher data,  from
-the field of educational psychology . In this example, apart from modelling the
-relation between the linear and circular component of a cylindrical variable we
-would also like to predict the two components from a set of covariates in a
-regression model. The teacher data will be further introduced in Sections
+Cylindrical data are data that consist of a linear variable and a
+directional variable. In this paper, the directional variable is
+circular meaning that it consists of a single angle instead of a set of
+angles. A circular variable is different from a linear variable in the
+sense that it is measured on a different scale. Figure \ref{circline}
+shows the difference between a circular scale (right) and a linear scale
+(left). The most important difference is that on a circular scale the
+datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in fact
+represent the same number while on a linear scale the two ends,
+\(-\infty\) and \(\infty\) are not connected. This difference requires
+us to use different statistical methods for circular variables (see e.g.
+Fisher (1995) for an introduction to circular data and Mardia \& Jupp
+(2000), Jammalamadaka \& Sengupta (2001) and Ley \& Verdebout (2017) for
+a more elaborate overview).
+
+Cylindrical data occur in several fields of research, such as for
+instance in meteorology (García-Portugués, Crujeiras, \&
+González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
+González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
+Maruotti, \& Cosoli, 2015). Several types of data in psychology are also
+of a cylindrical nature. For example, in research on human navigation in
+the field of cognitive psychology both distance, a linear variable, and
+direction, a circular variable, of movement are of interest (Chrastil \&
+Warren, 2017). In eye-tracking research, which can be used for
+investigating various cognitive processes (e.g. those involved in
+reading a text), we can also speak of cylindrical data. When measuring
+eye-movements we speak of the eye-movements themselves, the saccades,
+and fixations, the periods of time between movements when the eyes are
+looking at one point. Of the saccades both the direction, a circular
+variable, and the duration, a linear variable, are of interest (for a
+review of eye-tracking research see Rayner (2009)). Data from circumplex
+measurement instruments, e.g.~the interpersonal circumplex as used in
+personality psychology, are also of a cylindrical nature (see Section
+\ref{Example} for a more detailed explanation).
+
+In this paper we will discuss how a correct statistical treatment of
+such cylindrical data can lead to new insights. In particular we will
+show how cylindrical models pave the way for circular-linear and
+linear-circular regression. We will do this for a motivating example,
+the teacher data, from the field of educational psychology . In this
+example, apart from modelling the relation between the linear and
+circular component of a cylindrical variable we would also like to
+predict the two components from a set of covariates in a regression
+model. The teacher data will be further introduced in Sections
 \ref{Example} and \ref{DataAnalysis}.
@@
 
-As is the case for circular data, the analysis of cylindrical data requires 
-special methods. Several methods have been put forward to model the relation 
-between the linear and circular component of a cylindrical variable. Some of 
-these are based on regressing the linear component onto the circular component 
-using the following type of relation: $$y = \beta_0 + \beta_1*\cos(\theta) + 
-\beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear component and $\theta$
-the circular component [@mardia1978model;
-@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a 
-different way, e.g. by specifying a multivariate model for several linear and 
-circular variables and modelling their covariance matrix 
-[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For 
-example, @abe2017tractable introduce a cylindrical distribution based on a 
-Weibull distribution for the linear component and a sine-skewed von Mises 
-distribution for the circular component and link these through their shape and 
-concentration parameters respectively. However, none of the methods that have 
-been proposed thus far include additional covariates onto which both the
+As is the case for circular data, the analysis of cylindrical data
+requires special methods. Several methods have been put forward to model
+the relation between the linear and circular component of a cylindrical
+variable. Some of these are based on regressing the linear component
+onto the circular component using the following type of relation:
+\[y = \beta_0 + \beta_1*\cos(\theta) + 
+\beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the linear component
+and \(\theta\) the circular component (Johnson \& Wehrly, 1978; Mardia
+\& Sutton, 1978; Mastrantonio, Maruotti, \& Jona-Lasinio, 2015). Others
+model the relation in a different way, e.g.~by specifying a multivariate
+model for several linear and circular variables and modelling their
+covariance matrix (Mastrantonio, 2018) or by proposing a joint
+cylindrical distribution. For example, Abe \& Ley (2017) introduce a
+cylindrical distribution based on a Weibull distribution for the linear
+component and a sine-skewed von Mises distribution for the circular
+component and link these through their shape and concentration
+parameters respectively. However, none of the methods that have been
+proposed thus far include additional covariates onto which both the
 circular and linear component are regressed.
 
-Our aim in this paper is to fill this gap in the literature by adapting four
-existing cylindrical methods in such a way that they include a regression of
-both the linear and circular component of a cylindrical variable onto a set of
-covariates. First however, we will introduce the teacher data in Section
-\ref{Example}. In Section \ref{Models} we introduce the four modified models for
-cylindrical data that we use to analyze the data from the motivating example. We
-also choose a model selection criterion to compare the models. The teacher data
-will be analysed in Section \ref{DataAnalysis}. The paper will be concluded with
-a discussion in Section \ref{Discussion}. The Appendix contains technical
-details.
+Our aim in this paper is to fill this gap in the literature by adapting
+four existing cylindrical methods in such a way that they include a
+regression of both the linear and circular component of a cylindrical
+variable onto a set of covariates. First however, we will introduce the
+teacher data in Section \ref{Example}. In Section \ref{Models} we
+introduce the four modified models for cylindrical data that we use to
+analyze the data from the motivating example. We also choose a model
+selection criterion to compare the models. The teacher data will be
+analysed in Section \ref{DataAnalysis}. The paper will be concluded with
+a discussion in Section \ref{Discussion}. The Appendix contains
+technical details.
 
@@
 
-The motivating example for this article comes from the field of educational 
-psychology and was collected for the studies on classroom climate of
-@pennings2017interpersonal, @Claessens2016side and @vanderWant2015role. An 
-indicator of classroom climate is the students' perceptions of their teachers'
-interpersonal behavior. These interpersonal perceptions, both in educational
-psychology as well as in other areas of psychology, can be measured using
-circumplex measurement instruments (see
-@horowitz2010handbook for an overview of many such instruments).
-
-The Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] is
-one such circumplex measurement instrument. It is used to study student
-perceptions of their teachers' interpersonal behavior and contains
-items that load on two dimensions: Agency and Communion. Agency refers to the
-degree of power or control a teacher exerts in interaction with his/her
-students. Communion refers to the degree of friendliness or affiliation a
-teacher conveys in interaction with his/her students. The loadings on the two 
-dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
-this space are characterized by different teacher behavior, e.g. 'helpful' or
-'uncertain'. We call the two-dimensional space the interpersonal circle (IPC).
-The idea is that the IPC is ``a continuous order with no beginning or end''
-[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
-the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular variable.
+The motivating example for this article comes from the field of
+educational psychology and was collected for the studies on classroom
+climate of Pennings et al. (2017), Claessens (2016) and Want (2015). An
+indicator of classroom climate is the students' perceptions of their
+teachers' interpersonal behavior. These interpersonal perceptions, both
+in educational psychology as well as in other areas of psychology, can
+be measured using circumplex measurement instruments (see Horowitz \&
+Strack (2011) for an overview of many such instruments).
+
+The Questionnaire on Teacher Interaction (QTI) (Wubbels, Brekelmans,
+Brok, \& Tartwijk, 2006) is one such circumplex measurement instrument.
+It is used to study student perceptions of their teachers' interpersonal
+behavior and contains items that load on two dimensions: Agency and
+Communion. Agency refers to the degree of power or control a teacher
+exerts in interaction with his/her students. Communion refers to the
+degree of friendliness or affiliation a teacher conveys in interaction
+with his/her students. The loadings on the two dimensions of the QTI can
+be placed in a two-dimensional space formed by Agency (vertical) and
+Communion (horizontal), see Figure \ref{QTI}. Different parts of this
+space are characterized by different teacher behavior, e.g. `helpful' or
+`uncertain'. We call the two-dimensional space the interpersonal circle
+(IPC). The idea is that the IPC is ``a continuous order with no
+beginning or end'' (Gurtman, 2009, p. 2). We call such ordering a
+circumplex ordering and the IPC is therefore often called the
+interpersonal circumplex. The ordering also implies that scores on the
+IPC could be viewed as a circular variable.
 
@@
 
-@Cremers2018Assessing show how data from the IPC can be considered circular data
-and analyzed as such using a regression model. The two-dimension scores Agency
-and Communion can be converted to a circular score using the two-argument
-arctangent function in \eqref{PredVal}, where $A$ represents a score on
-the Agency dimension and $C$ represents a score on the Communion dimension. The
-resulting circular variable $\theta$ can then be modeled. However, when
-two-dimensional data are converted to the circle we lose some information, the
-length of the two-dimensional vector $(A, C)$, \emph{i.e.}, its euclidean norm
-$\mid\mid (A, C) \mid\mid$. This length represents the strength of the type of
-interpersonal behavior a teacher shows towards his/her students. In a
-cylindrical model we are able to incorporate this information, and model a
-circular variable $\theta$ together with a linear variable corresponding to
-$\mid\mid (A, C) \mid\mid$. This leads to an improved analysis of data from the
-IPC. In the next section we introduce several models that can be used for more
-accurate and informative regression analysis on the teacher data. Descriptives
-for the teacher data are given in Section \ref{DataAnalysis}.
+Cremers, Mainhard, \& Klugkist (2018a) show how data from the IPC can be
+considered circular data and analyzed as such using a regression model.
+The two-dimension scores Agency and Communion can be converted to a
+circular score using the two-argument arctangent function in
+\eqref{PredVal}, where \(A\) represents a score on the Agency dimension
+and \(C\) represents a score on the Communion dimension. The resulting
+circular variable \(\theta\) can then be modeled. However, when
+two-dimensional data are converted to the circle we lose some
+information, the length of the two-dimensional vector \((A, C)\),
+\emph{i.e.}, its euclidean norm \(\mid\mid (A, C) \mid\mid\). This
+length represents the strength of the type of interpersonal behavior a
+teacher shows towards his/her students. In a cylindrical model we are
+able to incorporate this information, and model a circular variable
+\(\theta\) together with a linear variable corresponding to
+\(\mid\mid (A, C) \mid\mid\). This leads to an improved analysis of data
+from the IPC. In the next section we introduce several models that can
+be used for more accurate and informative regression analysis on the
+teacher data. Descriptives for the teacher data are given in Section
+\ref{DataAnalysis}.
 
@@
 
-```{r loaddata, return = FALSE, echo = FALSE}
-Dat <- read_spss("Data Heleen/mergeddata.sav")
-
-#only select the first measurement occasion
-Dat <- subset(Dat, Time == 0)
-
-#sort on class size
-Dat <- Dat[order(-Dat$nklas), ]
-
-#Now remove duplicates 
-#(because dataframe is sorted on class size the smallest classes are removed)
-
-duplicates <- duplicated(Dat[,c("docnr", "Time")])
-Dat <- Dat[!duplicates,]
-
-#center and remove missings for self-efficacy
-Dat <- Dat %>% drop_na(Efficacy_CM)
-Dat$SEc <- Dat$Efficacy_CM- mean(Dat$Efficacy_CM)
-
-set.seed(50)
-
-#Create holdout and test sample
-Dat.hold <- sample_n(Dat, 15)
-Dat.test <- Dat[-as.numeric(rownames(Dat.hold)),]
-
-#Compute circular and linear outcome
-theta.test <- atan2(Dat.test$lAgency, Dat.test$lcommunion)
-y.test     <- sqrt(Dat.test$lAgency^2 + Dat.test$lcommunion^2)
-
-theta.hold <- atan2(Dat.hold$lAgency, Dat.hold$lcommunion)
-y.hold     <- sqrt(Dat.hold$lAgency^2 + Dat.hold$lcommunion^2)
-
-```
-
 \section{Four cylindrical regression models}\label{Models}
 
-In this section we introduce four cylindrical regression models to contain 
-predictors for the linear and circular outcomes, $\Theta$ and $Y$. We extend the models from @mastrantonio2018joint and
-@abe2017tractable. Additionally, we introduce two models where the relation
-between $\Theta$ and $Y$ is modelled as follows (following
-@mastrantonio2015bayesian):
+In this section we introduce four cylindrical regression models to
+contain predictors for the linear and circular outcomes, \(\Theta\) and
+\(Y\). We extend the models from Mastrantonio (2018) and Abe \& Ley
+(2017). Additionally, we introduce two models where the relation between
+\(\Theta\) and \(Y\) is modelled as follows (following Mastrantonio et
+al. (2015)):
 
@@
 
-In both of these models the relation between $\Theta \in [0, 2\pi)$ and $Y\in 
-(-\infty, + \infty)$ is specified as in \eqref{circlinlink} with the following
-conditional distribution:
+In both of these models the relation between \(\Theta \in [0, 2\pi)\)
+and \(Y\in (-\infty, + \infty)\) is specified as in \eqref{circlinlink}
+with the following conditional distribution:
 
@@
 
-In a regression setup the outcome $\theta_i$ for each individual $i = 1, \dots,
-n$, where $n$ is the sample size, is generated independently from \eqref{pnreg}.
-The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is defined as
-$\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
-$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and $\boldsymbol{B} = 
-(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ contain the regression 
-coefficients. Note however that the dimensions of $\boldsymbol{\beta}^{I}$ and 
-$\boldsymbol{\beta }^{I}$ need not necessarily be the same and we are thus 
-allowed to have a different set of predictor variables and vectors 
-$\boldsymbol{z}_i^I$ and $\boldsymbol{z}_i^{II}$ for the two components of 
-$\boldsymbol{\mu}_i$.
-
-Following @wang2012directional and @hernandez2016general the joint density of
-$r$ and $\theta$ for the GPN distribution equals:
+In a regression setup the outcome \(\theta_i\) for each individual
+\(i = 1, \dots, n\), where \(n\) is the sample size, is generated
+independently from \eqref{pnreg}. The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}\). The vector
+\(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values and
+\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
+contain the regression coefficients. Note however that the dimensions of
+\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{I}\) need not
+necessarily be the same and we are thus allowed to have a different set
+of predictor variables and vectors \(\boldsymbol{z}_i^I\) and
+\(\boldsymbol{z}_i^{II}\) for the two components of
+\(\boldsymbol{\mu}_i\).
+
+Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
+(2016) the joint density of \(r\) and \(\theta\) for the GPN
+distribution equals:
 
@@
 
-where $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2  & \rho\\ \rho & 1 
-\end{bmatrix}$, $\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta 
-\end{bmatrix}$. In a regression setup the outcome $\theta_i$ for each individual
-is generated independently from \eqref{gpnreg}. The mean vector 
-$\boldsymbol{\mu}_i \in \mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
-\boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. The vector 
-$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and each 
-$\boldsymbol{\beta}^{k}$ is a vector with two regression coefficients, one for 
-each of the two components of $\boldsymbol{\mu}_i$. Note that for the CL-GPN 
-model we do need to have the same predictors for both components of 
-$\boldsymbol{\mu}_i$.
-
-Both cylindrical models introduced here are estimated using MCMC methods based
-on @nunez2011bayesian, @wang2012directional and @hernandez2016general for the 
-regression of the circular outcome. A detailed description of the Bayesian
-estimation and MCMC samplers can be found in Appendices \ref{A1} and \ref{A2}.
+where
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
+\(\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta \end{bmatrix}\).
+In a regression setup the outcome \(\theta_i\) for each individual is
+generated independently from \eqref{gpnreg}. The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\).
+The vector \(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values
+and each \(\boldsymbol{\beta}^{k}\) is a vector with two regression
+coefficients, one for each of the two components of
+\(\boldsymbol{\mu}_i\). Note that for the CL-GPN model we do need to
+have the same predictors for both components of \(\boldsymbol{\mu}_i\).
+
+Both cylindrical models introduced here are estimated using MCMC methods
+based on Nuñez-Antonio et al. (2011), Wang \& Gelfand (2013) and
+Hernandez-Stumpfhauser et al. (2016) for the regression of the circular
+outcome. A detailed description of the Bayesian estimation and MCMC
+samplers can be found in Appendices \ref{A1} and \ref{A2}.
 
@@
 
-This model is an extension of the cylindrical model introduced by
-@abe2017tractable to the regression context. The joint density of $\Theta$ and
-$Y$, in this model defined on the positive real half-line $[0, + \infty)$, is:
+This model is an extension of the cylindrical model introduced by Abe \&
+Ley (2017) to the regression context. The joint density of \(\Theta\)
+and \(Y\), in this model defined on the positive real half-line
+\([0, + \infty)\), is:
 
@@
 
-In a regression setup the outcome vector $(\theta_i, y_i)^t$ for each individual
-is generated independently from \eqref{WeiSSVMdensity}. The parameters $\alpha >
-0$ and $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ are linear shape
-and scale parameters, $\mu_i = \eta_0 + 
-2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$, $\kappa > 0$ and
-$\lambda \in [-1, 1]$ are circular location, concentration and skewness 
-parameters. The parameter $\boldsymbol{\nu}$ is a vector of $q$ regression 
-coefficients $\nu_j \in (-\infty, +\infty)$ for the prediction of $y$ where $j =
-0, \dots, q$ and $\nu_0$ is the intercept. The parameter $\eta_0 \in [0, 2\pi)$
-is the intercept and $\boldsymbol{\eta}$ is a vector of $p$ regression
-coefficients $\eta_j \in (-\infty, +\infty)$ for the prediction of $\theta$
-where $j = 1, \dots, p$. The vector $\boldsymbol{x}_i$ is a vector of predictor
-values for the prediction of $y$ and $\boldsymbol{z}_i$ is a vector of predictor
-values for the prediction of $\theta$.
-
-As in @abe2017tractable, the conditional distribution of $y$ given $\theta$ is a
-Weibull distribution with scale $\alpha$ and shape
-$\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}$ and the conditional 
-distribution of $\theta$ given $y$ is a sine skewed von Mises distribution with 
-location parameter $\mu$ and concentration parameter $(\beta 
-y)^\alpha\tanh(\kappa)$.
+In a regression setup the outcome vector \((\theta_i, y_i)^t\) for each
+individual is generated independently from \eqref{WeiSSVMdensity}. The
+parameters \(\alpha > 0\) and
+\(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) are linear
+shape and scale parameters,
+\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)\),
+\(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular location,
+concentration and skewness parameters. The parameter
+\(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
+\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where
+\(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The parameter
+\(\eta_0 \in [0, 2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a
+vector of \(p\) regression coefficients
+\(\eta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\) where
+\(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector of
+predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\) is
+a vector of predictor values for the prediction of \(\theta\).
+
+As in Abe \& Ley (2017), the conditional distribution of \(y\) given
+\(\theta\) is a Weibull distribution with scale \(\alpha\) and shape
+\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}\) and the
+conditional distribution of \(\theta\) given \(y\) is a sine skewed von
+Mises distribution with location parameter \(\mu\) and concentration
+parameter \((\beta y)^\alpha\tanh(\kappa)\).
 
@@
 This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. It contains $p$ circular
-outcomes and $q$ linear outcomes. The circular outcomes $\boldsymbol{\Theta} = 
-(\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ are modeled together by a
-multivariate GPN distribution and the linear outcomes $\boldsymbol{Y} = 
-(\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled together by a 
-multivariate skew normal distribution [@sahu2003new]. Because the GPN 
-distribution is modelled using a so-called augmented representation (as in 
-\eqref{projection} and \eqref{gpnreg}) it is convenient to use a similar tactic
-for modelling the multivariate skew normal distribution. Following
-@mastrantonio2018joint the linear outcomes are represented as:
-
-$$\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},$$
-where where $\boldsymbol{\mu}_y$ is a mean vector for the linear outcome 
-$\boldsymbol{Y}$, $\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})$ is a $q 
-\times q$ diagonal matrix with diagonal elements $\boldsymbol{\lambda} = 
-\lambda_1, \dots, \lambda_q$ (skewness parameters), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
-\boldsymbol{I}_q)$, a q-dimensional half normal distribution
-[@olmos2012extension], and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
-\boldsymbol{\Sigma}_y)$. This means that conditional on the auxiliary data
-$\boldsymbol{D}$, $\boldsymbol{Y}$ is normally distributed with mean
-$\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}$ and covariance matrix
-$\boldsymbol{\Sigma}_y$. The joint density for $(\boldsymbol{Y},
-\boldsymbol{D})^t$ is defined as follows:
+Mastrantonio (2018) to the regression context. It contains \(p\)
+circular outcomes and \(q\) linear outcomes. The circular outcomes
+\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)\)
+are modeled together by a multivariate GPN distribution and the linear
+outcomes
+\(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)\) are
+modeled together by a multivariate skew normal distribution (Sahu, Dey,
+\& Branco, 2003). Because the GPN distribution is modelled using a
+so-called augmented representation (as in \eqref{projection} and
+\eqref{gpnreg}) it is convenient to use a similar tactic for modelling
+the multivariate skew normal distribution. Following Mastrantonio (2018)
+the linear outcomes are represented as:
+
+\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
+where where \(\boldsymbol{\mu}_y\) is a mean vector for the linear
+outcome \(\boldsymbol{Y}\),
+\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
+\(q \times q\) diagonal matrix with diagonal elements
+\(\boldsymbol{\lambda} = \lambda_1, \dots, \lambda_q\) (skewness
+parameters),
+\(\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, \boldsymbol{I}_q)\), a
+q-dimensional half normal distribution (Olmos, Varela, Gómez, \&
+Bolfarine, 2012), and
+\(\boldsymbol{H} \sim N_q(\boldsymbol{0}_q, \boldsymbol{\Sigma}_y)\).
+This means that conditional on the auxiliary data \(\boldsymbol{D}\),
+\(\boldsymbol{Y}\) is normally distributed with mean
+\(\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and
+covariance matrix \(\boldsymbol{\Sigma}_y\). The joint density for
+\((\boldsymbol{Y}, \boldsymbol{D})^t\) is defined as follows:
 
@@
 
-where the mean vector $\boldsymbol{\mu} = (\boldsymbol{\mu}_s,
-\boldsymbol{\mu}_y)^t$ and $\boldsymbol{\Sigma} = \left ( \begin{matrix}
-\boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} &
-\boldsymbol{\Sigma}_y \\ \end{matrix} \right )$. The matrix
-$\boldsymbol{\Sigma}_s$ is the covariance matrix for the variances of and
-covariances between the augmented representations of the circular outcome and
-the matrix $\boldsymbol{\Sigma}_{sy}$ contains covariances between the augmented
-representations of the circular outcome and the linear outcome.
-
-In our regression extension we have $i = 1, \dots, n$ observations of $p$ 
-circular outcomes, $q$ linear outcomes and $k$ covariates. The mean in the 
-density in \eqref{YDThetarjoint} then becomes $\boldsymbol{\mu}_i = 
-\boldsymbol{x}_i\boldsymbol{B}$ where $\boldsymbol{B}$ is a $k \times (2 
-\times p + q)$ matrix with regression coeffients. We estimate the model using 
-MCMC methods. A detailed description of these methods is given in Appendix
-\ref{A3}.
+where the mean vector
+\(\boldsymbol{\mu} = (\boldsymbol{\mu}_s, \boldsymbol{\mu}_y)^t\) and
+\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
+The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
+variances of and covariances between the augmented representations of
+the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
+contains covariances between the augmented representations of the
+circular outcome and the linear outcome.
+
+In our regression extension we have \(i = 1, \dots, n\) observations of
+\(p\) circular outcomes, \(q\) linear outcomes and \(k\) covariates. The
+mean in the density in \eqref{YDThetarjoint} then becomes
+\(\boldsymbol{\mu}_i = \boldsymbol{x}_i\boldsymbol{B}\) where
+\(\boldsymbol{B}\) is a \(k \times (2 \times p + q)\) matrix with
+regression coeffients. We estimate the model using MCMC methods. A
+detailed description of these methods is given in Appendix \ref{A3}.
 
@@
 
-where $B$ is the amount of posterior samples and $\boldsymbol{\vartheta}^{(j)}$ 
-are the posterior estimates of the model parameters for the $j^{th}$ iteration. 
-Note that although we fit the CL-PN, CL-GPN and joint GPN-SSN models using 
-Bayesian statistics, we do not take prior information into account when 
-assessing model fit. According to @BDA \textcolor{red}{(p. ???)} this is not 
-necessary since we are assessing the fit of a model to data, the holdout set,
-only. They argue that the prior in such case is only of interest for estimating
-the parameters of the model but not for determining the predictive accuracy. 
-For each of the cylindrical models we can then compute a PLSL for the circular
-and linear outcome by using the conditional log-likelihoods of the respective
-outcome.
+where \(B\) is the amount of posterior samples and
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
+model parameters for the \(j^{th}\) iteration. Note that although we fit
+the CL-PN, CL-GPN and joint GPN-SSN models using Bayesian statistics, we
+do not take prior information into account when assessing model fit.
+According to Gelman et al. (2014) \textcolor{red}{(p. ???)} this is not
+necessary since we are assessing the fit of a model to data, the holdout
+set, only. They argue that the prior in such case is only of interest
+for estimating the parameters of the model but not for determining the
+predictive accuracy. For each of the cylindrical models we can then
+compute a PLSL for the circular and linear outcome by using the
+conditional log-likelihoods of the respective outcome.
 
@@
 In this section we analyze the teacher data with the help of the four
-cylindrical models from Section \ref{Models}. First however, we give a more
-detailed description of the dataset.
-
-```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling CL-PN.R")
-
-its <- 20000
-set.seed(101)
-
-ZI.PN.test  <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-ZII.PN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.PN.test   <- as.matrix(cbind(rep(1, length(theta.test)),
-                               cos(theta.test),
-                               sin(theta.test),
-                               Dat.test$SEc))
-
-ZI.PN.hold  <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-ZII.PN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.PN.hold   <- as.matrix(cbind(rep(1, length(theta.hold)),
-                               cos(theta.hold),
-                               sin(theta.hold),
-                               Dat.hold$SEc))
-
-res_CLPN <- CLPN(theta.test, y.test, X.PN.test, ZI.PN.test, ZII.PN.test, its,
-                 theta.hold, y.hold, X.PN.hold, ZI.PN.hold, ZII.PN.hold)
-```
-
-```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling CL-GPN.R")
-
-its <- 20000
-set.seed(101)
-
-Z.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)),
-                              cos(theta.test),
-                              sin(theta.test),
-                              Dat.test$SEc))
-
-Z.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)),
-                              cos(theta.hold),
-                              sin(theta.hold),
-                              Dat.hold$SEc))
-
-res_CLGPN <- CLGPN(theta.test, y.test, X.GPN.test, Z.GPN.test, its, p = 2, 
-                   theta.hold, y.hold, X.GPN.hold, Z.GPN.hold)
-
-```
-
-```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling Joint GPN-SSN.R")
-
-its <- 20000
-set.seed(101)
-
-X.MGPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-
-X.MGPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-
-res_CLGPNM <- JGPNSSN(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
-                      theta.hold, y.hold, X.MGPN.hold)
-
-```
-
-```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Abe-Ley optimization.R")
-
-set.seed(101)
-
-Z.test     <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-X.test    <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
-
-Z.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-X.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
-
-colnames(X.test) <- c("ax", "bx")
-colnames(Z.test) <- c("az", "bz")
-colnames(X.hold) <- c("ax", "bx")
-colnames(Z.hold) <- c("az", "bz")
-
-dat.t <- as.data.frame(cbind(theta.test, y.test, X.test, Z.test))
-dat.h <- as.data.frame(cbind(theta.hold, y.hold, X.hold, Z.hold))
-
-ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
-               c(-1, 0, 0, 0, 0, 0, 0),
-               c(0, 0, 0, 0, 1, 0, 0),
-               c(0, 0, 0, 0, 0, 1, 0),
-               c(0, 0, 0, 0, 0, 0, 1),
-               c(0, 0, 0, 0, 0, 0, -1))
-
-ci.reg = c(-pi, -pi, 0, 0, -1, -1)
-
-param <- c(1,1,1,1,1,1,0)*0.9
-
-ui.reg %*% param - ci.reg
-
-resAL <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
-                     control = list(maxit = 1000000, fnscale = -1), data = dat.t)
-
-```
+cylindrical models from Section \ref{Models}. First however, we give a
+more detailed description of the dataset.
 
@@
 
-The teacher data was collected between 2010 and 2015 and contains several 
-repeated measures on the IPC of 161 teachers. Measurements were obtained using 
-the QTI and taken in different years and classes. For this paper we only 
-consider one measurement, the first occasion (2010) and largest class if data 
-for multiple classes were available. In addition to the score on the IPC, the 
-circular outcome, and the strength of the score on the IPC, the linear outcome, 
-a teachers' self-efficacy (\verb|SE|) concerning classroom management ia used as
-covariate in the analysis. After the removal of missings we have a sample of 148
-teachers of which 10\% (n = 15) is used as a holdout set for the computation of
-the PLSL. Table \ref{Tableteacherdescriptives} shows descriptives
-for the dataset. Figure \ref{dataplot} is a scatterplot showing the relation
+The teacher data was collected between 2010 and 2015 and contains
+several repeated measures on the IPC of 161 teachers. Measurements were
+obtained using the QTI and taken in different years and classes. For
+this paper we only consider one measurement, the first occasion (2010)
+and largest class if data for multiple classes were available. In
+addition to the score on the IPC, the circular outcome, and the strength
+of the score on the IPC, the linear outcome, a teachers' self-efficacy
+(\verb|SE|) concerning classroom management ia used as covariate in the
+analysis. After the removal of missings we have a sample of 148 teachers
+of which 10\% (n = 15) is used as a holdout set for the computation of
+the PLSL. Table \ref{Tableteacherdescriptives} shows descriptives for
+the dataset. Figure \ref{dataplot} is a scatterplot showing the relation
 between the linear and circular outcome of the teacher data.
-
-```{r dataplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/dataplot.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-plot(y.test, theta.test*(180/pi), 
-     ylab = "IPC location", xlab = "IPC strength",
-     bty = "n", ylim = c(-180, 180))
-
-dev.off()
-tools::texi2dvi("Plots/dataplot.tex", pdf=TRUE)
-
-```
-
-```{r descripitves, echo = FALSE, results = FALSE}
-#sample size
-nrow(Dat)
-
-#summary statistics
-summary(Dat$Efficacy_CM)
-summary(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
-summary(as.circular(atan2(Dat$lAgency, Dat$lcommunion)%%(2*pi)))
-
-sd(Dat$Efficacy_CM)
-sd(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
-
-```
 
@@
 
-Before analysis there were a couple of settings that we had to specify for the 
-cylindrical models. Starting values for the Abe-Ley model were the following 
-$\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 
-0.9, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
-was set to 2000. After we checked convergence via traceplots we concluded that 
-some of the parameters of the joint GPN-SSN model did not converge. Therefore
-we set the amount of iterations of the MCMC models to 20,000 and subtracted a 
-burn-in of 5000. With this specification the MCMC chains for all parameters 
-converge. Note that we choose the same amount of iterations for all three 
-Bayesian cylindrical models to make their comparison via the PLSL as fair as 
-possible. Lastly, the predictor \verb|SE| was centered before inclusion in the
-analysis.
-
-```{r estCLPN, echo = FALSE}
-
-CLPN.Gamma.m   <- apply(res_CLPN$Gamma[5000:20000,], 2, mode_est)
-CLPN.Gamma.hpd <- apply(res_CLPN$Gamma[5000:20000,], 2, hpd_est)
-
-CLPN.BI.m   <- apply(res_CLPN$BI[5000:20000,], 2, mode_est)
-CLPN.BI.hpd <- apply(res_CLPN$BI[5000:20000,], 2, hpd_est)
-
-CLPN.BII.m <- apply(res_CLPN$BII[5000:20000,], 2, mode_est)
-CLPN.BII.hpd <- apply(res_CLPN$BII[5000:20000,], 2, hpd_est)
-
-CLPN.Sigma.m   <- mode_est(res_CLPN$Sigma[5000:20000])
-CLPN.Sigma.hpd <- hpd_est(res_CLPN$Sigma[5000:20000])
-
-modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
-hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
-hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)
-
-CLPNres.tab <- cbind(modes.PN, hpd.LB.PN, hpd.UB.PN)
-
-```
-
-```{r estCLGPN, echo = FALSE}
-
-CLGPN.Sig.m   <- mode_est(res_CLGPN$Sig)
-CLGPN.Sig.hpd <- hpd_est(res_CLGPN$Sig)
-
-CLGPN.Gamma.m   <- apply(res_CLGPN$Gamma[5000:20000,], 2, mode_est)
-CLGPN.Gamma.hpd <- apply(res_CLGPN$Gamma[5000:20000,], 2, hpd_est)
-
-CLGPN.BI.m   <- apply(res_CLGPN$BI[5000:20000,], 2, mode_est)
-CLGPN.BI.hpd <- apply(res_CLGPN$BI[5000:20000,], 2, hpd_est)
-
-CLGPN.BII.m   <- apply(res_CLGPN$BII[5000:20000,], 2, mode_est)
-CLGPN.BII.hpd <- apply(res_CLGPN$BII[5000:20000,], 2, hpd_est)
-
-CLGPN.Sigma11.m   <- mode_est(res_CLGPN$Sigma[1,1,5000:20000])
-CLGPN.Sigma11.hpd <- hpd_est(res_CLGPN$Sigma[1,1,5000:20000])
-CLGPN.Sigma22.m   <- mode_est(res_CLGPN$Sigma[2,2,5000:20000])
-CLGPN.Sigma22.hpd <- hpd_est(res_CLGPN$Sigma[2,2,5000:20000])
-CLGPN.Sigma12.m   <- mode_est(res_CLGPN$Sigma[1,2,5000:20000])
-CLGPN.Sigma12.hpd <- hpd_est(res_CLGPN$Sigma[1,2,5000:20000])
-
-modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
-                CLGPN.Gamma.m, CLGPN.Sig.m,
-                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
-hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
-               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
-               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
-hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
-               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
-               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])
-
-CLGPNres.tab <- cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN)
-CLPNGPNres.tab <- round(cbind(CLPNres.tab, CLGPNres.tab), 2)
-
-rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_1^{I}$",
-                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
-                              "$\\gamma_0$", "$\\gamma_{cos}$",
-                              "$\\gamma_{sin}$", "$\\gamma_1$",
-                              "$\\sigma$", "$\\sum_{1,1}$",
-                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
-colnames(CLPNGPNres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")
-
-kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE,
-      caption = "Results for the modified CL-PN and CL-GPN model")%>%
-add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))
-
-```
-
-```{r estSL, echo = FALSE}
-
-ALres.tab <- matrix(round(c(resAL$par[1:4], resAL$par[6], resAL$par[5], resAL$par[7]), 2),7,1)
-rownames(ALres.tab) <- c("$\\eta_0$", "$\\eta_1$",
-                         "$\\nu_0$", "$\\nu_1$",
-                         "$\\alpha$", "$\\kappa$", "$\\lambda$")
-
-kable(ALres.tab, "latex", booktabs = T, escape = FALSE,
-      caption = "Results for the modified Abe-Ley model")%>%
-add_header_above(c("Parameter" = 1, "ML-estimate" = 1))
-
-```
-
-```{r estCLMGPN, echo = FALSE}
-
-CLGPNM.lambda.m   <- mode_est(res_CLGPNM$lambda[5000:20000, ])
-CLGPNM.lambda.hpd <- hpd_est(res_CLGPNM$lambda[5000:20000, ])
-
-CLGPNM.Beta11.m <- mode_est(res_CLGPNM$Beta[1,1,5000:20000])
-CLGPNM.Beta21.m <- mode_est(res_CLGPNM$Beta[2,1,5000:20000])
-CLGPNM.Beta12.m <- mode_est(res_CLGPNM$Beta[1,2,5000:20000])
-CLGPNM.Beta22.m <- mode_est(res_CLGPNM$Beta[2,2,5000:20000])
-CLGPNM.Beta13.m <- mode_est(res_CLGPNM$Beta[1,3,5000:20000])
-CLGPNM.Beta23.m <- mode_est(res_CLGPNM$Beta[2,3,5000:20000])
-
-CLGPNM.Beta11.hpd <- hpd_est(res_CLGPNM$Beta[1,1,5000:20000])
-CLGPNM.Beta21.hpd <- hpd_est(res_CLGPNM$Beta[2,1,5000:20000])
-CLGPNM.Beta12.hpd <- hpd_est(res_CLGPNM$Beta[1,2,5000:20000])
-CLGPNM.Beta22.hpd <- hpd_est(res_CLGPNM$Beta[2,2,5000:20000])
-CLGPNM.Beta13.hpd <- hpd_est(res_CLGPNM$Beta[1,3,5000:20000])
-CLGPNM.Beta23.hpd <- hpd_est(res_CLGPNM$Beta[2,3,5000:20000])
-
-CLGPNM.Betacon11.m <- mode_est(res_CLGPNM$Betacon[1,1,5000:20000])
-CLGPNM.Betacon21.m <- mode_est(res_CLGPNM$Betacon[2,1,5000:20000])
-CLGPNM.Betacon12.m <- mode_est(res_CLGPNM$Betacon[1,2,5000:20000])
-CLGPNM.Betacon22.m <- mode_est(res_CLGPNM$Betacon[2,2,5000:20000])
-CLGPNM.Betacon13.m <- mode_est(res_CLGPNM$Betacon[1,3,5000:20000])
-CLGPNM.Betacon23.m <- mode_est(res_CLGPNM$Betacon[2,3,5000:20000])
-
-CLGPNM.Betacon11.hpd <- hpd_est(res_CLGPNM$Betacon[1,1,5000:20000])
-CLGPNM.Betacon21.hpd <- hpd_est(res_CLGPNM$Betacon[2,1,5000:20000])
-CLGPNM.Betacon12.hpd <- hpd_est(res_CLGPNM$Betacon[1,2,5000:20000])
-CLGPNM.Betacon22.hpd <- hpd_est(res_CLGPNM$Betacon[2,2,5000:20000])
-CLGPNM.Betacon13.hpd <- hpd_est(res_CLGPNM$Betacon[1,3,5000:20000])
-CLGPNM.Betacon23.hpd <- hpd_est(res_CLGPNM$Betacon[2,3,5000:20000])
-
-CLGPNM.Sigma11.m <- mode_est(res_CLGPNM$Sigma[1,1,5000:20000])
-CLGPNM.Sigma12.m <- mode_est(res_CLGPNM$Sigma[1,2,5000:20000])
-CLGPNM.Sigma13.m <- mode_est(res_CLGPNM$Sigma[1,3,5000:20000])
-CLGPNM.Sigma22.m <- mode_est(res_CLGPNM$Sigma[2,2,5000:20000])
-CLGPNM.Sigma23.m <- mode_est(res_CLGPNM$Sigma[2,3,5000:20000])
-CLGPNM.Sigma33.m <- mode_est(res_CLGPNM$Sigma[3,3,5000:20000])
-
-CLGPNM.Sigma11.hpd <- hpd_est(res_CLGPNM$Sigma[1,1,5000:20000])
-CLGPNM.Sigma12.hpd <- hpd_est(res_CLGPNM$Sigma[1,2,5000:20000])
-CLGPNM.Sigma13.hpd <- hpd_est(res_CLGPNM$Sigma[1,3,5000:20000])
-CLGPNM.Sigma22.hpd <- hpd_est(res_CLGPNM$Sigma[2,2,5000:20000])
-CLGPNM.Sigma23.hpd <- hpd_est(res_CLGPNM$Sigma[2,3,5000:20000])
-CLGPNM.Sigma33.hpd <- hpd_est(res_CLGPNM$Sigma[3,3,5000:20000])
-
-CLGPNM.Sigmacon11.m <- mode_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
-CLGPNM.Sigmacon12.m <- mode_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
-CLGPNM.Sigmacon13.m <- mode_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
-CLGPNM.Sigmacon22.m <- mode_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
-CLGPNM.Sigmacon23.m <- mode_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
-CLGPNM.Sigmacon33.m <- mode_est(res_CLGPNM$Sigmacon[3,3,5000:20000])
-
-CLGPNM.Sigmacon11.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
-CLGPNM.Sigmacon12.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
-CLGPNM.Sigmacon13.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
-CLGPNM.Sigmacon22.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
-CLGPNM.Sigmacon23.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
-CLGPNM.Sigmacon33.hpd <- hpd_est(res_CLGPNM$Sigmacon[3,3,5000:20000])
-
-modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
-                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
-                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
-                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
-                CLGPNM.lambda.m)
-hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
-                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
-                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
-                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
-                CLGPNM.lambda.hpd[1])
-hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
-                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
-                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
-                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
-                CLGPNM.lambda.hpd[2])
-
-modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
-                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
-                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
-                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
-                   CLGPNM.lambda.m)
-hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
-                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
-                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
-                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
-                   CLGPNM.lambda.hpd[1])
-hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
-                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
-                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
-                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
-                   CLGPNM.lambda.hpd[2])
-
-CLGPNMres.tab <- round(cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN, modescon.GPN, hpdcon.LB.GPN, hpdcon.UB.GPN),2)
-
-rownames(CLGPNMres.tab) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
-                             "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
-                             "$\\beta_{1_s^{II}}$", "$\\beta_{1_y}$",
-                             "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
-                             "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
-                             "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
-                             "$\\lambda$")
-colnames(CLGPNMres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")
-
-kable(CLGPNMres.tab, "latex", booktabs = TRUE, escape = FALSE, caption = "Results for the modified joint projected and skew normal model")%>%
-add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))
-
-```
-
-Tables 2, 3 and 4 show results from the four models that were fit to the teacher
-dataset. For the models in which Bayesian estimation was used we show both the 
-estimated posterior mode and the 95\% highest posterior density (HPD) interval 
-for each parameter. Before we turn to an evaluation of the fit of the models, we
-discuss results of and the interpretation of parameters from the four models. We
-will cover parameters concerning the circular outcome, linear outcome and the
+Before analysis there were a couple of settings that we had to specify
+for the cylindrical models. Starting values for the Abe-Ley model were
+the following
+\(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
+The initial amount of iterations for the three MCMC samplers was set to
+2000. After we checked convergence via traceplots we concluded that some
+of the parameters of the joint GPN-SSN model did not converge. Therefore
+we set the amount of iterations of the MCMC models to 20,000 and
+subtracted a burn-in of 5000. With this specification the MCMC chains
+for all parameters converge. Note that we choose the same amount of
+iterations for all three Bayesian cylindrical models to make their
+comparison via the PLSL as fair as possible. Lastly, the predictor
+\verb|SE| was centered before inclusion in the analysis.
+
+\begin{table}
+
+\caption{\label{tab:estCLGPN}Results for the modified CL-PN and CL-GPN model}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{CL-PN} & \multicolumn{3}{c}{CL-GPN} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
+  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
+\midrule
+$\beta_0^{I}$ & 1.75 & 1.46 & 2.00 & 2.39 & 1.87 & 3.00\\
+$\beta_1^{I}$ & 0.62 & 0.37 & 0.84 & 0.74 & 0.36 & 1.19\\
+$\beta_0^{II}$ & 1.12 & 0.89 & 1.33 & 1.40 & 1.12 & 1.73\\
+$\beta_1^{II}$ & 0.54 & 0.36 & 0.76 & 0.65 & 0.43 & 0.90\\
+$\gamma_0$ & 0.37 & 0.31 & 0.43 & 0.36 & 0.30 & 0.41\\
+\addlinespace
+$\gamma_{cos}$ & 0.04 & 0.02 & 0.07 & 0.03 & 0.01 & 0.05\\
+$\gamma_{sin}$ & -0.02 & -0.05 & 0.02 & -0.01 & -0.04 & 0.02\\
+$\gamma_1$ & 0.03 & 0.00 & 0.07 & 0.03 & 0.00 & 0.06\\
+$\sigma$ & 0.14 & 0.12 & 0.16 & 0.13 & 0.12 & 0.16\\
+$\sum_{1,1}$ & NA & NA & NA & 3.02 & 1.87 & 4.97\\
+\addlinespace
+$\sum_{1,2}$ & NA & NA & NA & 0.59 & 0.20 & 0.88\\
+$\sum_{2,2}$ & NA & NA & NA & 1.00 & 1.00 & 1.00\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:estSL}Results for the modified Abe-Ley model}
+\centering
+\begin{tabular}[t]{lr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
+$\eta_0$ & 0.36\\
+$\eta_1$ & -0.02\\
+$\nu_0$ & 1.19\\
+$\nu_1$ & 0.01\\
+$\alpha$ & 3.82\\
+\addlinespace
+$\kappa$ & 1.58\\
+$\lambda$ & 0.68\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:estCLMGPN}Results for the modified joint projected and skew normal model}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
+  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
+\midrule
+$\beta_{0_s^{I}}$ & 0.31 & 0.26 & 0.34 & 2.07 & 1.77 & 2.50\\
+$\beta_{0_s^{II}}$ & 0.19 & 0.16 & 0.21 & 1.28 & 1.06 & 1.53\\
+$\beta_{0_y}$ & 0.33 & 0.30 & 0.36 & 0.33 & 0.30 & 0.36\\
+$\beta_{1_s^{I}}$ & 0.08 & 0.04 & 0.12 & 0.54 & 0.28 & 0.82\\
+$\beta_{1_s^{II}}$ & 0.07 & 0.04 & 0.09 & 0.46 & 0.29 & 0.65\\
+\addlinespace
+$\beta_{1_y}$ & 0.08 & 0.05 & 0.12 & 0.08 & 0.05 & 0.12\\
+$\sum_{s_{1,1}}$ & 0.05 & 0.04 & 0.06 & 2.30 & 1.66 & 3.34\\
+$\sum_{s_{2,2}}$ & 0.02 & 0.02 & 0.03 & 1.00 & 1.00 & 1.00\\
+$\sum_{y_{3,3}}$ & 0.03 & 0.02 & 0.04 & 0.03 & 0.02 & 0.04\\
+$\sum_{s_{1,2}}$ & 0.00 & 0.00 & 0.01 & 0.09 & -0.20 & 0.35\\
+\addlinespace
+$\sum_{sy_{1,3}}$ & 0.03 & 0.03 & 0.05 & 0.23 & 0.18 & 0.32\\
+$\sum_{sy_{2,3}}$ & 0.01 & 0.01 & 0.02 & 0.09 & 0.05 & 0.12\\
+$\lambda$ & 0.15 & 0.13 & 0.17 & 0.15 & 0.13 & 0.17\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Tables 2, 3 and 4 show results from the four models that were fit to the
+teacher dataset. For the models in which Bayesian estimation was used we
+show both the estimated posterior mode and the 95\% highest posterior
+density (HPD) interval for each parameter. Before we turn to an
+evaluation of the fit of the models, we discuss results of and the
+interpretation of parameters from the four models. We will cover
+parameters concerning the circular outcome, linear outcome and the
 association between them separately.
@@
 
-Firstly, we can compare the estimated mean of the circular outcome between the
-four models. In Table 5 we see the means of the posterior predictive
-distributions for the circular outcome of the CL-PN, CL-GPN and joint GPN-SSN 
-model. For the CL-PN model we can actually also compute this mean from the
-estimates in Table 2 as follows: $\mu_{circ} = atan2(\beta_0^{II}, \beta_0^{I})
-= 32.62^\circ$. The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are
-about equal and correspond quite well to the actual data mean of 33.22$^\circ$
-in Table \ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model
-which is 0.36 radians or 20.62$^\circ$ is different. This difference could be
-caused by the fact that the densities for the circular ouctome, projected normal
-or sine-skewed von Mises, differ between the models.
-
-```{r means, echo = FALSE}
-means <- round(rbind(mean_circ(res_CLPN$theta_pred[,5000:20000]),
-                     mean_circ(res_CLGPN$theta_pred[,5000:20000]),
-                     mean_circ(res_CLGPNM$theta_pred[,5000:20000]))*(180/pi), 2)
-
-rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
-colnames(means) <- "Mean (degrees)"
-kable(means, "latex", booktabs = T, escape = FALSE, caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models")%>%
-   kable_styling() %>%
-   add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", notation="alphabet")
-
-```
-
-We can also compare the effect of self-efficacy (\verb|SE|) on the circular
-outcome. For the CL-PN model and Abe-Ley models we can get estimates of a
-circular regression coefficient of this effect. For the Abe-Ley model this
-estimate is the parameter $\eta_1 = -0.02$. This means that for each unit
-increase in self-efficacy, at the inflection point of the circular regression
-line, the score of the teacher on the IPC decreases with $0.02*(180/\pi) =
-1.15^\circ$. \textcolor{red}{CHRISTOPHE: KLOPT DIT, OF MOET IK $\lambda$ HIER NOG MEENEMEN; NEGEER IK NU HET SINE-SKEWED GEDEELTE} The inflection point is the point at which the regression line
-starts flattening off, i.e. the steepness of the slope decreases. For the CL-PN
-model we can use methods from @CremersMulderKlugkist2017. The estimated
-posterior mode of $b_c$, a parameter that is comparable to $\eta_1$ in the
-Abe-Ley model, equals $1.53$ and its 95\% HPD interval is $(-17.37; 21.26)$.
-This means that for each unit increase in self-efficacy, at the inflection point
-of the circular regression line, the score of the teacher on the IPC increases
-with $1.53*(180/\pi) = 87.66^\circ$. The values of $b_c$ and $\eta_1$ are thus
-quite different even though they should represent the same effect. However, the
-slope at the inflection point that $b_c$ and $\eta_1$ describe is not
-necessarily representative of the effect of \verb|SE| in the data range. In
-Figure \ref{regline} we see that the inflection point (square) of the predicted
-circular regression line for the CL-PN model lies just ouside of the data range
-and that for the Abe-Ley model even further (outside the x-axis range). The
-slope of the regression lines in the data range is much more alike even though
-one is positive and the other negative. The slope at the mean self-efficicay,
-which has a value of 0 because it was centered, $SAM$ is estimated at $0.07
-(-0.03; 0.17)$. The $SAM$ can be interpreted as the amount of change on the
-circle (in radians) for a unit increase at the mean of the predictor variable.
-However, for the CL-PN model the 95\% HPD interval of $SAM$ for the effect of
-\verb|SE| includes 0 meaning that the value is not different from 0. THis
-explains in part how the slopes of the two regression lines are predicted to be
-different. For the Abe-Ley model standard errors of the parameters are not known
-so we cannot formally test whether $\eta_1$ differs from 0. Neither is there a
-known way of computing the effect at predictor values other than the one of the
+Firstly, we can compare the estimated mean of the circular outcome
+between the four models. In Table 5 we see the means of the posterior
+predictive distributions for the circular outcome of the CL-PN, CL-GPN
+and joint GPN-SSN model. For the CL-PN model we can actually also
+compute this mean from the estimates in Table 2 as follows:
+\(\mu_{circ} = atan2(\beta_0^{II}, \beta_0^{I}) = 32.62^\circ\). The
+estimates for the CL-PN, CL-GPN and joint GPN-SSN model are about equal
+and correspond quite well to the actual data mean of 33.22\(^\circ\) in
+Table \ref{Tableteacherdescriptives}. The estimate from the Abe-Ley
+model which is 0.36 radians or 20.62\(^\circ\) is different. This
+difference could be caused by the fact that the densities for the
+circular ouctome, projected normal or sine-skewed von Mises, differ
+between the models.
+
+\begin{table}
+
+\caption{\label{tab:means}Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models}
+\centering
+\begin{tabular}[t]{lr}
+\toprule
+  & Mean (degrees)\\
+\midrule
+Cl-PN & 31.79\\
+CL-GPN & 33.38\\
+GPN-SSN & 34.47\\
+\bottomrule
+\multicolumn{2}{l}{\textsuperscript{a} Note that these means}\\
+\multicolumn{2}{l}{are based on the}\\
+\multicolumn{2}{l}{posterior predictive}\\
+\multicolumn{2}{l}{distribution for the}\\
+\multicolumn{2}{l}{intercepts following}\\
+\multicolumn{2}{l}{(Wang \& Gelfand,}\\
+\multicolumn{2}{l}{2013).}\\
+\end{tabular}
+\end{table}
+
+We can also compare the effect of self-efficacy (\verb|SE|) on the
+circular outcome. For the CL-PN model and Abe-Ley models we can get
+estimates of a circular regression coefficient of this effect. For the
+Abe-Ley model this estimate is the parameter \(\eta_1 = -0.02\). This
+means that for each unit increase in self-efficacy, at the inflection
+point of the circular regression line, the score of the teacher on the
+IPC decreases with \(0.02*(180/\pi) = 1.15^\circ\).
+\textcolor{red}{CHRISTOPHE: KLOPT DIT, OF MOET IK $\lambda$ HIER NOG MEENEMEN; NEGEER IK NU HET SINE-SKEWED GEDEELTE}
+The inflection point is the point at which the regression line starts
+flattening off, i.e.~the steepness of the slope decreases. For the CL-PN
+model we can use methods from Cremers, Mulder, \& Klugkist (2018b). The
+estimated posterior mode of \(b_c\), a parameter that is comparable to
+\(\eta_1\) in the Abe-Ley model, equals \(1.53\) and its 95\% HPD
+interval is \((-17.37; 21.26)\). This means that for each unit increase
+in self-efficacy, at the inflection point of the circular regression
+line, the score of the teacher on the IPC increases with
+\(1.53*(180/\pi) = 87.66^\circ\). The values of \(b_c\) and \(\eta_1\)
+are thus quite different even though they should represent the same
+effect. However, the slope at the inflection point that \(b_c\) and
+\(\eta_1\) describe is not necessarily representative of the effect of
+\verb|SE| in the data range. In Figure \ref{regline} we see that the
+inflection point (square) of the predicted circular regression line for
+the CL-PN model lies just ouside of the data range and that for the
+Abe-Ley model even further (outside the x-axis range). The slope of the
+regression lines in the data range is much more alike even though one is
+positive and the other negative. The slope at the mean self-efficicay,
+which has a value of 0 because it was centered, \(SAM\) is estimated at
+\(0.07 (-0.03; 0.17)\). The \(SAM\) can be interpreted as the amount of
+change on the circle (in radians) for a unit increase at the mean of the
+predictor variable. However, for the CL-PN model the 95\% HPD interval
+of \(SAM\) for the effect of \verb|SE| includes 0 meaning that the value
+is not different from 0. THis explains in part how the slopes of the two
+regression lines are predicted to be different. For the Abe-Ley model
+standard errors of the parameters are not known so we cannot formally
+test whether \(\eta_1\) differs from 0. Neither is there a known way of
+computing the effect at predictor values other than the one of the
 inflection point.
 
-For the CL-GPN and GPN-SSN we cannot compute circular regression coefficients. 
-Instead, we will compute posterior predictive distributions for the predicted 
-circular outcome of individuals scoring the minimum, maximum and average 
-self-efficacy. The modes and 95\% HPD intervals of these posterior predictive 
-distributions in the CL-GPN model are $\hat{\theta}_{SE_{min}} = 211.31^\circ 
-(144.41^\circ \: 43.43^\circ)$, $\hat{\theta}_{SE_{mean}} = 25.82^\circ 
-(335.55^\circ \: 145.11^\circ)$ and $\hat{\theta}_{SE_{max}} =30.96^\circ 
-(6.72^\circ \: 74.35^\circ)$ respectively. For the GPN-SSN model the modes and
-95\% HPD intervals of the posterior predictive distributions are
-$\hat{\theta}_{SE_{min}} = 352.06^\circ (121.82^\circ \: 77.75^\circ)$,
-$\hat{\theta}_{SE_{mean}} = 22.74^\circ (336.71^\circ \: 132.25^\circ)$ and
-$\hat{\theta}_{SE_{max}} = 30.87^\circ (358.80^\circ \: 82.13^\circ)$
-respectively. For both models we can thus conclude that as the self-efficacy
-increases, the score of the teacher on the IPC moves counterlockwise. We can
-however not reach any conclusions about the size of this effect or whether it is
-different from zero.
-
-```{r ppreg, echo = FALSE, results = FALSE}
-mode_est_circ(res_CLGPN$theta_pred.min)
-mode_est_circ(res_CLGPN$theta_pred.mean)
-mode_est_circ(res_CLGPN$theta_pred.max)
-
-hpd_est_circ(res_CLGPN$theta_pred.min)
-hpd_est_circ(res_CLGPN$theta_pred.mean)
-hpd_est_circ(res_CLGPN$theta_pred.max)
-
-mode_est_circ(res_CLGPNM$theta_pred.min)
-mode_est_circ(res_CLGPNM$theta_pred.mean)
-mode_est_circ(res_CLGPNM$theta_pred.max)
-
-hpd_est_circ(res_CLGPNM$theta_pred.min)
-hpd_est_circ(res_CLGPNM$theta_pred.mean)
-hpd_est_circ(res_CLGPNM$theta_pred.max)
-
-```
-
-```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-a1 <- mode_est(res_CLPN$BI[5000:20000, 1])
-b1 <- mode_est(res_CLPN$BI[5000:20000, 2])
-a2 <- mode_est(res_CLPN$BII[5000:20000, 1])
-b2 <- mode_est(res_CLPN$BII[5000:20000, 2])
-ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-
-x <- seq(-5, 5, by = 0.1)
-
-pred.AL <- 0.37 + 2*atan(-0.01*x)
-pred.PN <- atan2(a2 + b2*x, a1 + b1*x)
-
-plot(Dat.test$SEc, theta.test*(180/pi), bty = "n", yaxt = "n",
-     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "self-efficacy")
-axis(side = 2,  at = c(-160, -80, 0, 80, 160))
-points(x, pred.AL*(180/pi), type = "l", lty = 1)
-points(x, pred.PN*(180/pi), type = "l", lty = 2)
-points(ax, ac*(180/pi), pch = 15)
-
-dev.off()
-tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)
-
-```
-
-```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}
-
-a1 <- res_CLPN$BI[5000:20000, 1]
-b1 <- res_CLPN$BI[5000:20000, 2]
-a2 <- res_CLPN$BII[5000:20000, 1]
-b2 <- res_CLPN$BII[5000:20000, 2]
-
-ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-bc <- -tan(atan2(a2, a1)-ac)/ax
-
-mode_est(ax)
-
-mode_est(bc)
-hpd_est(bc)
-
-mode_est(bc/(1 + (bc*(-ax))^2))
-hpd_est(bc/(1 + (bc*(-ax))^2))
-
-```
+For the CL-GPN and GPN-SSN we cannot compute circular regression
+coefficients. Instead, we will compute posterior predictive
+distributions for the predicted circular outcome of individuals scoring
+the minimum, maximum and average self-efficacy. The modes and 95\% HPD
+intervals of these posterior predictive distributions in the CL-GPN
+model are
+\(\hat{\theta}_{SE_{min}} = 211.31^\circ (144.41^\circ \: 43.43^\circ)\),
+\(\hat{\theta}_{SE_{mean}} = 25.82^\circ (335.55^\circ \: 145.11^\circ)\)
+and \(\hat{\theta}_{SE_{max}} =30.96^\circ (6.72^\circ \: 74.35^\circ)\)
+respectively. For the GPN-SSN model the modes and 95\% HPD intervals of
+the posterior predictive distributions are
+\(\hat{\theta}_{SE_{min}} = 352.06^\circ (121.82^\circ \: 77.75^\circ)\),
+\(\hat{\theta}_{SE_{mean}} = 22.74^\circ (336.71^\circ \: 132.25^\circ)\)
+and
+\(\hat{\theta}_{SE_{max}} = 30.87^\circ (358.80^\circ \: 82.13^\circ)\)
+respectively. For both models we can thus conclude that as the
+self-efficacy increases, the score of the teacher on the IPC moves
+counterlockwise. We can however not reach any conclusions about the size
+of this effect or whether it is different from zero.
 
@@
 
-For the linear outcome, we compare the regression coefficients of \verb|SE| for
-the linear outcome, the strength of the interpersonal behavior of a teacher. For
-the CL-PN, CL-GPN and joint GPN-SSN model this are the parameters $\gamma_1$ and
-$\beta_{1_y}$ respectively. Note that in the CL-PN and CL-GPN model the effect 
-of \verb|SE| is the effect controlled for $\cos\theta$ and $\sin\theta$ similar 
-to a 'normal' multiple regression model. For the GPN-SSN model the effect of
-\verb|SE| is also controlled for the circular outcome. However in this model the
-associaton is modelled in a way that is similar to modelling the association
-between several outcome variables in a multivariate regression or ANOVA model.
-Additionally, the distribution we assume for the linear outcome is different.
-For the CL-PN and CL-GPN models we assume a normal distribution while for the
-GPN-SSN model we assume a skewed normal distribution. These differences will
-lead to differences in the estimated effect of \verb|SE| on the linear outcome.
-The interpretation of the regression coefficients is however similar for all
-three models and is like the interpretation of a regression coefficient in any
-other standard linear regression model.
-
-For all three models there is an effect of \verb|SE| on the strength of the 
-interpersonal behavior of a teacher, the HPD interval does not include 0:
-$\gamma_1 = 0.03 \: (0.00; 0.07)$ (CL-PN), $\gamma_1 = 0.03 \: (0.00; 0.06)$
-(CL-GPN) and $\beta_{1_y} = 0.08 \: (0.05; 0.12)$ (GPN-SSN). This means that
-teachers with a higher self-efficacy for classroom management are stronger in
-their interpersonal behavior than teachers with a lower self-efficacy. In
-particular, for each unit increase in self-efficacy the strength of their
-interpersonal behavior is 0.03 or 0.08 (depending on the model) higher. This is
-an increase of 4.17 or 11.11 \% considering the range of the strength on the IPC
-in the data.
-
-For the Abe-Ley model we make use of the fact that the conditional distribution
-for the linear outcome is Weibull. This means that we can use methods from
-survival analsis in medical statistics to interpret the effect of self-efficacy.
-In survival analysis they make use of a 'survival' function in which the time is
-plotted against the probability of survival of subjects suffering from a
-specific medical condition. In our data however we plot the strength on the IPC
-(instead of time) against the probability of a teacher having such a strength.
-This probability can be computed using the so-called 'survival-function' which
-in our case is equal to $\exp(-\alpha y_i^{\beta})$ where $\beta = \exp(\nu_0 +
-\nu_1SE_i)$. In Figure \ref{reglineweib} we plot the survival function for the
-minimum, mean and maximum value of self-efficacy. From this Figure we conclude
-that the stronger interpersonal behaviors are less probable. We also see that
-the self-efficacy positively influences the strength on the IPC, the probability
-of having a stronger interpersonal behavior increases with increasing
-self-efficacy. This result is equivalent to the effect of self-efficacy found
-using the CL-PN, CL-GPN and GPN-SSN models.
-
-```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-summary(Dat.test$SEc)
-
-x <- c(-3.5, 0, 2)
-y <- seq(0, 1, by = 0.05)
-h <- matrix(NA, length(y), length(x))
-s <- matrix(NA, length(y), length(x))
-
-for(j in 1:length(x)){
-
-  for(i in 1:length(y)){
-    beta <- exp(resAL$par[3] + resAL$par[4]*x[j])
-    h[i,j] <- beta*resAL$par[6]*y[i]^(resAL$par[6]-1)
-    s[i,j] <- exp(-resAL$par[6]*y[i]^beta)
-  }
-
-}
-
-tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
-points(y, s[,2], type = "l", lty = 2)
-points(y, s[,3], type = "l", lty = 3)
-legend(0.7, 1, legend = c("min(SE)", "mean(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")
-
-dev.off()
-tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)
-
-```
+For the linear outcome, we compare the regression coefficients of
+\verb|SE| for the linear outcome, the strength of the interpersonal
+behavior of a teacher. For the CL-PN, CL-GPN and joint GPN-SSN model
+this are the parameters \(\gamma_1\) and \(\beta_{1_y}\) respectively.
+Note that in the CL-PN and CL-GPN model the effect of \verb|SE| is the
+effect controlled for \(\cos\theta\) and \(\sin\theta\) similar to a
+`normal' multiple regression model. For the GPN-SSN model the effect of
+\verb|SE| is also controlled for the circular outcome. However in this
+model the associaton is modelled in a way that is similar to modelling
+the association between several outcome variables in a multivariate
+regression or ANOVA model. Additionally, the distribution we assume for
+the linear outcome is different. For the CL-PN and CL-GPN models we
+assume a normal distribution while for the GPN-SSN model we assume a
+skewed normal distribution. These differences will lead to differences
+in the estimated effect of \verb|SE| on the linear outcome. The
+interpretation of the regression coefficients is however similar for all
+three models and is like the interpretation of a regression coefficient
+in any other standard linear regression model.
+
+For all three models there is an effect of \verb|SE| on the strength of
+the interpersonal behavior of a teacher, the HPD interval does not
+include 0: \(\gamma_1 = 0.03 \: (0.00; 0.07)\) (CL-PN),
+\(\gamma_1 = 0.03 \: (0.00; 0.06)\) (CL-GPN) and
+\(\beta_{1_y} = 0.08 \: (0.05; 0.12)\) (GPN-SSN). This means that
+teachers with a higher self-efficacy for classroom management are
+stronger in their interpersonal behavior than teachers with a lower
+self-efficacy. In particular, for each unit increase in self-efficacy
+the strength of their interpersonal behavior is 0.03 or 0.08 (depending
+on the model) higher. This is an increase of 4.17 or 11.11 \%
+considering the range of the strength on the IPC in the data.
+
+For the Abe-Ley model we make use of the fact that the conditional
+distribution for the linear outcome is Weibull. This means that we can
+use methods from survival analsis in medical statistics to interpret the
+effect of self-efficacy. In survival analysis they make use of a
+`survival' function in which the time is plotted against the probability
+of survival of subjects suffering from a specific medical condition. In
+our data however we plot the strength on the IPC (instead of time)
+against the probability of a teacher having such a strength. This
+probability can be computed using the so-called `survival-function'
+which in our case is equal to \(\exp(-\alpha y_i^{\beta})\) where
+\(\beta = \exp(\nu_0 + \nu_1SE_i)\). In Figure \ref{reglineweib} we plot
+the survival function for the minimum, mean and maximum value of
+self-efficacy. From this Figure we conclude that the stronger
+interpersonal behaviors are less probable. We also see that the
+self-efficacy positively influences the strength on the IPC, the
+probability of having a stronger interpersonal behavior increases with
+increasing self-efficacy. This result is equivalent to the effect of
+self-efficacy found using the CL-PN, CL-GPN and GPN-SSN models.
 
@@
 
-For all four models we can also investigate the association between the linear 
-and circular component. In the CL-PN and CL-GPN model we do this using the 
-parameters $\gamma_{\cos}$ and $\gamma_{sin}$, the coefficients of the
-regression of the linear outcome onto the sine and cosine of the circular
-outcome. In the joint SSN-GPN model we look at the covariances between the
-linear outcome and the sine and cosine of the circular outcome $\sum_{sy_{2,3}}$
-and $\sum_{sy_{1,3}}$. For the Abe-Ley model we can use the model parameters in
-Table 3 to compute a circular-linear correlation [@abe2017tractable].
-
-In both the CL-PN and CL-GPN model there is an effect of the cosine of the 
-circular outcome onto the linear outcome (HPD interval does not include 0). In 
-the teacher data the sine and cosine component have a substantive meaning. In 
-this case the Communion component of the IPC positively effects the strength of 
-a teachers' type of interpersonal behavior. Thus the teachers showing 
-interpersonal behavior types with higher communion scores (e.g. 'helpful' and 
-'understanding' in Figure \ref{QTI}) are stronger in their behavior. In the 
-joint SSN-GPN model we reach a sligtly different conclusion. In this model both 
-covariances, $\sum_{sy_{2,3}} = 0.09$ and $\sum_{sy_{1,3}} = 0.23$, are 
-different from 0, but the one between the cosine of the circular outcome and the
+For all four models we can also investigate the association between the
+linear and circular component. In the CL-PN and CL-GPN model we do this
+using the parameters \(\gamma_{\cos}\) and \(\gamma_{sin}\), the
+coefficients of the regression of the linear outcome onto the sine and
+cosine of the circular outcome. In the joint SSN-GPN model we look at
+the covariances between the linear outcome and the sine and cosine of
+the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). For
+the Abe-Ley model we can use the model parameters in Table 3 to compute
+a circular-linear correlation (Abe \& Ley, 2017).
+
+In both the CL-PN and CL-GPN model there is an effect of the cosine of
+the circular outcome onto the linear outcome (HPD interval does not
+include 0). In the teacher data the sine and cosine component have a
+substantive meaning. In this case the Communion component of the IPC
+positively effects the strength of a teachers' type of interpersonal
+behavior. Thus the teachers showing interpersonal behavior types with
+higher communion scores (e.g. `helpful' and `understanding' in Figure
+\ref{QTI}) are stronger in their behavior. In the joint SSN-GPN model we
+reach a sligtly different conclusion. In this model both covariances,
+\(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are different
+from 0, but the one between the cosine of the circular outcome and the
 linear outcome is larger. This means that both the Communion and Agency
-component of the IPC positively effect the strength of a teachers' type of 
-interpersonal behavior but the effect of the Communon component is larger. Thus,
-teachers scoring high on Agency but slightly higher on Communion (the 'helpful' 
-category) are strongest in their behavior. Following @abe2017tractable the
-estimated correlation between the circular and linear outcome is equal to
+component of the IPC positively effect the strength of a teachers' type
+of interpersonal behavior but the effect of the Communon component is
+larger. Thus, teachers scoring high on Agency but slightly higher on
+Communion (the `helpful' category) are strongest in their behavior.
+Following Abe \& Ley (2017) the estimated correlation between the
+circular and linear outcome is equal to
 \textcolor{red}{CHRISTOPHE: KAN JIJ DEZE BEREKENEN, IN R KAN IK GEEN MANIER
@@
 GEBRUIKT KRIJG IK IMAGINAIRE GETALLEN (voor m = 1 en m = 2), IK WEET NIET OF DAT
-DE BEDOELING IS..}. This correlation is not affected by $\mu$ or $\beta$ meaning
-that the covariate, self-efficacy, can not influence the correlation between the
-type and strength of interpersonal behavior.
-
-```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}
-
-require(pracma)
-
-alpha <- resAL$par[6]
-kappa <- resAL$par[5]
-lambda <- resAL$par[7]
-
-#R cannot compute legendre functions of non-integer degree
-#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
-#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
-#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]
-
-#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
-
-1/3.82
-cosh(1.58)
-
-```
+DE BEDOELING IS..}. This correlation is not affected by \(\mu\) or
+\(\beta\) meaning that the covariate, self-efficacy, can not influence
+the correlation between the type and strength of interpersonal behavior.
 
@@
 
-Instead of interpreting specific model parameters we can also look at the 
-overall fit of the models to the data. In Section \ref{Modelfit} we introduced 
-the criterion, PLSL, that we use to assess model fit in this paper. This
-criterion is focused on the out-of-sample predictive performance. Table 6 shows
-the values of this criterion for the linear and circular outcome of the four
-different models fit to the teacher data.
-
-The CL-PN and CL-GPN model have the best out-of-sample predictive performance 
-for the linear outcome. Note that they have roughly the same performance since 
-they model the linear outcome in the same way apart from the value of $r$ in 
-\eqref{circlinlink} which is a parameter computed in the estimation of the 
-parameters for the circular variable. This similarity in fit is also shown in 
-Figure \ref{preddist} which shows histograms of the test and holdout data
-together with the (posterior) predicitive distributions for all four cylindrical
-models. The posterior predictive distributions for the CL-PN and CL-GPN model
+Instead of interpreting specific model parameters we can also look at
+the overall fit of the models to the data. In Section \ref{Modelfit} we
+introduced the criterion, PLSL, that we use to assess model fit in this
+paper. This criterion is focused on the out-of-sample predictive
+performance. Table 6 shows the values of this criterion for the linear
+and circular outcome of the four different models fit to the teacher
+data.
+
+The CL-PN and CL-GPN model have the best out-of-sample predictive
+performance for the linear outcome. Note that they have roughly the same
+performance since they model the linear outcome in the same way apart
+from the value of \(r\) in \eqref{circlinlink} which is a parameter
+computed in the estimation of the parameters for the circular variable.
+This similarity in fit is also shown in Figure \ref{preddist} which
+shows histograms of the test and holdout data together with the
+(posterior) predicitive distributions for all four cylindrical models.
+The posterior predictive distributions for the CL-PN and CL-GPN model
 are almost the same for the linear outcome.
 
-The CL-GPN model has the best out-of-sample predicitve performance for the 
-circular outcome. The Abe-Ley model has the second best performance. This means 
-that for the circular variable a slightly skewed distribution (see Figure 
-\ref{preddist}) fits best. The GPN-SSN model fits worst even though it assumes
-the same distribution for the circular outcome as the CL-GPN model. This could
-be due to the fact that the relation between the circular and linear outcome is 
-modeled differently in these two models. In the CL-GPN model the linear outcome 
-depends on the circular one following the relation in \eqref{circlinlink} but 
-the circular outcome is not dependent on the linear one. In the GPN-SSN model 
-both outcomes are modelled with a joint variance-covariance matrix governing 
-their dependence. In Figure \ref{preddist} we see however that the actual 
-predictive densities do not differ that much between the four models. It is 
-possible that the large difference in PLSL are caused by the fact that they are 
-computed for a relatively small holdout dataset (n = 15) 
-\textcolor{red}{CHRISTOPHE: IS DIT LOGISCH?}.
-
-```{r ModelFit, echo = FALSE}
-
-CLGPNM.ll.circ <- -2*mean(res_CLGPNM$ll.circ[5000:20000])
-CLGPNM.ll.lin  <- -2*mean(res_CLGPNM$ll.lin[5000:20000])
-
-CLGPN.ll.circ <- -2*mean(res_CLGPN$ll.circ[5000:20000])
-CLGPN.ll.lin  <- -2*mean(res_CLGPN$ll.lin[5000:20000])
-
-CLPN.ll.circ <- -2*mean(res_CLPN$ll.circ[5000:20000])
-CLPN.ll.lin  <- -2*mean(res_CLPN$ll.lin[5000:20000])
-
-AL.ll.circ <- -2*func.reg.cond.circ(resAL$par, dat.h)
-AL.ll.lin  <- -2*func.reg.cond.lin(resAL$par, dat.h)
-
-PLSL <- round(rbind(c(CLPN.ll.circ, CLGPN.ll.circ, AL.ll.circ, CLGPNM.ll.circ),
-                    c(CLPN.ll.lin, CLGPN.ll.lin, AL.ll.lin, CLGPNM.ll.lin)), 2)
-rownames(PLSL) <- c("circular", "linear")
-colnames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "Joint GPN-SSN")
-
-kable(PLSL, "latex", booktabs = T, escape = FALSE, caption = "PLSL criteria for the circular and linear outcome in the four cylindrical models")
-
-```
-
-```{r preddist, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-source("R-code/Sample Abe-Ley.R")
-
-ALdy  <- matrix(NA, length(y.test), 1000)
-ALdth <- matrix(NA, length(theta.test), 1000)
-
-ALdy.h  <- matrix(NA, length(y.hold), 1000)
-ALdth.h <- matrix(NA, length(theta.hold), 1000)
-
-tikz("Plots/preddistdiffSE.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")
-
-layout(matrix(c(1,2,3, 4,5,5), 3, 2, byrow = TRUE),heights=c(2,2,1))
-
-par(mar = c(5,5,1,1))
-
-x.h <- Dat.hold$SEc
-x   <- Dat.test$SEc
-
-for(i in 1:length(theta.test)){
-  beta <- exp(resAL$par[3] + resAL$par[4]*x[i])
-  mu   <- resAL$par[1] + 2*atan(resAL$par[2]*x[i])
-
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu,
-                  kappa = resAL$par[5], lambda = resAL$par[7])
-
-  ALdy[i,] <- dat$x
-  ALdth[i,] <- dat$theta
-}
-
-for(i in 1:length(theta.hold)){
-  beta <- exp(resAL$par[3] + resAL$par[4]*x.h[i])
-  mu   <- resAL$par[1] + 2*atan(resAL$par[2]*x.h[i])
-
-  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu,
-                  kappa = resAL$par[5], lambda = resAL$par[7])
-
-  ALdy.h[i,]  <- dat$x
-  ALdth.h[i,] <- dat$theta
-
-}
-
-hist(theta.test, prob = TRUE, ylim = c(0, 1.25),
-     main = "", ylab = "Circular", xlab = "$\\theta_{test}$")
-lines(density(res_CLPN$theta_pred[,19000:20000]), lty = 1)
-lines(density(res_CLGPN$theta_pred[,19000:20000]), lty = 2)
-lines(density(ALdth), lty = 3)
-lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 4)
-
-hist(theta.hold, prob = TRUE, ylim = c(0, 1.25),
-     main = "", ylab = "Circular", xlab = "$\\theta_{holdout}$")
-lines(density(res_CLPN$theta_pred.hold[,19000:20000]), lty = 1)
-lines(density(res_CLGPN$theta_pred.hold[,19000:20000]), lty = 2)
-lines(density(ALdth.h), lty = 3)
-lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 4)
-
-hist(y.test, prob = TRUE, ylim = c(0, 3),
-     main = "", ylab = "Linear", xlab = "$y_{test}$")
-lines(density(res_CLPN$y_pred[,5000:20000]), lty = 1)
-lines(density(res_CLGPN$y_pred[,5000:20000]), lty = 2)
-lines(density(ALdy), lty = 3)
-lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 4)
-
-hist(y.hold, prob = TRUE, ylim = c(0, 4),
-     main = "", ylab = "Linear", xlab = "$y_{holdout}$")
-lines(density(res_CLPN$y_pred.hold[,5000:20000]), lty = 1)
-lines(density(res_CLGPN$y_pred.hold[,5000:20000]), lty = 2)
-lines(density(ALdy.h), lty = 3)
-lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 4)
-
-par(mar = c(2,2,1,1))
-
-plot(1, type = "n", axes=FALSE, xlab="", ylab="")
-legend(x = "top", inset = 0,
-       legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"),
-       lty = c(1,2,3,4), horiz = TRUE)
-
-dev.off()
-
-tools::texi2dvi("Plots/preddistdiffSE.tex",pdf=TRUE)
-
-```
+The CL-GPN model has the best out-of-sample predicitve performance for
+the circular outcome. The Abe-Ley model has the second best performance.
+This means that for the circular variable a slightly skewed distribution
+(see Figure \ref{preddist}) fits best. The GPN-SSN model fits worst even
+though it assumes the same distribution for the circular outcome as the
+CL-GPN model. This could be due to the fact that the relation between
+the circular and linear outcome is modeled differently in these two
+models. In the CL-GPN model the linear outcome depends on the circular
+one following the relation in \eqref{circlinlink} but the circular
+outcome is not dependent on the linear one. In the GPN-SSN model both
+outcomes are modelled with a joint variance-covariance matrix governing
+their dependence. In Figure \ref{preddist} we see however that the
+actual predictive densities do not differ that much between the four
+models. It is possible that the large difference in PLSL are caused by
+the fact that they are computed for a relatively small holdout dataset
+(n = 15) \textcolor{red}{CHRISTOPHE: IS DIT LOGISCH?}.
+
+\begin{table}
+
+\caption{\label{tab:ModelFit}PLSL criteria for the circular and linear outcome in the four cylindrical models}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+  & CL-PN & CL-GPN & Abe-Ley & Joint GPN-SSN\\
+\midrule
+circular & 86.49 & 46.83 & 59.34 & 111.63\\
+linear & -20.85 & -20.94 & 51.46 & 15.48\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-In this paper we have modified four models for cylindrical data such that they 
-include a regression of both the linear and circular components onto a set of 
-covariates. Subsequently we have used these four methods to analyze a dataset on
-the interpersonal behavior of teachers. Here we will comment on the differences
-between these models, the results from the analysis of the teacher data and how
-cylindrical models improve the analysis of this type of data; data from an
+In this paper we have modified four models for cylindrical data such
+that they include a regression of both the linear and circular
+components onto a set of covariates. Subsequently we have used these
+four methods to analyze a dataset on the interpersonal behavior of
+teachers. Here we will comment on the differences between these models,
+the results from the analysis of the teacher data and how cylindrical
+models improve the analysis of this type of data; data from an
 interpersonal circumplex.
 
-In terms of interpretability, the CL-PN and Abe-Ley model score best. As shown 
-in the previous section it is relatively easy to interpret the parameters of the
-linear and the circular component for these models. In the CL-GPN and joint 
-GPN-SSN model the interpretation of the parameters of the circular component it 
-not easy, if at all possible. This is caused by the fact that in addition to the
-mean vector the covariance matrix of the GPN distribution affects the location 
-of the circular data, making it difficult to compute regression coefficients on 
-the circle. @wang2012directional state that we can use Monte Carlo integration 
-to compute a circular mean and variance for the GPN distribution. This might 
-also be a solution to computing coefficients on the circle and could be applied 
-to the methods of @CremersMulderKlugkist2017 such that they can be used for 
-circular regression coefficients in GPN models as well. \textcolor{red}{HIER OOK
+In terms of interpretability, the CL-PN and Abe-Ley model score best. As
+shown in the previous section it is relatively easy to interpret the
+parameters of the linear and the circular component for these models. In
+the CL-GPN and joint GPN-SSN model the interpretation of the parameters
+of the circular component it not easy, if at all possible. This is
+caused by the fact that in addition to the mean vector the covariance
+matrix of the GPN distribution affects the location of the circular
+data, making it difficult to compute regression coefficients on the
+circle. Wang \& Gelfand (2013) state that we can use Monte Carlo
+integration to compute a circular mean and variance for the GPN
+distribution. This might also be a solution to computing coefficients on
+the circle and could be applied to the methods of Cremers et al. (2018b)
+such that they can be used for circular regression coefficients in GPN
+models as well. \textcolor{red}{HIER OOK
 NOG EEN OPMERKING OVER HOE DE ASSOCIATION TUSSEN CIRCULAR AND LINEAR COMPONENT
@@
 
-In terms of flexibility the joint GPN-SSN scores best. In this model multiple 
-linear and circular outcomes can be included and we can thus apply it to 
-multivariate cylindrical data. In addition for both the linear anc circular 
-components this model and the CL-GPN and CL-PN models are extendable to a 
-mixed-effects model and can thus also be fit to longitudinal data. Extension to 
-mixed-effects, longitudinal (see @nunez2014bayesian and @hernandez2016general 
-for hierarchical/mixed-effects models for the PN and GPN distributions 
-respectively). For the Abe-Ley model this may also be possible and has been done
-for the conditional distribution of its linear component, the Weibull 
-distribution, but has not been done in previous literature for the conditional 
-distribution of its circular component (sine-skewed von 
-Mises)\textcolor{red}{CHRISTOPHE: KLOPT DIT?}. In addition, the joint GPN-SSN
-model allows for non-symmetrical shapes of the distributions of both the linear
-(sine-skewed normal) and the circular (general projected normal) components. The
-Abe-Ley model also allows for this and the CL-GPN model allows for a
-non-symmetrical shape of the circular component.
-
-To investigate model fit for the teacher data we asessed the fit to the linear
-and circular component separately using the PLSL, a criterion that is designed
-to assess the out-of-sample preditive performance. The CL-PN and CL-GPN model
-have the best fit for the linear component and the CL-GPN model has the best fit
-for the circular component. Differences in fit, in addition to being a result of
-the different distributions that were used to model the linear and circular
-component of the data, may also be caused by the way in which the relation
-between the linear and circular component is modeled. Whereas in both the
-Abe-Ley and joint GPN-SSN model the distribution of the linear component is
-conditional on the circular component and vice versa, the distribution of the
-circular component in the CL-PN and CL-GPN model is independent of the linear
-component. In these models the linear component are related through a regression
-structure where the circular component serves as a predictor and the linear
-component as the outcome. We also investigated the (posterior) predictive
-distributions of each model and its fit to the training data, the part of the
-teacher data the model was fit on, and the holdout data, the part of the teacher
-data that we used to assess the out-of-sample predictive performance. From these
-plots we saw that whereas the PLSL criteria seem to indicate a substantive
-difference in fit, the (posterior) predicitve distributions seem to be quite
-similar (especially for the circular component). This discrepancy might be
-caused by the fact that the holdout dataset, for which the PLSL criteria were
-computed, is quite small and thus that small deviations in fit could lead to
-large deviations in the criterion value.
-
-The four cylindrical models that were modified to the regression context in this
-paper are not the only cylindrical distributions available from the literature.
-Several other cylindrical distributions, amongst which are @kato2008dependent,
-@sugasawa2015flexible and @fernandez2007models have been introduced (for more
-references we refer to chapter 2 of @ley2017modern). In the present research we
-have decided not to include these models for reasons of space, complexity of the
-models and ease of implementing a regression structure. In further research
-however it would be interesting to investigate other types of cylindrical
+In terms of flexibility the joint GPN-SSN scores best. In this model
+multiple linear and circular outcomes can be included and we can thus
+apply it to multivariate cylindrical data. In addition for both the
+linear anc circular components this model and the CL-GPN and CL-PN
+models are extendable to a mixed-effects model and can thus also be fit
+to longitudinal data. Extension to mixed-effects, longitudinal (see
+Nuñez-Antonio \& Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al.
+(2016) for hierarchical/mixed-effects models for the PN and GPN
+distributions respectively). For the Abe-Ley model this may also be
+possible and has been done for the conditional distribution of its
+linear component, the Weibull distribution, but has not been done in
+previous literature for the conditional distribution of its circular
+component (sine-skewed von
+Mises)\textcolor{red}{CHRISTOPHE: KLOPT DIT?}. In addition, the joint
+GPN-SSN model allows for non-symmetrical shapes of the distributions of
+both the linear (sine-skewed normal) and the circular (general projected
+normal) components. The Abe-Ley model also allows for this and the
+CL-GPN model allows for a non-symmetrical shape of the circular
+component.
+
+To investigate model fit for the teacher data we asessed the fit to the
+linear and circular component separately using the PLSL, a criterion
+that is designed to assess the out-of-sample preditive performance. The
+CL-PN and CL-GPN model have the best fit for the linear component and
+the CL-GPN model has the best fit for the circular component.
+Differences in fit, in addition to being a result of the different
+distributions that were used to model the linear and circular component
+of the data, may also be caused by the way in which the relation between
+the linear and circular component is modeled. Whereas in both the
+Abe-Ley and joint GPN-SSN model the distribution of the linear component
+is conditional on the circular component and vice versa, the
+distribution of the circular component in the CL-PN and CL-GPN model is
+independent of the linear component. In these models the linear
+component are related through a regression structure where the circular
+component serves as a predictor and the linear component as the outcome.
+We also investigated the (posterior) predictive distributions of each
+model and its fit to the training data, the part of the teacher data the
+model was fit on, and the holdout data, the part of the teacher data
+that we used to assess the out-of-sample predictive performance. From
+these plots we saw that whereas the PLSL criteria seem to indicate a
+substantive difference in fit, the (posterior) predicitve distributions
+seem to be quite similar (especially for the circular component). This
+discrepancy might be caused by the fact that the holdout dataset, for
+which the PLSL criteria were computed, is quite small and thus that
+small deviations in fit could lead to large deviations in the criterion
+value.
+
+The four cylindrical models that were modified to the regression context
+in this paper are not the only cylindrical distributions available from
+the literature. Several other cylindrical distributions, amongst which
+are Kato \& Shimizu (2008), Sugasawa (2015) and Fernández-Durán (2007)
+have been introduced (for more references we refer to chapter 2 of Ley
+\& Verdebout (2017)). In the present research we have decided not to
+include these models for reasons of space, complexity of the models and
+ease of implementing a regression structure. In further research however
+it would be interesting to investigate other types of cylindrical
 distributions as well.
 
-Even though some of the models have downsides regarding interpretation, e.g. the
-Abe-Ley model does not provide standard errors and parameters from the models 
-with a GPN distribution for the circular component are hard to interpret, 
-cylindrical models in general offer new insights into data of a cylindrical 
-nature in psychology. Concerning the particular example in this paper, the 
-teacher data that were measured using an interpersonal circumplex, we were able 
-to analyze all information in the data simultaneously. Before, the two 
-components from the interpersonal circumplex were analyzed separately. With such
-an approach we get information about the strength of a teachers' score on the 
-two axes of interpersonal behavior, in our case Agency and Communion. This 
-however does not allow us to distinguish between different types of behavior 
-that are specific combinations of Agency and Communion. A solution to that would
-be to treat circumplex data as circular [@Cremers2018Assessing]. This solution 
-however, does not retain the information about the strength of the specific type
-of interpersonal behavior a teacher displays. This is possible using the 
-cylindrical models in the present paper, we can simultaneosly model the 
-information about the type (a circular variable) and strength (a linear 
-variable) of interpersonal behavior that a teacher displays and the association
-between them.
+Even though some of the models have downsides regarding interpretation,
+e.g.~the Abe-Ley model does not provide standard errors and parameters
+from the models with a GPN distribution for the circular component are
+hard to interpret, cylindrical models in general offer new insights into
+data of a cylindrical nature in psychology. Concerning the particular
+example in this paper, the teacher data that were measured using an
+interpersonal circumplex, we were able to analyze all information in the
+data simultaneously. Before, the two components from the interpersonal
+circumplex were analyzed separately. With such an approach we get
+information about the strength of a teachers' score on the two axes of
+interpersonal behavior, in our case Agency and Communion. This however
+does not allow us to distinguish between different types of behavior
+that are specific combinations of Agency and Communion. A solution to
+that would be to treat circumplex data as circular (Cremers et al.,
+2018a). This solution however, does not retain the information about the
+strength of the specific type of interpersonal behavior a teacher
+displays. This is possible using the cylindrical models in the present
+paper, we can simultaneosly model the information about the type (a
+circular variable) and strength (a linear variable) of interpersonal
+behavior that a teacher displays and the association between them.
 
 \newpage
+
 \section*{References}
-<div id="refs"></div>
+
+\hypertarget{refs}{}
+\hypertarget{ref-abe2017tractable}{}
+Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
+for cylindrical data, with applications. \emph{Econometrics and
+Statistics}, \emph{4}, 91--104.
+doi:\href{https://doi.org/10.1016/j.ecosta.2016.04.001}{10.1016/j.ecosta.2016.04.001}
+
+\hypertarget{ref-chrastil2017rotational}{}
+Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
+integration: Encoding and execution errors in angle reproduction.
+\emph{Experimental Brain Research}, \emph{235}(6), 1885--1897.
+doi:\href{https://doi.org/10.1007/s00221-017-4910-y}{10.1007/s00221-017-4910-y}
+
+\hypertarget{ref-Claessens2016side}{}
+Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
+Teachers' perceptions of teacher--student relationships} (PhD thesis).
+
+\hypertarget{ref-Cremers2018Assessing}{}
+Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
+bayesian embedding approach to circular regression models.
+\emph{Methodology}, \emph{14}(2), 69--81.
+
+\hypertarget{ref-CremersMulderKlugkist2017}{}
+Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
+interpretation of regression coefficients. \emph{British Journal of
+Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
+doi:\href{https://doi.org/10.1111/bmsp.12108}{10.1111/bmsp.12108}
+
+\hypertarget{ref-fernandez2007models}{}
+Fernández-Durán, J. (2007). Models for circular--linear and
+circular--circular data constructed from circular distributions based on
+nonnegative trigonometric sums. \emph{Biometrics}, \emph{63}(2),
+579--585.
+doi:\href{https://doi.org/10.1111/j.1541-0420.2006.00716.x}{10.1111/j.1541-0420.2006.00716.x}
+
+\hypertarget{ref-fisher1995statistical}{}
+Fisher, N. I. (1995). \emph{Statistical analysis of circular data}.
+Cambridge: Cambridge University Press.
+
+\hypertarget{ref-garcia2014test}{}
+García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
+González-Manteiga, W., \& Pereira, J. (2014). A test for
+directional-linear independence, with applications to wildfire
+orientation and size. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{28}(5), 1261--1275.
+doi:\href{https://doi.org/10.1007/s00477-013-0819-6}{10.1007/s00477-013-0819-6}
+
+\hypertarget{ref-garcia2013exploring}{}
+García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
+Exploring wind direction and so2 concentration by circular--linear
+density estimation. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{27}(5), 1055--1067.
+doi:\href{https://doi.org/10.1007/s00477-012-0642-5}{10.1007/s00477-012-0642-5}
+
+\hypertarget{ref-BDA}{}
+Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
+(2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman
+\& Hall/CRC.
+
+\hypertarget{ref-gneiting2007strictly}{}
+Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
+prediction, and estimation. \emph{Journal of the American Statistical
+Association}, \emph{102}(477), 359--378.
+doi:\href{https://doi.org/10.1198/016214506000001437}{10.1198/016214506000001437}
+
+\hypertarget{ref-gurtman2009exploring}{}
+Gurtman, M. B. (2009). Exploring personality with the interpersonal
+circumplex. \emph{Social and Personality Psychology Compass},
+\emph{3}(4), 601--619.
+doi:\href{https://doi.org/10.1111/j.1751-9004.2009.00172.x}{10.1111/j.1751-9004.2009.00172.x}
+
+\hypertarget{ref-hernandez2016general}{}
+Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
+(2016). The general projected normal distribution of arbitrary
+dimension: Modeling and bayesian inference. \emph{Bayesian Analysis},
+\emph{12}(1), 113--133.
+doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}
+
+\hypertarget{ref-horowitz2010handbook}{}
+Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
+psychology: Theory, research, assessment, and therapeutic
+interventions}. Hoboken, NJ: John Wiley \& Sons.
+
+\hypertarget{ref-jammalamadaka2001topics}{}
+Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
+statistics} (Vol. 5). World Scientific.
+
+\hypertarget{ref-johnson1978some}{}
+Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
+distributions and related regression models. \emph{Journal of the
+American Statistical Association}, \emph{73}(363), 602--606.
+
+\hypertarget{ref-kato2008dependent}{}
+Kato, S., \& Shimizu, K. (2008). Dependent models for observations which
+include angular ones. \emph{Journal of Statistical Planning and
+Inference}, \emph{138}(11), 3538--3549.
+doi:\href{https://doi.org/10.1016/j.jspi.2006.12.009}{10.1016/j.jspi.2006.12.009}
+
+\hypertarget{ref-lagona2015hidden}{}
+Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
+markov approach to the analysis of space--time environmental data with
+linear and circular components. \emph{Stochastic Environmental Research
+and Risk Assessment}, \emph{29}(2), 397--409.
+doi:\href{https://doi.org/10.1007/s00477-014-0919-y}{10.1007/s00477-014-0919-y}
+
+\hypertarget{ref-ley2017modern}{}
+Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
+CRC Press.
+
+\hypertarget{ref-mardia2000directional}{}
+Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
+(Vol. 494). Chichester, England: Wiley.
+
+\hypertarget{ref-mardia1978model}{}
+Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical
+variables with applications. \emph{Journal of the Royal Statistical
+Society. Series B (Methodological)}, \emph{40}(2), 229--233.
+
+\hypertarget{ref-mastrantonio2018joint}{}
+Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
+distribution for poly-cylindrical data. \emph{Journal of Multivariate
+Analysis}, \emph{165}, 14--26.
+doi:\href{https://doi.org/10.1016/j.jmva.2017.11.006}{10.1016/j.jmva.2017.11.006}
+
+\hypertarget{ref-mastrantonio2015bayesian}{}
+Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
+hidden markov modelling using circular-linear general projected normal
+distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
+doi:\href{https://doi.org/10.1002/env.2326}{10.1002/env.2326}
+
+\hypertarget{ref-nunez2014bayesian}{}
+Nuñez-Antonio, G., \& Gutiérrez-Peña, E. (2014). A Bayesian model for
+longitudinal circular data based on the projected normal distribution.
+\emph{Computational Statistics \& Data Analysis}, \emph{71}, 506--519.
+doi:\href{https://doi.org/10.1016/j.csda.2012.07.025}{10.1016/j.csda.2012.07.025}
+
+\hypertarget{ref-nunez2011bayesian}{}
+Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
+Bayesian regression model for circular data based on the projected
+normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
+185--201.
+doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
+
+\hypertarget{ref-olmos2012extension}{}
+Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
+extension of the half-normal distribution. \emph{Statistical Papers},
+\emph{53}(4), 875--886.
+doi:\href{https://doi.org/10.1007/s00362-011-0391-4}{10.1007/s00362-011-0391-4}
+
+\hypertarget{ref-pennings2017interpersonal}{}
+Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C., Want,
+A. C. van der, \& Tartwijk, J. van. (2017). Interpersonal adaptation in
+teacher-student interaction. \emph{Learning and Instruction}, \emph{55},
+41--57.
+doi:\href{https://doi.org/10.1016/j.learninstruc.2017.09.005}{10.1016/j.learninstruc.2017.09.005}
+
+\hypertarget{ref-rayner200935th}{}
+Rayner, K. (2009). The 35th sir frederick bartlett lecture: Eye
+movements and attention in reading, scene perception, and visual search.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8),
+1457--1506.
+doi:\href{https://doi.org/10.1080/17470210902816461}{10.1080/17470210902816461}
+
+\hypertarget{ref-sahu2003new}{}
+Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
+multivariate skew distributions with applications to bayesian regression
+models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
+doi:\href{https://doi.org/10.2307/3316064}{10.2307/3316064}
+
+\hypertarget{ref-sugasawa2015flexible}{}
+Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the
+cylinder}. Retrieved from
+\href{arXiv:\%201501.06332v2}{arXiv: 1501.06332v2}
+
+\hypertarget{ref-wang2012directional}{}
+Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
+general projected normal distribution. \emph{Statistical Methodology},
+\emph{10}(1), 113--127.
+doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}
+
+\hypertarget{ref-vanderWant2015role}{}
+Want, A. C. van der. (2015). \emph{Teachers' interpersonal role
+identity.} (PhD thesis).
+
+\hypertarget{ref-wubbels2006interpersonal}{}
+Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
+An interpersonal perspective on classroom management in secondary
+classrooms in the netherlands. In C. Evertson \& C. S. Weinstein (Eds.),
+\emph{Handbook of classroom management: Research, practice, and
+contemporary issues} (pp. 1161--1191). Malwah, NJ: Lawrence Erlbaum
+Associates.
 
 \newpage
+
 \begin{appendices}

### 2018-07-02 Manuscript/Plots/dataplot.tex
Manuscript Update

% Created by tikzDevice version 0.11 on 2018-07-02 13:49:21
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (180.78,184.51) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.74,137.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.66,196.39) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12,137.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.16,142.64) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (300.32,152.77) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.19,147.01) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.65,144.99) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.42,169.19) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.48,141.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.48,102.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.38,141.15) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.22,165.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.28,141.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.93,142.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.14,140.40) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.46,141.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.70,137.38) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.02,142.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.66,159.30) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11,146.53) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.37,148.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.75,141.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.43,147.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.15,118.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.02,141.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.52,151.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.92,158.12) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.84,144.85) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.87,141.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.27,153.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.76,150.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.88,142.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.84,140.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.54,152.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.26,140.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.52, 81.92) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.08,152.67) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,116.63) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (124.33,154.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.57,143.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.47,142.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.27,134.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (216.19,168.30) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.80,143.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.48,144.16) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.81, 67.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.72,146.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.32,126.33) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.75,145.53) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.13,144.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.64,140.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.61,143.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.81,142.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.69,150.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.60,144.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.73, 91.85) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.57,145.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.02,152.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.91,140.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.97,136.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.97,143.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.23,133.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.56,146.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.41,132.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.22,152.07) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.25,142.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.31,169.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.34,140.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.72,137.19) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.96,130.36) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.40, 97.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.24,163.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.79,189.67) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.06,142.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.51, 77.91) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.26,181.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.12,147.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.20,136.51) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.21,139.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.13,145.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.60,138.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.48,152.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.86,142.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.37,144.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.01,141.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.61,147.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (297.18,134.55) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.04,137.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 78.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.82,140.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.81,144.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.90,149.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.74,147.67) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.65,149.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.88,120.54) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.38,138.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.32,142.69) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.63,153.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.51,135.76) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.31,129.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.59,138.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.83,109.16) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.55,155.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.83,143.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.63,156.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.48,136.09) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.44,147.25) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.65,180.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.56,156.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.32,146.74) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.44,151.99) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.48,130.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.20,164.64) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.19,157.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (335.89,140.67) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.89,159.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.22,139.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.95,144.27) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.21,153.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.28,151.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.36,130.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.85,136.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.12,143.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.83,133.88) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.86,145.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.61,143.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.15,141.54) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.05,138.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.50,143.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.10,141.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.67,137.53) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.51,138.43) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.75, 61.20) -- (395.39, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.75, 61.20) -- ( 70.75, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.13, 61.20) -- (117.13, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.51, 61.20) -- (163.51, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.88, 61.20) -- (209.88, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.26, 61.20) -- (256.26, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.64, 61.20) -- (302.64, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (349.01, 61.20) -- (349.01, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.39, 61.20) -- (395.39, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 70.75, 39.60) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (117.13, 39.60) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (163.51, 39.60) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (209.88, 39.60) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (256.26, 39.60) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (302.64, 39.60) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (349.01, 39.60) {0.7};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.39, 39.60) {0.8};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 77.48) -- ( 49.20,187.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 77.48) -- ( 43.20, 77.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 95.81) -- ( 43.20, 95.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,114.14) -- ( 43.20,114.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,132.47) -- ( 43.20,132.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,150.80) -- ( 43.20,150.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,169.14) -- ( 43.20,169.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,187.47) -- ( 43.20,187.47);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 77.48) {-150};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,114.14) {-50};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,132.47) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,150.80) {50};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,187.47) {150};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {IPC strength};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {IPC location};
\end{scope}
\end{tikzpicture}

\end{document}

### 2018-07-02 Manuscript/Plots/regline.tex
Manuscript Update

% Created by tikzDevice version 0.10.1.2 on 2018-06-13 10:26:07
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,239.88);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.85,218.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,157.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,233.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,157.92) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,164.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,177.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,169.84) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,167.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,198.32) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,163.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,112.96) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,162.31) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,194.16) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,163.16) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,164.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,161.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,163.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,157.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,164.61) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,185.63) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,169.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,171.90) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,162.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,169.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.25,133.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,163.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,175.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,184.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,167.07) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,163.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,177.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,174.39) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,163.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,161.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,176.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,160.90) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76, 86.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.31,177.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.13,130.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,179.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,165.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,164.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,153.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,197.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,165.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,166.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00, 67.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,169.11) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,143.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,167.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,167.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,161.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,165.27) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,164.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,174.74) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,167.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00, 98.99) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,167.33) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,177.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,161.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,156.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,165.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,152.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,169.76) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,151.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,176.33) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,163.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,198.11) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,161.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,157.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.25,148.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,106.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,190.96) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,224.63) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,164.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52, 81.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,213.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,170.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,156.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,160.03) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,167.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,159.31) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,177.09) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,163.40) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.19,166.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,162.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,169.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,153.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,156.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52, 81.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,161.07) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,166.77) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,173.15) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,170.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,172.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,135.84) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,158.85) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,164.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,178.53) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,155.38) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,147.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,158.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,121.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,180.64) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.31,165.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,181.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,155.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,170.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,212.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,181.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,169.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.19,176.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,149.19) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,192.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,183.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,161.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,185.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,160.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,166.32) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,177.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,175.76) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,148.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,155.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,165.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,152.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,167.38) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,164.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,162.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,159.32) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,165.36) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,162.19) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,157.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,158.82) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.02, 61.20) -- (110.02, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.54, 61.20) -- (157.54, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.05, 61.20) -- (205.05, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.57, 61.20) -- (252.57, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (300.08, 61.20) -- (300.08, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.60, 61.20) -- (347.60, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.50, 39.60) {-5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (110.02, 39.60) {-4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (157.54, 39.60) {-3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (205.05, 39.60) {-2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (252.57, 39.60) {-1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (300.08, 39.60) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (347.60, 39.60) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.12, 39.60) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {Self-efficacy};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,150.54) {$\theta_{test}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 75.82) -- ( 49.20,226.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 75.82) -- ( 43.20, 75.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,113.49) -- ( 43.20,113.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,151.16) -- ( 43.20,151.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,188.84) -- ( 43.20,188.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,226.51) -- ( 43.20,226.51);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 75.82) {-160};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,113.49) {-80};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,151.16) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,188.84) {80};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,226.51) {160};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,239.88);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,163.84) --
	( 67.26,163.79) --
	( 72.01,163.73) --
	( 76.76,163.68) --
	( 81.51,163.63) --
	( 86.26,163.57) --
	( 91.01,163.52) --
	( 95.77,163.47) --
	(100.52,163.41) --
	(105.27,163.36) --
	(110.02,163.30) --
	(114.77,163.25) --
	(119.52,163.20) --
	(124.28,163.14) --
	(129.03,163.09) --
	(133.78,163.03) --
	(138.53,162.98) --
	(143.28,162.93) --
	(148.03,162.87) --
	(152.78,162.82) --
	(157.54,162.77) --
	(162.29,162.71) --
	(167.04,162.66) --
	(171.79,162.60) --
	(176.54,162.55) --
	(181.29,162.50) --
	(186.05,162.44) --
	(190.80,162.39) --
	(195.55,162.33) --
	(200.30,162.28) --
	(205.05,162.23) --
	(209.80,162.17) --
	(214.56,162.12) --
	(219.31,162.06) --
	(224.06,162.01) --
	(228.81,161.96) --
	(233.56,161.90) --
	(238.31,161.85) --
	(243.06,161.79) --
	(247.82,161.74) --
	(252.57,161.69) --
	(257.32,161.63) --
	(262.07,161.58) --
	(266.82,161.52) --
	(271.57,161.47) --
	(276.33,161.42) --
	(281.08,161.36) --
	(285.83,161.31) --
	(290.58,161.25) --
	(295.33,161.20) --
	(300.08,161.15) --
	(304.84,161.09) --
	(309.59,161.04) --
	(314.34,160.99) --
	(319.09,160.93) --
	(323.84,160.88) --
	(328.59,160.82) --
	(333.34,160.77) --
	(338.10,160.72) --
	(342.85,160.66) --
	(347.60,160.61) --
	(352.35,160.55) --
	(357.10,160.50) --
	(361.85,160.45) --
	(366.61,160.39) --
	(371.36,160.34) --
	(376.11,160.28) --
	(380.86,160.23) --
	(385.61,160.18) --
	(390.36,160.12) --
	(395.12,160.07) --
	(399.87,160.01) --
	(404.62,159.96) --
	(409.37,159.91) --
	(414.12,159.85) --
	(418.87,159.80) --
	(423.63,159.74) --
	(428.38,159.69) --
	(433.13,159.64) --
	(433.62,159.63);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50, 89.31) --
	( 67.26, 89.47) --
	( 72.01, 89.63) --
	( 76.76, 89.81) --
	( 81.51, 90.01) --
	( 86.26, 90.23) --
	( 91.01, 90.47) --
	( 95.77, 90.74) --
	(100.52, 91.03) --
	(105.27, 91.36) --
	(110.02, 91.73) --
	(114.77, 92.16) --
	(119.52, 92.64) --
	(124.28, 93.19) --
	(129.03, 93.84) --
	(133.78, 94.61) --
	(138.53, 95.52) --
	(143.28, 96.63) --
	(148.03, 98.00) --
	(152.78, 99.71) --
	(157.54,101.91) --
	(162.29,104.79) --
	(167.04,108.59) --
	(171.79,113.62) --
	(176.54,120.04) --
	(181.29,127.51) --
	(186.05,135.04) --
	(190.80,141.60) --
	(195.55,146.79) --
	(200.30,150.72) --
	(205.05,153.69) --
	(209.80,155.95) --
	(214.56,157.71) --
	(219.31,159.11) --
	(224.06,160.25) --
	(228.81,161.18) --
	(233.56,161.96) --
	(238.31,162.62) --
	(243.06,163.19) --
	(247.82,163.68) --
	(252.57,164.11) --
	(257.32,164.49) --
	(262.07,164.82) --
	(266.82,165.12) --
	(271.57,165.39) --
	(276.33,165.63) --
	(281.08,165.85) --
	(285.83,166.05) --
	(290.58,166.23) --
	(295.33,166.40) --
	(300.08,166.56) --
	(304.84,166.70) --
	(309.59,166.83) --
	(314.34,166.96) --
	(319.09,167.07) --
	(323.84,167.18) --
	(328.59,167.28) --
	(333.34,167.37) --
	(338.10,167.46) --
	(342.85,167.55) --
	(347.60,167.63) --
	(352.35,167.70) --
	(357.10,167.77) --
	(361.85,167.84) --
	(366.61,167.90) --
	(371.36,167.96) --
	(376.11,168.02) --
	(380.86,168.07) --
	(385.61,168.13) --
	(390.36,168.18) --
	(395.12,168.22) --
	(399.87,168.27) --
	(404.62,168.31) --
	(409.37,168.35) --
	(414.12,168.39) --
	(418.87,168.43) --
	(423.63,168.47) --
	(428.38,168.51) --
	(433.13,168.54) --
	(433.62,168.54);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (179.31,125.69) --
	(183.81,125.69) --
	(183.81,130.19) --
	(179.31,130.19) --
	cycle;
\end{scope}
\end{tikzpicture}

\end{document}

### 2018-07-02 Manuscript/Plots/reglinediffSE.tex
Manuscript Update

% Created by tikzDevice version 0.11 on 2018-07-02 13:49:26
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.85,186.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,137.90) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,198.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,138.36) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,143.39) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,153.77) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,147.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,145.79) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,170.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,142.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,102.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,141.86) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,167.27) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,142.54) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,143.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,142.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,138.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,143.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,160.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,147.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,149.51) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,147.91) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.25,118.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,142.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,152.55) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,159.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,145.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,142.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,154.36) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,151.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,142.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,141.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,153.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,140.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76, 81.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.31,153.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.13,116.74) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,155.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,144.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,143.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,134.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,169.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,144.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,144.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00, 66.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,147.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,126.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,146.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,145.74) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,144.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,143.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,151.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,145.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00, 91.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,145.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,153.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,141.19) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,137.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,144.55) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,134.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,147.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,133.30) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,153.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,143.20) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,170.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,140.86) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,137.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.25,130.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16, 97.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,164.72) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,191.58) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,143.31) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52, 77.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,182.92) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,148.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,137.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,140.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,145.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,139.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,153.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,142.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.19,145.28) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.88) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,147.90) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,135.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,137.61) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52, 77.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,140.87) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,145.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,150.51) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,148.54) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,150.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,120.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,139.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,143.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,154.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,136.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,129.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,138.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,109.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,156.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.31,144.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,157.43) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,136.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,148.11) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,181.84) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,157.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,147.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.19,152.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,131.40) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,165.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,158.54) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,141.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,160.74) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,140.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,145.06) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,154.03) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,152.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,130.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,136.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,144.11) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,134.41) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,145.91) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,143.97) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,142.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,139.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.64,144.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,141.76) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,138.15) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,139.08) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.02, 61.20) -- (110.02, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.54, 61.20) -- (157.54, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.05, 61.20) -- (205.05, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.57, 61.20) -- (252.57, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (300.08, 61.20) -- (300.08, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.60, 61.20) -- (347.60, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.50, 39.60) {-5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (110.02, 39.60) {-4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (157.54, 39.60) {-3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (205.05, 39.60) {-2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (252.57, 39.60) {-1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (300.08, 39.60) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (347.60, 39.60) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.12, 39.60) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {self-efficacy};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {$\theta_{test}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 72.87) -- ( 49.20,193.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 72.87) -- ( 43.20, 72.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,102.92) -- ( 43.20,102.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,132.97) -- ( 43.20,132.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,163.02) -- ( 43.20,163.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,193.08) -- ( 43.20,193.08);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 72.87) {-160};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,102.92) {-80};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,132.97) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,163.02) {80};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,193.08) {160};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,143.08) --
	( 67.26,143.04) --
	( 72.01,143.00) --
	( 76.76,142.96) --
	( 81.51,142.91) --
	( 86.26,142.87) --
	( 91.01,142.83) --
	( 95.77,142.78) --
	(100.52,142.74) --
	(105.27,142.70) --
	(110.02,142.66) --
	(114.77,142.61) --
	(119.52,142.57) --
	(124.28,142.53) --
	(129.03,142.48) --
	(133.78,142.44) --
	(138.53,142.40) --
	(143.28,142.35) --
	(148.03,142.31) --
	(152.78,142.27) --
	(157.54,142.23) --
	(162.29,142.18) --
	(167.04,142.14) --
	(171.79,142.10) --
	(176.54,142.05) --
	(181.29,142.01) --
	(186.05,141.97) --
	(190.80,141.92) --
	(195.55,141.88) --
	(200.30,141.84) --
	(205.05,141.80) --
	(209.80,141.75) --
	(214.56,141.71) --
	(219.31,141.67) --
	(224.06,141.62) --
	(228.81,141.58) --
	(233.56,141.54) --
	(238.31,141.49) --
	(243.06,141.45) --
	(247.82,141.41) --
	(252.57,141.36) --
	(257.32,141.32) --
	(262.07,141.28) --
	(266.82,141.24) --
	(271.57,141.19) --
	(276.33,141.15) --
	(281.08,141.11) --
	(285.83,141.06) --
	(290.58,141.02) --
	(295.33,140.98) --
	(300.08,140.93) --
	(304.84,140.89) --
	(309.59,140.85) --
	(314.34,140.81) --
	(319.09,140.76) --
	(323.84,140.72) --
	(328.59,140.68) --
	(333.34,140.63) --
	(338.10,140.59) --
	(342.85,140.55) --
	(347.60,140.50) --
	(352.35,140.46) --
	(357.10,140.42) --
	(361.85,140.37) --
	(366.61,140.33) --
	(371.36,140.29) --
	(376.11,140.25) --
	(380.86,140.20) --
	(385.61,140.16) --
	(390.36,140.12) --
	(395.12,140.07) --
	(399.87,140.03) --
	(404.62,139.99) --
	(409.37,139.94) --
	(414.12,139.90) --
	(418.87,139.86) --
	(423.63,139.82) --
	(428.38,139.77) --
	(433.13,139.73) --
	(433.62,139.72);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50, 83.63) --
	( 67.26, 83.75) --
	( 72.01, 83.88) --
	( 76.76, 84.03) --
	( 81.51, 84.19) --
	( 86.26, 84.36) --
	( 91.01, 84.55) --
	( 95.77, 84.76) --
	(100.52, 85.00) --
	(105.27, 85.26) --
	(110.02, 85.56) --
	(114.77, 85.89) --
	(119.52, 86.28) --
	(124.28, 86.72) --
	(129.03, 87.24) --
	(133.78, 87.85) --
	(138.53, 88.58) --
	(143.28, 89.47) --
	(148.03, 90.56) --
	(152.78, 91.93) --
	(157.54, 93.68) --
	(162.29, 95.97) --
	(167.04, 99.01) --
	(171.79,103.02) --
	(176.54,108.14) --
	(181.29,114.10) --
	(186.05,120.11) --
	(190.80,125.34) --
	(195.55,129.48) --
	(200.30,132.62) --
	(205.05,134.98) --
	(209.80,136.79) --
	(214.56,138.19) --
	(219.31,139.31) --
	(224.06,140.22) --
	(228.81,140.96) --
	(233.56,141.58) --
	(238.31,142.11) --
	(243.06,142.56) --
	(247.82,142.95) --
	(252.57,143.30) --
	(257.32,143.60) --
	(262.07,143.86) --
	(266.82,144.10) --
	(271.57,144.32) --
	(276.33,144.51) --
	(281.08,144.69) --
	(285.83,144.85) --
	(290.58,144.99) --
	(295.33,145.13) --
	(300.08,145.25) --
	(304.84,145.36) --
	(309.59,145.47) --
	(314.34,145.57) --
	(319.09,145.66) --
	(323.84,145.75) --
	(328.59,145.83) --
	(333.34,145.90) --
	(338.10,145.97) --
	(342.85,146.04) --
	(347.60,146.10) --
	(352.35,146.16) --
	(357.10,146.22) --
	(361.85,146.27) --
	(366.61,146.32) --
	(371.36,146.37) --
	(376.11,146.42) --
	(380.86,146.46) --
	(385.61,146.50) --
	(390.36,146.54) --
	(395.12,146.58) --
	(399.87,146.62) --
	(404.62,146.65) --
	(409.37,146.68) --
	(414.12,146.72) --
	(418.87,146.75) --
	(423.63,146.78) --
	(428.38,146.80) --
	(433.13,146.83) --
	(433.62,146.83);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (179.31,112.20) --
	(183.81,112.20) --
	(183.81,116.70) --
	(179.31,116.70) --
	cycle;
\end{scope}
\end{tikzpicture}

\end{document}

### 2018-07-02 Manuscript/Plots/survivaldiffSE.tex
Manuscript Update

% Created by tikzDevice version 0.11 on 2018-07-02 13:49:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.43) --
	( 95.77,198.15) --
	(112.40,197.32) --
	(129.03,195.59) --
	(145.66,192.64) --
	(162.29,188.17) --
	(178.92,181.98) --
	(195.55,173.98) --
	(212.18,164.25) --
	(228.81,153.04) --
	(245.44,140.78) --
	(262.07,128.04) --
	(278.70,115.47) --
	(295.33,103.70) --
	(311.96, 93.25) --
	(328.59, 84.49) --
	(345.22, 77.56) --
	(361.85, 72.41) --
	(378.49, 68.82) --
	(395.12, 66.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.03, 61.20) -- (129.03, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.55, 61.20) -- (195.55, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.07, 61.20) -- (262.07, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.59, 61.20) -- (328.59, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.50, 39.60) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.03, 39.60) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (195.55, 39.60) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (262.07, 39.60) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (328.59, 39.60) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.12, 39.60) {1.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 49.20,198.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 43.20, 63.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 90.50) -- ( 43.20, 90.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,117.49) -- ( 43.20,117.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,144.48) -- ( 43.20,144.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,171.47) -- ( 43.20,171.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,198.47) -- ( 43.20,198.47);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 63.51) {0.0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 90.50) {0.2};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,117.49) {0.4};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,144.48) {0.6};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,171.47) {0.8};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,198.47) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {strength IPC};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {P(strength IPC)};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.44) --
	( 95.77,198.21) --
	(112.40,197.48) --
	(129.03,195.93) --
	(145.66,193.23) --
	(162.29,189.08) --
	(178.92,183.22) --
	(195.55,175.55) --
	(212.18,166.08) --
	(228.81,155.03) --
	(245.44,142.81) --
	(262.07,129.97) --
	(278.70,117.17) --
	(295.33,105.08) --
	(311.96, 94.28) --
	(328.59, 85.19) --
	(345.22, 77.97) --
	(361.85, 72.61) --
	(378.49, 68.89) --
	(395.12, 66.48);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.44) --
	( 95.77,198.23) --
	(112.40,197.56) --
	(129.03,196.11) --
	(145.66,193.55) --
	(162.29,189.57) --
	(178.92,183.91) --
	(195.55,176.42) --
	(212.18,167.10) --
	(228.81,156.16) --
	(245.44,143.97) --
	(262.07,131.08) --
	(278.70,118.15) --
	(295.33,105.89) --
	(311.96, 94.89) --
	(328.59, 85.60) --
	(345.22, 78.22) --
	(361.85, 72.73) --
	(378.49, 68.93) --
	(395.12, 66.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.33,186.47) -- (322.33,186.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (304.33,174.47) -- (322.33,174.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] (304.33,162.47) -- (322.33,162.47);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,183.02) {min(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {mean(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,159.02) {max(SE)};
\end{scope}
\end{tikzpicture}

\end{document}

### 2018-07-02 Manuscript/R-code/Abe-Ley optimization.R

@@ -1,0 +1,69 @@
+func.regII <- function(par, data){
+  
+  eta <- par[1:2]
+  nu <- par[3:4]
+  kappa <- par[5]
+  alpha <- par[6]
+  lambda <- par[7]
+  
+  y <- data$y
+  theta <- data$theta
+  x <- as.matrix(data[, c("ax", "bx")])
+  z <- as.matrix(data[, c("az", "bz")])
+  n <- length(data$y)
+  
+  p1 <- n*(log(alpha) - log(2*pi*cosh(kappa)))
+  p2 <- alpha * sum(log(exp(x%*%nu)))
+  p3 <- sum(log(1 + lambda*sin(theta - eta[1] + 2*atan(z[,2]*eta[2]))))
+  p4 <- (alpha - 1) * sum(log(y))
+  p5 <- sum((exp(x%*%nu)*y)^alpha*(1 - tanh(kappa)*cos(theta - eta[1] +  2*atan(z[,2]*eta[2]))))
+  
+  
+  p1+p2+p3+p4-p5
+  
+  
+}
+
+func.reg.cond.lin <- function(par, data){
+  
+  
+  eta <- par[1:2]
+  nu <- par[3:4]
+  kappa <- par[5]
+  alpha <- par[6]
+  lambda <- par[7]
+  
+  y <- data$y
+  theta <- data$theta
+  x <- as.matrix(data[, c("ax", "bx")])
+  z <- as.matrix(data[, c("az", "bz")])
+  n <- length(data$y)
+  
+  shape <- exp(x%*%nu)*(1-tanh(kappa)*cos(theta - (eta[1] +  2*atan(z[,2]*eta[2]))))^(1/alpha)
+  
+  log(alpha) + sum(log(shape^alpha)) + sum(log(y^(alpha - 1))) - sum((shape*y)^(alpha))
+
+}
+
+func.reg.cond.circ <- function(par, data){
+  
+  
+  eta <- par[1:2]
+  nu <- par[3:4]
+  kappa <- par[5]
+  alpha <- par[6]
+  lambda <- par[7]
+  
+  y <- data$y
+  theta <- data$theta
+  x <- as.matrix(data[, c("ax", "bx")])
+  z <- as.matrix(data[, c("az", "bz")])
+  n <- length(data$y)
+  
+  conc <- y^{alpha}*exp(x%*%nu)^{alpha}*tanh(kappa)
[…]

### 2018-07-02 Manuscript/R-code/Posterior Sampling CL-GPN.R

@@ -1,0 +1,301 @@
+#Functions for posterior distributions
+
+#Sampling Gamma and Sigma, the regression coefficients and residual variance of 
+#the regression on the linear outcome y.
+
+#N = sample size
+#X = design matrix
+#y = linear outcome
+#sig = residual variance (starting value or value from previous iteration)
+#p.gamma.mu = prior mean vector for the regression coefficients
+#p.gamma.prec = prior precision for the regression coefficients
+#p.sig.shape = prior shape parameter for the residual variance
+#p.sig.scale = prior scale parameter for the residual variance
+
+post.gamma.sig <- function(N, X, y, sig, p.gamma.mu, p.gamma.prec, p.sig.shape, p.sig.scale){
+  
+  prec.post <- t(X)%*%X + p.gamma.prec
+  var.post <- solve(prec.post)
+  mu.post <- solve(prec.post) %*% ((p.gamma.prec%*%p.gamma.mu) + (t(X)%*%y))
+  
+  gamma <- t(mvrnorm(1, mu.post, sig*var.post))
+  
+  shape.post <- N/2 + p.sig.shape
+  scale.post <- 0.5*t(y-X%*%t(gamma))%*%(y-X%*%t(gamma)) + p.sig.scale
+  
+  sig <- 1/rgamma(1, shape.post , scale.post)
+  
+  return(list("gamma" = gamma, "sig" = sig) )
+  
+}
+
+#Sampling rho, element of sigma, the variance-covariance matrix of the circular 
+#outcome
+
+post.rho <- function(Y, Z, Beta.I, Beta.II, tau, p.rho.mu, p.rho.var){
+  
+  num <- tau^(-1) * t(Y[,1] - Z%*%Beta.I)%*%(Y[,2] - Z%*%Beta.II) + p.rho.mu*p.rho.var^(-1)
+  den <- tau^(-1) * t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II) + p.rho.var^(-1)
+  
+  mu.rho.post <- num/den
+  var.rho.post <- 1/(tau^(-1)*(t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II)) + p.rho.var^(-1))
+  
+  rho <- rnorm(1, mu.rho.post, var.rho.post)
+  
+}
+
[…]

### 2018-07-02 Manuscript/R-code/Posterior Sampling CL-PN.R

@@ -1,0 +1,220 @@
+#Functions for posterior distributions
+
+#Sampling Gamma and Sigma, the regression coefficients and residual variance of 
+#the regression on the linear outcome y.
+
+#N = sample size
+#X = design matrix
+#y = linear outcome
+#sigma = residual variance (starting value or value from previous iteration)
+#p.gamma.mu = prior mean vector for the regression coefficients
+#p.gamma.prec = prior precision for the regression coefficients
+#p.sigma.shape = prior shape parameter for the residual variance
+#p.sigma.scale = prior scale parameter for the residual variance
+
+post.gamma.sigma <- function(N, X, y, sigma, p.gamma.mu, p.gamma.prec, p.sigma.shape, p.sigma.scale){
+  
+  prec.post <- t(X)%*%X + p.gamma.prec
+  var.post <- solve(prec.post)
+  mu.post <- solve(prec.post) %*% ((p.gamma.prec%*%p.gamma.mu) + (t(X)%*%y))
+  
+  gamma <- t(mvrnorm(1, mu.post, sigma*var.post))
+  
+  shape.post <- N/2 + p.sigma.shape
+  scale.post <- 0.5*t(y-X%*%t(gamma))%*%(y-X%*%t(gamma)) + p.sigma.scale
+  
+  sigma <- 1/rgamma(1, shape.post , scale.post)
+
+  
+  return(list("gamma" = gamma, "sigma" = sigma) )
+  
+}
+
+#Sampling Beta, the regression coefficients of one component (I, II) of the 
+#regression on the circular outcome theta.
+
+#Z = design matrix
+#Y = (cos(theta), sin(theta))*r
+#p.beta.mu = prior mean for the regression coefficients
+#p.beta.prec = prior precision for the regression coefficients
+
+
+post.beta <- function(Z, Y, p.beta.mu, p.beta.prec){
+  
+  mu.post <- solve(t(Z)%*%Z + p.beta.prec) %*% (p.beta.prec%*%p.beta.mu + t(Z)%*%Y)
+  var.post <- solve(t(Z)%*%Z + p.beta.prec)
+  
+  beta <- mvrnorm(1, mu.post, var.post)
+  
+}
+
[…]

### 2018-07-02 Manuscript/R-code/Posterior Sampling Joint GPN-SSN.R

@@ -1,0 +1,357 @@
+require(MCMCpack)
+require(truncnorm)
+#require(LaplacesDemon)
+require(bayesm)
+require(mvtnorm)
+
+#Constraint matrix function (for 1 circular and 1 linear variable)
+
+cmat <- function(var){diag(3)*(c(sqrt(var), sqrt(var), 1))}
+
+#sample from matrix normal (adapted from LaplacesDemon, function is.symmetric.matrix does not work properly; floating point problems...)
+
+rmatrixnorm <- function (M, U, V) 
+{
+  if (missing(M)) 
+    stop("Matrix M is missing.")
+  if (!is.matrix(M)) 
+    M <- matrix(M)
+  if (missing(U)) 
+    stop("Matrix U is missing.")
+  if (!is.matrix(U)) 
+    U <- matrix(U)
+    if (missing(V)) 
+    stop("Matrix V is missing.")
+  if (!is.matrix(V)) 
+    V <- matrix(V)
+  if (nrow(M) != nrow(U)) 
+    stop("Dimensions of M and U are incorrect.")
+  if (ncol(M) != ncol(V)) 
+    stop("Dimensions of M and V are incorrect.")
+  n <- nrow(U)
+  k <- ncol(V)
+  Z <- matrix(rnorm(n * k), n, k)
+  X <- M + t(chol(U)) %*% Z %*% chol(V)
+  return(X)
+}
+
+#Functions for posterior distributions
+
+#Sample Beta, a matrix of regression coefficients for the circular and linear 
+#variables
+
+#Y = the circular and linear outcomes
+#X = a design matrix
+#Beta = Beta matrix (previous iteration/starting value)
+#Kappa = Kappa matrix (previous iteration/starting value)
+#Sigma  = Sigma matrix (previous iteration/starting value)
+#p.Beta = 
+#p.Kappa = 
+
+post.Beta <- function(Y, X, Beta, Sigma, p.Beta, p.Kappa){
+
+Kappa.post <-  t(X)%*%X + p.Kappa 
[…]

### 2018-07-02 Manuscript/R-code/Sample Abe-Ley.R

@@ -1,0 +1,47 @@
+
+rwcauchy <- function(n, mu, rho){
+  
+  if (rho == 0) 
+    result <- runif(n, 0, 2 * pi)
+  else if (rho == 1) 
+    result <- rep(mu, n)
+  else {
+    scale <- -log(rho)
+    result <- rcauchy(n, mu, scale)%%(2 * pi)
+  }
+  return(result)
+  
+}
+
+rweiSSVM <- function(n, beta = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
+  
+  theta <- c()
+  x <- c()
+  
+  for(i in 1:n){
+    
+    u <- runif(1,0,1)
+    
+    theta.1 <- rwcauchy(1, mu, tanh(kappa/2))
+    
+    if(u < (1 + lambda*sin(theta.1 - mu))/2){
+      theta[i] <- theta.1 
+    }else{
+      theta[i] <- -theta.1
+    }
+    
+    shape <- beta * (1-tanh(kappa)*cos(theta-mu))^(1/alpha)
+    scale <- (1/shape)^(1/alpha) #(compute scale parameter for different parametrization)
+    
+    
+    x[i] <- rweibull(1, shape = alpha, scale = scale)
+    
+  }
+  
+  out <- cbind(theta %% (2*pi), x)
+  
+  colnames(out) <- c("theta", "x")
+  
+  return(as.data.frame(out))
+  
+}

### 2018-09-24 Manuscript/R-code/Posterior Sampling CL-GPN.R
Update Rcode

@@ -209,6 +209,8 @@ CLGPN <- function(theta, y, X, Z, its, p, theta.hold, y.hold, X.hold, Z.hold){
   theta_pred_min <- rep(NA, its)
   theta_pred_max <- rep(NA, its)
   theta_pred_mean <- rep(NA, its)
+  theta_pred_median <- rep(NA, its)
+  
 
   ll.lin <- rep(NA, its)
   ll.circ <- rep(NA, its)
@@ -252,6 +254,9 @@ CLGPN <- function(theta, y, X, Z, its, p, theta.hold, y.hold, X.hold, Z.hold){
 
     Y_pred <- mvrnorm(1, c(c(1, mean(Z[,2]))%*%Beta.I,  c(1, mean(Z[,2]))%*%Beta.II), sigma)
     theta_pred_mean[i] <- atan2(Y_pred[2], Y_pred[1])
+    
+    Y_pred <- mvrnorm(1, c(c(1, median(Z[,2]))%*%Beta.I,  c(1, median(Z[,2]))%*%Beta.II), sigma)
+    theta_pred_median[i] <- atan2(Y_pred[2], Y_pred[1])
 
     #compute predicted values
     for(j in 1:N){
@@ -295,7 +300,7 @@ CLGPN <- function(theta, y, X, Z, its, p, theta.hold, y.hold, X.hold, Z.hold){
               "theta_pred" = theta_pred, "y_pred" = y_pred,
               "theta_pred.hold" = theta_pred.hold, "y_pred.hold" = y_pred.hold,
               "theta_pred.min" = theta_pred_min,  "theta_pred.max" = theta_pred_max,
-              "theta_pred.mean" = theta_pred_mean))
+              "theta_pred.mean" = theta_pred_mean, "theta_pred.meadian" = theta_pred_median))
   
 }
 

### 2018-09-24 Manuscript/R-code/Posterior Sampling Joint GPN-SSN.R

@@ -278,6 +278,7 @@ JGPNSSN <- function(theta, y, X, its, p, q, theta.hold, y.hold, X.hold){
   theta_pred_min <- rep(NA, its)
   theta_pred_max <- rep(NA, its)
   theta_pred_mean <- rep(NA, its)
+  theta_pred_median <- rep(NA, its)
   
   ll.lin <- rep(NA, its)
   ll.circ <- rep(NA, its)
@@ -317,6 +318,9 @@ JGPNSSN <- function(theta, y, X, its, p, q, theta.hold, y.hold, X.hold){
     Y_pred <- mvrnorm(1, c(1, max(X[,2]))%*%Betacon.res[,,it]+c(0,0,lambda*D[1,]), Sigmacon.res[,,it])
     theta_pred_max[it] <- atan2(Y_pred[2], Y_pred[1])
     
+    Y_pred <- mvrnorm(1, c(1, median(X[,2]))%*%Betacon.res[,,it]+c(0,0,lambda*D[1,]), Sigmacon.res[,,it])
+    theta_pred_median[it] <- atan2(Y_pred[2], Y_pred[1])
+    
     #compute predicted values
     for(j in 1:N){
       Y_pred <- mvrnorm(1, X[j,]%*%Betacon.res[,,it]+c(0,0,lambda*D[j,]), Sigmacon.res[,,it])
@@ -350,7 +354,7 @@ JGPNSSN <- function(theta, y, X, its, p, q, theta.hold, y.hold, X.hold){
               "theta_pred" = theta_pred, "y_pred" = y_pred,
               "theta_pred.hold" = theta_pred.hold, "y_pred.hold" = y_pred.hold,
               "theta_pred.min" = theta_pred_min,  "theta_pred.max" = theta_pred_max,
-              "theta_pred.mean" = theta_pred_mean))
+              "theta_pred.mean" = theta_pred_mean, "theta_pred.median" = theta_pred_median))
   
 }
 

### 2018-09-24 Manuscript/R-code/Sample Abe-Ley.R

@@ -1,3 +1,4 @@
+#simulate from wrapped cauchy distribution
 
 rwcauchy <- function(n, mu, rho){
   
@@ -13,11 +14,13 @@ rwcauchy <- function(n, mu, rho){
   
 }
 
+#simulate from abe-ley distribution
 rweiSSVM <- function(n, beta = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
   
   theta <- c()
   x <- c()
   
+  #simulate theta
   for(i in 1:n){
     
     u <- runif(1,0,1)
@@ -31,8 +34,11 @@ rweiSSVM <- function(n, beta = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
     }
     
     shape <- beta * (1-tanh(kappa)*cos(theta-mu))^(1/alpha)
-    scale <- (1/shape)^(1/alpha) #(compute scale parameter for different parametrization)
-    
+    #compute scale parameter for different parametrization,
+    #shape = (scale_parWeibull in R)^(-alpha) = 1/(scale_parWeibull in R)^alpha dus 
+    #scale_parWeibull in R = kth root(1/shape) dus
+    #scale_parWeibull in R = (1/b)^(1/alpha)
+    scale <- (1/shape)^(1/alpha) 
     
     x[i] <- rweibull(1, shape = alpha, scale = scale)
     

### 2018-09-24 Manuscript/Plots/dataplot.tex
Plots update

@@
-% Created by tikzDevice version 0.11 on 2018-07-02 13:49:21
+% Created by tikzDevice version 0.11 on 2018-09-21 15:13:32
 % !TEX encoding = UTF-8 Unicode
@@
 \definecolor{drawColor}{RGB}{0,0,0}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.24,162.19) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.15,153.45) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.49,124.69) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.79,142.91) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.31,176.35) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.78,196.40) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.82,144.06) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.05,133.33) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.37,133.13) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.47,146.33) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.79,150.22) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.75,146.76) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.49,140.22) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.42,144.38) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.70,144.86) circle (  2.25);
 

### 2018-09-24 Manuscript/Plots/reglinediffSE.tex
Plots update

@@
-% Created by tikzDevice version 0.11 on 2018-07-02 13:49:26
+% Created by tikzDevice version 0.11 on 2018-08-10 10:45:22
 % !TEX encoding = UTF-8 Unicode
@@
 
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,163.42) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,154.46) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.13,124.99) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,143.66) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.88,177.93) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.76,198.47) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,144.84) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,133.85) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.52,133.65) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,147.17) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,151.15) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,147.61) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,140.91) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,145.17) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.79,145.66) circle (  2.25);
+
 \path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.85,186.29) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,137.90) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.67,137.89) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,198.47) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,198.46) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,142.46) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.40,142.45) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.10) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.09) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,138.00) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,137.99) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.81) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.80) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,154.36) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.92,154.35) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.28) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.43,141.27) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,137.58) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,137.57) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,164.72) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,164.71) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,147.90) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,147.89) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,136.34) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.16,136.33) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,157.43) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,157.42) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,147.59) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.28,147.58) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,140.02) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.04,140.01) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,154.03) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.55,154.02) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,136.81) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.00,136.80) circle (  2.25);
 
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {$\theta_{test}$};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {$\theta$};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 72.87) -- ( 49.20,193.08);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 72.87) -- ( 49.20,193.07);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,193.08) -- ( 43.20,193.08);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,193.07) -- ( 43.20,193.07);
 
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,193.08) {160};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,193.07) {160};
 \end{scope}
@@
 	(105.27,142.70) --
-	(110.02,142.66) --
+	(110.02,142.65) --
 	(114.77,142.61) --
@@
 	(152.78,142.27) --
-	(157.54,142.23) --
+	(157.54,142.22) --
 	(162.29,142.18) --
@@
 	(200.30,141.84) --
-	(205.05,141.80) --
+	(205.05,141.79) --
 	(209.80,141.75) --
@@
 	(262.07,141.28) --
-	(266.82,141.24) --
+	(266.82,141.23) --
 	(271.57,141.19) --
@@
 	(309.59,140.85) --
-	(314.34,140.81) --
+	(314.34,140.80) --
 	(319.09,140.76) --
@@
 	(371.36,140.29) --
-	(376.11,140.25) --
+	(376.11,140.24) --
 	(380.86,140.20) --
@@
 	(418.87,139.86) --
-	(423.63,139.82) --
+	(423.63,139.81) --
 	(428.38,139.77) --
@@
 
-\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50, 83.63) --
-	( 67.26, 83.75) --
-	( 72.01, 83.88) --
-	( 76.76, 84.03) --
-	( 81.51, 84.19) --
-	( 86.26, 84.36) --
-	( 91.01, 84.55) --
-	( 95.77, 84.76) --
-	(100.52, 85.00) --
-	(105.27, 85.26) --
-	(110.02, 85.56) --
-	(114.77, 85.89) --
-	(119.52, 86.28) --
-	(124.28, 86.72) --
-	(129.03, 87.24) --
-	(133.78, 87.85) --
-	(138.53, 88.58) --
-	(143.28, 89.47) --
-	(148.03, 90.56) --
-	(152.78, 91.93) --
-	(157.54, 93.68) --
-	(162.29, 95.97) --
-	(167.04, 99.01) --
-	(171.79,103.02) --
-	(176.54,108.14) --
-	(181.29,114.10) --
-	(186.05,120.11) --
-	(190.80,125.34) --
-	(195.55,129.48) --
-	(200.30,132.62) --
-	(205.05,134.98) --
-	(209.80,136.79) --
-	(214.56,138.19) --
-	(219.31,139.31) --
-	(224.06,140.22) --
-	(228.81,140.96) --
-	(233.56,141.58) --
-	(238.31,142.11) --
-	(243.06,142.56) --
-	(247.82,142.95) --
-	(252.57,143.30) --
-	(257.32,143.60) --
-	(262.07,143.86) --
-	(266.82,144.10) --
-	(271.57,144.32) --
-	(276.33,144.51) --
-	(281.08,144.69) --
-	(285.83,144.85) --
-	(290.58,144.99) --
-	(295.33,145.13) --
-	(300.08,145.25) --
-	(304.84,145.36) --
-	(309.59,145.47) --
-	(314.34,145.57) --
-	(319.09,145.66) --
-	(323.84,145.75) --
-	(328.59,145.83) --
-	(333.34,145.90) --
-	(338.10,145.97) --
-	(342.85,146.04) --
-	(347.60,146.10) --
-	(352.35,146.16) --
-	(357.10,146.22) --
-	(361.85,146.27) --
-	(366.61,146.32) --
-	(371.36,146.37) --
-	(376.11,146.42) --
-	(380.86,146.46) --
-	(385.61,146.50) --
-	(390.36,146.54) --
-	(395.12,146.58) --
-	(399.87,146.62) --
-	(404.62,146.65) --
-	(409.37,146.68) --
-	(414.12,146.72) --
-	(418.87,146.75) --
-	(423.63,146.78) --
-	(428.38,146.80) --
-	(433.13,146.83) --
-	(433.62,146.83);
+\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50, 84.15) --
+	( 67.26, 84.27) --
+	( 72.01, 84.40) --
+	( 76.76, 84.54) --
+	( 81.51, 84.69) --
+	( 86.26, 84.85) --
+	( 91.01, 85.03) --
+	( 95.77, 85.23) --
+	(100.52, 85.45) --
+	(105.27, 85.69) --
+	(110.02, 85.96) --
+	(114.77, 86.27) --
+	(119.52, 86.62) --
+	(124.28, 87.01) --
+	(129.03, 87.47) --
+	(133.78, 87.99) --
+	(138.53, 88.61) --
+	(143.28, 89.35) --
+	(148.03, 90.24) --
+	(152.78, 91.33) --
+	(157.54, 92.68) --
+	(162.29, 94.40) --
+	(167.04, 96.61) --
+	(171.79, 99.50) --
+	(176.54,103.26) --
+	(181.29,108.01) --
+	(186.05,113.56) --
+	(190.80,119.29) --
+	(195.55,124.47) --
+	(200.30,128.70) --
+	(205.05,131.98) --
+	(209.80,134.49) --
+	(214.56,136.43) --
+	(219.31,137.94) --
+	(224.06,139.14) --
+	(228.81,140.12) --
+	(233.56,140.92) --
+	(238.31,141.59) --
+	(243.06,142.16) --
+	(247.82,142.65) --
+	(252.57,143.07) --
+	(257.32,143.44) --
+	(262.07,143.77) --
+	(266.82,144.05) --
+	(271.57,144.31) --
+	(276.33,144.54) --
+	(281.08,144.75) --
+	(285.83,144.94) --
+	(290.58,145.11) --
+	(295.33,145.26) --
+	(300.08,145.41) --
+	(304.84,145.54) --
+	(309.59,145.66) --
+	(314.34,145.78) --
+	(319.09,145.88) --
+	(323.84,145.98) --
+	(328.59,146.07) --
+	(333.34,146.16) --
+	(338.10,146.24) --
+	(342.85,146.32) --
+	(347.60,146.39) --
+	(352.35,146.45) --
+	(357.10,146.52) --
+	(361.85,146.58) --
+	(366.61,146.63) --
+	(371.36,146.69) --
+	(376.11,146.74) --
+	(380.86,146.79) --
+	(385.61,146.84) --
+	(390.36,146.88) --
+	(395.12,146.92) --
+	(399.87,146.96) --
+	(404.62,147.00) --
+	(409.37,147.04) --
+	(414.12,147.07) --
+	(418.87,147.11) --
+	(423.63,147.14) --
+	(428.38,147.17) --
+	(433.13,147.20) --
+	(433.62,147.21);
 \definecolor{fillColor}{RGB}{0,0,0}
 
-\path[fill=fillColor] (179.31,112.20) --
-	(183.81,112.20) --
-	(183.81,116.70) --
-	(179.31,116.70) --
+\path[fill=fillColor] (184.93,112.69) --
+	(189.43,112.69) --
+	(189.43,117.19) --
+	(184.93,117.19) --
 	cycle;

### 2018-09-24 Manuscript/Plots/survivaldiffSE.tex
Plots update

@@
-% Created by tikzDevice version 0.11 on 2018-07-02 13:49:30
+% Created by tikzDevice version 0.11 on 2018-09-18 14:43:22
 % !TEX encoding = UTF-8 Unicode
@@
 \path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,198.43) --
-	( 95.77,198.15) --
-	(112.40,197.32) --
-	(129.03,195.59) --
-	(145.66,192.64) --
-	(162.29,188.17) --
-	(178.92,181.98) --
-	(195.55,173.98) --
-	(212.18,164.25) --
-	(228.81,153.04) --
-	(245.44,140.78) --
-	(262.07,128.04) --
-	(278.70,115.47) --
-	(295.33,103.70) --
-	(311.96, 93.25) --
-	(328.59, 84.49) --
-	(345.22, 77.56) --
-	(361.85, 72.41) --
-	(378.49, 68.82) --
+	( 79.14,198.38) --
+	( 95.77,197.84) --
+	(112.40,196.49) --
+	(129.03,194.01) --
+	(145.66,190.14) --
+	(162.29,184.70) --
+	(178.92,177.62) --
+	(195.55,168.93) --
+	(212.18,158.83) --
+	(228.81,147.63) --
+	(245.44,135.76) --
+	(262.07,123.72) --
+	(278.70,112.05) --
+	(295.33,101.23) --
+	(311.96, 91.68) --
+	(328.59, 83.63) --
+	(345.22, 77.20) --
+	(361.85, 72.33) --
+	(378.49, 68.84) --
 	(395.12, 66.48);
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 49.20,198.47);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 62.99) -- ( 49.20,198.47);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 43.20, 63.51);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 62.99) -- ( 43.20, 62.99);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 90.50) -- ( 43.20, 90.50);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 90.08) -- ( 43.20, 90.08);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,117.49) -- ( 43.20,117.49);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,117.18) -- ( 43.20,117.18);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,144.48) -- ( 43.20,144.48);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,144.27) -- ( 43.20,144.27);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,171.47) -- ( 43.20,171.47);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,171.37) -- ( 43.20,171.37);
 
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 63.51) {0.0};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 62.99) {0.0};
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 90.50) {0.2};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 90.08) {0.2};
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,117.49) {0.4};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,117.18) {0.4};
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,144.48) {0.6};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,144.27) {0.6};
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,171.47) {0.8};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,171.37) {0.8};
 
@@
 \path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,198.44) --
-	( 95.77,198.21) --
-	(112.40,197.48) --
-	(129.03,195.93) --
-	(145.66,193.23) --
-	(162.29,189.08) --
-	(178.92,183.22) --
-	(195.55,175.55) --
-	(212.18,166.08) --
-	(228.81,155.03) --
-	(245.44,142.81) --
-	(262.07,129.97) --
-	(278.70,117.17) --
-	(295.33,105.08) --
-	(311.96, 94.28) --
-	(328.59, 85.19) --
-	(345.22, 77.97) --
-	(361.85, 72.61) --
-	(378.49, 68.89) --
+	( 79.14,196.04) --
+	( 95.77,190.46) --
+	(112.40,182.63) --
+	(129.03,173.19) --
+	(145.66,162.73) --
+	(162.29,151.74) --
+	(178.92,140.68) --
+	(195.55,129.91) --
+	(212.18,119.75) --
+	(228.81,110.39) --
+	(245.44,101.97) --
+	(262.07, 94.57) --
+	(278.70, 88.20) --
+	(295.33, 82.82) --
+	(311.96, 78.36) --
+	(328.59, 74.74) --
+	(345.22, 71.84) --
+	(361.85, 69.57) --
+	(378.49, 67.81) --
 	(395.12, 66.48);
@@
 \path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,198.44) --
-	( 95.77,198.23) --
-	(112.40,197.56) --
-	(129.03,196.11) --
-	(145.66,193.55) --
-	(162.29,189.57) --
-	(178.92,183.91) --
-	(195.55,176.42) --
-	(212.18,167.10) --
-	(228.81,156.16) --
-	(245.44,143.97) --
-	(262.07,131.08) --
-	(278.70,118.15) --
-	(295.33,105.89) --
-	(311.96, 94.89) --
-	(328.59, 85.60) --
-	(345.22, 78.22) --
-	(361.85, 72.73) --
-	(378.49, 68.93) --
-	(395.12, 66.48);
+	( 79.14,198.08) --
+	( 95.77,196.53) --
+	(112.40,193.46) --
+	(129.03,188.72) --
+	(145.66,182.27) --
+	(162.29,174.23) --
+	(178.92,164.82) --
+	(195.55,154.36) --
+	(212.18,143.26) --
+	(228.81,131.93) --
+	(245.44,120.83) --
+	(262.07,110.32) --
+	(278.70,100.74) --
+	(295.33, 92.29) --
+	(311.96, 85.12) --
+	(328.59, 79.23) --
+	(345.22, 74.56) --
+	(361.85, 70.98) --
+	(378.49, 68.35) --
+	(395.12, 66.47);
 
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {mean(SE)};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {median(SE)};
 

### 2018-09-24 Manuscript/SupplementaryMaterial.Rmd
Supplementary material

---
title: " "
author: |
bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
  - \setlength\parindent{24pt}
  - \usepackage{setspace}\doublespacing
documentclass: article
pandoc_args: --natbib
---

\section{Supplementary Material}\label{Appendix}

In this Supplementary Material we outline the MCMC procedures to fit the cylindrical regression models. R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}. Note that the dimensions of the objects (design matrices, mean vectors, etc.) are those that were used in the analysis of the teacher data where we have 1 circular outcome, 1 linear outcome and estimate an intercept and regression coefficient for the covariate self-efficacy. Note that for the regression of the linear component in the CL-PN and CL-GPN models we also have the sine and cosine of the circular outcome in the regression equation, this makes the vector with regression coefficients, $\boldsymbol{\gamma}$, four-dimensional.

\subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k_{n \times 2}$ (for $k \in \{I,II\}$) and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set. 
\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with  $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.001$ and  $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_2$ for $k \in \{I,II\}$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$ and $\boldsymbol{\beta^{k}} = (0,0)^t$ for $k \in \{I,II\}$. Also set starting values $r_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i)\\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Sample $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse gamma posterior with $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_n, \boldsymbol{\Lambda}_n)$, with $\boldsymbol{\mu}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + (\boldsymbol{Z}^k)^t\boldsymbol{s}^k)$ and $\boldsymbol{\Lambda}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}_0)$.
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left(-\frac{1}{2}(r_i)^2 + b_ir_i\right)}$$ 
where $b_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z_i}$ and $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. 
\noindent We can sample from this posterior using a slice sampling technique (Cremers et al., 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^t\boldsymbol{z}_{i}) \propto r_{i} \textbf{I}\left(0 < v_i < \exp\left\{ -\frac{1}{2}(r_{i} - b_{i})^2\right\}\right)\textbf{I}(r_i > 0).$$
\noindent The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is
$$U\left(0, \exp\left\{-\frac{1}{2}(r_{i} -  b _{i})^2\right\}\right)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is proportional to
$$r_{i} \textbf{I}\left(b_{i} + \max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\} < r_{i} < b_{i} + \sqrt{-2\ln v_{i}}\right).$$
\noindent We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $ r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=b_{i} +\max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\}$ and $ r_{i_{2}}= b_{i} + \sqrt{-2\ln v_{i}}$.
\end{itemize}
\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset.
\item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We assess convergence visually using traceplots.
\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified CL-GPN model}\label{A2}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}_{n \times 2}$  and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set. 
\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.001$ and $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
\item Prior for $\rho$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
\item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$ for $j \in \{0,1\}$, $\rho = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$. Also set starting values $r_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$ for $j \in \{0,1\}$, $\rho$ and $\tau$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse gamma posterior where $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
\item Posterior for $\rho$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = (\boldsymbol{\beta}^{I})^t\boldsymbol{z}_i$ and $\mu_i^{II} = (\boldsymbol{\beta}^{II})^t\boldsymbol{z}_i$. 
\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{I} + \rho(s^{II}_{i} - \mu_i^{II})\})^2 + \beta_0$
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-\frac{1}{2}A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z_i}$, $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ and $A_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}$.
\noindent We can sample from this posterior using a slice sampling technique (Hernandez-Stumpfhauser et al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^t\boldsymbol{z}_{i}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -\frac{1}{2} A_i\left(r_{i} - \frac{B_i}{A_i}\right)^2\right\}\bigg)\textbf{I}(r_i > 0).$$
\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is
$$U\Bigg(0, \exp\left\{-\frac{1}{2}A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right).$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.
\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\rho$ and $\tau$ for the training dataset. 
\item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We visually assess convergence using traceplots.
\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified GPN-SSN model}\label{A3}

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort. 
\item Define the prior parameters for the training set. Since we have only one circular outcome, one linear outcome and one covariate, we have $m = 1$, $w = 1$ and $g = 1$. In this paper we use the following priors:

\begin{itemize}
\item Prior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_0, \nu_0)$, an inverse Wishart with $\boldsymbol{\Psi}_0 = 10^{-4}\boldsymbol{I}_{2m + w}$ and $\nu_0 = 1$.   
\item Prior for $\boldsymbol{B}$ in vectorized form: $N_{(g + 1)(2m + w)}(\boldsymbol{\beta}_0, \boldsymbol{\Sigma}  \otimes \boldsymbol{\kappa}_0)$, where $\otimes$ stands for the Kronecker product, $\boldsymbol{\beta}_0 = \text{vec}(\boldsymbol{B}_0)$, the matrix with prior values for the regression coefficients. We choose $\boldsymbol{\beta}_0 = \boldsymbol{0}_{(g + 1)(2m + w)}$, $\boldsymbol{B}_0 = \boldsymbol{0}_{(g + 1) \times (2m + w)}$ and $\boldsymbol{\kappa}_0 = 10^{-4}\boldsymbol{I}_{g + 1}$.
\item Prior for $\lambda$: $N(\gamma_0, \omega_0)$, with $\gamma_0 = 0$ and $\omega_0 = 10000$.
\end{itemize}

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:
$$\tilde{y}_i = \lambda d_i. $$
\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual $i$:
$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i^t,y_i)^t - (\boldsymbol{0}_{2m}^t, \lambda d_i)^{t}.$$

\item Sample $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_n, \nu_n)$, an inverse Wishart with $\boldsymbol{\Psi}_n = \boldsymbol{\Psi}_0 + (\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B})^t(\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B}) + (\boldsymbol{B} - \boldsymbol{B}_0)^t\boldsymbol{\kappa}_0(\boldsymbol{B} - \boldsymbol{B}_0)$ and $\nu_n = \nu_0 + n$ where $n$ is the sample size.
\item Posterior for $\boldsymbol{B}$ in matrix form: $MN(\boldsymbol{B}_n,   \boldsymbol{\kappa}_n, \boldsymbol{\Sigma})$, with $\boldsymbol{B}_n = \boldsymbol{\kappa}_n^{-1}\boldsymbol{X}^t\boldsymbol{\eta} + \boldsymbol{\kappa}_0\boldsymbol{B}_0$ and $\boldsymbol{\kappa}_n = \boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\kappa}_0$.
\item Posterior for $\lambda$: $N(\gamma_n, \omega_n)$, with $\omega_n = \big(\sum_{i = 1}^{n}d_i^2\sigma^{-2}_{y|s} + \omega_0^{-1}\big)^{-1}$ and $\gamma_n = \omega_n \big(\sum_{i = 1}^{n}d_i\sigma^{-2}_{y|s}(y_i - \mu_{y_i|s_i}) + \omega_0^{-1}\gamma_0 \big)$ where $\mu_{y_i|s_i} = \mu_y + \boldsymbol{\Sigma}_{sy}^{t}\boldsymbol{\Sigma}_{s}^{-1}(\boldsymbol{s_i - \boldsymbol{\mu}_s)}$ and $\sigma^2_{y|s} = \sigma^2_{y} - \boldsymbol{\Sigma}_{sy}^{t}\boldsymbol{\Sigma}_{s}^{-1}\boldsymbol{\Sigma}_{sy}$.
\end{itemize}

\item Sample new $d_i$ for the training and holdout dataset from the following posterior:
$$f(d_i) \propto \phi(y_i|\mu_{y_i|s_i} + \lambda d_i, \sigma^2_{y|s})\phi(d_i|0, 1),$$
where $\mu_{y_i|s_i} = \boldsymbol{B}_{y_i|s_i}^t\boldsymbol{x}_i$. We can see each $d_i$ as a positive regressor with $\lambda$ as covariate and $\phi(d_i|0, 1)$ as prior (Mastrantonio, 2018). The full conditional is then truncated normal with support $\mathbb{R}^{+}$ as follows:
$$N(m_{d_i}, v),$$ 
\noindent where $v = \big(\lambda^2\sigma^{-2}_{y|s} + 1\big)$ and $m_{d_i} = v\lambda\sigma^{-2}_{y|s}\big(y_i - \mu_{y_i|s_i}\big)$. 
\item Sample new $r_i$ for the training and holdout dataset from the following posterior
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i|y_i}^{-1}\boldsymbol{\mu}_{s_i|y_i}$, $\boldsymbol{\mu}_{s_i|y_i} = \boldsymbol{B}_{s_i|y_i}^t\boldsymbol{x}_i$ and $A_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i| y_i}^{-1}\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}$. The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i^t, y_i)^t \sim N_{2m+w}(\boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, \lambda d_i)^t, \boldsymbol{\Sigma})$. 
Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can in this model (where the variance-covariance matrix of the circular outcome is not constrained in the estimation procedure) simply define the $r_i$ as the Euclidean norm of the bivariate datapoints. However, for didactic purposes we continue with the explanation of the sampling procedure. We
can sample from the posterior for $r_i$ using a slice sampling technique (Hernandez-Stumpfhauser et al. 2018): 
\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^{t}\boldsymbol{x}_{i}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -\frac{1}{2}A_i\left(r_{i} - \frac{B_i}{A_i}\right)^2\right\}\bigg)\textbf{I}(r_i > 0).$$
\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is
$$U\Bigg(0, \exp\left\{-\frac{1}{2}A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to 
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.

\item Repeat steps 4 to 10 until the sampled parameter estimates have converged.

\item In the MCMC sampler we have estimated an unconstrained $\boldsymbol{\Sigma}$. However, for identification of the model we need to apply  constraints to both $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$. Therefore we need the matrix
$$\boldsymbol{C} = \begin{bmatrix} \boldsymbol{C}_s & \boldsymbol{0}_{2m \times w} \\ \boldsymbol{0}_{2m \times w}^t & \boldsymbol{I}_w \end{bmatrix}$$
where $\boldsymbol{C}_s$ is a $2m \times 2m$ diagonal matrix with every $(2(j-1) + k)^{th}$ entry $> 0$ where $k \in \{1,2\}$ and $j = 1, \dots, m$ (Mastrantonio, 2018). The estimates $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$ can then be related to their constrained versions $\tilde{\boldsymbol{\Sigma}}$ and $\tilde{\boldsymbol{\mu}}$ as follows:
$$\boldsymbol{\mu} = \boldsymbol{C}\tilde{\boldsymbol{\mu}}$$
$$\boldsymbol{\Sigma} = \boldsymbol{C}\tilde{\boldsymbol{\Sigma}}\boldsymbol{C}.$$

\end{enumerate}

### 2018-09-24 Manuscript/SupplementaryMaterial.tex
Supplementary material

--- 2018-09-24 Manuscript/SupplementaryMaterial.Rmd
@@
----
-title: " "
-author: |
-bibliography: CircularData.bib
-csl: apa.csl
-output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    number_sections: no
-fontsize: 12pt
-geometry: margin=1in
-header-includes:
-  - \usepackage{multirow}
-  - \usepackage{appendix}
-  - \usepackage{color}
-  - \usepackage{hyperref}
-  - \usepackage{subcaption}
-  - \setlength\parindent{24pt}
-  - \usepackage{setspace}\doublespacing
-documentclass: article
-pandoc_args: --natbib
----
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{}
+    \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+    \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
+
+\usepackage{multirow}
+\usepackage{appendix}
+\usepackage{color}
+\usepackage{hyperref}
+\usepackage{subcaption}
+\setlength\parindent{24pt}
+\usepackage{setspace}\doublespacing
+
+\begin{document}
 
@@
 
-In this Supplementary Material we outline the MCMC procedures to fit the cylindrical regression models.
-R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
-Note that the dimensions of the objects (design matrices, mean vectors, etc.) are those that were used in the analysis of the teacher data where we have 1 circular outcome, 1 linear outcome and estimate an intercept and regression coefficient for the covariate self-efficacy.
-Note that for the regression of the linear component in the CL-PN and CL-GPN models we also have the sine and cosine of the circular outcome in the regression equation, this makes the vector with regression coefficients, $\boldsymbol{\gamma}$, four-dimensional.
+In this Supplementary Material we outline the MCMC procedures to fit the
+cylindrical regression models. R-code for the MCMC sampler and the
+analysis of the teacher data can be found here:
+\url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
+Note that the dimensions of the objects (design matrices, mean vectors,
+etc.) are those that were used in the analysis of the teacher data where
+we have 1 circular outcome, 1 linear outcome and estimate an intercept
+and regression coefficient for the covariate self-efficacy. Note that
+for the regression of the linear component in the CL-PN and CL-GPN
+models we also have the sine and cosine of the circular outcome in the
+regression equation, this makes the vector with regression coefficients,
+\(\boldsymbol{\gamma}\), four-dimensional.
 
@@
 
-We use the following algorithm to obtain posterior estimates from the model:
+We use the following algorithm to obtain posterior estimates from the
+model:
 
@@
 \newpage
+
 \subsection{Bayesian Model and MCMC procedure for the modified CL-GPN model}\label{A2}
 
-We use the following algorithm to obtain posterior estimates from the model:
+We use the following algorithm to obtain posterior estimates from the
+model:
 
@@
 \newpage
+
 \subsection{Bayesian Model and MCMC procedure for the modified GPN-SSN model}\label{A3}
@@
 \end{enumerate}
+
+\end{document}

### 2018-09-24 Manuscript/ManuscriptMarkdownCV.Rmd
Manuscript update

---
title: "Regression models for Cylindrical data in Psychology"
author: #|
 # | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
 # | $^1$Department of Methodology and Statistics, Utrecht University
 # | $^2$TNO
  #| $^3$Department of Education, Utrecht University
  #| $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University

bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
  - \setlength\parindent{24pt}
  - \usepackage[nodisplayskipstretch]{setspace}\doublespacing
  #- \usepackage{endfloat}
  - \DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)

keywords: bla vla
documentclass: article
pandoc_args: --natbib
abstract: |
  Cylindrical data are multivariate data which consist of a directional,
  in this paper circular, and a linear component. Examples of cylindrical
  data in psychology include human navigation (direction and distance of 
  movement), eye-tracking research (direction and length of saccades) and
  data from an interpersonal circumplex (type and strength of
  interpersonal behavior). In this paper we adapt four models for
  cylindrical data to include a regression of the circular and linear
  component onto a set of covariates. Subsequently, we illustrate how to
  fit these models and interpret their results on a dataset on the
  interpersonal behavior of teachers.
---
```{r inlude = FALSE, echo = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
library(plotrix)
library(caret) #cross-validation
```

\section{Introduction}\label{Introduction}

Cylindrical data are data that consist of a linear variable and a directional
variable. In this paper, the directional variable is circular, meaning that it
consists of a single angle instead of a set of angles. A circular variable is
different from a linear variable in the sense that it is measured on a different
scale. Figure \ref{circline} shows the difference between a circular scale
(right) and a linear scale (left). The most important difference is that on a
circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in
fact represent the same number while on a linear scale the two ends, \(-\infty\)
and \(\infty\), are not connected and consequently the values 0\(^\circ\) and
360\(^\circ\) are located on different places on the scale. This difference
requires us to use special statistical methods for circular variables (see e.g.
@fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
  elaborate overview). For instance, a notion as simple as the sample average
  needs to be re-defined. As is the case for circular data, the analysis of
  cylindrical data also requires special methods.\newline
\indent Cylindrical data occur in several fields of research, such as for
instance in meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]. However, also in psychology several types
of cylindrical data can be found. For example, in research on human navigation
in the field of cognitive psychology both the distance, a linear variable, and
the direction, a circular variable, of movement are of interest
[@chrastil2017rotational]. In eye-tracking research, saccade data are an example
of cylindrical data, because both the direction (\emph{i.e.}, the circular variable)
and the duration (\emph{i.e.}, the linear variable) of the saccades are of interest
(for a review of eye-tracking research see  @rayner200935th).\newline
\indent The type of data that is used in the present study are also
psychological, namely data from circumplex measurement instruments. For
instance, data from the interpersonal circumplex as used in personality
psychology are by definition of a cylindrical nature (see Section \ref{Example}
for a more detailed explanation).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.\newline
\indent Our aim in this paper is to fill this gap in the literature by adapting
four existing cylindrical models in such a way that they include a regression of
both the linear and circular component of a cylindrical variable onto a set of
covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. Additionally, we will show how a
correct statistical treatment of such cylindrical data can lead to new insights.
We will do this for the teacher data, a dataset from the field of educational
psychology. In the teacher data, apart from modelling the relation between the
linear and circular component of a cylindrical variable we would also like to
predict the two components from a set of covariates in a regression model.

The paper is organized as follows: Section \ref{Example} describes the teacher
data, while Section \ref{Models} presents the four cylindrical models and our
associated adaptations to new regression models. In that same section we also
discuss the model fit criterion that we will use in Section \ref{DataAnalysis}
for the comparison of the four models. A detailed data analysis with interesting
new insights is also provided in Section \ref{DataAnalysis}. We conclude the
paper with a discussion in Section \ref{Discussion}, and the Supplementary Material collects
the technical details of the MCMC procedures.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/circline.pdf}
\caption{The difference between a linear scale (left) and a circular scale (right).}
\label{circline}
\end{figure}

\section{Teacher data}\label{Example}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/IPC-T.png}
\caption{The interpersonal circle for teachers (IPC-T). The words presented in
the circumference of the circle are anchor words to describe the type of
behavior located in each part of the IPC.}
\label{QTI}
\end{figure}

The motivating example for this article comes from the field of educational
psychology and was collected for the studies on classroom climate of
@vanderWant2015role, @Claessens2016side and @pennings2018interpersonal. An
indicator of the quality of the classroom climate is the students' perception of
their teachers' interpersonal behavior. These interpersonal perceptions, both in
educational psychology as well as in other areas of psychology, can be measured
using circumplex measurement instruments (see @horowitz2010handbook for an
overview of many such instruments).\newline
\indent The circumplex data used in this paper are measured using the
Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] which is
one such circumplex measurement instrument. The QTI is designed to measure
student perceptions of their teachers' interpersonal behavior and contains items
that load on two interpersonal dimensions: Agency and Communion. Agency refers
to the degree of power or control a teacher exerts in interaction with his/her
students. Communion refers to the degree of friendliness or affiliation a
teacher conveys in interaction with his/her students. The loadings on the two
dimensions of the QTI can be placed in a two-dimensional space formed by Agency
(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
this space are characterized by different teacher behavior, e.g. 'helpful' or
'uncertain'. This two-dimensional space is called the interpersonal
circle/circumplex (IPC). The IPC is ``a continuous order with no beginning or
end'' [@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering
and the IPC is therefore often called the interpersonal circumplex. The ordering
also implies that scores on the IPC could be viewed as a circular
variable.\newline
\indent @Cremers2018Assessing explain the circular nature of the IPC data and
analyze them as such using a circular regression model. The two-dimension scores
Agency and Communion can be converted to a circular score using the two-argument
arctangent function in \eqref{PredVal}, where \(A\) represents a score on the
Agency dimension and \(C\) represents a score on the Communion dimension
\begin{equation}\label{PredVal}
\theta          = \text{atan2}\left(A, \: C\right)  =
\left\{{\begin{array}{lcl}
                                                                       \arctan\left(\frac{A}{C}\right) & \text{if}  \quad&C > 0 \\
\arctan\left(\frac{A}{C}\right) + \pi & \text{if}  \quad& C  <  0  \:\: \&\:\: A \geq 0\\
 \arctan\left(\frac{A}{C}\right) - \pi & \text{if}  \quad&C  <  0 \:\:  \&\:\:A  < 0\\
 +\frac{\pi}{2} & \text{if}  \quad& C  =  0  \:\: \&\:\:A > 0\\
 -\frac{\pi}{2} & \text{if}  \quad& C =  0  \:\: \&\:\:A < 0\\
 \text{undefined} & \text{if} \quad& C =  0   \:\: \&\:\:A = 0.
 \end{array}}
\right.
\end{equation}
The resulting circular variable \(\theta\) can then be modelled and takes values
in the interval $[0, 2\pi)$. However, when two-dimensional data are converted
to the circle we lose some information, namely the length of the two-dimensional
vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\).
This length represents the strength of the type of interpersonal behavior a
teacher shows towards his/her students and can be considered as the linear
variable in a cylindrical model, allowing us to model a circular variable
\(\theta\) together with the linear variable corresponding to \(\mid\mid (A, C)^t
\mid\mid\). This leads to an improved analysis of interpersonal circumplex data
as we take all information into account. In the next section we introduce
several models that can be used for a more accurate and informative regression
analysis on the teacher data. First however we will provide descriptives for our
data set.

\subsection{Data description}\label{DataDescriptives}

```{r loaddata, return = FALSE, echo = FALSE}
Dat <- read_spss("Data Heleen/mergeddata.sav")

#only select the first measurement occasion
Dat <- subset(Dat, Time == 0)

#sort on class size
Dat <- Dat[order(-Dat$nklas), ]

#Now remove duplicates 
#(because dataframe is sorted on class size the smallest classes are removed)

duplicates <- duplicated(Dat[,c("docnr", "Time")])
Dat <- Dat[!duplicates,]

#center and remove missings for self-efficacy
Dat <- Dat %>% drop_na(Efficacy_CM)
Dat$SEc <- Dat$Efficacy_CM- mean(Dat$Efficacy_CM)
Dat$theta <- atan2(Dat$lAgency, Dat$lcommunion)
Dat$y <- sqrt(Dat$lAgency^2 + Dat$lcommunion^2)

set.seed(30)

#Create holdout and training samples for cross-validation

vec <-  seq(1:nrow(Dat))
flds <- createFolds(vec, k = 10, list = TRUE, returnTrain = FALSE)

for(i in 1:10) {

 #Create name for holdout data 
 nam.h <- paste("hold.", i, sep = "")

 #Get holdout dataset for fold i
 assign(nam.h, Dat[flds[[i]],])

 #Create name for training data
 nam.tr <- paste("train.", i, sep = "")

 #Get training dataset for fold i
 assign(nam.tr, Dat[-flds[[i]],])

}

```

```{r dataplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("Plots/dataplot.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

plot(Dat$y, Dat$theta*(180/pi), 
     ylab = "IPC location", xlab = "IPC strength",
     bty = "n", ylim = c(-180, 180))

dev.off()
tools::texi2dvi("Plots/dataplot.tex", pdf=TRUE)

```

```{r descripitves, echo = FALSE, results = FALSE}
#sample size
nrow(Dat)

#summary statistics
summary(Dat$Efficacy_CM)
summary(Dat$y)
summary(as.circular(Dat$theta%%(2*pi)))

sd(Dat$Efficacy_CM)
sd(Dat$y)

```

The teacher data was collected between 2010 and 2015 and contains several
repeated measures on the IPC of 161 teachers. Measurements were obtained using
the QTI and taken in different years and classes. For this paper we only
consider one measurement, the first occasion (2010) and largest class if data
for multiple classes were available. In addition to the score on the IPC, the
circular outcome, and the strength of the score on the IPC, the linear outcome,
a teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
covariate in the analysis. After listwise deletion of missings ($3$ in total,
only for the self-efficacy) we have a sample of 148 teachers. Table
\ref{Tableteacherdescriptives} shows descriptives for the dataset. Here
$\hat{\rho}$ is a sample estimate for the circular concentration where a value
of 0 means that the data is not concentrated at all, \emph{i.e.} spread over the
entire circle, and a value of 1 means that all data is concentrated at a single
point on the circle. Figure \ref{dataplot} is a scatterplot showing the relation
between the linear and circular outcome of the teacher data.

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/dataplot.pdf}
\caption{Plot showing the relation between the linear and circular outcome component (in degrees) of the teacher data.}
\label{dataplot}
\end{figure}

\begin{table}[h]
\centering
\caption{Descriptives for the teacher dataset.} 
\begin{tabular}{lrrrl}
  \noalign{\smallskip}\hline\noalign{\smallskip}
Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
IPC &33.22$^\circ$& 0.76 & - & Circular\\
strength IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
SE & 5.04 & 1.00 & 1.5 - 7.0 & Linear\\
   \hline
\end{tabular}
\label{Tableteacherdescriptives}
\end{table}

\section{Four cylindrical regression models}\label{Models}

In this section we present four cylindrical models and adapt them such
that they contain predictors for the linear and circular outcomes, \(Y\)
and \(\Theta\). The first two models are based on a construction by
@mastrantonio2015bayesian, while the other models are extensions of the
models from @abe2017tractable and @mastrantonio2018joint.

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

Following @mastrantonio2015bayesian we consider in this section two models where
the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
and $q$ covariates is specified as
\begin{equation}\label{circlinlink}
Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}
\noindent where the random variable \(R\geq0\) will be introduced below, the
error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
covariates. In both of these models the
conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
by
\begin{equation}\label{ycondtheta}
f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
\end{equation}
\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
in its location part. For the teacher dataset, the regression equation for the
linear outcome in the CL-PN and CL-GPN model is the following: \[\hat{y}_i =
\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i +
\gamma_1\text{SE}_i,\]
\noindent where $\text{SE}_i$ is the self-efficacy score of one individual $i = 1, \dots, n$ where $n$ is the sample size.\newline
\indent For the circular outcome we assume either a projected normal (PN) or a general
projected normal (GPN) distribution. These distributions arise from the radial
projection of a distribution defined on the plane onto the circle. The relation
between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
outcome \(\Theta\) is defined as follows
\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
\end{equation}
\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1
\end{bmatrix}\), \(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
@hernandez2016general). This leads to the
circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
will now detail how we modify both cylindrical distributions to also incorporate
covariates for the circular part.

\subsubsection{The modified CL-PN distribution}

Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
PN distribution equals
\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
\(n\) is the sample size, are generated independently from the distribution with
density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
contains the covariate values and the value 1 to estimate an intercept and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
the regression coefficients and intercepts. Note however that the dimensions of
\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{II}\) need not necessarily
be the same and we are thus allowed to have a different set of predictor
variables and vectors \(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for
the two components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
regression equation for the circular outcome in the CL-PN model is
\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix} \mu_{i}^{I}  \\ \mu_{i}^{II}
\end{pmatrix}=\begin{pmatrix} \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
\beta_0^{II} + \beta_1^{II}\text{SE}_i \end{pmatrix}.\]

\subsubsection{The modified CL-GPN distribution}

Following @wang2012directional and @hernandez2016general the joint density of
\(R\) and \(\Theta\) for the GPN distribution equals
\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
\rho^2 & \rho\\ \rho & 1 \end{bmatrix}\). In a regression setup the outcomes
\(\theta_i\) and \(r_i\) for each individual are generated independently from
\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
defined in the same way via covariates as for the modified CL-PN distribution.
Note in contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the same
predictors for both components of \(\boldsymbol{\mu}_i\) in the CL-GPN model.
This is due to the fact that the variance-covariance matrix
\(\boldsymbol{\Sigma}\) is no longer identity. For the teacher dataset, the
regression equation for the circular outcome in the CL-GPN model is the same as
in the CL-PN model.

\subsubsection{Parameter estimation}

Both cylindrical models introduced here are estimated using Markov Chain Monte
Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in the
Supplementary Material.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. The joint density of \(\Theta\) and
\(Y\), in this model defined only on the positive real half-line \([0, +
\infty)\), reads
\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\beta^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation}
\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
\(\kappa\) also regulating the circular-linear dependence. Our modification
occurs at the level of the linear scale parameter \(\beta>0\) and circular
location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
covariates: \(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\). The
parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\eta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a vector of \(p\)
regression coefficients \(\eta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
generated independently from the modified
density \eqref{WeiSSVMdensity}.\newline
\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
with location parameter \(\mu\) and concentration parameter \((\beta
y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\nu} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))).\nonumber
\end{align}
\noindent For the teacher data, $\boldsymbol{z} = \boldsymbol{x}$ and the regression equations for the circular and
linear outcomes in the Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
\tan^{-1}(\eta_1\text{SE}_i),\] and \[\hat{\beta}_{i} = \exp(\nu_0 +
\nu_1\text{SE}_i).\]
\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
  outcomes and \(w\) independent linear outcomes. The circular outcomes
  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
  so-called augmented representation (as in \eqref{projection} and
  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
  multivariate skew normal distribution. Following @mastrantonio2018joint the
  linear outcomes are represented as
\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear outcome
\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
is a \(w \times w\) diagonal matrix with diagonal elements
\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
$w$-dimensional half normal distribution [@olmos2012extension], and
\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
is normally distributed with mean \(\boldsymbol{\mu}_y +
\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
\boldsymbol{D}^t)^t\) is defined as:
\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
\end{equation}
\noindent where
\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
the \(\ell\)-dimensional normal density with mean vector
\(\boldsymbol{\mu}_\ell\) and
covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
modelling the augmented representations of \(\boldsymbol{\Theta}\) and
\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
distribution. The joint density of the model is then represented by:
\begin{equation}\label{YDThetarjoint} 
f(\boldsymbol{\theta}, \boldsymbol{r},
\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
\boldsymbol{y}^t)^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, ({\rm
diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
1}^{m}r_j, 
\end{equation}
\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{\mu} =
(\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\) and \(\boldsymbol{\Sigma} =
\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
of and covariances between the augmented representations of the circular outcome
and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
augmented representations of the circular outcome and the linear outcome.
\newline
\indent In our regression extension we have \(i = 1, \dots, n\) observations of
\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{\mu}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
and the $g$ covariate values. \newline
\indent For the teacher data, the regression equations for
the circular and linear outcomes in the GPN-SSN model are
\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +
\boldsymbol{\beta}_1\text{SE}_i,\] where
\(\hat{\boldsymbol{\mu}}_i = (\hat{\boldsymbol{\mu}}_{s_i}^t, \hat{\mu}_{y_i})^t\),
\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
and
\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\). Note that because $m = 1$ and $w = 1$, $\boldsymbol{\mu}_{s_i}$ is a 2 dimensional vector and $\mu_{y_i}$ is a scalar.
We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

\subsection{Model fit criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample predictive
performance to determine the fit of the model. To do so we use k-fold
cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
of the sample) is used once as a holdout set and 9 times as part of a training
set. The analysis will thus be performed 10 times, each time on a different
training set.\newline
\indent A proper criterion to compare out-of-sample predictive performance is
the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
value of this criterion, the better the predictive performance of the model.
Using ML estimates this criterion can be computed as follows:
\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
\end{equation}
\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) [@BDA, p. 169] and can be computed as:
\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
\end{equation}
\noindent where \(B\) is the amount of posterior samples and
\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
parameters for the \(j^{th}\) iteration. Because the joint density and thus also
the likelihood for the modified GPN-SSN model in (\ref{YDThetarjoint}) is not
available in closed form [@mastrantonio2018joint] we compute the PLSL for the
circular and linear outcome separately for all models. Note that although we fit
the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
prior information into account when assessing model fit with the PLSL. According
to @BDA this is not necessary since we are assessing the fit of a model
to data, the holdout set, only. They argue that the prior in such case is only
of interest for estimating the parameters of the model but not for determining
the predictive accuracy.\newline
\indent We use the loglikelihoods of the following conditional densities for the
computation of the PLSL in the teacher data:

\begin{itemize}
\item For the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.

\item For the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.

\item For the modified Abe-Ley model:

$y_i \mid \theta_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item For the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.
\end{itemize}

\noindent For each of the four cylindrical models and for each of the 10
cross-validation analyses we can then compute a PLSL for the circular and linear
outcome by using the conditional log-likelihoods of the respective outcome. To
evaluate the predictive performance we average across the PLSL criteria of the
cross-validation analyses. We also assess the cross-validation variability by
means of the standard deviations of the PLSL criteria.

\section{Data Analysis}\label{DataAnalysis}

In this section we analyze the teacher data with the help of the four
cylindrical models from Section \ref{Models}. We will present the results,
posterior estimates and their interpretation, per model and finish with a
section comparing the fit of the different models.

```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling CL-PN.R")

its <- 20000
set.seed(101)

res_CLPN <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold <- get(paste("hold.", i, sep = ""))

  ZI.PN.tr  <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  ZII.PN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.PN.tr   <- as.matrix(cbind(rep(1, length(train$theta)),
                                 cos(train$theta),
                                 sin(train$theta),
                                 train$SEc))

  ZI.PN.h  <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  ZII.PN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.PN.h   <- as.matrix(cbind(rep(1, length(hold$theta)),
                               cos(hold$theta),
                               sin(hold$theta),
                               hold$SEc))

  res_CLPN[[i]] <- CLPN(train$theta, train$y, X.PN.tr, ZI.PN.tr, ZII.PN.tr, its,
                        hold$theta, hold$y, X.PN.h, ZI.PN.h, ZII.PN.h)

}

```

```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling CL-GPN.R")

its <- 20000
set.seed(101)

res_CLGPN <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold  <- get(paste("hold.", i, sep = ""))

  Z.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)),
                              cos(train$theta),
                              sin(train$theta),
                              train$SEc))

  Z.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)),
                             cos(hold$theta),
                             sin(hold$theta),
                             hold$SEc))

  res_CLGPN[[i]] <- CLGPN(train$theta, train$y, X.GPN.tr, Z.GPN.tr, its, p = 2, 
                          hold$theta, hold$y, X.GPN.h, Z.GPN.h)

}

```

```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling Joint GPN-SSN.R")

its <- 20000
set.seed(101)

res_CLGPNM <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold  <- get(paste("hold.", i, sep = ""))

  X.MGPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))

  X.MGPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))

  res_CLGPNM[[i]] <- JGPNSSN(train$theta, train$y, X.MGPN.tr, its, p = 1, q = 1, 
                             hold$theta, hold$y, X.MGPN.h)

}

```

```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Abe-Ley optimization.R")

resAL <- list()

set.seed(101)

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold <- get(paste("hold.", i, sep = ""))

  Z.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))

  Z.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))

  theta.tr <- as.numeric(train$theta)
  y.tr     <- as.numeric(train$y)
  theta.h  <- as.numeric(hold$theta)
  y.h      <- as.numeric(hold$y)

  colnames(X.tr) <- c("ax", "bx")
  colnames(Z.tr) <- c("az", "bz")
  colnames(X.h) <- c("ax", "bx")
  colnames(Z.h) <- c("az", "bz")

  dat.t <- as.data.frame(cbind(theta.tr, y.tr, X.tr, Z.tr))
  dat.h <- as.data.frame(cbind(theta.h, y.h, X.h, Z.h))

  ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
                 c(-1, 0, 0, 0, 0, 0, 0),
                 c(0, 0, 0, 0, 1, 0, 0),
                 c(0, 0, 0, 0, 0, 1, 0),
                 c(0, 0, 0, 0, 0, 0, 1),
                 c(0, 0, 0, 0, 0, 0, -1))

  ci.reg = c(-pi, -pi, 0, 0, -1, -1)

  param <- c(1,1,1,1,1,1,0)*0.9

  ui.reg %*% param - ci.reg

  resAL[[i]] <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
                            control = list(maxit = 1000000, fnscale = -1), data = dat.t)

  resAL[[i]]$ll.circ <- -2*func.reg.cond.circ(resAL[[i]]$par, dat.h)
  resAL[[i]]$ll.lin  <- -2*func.reg.cond.lin(resAL[[i]]$par, dat.h)

}

```

\subsection{Results \& Analysis}\label{DataResults}

In the Supplementary Material we have described the starting values for the MCMC
procedures, hence it remains to specify  the starting  values for the
maximum likelihood based Abe-Ley model: \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 =
0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\). The initial amount
of iterations for the three MCMC samplers was set to 2000. After convergence
checks via traceplots we concluded that some of the parameters of the GPN-SSN
model did not converge. Therefore we set the amount of iterations of the MCMC
models to 20,000 and subtracted a burn-in of 5000 to reach convergence. Note
that we choose the same amount of iterations for all three Bayesian models to
make their comparison via the PLSL as fair as possible. Lastly, the predictor
\verb|SE| was centered before inclusion in the analysis as this allows the
intercepts to bear the classical meaning of average behavior.

\subsubsection{The modified CL-PN and CL-GPN models} \label{resCL(G)PN}

```{r estCLPN, cache = TRUE, echo = FALSE}

#Get 3 dimensional arrays (iteration, parameter, fold) for each parameter type
Gamma <- simplify2array(lapply(res_CLPN, "[[", "Gamma"))
BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
Sigma <- simplify2array(lapply(res_CLPN, "[[", "Sigma"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
CLPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est),1, sd)
CLPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
CLPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
CLPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
CLPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
CLPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.Sigma.m   <- mean(apply(Sigma[5000:20000,], 2, mode_est))
CLPN.Sigma.hpd <- rowMeans(apply(Sigma[5000:20000,], 2, hpd_est))
CLPN.Sigma.m.sd  <- sd(apply(Sigma[5000:20000,], 2, mode_est))
CLPN.Sigma.hpd.sd <- apply(apply(Sigma[5000:20000,], 2, hpd_est), 1, sd)

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)

modes.PN.sd <- c(CLPN.BI.m.sd, CLPN.BII.m.sd, CLPN.Gamma.m.sd, CLPN.Sigma.m.sd, NA, NA, NA)
hpd.LB.PN.sd <- c(CLPN.BI.hpd.sd[1,], CLPN.BII.hpd.sd[1,], CLPN.Gamma.hpd.sd[1,], CLPN.Sigma.hpd.sd[1], NA, NA, NA)
hpd.UB.PN.sd <- c(CLPN.BI.hpd.sd[2,], CLPN.BII.hpd.sd[2,], CLPN.Gamma.hpd.sd[2,], CLPN.Sigma.hpd.sd[2], NA, NA, NA)

CLPNres.tab <- cbind(modes.PN, modes.PN.sd, hpd.LB.PN, hpd.LB.PN.sd, hpd.UB.PN, hpd.UB.PN.sd)
CLPNres.tab = CLPNres.tab

CLPNres.tab.new = cbind(apply(CLPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
                        apply(CLPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
                        apply(CLPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                                       sprintf("%.2f", round(x[2], 2)), ")")))

```

```{r estCLGPN, cache = TRUE, echo = FALSE}

#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
Gamma <- simplify2array(lapply(res_CLGPN, "[[", "Gamma"))
BI    <- simplify2array(lapply(res_CLGPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLGPN, "[[", "BII"))
Sig   <- simplify2array(lapply(res_CLGPN, "[[", "Sig"))
Sigma <- simplify2array(lapply(res_CLGPN, "[[", "Sigma"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLGPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
CLGPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
CLGPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
CLGPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.Sig.m   <- mean(apply(Sig[5000:20000,], 2, mode_est))
CLGPN.Sig.m.sd   <- sd(apply(Sig[5000:20000,], 2, mode_est))
CLGPN.Sig.hpd <- rowMeans(apply(Sig[5000:20000,], 2, hpd_est))
CLGPN.Sig.hpd.sd <- apply(apply(Sig[5000:20000,], 2, hpd_est), 1, sd)

CLGPN.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPN.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPN.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPN.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPN.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPN.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]

CLGPN.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPN.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPN.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPN.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPN.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPN.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
                CLGPN.Gamma.m, CLGPN.Sig.m,
                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])

modes.GPN.sd  <- c(CLGPN.BI.m.sd, CLGPN.BII.m.sd,
                   CLGPN.Gamma.m.sd, CLGPN.Sig.m.sd,
                   CLGPN.Sigma11.m.sd, CLGPN.Sigma12.m.sd, CLGPN.Sigma22.m.sd)
hpd.LB.GPN.sd <- c(CLGPN.BI.hpd.sd[1,], CLGPN.BII.hpd.sd[1,],
                   CLGPN.Gamma.hpd.sd[1,], CLGPN.Sig.hpd.sd[1],
                   CLGPN.Sigma11.hpd.sd[1], CLGPN.Sigma12.hpd.sd[1], CLGPN.Sigma22.hpd.sd[1])
hpd.UB.GPN.sd <- c(CLGPN.BI.hpd.sd[2,], CLGPN.BII.hpd.sd[2,], 
                   CLGPN.Gamma.hpd.sd[2,], CLGPN.Sig.hpd.sd[2],
                   CLGPN.Sigma11.hpd.sd[2], CLGPN.Sigma12.hpd.sd[2], CLGPN.Sigma22.hpd.sd[2])

CLGPNres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd, hpd.UB.GPN, hpd.UB.GPN.sd)

CLGPNres.tab.new = cbind(apply(CLGPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
                         apply(CLGPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
                         apply(CLGPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")))

CLPNGPNres.tab <- cbind(CLPNres.tab.new, CLGPNres.tab.new)

rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_1^{I}$",
                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
                              "$\\gamma_0$", "$\\gamma_{cos}$",
                              "$\\gamma_{sin}$", "$\\gamma_1$",
                              "$\\sigma$", "$\\sum_{1,1}$",
                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
colnames(CLPNGPNres.tab) <- rep(c('Mode', "HPD LB", "HPD UB"), 2)

#Create table
#kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE, 
#      caption = "Results, cross-validation mean and standard deviation, for the modified CL-PN and CL-GPN model")%>%
#add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))

```

```{r means, cache = TRUE, echo = FALSE}

theta_pred_PN   <- simplify2array(lapply(res_CLPN, "[[", "theta_pred"))
theta_pred_GPN  <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred"))
theta_pred_GPNM <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred"))

theta_pred_PN <- array(unlist(theta_pred_PN),
                       dim = c(nrow(theta_pred_PN[[1]]), ncol(theta_pred_PN[[1]]), length(theta_pred_PN)))
theta_pred_GPN <- array(unlist(theta_pred_GPN),
                        dim = c(nrow(theta_pred_GPN[[1]]), ncol(theta_pred_GPN[[1]]), length(theta_pred_GPN)))
theta_pred_GPNM <- array(unlist(theta_pred_GPNM),
                         dim = c(nrow(theta_pred_GPNM[[1]]), ncol(theta_pred_GPNM[[1]]), length(theta_pred_GPNM)))

means <- round(rbind(mean_circ(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
                     mean_circ(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
                     mean_circ(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)))*(180/pi), 2)
hpds  <- round(rbind(apply(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
                     apply(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
                     apply(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ))*(180/pi), 2)

means = cbind(means, hpds)

rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
colnames(means) <- c('Mode', "HPD LB", "HPD UB")
#kable(means, "latex", booktabs = T, escape = FALSE, digits = 2,
#      caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and GPN-SSN models")%>%
#kable_styling() %>%
#add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", #notation="alphabet")

```

\begin{table}

\caption{\label{tab:estCLGPN}Results, cross-validation mean and standard deviation, for the modified CL-PN and CL-GPN models}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{CL-PN} & \multicolumn{3}{c}{CL-GPN} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 0.09 (0.09) & 2.42 (0.15) & 1.90 (0.10) & 3.05 (0.17)\\
$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.08 (0.08) & 0.85 (0.12) & 0.45 (0.09) & 1.30 (0.15)\\
$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 0.04 (0.04) & 1.47 (0.05) & 1.17 (0.04) & 1.79 (0.05)\\
$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.04 (0.04) & 0.70 (0.08) & 0.46 (0.05) & 0.96 (0.07)\\
$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.01 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
\addlinespace
$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.00 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.05 (0.00)\\
$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.00 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.01) & -0.00 (0.00) & 0.06 (0.00)\\
$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.00 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.02 (0.25) & 1.83 (0.14) & 5.02 (0.41)\\
\addlinespace
$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.46 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
$\sum_{2,2}$ & NA (NA) & NA (NA) & NA (NA) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean in the CL-PN, CL-GPN and GPN-SSN models}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  & Mode & HPD LB & HPD UB\\
  \midrule
Cl-PN & 32.29 & 24.81 & 39.71\\
CL-GPN & 33.70 & 26.72 & 41.15\\
GPN-SSN & 35.30 & 28.31 & 43.10\\
\bottomrule
\multicolumn{4}{l}{\textsuperscript{a} Note that these means are based on}\\
\multicolumn{4}{l}{the posterior predictive}\\
\multicolumn{4}{l}{distribution for the intercepts}\\
\multicolumn{4}{l}{following (Wang \& Gelfand, 2013).}\\
\end{tabular}
\end{table}

First recall the regression equations predicting the linear outcome
\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i\]
\noindent and circular outcome
\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
  \mu_{i}^{I}  \\
\mu_{i}^{II}
 \end{pmatrix}=\begin{pmatrix}
  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
  \beta_0^{II} + \beta_1^{II}\text{SE}_i
 \end{pmatrix}.\]
\noindent  For both models, \(\hat{y}\) is the predicted strength of
interpersonal behavior, \(\hat{\boldsymbol{\mu}}\) is the predicted mean vector
of the type of interpersonal behavior and the \(\gamma\)'s and \(\beta\)'s are
intercepts and regression coefficients.\newline \indent Table \ref{tab:estCLGPN}
shows the results for the modified CL-PN and CL-GPN models fit to the teacher
dataset. The estimates in this table are the averages and standard deviations of
the 10 cross-validation estimates. We show both the estimated posterior mode and
the 95\% highest posterior density (HPD) interval for each parameter. The
posterior estimates for the \(\beta\)'s and \(\gamma\)'s are quite similar for
both models. We also see this similarity when we look at the predicted circular
and linear means. The predicted linear mean is equal to \(\gamma_0\) which is
equal to 0.38 and 0.37 in the CL-PN and CL-GPN models, respectively. The
predicted circular means can be computed from \(\beta_0^{I}\) and
\(\beta_0^{II}\) using a double arctangent function \(\mbox{atan2}(\beta_0^{II},
\beta_0^{I})\), see \eqref{PredVal}. If we do this at each iteration of the MCMC
sampler we get the posterior distribution of the circular mean in both models.
Table \ref{tab:means} shows the posterior mode of the estimated circular means,
which are very similar (32.29\(^\circ\) and 33.54\(^\circ\)).\newline
\indent Next we  investigate the effect of self-efficacy. In the CL-PN model we
can transform the regression parameters on the two components \(I\) and \(II\)
to one circular regression coefficient \(b_c\) using the methods described by
@CremersMulderKlugkist2017. The dashed line in Figure \ref{regline} shows the
predicted regression line from the CL-PN model. Here $b_c$ is the slope of this
line at the inflection point (black square) and its posterior mode is estimated
at 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation to the
circular regression line representing the slope at a specific point. Therefore
it is possible for the HPD interval to be wider than $2\pi$. In this case the
interval is much wider and covers 0, indicating there is no evidence to reject
the null-hypothesis of no effect.}. Even though the HPD of \(b_c\) includes 0,
indicating there is no evidence to reject the null-hypotheis of no effect, we
continue with its interpretation for eductional purposes. The interpretation of
$b_c$ is that at the inflection point an increase of 1 unit in self-efficacy
leads to an increase of \(1.67*(180/\pi) = 95.68^\circ\) in the type of
interpersonal behavior..However, as we can see in Figure \ref{regline} the
inflection point lies almost outside the range of the actual data. Instead of
looking at the inflection point we might compute the slope of the regression
line at the average self-efficacy. This parameter we call the slope at the mean
(\(SAM\)) [@CremersMulderKlugkist2017] and it is estimated at 0.02 (0.01; 0.04)
for our data. This means that at the average self-efficacy an increase of 1 unit
only leads to an increase of \(0.02*(180/\pi) = 1.15^\circ\) in the type of
interpersonal behavior. The HPD of the \(SAM\) does not include 0 which means
that the effect at the average self-efficacy can be distinguished from
0.\newline
\indent In the CL-GPN model we cannot compute circular regression coefficients
such as the \(b_c\) and \(SAM\) computed above  due to the fact that not only
the mean vector $\boldsymbol{\mu}$ but also the covariance matrix
$\boldsymbol{\Sigma}$ influences the predicted value on the circle. Instead, we
will compute posterior predictive distributions for the predicted circular
outcome of individuals scoring the minimum, maximum and median self-efficacy.
The modes and 95\% HPD intervals of these posterior predictive distributions are
\(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\),
\(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\). Note
that we display the modes and HPD intervals for the posterior predictive
distributions on the interval $[0^\circ, 360^\circ)$. Note that  $44.49^\circ =
404.49^\circ$ due to the periodicity of a circular variable. The posterior mode estimate of
$215.74^\circ$ thus lies within its HPD interval $(147.36^\circ, \: 44.49^\circ)$.The
HPD intervals of the three posterior predictive distributions overlap. Had they
not overlapped we could have concluded that as the self-efficacy increases, the
score of the teacher on the IPC moves counterclockwise. \newline
\indent The effect of self-efficacy on the strength of interpersonal behavior is
quantified by \(\gamma_1\) in both the CL-PN and CL-GPN models. This parameter
can be interpreted in the same way as in a usual regression
model. The HPD interval of  \(\gamma_1\) includes 0 which means
that there is not enough evidence for an effect of self-efficacy on the strength
of interpersonal behavior. Note that the lower bounds of the HPD interval of
this parameter are  however very close to zero.\newline
\indent  The relation between the circular and linear outcome, that is, between
type of interpersonal behavior and strength of interpersonal behavior, is
described by the parameters $\gamma_{\cos}$ and $\gamma_{\sin}$. The HPD
interval of \(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
models, meaning that the cosine component of the type of interpersonal behavior
has an effect on the strength of interpersonal behavior. In the teacher data the
sine and cosine components have a substantive meaning. In this case the
Communion (cosine) component of the IPC positively effects the strength of a
teachers' type of interpersonal behavior, in plain words: teachers exhibiting
interpersonal behavior types with higher communion scores (e.g., 'helpful' and
'understanding' in Figure 2) are stronger in their behavior.

```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

a1 <- CLPN.BI.m[1]
b1 <- CLPN.BI.m[2]
a2 <- CLPN.BII.m[1]
b2 <- CLPN.BII.m[2]
ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
ac <- atan2(a2 + b2*ax, a1 + b1*ax)

x <- seq(-5, 5, by = 0.1)

pred.AL <- 0.37 + 2*atan(-0.01*x)
pred.PN <- atan2(a2 + b2*x, a1 + b1*x)

plot(Dat$SEc, Dat$theta*(180/pi), bty = "n", yaxt = "n",
     xlim = c(-5, 2), ylab = "$\\theta$", xlab = "self-efficacy")
axis(side = 2,  at = c(-160, -80, 0, 80, 160))
points(x, pred.AL*(180/pi), type = "l", lty = 1)
points(x, pred.PN*(180/pi), type = "l", lty = 2)
points(ax, ac*(180/pi), pch = 15)

dev.off()
tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)

```

```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}

BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))

bc <- matrix(NA, 15001, 10)
SAM <- matrix(NA, 15001, 10)

for(i in 1:10){

  a1 <- BI[5000:20000,1,i]
  b1 <- BI[5000:20000,2,i]
  a2 <- BII[5000:20000,1,i]
  b2 <- BII[5000:20000,2,i]

  ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
  ac <- atan2(a2 + b2*ax, a1 + b1*ax)
  bc[,i] <- -tan(atan2(a2, a1)-ac)/ax
  SAM[,i] <- bc[i]/(1 + (bc[i]*(-ax))^2)

}

mean(apply(bc, 2, mode_est))
rowMeans(apply(bc, 2, hpd_est))

mean(apply(SAM, 2, mode_est))
rowMeans(apply(SAM, 2, hpd_est))

```

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
\caption{Plot showing circular regression lines for the effect of self-efficacy
as predicted by the Abe-Ley model (solid line) and CL-PN model (dashed line). The
black square indicates the inflection point of the circular regression line for
the CL-PN model.}
\label{regline}
\end{figure}

```{r ppreg, echo = FALSE, results = FALSE, eval = FALSE}
theta_pred_GPN.min <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.min"))
theta_pred_GPN.median <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.meadian"))
theta_pred_GPN.max <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.max"))

mean_circ(apply(theta_pred_GPN.min[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPN.median[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPN.max[5000:20000,], 2, mode_est_circ))*(180/pi)

apply(apply(theta_pred_GPN.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPN.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPN.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)

```

\subsubsection{The modified Abe-Ley model} \label{resAL}

```{r estAL, cache = TRUE, echo = FALSE}
#Get 2 dimensional arrays (parameter, fold)
resAL_arr <- simplify2array(lapply(resAL, "[[", "par"))

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for table
ALres.tab <- matrix(round(c(rowMeans(resAL_arr[1:4,]), mean(resAL_arr[6,]), 
                            mean(resAL_arr[5,]), mean(resAL_arr[7,]), 
                            apply(resAL_arr[1:4,], 1, sd), sd(resAL_arr[6,]), 
                            sd(resAL_arr[5,]), sd(resAL_arr[7,])), 2), 7, 2)

ALres.tab.new = cbind(apply(ALres.tab, 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                             sprintf("%.2f", round(x[2], 2)), ")")))

rownames(ALres.tab.new) <- c("$\\eta_0$", "$\\eta_1$",
                             "$\\nu_0$", "$\\nu_1$",
                             "$\\alpha$", "$\\kappa$", "$\\lambda$")

#Create table
#kable(ALres.tab.new, "latex", booktabs = T, escape = FALSE, digits = 2,
#      caption = "Results, cross-validation mean and standard deviation, for the modified Abe-Ley model")%>%
#add_header_above(c("Parameter" = 1, "ML-estimate" = 1))

```

\begin{table}

\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation, for the modified Abe-Ley model}
\centering
\begin{tabular}[t]{ll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
$\eta_0$ & 0.36 (0.02)\\
$\eta_1$ & -0.03 (0.01)\\
$\nu_0$ & 1.17 (0.02)\\
$\nu_1$ & 0.04 (0.02)\\
$\alpha$ & 3.66 (0.12)\\
$\kappa$ & 1.51 (0.08)\\
$\lambda$ & 0.70 (0.05)\\
\bottomrule
\end{tabular}
\end{table}

First recall the regression equations predicting the linear outcome
\[\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i)\]

\noindent and circular outcome
\[\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).\]

\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type
of interpersonal behavior and \(\hat{\beta}\) is the predicted scale
parameter of the strength of interpersonal behavior, while the \(\eta\)'s and
\(\nu\)'s are regression parameters.\newline
\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley model
fit to the teacher dataset. The estimates in this table are the averages and
standard deviations of the 10 cross-validation estimates. The estimate of the
circular mean at an average self-efficacy, \(\eta_0\), equals 0.36 radians or
20.62\(^\circ\). For the Abe-Ley model we can also investigate the effect of
self-efficacy on the type of interpersonal behavior. The solid line in
Figure \ref{regline} shows this effect. Unlike for the regression line of the
CL-PN model, the inflection point of the regression line of the Abe-Ley model
lies outside the x-range of the figure. \newline
\indent For the Abe-Ley model the conditional distribution for the linear
outcome is Weibull. This means that we can use methods from survival analsis to
interpret the effect of self-efficacy. In survival analysis a 'survival'
function is used in which time is plotted against the probability of survival of
subjects suffering from a specific medical condition. In our data however we
plot the strength on the IPC against the probability of a teacher having such a
strength. This probability is computed using the
'survival-function' \(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i -
\mu_i))^{1/\alpha}})\) with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
\ref{reglineweib}, we plot the survival function for the minimum, median and
maximum value of self-efficacy. We conclude that stronger interpersonal
behaviors are less probable. We also see that the relation between self-efficacy and the strength on the IPC is not linear; the probability of having a stronger
interpersonal behavior is higher for both the minimum and maximum self-efficacy compared to a median self-efficacy score. Note however that the circular outcome also influences the survival function. The relation between the type of interpersonal behavior and the strength of interpersonal behavior may thus influence the shape of the survival function.\newline

```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
summary(Dat$SEc)

datafig =  rbind(Dat[Dat$SEc == min(Dat$SEc),], Dat[Dat$SEc == median(Dat$SEc),], Dat[Dat$SEc == max(Dat$SEc),])

x = unique(datafig$SEc)
theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
y <- seq(0, 1, by = 0.05)
#h <- matrix(NA, length(y), length(x))
s <- matrix(NA, length(y), length(x))
scross <- array(NA, dim = c(length(y), length(x), 10))

for(k in 1:10){

  train <- get(paste("train.", k, sep = ""))

  datafig =  rbind(train[train$SEc == min(train$SEc),],
                   train[train$SEc == median(train$SEc),],
                   train[train$SEc == max(train$SEc),])
  x = unique(datafig$SEc)
  theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])

    for(j in 1:length(x)){

      for(i in 1:length(y)){
          mu <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*x[j])
          beta <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*x[j])*(1 - tanh(mean(resAL_arr[5,k]))*cos(theta[j] - mu))^(1/mean(resAL_arr[6,k]))
    #h[i,j] <- beta*mean(resAL_arr[6,])*y[i]^(mean(resAL_arr[6,])-1)
          s[i,j] <- exp(-mean(resAL_arr[6,k])*y[i]^beta)

    }

     scross[,,k] <- s

  }

}

s = apply(scross, c(1,2), mean)

tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
points(y, s[,2], type = "l", lty = 2)
points(y, s[,3], type = "l", lty = 3)
legend(0.7, 1, legend = c("min(SE)", "median(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")

dev.off()
tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)

```

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
\label{reglineweib}
\end{figure}

```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}

require(pracma)

alpha <- mean(resAL_arr[6,])
kappa <- mean(resAL_arr[5,])
lambda <- mean(resAL_arr[7,])

#R cannot compute legendre functions of non-integer degree
#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]

#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General

1/3.82
cosh(1.58)

```

\subsubsection{The modified GPN-SSN model} \label{resGPNSSN}

```{r estCLGPNM, cache = TRUE, echo = FALSE}
#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
lambda   <- simplify2array(lapply(res_CLGPNM, "[[", "lambda"))
Beta <- simplify2array(lapply(res_CLGPNM, "[[", "Beta"))
Betacon <- simplify2array(lapply(res_CLGPNM, "[[", "Betacon"))
Sigma <- simplify2array(lapply(res_CLGPNM, "[[", "Sigma"))
Sigmacon <- simplify2array(lapply(res_CLGPNM, "[[", "Sigmacon"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLGPNM.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Sigma13.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Sigma13.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
CLGPNM.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Sigma23.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Sigma23.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
CLGPNM.Sigma33.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
CLGPNM.Sigma33.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]

CLGPNM.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Sigma13.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Sigma13.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
CLGPNM.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Sigma23.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Sigma23.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
CLGPNM.Sigma33.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
CLGPNM.Sigma33.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]

CLGPNM.Sigmacon11.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Sigmacon11.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Sigmacon12.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Sigmacon12.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Sigmacon13.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Sigmacon13.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
CLGPNM.Sigmacon22.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Sigmacon22.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Sigmacon23.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Sigmacon23.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
CLGPNM.Sigmacon33.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
CLGPNM.Sigmacon33.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]

CLGPNM.Sigmacon11.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Sigmacon11.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Sigmacon12.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Sigmacon12.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Sigmacon13.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Sigmacon13.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
CLGPNM.Sigmacon22.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Sigmacon22.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Sigmacon23.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Sigmacon23.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
CLGPNM.Sigmacon33.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
CLGPNM.Sigmacon33.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]

CLGPNM.Beta11.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Beta11.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Beta21.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
CLGPNM.Beta21.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Beta12.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Beta12.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
CLGPNM.Beta22.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Beta22.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Beta13.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Beta13.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
CLGPNM.Beta23.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Beta23.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]

CLGPNM.Beta11.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Beta11.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Beta21.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
CLGPNM.Beta21.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Beta12.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Beta12.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
CLGPNM.Beta22.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Beta22.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Beta13.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Beta13.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
CLGPNM.Beta23.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Beta23.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]

CLGPNM.Betacon11.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Betacon11.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Betacon21.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
CLGPNM.Betacon21.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Betacon12.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Betacon12.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
CLGPNM.Betacon22.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Betacon22.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Betacon13.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Betacon13.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
CLGPNM.Betacon23.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Betacon23.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]

CLGPNM.Betacon11.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Betacon11.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Betacon21.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
CLGPNM.Betacon21.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Betacon12.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Betacon12.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
CLGPNM.Betacon22.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Betacon22.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Betacon13.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Betacon13.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
CLGPNM.Betacon23.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Betacon23.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]

CLGPNM.lambda.m   <- mean(apply(lambda[5000:20000,,], 2, mode_est))
CLGPNM.lambda.m.sd   <- sd(apply(lambda[5000:20000,,], 2, mode_est))
CLGPNM.lambda.hpd <- rowMeans(apply(lambda[5000:20000,,], 2, hpd_est))
CLGPNM.lambda.hpd.sd <- apply(apply(lambda[5000:20000,,], 2, hpd_est), 1, sd)

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
                CLGPNM.lambda.m)
hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
                CLGPNM.lambda.hpd[1])
hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
                CLGPNM.lambda.hpd[2])

modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
                   CLGPNM.lambda.m)
hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
                   CLGPNM.lambda.hpd[1])
hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
                   CLGPNM.lambda.hpd[2])

modes.GPN.sd  <- c(CLGPNM.Beta11.m.sd, CLGPNM.Beta12.m.sd, CLGPNM.Beta13.m.sd,
                   CLGPNM.Beta21.m.sd, CLGPNM.Beta22.m.sd, CLGPNM.Beta23.m.sd,
                   CLGPNM.Sigma11.m.sd, CLGPNM.Sigma22.m.sd, CLGPNM.Sigma33.m.sd,
                   CLGPNM.Sigma12.m.sd, CLGPNM.Sigma13.m.sd, CLGPNM.Sigma23.m.sd,
                  CLGPNM.lambda.m.sd)
hpd.LB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[1], CLGPNM.Beta12.hpd.sd[1], CLGPNM.Beta13.hpd[1],
                   CLGPNM.Beta21.hpd.sd[1], CLGPNM.Beta22.hpd.sd[1], CLGPNM.Beta23.hpd[1],
                   CLGPNM.Sigma11.hpd.sd[1], CLGPNM.Sigma22.hpd.sd[1], CLGPNM.Sigma33.hpd[1],
                   CLGPNM.Sigma12.hpd.sd[1], CLGPNM.Sigma13.hpd.sd[1], CLGPNM.Sigma23.hpd[1],
                   CLGPNM.lambda.hpd.sd[1])
hpd.UB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[2], CLGPNM.Beta12.hpd.sd[2], CLGPNM.Beta13.hpd.sd[2],
                   CLGPNM.Beta21.hpd.sd[2], CLGPNM.Beta22.hpd.sd[2], CLGPNM.Beta23.hpd.sd[2],
                   CLGPNM.Sigma11.hpd.sd[2], CLGPNM.Sigma22.hpd.sd[2], CLGPNM.Sigma33.hpd.sd[2],
                   CLGPNM.Sigma12.hpd.sd[2], CLGPNM.Sigma13.hpd.sd[2], CLGPNM.Sigma23.hpd.sd[2],
                   CLGPNM.lambda.hpd.sd[2])

modescon.GPN.sd  <- c(CLGPNM.Betacon11.m.sd, CLGPNM.Betacon12.m.sd, CLGPNM.Betacon13.m.sd,
                      CLGPNM.Betacon21.m.sd, CLGPNM.Betacon22.m.sd, CLGPNM.Betacon23.m.sd,
                      CLGPNM.Sigmacon11.m.sd, CLGPNM.Sigmacon22.m.sd, CLGPNM.Sigmacon33.m.sd,
                      CLGPNM.Sigmacon12.m.sd, CLGPNM.Sigmacon13.m.sd, CLGPNM.Sigmacon23.m.sd,
                      CLGPNM.lambda.m.sd)
hpdcon.LB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[1], CLGPNM.Betacon12.hpd.sd[1], CLGPNM.Betacon13.hpd.sd[1],
                      CLGPNM.Betacon21.hpd.sd[1], CLGPNM.Betacon22.hpd.sd[1], CLGPNM.Betacon23.hpd.sd[1],
                      CLGPNM.Sigmacon11.hpd.sd[1], CLGPNM.Sigmacon22.hpd.sd[1], CLGPNM.Sigmacon33.hpd.sd[1],
                      CLGPNM.Sigmacon12.hpd.sd[1], CLGPNM.Sigmacon13.hpd.sd[1], CLGPNM.Sigmacon23.hpd.sd[1],
                      CLGPNM.lambda.hpd.sd[1])
hpdcon.UB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[2], CLGPNM.Betacon12.hpd.sd[2], CLGPNM.Betacon13.hpd.sd[2],
                      CLGPNM.Betacon21.hpd.sd[2], CLGPNM.Betacon22.hpd.sd[2], CLGPNM.Betacon23.hpd.sd[2],
                      CLGPNM.Sigmacon11.hpd.sd[2], CLGPNM.Sigmacon22.hpd.sd[2], CLGPNM.Sigmacon33.hpd.sd[2],
                      CLGPNM.Sigmacon12.hpd.sd[2], CLGPNM.Sigmacon13.hpd.sd[2], CLGPNM.Sigmacon23.hpd.sd[2],
                      CLGPNM.lambda.hpd.sd[2])

#Create table
CLGPNMres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd,
                       hpd.UB.GPN, hpd.UB.GPN.sd, modescon.GPN, modescon.GPN.sd, 
                       hpdcon.LB.GPN, hpdcon.LB.GPN.sd, hpdcon.UB.GPN, hpdcon.UB.GPN.sd)

CLGPNMres.tab.new = cbind(apply(CLGPNMres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,7:8], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,9:10], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                            sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,11:12], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                             sprintf("%.2f", round(x[2], 2)), ")")))

rownames(CLGPNMres.tab.new) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
                                 "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
                                 "$\\beta_{1_s^{II}}$", "$\\beta_{1_y}$",
                                 "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
                                 "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
                                 "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
                                 "$\\lambda$")
colnames(CLGPNMres.tab.new) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

#kable(CLGPNMres.tab.new, "latex", booktabs = TRUE, escape = FALSE, digits = 2,
#      caption = "Results, cross-validation mean and standard deviation, for the GPN-SSN model")%>%
#add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))

```

```{r ppreggpnm, echo = FALSE, results = FALSE, eval = FALSE}
theta_pred_GPNM.min <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.min"))
theta_pred_GPNM.median <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.median"))
theta_pred_GPNM.max <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.max"))

mean_circ(apply(theta_pred_GPNM.min[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPNM.median[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPNM.max[5000:20000,], 2, mode_est_circ))*(180/pi)

dim(theta_pred_GPNM.median)

apply(apply(theta_pred_GPNM.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPNM.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPNM.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)

```

\begin{table}
\caption{\label{tab:estCLGPNM}Results, cross-validation mean and standard deviation, for the GPN-SSN model}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.10) & 1.75 (0.09) & 2.50 (0.11)\\
$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.33 (0.07) & 1.10 (0.05) & 1.57 (0.06)\\
$\beta_{0_y}$ & 0.33 (0.01) & 0.30 (0.30) & 0.36 (0.01) & 0.33 (0.01) & 0.30 (0.01) & 0.36 (0.01)\\
$\beta_{1_s^{I}}$ & 0.09 (0.01) & 0.05 (0.01) & 0.13 (0.01) & 0.60 (0.06) & 0.33 (0.05) & 0.90 (0.06)\\
$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.03) & 0.66 (0.04)\\
\addlinespace
$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.43 (0.14) & 1.72 (0.07) & 3.46 (0.13)\\
$\sum_{s_{2,2}}$ & 0.02 (0.00) & 0.02 (0.00) & 0.03 (0.00) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
$\sum_{y_{3,3}}$ & 0.03 (0.00) & 0.02 (0.02) & 0.04 (0.00) & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00)\\
$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.07 (0.06) & -0.20 (0.06) & 0.35 (0.06)\\
\addlinespace
$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.01) & 0.32 (0.01)\\
$\sum_{sy_{2,3}}$ & 0.01 (0.00) & 0.01 (0.01) & 0.02 (0.00) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
$\lambda$ & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01) & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01)\\
\bottomrule
\end{tabular}
\end{table}

First recall the regression equation predicting the circular and linear
outcome:

\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +\boldsymbol{\beta}_1\text{SE}_i,\]

\noindent where
\(\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t\),
\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
and
\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
The parameters \(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\),
\(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) are the intercepts and
regression coefficients for the circular outcome and \(\beta_{0_y}\) and
\(\beta_{1_y}\) are the intercept and regression coefficient for the
linear outcome.\newline
\indent Table \ref{tab:estCLGPNM} shows the results for the modified GPN-SSN model fit to
the teacher dataset. The estimates in this table are the averages and
standard deviations of the 10 cross-validation estimates. We show both
the estimated posterior mode and the 95\% highest posterior density
(HPD) interval for each parameter. The predicted circular means can be
computed from \(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) in a
similar fashion as for the CL-PN and CL-GPN models. Table \ref{tab:means} shows the
posterior mode of the estimated circular mean, which equals 35.30\(^\circ\) and hence
is very similar to those of the CL-PN and CL-GPN models.\newline
\indent For the same reason as in the CL-GPN model we cannot compute circular
regression coefficients for the effect of self-efficacy on the type of
interpersonal behavior such as the \(b_c\) and \(SAM\). Instead, we will again
compute posterior predictive distributions for the predicted circular outcome of
individuals scoring the minimum, maximum and median self-efficacy. The modes
and 95\% HPD intervals of these posterior predictive distributions are
\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.11^\circ, \: 72.02^\circ)\),
\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\). On a
circle the HPD intervals of the three posterior predictive distributions
overlap. Had they not overlapped we could have concluded that as the
self-efficacy increases, the score of the teacher on the IPC moves
counterclockwise. The effect of self-efficacy on the strength of interpersonal
behavior is quantified by \(\beta_{1_y}\), and we learn that for a 1 unit
increase in self-efficacy the strength of interpersonal behavior increases by
0.09. The average strength is quantified by \(\beta_{0_y}\) and equals
0.33.\newline
\indent To investigate the association between the linear and circular outcome
we look at the covariances between the linear outcome and the sine and cosine of
the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). Both
covariances, \(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are
different from zero, but the one of the cosine component is larger. This means
the correlation with the Communion component is larger and that teachers scoring
both high on Communion and Agency show stronger behavior. This is a slightly
different conclusion from the one in the CL-PN and CL-GPN models.

\subsubsection{Model fit}

In this section we will assess the overall fit of the cylindrical models to the
data via the PLSL criterion described in Section \ref{Modelfit}. Table
\ref{tab:ModelFit} shows the values of this criterion for the linear and
circular outcomes of the four models.\newline
\indent The CL-PN and CL-GPN models have the best out-of-sample predictive
performance for the linear outcome. They show roughly the same performance because
they model the linear outcome in the same way. Only the value of \(r\) in
\eqref{circlinlink} differs. We should note that even though the predictive
performance of the Abe-Ley model for the linear outcome is worst on average, the
standard deviation of the cross-validation estimates is rather large. This means
that in some samples, the Abe-Ley model shows a lower PLSL value than the
average of 25.49\newline
\indent The Abe-Ley model has the best out-of-sample predictive performance for
the circular outcome. This would suggest that for the circular variable a
slightly skewed distribution fits best. However, both the GPN-SSN  and the
CL-GPN models fit much worse even though the distribution for the circular
outcome in these models can also take a skewed shape. It should be noted that
the  standard deviation of the cross-validation estimates is rather large for
the Abe-Ley and the CL-GPN model. It is possible that these large standard
deviations for the PLSL are caused by the fact that they are computed for a
relatively small sample size, but this does not explain why the PLSL has a large
standard deviation for only a few cylindrical models and not for all.

```{r ModelFit, cache = TRUE, echo = FALSE}

CLGPNM.ll.circ <- simplify2array(lapply(res_CLGPNM, "[[", "ll.circ"))
CLGPNM.ll.lin  <- simplify2array(lapply(res_CLGPNM, "[[", "ll.lin"))
CLGPN.ll.circ  <- simplify2array(lapply(res_CLGPN, "[[", "ll.circ"))
CLGPN.ll.lin   <- simplify2array(lapply(res_CLGPN, "[[", "ll.lin"))
CLPN.ll.circ   <- simplify2array(lapply(res_CLPN, "[[", "ll.circ"))
CLPN.ll.lin    <- simplify2array(lapply(res_CLPN, "[[", "ll.lin"))

CLGPNM.ll.circ.m <- mean(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
CLGPNM.ll.lin.m  <- mean(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
CLGPN.ll.circ.m  <- mean(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
CLGPN.ll.lin.m   <- mean(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
CLPN.ll.circ.m   <- mean(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
CLPN.ll.lin.m    <- mean(-2*apply(CLPN.ll.lin[5:20,], 2, mean))

CLGPNM.ll.circ.sd <- sd(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
CLGPNM.ll.lin.sd  <- sd(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
CLGPN.ll.circ.sd  <- sd(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
CLGPN.ll.lin.sd   <- sd(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
CLPN.ll.circ.sd   <- sd(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
CLPN.ll.lin.sd    <- sd(-2*apply(CLPN.ll.lin[5:20,], 2, mean))

AL.ll.circ <- simplify2array(lapply(resAL, "[[", "ll.circ"))
AL.ll.lin <- simplify2array(lapply(resAL, "[[", "ll.lin"))

AL.ll.circ.m <- mean(AL.ll.circ)
AL.ll.lin.m  <- mean(AL.ll.lin)

AL.ll.circ.sd <- sd(AL.ll.circ)
AL.ll.lin.sd  <- sd(AL.ll.lin)

PLSL.m <- round(rbind(c(CLPN.ll.circ.m, CLGPN.ll.circ.m, AL.ll.circ.m, CLGPNM.ll.circ.m),
                      c(CLPN.ll.lin.m, CLGPN.ll.lin.m, AL.ll.lin.m, CLGPNM.ll.lin.m)), 2)
PLSL.sd <- round(rbind(c(CLPN.ll.circ.sd, CLGPN.ll.circ.sd, AL.ll.circ.sd, CLGPNM.ll.circ.sd),
                      c(CLPN.ll.lin.sd, CLGPN.ll.lin.sd, AL.ll.lin.sd, CLGPNM.ll.lin.sd)), 2)
PLSL.sd <- apply(PLSL.sd, 1:2, function(x) paste0("(", sprintf("%.2f", round(x, 2)), ")"))

PLSL <- data.frame(PLSL.m[1,], PLSL.sd[1,], PLSL.m[2,], PLSL.sd[2,])

rownames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN")
colnames(PLSL) <- c("mean", "sd", "mean", "sd")

#kable(PLSL, "latex", booktabs = T, escape = FALSE,
#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four #cylindrical models")%>%
#add_header_above(c("Model" = 1, "Circular" = 2, "Linear" = 2))

```

\begin{table}

\caption{\label{tab:ModelFit}PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four cylindrical models}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{Model} & \multicolumn{2}{c}{Circular} & \multicolumn{2}{c}{Linear} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
  & mean & sd & mean & sd\\
\midrule
CL-PN & 82.96 & (9.47) & -17.65 & (3.70)\\
CL-GPN & 85.22 & (18.12) & -18.12 & (3.57)\\
Abe-Ley & 31.97 & (22.07) & 25.49 & (17.46)\\
GPN-SSN & 106.38 & (8.84) & -2.14 & (6.78)\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}\label{Discussion}

In this paper we modified four models for cylindrical data in such a way that
they include a regression of both the linear and circular outcome onto a
set of covariates. Subsequently we have shown how these four methods can
be used to analyze a dataset on the interpersonal behavior of teachers.
In this final section we will comment on the differences between these models, the
results from the analysis of the teacher data and how cylindrical models
improve the analysis of such cylindrical data.\newline
\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
best.  In the
CL-GPN and GPN-SSN models the interpretation of the parameters of the
circular outcome component is not straightforward, if at all possible.
This is caused by the fact that in addition to the mean vector the
covariance matrix of the GPN distribution affects the location of the
circular data, making it difficult to compute regression coefficients on
the circle. @wang2012directional state that Monte Carlo integration
can be used to compute a circular mean and variance for the GPN
distribution. In future research, this solution might be applied to the
methods of @CremersMulderKlugkist2017 in order to compute circular
coefficients for GPN models.\newline
\indent In terms of flexibility the GPN-SSN model scores best. Multiple
linear and circular outcomes can be included and we can thus apply the
model to multivariate cylindrical data. In addition the GPN-SSN, the
CL-GPN and CL-PN models are extendable to a mixed-effects structure and
can thus also be fit to longitudinal data (see @nunez2014bayesian and @hernandez2016general for
hierarchical/mixed-effects models for the PN and GPN distributions
respectively). For the Abe-Ley model this may also be possible but has
not been done in previous research for the conditional distribution of
its circular outcome (sine-skewed von
Mises). Concerning asymmetry, both the
GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
the distributions of both the linear and circular outcome, while the CL-GPN
model permits an asymmetric  circular outcome.\newline
\indent To investigate model fit for the teacher data, we assessed
out-of-sample predictive performance for both the linear and circular
outcome. The CL-PN and CL-GPN models have the best fit for the linear
outcome while the Abe-Ley model fits best  the circular outcome.
Differences in fit, in addition to being a result of different
distributional assumptions, may also be caused by the way in which the
relation between the linear and circular outcome is modeled. Whereas in
both the Abe-Ley and GPN-SSN models the distribution of the linear
outcome is conditioned on the circular outcome and vice versa, the
distribution of the circular outcome in the CL-PN and CL-GPN models is
independent of the linear outcome. In these models the circular outcome
is regressed onto the linear outcome.\newline
\indent The four cylindrical models that were modified to the regression
context in this paper are not the only cylindrical distributions
available from the literature. Other interesting cylindrical distributions have been introduced by
@fernandez2007models, @kato2008dependent and @sugasawa2015flexible (for more references we
refer to Chapter 2 of @ley2017modern). In the present study we
have decided not to include these distributions for reasons of space,
complexity of the models and ease of implementing a regression
structure. In future research however it would be interesting to
investigate other types of cylindrical distributions as well in order to
compare the interpretability, flexibility and model fit to the models
developed in the present study.\newline
\indent We conclude the paper on a general note regarding cylindrical models.
They  offer new insights into data of a cylindrical nature in psychology.
Concerning the example used in this paper, the advantage of cylindrical data
analysis is that we were able to analyze all circular and linear information in
the data simultaneously. In previous research, the two components of the
interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
separately. Such an approach also provides information about the strength of
teachers' score on Agency and Communion, yet a large portion of information
about the combination of Agency and Communion, which describes the kind of
behavior that is observed,  gets lost. A first solution to include both
dimensions as a circular variable in data analysis was described by
@Cremers2018Assessing. A downside of that analysis was that information about
the strength of the specific type of interpersonal behavior could not be
retained. In the present study, we have shown how using cylindrical models can
simultaneously model the information about the type of and strength of
interpersonal behavior and how these are influenced by teachers' self-efficacy
in classroom management. The results of the present study therefore provide an
answer to an often stated problem in data analysis of interpersonal circumplex
data.

\section*{References}
<div id="refs"></div>

### 2018-09-24 Manuscript/ManuscriptMarkdownCV.tex
Manuscript update

--- 2018-09-24 Manuscript/ManuscriptMarkdownCV.Rmd
@@
----
-title: "Regression models for Cylindrical data in Psychology"
-author: #|
- # | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
- # | $^1$Department of Methodology and Statistics, Utrecht University
- # | $^2$TNO
-  #| $^3$Department of Education, Utrecht University
-  #| $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
-
-bibliography: CircularData.bib
-csl: apa.csl
-output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    number_sections: yes
-  html_document: default
-  word_document: default
-fontsize: 12pt
-geometry: margin=1in
-header-includes:
-  - \usepackage{multirow}
-  - \usepackage{appendix}
-  - \usepackage{color}
-  - \usepackage{hyperref}
-  - \usepackage{subcaption}
-  - \setlength\parindent{24pt}
-  - \usepackage[nodisplayskipstretch]{setspace}\doublespacing
-  #- \usepackage{endfloat}
-  - \DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
-
-keywords: bla vla
-documentclass: article
-pandoc_args: --natbib
-abstract: |
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Regression models for Cylindrical data in Psychology},
+            pdfkeywords={bla vla},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{Regression models for Cylindrical data in Psychology}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+
+\usepackage{multirow}
+\usepackage{appendix}
+\usepackage{color}
+\usepackage{hyperref}
+\usepackage{subcaption}
+\setlength\parindent{24pt}
+\usepackage[nodisplayskipstretch]{setspace}\doublespacing
+\DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
+
+\begin{document}
+\maketitle
+\begin{abstract}
 Cylindrical data are multivariate data which consist of a directional,
@@
 interpersonal behavior of teachers.
----
-```{r inlude = FALSE, echo = FALSE}
-
-knitr::opts_chunk$set(message = FALSE, warning = FALSE)
-```
-
-```{r, echo = FALSE}
-library(circular)
-library(bpnreg)
-library(haven)
-library(tikzDevice)
-library(plotrix)
-library(tidyr)
-library(dplyr) #sample_n()
-library(MASS) #mvrnorm()
-library(kableExtra) #latex tables
-library(plotrix)
-library(caret) #cross-validation
-```
+\end{abstract}
 
@@
 
-Cylindrical data are data that consist of a linear variable and a directional
-variable. In this paper, the directional variable is circular, meaning that it
-consists of a single angle instead of a set of angles. A circular variable is
-different from a linear variable in the sense that it is measured on a different
-scale. Figure \ref{circline} shows the difference between a circular scale
-(right) and a linear scale (left). The most important difference is that on a
-circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in
-fact represent the same number while on a linear scale the two ends, \(-\infty\)
-and \(\infty\), are not connected and consequently the values 0\(^\circ\) and
-360\(^\circ\) are located on different places on the scale. This difference
-requires us to use special statistical methods for circular variables (see e.g.
-@fisher1995statistical for an introduction to circular data and
-@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
-  elaborate overview). For instance, a notion as simple as the sample average
-  needs to be re-defined. As is the case for circular data, the analysis of
+Cylindrical data are data that consist of a linear variable and a
+directional variable. In this paper, the directional variable is
+circular, meaning that it consists of a single angle instead of a set of
+angles. A circular variable is different from a linear variable in the
+sense that it is measured on a different scale. Figure \ref{circline}
+shows the difference between a circular scale (right) and a linear scale
+(left). The most important difference is that on a circular scale the
+datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in fact
+represent the same number while on a linear scale the two ends,
+\(-\infty\) and \(\infty\), are not connected and consequently the
+values 0\(^\circ\) and 360\(^\circ\) are located on different places on
+the scale. This difference requires us to use special statistical
+methods for circular variables (see e.g. Fisher (1995) for an
+introduction to circular data and Mardia \& Jupp (2000), Jammalamadaka
+\& Sengupta (2001) and Ley \& Verdebout (2017) for a more elaborate
+overview). For instance, a notion as simple as the sample average needs
+to be re-defined. As is the case for circular data, the analysis of
 cylindrical data also requires special methods.\newline
-\indent Cylindrical data occur in several fields of research, such as for
-instance in meteorology [@garcia2013exploring], ecology [@garcia2014test] or
-marine research [@lagona2015hidden]. However, also in psychology several types
-of cylindrical data can be found. For example, in research on human navigation
-in the field of cognitive psychology both the distance, a linear variable, and
-the direction, a circular variable, of movement are of interest
-[@chrastil2017rotational]. In eye-tracking research, saccade data are an example
-of cylindrical data, because both the direction (\emph{i.e.}, the circular variable)
-and the duration (\emph{i.e.}, the linear variable) of the saccades are of interest
-(for a review of eye-tracking research see  @rayner200935th).\newline
+\indent Cylindrical data occur in several fields of research, such as
+for instance in meteorology (García-Portugués, Crujeiras, \&
+González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
+González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
+Maruotti, \& Cosoli, 2015). However, also in psychology several types of
+cylindrical data can be found. For example, in research on human
+navigation in the field of cognitive psychology both the distance, a
+linear variable, and the direction, a circular variable, of movement are
+of interest (Chrastil \& Warren, 2017). In eye-tracking research,
+saccade data are an example of cylindrical data, because both the
+direction (\emph{i.e.}, the circular variable) and the duration
+(\emph{i.e.}, the linear variable) of the saccades are of interest (for
+a review of eye-tracking research see Rayner (2009)).\newline
 \indent The type of data that is used in the present study are also
@@
 instance, data from the interpersonal circumplex as used in personality
-psychology are by definition of a cylindrical nature (see Section \ref{Example}
-for a more detailed explanation).\newline
-\indent In the literature, several methods have been put forward to model the
-relation between the linear and circular component of a cylindrical variable.
-Some of these are based on regressing the linear component onto the circular
-component using the following type of relation: \[y = \beta_0 +
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
-linear component and \(\theta\) the circular component [@mardia1978model;
-@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
-different way, e.g. by specifying a multivariate model for several linear and
-circular variables and modelling their covariance matrix
-[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
-example, @abe2017tractable introduce a cylindrical distribution based on a
-Weibull distribution for the linear component and a sine-skewed von Mises
-distribution for the circular component and link these through their respective
-shape and concentration parameters. However, none of the methods that have been
-proposed thus far include additional covariates onto which both the circular and
-linear component are regressed.\newline
-\indent Our aim in this paper is to fill this gap in the literature by adapting
-four existing cylindrical models in such a way that they include a regression of
-both the linear and circular component of a cylindrical variable onto a set of
-covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. Additionally, we will show how a
-correct statistical treatment of such cylindrical data can lead to new insights.
-We will do this for the teacher data, a dataset from the field of educational
-psychology. In the teacher data, apart from modelling the relation between the
-linear and circular component of a cylindrical variable we would also like to
-predict the two components from a set of covariates in a regression model.
-
-The paper is organized as follows: Section \ref{Example} describes the teacher
-data, while Section \ref{Models} presents the four cylindrical models and our
-associated adaptations to new regression models. In that same section we also
-discuss the model fit criterion that we will use in Section \ref{DataAnalysis}
-for the comparison of the four models. A detailed data analysis with interesting
-new insights is also provided in Section \ref{DataAnalysis}. We conclude the
-paper with a discussion in Section \ref{Discussion}, and the Supplementary Material collects
-the technical details of the MCMC procedures.
+psychology are by definition of a cylindrical nature (see Section
+\ref{Example} for a more detailed explanation).\newline
+\indent In the literature, several methods have been put forward to
+model the relation between the linear and circular component of a
+cylindrical variable. Some of these are based on regressing the linear
+component onto the circular component using the following type of
+relation: \[y = \beta_0 +
+\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is
+the linear component and \(\theta\) the circular component (Johnson \&
+Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \&
+Jona-Lasinio, 2015). Others model the relation in a different way,
+e.g.~by specifying a multivariate model for several linear and circular
+variables and modelling their covariance matrix (Mastrantonio, 2018) or
+by proposing a joint cylindrical distribution. For example, Abe \& Ley
+(2017) introduce a cylindrical distribution based on a Weibull
+distribution for the linear component and a sine-skewed von Mises
+distribution for the circular component and link these through their
+respective shape and concentration parameters. However, none of the
+methods that have been proposed thus far include additional covariates
+onto which both the circular and linear component are regressed.\newline
+\indent Our aim in this paper is to fill this gap in the literature by
+adapting four existing cylindrical models in such a way that they
+include a regression of both the linear and circular component of a
+cylindrical variable onto a set of covariates. From now on we will
+therefore refer to the components of the cylindrical variable as outcome
+components. Additionally, we will show how a correct statistical
+treatment of such cylindrical data can lead to new insights. We will do
+this for the teacher data, a dataset from the field of educational
+psychology. In the teacher data, apart from modelling the relation
+between the linear and circular component of a cylindrical variable we
+would also like to predict the two components from a set of covariates
+in a regression model.
+
+The paper is organized as follows: Section \ref{Example} describes the
+teacher data, while Section \ref{Models} presents the four cylindrical
+models and our associated adaptations to new regression models. In that
+same section we also discuss the model fit criterion that we will use in
+Section \ref{DataAnalysis} for the comparison of the four models. A
+detailed data analysis with interesting new insights is also provided in
+Section \ref{DataAnalysis}. We conclude the paper with a discussion in
+Section \ref{Discussion}, and the Supplementary Material collects the
+technical details of the MCMC procedures.
 
@@
 
-The motivating example for this article comes from the field of educational
-psychology and was collected for the studies on classroom climate of
-@vanderWant2015role, @Claessens2016side and @pennings2018interpersonal. An
-indicator of the quality of the classroom climate is the students' perception of
-their teachers' interpersonal behavior. These interpersonal perceptions, both in
-educational psychology as well as in other areas of psychology, can be measured
-using circumplex measurement instruments (see @horowitz2010handbook for an
-overview of many such instruments).\newline
+The motivating example for this article comes from the field of
+educational psychology and was collected for the studies on classroom
+climate of \VANDER{Want}{Van der}{van der} Want (2015), Claessens (2016)
+and Pennings et al. (2018). An indicator of the quality of the classroom
+climate is the students' perception of their teachers' interpersonal
+behavior. These interpersonal perceptions, both in educational
+psychology as well as in other areas of psychology, can be measured
+using circumplex measurement instruments (see Horowitz \& Strack (2011)
+for an overview of many such instruments).\newline
 \indent The circumplex data used in this paper are measured using the
-Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] which is
-one such circumplex measurement instrument. The QTI is designed to measure
-student perceptions of their teachers' interpersonal behavior and contains items
-that load on two interpersonal dimensions: Agency and Communion. Agency refers
-to the degree of power or control a teacher exerts in interaction with his/her
-students. Communion refers to the degree of friendliness or affiliation a
-teacher conveys in interaction with his/her students. The loadings on the two
-dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
-this space are characterized by different teacher behavior, e.g. 'helpful' or
-'uncertain'. This two-dimensional space is called the interpersonal
-circle/circumplex (IPC). The IPC is ``a continuous order with no beginning or
-end'' [@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering
-and the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular
-variable.\newline
-\indent @Cremers2018Assessing explain the circular nature of the IPC data and
-analyze them as such using a circular regression model. The two-dimension scores
-Agency and Communion can be converted to a circular score using the two-argument
-arctangent function in \eqref{PredVal}, where \(A\) represents a score on the
-Agency dimension and \(C\) represents a score on the Communion dimension
+Questionnaire on Teacher Interaction (QTI) (Wubbels, Brekelmans, Brok,
+\& Tartwijk, 2006) which is one such circumplex measurement instrument.
+The QTI is designed to measure student perceptions of their teachers'
+interpersonal behavior and contains items that load on two interpersonal
+dimensions: Agency and Communion. Agency refers to the degree of power
+or control a teacher exerts in interaction with his/her students.
+Communion refers to the degree of friendliness or affiliation a teacher
+conveys in interaction with his/her students. The loadings on the two
+dimensions of the QTI can be placed in a two-dimensional space formed by
+Agency (vertical) and Communion (horizontal), see Figure \ref{QTI}.
+Different parts of this space are characterized by different teacher
+behavior, e.g. `helpful' or `uncertain'. This two-dimensional space is
+called the interpersonal circle/circumplex (IPC). The IPC is ``a
+continuous order with no beginning or end'' (Gurtman, 2009, p. 2). We
+call such ordering a circumplex ordering and the IPC is therefore often
+called the interpersonal circumplex. The ordering also implies that
+scores on the IPC could be viewed as a circular variable.\newline
+\indent Cremers, Mainhard, \& Klugkist (2018a) explain the circular
+nature of the IPC data and analyze them as such using a circular
+regression model. The two-dimension scores Agency and Communion can be
+converted to a circular score using the two-argument arctangent function
+in \eqref{PredVal}, where \(A\) represents a score on the Agency
+dimension and \(C\) represents a score on the Communion dimension
+
 \begin{equation}\label{PredVal}
@@
 \end{equation}
-The resulting circular variable \(\theta\) can then be modelled and takes values
-in the interval $[0, 2\pi)$. However, when two-dimensional data are converted
-to the circle we lose some information, namely the length of the two-dimensional
-vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\).
-This length represents the strength of the type of interpersonal behavior a
-teacher shows towards his/her students and can be considered as the linear
-variable in a cylindrical model, allowing us to model a circular variable
-\(\theta\) together with the linear variable corresponding to \(\mid\mid (A, C)^t
-\mid\mid\). This leads to an improved analysis of interpersonal circumplex data
-as we take all information into account. In the next section we introduce
-several models that can be used for a more accurate and informative regression
-analysis on the teacher data. First however we will provide descriptives for our
-data set.
+
+The resulting circular variable \(\theta\) can then be modelled and
+takes values in the interval \([0, 2\pi)\). However, when
+two-dimensional data are converted to the circle we lose some
+information, namely the length of the two-dimensional vector
+\((A, C)^t\), \emph{i.e.}, its Euclidean norm
+\(\mid\mid (A, C)^t \mid\mid\). This length represents the strength of
+the type of interpersonal behavior a teacher shows towards his/her
+students and can be considered as the linear variable in a cylindrical
+model, allowing us to model a circular variable \(\theta\) together with
+the linear variable corresponding to \(\mid\mid (A, C)^t \mid\mid\).
+This leads to an improved analysis of interpersonal circumplex data as
+we take all information into account. In the next section we introduce
+several models that can be used for a more accurate and informative
+regression analysis on the teacher data. First however we will provide
+descriptives for our data set.
 
@@
 
-```{r loaddata, return = FALSE, echo = FALSE}
-Dat <- read_spss("Data Heleen/mergeddata.sav")
-
-#only select the first measurement occasion
-Dat <- subset(Dat, Time == 0)
-
-#sort on class size
-Dat <- Dat[order(-Dat$nklas), ]
-
-#Now remove duplicates 
-#(because dataframe is sorted on class size the smallest classes are removed)
-
-duplicates <- duplicated(Dat[,c("docnr", "Time")])
-Dat <- Dat[!duplicates,]
-
-#center and remove missings for self-efficacy
-Dat <- Dat %>% drop_na(Efficacy_CM)
-Dat$SEc <- Dat$Efficacy_CM- mean(Dat$Efficacy_CM)
-Dat$theta <- atan2(Dat$lAgency, Dat$lcommunion)
-Dat$y <- sqrt(Dat$lAgency^2 + Dat$lcommunion^2)
-
-set.seed(30)
-
-#Create holdout and training samples for cross-validation
-
-vec <-  seq(1:nrow(Dat))
-flds <- createFolds(vec, k = 10, list = TRUE, returnTrain = FALSE)
-
-for(i in 1:10) {
-
- #Create name for holdout data 
- nam.h <- paste("hold.", i, sep = "")
-
- #Get holdout dataset for fold i
- assign(nam.h, Dat[flds[[i]],])
-
- #Create name for training data
- nam.tr <- paste("train.", i, sep = "")
-
- #Get training dataset for fold i
- assign(nam.tr, Dat[-flds[[i]],])
-
-}
-
-```
-
-```{r dataplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/dataplot.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-plot(Dat$y, Dat$theta*(180/pi), 
-     ylab = "IPC location", xlab = "IPC strength",
-     bty = "n", ylim = c(-180, 180))
-
-dev.off()
-tools::texi2dvi("Plots/dataplot.tex", pdf=TRUE)
-
-```
-
-```{r descripitves, echo = FALSE, results = FALSE}
-#sample size
-nrow(Dat)
-
-#summary statistics
-summary(Dat$Efficacy_CM)
-summary(Dat$y)
-summary(as.circular(Dat$theta%%(2*pi)))
-
-sd(Dat$Efficacy_CM)
-sd(Dat$y)
-
-```
-
-The teacher data was collected between 2010 and 2015 and contains several
-repeated measures on the IPC of 161 teachers. Measurements were obtained using
-the QTI and taken in different years and classes. For this paper we only
-consider one measurement, the first occasion (2010) and largest class if data
-for multiple classes were available. In addition to the score on the IPC, the
-circular outcome, and the strength of the score on the IPC, the linear outcome,
-a teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
-covariate in the analysis. After listwise deletion of missings ($3$ in total,
-only for the self-efficacy) we have a sample of 148 teachers. Table
+The teacher data was collected between 2010 and 2015 and contains
+several repeated measures on the IPC of 161 teachers. Measurements were
+obtained using the QTI and taken in different years and classes. For
+this paper we only consider one measurement, the first occasion (2010)
+and largest class if data for multiple classes were available. In
+addition to the score on the IPC, the circular outcome, and the strength
+of the score on the IPC, the linear outcome, a teachers' self-efficacy
+(\verb|SE|) concerning classroom management is used as covariate in the
+analysis. After listwise deletion of missings (\(3\) in total, only for
+the self-efficacy) we have a sample of 148 teachers. Table
 \ref{Tableteacherdescriptives} shows descriptives for the dataset. Here
-$\hat{\rho}$ is a sample estimate for the circular concentration where a value
-of 0 means that the data is not concentrated at all, \emph{i.e.} spread over the
-entire circle, and a value of 1 means that all data is concentrated at a single
-point on the circle. Figure \ref{dataplot} is a scatterplot showing the relation
-between the linear and circular outcome of the teacher data.
+\(\hat{\rho}\) is a sample estimate for the circular concentration where
+a value of 0 means that the data is not concentrated at all, \emph{i.e.}
+spread over the entire circle, and a value of 1 means that all data is
+concentrated at a single point on the circle. Figure \ref{dataplot} is a
+scatterplot showing the relation between the linear and circular outcome
+of the teacher data.
 
@@
 and \(\Theta\). The first two models are based on a construction by
-@mastrantonio2015bayesian, while the other models are extensions of the
-models from @abe2017tractable and @mastrantonio2018joint.
+Mastrantonio et al. (2015), while the other models are extensions of the
+models from Abe \& Ley (2017) and Mastrantonio (2018).
 
@@
 
-Following @mastrantonio2015bayesian we consider in this section two models where
-the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
-and $q$ covariates is specified as
+Following Mastrantonio et al. (2015) we consider in this section two
+models where the relation between \(\Theta \in [0, 2\pi)\) and
+\(Y\in (-\infty, + \infty)\) and \(q\) covariates is specified as
+
 \begin{equation}\label{circlinlink}
@@
 \end{equation}
-\noindent where the random variable \(R\geq0\) will be introduced below, the
-error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
-\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
-intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
-covariates. In both of these models the
-conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
-by
+
+\noindent where the random variable \(R\geq0\) will be introduced below,
+the error term \(\epsilon \sim N(0, \sigma^2)\) with variance
+\(\sigma^2>0\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
+the intercept and regression coefficients and \(x_1, \dots, x_q\) are
+the \(q\) covariates. In both of these models the conditional
+distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given by
+
 \begin{equation}\label{ycondtheta}
@@
 \end{equation}
-\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
-\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
-\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
-\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
-in its location part. For the teacher dataset, the regression equation for the
-linear outcome in the CL-PN and CL-GPN model is the following: \[\hat{y}_i =
+
+\noindent where
+\(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
+\(r \geq 0\). The linear outcome thus has a normal distribution
+conditional on \(\Theta\) and \(R\) and contains already linear
+covariates \(x_1, \dots, x_q\) in its location part. For the teacher
+dataset, the regression equation for the linear outcome in the CL-PN and
+CL-GPN model is the following: \[\hat{y}_i =
 \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i +
-\gamma_1\text{SE}_i,\]
-\noindent where $\text{SE}_i$ is the self-efficacy score of one individual $i = 1, \dots, n$ where $n$ is the sample size.\newline
-\indent For the circular outcome we assume either a projected normal (PN) or a general
-projected normal (GPN) distribution. These distributions arise from the radial
-projection of a distribution defined on the plane onto the circle. The relation
-between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
-outcome \(\Theta\) is defined as follows
+\gamma_1\text{SE}_i,\] \noindent where \(\text{SE}_i\) is the
+self-efficacy score of one individual \(i = 1, \dots, n\) where \(n\) is
+the sample size.\newline
+\indent For the circular outcome we assume either a projected normal
+(PN) or a general projected normal (GPN) distribution. These
+distributions arise from the radial projection of a distribution defined
+on the plane onto the circle. The relation between a bivariate vector
+\(\boldsymbol{S}\) in the plane and the circular outcome \(\Theta\) is
+defined as follows
+
 \begin{equation}\label{projection}
@@
 \end{equation}
-\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
-the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
-\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
-assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
-$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1
-\end{bmatrix}\), \(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
-@hernandez2016general). This leads to the
-circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
-will now detail how we modify both cylindrical distributions to also incorporate
-covariates for the circular part.
+
+\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean
+norm of the bivariate vector \(\boldsymbol{S}\). In the PN distribution
+we assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\)
+and in the GPN we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
+\(\boldsymbol{\mu} \in \mathbb{R}^2\),
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
+\(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
+Hernandez-Stumpfhauser, Breidt, \& Woerd (2016)). This leads to the
+circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN)
+distributions. We will now detail how we modify both cylindrical
+distributions to also incorporate covariates for the circular part.
 
@@
 
-Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
-PN distribution equals
+Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
+density of \(\Theta\) and \(R\) for the PN distribution equals
+
 \begin{equation}\label{pnreg}
@@
 \end{equation}
-\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
-\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
-\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
-the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
-\(n\) is the sample size, are generated independently from the distribution with
-density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
-is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
-where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
-contains the covariate values and the value 1 to estimate an intercept and
-\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
-the regression coefficients and intercepts. Note however that the dimensions of
-\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{II}\) need not necessarily
-be the same and we are thus allowed to have a different set of predictor
-variables and vectors \(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for
-the two components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
+
+\noindent where
+\(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
+and \(r\) is the same as in \eqref{circlinlink} and \eqref{ycondtheta}
+and is defined in \eqref{projection}. In a regression setup the outcomes
+\(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where \(n\) is
+the sample size, are generated independently from the distribution with
+density \eqref{pnreg}. The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
+vector \(\boldsymbol{z}_i\) is a vector of dimension \(p + 1\) that
+contains the covariate values and the value 1 to estimate an intercept
+and
+\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
+contains the regression coefficients and intercepts. Note however that
+the dimensions of \(\boldsymbol{\beta}^{I}\) and
+\(\boldsymbol{\beta }^{II}\) need not necessarily be the same and we are
+thus allowed to have a different set of predictor variables and vectors
+\(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for the two
+components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
 regression equation for the circular outcome in the CL-PN model is
@@
 
-Following @wang2012directional and @hernandez2016general the joint density of
-\(R\) and \(\Theta\) for the GPN distribution equals
+Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
+(2016) the joint density of \(R\) and \(\Theta\) for the GPN
+distribution equals
+
 \begin{equation}\label{gpnreg}
@@
 \end{equation}
-\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
-\rho^2 & \rho\\ \rho & 1 \end{bmatrix}\). In a regression setup the outcomes
-\(\theta_i\) and \(r_i\) for each individual are generated independently from
-\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
-defined in the same way via covariates as for the modified CL-PN distribution.
-Note in contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
-\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the same
-predictors for both components of \(\boldsymbol{\mu}_i\) in the CL-GPN model.
-This is due to the fact that the variance-covariance matrix
-\(\boldsymbol{\Sigma}\) is no longer identity. For the teacher dataset, the
-regression equation for the circular outcome in the CL-GPN model is the same as
-in the CL-PN model.
+
+\noindent where we recall that
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\).
+In a regression setup the outcomes \(\theta_i\) and \(r_i\) for each
+individual are generated independently from \eqref{gpnreg}. The mean
+vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the same
+way via covariates as for the modified CL-PN distribution. Note in
+contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
+\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the
+same predictors for both components of \(\boldsymbol{\mu}_i\) in the
+CL-GPN model. This is due to the fact that the variance-covariance
+matrix \(\boldsymbol{\Sigma}\) is no longer identity. For the teacher
+dataset, the regression equation for the circular outcome in the CL-GPN
+model is the same as in the CL-PN model.
 
@@
 
-Both cylindrical models introduced here are estimated using Markov Chain Monte
-Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in the
-Supplementary Material.
+Both cylindrical models introduced here are estimated using Markov Chain
+Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
+Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
+regression of the circular outcome. A detailed description of the
+Bayesian estimation and MCMC samplers can be found in the Supplementary
+Material.
 
@@
 
-This model is an extension of the cylindrical model introduced in
-@abe2017tractable to the regression context. The joint density of \(\Theta\) and
-\(Y\), in this model defined only on the positive real half-line \([0, +
-\infty)\), reads
+This model is an extension of the cylindrical model introduced in Abe \&
+Ley (2017) to the regression context. The joint density of \(\Theta\)
+and \(Y\), in this model defined only on the positive real half-line
+\([0, + \infty)\), reads
+
 \begin{equation}\label{WeiSSVMdensity}
@@
 \end{equation}
-\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
-\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
-\(\kappa\) also regulating the circular-linear dependence. Our modification
-occurs at the level of the linear scale parameter \(\beta>0\) and circular
-location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
-covariates: \(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
-\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\). The
-parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
-\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
-\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\eta_0 \in [0,
-2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a vector of \(p\)
-regression coefficients \(\eta_j \in (-\infty, +\infty)\) for the prediction of
-\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
-vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
-is a vector of predictor values for the prediction of \(\theta\). In a
-regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
-generated independently from the modified
-density \eqref{WeiSSVMdensity}.\newline
-\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
-\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
-\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
-distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
-with location parameter \(\mu\) and concentration parameter \((\beta
-y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
+
+\noindent where \(\alpha > 0\) is a linear shape parameter,
+\(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular concentration
+and skewness parameters with \(\kappa\) also regulating the
+circular-linear dependence. Our modification occurs at the level of the
+linear scale parameter \(\beta>0\) and circular location parameter
+\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
+\(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
+\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\).
+The parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression
+coefficients \(\nu_j \in (-\infty, +\infty)\) for the prediction of
+\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
+parameter \(\eta_0 \in [0, 2\pi)\) is the intercept and
+\(\boldsymbol{\eta}\) is a vector of \(p\) regression coefficients
+\(\eta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\) where
+\(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector of
+predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\) is
+a vector of predictor values for the prediction of \(\theta\). In a
+regression setup the outcome vector \((\theta_i, y_i)^t\) for each
+individual is generated independently from the modified density
+\eqref{WeiSSVMdensity}.\newline
+\indent As in Abe \& Ley (2017), the conditional distribution of \(Y\)
+given \(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\)
+and scale \(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and
+the conditional distribution of \(\Theta\) given \(Y=y\) is a sine
+skewed von Mises distribution with location parameter \(\mu\) and
+concentration parameter \((\beta y)^\alpha\tanh(\kappa)\). The
+log-likelihood for this model equals
+
 \begin{align}\label{WeiSSVMLikelihood}
@@
 \end{align}
-\noindent For the teacher data, $\boldsymbol{z} = \boldsymbol{x}$ and the regression equations for the circular and
-linear outcomes in the Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
+
+\noindent For the teacher data, \(\boldsymbol{z} = \boldsymbol{x}\) and
+the regression equations for the circular and linear outcomes in the
+Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
 \tan^{-1}(\eta_1\text{SE}_i),\] and \[\hat{\beta}_{i} = \exp(\nu_0 +
-\nu_1\text{SE}_i).\]
-\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
-the maximum likelihood (ML) estimates for the parameters of the model.
+\nu_1\text{SE}_i).\] \noindent We can use numerical optimization
+(Nelder-Mead) to find solutions for the maximum likelihood (ML)
+estimates for the parameters of the model.
 
@@
 This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
-  outcomes and \(w\) independent linear outcomes. The circular outcomes
-  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
-  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
-  distribution.
-The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes.
-The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
-  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
-  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
+Mastrantonio (2018) to the regression context. Both models contain \(m\)
+independent circular outcomes and \(w\) independent linear outcomes. The
+circular outcomes
+\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\)
+are modelled together by a multivariate GPN distribution. The joint
+distribution of \(\boldsymbol{\Theta}\) and \(\boldsymbol{R}\) can thus
+be modeled as the product of (\ref{gpnreg}) for each of the \(m\)
+circular outcomes. The linear outcomes
+\(\boldsymbol{Y} = (\boldsymbol{Y}_1,  \dots, \boldsymbol{Y}_w)\) are
+modelled together by a multivariate skew normal distribution (Sahu, Dey,
+\& Branco, 2003). Because the GPN distribution is modelled using a
 so-called augmented representation (as in \eqref{projection} and
-  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
-  multivariate skew normal distribution. Following @mastrantonio2018joint the
-  linear outcomes are represented as
+\eqref{gpnreg}) it is convenient to use a similar tactic for modelling
+the multivariate skew normal distribution. Following Mastrantonio (2018)
+the linear outcomes are represented as
 \[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
-\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear outcome
-\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
-is a \(w \times w\) diagonal matrix with diagonal elements
+\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear
+outcome \(\boldsymbol{Y}\),
+\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
+\(w \times w\) diagonal matrix with diagonal elements
 \(\lambda_1, \dots, \lambda_w\) (skewness parameters),
 \(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
-$w$-dimensional half normal distribution [@olmos2012extension], and
-\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
-that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
-is normally distributed with mean \(\boldsymbol{\mu}_y +
-\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
-\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
-\boldsymbol{D}^t)^t\) is defined as:
+\(w\)-dimensional half normal distribution (Olmos, Varela, Gómez, \&
+Bolfarine, 2012), and
+\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\).
+This means that, conditional on the auxiliary data \(\boldsymbol{D}\),
+\(\boldsymbol{Y}\) is normally distributed with mean
+\(\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and
+covariance matrix \(\boldsymbol{\Sigma}_y\). The joint density for
+\((\boldsymbol{Y}^t, \boldsymbol{D}^t)^t\) is defined as:
+
 \begin{equation}\label{YDjoint}
@@
 \end{equation}
+
 \noindent where
-\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
-the \(\ell\)-dimensional normal density with mean vector
-\(\boldsymbol{\mu}_\ell\) and
-covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
-modelling the augmented representations of \(\boldsymbol{\Theta}\) and
-\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
-distribution. The joint density of the model is then represented by:
+\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\)
+stands for the \(\ell\)-dimensional normal density with mean vector
+\(\boldsymbol{\mu}_\ell\) and covariance \(\boldsymbol{\Sigma}_\ell\).
+As in Mastrantonio (2018) dependence between the linear and circular
+outcome is created by modelling the augmented representations of
+\(\boldsymbol{\Theta}\) and \(\boldsymbol{Y}\) together in a \(2m + w\)
+dimensional normal distribution. The joint density of the model is then
+represented by:
+
 \begin{equation}\label{YDThetarjoint} 
@@
 \end{equation}
-\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
-r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{\mu} =
-(\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\) and \(\boldsymbol{\Sigma} =
-\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
-\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
-The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
-of and covariances between the augmented representations of the circular outcome
-and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
-augmented representations of the circular outcome and the linear outcome.
-\newline
-\indent In our regression extension we have \(i = 1, \dots, n\) observations of
-\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
-the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{\mu}_i =
-\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
-\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
-is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
-and the $g$ covariate values. \newline
-\indent For the teacher data, the regression equations for
-the circular and linear outcomes in the GPN-SSN model are
+
+\noindent where
+\(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots, r_m(\cos(\theta_m), \sin(\theta_m)))^t\),
+the mean vector
+\(\boldsymbol{\mu} = (\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\)
+and
+\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
+The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
+variances of and covariances between the augmented representations of
+the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
+contains covariances between the augmented representations of the
+circular outcome and the linear outcome. \newline
+\indent In our regression extension we have \(i = 1, \dots, n\)
+observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
+covariates. The mean in the density in \eqref{YDThetarjoint} then
+becomes \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
+\(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with
+regression coefficients and intercepts and \(\boldsymbol{x}_i\) is a
+\(g + 1\) dimensional vector containing the value 1 to estimate an
+intercept and the \(g\) covariate values. \newline
+\indent For the teacher data, the regression equations for the circular
+and linear outcomes in the GPN-SSN model are
 \[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +
@@
 and
-\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\). Note that because $m = 1$ and $w = 1$, $\boldsymbol{\mu}_{s_i}$ is a 2 dimensional vector and $\mu_{y_i}$ is a scalar.
-We estimate the model using MCMC methods. A detailed description of these
-methods is given in the Supplementary Material.
+\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
+Note that because \(m = 1\) and \(w = 1\), \(\boldsymbol{\mu}_{s_i}\) is
+a 2 dimensional vector and \(\mu_{y_i}\) is a scalar. We estimate the
+model using MCMC methods. A detailed description of these methods is
+given in the Supplementary Material.
 
@@
 
-For the four cylindrical models we focus on their out-of-sample predictive
-performance to determine the fit of the model. To do so we use k-fold
-cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
-of the sample) is used once as a holdout set and 9 times as part of a training
-set. The analysis will thus be performed 10 times, each time on a different
-training set.\newline
-\indent A proper criterion to compare out-of-sample predictive performance is
-the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
-value of this criterion, the better the predictive performance of the model.
-Using ML estimates this criterion can be computed as follows:
+For the four cylindrical models we focus on their out-of-sample
+predictive performance to determine the fit of the model. To do so we
+use k-fold cross-validation and split our data into 10 folds. Each of
+these folds (10 \(\%\) of the sample) is used once as a holdout set and
+9 times as part of a training set. The analysis will thus be performed
+10 times, each time on a different training set.\newline
+\indent A proper criterion to compare out-of-sample predictive
+performance is the Predictive Log Scoring Loss (PLSL) (Gneiting \&
+Raftery, 2007). The lower the value of this criterion, the better the
+predictive performance of the model. Using ML estimates this criterion
+can be computed as follows:
+
 \begin{equation}\label{PLSLML}
@@
 \end{equation}
-\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
-holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
-\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
-Using posterior samples the criterion is similar to the log pointwise predictive
-density (lppd) [@BDA, p. 169] and can be computed as:
+
+\noindent where \(l\) is the model likelihood, \(M\) is the sample size
+of the holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout
+set and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the
+model parameters. Using posterior samples the criterion is similar to
+the log pointwise predictive density (lppd) (Gelman et al., 2014, p.
+169) and can be computed as:
+
 \begin{equation}\label{PLSLBayes}
@@
 \end{equation}
+
 \noindent where \(B\) is the amount of posterior samples and
-\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
-parameters for the \(j^{th}\) iteration. Because the joint density and thus also
-the likelihood for the modified GPN-SSN model in (\ref{YDThetarjoint}) is not
-available in closed form [@mastrantonio2018joint] we compute the PLSL for the
-circular and linear outcome separately for all models. Note that although we fit
-the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
-prior information into account when assessing model fit with the PLSL. According
-to @BDA this is not necessary since we are assessing the fit of a model
-to data, the holdout set, only. They argue that the prior in such case is only
-of interest for estimating the parameters of the model but not for determining
-the predictive accuracy.\newline
-\indent We use the loglikelihoods of the following conditional densities for the
-computation of the PLSL in the teacher data:
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
+model parameters for the \(j^{th}\) iteration. Because the joint density
+and thus also the likelihood for the modified GPN-SSN model in
+(\ref{YDThetarjoint}) is not available in closed form (Mastrantonio,
+2018) we compute the PLSL for the circular and linear outcome separately
+for all models. Note that although we fit the CL-PN, CL-GPN and GPN-SSN
+models using Bayesian statistics, we do not take prior information into
+account when assessing model fit with the PLSL. According to Gelman et
+al. (2014) this is not necessary since we are assessing the fit of a
+model to data, the holdout set, only. They argue that the prior in such
+case is only of interest for estimating the parameters of the model but
+not for determining the predictive accuracy.\newline
+\indent We use the loglikelihoods of the following conditional densities
+for the computation of the PLSL in the teacher data:
 
@@
 \noindent For each of the four cylindrical models and for each of the 10
-cross-validation analyses we can then compute a PLSL for the circular and linear
-outcome by using the conditional log-likelihoods of the respective outcome. To
-evaluate the predictive performance we average across the PLSL criteria of the
-cross-validation analyses. We also assess the cross-validation variability by
-means of the standard deviations of the PLSL criteria.
+cross-validation analyses we can then compute a PLSL for the circular
+and linear outcome by using the conditional log-likelihoods of the
+respective outcome. To evaluate the predictive performance we average
+across the PLSL criteria of the cross-validation analyses. We also
+assess the cross-validation variability by means of the standard
+deviations of the PLSL criteria.
 
@@
 In this section we analyze the teacher data with the help of the four
-cylindrical models from Section \ref{Models}. We will present the results,
-posterior estimates and their interpretation, per model and finish with a
-section comparing the fit of the different models.
-
-```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling CL-PN.R")
-
-its <- 20000
-set.seed(101)
-
-res_CLPN <- list()
-
-for(i in 1:10){
-
-  train <- get(paste("train.", i, sep = ""))
-  hold <- get(paste("hold.", i, sep = ""))
-
-  ZI.PN.tr  <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-  ZII.PN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-  X.PN.tr   <- as.matrix(cbind(rep(1, length(train$theta)),
-                                 cos(train$theta),
-                                 sin(train$theta),
-                                 train$SEc))
-
-  ZI.PN.h  <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-  ZII.PN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-  X.PN.h   <- as.matrix(cbind(rep(1, length(hold$theta)),
-                               cos(hold$theta),
-                               sin(hold$theta),
-                               hold$SEc))
-
-  res_CLPN[[i]] <- CLPN(train$theta, train$y, X.PN.tr, ZI.PN.tr, ZII.PN.tr, its,
-                        hold$theta, hold$y, X.PN.h, ZI.PN.h, ZII.PN.h)
-
-}
-
-```
-
-```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling CL-GPN.R")
-
-its <- 20000
-set.seed(101)
-
-res_CLGPN <- list()
-
-for(i in 1:10){
-
-  train <- get(paste("train.", i, sep = ""))
-  hold  <- get(paste("hold.", i, sep = ""))
-
-  Z.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-  X.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)),
-                              cos(train$theta),
-                              sin(train$theta),
-                              train$SEc))
-
-  Z.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-  X.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)),
-                             cos(hold$theta),
-                             sin(hold$theta),
-                             hold$SEc))
-
-  res_CLGPN[[i]] <- CLGPN(train$theta, train$y, X.GPN.tr, Z.GPN.tr, its, p = 2, 
-                          hold$theta, hold$y, X.GPN.h, Z.GPN.h)
-
-}
-
-```
-
-```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Posterior Sampling Joint GPN-SSN.R")
-
-its <- 20000
-set.seed(101)
-
-res_CLGPNM <- list()
-
-for(i in 1:10){
-
-  train <- get(paste("train.", i, sep = ""))
-  hold  <- get(paste("hold.", i, sep = ""))
-
-  X.MGPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-
-  X.MGPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-
-  res_CLGPNM[[i]] <- JGPNSSN(train$theta, train$y, X.MGPN.tr, its, p = 1, q = 1, 
-                             hold$theta, hold$y, X.MGPN.h)
-
-}
-
-```
-
-```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}
-
-source("R-code/Abe-Ley optimization.R")
-
-resAL <- list()
-
-set.seed(101)
-
-for(i in 1:10){
-
-  train <- get(paste("train.", i, sep = ""))
-  hold <- get(paste("hold.", i, sep = ""))
-
-  Z.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-  X.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
-
-  Z.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-  X.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
-
-  theta.tr <- as.numeric(train$theta)
-  y.tr     <- as.numeric(train$y)
-  theta.h  <- as.numeric(hold$theta)
-  y.h      <- as.numeric(hold$y)
-
-  colnames(X.tr) <- c("ax", "bx")
-  colnames(Z.tr) <- c("az", "bz")
-  colnames(X.h) <- c("ax", "bx")
-  colnames(Z.h) <- c("az", "bz")
-
-  dat.t <- as.data.frame(cbind(theta.tr, y.tr, X.tr, Z.tr))
-  dat.h <- as.data.frame(cbind(theta.h, y.h, X.h, Z.h))
-
-  ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
-                 c(-1, 0, 0, 0, 0, 0, 0),
-                 c(0, 0, 0, 0, 1, 0, 0),
-                 c(0, 0, 0, 0, 0, 1, 0),
-                 c(0, 0, 0, 0, 0, 0, 1),
-                 c(0, 0, 0, 0, 0, 0, -1))
-
-  ci.reg = c(-pi, -pi, 0, 0, -1, -1)
-
-  param <- c(1,1,1,1,1,1,0)*0.9
-
-  ui.reg %*% param - ci.reg
-
-  resAL[[i]] <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
-                            control = list(maxit = 1000000, fnscale = -1), data = dat.t)
-
-  resAL[[i]]$ll.circ <- -2*func.reg.cond.circ(resAL[[i]]$par, dat.h)
-  resAL[[i]]$ll.lin  <- -2*func.reg.cond.lin(resAL[[i]]$par, dat.h)
-
-}
-
-```
+cylindrical models from Section \ref{Models}. We will present the
+results, posterior estimates and their interpretation, per model and
+finish with a section comparing the fit of the different models.
 
@@
 
-In the Supplementary Material we have described the starting values for the MCMC
-procedures, hence it remains to specify  the starting  values for the
-maximum likelihood based Abe-Ley model: \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 =
-0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\). The initial amount
-of iterations for the three MCMC samplers was set to 2000. After convergence
-checks via traceplots we concluded that some of the parameters of the GPN-SSN
-model did not converge. Therefore we set the amount of iterations of the MCMC
-models to 20,000 and subtracted a burn-in of 5000 to reach convergence. Note
-that we choose the same amount of iterations for all three Bayesian models to
-make their comparison via the PLSL as fair as possible. Lastly, the predictor
-\verb|SE| was centered before inclusion in the analysis as this allows the
-intercepts to bear the classical meaning of average behavior.
+In the Supplementary Material we have described the starting values for
+the MCMC procedures, hence it remains to specify the starting values for
+the maximum likelihood based Abe-Ley model:
+\(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
+The initial amount of iterations for the three MCMC samplers was set to
+2000. After convergence checks via traceplots we concluded that some of
+the parameters of the GPN-SSN model did not converge. Therefore we set
+the amount of iterations of the MCMC models to 20,000 and subtracted a
+burn-in of 5000 to reach convergence. Note that we choose the same
+amount of iterations for all three Bayesian models to make their
+comparison via the PLSL as fair as possible. Lastly, the predictor
+\verb|SE| was centered before inclusion in the analysis as this allows
+the intercepts to bear the classical meaning of average behavior.
 
 \subsubsection{The modified CL-PN and CL-GPN models} \label{resCL(G)PN}
-
-```{r estCLPN, cache = TRUE, echo = FALSE}
-
-#Get 3 dimensional arrays (iteration, parameter, fold) for each parameter type
-Gamma <- simplify2array(lapply(res_CLPN, "[[", "Gamma"))
-BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
-BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
-Sigma <- simplify2array(lapply(res_CLPN, "[[", "Sigma"))
-
-#Averge over the folds and compute the posterior modes and hpd intervals
-CLPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
-CLPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est),1, sd)
-CLPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
-CLPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
-CLPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
-CLPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
-CLPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLPN.Sigma.m   <- mean(apply(Sigma[5000:20000,], 2, mode_est))
-CLPN.Sigma.hpd <- rowMeans(apply(Sigma[5000:20000,], 2, hpd_est))
-CLPN.Sigma.m.sd  <- sd(apply(Sigma[5000:20000,], 2, mode_est))
-CLPN.Sigma.hpd.sd <- apply(apply(Sigma[5000:20000,], 2, hpd_est), 1, sd)
-
-#Compute a cross-validation variance/standard deviation
-
-#Put estimates in correct format for tables
-modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
-hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
-hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)
-
-modes.PN.sd <- c(CLPN.BI.m.sd, CLPN.BII.m.sd, CLPN.Gamma.m.sd, CLPN.Sigma.m.sd, NA, NA, NA)
-hpd.LB.PN.sd <- c(CLPN.BI.hpd.sd[1,], CLPN.BII.hpd.sd[1,], CLPN.Gamma.hpd.sd[1,], CLPN.Sigma.hpd.sd[1], NA, NA, NA)
-hpd.UB.PN.sd <- c(CLPN.BI.hpd.sd[2,], CLPN.BII.hpd.sd[2,], CLPN.Gamma.hpd.sd[2,], CLPN.Sigma.hpd.sd[2], NA, NA, NA)
-
-CLPNres.tab <- cbind(modes.PN, modes.PN.sd, hpd.LB.PN, hpd.LB.PN.sd, hpd.UB.PN, hpd.UB.PN.sd)
-CLPNres.tab = CLPNres.tab
-
-CLPNres.tab.new = cbind(apply(CLPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
-                        apply(CLPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
-                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
-                        apply(CLPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
-                                                                       sprintf("%.2f", round(x[2], 2)), ")")))
-
-```
-
-```{r estCLGPN, cache = TRUE, echo = FALSE}
-
-#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
-Gamma <- simplify2array(lapply(res_CLGPN, "[[", "Gamma"))
-BI    <- simplify2array(lapply(res_CLGPN, "[[", "BI"))
-BII   <- simplify2array(lapply(res_CLGPN, "[[", "BII"))
-Sig   <- simplify2array(lapply(res_CLGPN, "[[", "Sig"))
-Sigma <- simplify2array(lapply(res_CLGPN, "[[", "Sigma"))
-
-#Averge over the folds and compute the posterior modes and hpd intervals
-CLGPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
-CLGPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est), 1, sd)
-CLGPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLGPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLGPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
-CLGPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
-CLGPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLGPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLGPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
-CLGPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
-CLGPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
-CLGPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)
-
-CLGPN.Sig.m   <- mean(apply(Sig[5000:20000,], 2, mode_est))
-CLGPN.Sig.m.sd   <- sd(apply(Sig[5000:20000,], 2, mode_est))
-CLGPN.Sig.hpd <- rowMeans(apply(Sig[5000:20000,], 2, hpd_est))
-CLGPN.Sig.hpd.sd <- apply(apply(Sig[5000:20000,], 2, hpd_est), 1, sd)
-
-CLGPN.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
-CLGPN.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
-CLGPN.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
-CLGPN.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
-CLGPN.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
-CLGPN.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
-
-CLGPN.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
-CLGPN.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
-CLGPN.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
-CLGPN.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
-CLGPN.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
-CLGPN.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
-
-#Compute a cross-validation variance/standard deviation
-
-#Put estimates in correct format for tables
-modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
-                CLGPN.Gamma.m, CLGPN.Sig.m,
-                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
-hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
-               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
-               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
-hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
-               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
-               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])
-
-modes.GPN.sd  <- c(CLGPN.BI.m.sd, CLGPN.BII.m.sd,
-                   CLGPN.Gamma.m.sd, CLGPN.Sig.m.sd,
-                   CLGPN.Sigma11.m.sd, CLGPN.Sigma12.m.sd, CLGPN.Sigma22.m.sd)
-hpd.LB.GPN.sd <- c(CLGPN.BI.hpd.sd[1,], CLGPN.BII.hpd.sd[1,],
-                   CLGPN.Gamma.hpd.sd[1,], CLGPN.Sig.hpd.sd[1],
-                   CLGPN.Sigma11.hpd.sd[1], CLGPN.Sigma12.hpd.sd[1], CLGPN.Sigma22.hpd.sd[1])
-hpd.UB.GPN.sd <- c(CLGPN.BI.hpd.sd[2,], CLGPN.BII.hpd.sd[2,], 
-                   CLGPN.Gamma.hpd.sd[2,], CLGPN.Sig.hpd.sd[2],
-                   CLGPN.Sigma11.hpd.sd[2], CLGPN.Sigma12.hpd.sd[2], CLGPN.Sigma22.hpd.sd[2])
-
-CLGPNres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd, hpd.UB.GPN, hpd.UB.GPN.sd)
-
-CLGPNres.tab.new = cbind(apply(CLGPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
-                         apply(CLGPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
-                         apply(CLGPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                         sprintf("%.2f", round(x[2], 2)), ")")))
-
-CLPNGPNres.tab <- cbind(CLPNres.tab.new, CLGPNres.tab.new)
-
-rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_1^{I}$",
-                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
-                              "$\\gamma_0$", "$\\gamma_{cos}$",
-                              "$\\gamma_{sin}$", "$\\gamma_1$",
-                              "$\\sigma$", "$\\sum_{1,1}$",
-                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
-colnames(CLPNGPNres.tab) <- rep(c('Mode', "HPD LB", "HPD UB"), 2)
-
-#Create table
-#kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE, 
-#      caption = "Results, cross-validation mean and standard deviation, for the modified CL-PN and CL-GPN model")%>%
-#add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))
-
-```
-
-```{r means, cache = TRUE, echo = FALSE}
-
-theta_pred_PN   <- simplify2array(lapply(res_CLPN, "[[", "theta_pred"))
-theta_pred_GPN  <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred"))
-theta_pred_GPNM <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred"))
-
-theta_pred_PN <- array(unlist(theta_pred_PN),
-                       dim = c(nrow(theta_pred_PN[[1]]), ncol(theta_pred_PN[[1]]), length(theta_pred_PN)))
-theta_pred_GPN <- array(unlist(theta_pred_GPN),
-                        dim = c(nrow(theta_pred_GPN[[1]]), ncol(theta_pred_GPN[[1]]), length(theta_pred_GPN)))
-theta_pred_GPNM <- array(unlist(theta_pred_GPNM),
-                         dim = c(nrow(theta_pred_GPNM[[1]]), ncol(theta_pred_GPNM[[1]]), length(theta_pred_GPNM)))
-
-means <- round(rbind(mean_circ(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
-                     mean_circ(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
-                     mean_circ(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)))*(180/pi), 2)
-hpds  <- round(rbind(apply(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
-                     apply(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
-                     apply(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ))*(180/pi), 2)
-
-means = cbind(means, hpds)
-
-rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
-colnames(means) <- c('Mode', "HPD LB", "HPD UB")
-#kable(means, "latex", booktabs = T, escape = FALSE, digits = 2,
-#      caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and GPN-SSN models")%>%
-#kable_styling() %>%
-#add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", #notation="alphabet")
-
-```
 
@@
   \beta_0^{II} + \beta_1^{II}\text{SE}_i
- \end{pmatrix}.\]
-\noindent  For both models, \(\hat{y}\) is the predicted strength of
-interpersonal behavior, \(\hat{\boldsymbol{\mu}}\) is the predicted mean vector
-of the type of interpersonal behavior and the \(\gamma\)'s and \(\beta\)'s are
-intercepts and regression coefficients.\newline \indent Table \ref{tab:estCLGPN}
-shows the results for the modified CL-PN and CL-GPN models fit to the teacher
-dataset. The estimates in this table are the averages and standard deviations of
-the 10 cross-validation estimates. We show both the estimated posterior mode and
-the 95\% highest posterior density (HPD) interval for each parameter. The
-posterior estimates for the \(\beta\)'s and \(\gamma\)'s are quite similar for
-both models. We also see this similarity when we look at the predicted circular
-and linear means. The predicted linear mean is equal to \(\gamma_0\) which is
-equal to 0.38 and 0.37 in the CL-PN and CL-GPN models, respectively. The
-predicted circular means can be computed from \(\beta_0^{I}\) and
-\(\beta_0^{II}\) using a double arctangent function \(\mbox{atan2}(\beta_0^{II},
-\beta_0^{I})\), see \eqref{PredVal}. If we do this at each iteration of the MCMC
-sampler we get the posterior distribution of the circular mean in both models.
-Table \ref{tab:means} shows the posterior mode of the estimated circular means,
-which are very similar (32.29\(^\circ\) and 33.54\(^\circ\)).\newline
-\indent Next we  investigate the effect of self-efficacy. In the CL-PN model we
-can transform the regression parameters on the two components \(I\) and \(II\)
-to one circular regression coefficient \(b_c\) using the methods described by
-@CremersMulderKlugkist2017. The dashed line in Figure \ref{regline} shows the
-predicted regression line from the CL-PN model. Here $b_c$ is the slope of this
-line at the inflection point (black square) and its posterior mode is estimated
-at 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation to the
+ \end{pmatrix}.\] \noindent  For both models, \(\hat{y}\) is the
+predicted strength of interpersonal behavior, \(\hat{\boldsymbol{\mu}}\)
+is the predicted mean vector of the type of interpersonal behavior and
+the \(\gamma\)'s and \(\beta\)'s are intercepts and regression
+coefficients.\newline \indent Table \ref{tab:estCLGPN} shows the results
+for the modified CL-PN and CL-GPN models fit to the teacher dataset. The
+estimates in this table are the averages and standard deviations of the
+10 cross-validation estimates. We show both the estimated posterior mode
+and the 95\% highest posterior density (HPD) interval for each
+parameter. The posterior estimates for the \(\beta\)'s and \(\gamma\)'s
+are quite similar for both models. We also see this similarity when we
+look at the predicted circular and linear means. The predicted linear
+mean is equal to \(\gamma_0\) which is equal to 0.38 and 0.37 in the
+CL-PN and CL-GPN models, respectively. The predicted circular means can
+be computed from \(\beta_0^{I}\) and \(\beta_0^{II}\) using a double
+arctangent function \(\mbox{atan2}(\beta_0^{II}, \beta_0^{I})\), see
+\eqref{PredVal}. If we do this at each iteration of the MCMC sampler we
+get the posterior distribution of the circular mean in both models.
+Table \ref{tab:means} shows the posterior mode of the estimated circular
+means, which are very similar (32.29\(^\circ\) and
+33.54\(^\circ\)).\newline
+\indent Next we investigate the effect of self-efficacy. In the CL-PN
+model we can transform the regression parameters on the two components
+\(I\) and \(II\) to one circular regression coefficient \(b_c\) using
+the methods described by Cremers, Mulder, \& Klugkist (2018b). The
+dashed line in Figure \ref{regline} shows the predicted regression line
+from the CL-PN model. Here \(b_c\) is the slope of this line at the
+inflection point (black square) and its posterior mode is estimated at
+1.67 (-24.66,
+29.33)\footnote{Note that this is a linear approximation to the
 circular regression line representing the slope at a specific point. Therefore
@@
 interval is much wider and covers 0, indicating there is no evidence to reject
-the null-hypothesis of no effect.}. Even though the HPD of \(b_c\) includes 0,
-indicating there is no evidence to reject the null-hypotheis of no effect, we
-continue with its interpretation for eductional purposes. The interpretation of
-$b_c$ is that at the inflection point an increase of 1 unit in self-efficacy
-leads to an increase of \(1.67*(180/\pi) = 95.68^\circ\) in the type of
-interpersonal behavior..However, as we can see in Figure \ref{regline} the
-inflection point lies almost outside the range of the actual data. Instead of
-looking at the inflection point we might compute the slope of the regression
-line at the average self-efficacy. This parameter we call the slope at the mean
-(\(SAM\)) [@CremersMulderKlugkist2017] and it is estimated at 0.02 (0.01; 0.04)
-for our data. This means that at the average self-efficacy an increase of 1 unit
-only leads to an increase of \(0.02*(180/\pi) = 1.15^\circ\) in the type of
-interpersonal behavior. The HPD of the \(SAM\) does not include 0 which means
-that the effect at the average self-efficacy can be distinguished from
-0.\newline
-\indent In the CL-GPN model we cannot compute circular regression coefficients
-such as the \(b_c\) and \(SAM\) computed above  due to the fact that not only
-the mean vector $\boldsymbol{\mu}$ but also the covariance matrix
-$\boldsymbol{\Sigma}$ influences the predicted value on the circle. Instead, we
-will compute posterior predictive distributions for the predicted circular
-outcome of individuals scoring the minimum, maximum and median self-efficacy.
-The modes and 95\% HPD intervals of these posterior predictive distributions are
+the null-hypothesis of no effect.}. Even though the HPD of \(b_c\)
+includes 0, indicating there is no evidence to reject the null-hypotheis
+of no effect, we continue with its interpretation for eductional
+purposes. The interpretation of \(b_c\) is that at the inflection point
+an increase of 1 unit in self-efficacy leads to an increase of
+\(1.67*(180/\pi) = 95.68^\circ\) in the type of interpersonal
+behavior..However, as we can see in Figure \ref{regline} the inflection
+point lies almost outside the range of the actual data. Instead of
+looking at the inflection point we might compute the slope of the
+regression line at the average self-efficacy. This parameter we call the
+slope at the mean (\(SAM\)) (Cremers et al., 2018b) and it is estimated
+at 0.02 (0.01; 0.04) for our data. This means that at the average
+self-efficacy an increase of 1 unit only leads to an increase of
+\(0.02*(180/\pi) = 1.15^\circ\) in the type of interpersonal behavior.
+The HPD of the \(SAM\) does not include 0 which means that the effect at
+the average self-efficacy can be distinguished from 0.\newline
+\indent In the CL-GPN model we cannot compute circular regression
+coefficients such as the \(b_c\) and \(SAM\) computed above due to the
+fact that not only the mean vector \(\boldsymbol{\mu}\) but also the
+covariance matrix \(\boldsymbol{\Sigma}\) influences the predicted value
+on the circle. Instead, we will compute posterior predictive
+distributions for the predicted circular outcome of individuals scoring
+the minimum, maximum and median self-efficacy. The modes and 95\% HPD
+intervals of these posterior predictive distributions are
 \(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\),
 \(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\). Note
-that we display the modes and HPD intervals for the posterior predictive
-distributions on the interval $[0^\circ, 360^\circ)$. Note that  $44.49^\circ =
-404.49^\circ$ due to the periodicity of a circular variable. The posterior mode estimate of
-$215.74^\circ$ thus lies within its HPD interval $(147.36^\circ, \: 44.49^\circ)$.The
-HPD intervals of the three posterior predictive distributions overlap. Had they
-not overlapped we could have concluded that as the self-efficacy increases, the
-score of the teacher on the IPC moves counterclockwise. \newline
-\indent The effect of self-efficacy on the strength of interpersonal behavior is
-quantified by \(\gamma_1\) in both the CL-PN and CL-GPN models. This parameter
-can be interpreted in the same way as in a usual regression
-model. The HPD interval of  \(\gamma_1\) includes 0 which means
-that there is not enough evidence for an effect of self-efficacy on the strength
-of interpersonal behavior. Note that the lower bounds of the HPD interval of
-this parameter are  however very close to zero.\newline
-\indent  The relation between the circular and linear outcome, that is, between
-type of interpersonal behavior and strength of interpersonal behavior, is
-described by the parameters $\gamma_{\cos}$ and $\gamma_{\sin}$. The HPD
-interval of \(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
-models, meaning that the cosine component of the type of interpersonal behavior
-has an effect on the strength of interpersonal behavior. In the teacher data the
-sine and cosine components have a substantive meaning. In this case the
-Communion (cosine) component of the IPC positively effects the strength of a
-teachers' type of interpersonal behavior, in plain words: teachers exhibiting
-interpersonal behavior types with higher communion scores (e.g., 'helpful' and
-'understanding' in Figure 2) are stronger in their behavior.
-
-```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-a1 <- CLPN.BI.m[1]
-b1 <- CLPN.BI.m[2]
-a2 <- CLPN.BII.m[1]
-b2 <- CLPN.BII.m[2]
-ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-
-x <- seq(-5, 5, by = 0.1)
-
-pred.AL <- 0.37 + 2*atan(-0.01*x)
-pred.PN <- atan2(a2 + b2*x, a1 + b1*x)
-
-plot(Dat$SEc, Dat$theta*(180/pi), bty = "n", yaxt = "n",
-     xlim = c(-5, 2), ylab = "$\\theta$", xlab = "self-efficacy")
-axis(side = 2,  at = c(-160, -80, 0, 80, 160))
-points(x, pred.AL*(180/pi), type = "l", lty = 1)
-points(x, pred.PN*(180/pi), type = "l", lty = 2)
-points(ax, ac*(180/pi), pch = 15)
-
-dev.off()
-tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)
-
-```
-
-```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}
-
-BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
-BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
-
-bc <- matrix(NA, 15001, 10)
-SAM <- matrix(NA, 15001, 10)
-
-for(i in 1:10){
-
-  a1 <- BI[5000:20000,1,i]
-  b1 <- BI[5000:20000,2,i]
-  a2 <- BII[5000:20000,1,i]
-  b2 <- BII[5000:20000,2,i]
-
-  ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-  ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-  bc[,i] <- -tan(atan2(a2, a1)-ac)/ax
-  SAM[,i] <- bc[i]/(1 + (bc[i]*(-ax))^2)
-
-}
-
-mean(apply(bc, 2, mode_est))
-rowMeans(apply(bc, 2, hpd_est))
-
-mean(apply(SAM, 2, mode_est))
-rowMeans(apply(SAM, 2, hpd_est))
-
-```
+\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\).
+Note that we display the modes and HPD intervals for the posterior
+predictive distributions on the interval \([0^\circ, 360^\circ)\). Note
+that \(44.49^\circ = 404.49^\circ\) due to the periodicity of a circular
+variable. The posterior mode estimate of \(215.74^\circ\) thus lies
+within its HPD interval \((147.36^\circ, \: 44.49^\circ)\).The HPD
+intervals of the three posterior predictive distributions overlap. Had
+they not overlapped we could have concluded that as the self-efficacy
+increases, the score of the teacher on the IPC moves counterclockwise.
+\newline
+\indent The effect of self-efficacy on the strength of interpersonal
+behavior is quantified by \(\gamma_1\) in both the CL-PN and CL-GPN
+models. This parameter can be interpreted in the same way as in a usual
+regression model. The HPD interval of \(\gamma_1\) includes 0 which
+means that there is not enough evidence for an effect of self-efficacy
+on the strength of interpersonal behavior. Note that the lower bounds of
+the HPD interval of this parameter are however very close to
+zero.\newline
+\indent  The relation between the circular and linear outcome, that is,
+between type of interpersonal behavior and strength of interpersonal
+behavior, is described by the parameters \(\gamma_{\cos}\) and
+\(\gamma_{\sin}\). The HPD interval of \(\gamma_{\cos}\) does not
+include 0 for both the CL-PN and CL-GPN models, meaning that the cosine
+component of the type of interpersonal behavior has an effect on the
+strength of interpersonal behavior. In the teacher data the sine and
+cosine components have a substantive meaning. In this case the Communion
+(cosine) component of the IPC positively effects the strength of a
+teachers' type of interpersonal behavior, in plain words: teachers
+exhibiting interpersonal behavior types with higher communion scores
+(e.g., `helpful' and `understanding' in Figure 2) are stronger in their
+behavior.
 
@@
 
-```{r ppreg, echo = FALSE, results = FALSE, eval = FALSE}
-theta_pred_GPN.min <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.min"))
-theta_pred_GPN.median <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.meadian"))
-theta_pred_GPN.max <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.max"))
-
-mean_circ(apply(theta_pred_GPN.min[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPN.median[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPN.max[5000:20000,], 2, mode_est_circ))*(180/pi)
-
-apply(apply(theta_pred_GPN.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPN.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPN.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-
-```
-
 \subsubsection{The modified Abe-Ley model} \label{resAL}
-
-```{r estAL, cache = TRUE, echo = FALSE}
-#Get 2 dimensional arrays (parameter, fold)
-resAL_arr <- simplify2array(lapply(resAL, "[[", "par"))
-
-#Compute a cross-validation variance/standard deviation
-
-#Put estimates in correct format for table
-ALres.tab <- matrix(round(c(rowMeans(resAL_arr[1:4,]), mean(resAL_arr[6,]), 
-                            mean(resAL_arr[5,]), mean(resAL_arr[7,]), 
-                            apply(resAL_arr[1:4,], 1, sd), sd(resAL_arr[6,]), 
-                            sd(resAL_arr[5,]), sd(resAL_arr[7,])), 2), 7, 2)
-
-ALres.tab.new = cbind(apply(ALres.tab, 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
-                                                             sprintf("%.2f", round(x[2], 2)), ")")))
-
-rownames(ALres.tab.new) <- c("$\\eta_0$", "$\\eta_1$",
-                             "$\\nu_0$", "$\\nu_1$",
-                             "$\\alpha$", "$\\kappa$", "$\\lambda$")
-
-#Create table
-#kable(ALres.tab.new, "latex", booktabs = T, escape = FALSE, digits = 2,
-#      caption = "Results, cross-validation mean and standard deviation, for the modified Abe-Ley model")%>%
-#add_header_above(c("Parameter" = 1, "ML-estimate" = 1))
-
-```
 
@@
 
-\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type
-of interpersonal behavior and \(\hat{\beta}\) is the predicted scale
-parameter of the strength of interpersonal behavior, while the \(\eta\)'s and
-\(\nu\)'s are regression parameters.\newline
-\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley model
-fit to the teacher dataset. The estimates in this table are the averages and
-standard deviations of the 10 cross-validation estimates. The estimate of the
-circular mean at an average self-efficacy, \(\eta_0\), equals 0.36 radians or
-20.62\(^\circ\). For the Abe-Ley model we can also investigate the effect of
-self-efficacy on the type of interpersonal behavior. The solid line in
-Figure \ref{regline} shows this effect. Unlike for the regression line of the
-CL-PN model, the inflection point of the regression line of the Abe-Ley model
-lies outside the x-range of the figure. \newline
-\indent For the Abe-Ley model the conditional distribution for the linear
-outcome is Weibull. This means that we can use methods from survival analsis to
-interpret the effect of self-efficacy. In survival analysis a 'survival'
-function is used in which time is plotted against the probability of survival of
-subjects suffering from a specific medical condition. In our data however we
-plot the strength on the IPC against the probability of a teacher having such a
-strength. This probability is computed using the
-'survival-function' \(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i -
-\mu_i))^{1/\alpha}})\) with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
-\ref{reglineweib}, we plot the survival function for the minimum, median and
-maximum value of self-efficacy. We conclude that stronger interpersonal
-behaviors are less probable. We also see that the relation between self-efficacy and the strength on the IPC is not linear; the probability of having a stronger
-interpersonal behavior is higher for both the minimum and maximum self-efficacy compared to a median self-efficacy score.
-Note however that the circular outcome also influences the survival function.
-The relation between the type of interpersonal behavior and the strength of interpersonal behavior may thus influence the shape of the survival function.\newline
-
-```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-summary(Dat$SEc)
-
-datafig =  rbind(Dat[Dat$SEc == min(Dat$SEc),], Dat[Dat$SEc == median(Dat$SEc),], Dat[Dat$SEc == max(Dat$SEc),])
-
-x = unique(datafig$SEc)
-theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
-y <- seq(0, 1, by = 0.05)
-#h <- matrix(NA, length(y), length(x))
-s <- matrix(NA, length(y), length(x))
-scross <- array(NA, dim = c(length(y), length(x), 10))
-
-for(k in 1:10){
-
-  train <- get(paste("train.", k, sep = ""))
-
-  datafig =  rbind(train[train$SEc == min(train$SEc),],
-                   train[train$SEc == median(train$SEc),],
-                   train[train$SEc == max(train$SEc),])
-  x = unique(datafig$SEc)
-  theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
-
-    for(j in 1:length(x)){
-
-      for(i in 1:length(y)){
-          mu <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*x[j])
-          beta <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*x[j])*(1 - tanh(mean(resAL_arr[5,k]))*cos(theta[j] - mu))^(1/mean(resAL_arr[6,k]))
-    #h[i,j] <- beta*mean(resAL_arr[6,])*y[i]^(mean(resAL_arr[6,])-1)
-          s[i,j] <- exp(-mean(resAL_arr[6,k])*y[i]^beta)
-
-    }
-
-     scross[,,k] <- s
-
-  }
-
-}
-
-s = apply(scross, c(1,2), mean)
-
-tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
-points(y, s[,2], type = "l", lty = 2)
-points(y, s[,3], type = "l", lty = 3)
-legend(0.7, 1, legend = c("min(SE)", "median(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")
-
-dev.off()
-tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)
-
-```
+\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type of
+interpersonal behavior and \(\hat{\beta}\) is the predicted scale
+parameter of the strength of interpersonal behavior, while the
+\(\eta\)'s and \(\nu\)'s are regression parameters.\newline
+\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley
+model fit to the teacher dataset. The estimates in this table are the
+averages and standard deviations of the 10 cross-validation estimates.
+The estimate of the circular mean at an average self-efficacy,
+\(\eta_0\), equals 0.36 radians or 20.62\(^\circ\). For the Abe-Ley
+model we can also investigate the effect of self-efficacy on the type of
+interpersonal behavior. The solid line in Figure \ref{regline} shows
+this effect. Unlike for the regression line of the CL-PN model, the
+inflection point of the regression line of the Abe-Ley model lies
+outside the x-range of the figure. \newline
+\indent For the Abe-Ley model the conditional distribution for the
+linear outcome is Weibull. This means that we can use methods from
+survival analsis to interpret the effect of self-efficacy. In survival
+analysis a `survival' function is used in which time is plotted against
+the probability of survival of subjects suffering from a specific
+medical condition. In our data however we plot the strength on the IPC
+against the probability of a teacher having such a strength. This
+probability is computed using the `survival-function'
+\(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}})\)
+with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
+\ref{reglineweib}, we plot the survival function for the minimum, median
+and maximum value of self-efficacy. We conclude that stronger
+interpersonal behaviors are less probable. We also see that the relation
+between self-efficacy and the strength on the IPC is not linear; the
+probability of having a stronger interpersonal behavior is higher for
+both the minimum and maximum self-efficacy compared to a median
+self-efficacy score. Note however that the circular outcome also
+influences the survival function. The relation between the type of
+interpersonal behavior and the strength of interpersonal behavior may
+thus influence the shape of the survival function.\newline
 
@@
 
-```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}
-
-require(pracma)
-
-alpha <- mean(resAL_arr[6,])
-kappa <- mean(resAL_arr[5,])
-lambda <- mean(resAL_arr[7,])
-
-#R cannot compute legendre functions of non-integer degree
-#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
-#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
-#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]
-
-#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
-
-1/3.82
-cosh(1.58)
-
-```
-
 \subsubsection{The modified GPN-SSN model} \label{resGPNSSN}
-
-```{r estCLGPNM, cache = TRUE, echo = FALSE}
-#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
-lambda   <- simplify2array(lapply(res_CLGPNM, "[[", "lambda"))
-Beta <- simplify2array(lapply(res_CLGPNM, "[[", "Beta"))
-Betacon <- simplify2array(lapply(res_CLGPNM, "[[", "Betacon"))
-Sigma <- simplify2array(lapply(res_CLGPNM, "[[", "Sigma"))
-Sigmacon <- simplify2array(lapply(res_CLGPNM, "[[", "Sigmacon"))
-
-#Averge over the folds and compute the posterior modes and hpd intervals
-CLGPNM.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
-CLGPNM.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
-CLGPNM.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
-CLGPNM.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
-CLGPNM.Sigma13.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
-CLGPNM.Sigma13.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
-CLGPNM.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
-CLGPNM.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
-CLGPNM.Sigma23.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
-CLGPNM.Sigma23.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
-CLGPNM.Sigma33.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
-CLGPNM.Sigma33.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]
-
-CLGPNM.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
-CLGPNM.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
-CLGPNM.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
-CLGPNM.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
-CLGPNM.Sigma13.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
-CLGPNM.Sigma13.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
-CLGPNM.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
-CLGPNM.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
-CLGPNM.Sigma23.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
-CLGPNM.Sigma23.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
-CLGPNM.Sigma33.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
-CLGPNM.Sigma33.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]
-
-CLGPNM.Sigmacon11.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
-CLGPNM.Sigmacon11.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
-CLGPNM.Sigmacon12.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
-CLGPNM.Sigmacon12.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
-CLGPNM.Sigmacon13.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
-CLGPNM.Sigmacon13.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
-CLGPNM.Sigmacon22.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
-CLGPNM.Sigmacon22.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
-CLGPNM.Sigmacon23.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
-CLGPNM.Sigmacon23.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
-CLGPNM.Sigmacon33.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
-CLGPNM.Sigmacon33.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]
-
-CLGPNM.Sigmacon11.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
-CLGPNM.Sigmacon11.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
-CLGPNM.Sigmacon12.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
-CLGPNM.Sigmacon12.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
-CLGPNM.Sigmacon13.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
-CLGPNM.Sigmacon13.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
-CLGPNM.Sigmacon22.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
-CLGPNM.Sigmacon22.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
-CLGPNM.Sigmacon23.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
-CLGPNM.Sigmacon23.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
-CLGPNM.Sigmacon33.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
-CLGPNM.Sigmacon33.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]
-
-CLGPNM.Beta11.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
-CLGPNM.Beta11.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
-CLGPNM.Beta21.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
-CLGPNM.Beta21.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
-CLGPNM.Beta12.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
-CLGPNM.Beta12.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
-CLGPNM.Beta22.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
-CLGPNM.Beta22.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
-CLGPNM.Beta13.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
-CLGPNM.Beta13.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
-CLGPNM.Beta23.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
-CLGPNM.Beta23.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]
-
-CLGPNM.Beta11.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
-CLGPNM.Beta11.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
-CLGPNM.Beta21.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
-CLGPNM.Beta21.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
-CLGPNM.Beta12.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
-CLGPNM.Beta12.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
-CLGPNM.Beta22.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
-CLGPNM.Beta22.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
-CLGPNM.Beta13.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
-CLGPNM.Beta13.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
-CLGPNM.Beta23.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
-CLGPNM.Beta23.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]
-
-CLGPNM.Betacon11.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
-CLGPNM.Betacon11.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
-CLGPNM.Betacon21.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
-CLGPNM.Betacon21.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
-CLGPNM.Betacon12.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
-CLGPNM.Betacon12.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
-CLGPNM.Betacon22.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
-CLGPNM.Betacon22.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
-CLGPNM.Betacon13.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
-CLGPNM.Betacon13.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
-CLGPNM.Betacon23.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
-CLGPNM.Betacon23.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]
-
-CLGPNM.Betacon11.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
-CLGPNM.Betacon11.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
-CLGPNM.Betacon21.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
-CLGPNM.Betacon21.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
-CLGPNM.Betacon12.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
-CLGPNM.Betacon12.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
-CLGPNM.Betacon22.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
-CLGPNM.Betacon22.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
-CLGPNM.Betacon13.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
-CLGPNM.Betacon13.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
-CLGPNM.Betacon23.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
-CLGPNM.Betacon23.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]
-
-CLGPNM.lambda.m   <- mean(apply(lambda[5000:20000,,], 2, mode_est))
-CLGPNM.lambda.m.sd   <- sd(apply(lambda[5000:20000,,], 2, mode_est))
-CLGPNM.lambda.hpd <- rowMeans(apply(lambda[5000:20000,,], 2, hpd_est))
-CLGPNM.lambda.hpd.sd <- apply(apply(lambda[5000:20000,,], 2, hpd_est), 1, sd)
-
-#Compute a cross-validation variance/standard deviation
-
-#Put estimates in correct format for tables
-modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
-                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
-                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
-                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
-                CLGPNM.lambda.m)
-hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
-                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
-                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
-                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
-                CLGPNM.lambda.hpd[1])
-hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
-                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
-                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
-                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
-                CLGPNM.lambda.hpd[2])
-
-modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
-                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
-                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
-                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
-                   CLGPNM.lambda.m)
-hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
-                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
-                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
-                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
-                   CLGPNM.lambda.hpd[1])
-hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
-                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
-                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
-                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
-                   CLGPNM.lambda.hpd[2])
-
-modes.GPN.sd  <- c(CLGPNM.Beta11.m.sd, CLGPNM.Beta12.m.sd, CLGPNM.Beta13.m.sd,
-                   CLGPNM.Beta21.m.sd, CLGPNM.Beta22.m.sd, CLGPNM.Beta23.m.sd,
-                   CLGPNM.Sigma11.m.sd, CLGPNM.Sigma22.m.sd, CLGPNM.Sigma33.m.sd,
-                   CLGPNM.Sigma12.m.sd, CLGPNM.Sigma13.m.sd, CLGPNM.Sigma23.m.sd,
-                  CLGPNM.lambda.m.sd)
-hpd.LB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[1], CLGPNM.Beta12.hpd.sd[1], CLGPNM.Beta13.hpd[1],
-                   CLGPNM.Beta21.hpd.sd[1], CLGPNM.Beta22.hpd.sd[1], CLGPNM.Beta23.hpd[1],
-                   CLGPNM.Sigma11.hpd.sd[1], CLGPNM.Sigma22.hpd.sd[1], CLGPNM.Sigma33.hpd[1],
-                   CLGPNM.Sigma12.hpd.sd[1], CLGPNM.Sigma13.hpd.sd[1], CLGPNM.Sigma23.hpd[1],
-                   CLGPNM.lambda.hpd.sd[1])
-hpd.UB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[2], CLGPNM.Beta12.hpd.sd[2], CLGPNM.Beta13.hpd.sd[2],
-                   CLGPNM.Beta21.hpd.sd[2], CLGPNM.Beta22.hpd.sd[2], CLGPNM.Beta23.hpd.sd[2],
-                   CLGPNM.Sigma11.hpd.sd[2], CLGPNM.Sigma22.hpd.sd[2], CLGPNM.Sigma33.hpd.sd[2],
-                   CLGPNM.Sigma12.hpd.sd[2], CLGPNM.Sigma13.hpd.sd[2], CLGPNM.Sigma23.hpd.sd[2],
-                   CLGPNM.lambda.hpd.sd[2])
-
-modescon.GPN.sd  <- c(CLGPNM.Betacon11.m.sd, CLGPNM.Betacon12.m.sd, CLGPNM.Betacon13.m.sd,
-                      CLGPNM.Betacon21.m.sd, CLGPNM.Betacon22.m.sd, CLGPNM.Betacon23.m.sd,
-                      CLGPNM.Sigmacon11.m.sd, CLGPNM.Sigmacon22.m.sd, CLGPNM.Sigmacon33.m.sd,
-                      CLGPNM.Sigmacon12.m.sd, CLGPNM.Sigmacon13.m.sd, CLGPNM.Sigmacon23.m.sd,
-                      CLGPNM.lambda.m.sd)
-hpdcon.LB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[1], CLGPNM.Betacon12.hpd.sd[1], CLGPNM.Betacon13.hpd.sd[1],
-                      CLGPNM.Betacon21.hpd.sd[1], CLGPNM.Betacon22.hpd.sd[1], CLGPNM.Betacon23.hpd.sd[1],
-                      CLGPNM.Sigmacon11.hpd.sd[1], CLGPNM.Sigmacon22.hpd.sd[1], CLGPNM.Sigmacon33.hpd.sd[1],
-                      CLGPNM.Sigmacon12.hpd.sd[1], CLGPNM.Sigmacon13.hpd.sd[1], CLGPNM.Sigmacon23.hpd.sd[1],
-                      CLGPNM.lambda.hpd.sd[1])
-hpdcon.UB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[2], CLGPNM.Betacon12.hpd.sd[2], CLGPNM.Betacon13.hpd.sd[2],
-                      CLGPNM.Betacon21.hpd.sd[2], CLGPNM.Betacon22.hpd.sd[2], CLGPNM.Betacon23.hpd.sd[2],
-                      CLGPNM.Sigmacon11.hpd.sd[2], CLGPNM.Sigmacon22.hpd.sd[2], CLGPNM.Sigmacon33.hpd.sd[2],
-                      CLGPNM.Sigmacon12.hpd.sd[2], CLGPNM.Sigmacon13.hpd.sd[2], CLGPNM.Sigmacon23.hpd.sd[2],
-                      CLGPNM.lambda.hpd.sd[2])
-
-#Create table
-CLGPNMres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd,
-                       hpd.UB.GPN, hpd.UB.GPN.sd, modescon.GPN, modescon.GPN.sd, 
-                       hpdcon.LB.GPN, hpdcon.LB.GPN.sd, hpdcon.UB.GPN, hpdcon.UB.GPN.sd)
-
-CLGPNMres.tab.new = cbind(apply(CLGPNMres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
-                          apply(CLGPNMres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
-                          apply(CLGPNMres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
-                          apply(CLGPNMres.tab[,7:8], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
-                          apply(CLGPNMres.tab[,9:10], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                            sprintf("%.2f", round(x[2], 2)), ")")),
-                          apply(CLGPNMres.tab[,11:12], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
-                                                                             sprintf("%.2f", round(x[2], 2)), ")")))
-
-rownames(CLGPNMres.tab.new) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
-                                 "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
-                                 "$\\beta_{1_s^{II}}$", "$\\beta_{1_y}$",
-                                 "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
-                                 "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
-                                 "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
-                                 "$\\lambda$")
-colnames(CLGPNMres.tab.new) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")
-
-#kable(CLGPNMres.tab.new, "latex", booktabs = TRUE, escape = FALSE, digits = 2,
-#      caption = "Results, cross-validation mean and standard deviation, for the GPN-SSN model")%>%
-#add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))
-
-```
-
-```{r ppreggpnm, echo = FALSE, results = FALSE, eval = FALSE}
-theta_pred_GPNM.min <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.min"))
-theta_pred_GPNM.median <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.median"))
-theta_pred_GPNM.max <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.max"))
-
-mean_circ(apply(theta_pred_GPNM.min[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPNM.median[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPNM.max[5000:20000,], 2, mode_est_circ))*(180/pi)
-
-dim(theta_pred_GPNM.median)
-
-apply(apply(theta_pred_GPNM.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPNM.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPNM.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-
-```
 
@@
 linear outcome.\newline
-\indent Table \ref{tab:estCLGPNM} shows the results for the modified GPN-SSN model fit to
-the teacher dataset. The estimates in this table are the averages and
-standard deviations of the 10 cross-validation estimates. We show both
-the estimated posterior mode and the 95\% highest posterior density
-(HPD) interval for each parameter. The predicted circular means can be
-computed from \(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) in a
-similar fashion as for the CL-PN and CL-GPN models. Table \ref{tab:means} shows the
-posterior mode of the estimated circular mean, which equals 35.30\(^\circ\) and hence
-is very similar to those of the CL-PN and CL-GPN models.\newline
-\indent For the same reason as in the CL-GPN model we cannot compute circular
-regression coefficients for the effect of self-efficacy on the type of
-interpersonal behavior such as the \(b_c\) and \(SAM\). Instead, we will again
-compute posterior predictive distributions for the predicted circular outcome of
-individuals scoring the minimum, maximum and median self-efficacy. The modes
-and 95\% HPD intervals of these posterior predictive distributions are
+\indent Table \ref{tab:estCLGPNM} shows the results for the modified
+GPN-SSN model fit to the teacher dataset. The estimates in this table
+are the averages and standard deviations of the 10 cross-validation
+estimates. We show both the estimated posterior mode and the 95\%
+highest posterior density (HPD) interval for each parameter. The
+predicted circular means can be computed from \(\beta_{0_{s^{I}}}\) and
+\(\beta_{0_{s^{II}}}\) in a similar fashion as for the CL-PN and CL-GPN
+models. Table \ref{tab:means} shows the posterior mode of the estimated
+circular mean, which equals 35.30\(^\circ\) and hence is very similar to
+those of the CL-PN and CL-GPN models.\newline
+\indent For the same reason as in the CL-GPN model we cannot compute
+circular regression coefficients for the effect of self-efficacy on the
+type of interpersonal behavior such as the \(b_c\) and \(SAM\). Instead,
+we will again compute posterior predictive distributions for the
+predicted circular outcome of individuals scoring the minimum, maximum
+and median self-efficacy. The modes and 95\% HPD intervals of these
+posterior predictive distributions are
 \(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.11^\circ, \: 72.02^\circ)\),
 \(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\). On a
-circle the HPD intervals of the three posterior predictive distributions
-overlap. Had they not overlapped we could have concluded that as the
-self-efficacy increases, the score of the teacher on the IPC moves
-counterclockwise. The effect of self-efficacy on the strength of interpersonal
-behavior is quantified by \(\beta_{1_y}\), and we learn that for a 1 unit
-increase in self-efficacy the strength of interpersonal behavior increases by
-0.09. The average strength is quantified by \(\beta_{0_y}\) and equals
-0.33.\newline
-\indent To investigate the association between the linear and circular outcome
-we look at the covariances between the linear outcome and the sine and cosine of
-the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). Both
-covariances, \(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are
-different from zero, but the one of the cosine component is larger. This means
-the correlation with the Communion component is larger and that teachers scoring
-both high on Communion and Agency show stronger behavior. This is a slightly
+\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\).
+On a circle the HPD intervals of the three posterior predictive
+distributions overlap. Had they not overlapped we could have concluded
+that as the self-efficacy increases, the score of the teacher on the IPC
+moves counterclockwise. The effect of self-efficacy on the strength of
+interpersonal behavior is quantified by \(\beta_{1_y}\), and we learn
+that for a 1 unit increase in self-efficacy the strength of
+interpersonal behavior increases by 0.09. The average strength is
+quantified by \(\beta_{0_y}\) and equals 0.33.\newline
+\indent To investigate the association between the linear and circular
+outcome we look at the covariances between the linear outcome and the
+sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
+\(\sum_{sy_{1,3}}\). Both covariances, \(\sum_{sy_{2,3}} = 0.09\) and
+\(\sum_{sy_{1,3}} = 0.23\), are different from zero, but the one of the
+cosine component is larger. This means the correlation with the
+Communion component is larger and that teachers scoring both high on
+Communion and Agency show stronger behavior. This is a slightly
 different conclusion from the one in the CL-PN and CL-GPN models.
@@
 
-In this section we will assess the overall fit of the cylindrical models to the
-data via the PLSL criterion described in Section \ref{Modelfit}. Table
-\ref{tab:ModelFit} shows the values of this criterion for the linear and
-circular outcomes of the four models.\newline
-\indent The CL-PN and CL-GPN models have the best out-of-sample predictive
-performance for the linear outcome. They show roughly the same performance because
-they model the linear outcome in the same way. Only the value of \(r\) in
-\eqref{circlinlink} differs. We should note that even though the predictive
-performance of the Abe-Ley model for the linear outcome is worst on average, the
-standard deviation of the cross-validation estimates is rather large. This means
-that in some samples, the Abe-Ley model shows a lower PLSL value than the
-average of 25.49\newline
-\indent The Abe-Ley model has the best out-of-sample predictive performance for
-the circular outcome. This would suggest that for the circular variable a
-slightly skewed distribution fits best. However, both the GPN-SSN  and the
-CL-GPN models fit much worse even though the distribution for the circular
-outcome in these models can also take a skewed shape. It should be noted that
-the  standard deviation of the cross-validation estimates is rather large for
-the Abe-Ley and the CL-GPN model. It is possible that these large standard
-deviations for the PLSL are caused by the fact that they are computed for a
-relatively small sample size, but this does not explain why the PLSL has a large
+In this section we will assess the overall fit of the cylindrical models
+to the data via the PLSL criterion described in Section \ref{Modelfit}.
+Table \ref{tab:ModelFit} shows the values of this criterion for the
+linear and circular outcomes of the four models.\newline
+\indent The CL-PN and CL-GPN models have the best out-of-sample
+predictive performance for the linear outcome. They show roughly the
+same performance because they model the linear outcome in the same way.
+Only the value of \(r\) in \eqref{circlinlink} differs. We should note
+that even though the predictive performance of the Abe-Ley model for the
+linear outcome is worst on average, the standard deviation of the
+cross-validation estimates is rather large. This means that in some
+samples, the Abe-Ley model shows a lower PLSL value than the average of
+25.49\newline
+\indent The Abe-Ley model has the best out-of-sample predictive
+performance for the circular outcome. This would suggest that for the
+circular variable a slightly skewed distribution fits best. However,
+both the GPN-SSN and the CL-GPN models fit much worse even though the
+distribution for the circular outcome in these models can also take a
+skewed shape. It should be noted that the standard deviation of the
+cross-validation estimates is rather large for the Abe-Ley and the
+CL-GPN model. It is possible that these large standard deviations for
+the PLSL are caused by the fact that they are computed for a relatively
+small sample size, but this does not explain why the PLSL has a large
 standard deviation for only a few cylindrical models and not for all.
-
-```{r ModelFit, cache = TRUE, echo = FALSE}
-
-CLGPNM.ll.circ <- simplify2array(lapply(res_CLGPNM, "[[", "ll.circ"))
-CLGPNM.ll.lin  <- simplify2array(lapply(res_CLGPNM, "[[", "ll.lin"))
-CLGPN.ll.circ  <- simplify2array(lapply(res_CLGPN, "[[", "ll.circ"))
-CLGPN.ll.lin   <- simplify2array(lapply(res_CLGPN, "[[", "ll.lin"))
-CLPN.ll.circ   <- simplify2array(lapply(res_CLPN, "[[", "ll.circ"))
-CLPN.ll.lin    <- simplify2array(lapply(res_CLPN, "[[", "ll.lin"))
-
-CLGPNM.ll.circ.m <- mean(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
-CLGPNM.ll.lin.m  <- mean(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
-CLGPN.ll.circ.m  <- mean(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
-CLGPN.ll.lin.m   <- mean(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
-CLPN.ll.circ.m   <- mean(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
-CLPN.ll.lin.m    <- mean(-2*apply(CLPN.ll.lin[5:20,], 2, mean))
-
-CLGPNM.ll.circ.sd <- sd(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
-CLGPNM.ll.lin.sd  <- sd(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
-CLGPN.ll.circ.sd  <- sd(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
-CLGPN.ll.lin.sd   <- sd(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
-CLPN.ll.circ.sd   <- sd(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
-CLPN.ll.lin.sd    <- sd(-2*apply(CLPN.ll.lin[5:20,], 2, mean))
-
-AL.ll.circ <- simplify2array(lapply(resAL, "[[", "ll.circ"))
-AL.ll.lin <- simplify2array(lapply(resAL, "[[", "ll.lin"))
-
-AL.ll.circ.m <- mean(AL.ll.circ)
-AL.ll.lin.m  <- mean(AL.ll.lin)
-
-AL.ll.circ.sd <- sd(AL.ll.circ)
-AL.ll.lin.sd  <- sd(AL.ll.lin)
-
-PLSL.m <- round(rbind(c(CLPN.ll.circ.m, CLGPN.ll.circ.m, AL.ll.circ.m, CLGPNM.ll.circ.m),
-                      c(CLPN.ll.lin.m, CLGPN.ll.lin.m, AL.ll.lin.m, CLGPNM.ll.lin.m)), 2)
-PLSL.sd <- round(rbind(c(CLPN.ll.circ.sd, CLGPN.ll.circ.sd, AL.ll.circ.sd, CLGPNM.ll.circ.sd),
-                      c(CLPN.ll.lin.sd, CLGPN.ll.lin.sd, AL.ll.lin.sd, CLGPNM.ll.lin.sd)), 2)
-PLSL.sd <- apply(PLSL.sd, 1:2, function(x) paste0("(", sprintf("%.2f", round(x, 2)), ")"))
-
-PLSL <- data.frame(PLSL.m[1,], PLSL.sd[1,], PLSL.m[2,], PLSL.sd[2,])
-
-rownames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN")
-colnames(PLSL) <- c("mean", "sd", "mean", "sd")
-
-#kable(PLSL, "latex", booktabs = T, escape = FALSE,
-#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four #cylindrical models")%>%
-#add_header_above(c("Model" = 1, "Circular" = 2, "Linear" = 2))
-
-```
 
@@
 
-In this paper we modified four models for cylindrical data in such a way that
-they include a regression of both the linear and circular outcome onto a
-set of covariates. Subsequently we have shown how these four methods can
-be used to analyze a dataset on the interpersonal behavior of teachers.
-In this final section we will comment on the differences between these models, the
-results from the analysis of the teacher data and how cylindrical models
-improve the analysis of such cylindrical data.\newline
-\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
-best.  In the
-CL-GPN and GPN-SSN models the interpretation of the parameters of the
-circular outcome component is not straightforward, if at all possible.
-This is caused by the fact that in addition to the mean vector the
-covariance matrix of the GPN distribution affects the location of the
-circular data, making it difficult to compute regression coefficients on
-the circle. @wang2012directional state that Monte Carlo integration
-can be used to compute a circular mean and variance for the GPN
-distribution. In future research, this solution might be applied to the
-methods of @CremersMulderKlugkist2017 in order to compute circular
-coefficients for GPN models.\newline
+In this paper we modified four models for cylindrical data in such a way
+that they include a regression of both the linear and circular outcome
+onto a set of covariates. Subsequently we have shown how these four
+methods can be used to analyze a dataset on the interpersonal behavior
+of teachers. In this final section we will comment on the differences
+between these models, the results from the analysis of the teacher data
+and how cylindrical models improve the analysis of such cylindrical
+data.\newline
+\indent In terms of interpretability, the CL-PN and Abe-Ley models
+perform best. In the CL-GPN and GPN-SSN models the interpretation of the
+parameters of the circular outcome component is not straightforward, if
+at all possible. This is caused by the fact that in addition to the mean
+vector the covariance matrix of the GPN distribution affects the
+location of the circular data, making it difficult to compute regression
+coefficients on the circle. Wang \& Gelfand (2013) state that Monte
+Carlo integration can be used to compute a circular mean and variance
+for the GPN distribution. In future research, this solution might be
+applied to the methods of Cremers et al. (2018b) in order to compute
+circular coefficients for GPN models.\newline
 \indent In terms of flexibility the GPN-SSN model scores best. Multiple
@@
 CL-GPN and CL-PN models are extendable to a mixed-effects structure and
-can thus also be fit to longitudinal data (see @nunez2014bayesian and @hernandez2016general for
+can thus also be fit to longitudinal data (see Nuñez-Antonio \&
+Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al. (2016) for
 hierarchical/mixed-effects models for the PN and GPN distributions
@@
 not been done in previous research for the conditional distribution of
-its circular outcome (sine-skewed von
-Mises). Concerning asymmetry, both the
-GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
-the distributions of both the linear and circular outcome, while the CL-GPN
-model permits an asymmetric  circular outcome.\newline
+its circular outcome (sine-skewed von Mises). Concerning asymmetry, both
+the GPN-SSN as well as the Abe-Ley model allow for non-symmetrical
+shapes of the distributions of both the linear and circular outcome,
+while the CL-GPN model permits an asymmetric circular outcome.\newline
 \indent To investigate model fit for the teacher data, we assessed
@@
 context in this paper are not the only cylindrical distributions
-available from the literature. Other interesting cylindrical distributions have been introduced by
-@fernandez2007models, @kato2008dependent and @sugasawa2015flexible (for more references we
-refer to Chapter 2 of @ley2017modern). In the present study we
-have decided not to include these distributions for reasons of space,
+available from the literature. Other interesting cylindrical
+distributions have been introduced by Fernández-Durán (2007), Kato \&
+Shimizu (2008) and Sugasawa (2015) (for more references we refer to
+Chapter 2 of Ley \& Verdebout (2017)). In the present study we have
+decided not to include these distributions for reasons of space,
 complexity of the models and ease of implementing a regression
@@
 developed in the present study.\newline
-\indent We conclude the paper on a general note regarding cylindrical models.
-They  offer new insights into data of a cylindrical nature in psychology.
-Concerning the example used in this paper, the advantage of cylindrical data
-analysis is that we were able to analyze all circular and linear information in
-the data simultaneously. In previous research, the two components of the
-interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
-separately. Such an approach also provides information about the strength of
-teachers' score on Agency and Communion, yet a large portion of information
-about the combination of Agency and Communion, which describes the kind of
-behavior that is observed,  gets lost. A first solution to include both
-dimensions as a circular variable in data analysis was described by
-@Cremers2018Assessing. A downside of that analysis was that information about
-the strength of the specific type of interpersonal behavior could not be
-retained. In the present study, we have shown how using cylindrical models can
-simultaneously model the information about the type of and strength of
-interpersonal behavior and how these are influenced by teachers' self-efficacy
-in classroom management. The results of the present study therefore provide an
-answer to an often stated problem in data analysis of interpersonal circumplex
-data.
+\indent We conclude the paper on a general note regarding cylindrical
+models. They offer new insights into data of a cylindrical nature in
+psychology. Concerning the example used in this paper, the advantage of
+cylindrical data analysis is that we were able to analyze all circular
+and linear information in the data simultaneously. In previous research,
+the two components of the interpersonal circumplex (\emph{i.e.}, Agency
+and Communion) were analyzed separately. Such an approach also provides
+information about the strength of teachers' score on Agency and
+Communion, yet a large portion of information about the combination of
+Agency and Communion, which describes the kind of behavior that is
+observed, gets lost. A first solution to include both dimensions as a
+circular variable in data analysis was described by Cremers et al.
+(2018a). A downside of that analysis was that information about the
+strength of the specific type of interpersonal behavior could not be
+retained. In the present study, we have shown how using cylindrical
+models can simultaneously model the information about the type of and
+strength of interpersonal behavior and how these are influenced by
+teachers' self-efficacy in classroom management. The results of the
+present study therefore provide an answer to an often stated problem in
+data analysis of interpersonal circumplex data.
 
 \section*{References}
-<div id="refs"></div>
+\DeclareRobustCommand{\VANDER}[3]{#3}
+
+\hypertarget{ref-abe2017tractable}{}
+Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
+for cylindrical data, with applications. \emph{Econometrics and
+Statistics}, \emph{4}, 91--104.
+doi:\href{https://doi.org/10.1016/j.ecosta.2016.04.001}{10.1016/j.ecosta.2016.04.001}
+
+\hypertarget{ref-chrastil2017rotational}{}
+Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
+integration: Encoding and execution errors in angle reproduction.
+\emph{Experimental Brain Research}, \emph{235}(6), 1885--1897.
+doi:\href{https://doi.org/10.1007/s00221-017-4910-y}{10.1007/s00221-017-4910-y}
+
+\hypertarget{ref-Claessens2016side}{}
+Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
+Teachers' perceptions of teacher--student relationships} (PhD thesis).
+
+\hypertarget{ref-Cremers2018Assessing}{}
+Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
+Bayesian embedding approach to circular regression models.
+\emph{Methodology}, \emph{14}(2), 69--81.
+
+\hypertarget{ref-CremersMulderKlugkist2017}{}
+Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
+interpretation of regression coefficients. \emph{British Journal of
+Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
+doi:\href{https://doi.org/10.1111/bmsp.12108}{10.1111/bmsp.12108}
+
+\hypertarget{ref-fernandez2007models}{}
+Fernández-Durán, J. (2007). Models for circular--linear and
+circular--circular data constructed from circular distributions based on
+nonnegative trigonometric sums. \emph{Biometrics}, \emph{63}(2),
+579--585.
+doi:\href{https://doi.org/10.1111/j.1541-0420.2006.00716.x}{10.1111/j.1541-0420.2006.00716.x}
+
+\hypertarget{ref-fisher1995statistical}{}
+Fisher, N. I. (1995). \emph{Statistical analysis of circular data}.
+Cambridge: Cambridge University Press.
+
+\hypertarget{ref-garcia2014test}{}
+García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
+González-Manteiga, W., \& Pereira, J. (2014). A test for
+directional-linear independence, with applications to wildfire
+orientation and size. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{28}(5), 1261--1275.
+doi:\href{https://doi.org/10.1007/s00477-013-0819-6}{10.1007/s00477-013-0819-6}
+
+\hypertarget{ref-garcia2013exploring}{}
+García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
+Exploring wind direction and SO2 concentration by circular--linear
+density estimation. \emph{Stochastic Environmental Research and Risk
+Assessment}, \emph{27}(5), 1055--1067.
+doi:\href{https://doi.org/10.1007/s00477-012-0642-5}{10.1007/s00477-012-0642-5}
+
+\hypertarget{ref-BDA}{}
+Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
+(2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman
+\& Hall/CRC.
+
+\hypertarget{ref-gneiting2007strictly}{}
+Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
+prediction, and estimation. \emph{Journal of the American Statistical
+Association}, \emph{102}(477), 359--378.
+doi:\href{https://doi.org/10.1198/016214506000001437}{10.1198/016214506000001437}
+
+\hypertarget{ref-gurtman2009exploring}{}
+Gurtman, M. B. (2009). Exploring personality with the interpersonal
+circumplex. \emph{Social and Personality Psychology Compass},
+\emph{3}(4), 601--619.
+doi:\href{https://doi.org/10.1111/j.1751-9004.2009.00172.x}{10.1111/j.1751-9004.2009.00172.x}
+
+\hypertarget{ref-hernandez2016general}{}
+Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
+(2016). The general projected normal distribution of arbitrary
+dimension: Modeling and Bayesian inference. \emph{Bayesian Analysis},
+\emph{12}(1), 113--133.
+doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}
+
+\hypertarget{ref-horowitz2010handbook}{}
+Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
+psychology: Theory, research, assessment, and therapeutic
+interventions}. Hoboken, NJ: John Wiley \& Sons.
+
+\hypertarget{ref-jammalamadaka2001topics}{}
+Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
+statistics} (Vol. 5). World Scientific.
+
+\hypertarget{ref-johnson1978some}{}
+Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
+distributions and related regression models. \emph{Journal of the
+American Statistical Association}, \emph{73}(363), 602--606.
+
+\hypertarget{ref-kato2008dependent}{}
+Kato, S., \& Shimizu, K. (2008). Dependent models for observations which
+include angular ones. \emph{Journal of Statistical Planning and
+Inference}, \emph{138}(11), 3538--3549.
+doi:\href{https://doi.org/10.1016/j.jspi.2006.12.009}{10.1016/j.jspi.2006.12.009}
+
+\hypertarget{ref-lagona2015hidden}{}
+Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
+Markov approach to the analysis of space--time environmental data with
+linear and circular components. \emph{Stochastic Environmental Research
+and Risk Assessment}, \emph{29}(2), 397--409.
+doi:\href{https://doi.org/10.1007/s00477-014-0919-y}{10.1007/s00477-014-0919-y}
+
+\hypertarget{ref-ley2017modern}{}
+Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
+Chapman \& Hall/CRC Press. Boca Raton, FL.
+
+\hypertarget{ref-mardia2000directional}{}
+Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
+(Vol. 494). Chichester, England: Wiley.
+
+\hypertarget{ref-mardia1978model}{}
+Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical
+variables with applications. \emph{Journal of the Royal Statistical
+Society Series B (Methodological)}, \emph{40}(2), 229--233.
+
+\hypertarget{ref-mastrantonio2018joint}{}
+Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
+distribution for poly-cylindrical data. \emph{Journal of Multivariate
+Analysis}, \emph{165}, 14--26.
+doi:\href{https://doi.org/10.1016/j.jmva.2017.11.006}{10.1016/j.jmva.2017.11.006}
+
+\hypertarget{ref-mastrantonio2015bayesian}{}
+Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
+hidden Markov modelling using circular-linear general projected normal
+distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
+doi:\href{https://doi.org/10.1002/env.2326}{10.1002/env.2326}
+
+\hypertarget{ref-nunez2014bayesian}{}
+Nuñez-Antonio, G., \& Gutiérrez-Peña, E. (2014). A Bayesian model for
+longitudinal circular data based on the projected normal distribution.
+\emph{Computational Statistics \& Data Analysis}, \emph{71}, 506--519.
+doi:\href{https://doi.org/10.1016/j.csda.2012.07.025}{10.1016/j.csda.2012.07.025}
+
+\hypertarget{ref-nunez2011bayesian}{}
+Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
+Bayesian regression model for circular data based on the projected
+normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
+185--201.
+doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
+
+\hypertarget{ref-olmos2012extension}{}
+Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
+extension of the half-normal distribution. \emph{Statistical Papers},
+\emph{53}(4), 875--886.
+doi:\href{https://doi.org/10.1007/s00362-011-0391-4}{10.1007/s00362-011-0391-4}
+
+\hypertarget{ref-pennings2018interpersonal}{}
+Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C.,
+\VANDER{Want}{Van der}{van der} Want, A. C., \& Tartwijk, J. van.
+(2018). Interpersonal adaptation in teacher-student interaction.
+\emph{Learning and Instruction}, \emph{55}, 41--57.
+doi:\href{https://doi.org/10.1016/j.learninstruc.2017.09.005}{10.1016/j.learninstruc.2017.09.005}
+
+\hypertarget{ref-rayner200935th}{}
+Rayner, K. (2009). The 35th Sir Frederick Bartlett lecture: Eye
+movements and attention in reading, scene perception, and visual search.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8),
+1457--1506.
+doi:\href{https://doi.org/10.1080/17470210902816461}{10.1080/17470210902816461}
+
+\hypertarget{ref-sahu2003new}{}
+Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
+multivariate skew distributions with applications to Bayesian regression
+models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
+doi:\href{https://doi.org/10.2307/3316064}{10.2307/3316064}
+
+\hypertarget{ref-sugasawa2015flexible}{}
+Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the
+cylinder}. Retrieved from
+\href{arXiv:\%201501.06332v2}{arXiv: 1501.06332v2}
+
+\hypertarget{ref-wang2012directional}{}
+Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
+general projected normal distribution. \emph{Statistical Methodology},
+\emph{10}(1), 113--127.
+doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}
+
+\hypertarget{refs}{}
+\hypertarget{ref-vanderWant2015role}{}
+\VANDER{Want}{Van der}{van der} Want, A. C. (2015). \emph{Teachers'
+interpersonal role identity.} (PhD thesis).
+
+\hypertarget{ref-wubbels2006interpersonal}{}
+Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
+An interpersonal perspective on classroom management in secondary
+classrooms in the Netherlands. In C. Evertson \& C. S. Weinstein (Eds.),
+\emph{Handbook of classroom management: Research, practice, and
+contemporary issues} (pp. 1161--1191). Malwah, NJ: Lawrence Erlbaum
+Associates.
+
+\end{document}

### 2019-01-28 Manuscript/ManuscriptMarkdownCV.Rmd
Update manuscript

@@
 ---
-title: "Regression models for Cylindrical data in Psychology"
-author: #|
- # | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
- # | $^1$Department of Methodology and Statistics, Utrecht University
- # | $^2$TNO
-  #| $^3$Department of Education, Utrecht University
-  #| $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
-
-bibliography: CircularData.bib
-csl: apa.csl
-output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    number_sections: yes
-  html_document: default
-  word_document: default
-fontsize: 12pt
-geometry: margin=1in
-header-includes:
-  - \usepackage{multirow}
-  - \usepackage{appendix}
-  - \usepackage{color}
-  - \usepackage{hyperref}
-  - \usepackage{subcaption}
-  - \setlength\parindent{24pt}
-  - \usepackage[nodisplayskipstretch]{setspace}\doublespacing
-  #- \usepackage{endfloat}
-  - \DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
-
-keywords: bla vla
-documentclass: article
-pandoc_args: --natbib
+title             : "Regression models for Cylindrical data in Psychology"
+shorttitle        : "Cylindrical data in Psychology"
+
+author: 
+  - name          : "Jolien Cremers"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Postal address"
+    email         : "joliencremers@gmail.com"
+  - name          : "Helena J.M. Pennings"
+    affiliation   : "2,3"
+  - name          : "Christophe Ley"
+    affiliation   : "4"
+
+affiliation:
+  - id            : "1"
+    institution   : "Department of Methodology and Statistics, Utrecht University"
+  - id            : "2"
+    institution   : "TNO"
+  - id            : "3"
+    institution   : "Department of Education, Utrecht University"
+  - id            : "4"
+    institution   : "Department of Applied Mathematics, Computer Science and Statistics, Ghent University"
+
+authornote: |
+  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+
+  Enter author note here.
+
 abstract: |
@@
   interpersonal behavior of teachers.
+
+keywords          : "cylindrical data, regression, interpersonal behavior"
+wordcount         : "X"
+
+bibliography      : ["CircularData.bib"]
+
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_pdf
+
+header-includes:
+  - \DeclareRobustCommand{\VANDER}[3]{#2} 
+  - \usepackage{multirow}
+  - \usepackage{rotating}
+  - \usepackage{color}
+  - \usepackage{subcaption}
+   #% set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
+
 ---
@@
 ```{r, echo = FALSE}
+library(papaja)
 library(circular)
@@
 library(caret) #cross-validation
+library(coda) #geweke diagnostic
 ```
@@
 
-Cylindrical data are data that consist of a linear variable and a directional
-variable. In this paper, the directional variable is circular, meaning that it
-consists of a single angle instead of a set of angles. A circular variable is
-different from a linear variable in the sense that it is measured on a different
-scale. Figure \ref{circline} shows the difference between a circular scale
-(right) and a linear scale (left). The most important difference is that on a
-circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in
-fact represent the same number while on a linear scale the two ends, \(-\infty\)
-and \(\infty\), are not connected and consequently the values 0\(^\circ\) and
-360\(^\circ\) are located on different places on the scale. This difference
-requires us to use special statistical methods for circular variables (see e.g.
-@fisher1995statistical for an introduction to circular data and
+In social sciences the use of cylindrical data is very common. Such data consist
+of a linear and a circular component. @gurtman2011reasoning refers to such data
+as vectors, with a directional measure (i.e., the circular component) and a
+measure indicating the magnitude (i.e., linear component). Many established
+models in psychology are often referred to as circular or circumplex models, but
+those models are cylindrical. Examples of such cylindrical models are the
+interpersonal circle/circumplex
+[@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
+affect [@russell1980circumplex], the circumplex of human emotion
+[@plutchik1997general] or the model of human values [@schwartz1992values].
+\newline
+\indent Also, many of the more recent types of data that are studied in psychology are
+cylindrical. For example, research on human navigation uses data where distance
+(i.e., linear component) and direction (i.e., the linear component) are of
+interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
+consist of both the direction (i.e., the circular variable) and the duration
+(i.e., the linear variable)(e.g., @rayner200935th). Apart from the social
+sciences data with a circular and linear outcome more commonly occur in
+meteorology meteorology [@garcia2013exploring], ecology [@garcia2014test] or
+marine research [@lagona2015hidden]
+\newline
+\indent Up until now researchers studying cylindrical data had to rely on the linear
+statistical methods to analyze their research results. Focusing on only the
+circular variable or only on the linear variable. However, lately more and more
+of these researchers acknowledge that linear methods are not sufficient and call
+for new methods [@gurtman2011reasoning; @pennings2017complexity;
+@wright2009integrating] that take into account both the circular and the linear
+component of these data.
+\newline
+\indent The aim in the present paper is twofold. Firstly, we intend to fill this gap in
+the literature on cylindrical models by adapting four existing cylindrical
+models in such a way that they include a regression of both the linear and
+circular component of a cylindrical variable onto a set of covariates. From now
+on we will therefore refer to the components of the cylindrical variable as
+outcome components. Secondly, we will show how using these models can benefit
+the analysis of circumplex data and cylindrical data in psychology in general.
+More specifically we will show these benefits for interpersonal teacher data
+from the field of educational psychology. Apart from modelling the dependence
+between the linear and circular component of a cylindrical variable we would
+also like to predict the two components from a set of covariates in a regression
+model. In our analysis, we will show how this can be done in four different ways
+using the adapted cylindrical models.
+
+\subsection{Cylindrical data}
+
+Thus, data that consist of a variable and a circular variable is called
+cylindrical data. A circular variable is different from the linear variable in
+the sense that it is measured on a different scale. Figure \ref{circline} shows
+the difference between a circular scale (right) and a linear scale (left). The
+most important difference is that on a circular scale the datapoints 0\(^\circ\)
+and 360\(^\circ\) are connected and in fact represent the same number while on a
+linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
+consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
+places on the scale. Both circular data and cylindrical data require special
+analysis methods due to this periodicity in the scale of a circular variable
+(see e.g. @fisher1995statistical for an introduction to circular data and
 @mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
-  elaborate overview). For instance, a notion as simple as the sample average
-  needs to be re-defined. As is the case for circular data, the analysis of
-  cylindrical data also requires special methods.\newline
-\indent Cylindrical data occur in several fields of research, such as for
-instance in meteorology [@garcia2013exploring], ecology [@garcia2014test] or
-marine research [@lagona2015hidden]. However, also in psychology several types
-of cylindrical data can be found. For example, in research on human navigation
-in the field of cognitive psychology both the distance, a linear variable, and
-the direction, a circular variable, of movement are of interest
-[@chrastil2017rotational]. In eye-tracking research, saccade data are an example
-of cylindrical data, because both the direction (\emph{i.e.}, the circular variable)
-and the duration (\emph{i.e.}, the linear variable) of the saccades are of interest
-(for a review of eye-tracking research see  @rayner200935th).\newline
-\indent The type of data that is used in the present study are also
-psychological, namely data from circumplex measurement instruments. For
-instance, data from the interpersonal circumplex as used in personality
-psychology are by definition of a cylindrical nature (see Section \ref{Example}
-for a more detailed explanation).\newline
+elaborate overview).\newline
 \indent In the literature, several methods have been put forward to model the
@@
 proposed thus far include additional covariates onto which both the circular and
-linear component are regressed.\newline
-\indent Our aim in this paper is to fill this gap in the literature by adapting
-four existing cylindrical models in such a way that they include a regression of
-both the linear and circular component of a cylindrical variable onto a set of
-covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. Additionally, we will show how a
-correct statistical treatment of such cylindrical data can lead to new insights.
-We will do this for the teacher data, a dataset from the field of educational
-psychology. In the teacher data, apart from modelling the relation between the
-linear and circular component of a cylindrical variable we would also like to
-predict the two components from a set of covariates in a regression model.
-
-The paper is organized as follows: Section \ref{Example} describes the teacher
-data, while Section \ref{Models} presents the four cylindrical models and our
-associated adaptations to new regression models. In that same section we also
-discuss the model fit criterion that we will use in Section \ref{DataAnalysis}
-for the comparison of the four models. A detailed data analysis with interesting
-new insights is also provided in Section \ref{DataAnalysis}. We conclude the
-paper with a discussion in Section \ref{Discussion}, and the Supplementary Material collects
-the technical details of the MCMC procedures.
+linear component are regressed.
 
@@
 
-\section{Teacher data}\label{Example}
+\section{Teacher Data}\label{Example}
 
@@
 psychology and was collected for the studies on classroom climate of
-@vanderWant2015role, @Claessens2016side and @pennings2018interpersonal. An
+@vanderWant2015role, @Claessens2016side and @pennings2017phd. An
 indicator of the quality of the classroom climate is the students' perception of
@@
 and the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular
-variable.\newline
+also implies that scores on the IPC could be viewed as a circular variable. This
+circular variable represents the type of interpersonal behavior that a teacher
+shows towards his/her students.\newline
 \indent @Cremers2018Assessing explain the circular nature of the IPC data and
-analyze them as such using a circular regression model. The two-dimension scores
-Agency and Communion can be converted to a circular score using the two-argument
-arctangent function in \eqref{PredVal}, where \(A\) represents a score on the
-Agency dimension and \(C\) represents a score on the Communion dimension
+analyze them as such using a circular regression model. The two dimension
+scores, Agency and Communion, can be converted to a circular score using the
+two-argument arctangent function in \eqref{PredVal}\footnote{The selection of
+the origin in circumplex data depends on the scaling of the Agency and Communion
+scores; their respective 0 scores form the origin. Agency and Communion are
+measured on a scale from 1 to 5 and for analysis purposes they are later
+converted to scale ranging from -1 to 1. Scaling is however only considered an
+issue in those instances where cylindrical data is derived from measurements in
+bivariate space.}, where \(A\) represents a
+score on the Agency dimension and \(C\) represents a score on the Communion
+dimension. Note that when placing a unit circle on Figure \ref{QTI} we see that
+the Agency dimension is related to the sine of the circular score and the
+Communion dimension is related to the cosine of the circular score.
 \begin{equation}\label{PredVal}
@@
 The resulting circular variable \(\theta\) can then be modelled and takes values
-in the interval $[0, 2\pi)$. However, when two-dimensional data are converted
-to the circle we lose some information, namely the length of the two-dimensional
-vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\).
-This length represents the strength of the type of interpersonal behavior a
-teacher shows towards his/her students and can be considered as the linear
-variable in a cylindrical model, allowing us to model a circular variable
-\(\theta\) together with the linear variable corresponding to \(\mid\mid (A, C)^t
-\mid\mid\). This leads to an improved analysis of interpersonal circumplex data
-as we take all information into account. In the next section we introduce
-several models that can be used for a more accurate and informative regression
-analysis on the teacher data. First however we will provide descriptives for our
-data set.
-
-\subsection{Data description}\label{DataDescriptives}
-
-```{r loaddata, return = FALSE, echo = FALSE}
+in the interval $[0, 2\pi)$ or $[0^\circ, 360^\circ)$. Note that the round
+brackets mean that $2\pi$ and $360^\circ$ are not included in the interval since
+these represent the same value as 0 as a result of periodicity. \newline
+\indent A circular analysis of circumplex data has several benefits: it is more
+in line with its theoretical definition and it allows us to analyse the blend of
+the two dimensions Agency and Communion instead of both dimensions separately
+providing us with new insights compared to a separate analysis of the two
+dimensions that is standard in the literature (see e.g. @pennings2018interpersonal, @wright2009integrating or @wubbels2006interpersonal). There is however one main drawback: when two-dimensional
+data are converted to the circle we lose some information, namely the length of
+the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm
+\(\mid\mid (A, C)^t \mid\mid\). This length represents the strength of the
+interpersonal behavior a teacher shows towards his/her students. In a
+cylindrical model this strength (the linear outcome) can be modeled together
+with the type of interpersonal behavior of a teacher (the circular outcome).
+This leads to an improved analysis of interpersonal circumplex data, over either
+analyzing the two dimensions separately or using a circular model, because we
+take all information, circular and linear, into account. In the next section we
+introduce several cylindrical models that can be used to analyze the teacher
+data. First however we will provide descriptives for the teacher data.
+
+\subsection{Data Description}\label{DataDescriptives}
+
+```{r loaddata, return = FALSE, echo = FALSE, results = FALSE}
 Dat <- read_spss("Data Heleen/mergeddata.sav")
@@
 Dat <- Dat[!duplicates,]
+
+nrow(Dat)
 
@@
 
-plot(Dat$y, Dat$theta*(180/pi), 
-     ylab = "IPC location", xlab = "IPC strength",
+SE_low = Dat[Dat$Efficacy_CM <= 4.04,]
+SE_mid = Dat[Dat$Efficacy_CM > 4.04 & Dat$Efficacy_CM < 6.04,]
+SE_high = Dat[Dat$Efficacy_CM >= 6.04,]
+
+summary(SE_low$y)
+summary(SE_mid$y)
+summary(SE_high$y)
+
+plot(SE_low$y, SE_low$theta*(180/pi), 
+     ylab = "IPC", xlab = "IPC strength",
      bty = "n", ylim = c(-180, 180))
+points(SE_mid$y, SE_mid$theta*(180/pi), pch = 2)
+points(SE_high$y, SE_high$theta*(180/pi), pch = 3)
+legend(0.45, -75, legend = c("low SE", "average SE", "high SE"), pch = c(1,2,3), bty = "n")
 
@@
 consider one measurement, the first occasion (2010) and largest class if data
-for multiple classes were available. In addition to the score on the IPC, the
-circular outcome, and the strength of the score on the IPC, the linear outcome,
-a teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
-covariate in the analysis. After listwise deletion of missings ($3$ in total,
-only for the self-efficacy) we have a sample of 148 teachers. Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset. Here
-$\hat{\rho}$ is a sample estimate for the circular concentration where a value
-of 0 means that the data is not concentrated at all, \emph{i.e.} spread over the
-entire circle, and a value of 1 means that all data is concentrated at a single
-point on the circle. Figure \ref{dataplot} is a scatterplot showing the relation
-between the linear and circular outcome of the teacher data.
+for multiple classes were available. This results in a sample of 151 teachers. In addition to the type of interpersonal
+behavior (IPC), the circular outcome, and the strength of interpersonal behavior
+(IPC strength), the linear outcome, a teachers' self-efficacy (\verb|SE|)
+concerning classroom management is used as covariate in the analysis.
+In previous research, in psychology and education it had been shown that higher self-efficacy is related to the quality of interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
+After listwise
+deletion of missings ($3$ in total, only for the self-efficacy) we have a sample
+of 148 teachers. Table \ref{Tableteacherdescriptives} shows descriptives for the
+dataset. For the circular variable IPC we show sample estimates for the circular
+mean $\bar{\theta}$ and concentration $\hat{\rho}$. The circular concentration
+lies between 0, meaning the data is not concentrated at all \emph{i.e.} spread
+over the entire circle, and 1, meaning all data is concentrated at a single
+point on the circle The population values of these parameters are usually, and
+also in this paper, referred to as $\mu$ (circular location) and $\kappa$
+(circular concentration). For the linear variables (strength IPC and SE) we show
+sample estimates of the linear mean and standard deviation. Figure
+\ref{dataplot} is a scatterplot showing the relation between the linear and
+circular outcome of the teacher data for teachers with low SE (below 1 sd below
+the mean), average SE (between 1 sd below and 1 sd above the mean) and high SE
+(above 1 sd above the mean).
 
@@
    \hline
+\multicolumn{5}{l}{Note: For the circular variable IPC we show sample }\\
+\multicolumn{5}{l}{estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.}\\
+\multicolumn{5}{l}{For the linear variable we show the sample mean,}\\
+\multicolumn{5}{l}{standard deviation and range.}
 \end{tabular}
@@
 
-\section{Four cylindrical regression models}\label{Models}
-
-In this section we present four cylindrical models and adapt them such
-that they contain predictors for the linear and circular outcomes, \(Y\)
-and \(\Theta\). The first two models are based on a construction by
-@mastrantonio2015bayesian, while the other models are extensions of the
-models from @abe2017tractable and @mastrantonio2018joint.
-
-\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}
-
-Following @mastrantonio2015bayesian we consider in this section two models where
+\section{Four Cylindrical Regression Models}\label{Models}
+
+One of the goals of this paper is to show the benefits of cylindrical methods
+for the analysis of circumplex data and cylindrical data in psychology in
+general. To do so we decide to focus on four cylindrical models. The models were
+selected for their relatively low complexity and the ease with which a
+regression structure could be incorporated. But also because they show different
+ways of modelling the the linear and circular outcome and thereby illustrate a
+wider range of cylindrical models available in the literature. All four
+cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
+x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_q$ for the linear and
+circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
+based on a construction by @mastrantonio2015bayesian, while the other models are
+extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
+four cylindrical  models are introduced separately in the subsections below.
+However, to provide a more succinct overview and comparison of the four models,
+\ref{TableModels} gives an overview of the similarities and differences between
+the models.
+
+\begin{sidewaystable}[h]
+\centering
+\caption{Comparison of the four cylindrical regression models} 
+\begin{tabular}{lllll}
+  \noalign{\smallskip}\hline\noalign{\smallskip}
+\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
+$\Theta$ & &&&\\
+$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
+$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
+$\:\:$Shape & symmetric, & assymetric, & assymetric, & assymetric, \\
+            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
+$Y$& &&&\\
+$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
+$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(0, + \infty)$\\
+$\:\:$Shape & symmetric, & symmetric, & assymetric, & assymetric, \\
+            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
+$\Theta$-$Y$ dependence &                                   &                                   & & \\
+                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
+                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
+\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
+\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
+
+\end{tabular}
+\label{TableModels}
+\end{sidewaystable}
+
+\subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+
+Following @mastrantonio2015bayesian we consider two models where
 the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
@@
 intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
-covariates. In both of these models the
-conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
-by
-\begin{equation}\label{ycondtheta}
-f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
-\end{equation}
-\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
-\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
-\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
-\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
-in its location part. For the teacher dataset, the regression equation for the
-linear outcome in the CL-PN and CL-GPN model is the following: \[\hat{y}_i =
-\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i +
-\gamma_1\text{SE}_i,\]
-\noindent where $\text{SE}_i$ is the self-efficacy score of one individual $i = 1, \dots, n$ where $n$ is the sample size.\newline
-\indent For the circular outcome we assume either a projected normal (PN) or a general
-projected normal (GPN) distribution. These distributions arise from the radial
-projection of a distribution defined on the plane onto the circle. The relation
-between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
-outcome \(\Theta\) is defined as follows
+covariates for the prediction of the linear outcome. We thus assume a normal
+distribution for the linear outcome.\newline
+\indent For the circular outcome we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise from a
+projection of a distribution defined in bivariate space onto the circle. Figure
+\ref{projection} represents this projection. In the left plot of Figure
+\ref{projection} we see datapoints from the bivariate variable
+\(\boldsymbol{S}\) that in the middle plot are projected to form the circular
+outcome \(\Theta\) in the right plot. Mathematically the relation between
+\(\boldsymbol{S}\) and \(\Theta\) is defined as follows
 \begin{equation}\label{projection}
@@
 \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
-the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
-\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
-assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
-$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1
-\end{bmatrix}\), \(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
-@hernandez2016general). This leads to the
-circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
-will now detail how we modify both cylindrical distributions to also incorporate
-covariates for the circular part.
-
-\subsubsection{The modified CL-PN distribution}
-
-Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
-PN distribution equals
-\begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
-\end{equation}
-\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
-\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
-\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
-the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
-\(n\) is the sample size, are generated independently from the distribution with
-density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
-is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
-where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
-contains the covariate values and the value 1 to estimate an intercept and
-\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
-the regression coefficients and intercepts. Note however that the dimensions of
-\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{II}\) need not necessarily
-be the same and we are thus allowed to have a different set of predictor
-variables and vectors \(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for
-the two components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
-regression equation for the circular outcome in the CL-PN model is
-\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix} \mu_{i}^{I}  \\ \mu_{i}^{II}
-\end{pmatrix}=\begin{pmatrix} \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
-\beta_0^{II} + \beta_1^{II}\text{SE}_i \end{pmatrix}.\]
-
-\subsubsection{The modified CL-GPN distribution}
-
-Following @wang2012directional and @hernandez2016general the joint density of
-\(R\) and \(\Theta\) for the GPN distribution equals
-\begin{equation}\label{gpnreg}
-f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
-\end{equation}
-\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
-\rho^2 & \rho\\ \rho & 1 \end{bmatrix}\). In a regression setup the outcomes
-\(\theta_i\) and \(r_i\) for each individual are generated independently from
-\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
-defined in the same way via covariates as for the modified CL-PN distribution.
-Note in contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
-\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the same
-predictors for both components of \(\boldsymbol{\mu}_i\) in the CL-GPN model.
-This is due to the fact that the variance-covariance matrix
-\(\boldsymbol{\Sigma}\) is no longer identity. For the teacher dataset, the
-regression equation for the circular outcome in the CL-GPN model is the same as
-in the CL-PN model.
-
-\subsubsection{Parameter estimation}
-
-Both cylindrical models introduced here are estimated using Markov Chain Monte
+\(\boldsymbol{S}\); the lines connecting the bivariate datapoints to the origin
+in the middle plot. We call $\boldsymbol{S}$ the augmented representation of the
+circular outcome, it is a variable that in contrast to \(\Theta\) is not
+observed and thus considered latent or auxiliary. This then means that we do not
+model \(\Theta\) directly but indirectly through $\boldsymbol{S}$. \newline
+\indent For both the PN and GPN distribution the circular location parameter
+\(\mu\in [0, 2\pi)\) is modeled as \(\hat{\mu}_i =
+\mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) =
+\mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i,
+\boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
+model the circular location parameter also depends on the variance-covariance
+matrix and the circular predicted values should be computed using numerical
+integretion or Monte Carlo methods because a closed form expression for the mean
+direction is not available.} where $\boldsymbol{\beta}^{I}$ and
+$\boldsymbol{\beta}^{II}$ are vectors with intercepts and regression
+coefficients for the prediction of $\boldsymbol{S}$ and $\boldsymbol{z}_i$ is a
+vector with predictor values for each individual $i \in 1, \dots, n$ where $n$
+is the sample size. Note that as a result of the augmented representation of the
+circular outcome we have two sets of regression coefficients and intercepts, one
+for each bivariate component of $\boldsymbol{S}$. This leads to problems when we
+want to interpret the effect of a covariate on the circle. A circular regression
+line is shown in Figure \ref{circregline}, with covariate values on the x-axis
+and the predicted circular outcome on the y-axis. As can be seen it is of a
+non-linear character meaning that the effect of a covariate is different at
+different values of the covariate. A circular regression line is usually
+described by the slope at the inflection point, the point at which the slope of
+the regression line starts flattening off (indicated with a square in Figure
+\ref{circregline}). By default, the parameters from the PN and GPN model do not
+directly describe this inflection point. For the PN distribution however,
+@CremersMulderKlugkist2017 solved this interpretation problem and introduce new
+  circular regression coefficients. They introduce a new parameter $b_c$ that
+  describes the slope at the inflection point of the regression line. For the
+  GPN distrbution the interpretation problem however remains.\newline
+\indent The main difference between the PN and GPN distribution lies in the
+definition of their covariance matrix. For the PN distribution this is an
+identity matrix, causing the distribution to be unimodal and symmetric, whereas
+for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2
+& \xi\\ \xi & 1 \end{bmatrix}\), allowing for multimodality and
+assymetry/skewness.\newline
+\indent For the teacher dataset the predicted linear outcome, strengths of
+interpersonal behavior, in the CL-PN and CL-GPN model is the following:
+\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted circular
+outcome, type of interpersonal behavior, equals: \[\hat{\theta}_i = \mu_i =
+\mbox{atan2} (\beta_0^{II} + \beta_1^{II}\text{SE}_i, \beta_0^{I} +
+\beta_1^{I}\text{SE}_i).\] \noindent where $\text{SE}_i$ is the self-efficacy
+score of one individual. The CL-PN and CL-GPN models thus allow us to assess the
+average type and strength of interpersonal behavior through the parameters
+$\beta_{0}^{I}$, $\beta_{0}^{II}$ and $\gamma_0$ as well as the effect of
+self-efficacy on type and strength of teacher behavior through the parameters,
+$\beta_{1}^{I}$, $\beta_{1}^{II}$ and $\gamma_1$. In addition, because the type
+and strength of interpersonal behavior are modelled together via the regression
+in \eqref{circlinlink} we can assess the effect of the type of interpersonal
+behavior on the strength through the parameters $\gamma_{sin}$ and
+$\gamma_{cos}$. In the teacher data these are the regression coefficients for
+the effect of the sine and cosine of the type of behavior which are related to
+the scores on the Agency an Communion dimensions respectively.\newline
+\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
 Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@@
 
-\subsection{The modified Abe-Ley model}\label{WeiSSVM}
+```{r plotprojecting, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+library(MASS)
+library(plotrix)
+
+set.seed(10200)
+
+Y <- mvrnorm(10, c(2,2), (diag(2)))
+
+theta <- atan2(Y[,2], Y[,1])
+
+tikz("Plots/plotprojecting.tex", standAlone =TRUE, height = 3, width = 9, pointsize = 12, engine = "pdftex")
+
+par(mfrow = c(1,3), oma =c(0,0,0,0), mar = c(0,0,0,0))
+
+plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+points(x = cos(theta), y = sin(theta))
+segments(x0 = cos(theta), x1 = Y[,1], y0 = sin(theta), y1 = Y[,2])
+segments(x0 = cos(theta), x1 = 0, y0 = sin(theta), y1 = 0, lty = 2)
+
+plot(x = cos(theta), y = sin(theta), ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+dev.off()
+
+tools::texi2dvi("Plots/plotprojecting.tex",pdf=TRUE)
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
+\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
+The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norm of the bivariate datapoints, the random variable $R$.}
+\label{projection}
+\end{figure}
+
+```{r circregline, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+
+tikz("Plots/circregline.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+set.seed(101)
+
+Xreal <- rnorm(100, -2, 1.5)
+
+CompI  <- 1 + 0.3*Xreal + rnorm(100, 0.1,1)
+CompII <- 2 + -0.4*Xreal + rnorm(100, 0.1,1)
+
+Outreal <- atan2(CompII, CompI)
+
+X <- seq(-5, 5, 0.1)
+
+predCompI  <- 1.1 + 0.25*X
+predCompII <- 2.1 + -0.4*X
+
+PredCirc <- atan2(predCompII, predCompI)
+
+axval <- ((1.1*0.25)+(2.1*-0.4))/(1.1^2 + 2.1^2)
+acval <- atan2(2.1 + -0.4*axval, 1.1 + 0.25*axval)
+
+plot(X, PredCirc*(180/pi), ylim = c(-200,200), type = "l", xlab = "Predictor",
+     ylab = "Circular Outcome", bty = 'n', yaxt = "n", xaxt = "n")
+axis(1, xaxp = c(-5, 0, 5), at = c(seq(-5, 5, by = 1)),
+     labels = c(seq(-5, 5, by = 1)), las = 2)
+axis(2, xaxp = c(-180, 0,180), at = c(seq(-180, 180, by = 60)),
+     labels = c(seq(-180, 180, by = 60)), las = 2)
+points(Xreal, Outreal*(180/pi), cex=0.5)
+points(axval,
+       acval*(180/pi),
+       pch = 0)
+
+dev.off()
+tools::texi2dvi("Plots/circregline.tex",pdf=TRUE)
+
+```
+
+\begin{figure}[]
+  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
+  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to.
+The square indicates the inflection point of the regression live.}
+  \label{circregline}
+\end{figure}
+
+\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
 
 This model is an extension of the cylindrical model introduced in
-@abe2017tractable to the regression context. The joint density of \(\Theta\) and
-\(Y\), in this model defined only on the positive real half-line \([0, +
-\infty)\), reads
-\begin{equation}\label{WeiSSVMdensity}
-f(\theta, y) = \frac{\alpha\beta^\alpha}{2\pi\cosh(\kappa)}
-                 (1 +\lambda\sin(\theta - \mu))
-                 y^{\alpha-1}
-                 \exp[-(\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
-\end{equation}
-\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
-\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
-\(\kappa\) also regulating the circular-linear dependence. Our modification
-occurs at the level of the linear scale parameter \(\beta>0\) and circular
-location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
-covariates: \(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
-\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\). The
-parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
-\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
-\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\eta_0 \in [0,
-2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a vector of \(p\)
-regression coefficients \(\eta_j \in (-\infty, +\infty)\) for the prediction of
+@abe2017tractable to the regression context. It concerns a combination
+of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
+\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
+with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
+\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
+contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
+model defined only on the positive real half-line \([0, + \infty)\) and thus can
+not be negative.\newline
+\indent In this model we predict the linear scale parameter and circular
+location parameter, both of which we can express in terms of
+covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
+\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
+parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
+\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
+\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
+2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
+regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
 \(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
 vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
-is a vector of predictor values for the prediction of \(\theta\). In a
-regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
-generated independently from the modified
-density \eqref{WeiSSVMdensity}.\newline
-\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
-\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
-\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
-distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
-with location parameter \(\mu\) and concentration parameter \((\beta
-y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
-\begin{align}\label{WeiSSVMLikelihood}
-l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
-   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\nu} \nonumber\\
-   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
-   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
-   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))).\nonumber
-\end{align}
-\noindent For the teacher data, $\boldsymbol{z} = \boldsymbol{x}$ and the regression equations for the circular and
-linear outcomes in the Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
-\tan^{-1}(\eta_1\text{SE}_i),\] and \[\hat{\beta}_{i} = \exp(\nu_0 +
-\nu_1\text{SE}_i).\]
-\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
+is a vector of predictor values for the prediction of \(\theta\).\newline
+\indent For the teacher data, the predicted values for the circular outcome in
+the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
+\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
+The conditional distribution for the linear outcome is Weibull, meaning that we
+can use methods from survival analsis to interpret the effect of a predictor. In
+survival analysis a 'survival' function is used in which time is plotted against
+the probability of survival of subjects suffering from a specific medical
+condition. In the teacher data we can thus compute the probability of a teacher
+having a specific strength on the IPC. This probability is computed using the
+'survival-function' defined as  \[\exp(-\alpha
+y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
+with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
+function we also see that the circular concentration parameter \(\kappa\) and
+the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
+the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
+and strength of interpersonal behavior through the parameters $\beta_{0}$ and
+$\gamma_0$ as well as the effect of self-efficacy on type and strength of
+teacher behavior through the parameters, $\beta_{1}$ and $\gamma_1$.\newline
+\indent We can use numerical optimization (Nelder-Mead) to find solutions for
 the maximum likelihood (ML) estimates for the parameters of the model.
 
-\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}
+\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
 
 This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
-  outcomes and \(w\) independent linear outcomes. The circular outcomes
-  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
+@mastrantonio2018joint to the regression context. The model contains \(m\)
+  independent circular outcomes and \(w\) independent linear outcomes. The
+  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
   \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
-  distribution.
-The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes.
-The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
-  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
-  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
-  so-called augmented representation (as in \eqref{projection} and
-  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
-  multivariate skew normal distribution. Following @mastrantonio2018joint the
-  linear outcomes are represented as
-\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
-\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear outcome
-\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
-is a \(w \times w\) diagonal matrix with diagonal elements
-\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
-\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
-$w$-dimensional half normal distribution [@olmos2012extension], and
-\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
-that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
-is normally distributed with mean \(\boldsymbol{\mu}_y +
-\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
-\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
-\boldsymbol{D}^t)^t\) is defined as:
-\begin{equation}\label{YDjoint}
-f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
-\end{equation}
-\noindent where
-\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
-the \(\ell\)-dimensional normal density with mean vector
-\(\boldsymbol{\mu}_\ell\) and
-covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
-modelling the augmented representations of \(\boldsymbol{\Theta}\) and
-\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
-distribution. The joint density of the model is then represented by:
-\begin{equation}\label{YDThetarjoint} 
-f(\boldsymbol{\theta}, \boldsymbol{r},
-\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
-\boldsymbol{y}^t)^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, ({\rm
-diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
-\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
-1}^{m}r_j, 
-\end{equation}
-\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
-r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{\mu} =
-(\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\) and \(\boldsymbol{\Sigma} =
-\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
-\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
-The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
-of and covariances between the augmented representations of the circular outcome
-and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
-augmented representations of the circular outcome and the linear outcome.
-\newline
-\indent In our regression extension we have \(i = 1, \dots, n\) observations of
-\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
-the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{\mu}_i =
+  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
+  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
+  distribution [@sahu2003new]. Thus also in this model the linear outcome is
+  defined constrained to be positive. Because the GPN distribution is modelled
+  using a so-called augmented representation (see also the description of the
+  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
+  modelling the multivariate skew normal distribution. As in
+@mastrantonio2018joint dependence between the linear and circular outcome is
+  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
+  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
+  distribution.\newline
+\indent This means that we have a shared mean vector and variance-covariance
+matrix for the linear and circular outcome(s), much like having multiple
+outcomes in a MANOVA (multivariate analysis of variance) model. In our
+regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
+observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
+covariates. The mean vector then becomes \(\boldsymbol{M}_i =
 \boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
-\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
-is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
-and the $g$ covariate values. \newline
-\indent For the teacher data, the regression equations for
-the circular and linear outcomes in the GPN-SSN model are
-\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +
-\boldsymbol{\beta}_1\text{SE}_i,\] where
-\(\hat{\boldsymbol{\mu}}_i = (\hat{\boldsymbol{\mu}}_{s_i}^t, \hat{\mu}_{y_i})^t\),
-\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
-and
-\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\). Note that because $m = 1$ and $w = 1$, $\boldsymbol{\mu}_{s_i}$ is a 2 dimensional vector and $\mu_{y_i}$ is a scalar.
-We estimate the model using MCMC methods. A detailed description of these
+\times (2m + w)\) matrix with regression coefficients and intercepts and
+$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
+estimate an intercept and the $g$ covariate values. This means that in contrast
+to the other three models, we have to use the same set of predictors for the
+circular and linear outcome.\newline
+\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
+\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
+\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
+The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
+linear outcomes in the GPN-SSN model are
+\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
+\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
+\beta_{1_{s^{I}}}\text{SE}_i),\]
+
+\noindent and
+
+\[\hat{y_i} = \beta_{0_{y_i}} +
+\beta_{1_{y_i}}\text{SE}_i.\]
+
+\noindent The GPN-SSN model thus allows us to assess the average type and
+strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
+$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
+on type and strength of teacher behavior through the parameters,
+$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
+because the type and strength of interpersonal behavior are modelled together
+using a multivariate normal distribution we can through its variance-covariance
+matrix also assess the dependence between the type and strength of interpersonal
+behavior. \newline
+\indent We estimate the model using MCMC methods. A detailed description of these
 methods is given in the Supplementary Material.
 
-\subsection{Model fit criterion}\label{Modelfit}
+\subsection{Model Fit Criterion}\label{Modelfit}
 
@@
 parameters for the \(j^{th}\) iteration. Because the joint density and thus also
-the likelihood for the modified GPN-SSN model in (\ref{YDThetarjoint}) is not
+the likelihood for the modified GPN-SSN model is not
 available in closed form [@mastrantonio2018joint] we compute the PLSL for the
@@
 the predictive accuracy.\newline
-\indent We use the loglikelihoods of the following conditional densities for the
-computation of the PLSL in the teacher data:
-
-\begin{itemize}
-\item For the modified CL-PN model:
-
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
-
-\item For the modified CL-GPN model:
-
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
-
-\item For the modified Abe-Ley model:
-
-$y_i \mid \theta_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.
-
-$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.
-
-\item For the modified joint projected and skew normal model:
-
-$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$
-
-$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$
-
-where $SSN$ is the skew normal distribution.
-\end{itemize}
-
-\noindent For each of the four cylindrical models and for each of the 10
+\indent For each of the four cylindrical models and for each of the 10
 cross-validation analyses we can then compute a PLSL for the circular and linear
-outcome by using the conditional log-likelihoods of the respective outcome. To
-evaluate the predictive performance we average across the PLSL criteria of the
+outcome by using the conditional log-likelihoods of the respective outcome (see
+Supplementary Material for a definition of the loglikelihoods). To evaluate the
+predictive performance we average across the PLSL criteria of the
 cross-validation analyses. We also assess the cross-validation variability by
@@
 In this section we analyze the teacher data with the help of the four
-cylindrical models from Section \ref{Models}. We will present the results,
-posterior estimates and their interpretation, per model and finish with a
-section comparing the fit of the different models.
+cylindrical models from the previous section. We will present the results,
+posterior estimates and their interpretation for all four models.
 
@@
 
+```{r geweke, results = FALSE, echo = FALSE, cache = TRUE}
+
+for(i in 1:10){
+  print(paste('fold', i))
+  print(geweke.diag(mcmc(cbind(res_CLPN[[i]]$Gamma,res_CLPN[[i]]$BI,
+                               res_CLPN[[i]]$BII,res_CLPN[[i]]$Sigma)), 0.25, 0.75))
+  print(geweke.diag(mcmc(cbind(res_CLGPN[[i]]$Gamma,res_CLGPN[[i]]$BI,
+                               res_CLGPN[[i]]$BII,res_CLGPN[[i]]$Sigma,res_CLGPN[[i]]$Sig)), 0.25, 0.75))
+  print(geweke.diag(mcmc(cbind(res_CLGPNM[[i]]$lambda,res_CLGPNM[[i]]$Betacon[1,1,],
+                               res_CLGPNM[[i]]$Betacon[1,2,], res_CLGPNM[[i]]$Betacon[1,3,],
+                               res_CLGPNM[[i]]$Betacon[2,1,], res_CLGPNM[[i]]$Betacon[2,2,],
+                               res_CLGPNM[[i]]$Betacon[2,3,], res_CLGPNM[[i]]$Sigmacon[1,1,],
+                               res_CLGPNM[[i]]$Sigmacon[1,2,], res_CLGPNM[[i]]$Sigmacon[1,3,],
+                               res_CLGPNM[[i]]$Sigmacon[2,1,], res_CLGPNM[[i]]$Sigmacon[2,2,],
+                               res_CLGPNM[[i]]$Sigmacon[2,3,], res_CLGPNM[[i]]$Sigmacon[3,1,],
+                               res_CLGPNM[[i]]$Sigmacon[3,2,], res_CLGPNM[[i]]$Sigmacon[3,3,])), 0.25, 0.75))
+}
+
+```
+
 \subsection{Results \& Analysis}\label{DataResults}
@@
 In the Supplementary Material we have described the starting values for the MCMC
-procedures, hence it remains to specify  the starting  values for the
-maximum likelihood based Abe-Ley model: \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 =
-0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\). The initial amount
-of iterations for the three MCMC samplers was set to 2000. After convergence
-checks via traceplots we concluded that some of the parameters of the GPN-SSN
-model did not converge. Therefore we set the amount of iterations of the MCMC
-models to 20,000 and subtracted a burn-in of 5000 to reach convergence. Note
-that we choose the same amount of iterations for all three Bayesian models to
-make their comparison via the PLSL as fair as possible. Lastly, the predictor
+procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it remains to specify
+the starting  values for the maximum likelihood based Abe-Ley model: \(\eta_0 =
+0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda
+= 0\). The initial number of iterations for the three MCMC samplers was set to
+2000. After convergence checks via traceplots we concluded that some of the
+parameters of the GPN-SSN model did not converge. Therefore we set the number of
+iterations of the MCMC models to 20,000 and subtracted a burn-in of 5000 to
+reach convergence (the geweke diagnostics show absolute z-scores over 1.96 in
+6\% of the estimated parameters). Note that we choose the same number of
+iterations for all three models models estimated using MCMC prodedures to make
+their comparison via the PLSL as fair as possible. Lastly, the predictor
 \verb|SE| was centered before inclusion in the analysis as this allows the
-intercepts to bear the classical meaning of average behavior.
-
-\subsubsection{The modified CL-PN and CL-GPN models} \label{resCL(G)PN}
+intercepts to bear the classical meaning of average behavior.\newline
+\indent Tables \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show
+the results for the four cylindrical models that were fit to the teacher data.
+For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN we show
+descriptives of the posterior of the estimated parameters (posterior mode and
+lower and upper bound of the 95\% highest posterior density (HPD) interval). For
+the Abe-Ley model we show the maximum likelihood estimates of the parameters. To
+compare the results of the four models we focus on the following aspects: the
+estimated average scores (intercept) on the type and strength of interpersonal
+behavior (1), the effect of self-efficacy on the type and strength of
+interpersonal behavior (2), the dependence between the type and strength of
+interpersonal behavior (3) and the model fit (4).
 
@@
 
-```{r means, cache = TRUE, echo = FALSE}
-
-theta_pred_PN   <- simplify2array(lapply(res_CLPN, "[[", "theta_pred"))
-theta_pred_GPN  <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred"))
-theta_pred_GPNM <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred"))
-
-theta_pred_PN <- array(unlist(theta_pred_PN),
-                       dim = c(nrow(theta_pred_PN[[1]]), ncol(theta_pred_PN[[1]]), length(theta_pred_PN)))
-theta_pred_GPN <- array(unlist(theta_pred_GPN),
-                        dim = c(nrow(theta_pred_GPN[[1]]), ncol(theta_pred_GPN[[1]]), length(theta_pred_GPN)))
-theta_pred_GPNM <- array(unlist(theta_pred_GPNM),
-                         dim = c(nrow(theta_pred_GPNM[[1]]), ncol(theta_pred_GPNM[[1]]), length(theta_pred_GPNM)))
-
-means <- round(rbind(mean_circ(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
-                     mean_circ(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
-                     mean_circ(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)))*(180/pi), 2)
-hpds  <- round(rbind(apply(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
-                     apply(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
-                     apply(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ))*(180/pi), 2)
-
-means = cbind(means, hpds)
-
-rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
-colnames(means) <- c('Mode', "HPD LB", "HPD UB")
-#kable(means, "latex", booktabs = T, escape = FALSE, digits = 2,
-#      caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and GPN-SSN models")%>%
-#kable_styling() %>%
-#add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", #notation="alphabet")
-
-```
-
 \begin{table}
@@
 \midrule
-$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 0.09 (0.09) & 2.42 (0.15) & 1.90 (0.10) & 3.05 (0.17)\\
-$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.08 (0.08) & 0.85 (0.12) & 0.45 (0.09) & 1.30 (0.15)\\
-$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 0.04 (0.04) & 1.47 (0.05) & 1.17 (0.04) & 1.79 (0.05)\\
-$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.04 (0.04) & 0.70 (0.08) & 0.46 (0.05) & 0.96 (0.07)\\
-$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.01 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
+$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 2.03 (0.09) & 2.43 (0.12) & 1.91 (0.10) & 3.05 (0.17)\\
+$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.90 (0.08) & 0.84 (0.11) & 0.45 (0.09) & 1.29 (0.15)\\
+$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 1.37 (0.04) & 1.47 (0.05) & 1.16 (0.04) & 1.78 (0.05)\\
+$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.79 (0.04) & 0.70 (0.06) & 0.47 (0.05) & 0.96 (0.08)\\
+$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.44 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
 \addlinespace
-$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.00 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.05 (0.00)\\
-$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.00 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
-$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.01) & -0.00 (0.00) & 0.06 (0.00)\\
-$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.00 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
-$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.02 (0.25) & 1.83 (0.14) & 5.02 (0.41)\\
+$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.06 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.04 (0.00)\\
+$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.02 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
+$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.00) & -0.00 (0.00) & 0.06 (0.00)\\
+$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
+$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.04 (0.29) & 1.85 (0.13) & 5.00 (0.41)\\
 \addlinespace
-$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.46 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
+$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.47 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
 $\sum_{2,2}$ & NA (NA) & NA (NA) & NA (NA) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
 \bottomrule
+\multicolumn{7}{l}{Note: $\beta_0^{I}$, $\beta_0^{II}$ and $\gamma_0$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{at the average self-efficacy. $\beta_1^{I}$, $\beta_1^{II}$ and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
+\multicolumn{7}{l}{type and strength of interpersonal behavior. $\gamma_{cos}$ and $\gamma_{sin}$ inform us about the dependence}\\
+\multicolumn{7}{l}{ between the type and strength of interpersonal behavior. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are are}\\
+\multicolumn{7}{l}{elements of the variance-covariance matrix of the type of interpersonal behavior in the}\\
+\multicolumn{7}{l}{CL-GPN model and $\sigma$ is the error standard deviation of the strength of interpersonal behavior.}\\
+
 \end{tabular}
 \end{table}
-
-\begin{table}
-\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean in the CL-PN, CL-GPN and GPN-SSN models}
-\centering
-\begin{tabular}[t]{lrrr}
-\toprule
-  & Mode & HPD LB & HPD UB\\
-  \midrule
-Cl-PN & 32.29 & 24.81 & 39.71\\
-CL-GPN & 33.70 & 26.72 & 41.15\\
-GPN-SSN & 35.30 & 28.31 & 43.10\\
-\bottomrule
-\multicolumn{4}{l}{\textsuperscript{a} Note that these means are based on}\\
-\multicolumn{4}{l}{the posterior predictive}\\
-\multicolumn{4}{l}{distribution for the intercepts}\\
-\multicolumn{4}{l}{following (Wang \& Gelfand, 2013).}\\
-\end{tabular}
-\end{table}
-
-First recall the regression equations predicting the linear outcome
-\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i\]
-\noindent and circular outcome
-\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
-  \mu_{i}^{I}  \\
-\mu_{i}^{II}
- \end{pmatrix}=\begin{pmatrix}
-  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
-  \beta_0^{II} + \beta_1^{II}\text{SE}_i
- \end{pmatrix}.\]
-\noindent  For both models, \(\hat{y}\) is the predicted strength of
-interpersonal behavior, \(\hat{\boldsymbol{\mu}}\) is the predicted mean vector
-of the type of interpersonal behavior and the \(\gamma\)'s and \(\beta\)'s are
-intercepts and regression coefficients.\newline \indent Table \ref{tab:estCLGPN}
-shows the results for the modified CL-PN and CL-GPN models fit to the teacher
-dataset. The estimates in this table are the averages and standard deviations of
-the 10 cross-validation estimates. We show both the estimated posterior mode and
-the 95\% highest posterior density (HPD) interval for each parameter. The
-posterior estimates for the \(\beta\)'s and \(\gamma\)'s are quite similar for
-both models. We also see this similarity when we look at the predicted circular
-and linear means. The predicted linear mean is equal to \(\gamma_0\) which is
-equal to 0.38 and 0.37 in the CL-PN and CL-GPN models, respectively. The
-predicted circular means can be computed from \(\beta_0^{I}\) and
-\(\beta_0^{II}\) using a double arctangent function \(\mbox{atan2}(\beta_0^{II},
-\beta_0^{I})\), see \eqref{PredVal}. If we do this at each iteration of the MCMC
-sampler we get the posterior distribution of the circular mean in both models.
-Table \ref{tab:means} shows the posterior mode of the estimated circular means,
-which are very similar (32.29\(^\circ\) and 33.54\(^\circ\)).\newline
-\indent Next we  investigate the effect of self-efficacy. In the CL-PN model we
-can transform the regression parameters on the two components \(I\) and \(II\)
-to one circular regression coefficient \(b_c\) using the methods described by
-@CremersMulderKlugkist2017. The dashed line in Figure \ref{regline} shows the
-predicted regression line from the CL-PN model. Here $b_c$ is the slope of this
-line at the inflection point (black square) and its posterior mode is estimated
-at 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation to the
-circular regression line representing the slope at a specific point. Therefore
-it is possible for the HPD interval to be wider than $2\pi$. In this case the
-interval is much wider and covers 0, indicating there is no evidence to reject
-the null-hypothesis of no effect.}. Even though the HPD of \(b_c\) includes 0,
-indicating there is no evidence to reject the null-hypotheis of no effect, we
-continue with its interpretation for eductional purposes. The interpretation of
-$b_c$ is that at the inflection point an increase of 1 unit in self-efficacy
-leads to an increase of \(1.67*(180/\pi) = 95.68^\circ\) in the type of
-interpersonal behavior..However, as we can see in Figure \ref{regline} the
-inflection point lies almost outside the range of the actual data. Instead of
-looking at the inflection point we might compute the slope of the regression
-line at the average self-efficacy. This parameter we call the slope at the mean
-(\(SAM\)) [@CremersMulderKlugkist2017] and it is estimated at 0.02 (0.01; 0.04)
-for our data. This means that at the average self-efficacy an increase of 1 unit
-only leads to an increase of \(0.02*(180/\pi) = 1.15^\circ\) in the type of
-interpersonal behavior. The HPD of the \(SAM\) does not include 0 which means
-that the effect at the average self-efficacy can be distinguished from
-0.\newline
-\indent In the CL-GPN model we cannot compute circular regression coefficients
-such as the \(b_c\) and \(SAM\) computed above  due to the fact that not only
-the mean vector $\boldsymbol{\mu}$ but also the covariance matrix
-$\boldsymbol{\Sigma}$ influences the predicted value on the circle. Instead, we
-will compute posterior predictive distributions for the predicted circular
-outcome of individuals scoring the minimum, maximum and median self-efficacy.
-The modes and 95\% HPD intervals of these posterior predictive distributions are
-\(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\),
-\(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\). Note
-that we display the modes and HPD intervals for the posterior predictive
-distributions on the interval $[0^\circ, 360^\circ)$. Note that  $44.49^\circ =
-404.49^\circ$ due to the periodicity of a circular variable. The posterior mode estimate of
-$215.74^\circ$ thus lies within its HPD interval $(147.36^\circ, \: 44.49^\circ)$.The
-HPD intervals of the three posterior predictive distributions overlap. Had they
-not overlapped we could have concluded that as the self-efficacy increases, the
-score of the teacher on the IPC moves counterclockwise. \newline
-\indent The effect of self-efficacy on the strength of interpersonal behavior is
-quantified by \(\gamma_1\) in both the CL-PN and CL-GPN models. This parameter
-can be interpreted in the same way as in a usual regression
-model. The HPD interval of  \(\gamma_1\) includes 0 which means
-that there is not enough evidence for an effect of self-efficacy on the strength
-of interpersonal behavior. Note that the lower bounds of the HPD interval of
-this parameter are  however very close to zero.\newline
-\indent  The relation between the circular and linear outcome, that is, between
-type of interpersonal behavior and strength of interpersonal behavior, is
-described by the parameters $\gamma_{\cos}$ and $\gamma_{\sin}$. The HPD
-interval of \(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
-models, meaning that the cosine component of the type of interpersonal behavior
-has an effect on the strength of interpersonal behavior. In the teacher data the
-sine and cosine components have a substantive meaning. In this case the
-Communion (cosine) component of the IPC positively effects the strength of a
-teachers' type of interpersonal behavior, in plain words: teachers exhibiting
-interpersonal behavior types with higher communion scores (e.g., 'helpful' and
-'understanding' in Figure 2) are stronger in their behavior.
-
-```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-a1 <- CLPN.BI.m[1]
-b1 <- CLPN.BI.m[2]
-a2 <- CLPN.BII.m[1]
-b2 <- CLPN.BII.m[2]
-ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-
-x <- seq(-5, 5, by = 0.1)
-
-pred.AL <- 0.37 + 2*atan(-0.01*x)
-pred.PN <- atan2(a2 + b2*x, a1 + b1*x)
-
-plot(Dat$SEc, Dat$theta*(180/pi), bty = "n", yaxt = "n",
-     xlim = c(-5, 2), ylab = "$\\theta$", xlab = "self-efficacy")
-axis(side = 2,  at = c(-160, -80, 0, 80, 160))
-points(x, pred.AL*(180/pi), type = "l", lty = 1)
-points(x, pred.PN*(180/pi), type = "l", lty = 2)
-points(ax, ac*(180/pi), pch = 15)
-
-dev.off()
-tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)
-
-```
-
-```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}
-
-BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
-BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
-
-bc <- matrix(NA, 15001, 10)
-SAM <- matrix(NA, 15001, 10)
-
-for(i in 1:10){
-
-  a1 <- BI[5000:20000,1,i]
-  b1 <- BI[5000:20000,2,i]
-  a2 <- BII[5000:20000,1,i]
-  b2 <- BII[5000:20000,2,i]
-
-  ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
-  ac <- atan2(a2 + b2*ax, a1 + b1*ax)
-  bc[,i] <- -tan(atan2(a2, a1)-ac)/ax
-  SAM[,i] <- bc[i]/(1 + (bc[i]*(-ax))^2)
-
-}
-
-mean(apply(bc, 2, mode_est))
-rowMeans(apply(bc, 2, hpd_est))
-
-mean(apply(SAM, 2, mode_est))
-rowMeans(apply(SAM, 2, hpd_est))
-
-```
-
-\begin{figure}
-\centering
-\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
-\caption{Plot showing circular regression lines for the effect of self-efficacy
-as predicted by the Abe-Ley model (solid line) and CL-PN model (dashed line). The
-black square indicates the inflection point of the circular regression line for
-the CL-PN model.}
-\label{regline}
-\end{figure}
-
-```{r ppreg, echo = FALSE, results = FALSE, eval = FALSE}
-theta_pred_GPN.min <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.min"))
-theta_pred_GPN.median <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.meadian"))
-theta_pred_GPN.max <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.max"))
-
-mean_circ(apply(theta_pred_GPN.min[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPN.median[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPN.max[5000:20000,], 2, mode_est_circ))*(180/pi)
-
-apply(apply(theta_pred_GPN.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPN.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPN.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-
-```
-
-\subsubsection{The modified Abe-Ley model} \label{resAL}
 
@@
 
-rownames(ALres.tab.new) <- c("$\\eta_0$", "$\\eta_1$",
-                             "$\\nu_0$", "$\\nu_1$",
+rownames(ALres.tab.new) <- c("$\\beta_0$", "$\\beta_1$",
+                             "$\\gamma_0$", "$\\gamma_1$",
                              "$\\alpha$", "$\\kappa$", "$\\lambda$")
@@
 
-\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation, for the modified Abe-Ley model}
+\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation (SD), for the modified Abe-Ley model}
 \centering
-\begin{tabular}[t]{ll}
+\begin{tabular}[t]{llllllll}
 \toprule
-\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
-\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
-$\eta_0$ & 0.36 (0.02)\\
-$\eta_1$ & -0.03 (0.01)\\
-$\nu_0$ & 1.17 (0.02)\\
-$\nu_1$ & 0.04 (0.02)\\
-$\alpha$ & 3.66 (0.12)\\
-$\kappa$ & 1.51 (0.08)\\
-$\lambda$ & 0.70 (0.05)\\
+& $\beta_0$ & $\beta_1$ & $\gamma_0$  & $\gamma_1$ & $\alpha$ & $\kappa$ & $\lambda$\\
+Mean & 0.36 & -0.03 & 1.17 & 0.04 & 3.66 & 1.51 & 0.70 \\
+SD & 0.02) & 0.01 & 0.02 & 0.02 & 0.12 & 0.08 & 0.05\\
 \bottomrule
+\multicolumn{8}{l}{Note: $\beta_0$  and $\gamma_0$ inform us about the type and strength }\\
+\multicolumn{8}{l}{of interpersonal behavior at the average self-efficacy. $\beta_1$ }\\
+\multicolumn{8}{l}{and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
+\multicolumn{8}{l}{type and strength of interpersonal behavior. $\alpha$ is the } \\
+\multicolumn{8}{l}{shape parameter of the distribution of the strength }\\
+\multicolumn{8}{l}{of interpersonal bahavior. $\kappa$ and $\lambda$ are the concentration}\\
+\multicolumn{8}{l}{and skewness parameters for the distribution of the type}\\
+\multicolumn{8}{l}{of interpersonal behavior.}\\
 \end{tabular}
 \end{table}
-
-First recall the regression equations predicting the linear outcome
-\[\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i)\]
-
-\noindent and circular outcome
-\[\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).\]
-
-\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type
-of interpersonal behavior and \(\hat{\beta}\) is the predicted scale
-parameter of the strength of interpersonal behavior, while the \(\eta\)'s and
-\(\nu\)'s are regression parameters.\newline
-\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley model
-fit to the teacher dataset. The estimates in this table are the averages and
-standard deviations of the 10 cross-validation estimates. The estimate of the
-circular mean at an average self-efficacy, \(\eta_0\), equals 0.36 radians or
-20.62\(^\circ\). For the Abe-Ley model we can also investigate the effect of
-self-efficacy on the type of interpersonal behavior. The solid line in
-Figure \ref{regline} shows this effect. Unlike for the regression line of the
-CL-PN model, the inflection point of the regression line of the Abe-Ley model
-lies outside the x-range of the figure. \newline
-\indent For the Abe-Ley model the conditional distribution for the linear
-outcome is Weibull. This means that we can use methods from survival analsis to
-interpret the effect of self-efficacy. In survival analysis a 'survival'
-function is used in which time is plotted against the probability of survival of
-subjects suffering from a specific medical condition. In our data however we
-plot the strength on the IPC against the probability of a teacher having such a
-strength. This probability is computed using the
-'survival-function' \(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i -
-\mu_i))^{1/\alpha}})\) with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
-\ref{reglineweib}, we plot the survival function for the minimum, median and
-maximum value of self-efficacy. We conclude that stronger interpersonal
-behaviors are less probable. We also see that the relation between self-efficacy and the strength on the IPC is not linear; the probability of having a stronger
-interpersonal behavior is higher for both the minimum and maximum self-efficacy compared to a median self-efficacy score.
-Note however that the circular outcome also influences the survival function.
-The relation between the type of interpersonal behavior and the strength of interpersonal behavior may thus influence the shape of the survival function.\newline
-
-```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-summary(Dat$SEc)
-
-datafig =  rbind(Dat[Dat$SEc == min(Dat$SEc),], Dat[Dat$SEc == median(Dat$SEc),], Dat[Dat$SEc == max(Dat$SEc),])
-
-x = unique(datafig$SEc)
-theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
-y <- seq(0, 1, by = 0.05)
-#h <- matrix(NA, length(y), length(x))
-s <- matrix(NA, length(y), length(x))
-scross <- array(NA, dim = c(length(y), length(x), 10))
-
-for(k in 1:10){
-
-  train <- get(paste("train.", k, sep = ""))
-
-  datafig =  rbind(train[train$SEc == min(train$SEc),],
-                   train[train$SEc == median(train$SEc),],
-                   train[train$SEc == max(train$SEc),])
-  x = unique(datafig$SEc)
-  theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
-
-    for(j in 1:length(x)){
-
-      for(i in 1:length(y)){
-          mu <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*x[j])
-          beta <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*x[j])*(1 - tanh(mean(resAL_arr[5,k]))*cos(theta[j] - mu))^(1/mean(resAL_arr[6,k]))
-    #h[i,j] <- beta*mean(resAL_arr[6,])*y[i]^(mean(resAL_arr[6,])-1)
-          s[i,j] <- exp(-mean(resAL_arr[6,k])*y[i]^beta)
-
-    }
-
-     scross[,,k] <- s
-
-  }
-
-}
-
-s = apply(scross, c(1,2), mean)
-
-tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
-points(y, s[,2], type = "l", lty = 2)
-points(y, s[,3], type = "l", lty = 3)
-legend(0.7, 1, legend = c("min(SE)", "median(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")
-
-dev.off()
-tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)
-
-```
-
-\begin{figure}
-\centering
-\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
-\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
-\label{reglineweib}
-\end{figure}
-
-```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}
-
-require(pracma)
-
-alpha <- mean(resAL_arr[6,])
-kappa <- mean(resAL_arr[5,])
-lambda <- mean(resAL_arr[7,])
-
-#R cannot compute legendre functions of non-integer degree
-#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
-#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
-#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]
-
-#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
-
-1/3.82
-cosh(1.58)
-
-```
-
-\subsubsection{The modified GPN-SSN model} \label{resGPNSSN}
 
@@
 
-```{r ppreggpnm, echo = FALSE, results = FALSE, eval = FALSE}
-theta_pred_GPNM.min <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.min"))
-theta_pred_GPNM.median <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.median"))
-theta_pred_GPNM.max <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.max"))
-
-mean_circ(apply(theta_pred_GPNM.min[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPNM.median[5000:20000,], 2, mode_est_circ))*(180/pi)
-mean_circ(apply(theta_pred_GPNM.max[5000:20000,], 2, mode_est_circ))*(180/pi)
-
-dim(theta_pred_GPNM.median)
-
-apply(apply(theta_pred_GPNM.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPNM.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-apply(apply(theta_pred_GPNM.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
-
-```
-
 \begin{table}
@@
 \midrule
-$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.10) & 1.75 (0.09) & 2.50 (0.11)\\
-$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.33 (0.07) & 1.10 (0.05) & 1.57 (0.06)\\
+$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.11) & 1.75 (0.09) & 2.50 (0.11)\\
+$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.34 (0.06) & 1.10 (0.05) & 1.57 (0.06)\\
 $\beta_{0_y}$ & 0.33 (0.01) & 0.30 (0.30) & 0.36 (0.01) & 0.33 (0.01) & 0.30 (0.01) & 0.36 (0.01)\\
+\addlinespace
 $\beta_{1_s^{I}}$ & 0.09 (0.01) & 0.05 (0.01) & 0.13 (0.01) & 0.60 (0.06) & 0.33 (0.05) & 0.90 (0.06)\\
-$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.03) & 0.66 (0.04)\\
+$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.04) & 0.66 (0.04)\\
+$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
 \addlinespace
-$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
-$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.43 (0.14) & 1.72 (0.07) & 3.46 (0.13)\\
+$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.44 (0.15) & 1.72 (0.07) & 3.46 (0.14)\\
 $\sum_{s_{2,2}}$ & 0.02 (0.00) & 0.02 (0.00) & 0.03 (0.00) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
 $\sum_{y_{3,3}}$ & 0.03 (0.00) & 0.02 (0.02) & 0.04 (0.00) & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00)\\
-$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.07 (0.06) & -0.20 (0.06) & 0.35 (0.06)\\
-\addlinespace
-$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.01) & 0.32 (0.01)\\
+$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.08 (0.06) & -0.20 (0.06) & 0.34 (0.06)\\
+$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.00) & 0.32 (0.01)\\
 $\sum_{sy_{2,3}}$ & 0.01 (0.00) & 0.01 (0.01) & 0.02 (0.00) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
@@
 \bottomrule
+\multicolumn{7}{l}{Note: $\beta_{0_s^{I}}$, $\beta_{0_s^{II}}$ and $\beta_{0_y}$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{at the average self-efficacy. $\beta_{1_s^{I}}$, $\beta_{1_s^{II}}$ and $\beta_{1_y}$ inform us about the effect of self-efficacy }\\
+\multicolumn{7}{l}{on the type and strength of interpersonal behavior. $\sum_{s_{1,1}}$, $\sum_{s_{1,2}}$, $\sum_{s_{2,2}}$, $\sum_{y_{3,3}}$, $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ }\\
+\multicolumn{7}{l}{are elements of the variance-covariance matrix of which $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ inform us about}\\
+\multicolumn{7}{l}{the dependence between the type and strength of interpersonal behavior.}\\
+\multicolumn{7}{l}{$\lambda$ is the skewness parameter of the distribution of the strengths of interpersonal behavior.}\\
 \end{tabular}
@@
 
-First recall the regression equation predicting the circular and linear
-outcome:
-
-\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +\boldsymbol{\beta}_1\text{SE}_i,\]
-
-\noindent where
-\(\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t\),
-\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
-and
-\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
-The parameters \(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\),
-\(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) are the intercepts and
-regression coefficients for the circular outcome and \(\beta_{0_y}\) and
-\(\beta_{1_y}\) are the intercept and regression coefficient for the
-linear outcome.\newline
-\indent Table \ref{tab:estCLGPNM} shows the results for the modified GPN-SSN model fit to
-the teacher dataset. The estimates in this table are the averages and
-standard deviations of the 10 cross-validation estimates. We show both
-the estimated posterior mode and the 95\% highest posterior density
-(HPD) interval for each parameter. The predicted circular means can be
-computed from \(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) in a
-similar fashion as for the CL-PN and CL-GPN models. Table \ref{tab:means} shows the
-posterior mode of the estimated circular mean, which equals 35.30\(^\circ\) and hence
-is very similar to those of the CL-PN and CL-GPN models.\newline
-\indent For the same reason as in the CL-GPN model we cannot compute circular
-regression coefficients for the effect of self-efficacy on the type of
-interpersonal behavior such as the \(b_c\) and \(SAM\). Instead, we will again
-compute posterior predictive distributions for the predicted circular outcome of
-individuals scoring the minimum, maximum and median self-efficacy. The modes
-and 95\% HPD intervals of these posterior predictive distributions are
-\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.11^\circ, \: 72.02^\circ)\),
-\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\). On a
-circle the HPD intervals of the three posterior predictive distributions
-overlap. Had they not overlapped we could have concluded that as the
-self-efficacy increases, the score of the teacher on the IPC moves
-counterclockwise. The effect of self-efficacy on the strength of interpersonal
-behavior is quantified by \(\beta_{1_y}\), and we learn that for a 1 unit
-increase in self-efficacy the strength of interpersonal behavior increases by
-0.09. The average strength is quantified by \(\beta_{0_y}\) and equals
-0.33.\newline
-\indent To investigate the association between the linear and circular outcome
-we look at the covariances between the linear outcome and the sine and cosine of
-the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). Both
-covariances, \(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are
-different from zero, but the one of the cosine component is larger. This means
-the correlation with the Communion component is larger and that teachers scoring
-both high on Communion and Agency show stronger behavior. This is a slightly
-different conclusion from the one in the CL-PN and CL-GPN models.
+\subsubsection{Average type and strength of interpersonal behavior} 
+The parameters $\gamma_0$ in the CL-PN, CL-GPN and Abe-Ley model and the
+parameter $\beta_{0_y}$ in the GPN-SSN model inform us about the strength of
+interpersonal behavior at the average self-efficacy. For the CL-PN, CL-GPN and
+GPN-SSN model the parameters are estimated at 0.38, 0.37 and 0.30 respectively
+and are a direct prediction of the strength of interpersonal behavior at the
+average self-efficacy. The estimate for the GPN-SSN model is notably lower and
+likely to be caused by its skewed distribution for the strengths of
+interpersonal behavior. In the Abe-Ley model, $\gamma_0$ influences the shape
+parameter of the distribution of the strength of interpersonal behavior and does
+not directly estimate the average strength. Instead we can use the survival
+function to say something about the probability of having a certain strength of
+interpersonal behavior. Figure \ref{reglineweib} shows this function for several
+values of self-efficacy. We look at the survival function at average values of
+self-efficacy. Note that this function is the average of all survival functions
+for observations that fall within 1 standard deviation of the mean. The survival
+function indicates that the probability of having a low strength of
+interpersonal behavior is higher than having a high strength. We however can not
+make any direct statement about the estimated strength using the Abe-Ley
+model.\newline
+\indent The parameters $\beta_0^{I}$, $\beta_0^{II}$, $\beta_0$,
+$\beta_{0_{s^{I}}}$ and $\beta_{0_{s^{II}}}$ inform us about the type of
+interpersonal behavior at the average self-efficacy for the CL-PN, CL-GPN,
+Abe-Ley and GPN-SSN model respectively. For the CL-PN, CL-GPN and GPN-SSN model
+we need to combine the estimates for the underlying bivariate components $\{I,
+II\}$ into one circular estimate using the double arctangent
+function\footnote{\(\mbox{atan2}(\beta_0^{II}, \beta_0^{I})\) or
+\(\mbox{atan2}(\beta_{0_{s^{II}}}, \beta_{0_{s^{I}}})\)}.
+Table \ref{tab:means} shows that these circular estimates are similar for the
+three models at 32.29$^\circ$, 33.70$^\circ$ and 35.53$^\circ$. In the Abe-Ley
+model the type of interpersonal behavior at the average self-efficacy is
+estimated at 0.36 radians or 20.63$^\circ$.
+
+```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+summary(Dat$SEc)
+
+y <- seq(0, 1, by = 0.05)
+scross <- array(NA, dim = c(3, length(y), 10))
+
+SE_low = Dat[Dat$Efficacy_CM <= 4.04,]
+SE_mid = Dat[Dat$Efficacy_CM > 4.04 & Dat$Efficacy_CM < 6.04,]
+SE_high = Dat[Dat$Efficacy_CM >= 6.04,]
+
+for(i in 1:length(y)){
+  mu_low <- mean(resAL_arr[1,]) + 2*atan(mean(resAL_arr[2,])*SE_low$SEc)
+  beta_low <- exp(mean(resAL_arr[3,]) + mean(resAL_arr[4,])*SE_low$SEc)*(1 -     tanh(mean(resAL_arr[5,]))*cos(SE_low$theta - mu_low))^(1/mean(resAL_arr[6,]))
+
+  mu_mid <- mean(resAL_arr[1,]) + 2*atan(mean(resAL_arr[2,])*SE_mid$SEc)
+  beta_mid <- exp(mean(resAL_arr[3,]) + mean(resAL_arr[4,])*SE_mid$SEc)*(1 -     tanh(mean(resAL_arr[5,]))*cos(SE_mid$theta - mu_mid))^(1/mean(resAL_arr[6,]))
+
+  mu_high <- mean(resAL_arr[1,]) + 2*atan(mean(resAL_arr[2,])*SE_high$SEc)
+  beta_high <- exp(mean(resAL_arr[3,]) + mean(resAL_arr[4,])*SE_high$SEc)*(1 -     tanh(mean(resAL_arr[5,]))*cos(SE_high$theta - mu_high))^(1/mean(resAL_arr[6,]))
+
+  if(i == 1){
+    res_low = exp(-mean(resAL_arr[6,])*y[i]^beta_low)
+    res_mid = exp(-mean(resAL_arr[6,])*y[i]^beta_mid)
+    res_high = exp(-mean(resAL_arr[6,])*y[i]^beta_high)
+  }else{
+    res_low = cbind(res_low, exp(-mean(resAL_arr[6,])*y[i]^beta_low))
+    res_mid = cbind(res_mid, exp(-mean(resAL_arr[6,])*y[i]^beta_mid))
+    res_high = cbind(res_high, exp(-mean(resAL_arr[6,])*y[i]^beta_high))
+  }
+
+}
+
+for(k in 1:10){
+
+  train <- get(paste("train.", k, sep = ""))
+
+  SE_low = train[train$Efficacy_CM <= 4.04,]
+  SE_mid = train[train$Efficacy_CM > 4.04 & train$Efficacy_CM < 6.04,]
+  SE_high = train[train$Efficacy_CM >= 6.04,]
+
+  for(i in 1:length(y)){
+    mu_low <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*SE_low$SEc)
+    beta_low <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*SE_low$SEc)*(1 -     tanh(mean(resAL_arr[5,k]))*cos(SE_low$theta - mu_low))^(1/mean(resAL_arr[6,k]))
+
+    mu_mid <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*SE_mid$SEc)
+    beta_mid <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*SE_mid$SEc)*(1 -     tanh(mean(resAL_arr[5,k]))*cos(SE_mid$theta - mu_mid))^(1/mean(resAL_arr[6,k]))
+
+    mu_high <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*SE_high$SEc)
+    beta_high <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*SE_high$SEc)*(1 -     tanh(mean(resAL_arr[5,k]))*cos(SE_high$theta - mu_high))^(1/mean(resAL_arr[6,k]))
+
+    if(i == 1){
+      res_low = exp(-mean(resAL_arr[6,k])*y[i]^beta_low)
+      res_mid = exp(-mean(resAL_arr[6,k])*y[i]^beta_mid)
+      res_high = exp(-mean(resAL_arr[6,k])*y[i]^beta_high)
+    }else{
+      res_low = cbind(res_low, exp(-mean(resAL_arr[6,k])*y[i]^beta_low))
+      res_mid = cbind(res_mid, exp(-mean(resAL_arr[6,k])*y[i]^beta_mid))
+      res_high = cbind(res_high, exp(-mean(resAL_arr[6,k])*y[i]^beta_high))
+    }
+
+  }
+
+  scross[1,,k] <- apply(res_low, 2, mean)
+  scross[2,,k] <- apply(res_mid, 2, mean)
+  scross[3,,k] <- apply(res_high, 2, mean)
+}
+
+s = apply(scross, c(1,2), mean)
+
+tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+plot(y, s[1,], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
+points(y, s[2,], type = "l", lty = 2)
+points(y, s[3,], type = "l", lty = 3)
+legend(0.7, 1, legend = c("low SE", "average SE", "high SE"), lty = c(1,2,3), bty = "n")
+
+dev.off()
+tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
+\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
+\label{reglineweib}
+\end{figure}
+
+```{r means, cache = TRUE, echo = FALSE}
+
+theta_pred_PN   <- simplify2array(lapply(res_CLPN, "[[", "theta_pred"))
+theta_pred_GPN  <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred"))
+theta_pred_GPNM <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred"))
+
+theta_pred_PN <- array(unlist(theta_pred_PN),
+                       dim = c(nrow(theta_pred_PN[[1]]), ncol(theta_pred_PN[[1]]), length(theta_pred_PN)))
+theta_pred_GPN <- array(unlist(theta_pred_GPN),
+                        dim = c(nrow(theta_pred_GPN[[1]]), ncol(theta_pred_GPN[[1]]), length(theta_pred_GPN)))
+theta_pred_GPNM <- array(unlist(theta_pred_GPNM),
+                         dim = c(nrow(theta_pred_GPNM[[1]]), ncol(theta_pred_GPNM[[1]]), length(theta_pred_GPNM)))
+
+means <- round(rbind(mean_circ(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
+                     mean_circ(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
+                     mean_circ(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)))*(180/pi), 2)
+hpds  <- round(rbind(apply(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
+                     apply(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
+                     apply(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ))*(180/pi), 2)
+
+means = cbind(means, hpds)
+
+#rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
+#colnames(means) <- c('Mode', "HPD LB", "HPD UB")
+#kable(means, "latex", booktabs = T, escape = FALSE, digits = 2,
+#      caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and GPN-SSN models")%>%
+#kable_styling() #%>%
+#add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", #notation="alphabet")
+
+```
+
+\begin{table}
+\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean (at SE = 0) in the CL-PN, CL-GPN and GPN-SSN models}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+  & Mode & HPD LB & HPD UB\\
+  \midrule
+CL-PN & 32.29 & 24.81 & 39.71\\
+CL-GPN & 33.70 & 26.72 & 41.15\\
+GPN-SSN & 35.53 & 28.40 & 43.30\\
+\bottomrule
+\multicolumn{4}{l}{Note that these means are based on}\\
+\multicolumn{4}{l}{their posterior predictive distribution }\\
+\multicolumn{4}{l}{following (Wang and Gelfand, 2013)}\\
+\end{tabular}
+\end{table}
+
+\subsubsection{The effect of self-efficacy}
+
+The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley model and $\beta_{1_y}$
+in the GPN-SSN model inform us about the effect of self-efficacy on the strength
+of interpersonal behavior. For the CL-PN, CL-GPN and GPN-SSN model the
+parameters are estimated at 0.03, 0.03 and 0.09 respectively and are a direct
+estimate of the effect of self-efficacy on the strength of interpersonal
+behavior, \emph{i.e.} an increase of 1 unit in self-efficacy leads to an
+increase of 0.09 units in the strength of interpersonal behavior according to
+the GPN-SSN model. These estimates are however quite small and only different
+from 0 (the HPD interval does not contain 0) in the GPN-SSN model. It is hard to
+say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to base our
+conclusions on. The model CL-GPN and CL-PN fit the linear outcome best according
+to the model fit in Table \ref{tab:ModelFit}. In these models the linear outcome has a
+symmetric distribution whereas in the GPN-SSN the distribution of the linear
+outcome is skewed. However, the effect of self-efficacy is different from 0 only
+in the GPN-SSN model which does not seem to match with its lower model fit.\newline
+\indent In the Abe-Ley model, $\gamma_1$ influences the shape parameter of the distribution of
+the strength of interpersonal behavior and does not directly estimate the effect
+of self-efficacy. Instead we can use the survival function to say something
+about the probability of having a certain strength of interpersonal behavior for
+different values of self-efficacy. Figure \ref{reglineweib} shows this function
+for low, average and high values of self-efficacy (as defined in Figure
+\ref{dataplot}). This function indicates that the effect of self-efficacy on the
+strength of interpersonal behavior is not linear. The probability of having a
+higher strength of interpersonal behavior is highest for low self-efficacy and
+lowest for average self-efficacy.\newline
+\indent The parameters $\beta_1^{I}$, $\beta_1^{II}$, $\beta_1$,
+$\beta_{1_{s^{I}}}$ and $\beta_{1_{s^{II}}}$ inform us about the effect of
+self-efficacy on the type of interpersonal behavior in the CL-PN, CL-GPN,
+Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley model we have
+drawn the circular regression lines for this effect in Figure \ref{regline} (see
+the description of the CL-PN and CL-GPN models). For the CL-PN model the
+inflection point is indicated with a square in Figure \ref{regline}. The
+inflection point for the Abe-Ley model falls outside the bounds of the plot and
+is therefore not displayed. The slope at the inflection point, $b_c$, for the
+CL-PN model is computed by using methods from @CremersMulderKlugkist2017 and is
+equal to 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation
+to the circular regression line representing the slope at a specific point.
+Therefore it is possible for the HPD interval to be wider than $2\pi$. In this
+case the interval is much wider and covers 0, indicating there is no evidence
+for an effect.} The parameter $\beta_1$ is the slope at the inflection point for
+the Abe-Ley model and is equal to -0.03. Even though these slopes are quite
+different, the regression lines in Figure \ref{regline} are quite similar in the
+data range. Both the regression line of the Abe-Ley model and the CL-PN model
+show slopes that are not very steep in the range of the data indicating that the
+effect of self-efficacy on the type of interpersonal behavior is not large.
+\newline
+\indent In the CL-GPN and GPN-SSN model we cannot compute circular regression
+coefficients due to the fact that not only the mean vector of the GPN
+distribution but also the covariance matrix influences the predicted value on
+the circle. Instead, we will compute posterior predictive distributions for the
+predicted circular outcome of individuals scoring the minimum, maximum and
+median self-efficacy. The modes and 95\% HPD intervals of these posterior
+predictive distributions are \(\hat{\theta}_{SE_{min}} = 215.74^\circ
+(147.36^\circ, \: 44.49^\circ)\), \(\hat{\theta}_{SE_{median}} = 25.93^\circ
+(337.02^\circ, \: 138.59^\circ)\), \(\hat{\theta}_{SE_{max}} = 30.86^\circ
+(8.63^\circ, \: 72.19^\circ)\) for the CL-GPN model. Note that we display the
+modes and HPD intervals for the posterior predictive distributions on the
+interval $[0^\circ, 360^\circ)$ and that $44.49^\circ = 404.49^\circ$ due to the
+periodicity of a circular variable. The posterior mode estimate of
+$215.74^\circ$ thus lies within its HPD interval $(147.36^\circ, \:
+44.49^\circ)$. For the GPN-SSN model the modes and 95\% HPD intervals of the
+posterior predictive distributiona are \(\hat{\theta}_{SE_{min}} = 206.87^\circ
+(117.12^\circ, \: 72.02^\circ)\), \(\hat{\theta}_{SE_{median}} = 24.68^\circ
+(334.73^\circ, \: 128.27^\circ)\), \(\hat{\theta}_{SE_{max}} = 29.81^\circ
+(0.74^\circ, \: 80.61^\circ)\). For both the CL-GPN and GPN-SSN model the HPD
+intervals of the mode of the posterior predictive intervals of individuals
+scoring the minimum, median and maximum self-efficacy overlap. This indicates
+that the effect of self-efficacy, if there is any, on the type of interpersonal
+behavior a teacher shows is not expected to be strong. Had the HPD intervals not
+overlapped we could have concluded that as the self-efficacy increases, the
+score of the teacher on the IPC moves counterclockwise.\newline
+
+```{r ppreg, echo = FALSE, results = FALSE, eval = FALSE}
+theta_pred_GPN.min <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.min"))
+theta_pred_GPN.median <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.meadian"))
+theta_pred_GPN.max <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.max"))
+
+mean_circ(apply(theta_pred_GPN.min[5000:20000,], 2, mode_est_circ))*(180/pi)
+mean_circ(apply(theta_pred_GPN.median[5000:20000,], 2, mode_est_circ))*(180/pi)
+mean_circ(apply(theta_pred_GPN.max[5000:20000,], 2, mode_est_circ))*(180/pi)
+
+apply(apply(theta_pred_GPN.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+apply(apply(theta_pred_GPN.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+apply(apply(theta_pred_GPN.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+
+```
+
+```{r ppreggpnm, echo = FALSE, results = FALSE, eval = FALSE}
+theta_pred_GPNM.min <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.min"))
+theta_pred_GPNM.median <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.median"))
+theta_pred_GPNM.max <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.max"))
+
+mean_circ(apply(theta_pred_GPNM.min[5000:20000,], 2, mode_est_circ))*(180/pi)
+mean_circ(apply(theta_pred_GPNM.median[5000:20000,], 2, mode_est_circ))*(180/pi)
+mean_circ(apply(theta_pred_GPNM.max[5000:20000,], 2, mode_est_circ))*(180/pi)
+
+dim(theta_pred_GPNM.median)
+
+apply(apply(theta_pred_GPNM.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+apply(apply(theta_pred_GPNM.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+apply(apply(theta_pred_GPNM.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
+
+```
+
+```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+
+tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+a1 <- CLPN.BI.m[1]
+b1 <- CLPN.BI.m[2]
+a2 <- CLPN.BII.m[1]
+b2 <- CLPN.BII.m[2]
+ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
+ac <- atan2(a2 + b2*ax, a1 + b1*ax)
+
+x <- seq(-5, 5, by = 0.1)
+
+pred.AL <- 0.37 + 2*atan(-0.01*x)
+pred.PN <- atan2(a2 + b2*x, a1 + b1*x)
+
+plot(Dat$SEc, Dat$theta*(180/pi), bty = "n", yaxt = "n",
+     xlim = c(-5, 2), ylab = "$\\theta$", xlab = "self-efficacy")
+axis(side = 2,  at = c(-160, -80, 0, 80, 160))
+points(x, pred.AL*(180/pi), type = "l", lty = 1)
+points(x, pred.PN*(180/pi), type = "l", lty = 2)
+points(ax, ac*(180/pi), pch = 15)
+
+dev.off()
+tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)
+
+```
+
+```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}
+
+BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
+BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
+
+bc <- matrix(NA, 15001, 10)
+SAM <- matrix(NA, 15001, 10)
+
+for(i in 1:10){
+
+  a1 <- BI[5000:20000,1,i]
+  b1 <- BI[5000:20000,2,i]
+  a2 <- BII[5000:20000,1,i]
+  b2 <- BII[5000:20000,2,i]
+
+  ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
+  ac <- atan2(a2 + b2*ax, a1 + b1*ax)
+  bc[,i] <- -tan(atan2(a2, a1)-ac)/ax
+  SAM[,i] <- bc[i]/(1 + (bc[i]*(-ax))^2)
+
+}
+
+mean(apply(bc, 2, mode_est))
+rowMeans(apply(bc, 2, hpd_est))
+
+mean(apply(SAM, 2, mode_est))
+rowMeans(apply(SAM, 2, hpd_est))
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
+\caption{Plot showing circular regression lines for the effect of self-efficacy
+as predicted by the Abe-Ley model (solid line) and CL-PN model (dashed line). The
+black square indicates the inflection point of the circular regression line for
+the CL-PN model.}
+\label{regline}
+\end{figure}
+
+\subsubsection{Dependence between type and strength of interpersonal behavior.}
+
+The relation between the type and strength of interpersonal behavior in the
+CL-PN and CL-GPN model, is described by the parameters $\gamma_{\cos}$ and
+$\gamma_{\sin}$. The HPD interval of \(\gamma_{\cos}\) does not include 0 for
+both the CL-PN and CL-GPN models, meaning that the cosine component of the type
+of interpersonal behavior has an effect on the strength of interpersonal
+behavior.\newline
+\indent In the teacher data the sine and cosine components have a substantive
+meaning. This is illustrated in Figure \ref{QTI}. In a unit circle the
+horizontal axis (Communion) represents the cosine of and the vertical axis
+(Agency) represents the sine of an angle. For the teacher data this means that
+the Communion (cosine) dimension of the IPC positively effects the strength of a
+teachers' type of interpersonal behavior, in plain words: teachers exhibiting
+interpersonal behavior types with higher communion scores (e.g., 'helpful' and
+'understanding' in Figure 2) are stronger in their interpersonal behavior.\newline 
+\indent In the GPN-SSN model the dependence between the type and strengths of
+interpersonal behavior is modelled through the covariances between the linear
+outcome and the sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
+\(\sum_{sy_{1,3}}\). Both covariances, \(\sum_{sy_{2,3}} = 0.09\) and
+\(\sum_{sy_{1,3}} = 0.23\), are different from zero, but the one of the cosine
+component, and thus the correlation with the Communion dimension, is larger.
+This mean that teachers scoring both high on Communion and Agency show stronger
+behavior. Together with the results from the CL-PN and CL-GPN models in the
+previous paragraph this translates to the conclusion that teachers with the
+strongest interpersonal behavior have a type of interpersonal behavior between
+0$^\circ$ and 90$^\circ$. To get these scores on the circle both the Agency and
+the Communion score of a Teacher have to be positive (see \eqref{PredVal}). This
+corresponds to the pattern observed in the teacher data in Figure \ref{QTI}. At
+a strength of 0.4 and up we see that the scores on the circle range on average
+between 0$^\circ$ and 100$^\circ$.
+
+```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}
+
+require(pracma)
+
+alpha <- mean(resAL_arr[6,])
+kappa <- mean(resAL_arr[5,])
+lambda <- mean(resAL_arr[7,])
+
+#R cannot compute legendre functions of non-integer degree
+#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
+#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
+#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]
+
+#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
+
+1/3.82
+cosh(1.58)
+
+```
 
@@
 
-In this section we will assess the overall fit of the cylindrical models to the
-data via the PLSL criterion described in Section \ref{Modelfit}. Table
-\ref{tab:ModelFit} shows the values of this criterion for the linear and
-circular outcomes of the four models.\newline
-\indent The CL-PN and CL-GPN models have the best out-of-sample predictive
-performance for the linear outcome. They show roughly the same performance because
-they model the linear outcome in the same way. Only the value of \(r\) in
-\eqref{circlinlink} differs. We should note that even though the predictive
-performance of the Abe-Ley model for the linear outcome is worst on average, the
-standard deviation of the cross-validation estimates is rather large. This means
-that in some samples, the Abe-Ley model shows a lower PLSL value than the
-average of 25.49\newline
+Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the linear
+and circular outcomes of the four cylindrical models fit to the teacher
+data.\newline \indent The CL-PN and CL-GPN models have the best out-of-sample
+predictive performance for the linear outcome. They show roughly the same
+performance because they model the linear outcome in the same way. We should
+note that even though the predictive performance of the Abe-Ley model for the
+linear outcome is worst on average, the standard deviation of the
+cross-validation estimates is rather large. This means that in some samples, the
+Abe-Ley model shows a lower PLSL value than the average of 25.49\newline
 \indent The Abe-Ley model has the best out-of-sample predictive performance for
@@
 relatively small sample size, but this does not explain why the PLSL has a large
-standard deviation for only a few cylindrical models and not for all.
+standard deviation for only a few cylindrical models and not for all.\newline
+\indent In this situation, where one model fits the linear outcome best and
+another one fits the circular outcome best, it is hard to determine which model
+we should choose. In this case the results for the CL-PN /CL-GPN and Abe-Ley
+model are quite different regarding the effect of self-efficacy on the linear
+outcome (strength of interpersonal behavior). Because the Abe-Ley fit for the
+linear part is worst we would choose to trust the results for the CL-PN and
+CL-GPN model here. For the circular part however the results of the CL-PN/CL-GPN
+model do not differ as much from the Abe-Ley model and we reach the same
+conclusion for both models, namely that the effect of self-efficacy on type of
+interpersonal behavior is not very strong. Therefore we would prefer the
+CL-PN/CL-GPN models in this case because where it matters in terms of
+interpretation (the linear part) they show better fit.
 
@@
 #kable(PLSL, "latex", booktabs = T, escape = FALSE,
-#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four #cylindrical models")%>%
+#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four cylindrical models")%>%
 #add_header_above(c("Model" = 1, "Circular" = 2, "Linear" = 2))
@@
 CL-PN & 82.96 & (9.47) & -17.65 & (3.70)\\
-CL-GPN & 85.22 & (18.12) & -18.12 & (3.57)\\
+CL-GPN & 86.05 & (16.63) & -18.30 & (3.00)\\
 Abe-Ley & 31.97 & (22.07) & 25.49 & (17.46)\\
-GPN-SSN & 106.38 & (8.84) & -2.14 & (6.78)\\
+GPN-SSN & 107.10 & (10.52) & -2.37 & (7.01)\\
 \bottomrule
@@
 In this paper we modified four models for cylindrical data in such a way that
-they include a regression of both the linear and circular outcome onto a
-set of covariates. Subsequently we have shown how these four methods can
-be used to analyze a dataset on the interpersonal behavior of teachers.
-In this final section we will comment on the differences between these models, the
-results from the analysis of the teacher data and how cylindrical models
-improve the analysis of such cylindrical data.\newline
+they include a regression of both the linear and circular outcome onto a set of
+covariates. Subsequently we have shown how these four methods can be used to
+analyze a dataset on the interpersonal behavior of teachers. In this final
+section we will first comment on what researchers can gain by using cylindrical
+models for the teacher data. Subsequently we will comment on the differences
+between the cylindrical models that were introduced in this paper.\newline
+\indent Concerning the teacher data, the advantage of cylindrical data analysis
+is that we were able to analyze the information about the type and strength of
+interpersonal behavior simultaneously. In previous research, the two components
+of the interpersonal circumplex (\emph{i.e.}, Agency and Communion) were
+analyzed separately. Such an approach also provides information about the
+strength of teachers' score on Agency and Communion, yet a large portion of
+information about the combination of Agency and Communion, which describes the
+type of behavior that is observed, gets lost. A first solution to include both
+dimensions as a circular variable in data analysis was described by
+@Cremers2018Assessing. A downside of that analysis was that information about
+the strength of the specific type of interpersonal behavior could not be
+retained. In the present study, we have shown how using cylindrical models can
+simultaneously model the information about the type of and strength of
+interpersonal behavior and how these are influenced by teachers' self-efficacy
+in classroom management. Although we do not find any strong effects of
+self-efficacy on either the type or strength of behavior, the four cylindrical
+models do provide a way of analyzing and interpreting this effect. This is
+beneficial for future research in which we may want to investigate the effect of
+further covariates on data from the circumplex. Furthermore, in addition to
+being able to assess the influence of covariates, the cylindrical models also
+provide information about the dependence between the type and strength of
+interpersonal behavior. We found that stronger behavior is associated with those
+types of behavior that show higher scores on the Communion dimension (and also
+on the Agency dimension). This translates to stronger behavior being shown by
+teachers whose type of interpersonal behavior ranges between 0$^\circ$ and
+90$^\circ$, the 'helpful' and 'directing' subtypes according to Figure
+\ref{QTI}. \textcolor{red}{Is dit nieuw/zijn er papers die hier iets over
+zeggen?}\newline
+\indent As mentioned in the introducion data from the interpersonal circumplex
+is not the only type of cylindrical data that occurs in psychology. The methods
+presented in this paper are also of use for research on human navigation and
+eye-tracking research. Furthermore, even though cylindrical models are already
+used in fields outside of psychology, the addition of a regression structure to
+the models is of use in these fields as well. \newline
 \indent In terms of interpretability, the CL-PN and Abe-Ley models perform
-best.  In the
+best out of the four cylindrical models.  In the
 CL-GPN and GPN-SSN models the interpretation of the parameters of the
@@
 model permits an asymmetric  circular outcome.\newline
-\indent To investigate model fit for the teacher data, we assessed
-out-of-sample predictive performance for both the linear and circular
-outcome. The CL-PN and CL-GPN models have the best fit for the linear
-outcome while the Abe-Ley model fits best  the circular outcome.
-Differences in fit, in addition to being a result of different
-distributional assumptions, may also be caused by the way in which the
-relation between the linear and circular outcome is modeled. Whereas in
-both the Abe-Ley and GPN-SSN models the distribution of the linear
-outcome is conditioned on the circular outcome and vice versa, the
-distribution of the circular outcome in the CL-PN and CL-GPN models is
-independent of the linear outcome. In these models the circular outcome
-is regressed onto the linear outcome.\newline
 \indent The four cylindrical models that were modified to the regression
@@
 developed in the present study.\newline
-\indent We conclude the paper on a general note regarding cylindrical models.
-They  offer new insights into data of a cylindrical nature in psychology.
-Concerning the example used in this paper, the advantage of cylindrical data
-analysis is that we were able to analyze all circular and linear information in
-the data simultaneously. In previous research, the two components of the
-interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
-separately. Such an approach also provides information about the strength of
-teachers' score on Agency and Communion, yet a large portion of information
-about the combination of Agency and Communion, which describes the kind of
-behavior that is observed,  gets lost. A first solution to include both
-dimensions as a circular variable in data analysis was described by
-@Cremers2018Assessing. A downside of that analysis was that information about
-the strength of the specific type of interpersonal behavior could not be
-retained. In the present study, we have shown how using cylindrical models can
-simultaneously model the information about the type of and strength of
-interpersonal behavior and how these are influenced by teachers' self-efficacy
-in classroom management. The results of the present study therefore provide an
-answer to an often stated problem in data analysis of interpersonal circumplex
-data.
 

### 2019-01-28 Manuscript/ManuscriptMarkdownCV.tex
Update manuscript

@@
-\documentclass[12pt,]{article}
+\documentclass[man]{apa6}
 \usepackage{lmodern}
@@
 }{}
-\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
             pdftitle={Regression models for Cylindrical data in Psychology},
-            pdfkeywords={bla vla},
+            pdfauthor={Jolien Cremers, Helena J.M. Pennings, \& Christophe Ley},
+            pdfkeywords={cylindrical data, regression, interpersonal behavior},
             pdfborder={0 0 0},
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{0}
 % Redefines (sub)paragraphs to behave more like sections
@@
 
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
   \title{Regression models for Cylindrical data in Psychology}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{}
-    \preauthor{}\postauthor{}
+    \author{Jolien Cremers\textsuperscript{1}, Helena J.M.
+Pennings\textsuperscript{2,3}, \& Christophe Ley\textsuperscript{4}}
     \date{}
-    \predate{}\postdate{}
-
-\usepackage{booktabs}
+
+\shorttitle{Cylindrical data in Psychology}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} Department of Methodology and Statistics, Utrecht University\\\textsuperscript{2} TNO\\\textsuperscript{3} Department of Education, Utrecht University\\\textsuperscript{4} Department of Applied Mathematics, Computer Science and Statistics, Ghent University}
+\keywords{cylindrical data, regression, interpersonal behavior\newline\indent Word count: X}
+\usepackage{csquotes}
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
+
 \usepackage{longtable}
-\usepackage{array}
+\usepackage{lscape}
 \usepackage{multirow}
-\usepackage[table]{xcolor}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
+\usepackage{tabularx}
+\usepackage[flushleft]{threeparttable}
 \usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-
+
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
+
+\usepackage{lineno}
+
+\linenumbers
+\DeclareRobustCommand{\VANDER}[3]{#2}
 \usepackage{multirow}
-\usepackage{appendix}
+\usepackage{rotating}
 \usepackage{color}
-\usepackage{hyperref}
 \usepackage{subcaption}
-\setlength\parindent{24pt}
-\usepackage[nodisplayskipstretch]{setspace}\doublespacing
-\DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
-
-\begin{document}
-\maketitle
-\begin{abstract}
+
+\authornote{Add complete departmental affiliations for each
+author here. Each new line herein must be indented, like this line.
+
+Enter author note here.
+
+Correspondence concerning this article should be addressed to Jolien
+Cremers, Postal address. E-mail:
+\href{mailto:joliencremers@gmail.com}{\nolinkurl{joliencremers@gmail.com}}}
+
+\abstract{
 Cylindrical data are multivariate data which consist of a directional,
@@
 interpersonal behavior of teachers.
-\end{abstract}
+
+}
+
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
+\begin{document}
+\maketitle
 
@@
 
-Cylindrical data are data that consist of a linear variable and a
-directional variable. In this paper, the directional variable is
-circular, meaning that it consists of a single angle instead of a set of
-angles. A circular variable is different from a linear variable in the
-sense that it is measured on a different scale. Figure \ref{circline}
-shows the difference between a circular scale (right) and a linear scale
-(left). The most important difference is that on a circular scale the
-datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in fact
-represent the same number while on a linear scale the two ends,
-\(-\infty\) and \(\infty\), are not connected and consequently the
-values 0\(^\circ\) and 360\(^\circ\) are located on different places on
-the scale. This difference requires us to use special statistical
-methods for circular variables (see e.g. Fisher (1995) for an
-introduction to circular data and Mardia \& Jupp (2000), Jammalamadaka
-\& Sengupta (2001) and Ley \& Verdebout (2017) for a more elaborate
-overview). For instance, a notion as simple as the sample average needs
-to be re-defined. As is the case for circular data, the analysis of
-cylindrical data also requires special methods.\newline
-\indent Cylindrical data occur in several fields of research, such as
-for instance in meteorology (García-Portugués, Crujeiras, \&
-González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
-González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
-Maruotti, \& Cosoli, 2015). However, also in psychology several types of
-cylindrical data can be found. For example, in research on human
-navigation in the field of cognitive psychology both the distance, a
-linear variable, and the direction, a circular variable, of movement are
-of interest (Chrastil \& Warren, 2017). In eye-tracking research,
-saccade data are an example of cylindrical data, because both the
-direction (\emph{i.e.}, the circular variable) and the duration
-(\emph{i.e.}, the linear variable) of the saccades are of interest (for
-a review of eye-tracking research see Rayner (2009)).\newline
-\indent The type of data that is used in the present study are also
-psychological, namely data from circumplex measurement instruments. For
-instance, data from the interpersonal circumplex as used in personality
-psychology are by definition of a cylindrical nature (see Section
-\ref{Example} for a more detailed explanation).\newline
-\indent In the literature, several methods have been put forward to
-model the relation between the linear and circular component of a
-cylindrical variable. Some of these are based on regressing the linear
-component onto the circular component using the following type of
-relation: \[y = \beta_0 +
+In social sciences the use of cylindrical data is very common. Such data
+consist of a linear and a circular component. Gurtman (2011) refers to
+such data as vectors, with a directional measure (i.e., the circular
+component) and a measure indicating the magnitude (i.e., linear
+component). Many established models in psychology are often referred to
+as circular or circumplex models, but those models are cylindrical.
+Examples of such cylindrical models are the interpersonal
+circle/circumplex (Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans,
+Brok, \& Tartwijk, 2006), the circumplex of affect (Russell, 1980), the
+circumplex of human emotion (Plutchik, 1997) or the model of human
+values (Schwartz, 1992). \newline \indent Also, many of the more recent
+types of data that are studied in psychology are cylindrical. For
+example, research on human navigation uses data where distance (i.e.,
+linear component) and direction (i.e., the linear component) are of
+interest (Chrastil \& Warren, 2017) or in eye-tracking, the saccade data
+also consist of both the direction (i.e., the circular variable) and the
+duration (i.e., the linear variable)(e.g., Rayner (2009)). Apart from
+the social sciences data with a circular and linear outcome more
+commonly occur in meteorology meteorology (García-Portugués, Crujeiras,
+\& González-Manteiga, 2013), ecology (García-Portugués, Barros,
+Crujeiras, González-Manteiga, \& Pereira, 2014) or marine research
+(Lagona, Picone, Maruotti, \& Cosoli, 2015) \newline \indent Up until
+now researchers studying cylindrical data had to rely on the linear
+statistical methods to analyze their research results. Focusing on only
+the circular variable or only on the linear variable. However, lately
+more and more of these researchers acknowledge that linear methods are
+not sufficient and call for new methods (Gurtman, 2011; Pennings, 2017b;
+Wright, Pincus, Conroy, \& Hilsenroth, 2009) that take into account both
+the circular and the linear component of these data. \newline
+\indent The aim in the present paper is twofold. Firstly, we intend to
+fill this gap in the literature on cylindrical models by adapting four
+existing cylindrical models in such a way that they include a regression
+of both the linear and circular component of a cylindrical variable onto
+a set of covariates. From now on we will therefore refer to the
+components of the cylindrical variable as outcome components. Secondly,
+we will show how using these models can benefit the analysis of
+circumplex data and cylindrical data in psychology in general. More
+specifically we will show these benefits for interpersonal teacher data
+from the field of educational psychology. Apart from modelling the
+dependence between the linear and circular component of a cylindrical
+variable we would also like to predict the two components from a set of
+covariates in a regression model. In our analysis, we will show how this
+can be done in four different ways using the adapted cylindrical models.
+
+\subsection{Cylindrical data}
+
+Thus, data that consist of a variable and a circular variable is called
+cylindrical data. A circular variable is different from the linear
+variable in the sense that it is measured on a different scale. Figure
+\ref{circline} shows the difference between a circular scale (right) and
+a linear scale (left). The most important difference is that on a
+circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are
+connected and in fact represent the same number while on a linear scale
+the two ends, \(-\infty\) and \(\infty\), are not connected and
+consequently the values 0\(^\circ\) and 360\(^\circ\) are located on
+different places on the scale. Both circular data and cylindrical data
+require special analysis methods due to this periodicity in the scale of
+a circular variable (see e.g.~Fisher (1995) for an introduction to
+circular data and Mardia and Jupp (2000), Jammalamadaka and Sengupta
+(2001) and Ley and Verdebout (2017) for a more elaborate
+overview).\newline \indent In the literature, several methods have been
+put forward to model the relation between the linear and circular
+component of a cylindrical variable. Some of these are based on
+regressing the linear component onto the circular component using the
+following type of relation: \[y = \beta_0 +
 \beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is
@@
 variables and modelling their covariance matrix (Mastrantonio, 2018) or
-by proposing a joint cylindrical distribution. For example, Abe \& Ley
+by proposing a joint cylindrical distribution. For example, Abe and Ley
 (2017) introduce a cylindrical distribution based on a Weibull
@@
 methods that have been proposed thus far include additional covariates
-onto which both the circular and linear component are regressed.\newline
-\indent Our aim in this paper is to fill this gap in the literature by
-adapting four existing cylindrical models in such a way that they
-include a regression of both the linear and circular component of a
-cylindrical variable onto a set of covariates. From now on we will
-therefore refer to the components of the cylindrical variable as outcome
-components. Additionally, we will show how a correct statistical
-treatment of such cylindrical data can lead to new insights. We will do
-this for the teacher data, a dataset from the field of educational
-psychology. In the teacher data, apart from modelling the relation
-between the linear and circular component of a cylindrical variable we
-would also like to predict the two components from a set of covariates
-in a regression model.
-
-The paper is organized as follows: Section \ref{Example} describes the
-teacher data, while Section \ref{Models} presents the four cylindrical
-models and our associated adaptations to new regression models. In that
-same section we also discuss the model fit criterion that we will use in
-Section \ref{DataAnalysis} for the comparison of the four models. A
-detailed data analysis with interesting new insights is also provided in
-Section \ref{DataAnalysis}. We conclude the paper with a discussion in
-Section \ref{Discussion}, and the Supplementary Material collects the
-technical details of the MCMC procedures.
+onto which both the circular and linear component are regressed.
 
@@
 climate of \VANDER{Want}{Van der}{van der} Want (2015), Claessens (2016)
-and Pennings et al. (2018). An indicator of the quality of the classroom
+and Pennings (2017a). An indicator of the quality of the classroom
 climate is the students' perception of their teachers' interpersonal
@@
 psychology as well as in other areas of psychology, can be measured
-using circumplex measurement instruments (see Horowitz \& Strack (2011)
-for an overview of many such instruments).\newline
-\indent The circumplex data used in this paper are measured using the
-Questionnaire on Teacher Interaction (QTI) (Wubbels, Brekelmans, Brok,
-\& Tartwijk, 2006) which is one such circumplex measurement instrument.
-The QTI is designed to measure student perceptions of their teachers'
-interpersonal behavior and contains items that load on two interpersonal
-dimensions: Agency and Communion. Agency refers to the degree of power
-or control a teacher exerts in interaction with his/her students.
-Communion refers to the degree of friendliness or affiliation a teacher
-conveys in interaction with his/her students. The loadings on the two
-dimensions of the QTI can be placed in a two-dimensional space formed by
-Agency (vertical) and Communion (horizontal), see Figure \ref{QTI}.
-Different parts of this space are characterized by different teacher
-behavior, e.g. `helpful' or `uncertain'. This two-dimensional space is
-called the interpersonal circle/circumplex (IPC). The IPC is ``a
-continuous order with no beginning or end'' (Gurtman, 2009, p. 2). We
-call such ordering a circumplex ordering and the IPC is therefore often
-called the interpersonal circumplex. The ordering also implies that
-scores on the IPC could be viewed as a circular variable.\newline
-\indent Cremers, Mainhard, \& Klugkist (2018a) explain the circular
-nature of the IPC data and analyze them as such using a circular
-regression model. The two-dimension scores Agency and Communion can be
-converted to a circular score using the two-argument arctangent function
-in \eqref{PredVal}, where \(A\) represents a score on the Agency
-dimension and \(C\) represents a score on the Communion dimension
-
+using circumplex measurement instruments (see Horowitz and Strack (2011)
+for an overview of many such instruments).\newline \indent The
+circumplex data used in this paper are measured using the Questionnaire
+on Teacher Interaction (QTI) (Wubbels et al., 2006) which is one such
+circumplex measurement instrument. The QTI is designed to measure
+student perceptions of their teachers' interpersonal behavior and
+contains items that load on two interpersonal dimensions: Agency and
+Communion. Agency refers to the degree of power or control a teacher
+exerts in interaction with his/her students. Communion refers to the
+degree of friendliness or affiliation a teacher conveys in interaction
+with his/her students. The loadings on the two dimensions of the QTI can
+be placed in a two-dimensional space formed by Agency (vertical) and
+Communion (horizontal), see Figure \ref{QTI}. Different parts of this
+space are characterized by different teacher behavior,
+e.g.~\enquote{helpful} or \enquote{uncertain}. This two-dimensional
+space is called the interpersonal circle/circumplex (IPC). The IPC is
+``a continuous order with no beginning or end'' (Gurtman, 2009, p. 2).
+We call such ordering a circumplex ordering and the IPC is therefore
+often called the interpersonal circumplex. The ordering also implies
+that scores on the IPC could be viewed as a circular variable. This
+circular variable represents the type of interpersonal behavior that a
+teacher shows towards his/her students.\newline \indent Cremers et al.
+(2018a) explain the circular nature of the IPC data and analyze them as
+such using a circular regression model. The two dimension scores, Agency
+and Communion, can be converted to a circular score using the
+two-argument arctangent function in
+\eqref{PredVal}\footnote{The selection of
+the origin in circumplex data depends on the scaling of the Agency and Communion
+scores; their respective 0 scores form the origin. Agency and Communion are
+measured on a scale from 1 to 5 and for analysis purposes they are later
+converted to scale ranging from -1 to 1. Scaling is however only considered an
+issue in those instances where cylindrical data is derived from measurements in
+bivariate space.}, where \(A\) represents a score on the Agency
+dimension and \(C\) represents a score on the Communion dimension. Note
+that when placing a unit circle on Figure \ref{QTI} we see that the
+Agency dimension is related to the sine of the circular score and the
+Communion dimension is related to the cosine of the circular score.
 \begin{equation}\label{PredVal}
@@
 \right.
-\end{equation}
-
-The resulting circular variable \(\theta\) can then be modelled and
-takes values in the interval \([0, 2\pi)\). However, when
+\end{equation} The resulting circular variable \(\theta\) can then be
+modelled and takes values in the interval \([0, 2\pi)\) or
+\([0^\circ, 360^\circ)\). Note that the round brackets mean that
+\(2\pi\) and \(360^\circ\) are not included in the interval since these
+represent the same value as 0 as a result of periodicity. \newline
+\indent A circular analysis of circumplex data has several benefits: it
+is more in line with its theoretical definition and it allows us to
+analyse the blend of the two dimensions Agency and Communion instead of
+both dimensions separately providing us with new insights compared to a
+separate analysis of the two dimensions that is standard in the
+literature (see e.g.~Pennings et al. (2018), Wright et al. (2009) or
+Wubbels et al. (2006)). There is however one main drawback: when
 two-dimensional data are converted to the circle we lose some
@@
 \(\mid\mid (A, C)^t \mid\mid\). This length represents the strength of
-the type of interpersonal behavior a teacher shows towards his/her
-students and can be considered as the linear variable in a cylindrical
-model, allowing us to model a circular variable \(\theta\) together with
-the linear variable corresponding to \(\mid\mid (A, C)^t \mid\mid\).
-This leads to an improved analysis of interpersonal circumplex data as
-we take all information into account. In the next section we introduce
-several models that can be used for a more accurate and informative
-regression analysis on the teacher data. First however we will provide
-descriptives for our data set.
-
-\subsection{Data description}\label{DataDescriptives}
+the interpersonal behavior a teacher shows towards his/her students. In
+a cylindrical model this strength (the linear outcome) can be modeled
+together with the type of interpersonal behavior of a teacher (the
+circular outcome). This leads to an improved analysis of interpersonal
+circumplex data, over either analyzing the two dimensions separately or
+using a circular model, because we take all information, circular and
+linear, into account. In the next section we introduce several
+cylindrical models that can be used to analyze the teacher data. First
+however we will provide descriptives for the teacher data.
+
+\subsection{Data Description}\label{DataDescriptives}
 
@@
 this paper we only consider one measurement, the first occasion (2010)
-and largest class if data for multiple classes were available. In
-addition to the score on the IPC, the circular outcome, and the strength
-of the score on the IPC, the linear outcome, a teachers' self-efficacy
-(\verb|SE|) concerning classroom management is used as covariate in the
-analysis. After listwise deletion of missings (\(3\) in total, only for
-the self-efficacy) we have a sample of 148 teachers. Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset. Here
-\(\hat{\rho}\) is a sample estimate for the circular concentration where
-a value of 0 means that the data is not concentrated at all, \emph{i.e.}
-spread over the entire circle, and a value of 1 means that all data is
-concentrated at a single point on the circle. Figure \ref{dataplot} is a
-scatterplot showing the relation between the linear and circular outcome
-of the teacher data.
+and largest class if data for multiple classes were available. This
+results in a sample of 151 teachers. In addition to the type of
+interpersonal behavior (IPC), the circular outcome, and the strength of
+interpersonal behavior (IPC strength), the linear outcome, a teachers'
+self-efficacy (\verb|SE|) concerning classroom management is used as
+covariate in the analysis. In previous research, in psychology and
+education it had been shown that higher self-efficacy is related to the
+quality of interpersonal interactions (Locke \& Sadler, 2007;
+\VANDER{Want}{Van der}{van der} Want et al., n.d.). After listwise
+deletion of missings (\(3\) in total, only for the self-efficacy) we
+have a sample of 148 teachers. Table \ref{Tableteacherdescriptives}
+shows descriptives for the dataset. For the circular variable IPC we
+show sample estimates for the circular mean \(\bar{\theta}\) and
+concentration \(\hat{\rho}\). The circular concentration lies between 0,
+meaning the data is not concentrated at all \emph{i.e.} spread over the
+entire circle, and 1, meaning all data is concentrated at a single point
+on the circle The population values of these parameters are usually, and
+also in this paper, referred to as \(\mu\) (circular location) and
+\(\kappa\) (circular concentration). For the linear variables (strength
+IPC and SE) we show sample estimates of the linear mean and standard
+deviation. Figure \ref{dataplot} is a scatterplot showing the relation
+between the linear and circular outcome of the teacher data for teachers
+with low SE (below 1 sd below the mean), average SE (between 1 sd below
+and 1 sd above the mean) and high SE (above 1 sd above the mean).
 
@@
 
-\section{Four cylindrical regression models}\label{Models}
-
-In this section we present four cylindrical models and adapt them such
-that they contain predictors for the linear and circular outcomes, \(Y\)
-and \(\Theta\). The first two models are based on a construction by
-Mastrantonio et al. (2015), while the other models are extensions of the
-models from Abe \& Ley (2017) and Mastrantonio (2018).
-
-\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}
-
-Following Mastrantonio et al. (2015) we consider in this section two
-models where the relation between \(\Theta \in [0, 2\pi)\) and
+\section{Four Cylindrical Regression Models}\label{Models}
+
+One of the goals of this paper is to show the benefits of cylindrical
+methods for the analysis of circumplex data and cylindrical data in
+psychology in general. To do so we decide to focus on four cylindrical
+models. The models were selected for their relatively low complexity and
+the ease with which a regression structure could be incorporated. But
+also because they show different ways of modelling the the linear and
+circular outcome and thereby illustrate a wider range of cylindrical
+models available in the literature. All four cylindrical models contain
+a set of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\)
+predictors \(\boldsymbol{z} = z_1, \dots, z_q\) for the linear and
+circular outcomes, \(Y\) and \(\Theta\), respectively. The first two
+models are based on a construction by Mastrantonio et al. (2015), while
+the other models are extensions of the models from Abe and Ley (2017)
+and Mastrantonio (2018). The four cylindrical models are introduced
+separately in the subsections below. However, to provide a more succinct
+overview and comparison of the four models, \ref{TableModels} gives an
+overview of the similarities and differences between the models.
+
+\begin{sidewaystable}[h]
+\centering
+\caption{Comparison of the four cylindrical regression models} 
+\begin{tabular}{lllll}
+  \noalign{\smallskip}\hline\noalign{\smallskip}
+\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
+$\Theta$ & &&&\\
+$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
+$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
+$\:\:$Shape & symmetric, & assymetric, & assymetric, & assymetric, \\
+            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
+$Y$& &&&\\
+$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
+$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(0, + \infty)$\\
+$\:\:$Shape & symmetric, & symmetric, & assymetric, & assymetric, \\
+            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
+$\Theta$-$Y$ dependence &                                   &                                   & & \\
+                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
+                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
+\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
+\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
+
+\end{tabular}
+\label{TableModels}
+\end{sidewaystable}
+
+\subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+
+Following Mastrantonio et al. (2015) we consider two models where the
+relation between \(\Theta \in [0, 2\pi)\) and
 \(Y\in (-\infty, + \infty)\) and \(q\) covariates is specified as
-
 \begin{equation}\label{circlinlink}
 Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
-\end{equation}
-
-\noindent where the random variable \(R\geq0\) will be introduced below,
-the error term \(\epsilon \sim N(0, \sigma^2)\) with variance
-\(\sigma^2>0\),
+\end{equation} \noindent where the random variable \(R\geq0\) will be
+introduced below, the error term \(\epsilon \sim N(0, \sigma^2)\) with
+variance \(\sigma^2>0\),
 \(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
 the intercept and regression coefficients and \(x_1, \dots, x_q\) are
-the \(q\) covariates. In both of these models the conditional
-distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given by
-
-\begin{equation}\label{ycondtheta}
-f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
-\end{equation}
-
-\noindent where
-\(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
-\(r \geq 0\). The linear outcome thus has a normal distribution
-conditional on \(\Theta\) and \(R\) and contains already linear
-covariates \(x_1, \dots, x_q\) in its location part. For the teacher
-dataset, the regression equation for the linear outcome in the CL-PN and
-CL-GPN model is the following: \[\hat{y}_i =
-\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i +
-\gamma_1\text{SE}_i,\] \noindent where \(\text{SE}_i\) is the
-self-efficacy score of one individual \(i = 1, \dots, n\) where \(n\) is
-the sample size.\newline
-\indent For the circular outcome we assume either a projected normal
-(PN) or a general projected normal (GPN) distribution. These
-distributions arise from the radial projection of a distribution defined
-on the plane onto the circle. The relation between a bivariate vector
-\(\boldsymbol{S}\) in the plane and the circular outcome \(\Theta\) is
-defined as follows
-
-\begin{equation}\label{projection}
+the \(q\) covariates for the prediction of the linear outcome. We thus
+assume a normal distribution for the linear outcome.\newline \indent For
+the circular outcome we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise
+from a projection of a distribution defined in bivariate space onto the
+circle. Figure \ref{projection} represents this projection. In the left
+plot of Figure \ref{projection} we see datapoints from the bivariate
+variable \(\boldsymbol{S}\) that in the middle plot are projected to
+form the circular outcome \(\Theta\) in the right plot. Mathematically
+the relation between \(\boldsymbol{S}\) and \(\Theta\) is defined as
+follows \begin{equation}\label{projection}
 \boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
-\end{equation}
-
-\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean
-norm of the bivariate vector \(\boldsymbol{S}\). In the PN distribution
-we assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\)
-and in the GPN we assume
-\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
-\(\boldsymbol{\mu} \in \mathbb{R}^2\),
-\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
-\(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
-Hernandez-Stumpfhauser, Breidt, \& Woerd (2016)). This leads to the
-circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN)
-distributions. We will now detail how we modify both cylindrical
-distributions to also incorporate covariates for the circular part.
-
-\subsubsection{The modified CL-PN distribution}
-
-Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
-density of \(\Theta\) and \(R\) for the PN distribution equals
-
-\begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
-\end{equation}
-
-\noindent where
-\(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
-and \(r\) is the same as in \eqref{circlinlink} and \eqref{ycondtheta}
-and is defined in \eqref{projection}. In a regression setup the outcomes
-\(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where \(n\) is
-the sample size, are generated independently from the distribution with
-density \eqref{pnreg}. The mean vector
-\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
-\(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
-vector \(\boldsymbol{z}_i\) is a vector of dimension \(p + 1\) that
-contains the covariate values and the value 1 to estimate an intercept
-and
-\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
-contains the regression coefficients and intercepts. Note however that
-the dimensions of \(\boldsymbol{\beta}^{I}\) and
-\(\boldsymbol{\beta }^{II}\) need not necessarily be the same and we are
-thus allowed to have a different set of predictor variables and vectors
-\(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for the two
-components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
-regression equation for the circular outcome in the CL-PN model is
-\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix} \mu_{i}^{I}  \\ \mu_{i}^{II}
-\end{pmatrix}=\begin{pmatrix} \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
-\beta_0^{II} + \beta_1^{II}\text{SE}_i \end{pmatrix}.\]
-
-\subsubsection{The modified CL-GPN distribution}
-
-Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
-(2016) the joint density of \(R\) and \(\Theta\) for the GPN
-distribution equals
-
-\begin{equation}\label{gpnreg}
-f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
-\end{equation}
-
-\noindent where we recall that
-\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\).
-In a regression setup the outcomes \(\theta_i\) and \(r_i\) for each
-individual are generated independently from \eqref{gpnreg}. The mean
-vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the same
-way via covariates as for the modified CL-PN distribution. Note in
-contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
-\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the
-same predictors for both components of \(\boldsymbol{\mu}_i\) in the
-CL-GPN model. This is due to the fact that the variance-covariance
-matrix \(\boldsymbol{\Sigma}\) is no longer identity. For the teacher
-dataset, the regression equation for the circular outcome in the CL-GPN
-model is the same as in the CL-PN model.
-
-\subsubsection{Parameter estimation}
-
-Both cylindrical models introduced here are estimated using Markov Chain
-Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
-Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
-regression of the circular outcome. A detailed description of the
-Bayesian estimation and MCMC samplers can be found in the Supplementary
-Material.
-
-\subsection{The modified Abe-Ley model}\label{WeiSSVM}
-
-This model is an extension of the cylindrical model introduced in Abe \&
-Ley (2017) to the regression context. The joint density of \(\Theta\)
-and \(Y\), in this model defined only on the positive real half-line
-\([0, + \infty)\), reads
-
-\begin{equation}\label{WeiSSVMdensity}
-f(\theta, y) = \frac{\alpha\beta^\alpha}{2\pi\cosh(\kappa)}
-                 (1 +\lambda\sin(\theta - \mu))
-                 y^{\alpha-1}
-                 \exp[-(\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
-\end{equation}
-
-\noindent where \(\alpha > 0\) is a linear shape parameter,
-\(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular concentration
-and skewness parameters with \(\kappa\) also regulating the
-circular-linear dependence. Our modification occurs at the level of the
-linear scale parameter \(\beta>0\) and circular location parameter
-\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
-\(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
-\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\).
-The parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression
-coefficients \(\nu_j \in (-\infty, +\infty)\) for the prediction of
-\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
-parameter \(\eta_0 \in [0, 2\pi)\) is the intercept and
-\(\boldsymbol{\eta}\) is a vector of \(p\) regression coefficients
-\(\eta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\) where
-\(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector of
-predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\) is
-a vector of predictor values for the prediction of \(\theta\). In a
-regression setup the outcome vector \((\theta_i, y_i)^t\) for each
-individual is generated independently from the modified density
-\eqref{WeiSSVMdensity}.\newline
-\indent As in Abe \& Ley (2017), the conditional distribution of \(Y\)
-given \(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\)
-and scale \(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and
-the conditional distribution of \(\Theta\) given \(Y=y\) is a sine
-skewed von Mises distribution with location parameter \(\mu\) and
-concentration parameter \((\beta y)^\alpha\tanh(\kappa)\). The
-log-likelihood for this model equals
-
-\begin{align}\label{WeiSSVMLikelihood}
-l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
-   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\nu} \nonumber\\
-   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
-   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
-   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))).\nonumber
-\end{align}
-
-\noindent For the teacher data, \(\boldsymbol{z} = \boldsymbol{x}\) and
-the regression equations for the circular and linear outcomes in the
-Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
-\tan^{-1}(\eta_1\text{SE}_i),\] and \[\hat{\beta}_{i} = \exp(\nu_0 +
-\nu_1\text{SE}_i).\] \noindent We can use numerical optimization
-(Nelder-Mead) to find solutions for the maximum likelihood (ML)
-estimates for the parameters of the model.
-
-\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}
+\end{equation} \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\),
+the Euclidean norm of \(\boldsymbol{S}\); the lines connecting the
+bivariate datapoints to the origin in the middle plot. We call
+\(\boldsymbol{S}\) the augmented representation of the circular outcome,
+it is a variable that in contrast to \(\Theta\) is not observed and thus
+considered latent or auxiliary. This then means that we do not model
+\(\Theta\) directly but indirectly through \(\boldsymbol{S}\). \newline
+\indent For both the PN and GPN distribution the circular location
+parameter \(\mu\in [0, 2\pi)\) is modeled as
+\(\hat{\mu}_i = \mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) = \mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i, \boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
+model the circular location parameter also depends on the variance-covariance
+matrix and the circular predicted values should be computed using numerical
+integretion or Monte Carlo methods because a closed form expression for the mean
+direction is not available.} where \(\boldsymbol{\beta}^{I}\) and
+\(\boldsymbol{\beta}^{II}\) are vectors with intercepts and regression
+coefficients for the prediction of \(\boldsymbol{S}\) and
+\(\boldsymbol{z}_i\) is a vector with predictor values for each
+individual \(i \in 1, \dots, n\) where \(n\) is the sample size. Note
+that as a result of the augmented representation of the circular outcome
+we have two sets of regression coefficients and intercepts, one for each
+bivariate component of \(\boldsymbol{S}\). This leads to problems when
+we want to interpret the effect of a covariate on the circle. A circular
+regression line is shown in Figure \ref{circregline}, with covariate
+values on the x-axis and the predicted circular outcome on the y-axis.
+As can be seen it is of a non-linear character meaning that the effect
+of a covariate is different at different values of the covariate. A
+circular regression line is usually described by the slope at the
+inflection point, the point at which the slope of the regression line
+starts flattening off (indicated with a square in Figure
+\ref{circregline}). By default, the parameters from the PN and GPN model
+do not directly describe this inflection point. For the PN distribution
+however, Cremers et al. (2018b) solved this interpretation problem and
+introduce new circular regression coefficients. They introduce a new
+parameter \(b_c\) that describes the slope at the inflection point of
+the regression line. For the GPN distrbution the interpretation problem
+however remains.\newline \indent The main difference between the PN and
+GPN distribution lies in the definition of their covariance matrix. For
+the PN distribution this is an identity matrix, causing the distribution
+to be unimodal and symmetric, whereas for the GPN distribution
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\),
+allowing for multimodality and assymetry/skewness.\newline \indent For
+the teacher dataset the predicted linear outcome, strengths of
+interpersonal behavior, in the CL-PN and CL-GPN model is the following:
+\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted
+circular outcome, type of interpersonal behavior, equals:
+\[\hat{\theta}_i = \mu_i =
+\mbox{atan2} (\beta_0^{II} + \beta_1^{II}\text{SE}_i, \beta_0^{I} +
+\beta_1^{I}\text{SE}_i).\] \noindent where \(\text{SE}_i\) is the
+self-efficacy score of one individual. The CL-PN and CL-GPN models thus
+allow us to assess the average type and strength of interpersonal
+behavior through the parameters \(\beta_{0}^{I}\), \(\beta_{0}^{II}\)
+and \(\gamma_0\) as well as the effect of self-efficacy on type and
+strength of teacher behavior through the parameters, \(\beta_{1}^{I}\),
+\(\beta_{1}^{II}\) and \(\gamma_1\). In addition, because the type and
+strength of interpersonal behavior are modelled together via the
+regression in \eqref{circlinlink} we can assess the effect of the type
+of interpersonal behavior on the strength through the parameters
+\(\gamma_{sin}\) and \(\gamma_{cos}\). In the teacher data these are the
+regression coefficients for the effect of the sine and cosine of the
+type of behavior which are related to the scores on the Agency an
+Communion dimensions respectively.\newline \indent Both the CL-PN and
+CL-GPN models are estimated using Markov Chain Monte Carlo (MCMC)
+methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011),
+Wang and Gelfand (2013) and Hernandez-Stumpfhauser, Breidt, and Woerd
+(2016) for the regression of the circular outcome. A detailed
+description of the Bayesian estimation and MCMC samplers can be found in
+the Supplementary Material.
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
+\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
+The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norm of the bivariate datapoints, the random variable $R$.}
+\label{projection}
+\end{figure}
+
+\begin{figure}[]
+  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
+  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to.
+The square indicates the inflection point of the regression live.}
+  \label{circregline}
+\end{figure}
+
+\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+
+This model is an extension of the cylindrical model introduced in Abe
+and Ley (2017) to the regression context. It concerns a combination of a
+Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
+\(\alpha\), for the linear outcome and a sine-skewed von Mises
+distribution, with location parameter \(\mu\in [0, 2\pi)\),
+concentration parameter \(\kappa>0\) and skewness
+\(\lambda \in [-1,1]\), for the circular outcome. In contrast to the
+CL-PN and CL-GPN models, the linear outcome \(Y\) is in this model
+defined only on the positive real half-line \([0, + \infty)\) and thus
+can not be negative.\newline \indent In this model we predict the linear
+scale parameter and circular location parameter, both of which we can
+express in terms of covariates:
+\(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
+\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
+The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
+coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
+\(y\) where \(j = 0, \dots, q\) and \(\gamma_0\) is the intercept. The
+parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
+\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
+\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
+where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
+of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
+is a vector of predictor values for the prediction of
+\(\theta\).\newline \indent For the teacher data, the predicted values
+for the circular outcome in the Abe-Ley model are:
+\[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
+\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear
+outcome. The conditional distribution for the linear outcome is Weibull,
+meaning that we can use methods from survival analsis to interpret the
+effect of a predictor. In survival analysis a \enquote{survival}
+function is used in which time is plotted against the probability of
+survival of subjects suffering from a specific medical condition. In the
+teacher data we can thus compute the probability of a teacher having a
+specific strength on the IPC. This probability is computed using the
+\enquote{survival-function} defined as \[\exp(-\alpha
+y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
+with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the
+survival function we also see that the circular concentration parameter
+\(\kappa\) and the linear shape parameter \(\alpha\) regulate the
+circular-linear dependence in the Abe-Ley model. The Abe-Ley model thus
+allows us to assess the average type and strength of interpersonal
+behavior through the parameters \(\beta_{0}\) and \(\gamma_0\) as well
+as the effect of self-efficacy on type and strength of teacher behavior
+through the parameters, \(\beta_{1}\) and \(\gamma_1\).\newline
+\indent We can use numerical optimization (Nelder-Mead) to find
+solutions for the maximum likelihood (ML) estimates for the parameters
+of the model.
+
+\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
 
 This model is an extension of the cylindrical model introduced by
-Mastrantonio (2018) to the regression context. Both models contain \(m\)
+Mastrantonio (2018) to the regression context. The model contains \(m\)
 independent circular outcomes and \(w\) independent linear outcomes. The
@@
 \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\)
-are modelled together by a multivariate GPN distribution. The joint
-distribution of \(\boldsymbol{\Theta}\) and \(\boldsymbol{R}\) can thus
-be modeled as the product of (\ref{gpnreg}) for each of the \(m\)
-circular outcomes. The linear outcomes
+are modelled together by a multivariate GPN distribution. The linear
+outcomes
 \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,  \boldsymbol{Y}_w)\) are
 modelled together by a multivariate skew normal distribution (Sahu, Dey,
-\& Branco, 2003). Because the GPN distribution is modelled using a
-so-called augmented representation (as in \eqref{projection} and
-\eqref{gpnreg}) it is convenient to use a similar tactic for modelling
-the multivariate skew normal distribution. Following Mastrantonio (2018)
-the linear outcomes are represented as
-\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
-\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear
-outcome \(\boldsymbol{Y}\),
-\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
-\(w \times w\) diagonal matrix with diagonal elements
-\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
-\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
-\(w\)-dimensional half normal distribution (Olmos, Varela, Gómez, \&
-Bolfarine, 2012), and
-\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\).
-This means that, conditional on the auxiliary data \(\boldsymbol{D}\),
-\(\boldsymbol{Y}\) is normally distributed with mean
-\(\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and
-covariance matrix \(\boldsymbol{\Sigma}_y\). The joint density for
-\((\boldsymbol{Y}^t, \boldsymbol{D}^t)^t\) is defined as:
-
-\begin{equation}\label{YDjoint}
-f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
-\end{equation}
-
-\noindent where
-\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\)
-stands for the \(\ell\)-dimensional normal density with mean vector
-\(\boldsymbol{\mu}_\ell\) and covariance \(\boldsymbol{\Sigma}_\ell\).
-As in Mastrantonio (2018) dependence between the linear and circular
-outcome is created by modelling the augmented representations of
+\& Branco, 2003). Thus also in this model the linear outcome is defined
+constrained to be positive. Because the GPN distribution is modelled
+using a so-called augmented representation (see also the description of
+the CL-PN and CL-GPN models) it is convenient to use a similar tactic
+for modelling the multivariate skew normal distribution. As in
+Mastrantonio (2018) dependence between the linear and circular outcome
+is created by modelling the augmented representations of
 \(\boldsymbol{\Theta}\) and \(\boldsymbol{Y}\) together in a \(2m + w\)
-dimensional normal distribution. The joint density of the model is then
-represented by:
-
-\begin{equation}\label{YDThetarjoint} 
-f(\boldsymbol{\theta}, \boldsymbol{r},
-\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
-\boldsymbol{y}^t)^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, ({\rm
-diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
-\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
-1}^{m}r_j, 
-\end{equation}
-
-\noindent where
-\(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots, r_m(\cos(\theta_m), \sin(\theta_m)))^t\),
-the mean vector
-\(\boldsymbol{\mu} = (\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\)
-and
-\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
-The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
-variances of and covariances between the augmented representations of
-the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
-contains covariances between the augmented representations of the
-circular outcome and the linear outcome. \newline
-\indent In our regression extension we have \(i = 1, \dots, n\)
-observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
-covariates. The mean in the density in \eqref{YDThetarjoint} then
-becomes \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
+dimensional normal distribution.\newline \indent This means that we have
+a shared mean vector and variance-covariance matrix for the linear and
+circular outcome(s), much like having multiple outcomes in a MANOVA
+(multivariate analysis of variance) model. In our regression extension
+of the GPN-SSN model we have \(i = 1, \dots, n\) observations of \(m\)
+circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean
+vector then becomes
+\(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
 \(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with
@@
 \(g + 1\) dimensional vector containing the value 1 to estimate an
-intercept and the \(g\) covariate values. \newline
-\indent For the teacher data, the regression equations for the circular
+intercept and the \(g\) covariate values. This means that in contrast to
+the other three models, we have to use the same set of predictors for
+the circular and linear outcome.\newline \indent For the teacher data,
+\(\boldsymbol{B} = \begin{bmatrix} \beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
+The predicted
+circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.}
 and linear outcomes in the GPN-SSN model are
-\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +
-\boldsymbol{\beta}_1\text{SE}_i,\] where
-\(\hat{\boldsymbol{\mu}}_i = (\hat{\boldsymbol{\mu}}_{s_i}^t, \hat{\mu}_{y_i})^t\),
-\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
-and
-\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
-Note that because \(m = 1\) and \(w = 1\), \(\boldsymbol{\mu}_{s_i}\) is
-a 2 dimensional vector and \(\mu_{y_i}\) is a scalar. We estimate the
+\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
+\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
+\beta_{1_{s^{I}}}\text{SE}_i),\]
+
+\noindent and
+
+\[\hat{y_i} = \beta_{0_{y_i}} +
+\beta_{1_{y_i}}\text{SE}_i.\]
+
+\noindent The GPN-SSN model thus allows us to assess the average type
+and strength of interpersonal behavior through the parameters
+\(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\) and \(\beta_{0_{y}}\) as
+well as the effect of self-efficacy on type and strength of teacher
+behavior through the parameters, \(\beta_{1_{s^{I}}}\),
+\(\beta_{1_{s^{II}}}\) and \(\beta_{1_{y}}\). In addition, because the
+type and strength of interpersonal behavior are modelled together using
+a multivariate normal distribution we can through its
+variance-covariance matrix also assess the dependence between the type
+and strength of interpersonal behavior. \newline \indent We estimate the
 model using MCMC methods. A detailed description of these methods is
@@
 
-\subsection{Model fit criterion}\label{Modelfit}
+\subsection{Model Fit Criterion}\label{Modelfit}
 
@@
 9 times as part of a training set. The analysis will thus be performed
-10 times, each time on a different training set.\newline
-\indent A proper criterion to compare out-of-sample predictive
-performance is the Predictive Log Scoring Loss (PLSL) (Gneiting \&
-Raftery, 2007). The lower the value of this criterion, the better the
-predictive performance of the model. Using ML estimates this criterion
-can be computed as follows:
-
-\begin{equation}\label{PLSLML}
+10 times, each time on a different training set.\newline \indent A
+proper criterion to compare out-of-sample predictive performance is the
+Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007). The
+lower the value of this criterion, the better the predictive performance
+of the model. Using ML estimates this criterion can be computed as
+follows: \begin{equation}\label{PLSLML}
 PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
-\end{equation}
-
-\noindent where \(l\) is the model likelihood, \(M\) is the sample size
-of the holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout
-set and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the
-model parameters. Using posterior samples the criterion is similar to
-the log pointwise predictive density (lppd) (Gelman et al., 2014, p.
-169) and can be computed as:
-
+\end{equation} \noindent where \(l\) is the model likelihood, \(M\) is
+the sample size of the holdout set, \(x_i\) is the \(i^{th}\) datapoint
+from the holdout set and \(\hat{\boldsymbol{\vartheta}}\) are the ML
+estimates of the model parameters. Using posterior samples the criterion
+is similar to the log pointwise predictive density (lppd) (Gelman et
+al., 2014, p. 169) and can be computed as:
 \begin{equation}\label{PLSLBayes}
 PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
-\end{equation}
-
-\noindent where \(B\) is the amount of posterior samples and
-\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
+\end{equation} \noindent where \(B\) is the amount of posterior samples
+and \(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
 model parameters for the \(j^{th}\) iteration. Because the joint density
-and thus also the likelihood for the modified GPN-SSN model in
-(\ref{YDThetarjoint}) is not available in closed form (Mastrantonio,
-2018) we compute the PLSL for the circular and linear outcome separately
-for all models. Note that although we fit the CL-PN, CL-GPN and GPN-SSN
-models using Bayesian statistics, we do not take prior information into
-account when assessing model fit with the PLSL. According to Gelman et
-al. (2014) this is not necessary since we are assessing the fit of a
-model to data, the holdout set, only. They argue that the prior in such
-case is only of interest for estimating the parameters of the model but
-not for determining the predictive accuracy.\newline
-\indent We use the loglikelihoods of the following conditional densities
-for the computation of the PLSL in the teacher data:
-
-\begin{itemize}
-\item For the modified CL-PN model:
-
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
-
-\item For the modified CL-GPN model:
-
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
-
-\item For the modified Abe-Ley model:
-
-$y_i \mid \theta_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.
-
-$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.
-
-\item For the modified joint projected and skew normal model:
-
-$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$
-
-$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$
-
-where $SSN$ is the skew normal distribution.
-\end{itemize}
-
-\noindent For each of the four cylindrical models and for each of the 10
-cross-validation analyses we can then compute a PLSL for the circular
-and linear outcome by using the conditional log-likelihoods of the
-respective outcome. To evaluate the predictive performance we average
-across the PLSL criteria of the cross-validation analyses. We also
-assess the cross-validation variability by means of the standard
-deviations of the PLSL criteria.
+and thus also the likelihood for the modified GPN-SSN model is not
+available in closed form (Mastrantonio, 2018) we compute the PLSL for
+the circular and linear outcome separately for all models. Note that
+although we fit the CL-PN, CL-GPN and GPN-SSN models using Bayesian
+statistics, we do not take prior information into account when assessing
+model fit with the PLSL. According to Gelman et al. (2014) this is not
+necessary since we are assessing the fit of a model to data, the holdout
+set, only. They argue that the prior in such case is only of interest
+for estimating the parameters of the model but not for determining the
+predictive accuracy.\newline \indent For each of the four cylindrical
+models and for each of the 10 cross-validation analyses we can then
+compute a PLSL for the circular and linear outcome by using the
+conditional log-likelihoods of the respective outcome (see Supplementary
+Material for a definition of the loglikelihoods). To evaluate the
+predictive performance we average across the PLSL criteria of the
+cross-validation analyses. We also assess the cross-validation
+variability by means of the standard deviations of the PLSL criteria.
 
@@
 In this section we analyze the teacher data with the help of the four
-cylindrical models from Section \ref{Models}. We will present the
-results, posterior estimates and their interpretation, per model and
-finish with a section comparing the fit of the different models.
+cylindrical models from the previous section. We will present the
+results, posterior estimates and their interpretation for all four
+models.
 
@@
 In the Supplementary Material we have described the starting values for
-the MCMC procedures, hence it remains to specify the starting values for
-the maximum likelihood based Abe-Ley model:
+the MCMC procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it
+remains to specify the starting values for the maximum likelihood based
+Abe-Ley model:
 \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
-The initial amount of iterations for the three MCMC samplers was set to
+The initial number of iterations for the three MCMC samplers was set to
 2000. After convergence checks via traceplots we concluded that some of
 the parameters of the GPN-SSN model did not converge. Therefore we set
-the amount of iterations of the MCMC models to 20,000 and subtracted a
-burn-in of 5000 to reach convergence. Note that we choose the same
-amount of iterations for all three Bayesian models to make their
-comparison via the PLSL as fair as possible. Lastly, the predictor
-\verb|SE| was centered before inclusion in the analysis as this allows
-the intercepts to bear the classical meaning of average behavior.
-
-\subsubsection{The modified CL-PN and CL-GPN models} \label{resCL(G)PN}
+the number of iterations of the MCMC models to 20,000 and subtracted a
+burn-in of 5000 to reach convergence (the geweke diagnostics show
+absolute z-scores over 1.96 in 6\% of the estimated parameters). Note
+that we choose the same number of iterations for all three models models
+estimated using MCMC prodedures to make their comparison via the PLSL as
+fair as possible. Lastly, the predictor \verb|SE| was centered before
+inclusion in the analysis as this allows the intercepts to bear the
+classical meaning of average behavior.\newline \indent Tables
+\ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show the
+results for the four cylindrical models that were fit to the teacher
+data. For the models estimated using MCMC methods, CL-PN, CL-GPN and
+GPN-SSN we show descriptives of the posterior of the estimated
+parameters (posterior mode and lower and upper bound of the 95\% highest
+posterior density (HPD) interval). For the Abe-Ley model we show the
+maximum likelihood estimates of the parameters. To compare the results
+of the four models we focus on the following aspects: the estimated
+average scores (intercept) on the type and strength of interpersonal
+behavior (1), the effect of self-efficacy on the type and strength of
+interpersonal behavior (2), the dependence between the type and strength
+of interpersonal behavior (3) and the model fit (4).
 
@@
 \midrule
-$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 0.09 (0.09) & 2.42 (0.15) & 1.90 (0.10) & 3.05 (0.17)\\
-$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.08 (0.08) & 0.85 (0.12) & 0.45 (0.09) & 1.30 (0.15)\\
-$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 0.04 (0.04) & 1.47 (0.05) & 1.17 (0.04) & 1.79 (0.05)\\
-$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.04 (0.04) & 0.70 (0.08) & 0.46 (0.05) & 0.96 (0.07)\\
-$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.01 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
+$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 2.03 (0.09) & 2.43 (0.12) & 1.91 (0.10) & 3.05 (0.17)\\
+$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.90 (0.08) & 0.84 (0.11) & 0.45 (0.09) & 1.29 (0.15)\\
+$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 1.37 (0.04) & 1.47 (0.05) & 1.16 (0.04) & 1.78 (0.05)\\
+$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.79 (0.04) & 0.70 (0.06) & 0.47 (0.05) & 0.96 (0.08)\\
+$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.44 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
 \addlinespace
-$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.00 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.05 (0.00)\\
-$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.00 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
-$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.01) & -0.00 (0.00) & 0.06 (0.00)\\
-$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.00 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
-$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.02 (0.25) & 1.83 (0.14) & 5.02 (0.41)\\
+$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.06 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.04 (0.00)\\
+$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.02 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
+$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.00) & -0.00 (0.00) & 0.06 (0.00)\\
+$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
+$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.04 (0.29) & 1.85 (0.13) & 5.00 (0.41)\\
 \addlinespace
-$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.46 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
+$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.47 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
 $\sum_{2,2}$ & NA (NA) & NA (NA) & NA (NA) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
 \bottomrule
+\multicolumn{7}{l}{Note: $\beta_0^{I}$, $\beta_0^{II}$ and $\gamma_0$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{at the average self-efficacy. $\beta_1^{I}$, $\beta_1^{II}$ and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
+\multicolumn{7}{l}{type and strength of interpersonal behavior. $\gamma_{cos}$ and $\gamma_{sin}$ inform us about the dependence}\\
+\multicolumn{7}{l}{ between the type and strength of interpersonal behavior. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are are}\\
+\multicolumn{7}{l}{elements of the variance-covariance matrix of the type of interpersonal behavior in the}\\
+\multicolumn{7}{l}{CL-GPN model and $\sigma$ is the error standard deviation of the strength of interpersonal behavior.}\\
+
 \end{tabular}
@@
 \begin{table}
-\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean in the CL-PN, CL-GPN and GPN-SSN models}
+
+\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation (SD), for the modified Abe-Ley model}
+\centering
+\begin{tabular}[t]{llllllll}
+\toprule
+& $\beta_0$ & $\beta_1$ & $\gamma_0$  & $\gamma_1$ & $\alpha$ & $\kappa$ & $\lambda$\\
+Mean & 0.36 & -0.03 & 1.17 & 0.04 & 3.66 & 1.51 & 0.70 \\
+SD & 0.02) & 0.01 & 0.02 & 0.02 & 0.12 & 0.08 & 0.05\\
+\bottomrule
+\multicolumn{8}{l}{Note: $\beta_0$  and $\gamma_0$ inform us about the type and strength }\\
+\multicolumn{8}{l}{of interpersonal behavior at the average self-efficacy. $\beta_1$ }\\
+\multicolumn{8}{l}{and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
+\multicolumn{8}{l}{type and strength of interpersonal behavior. $\alpha$ is the } \\
+\multicolumn{8}{l}{shape parameter of the distribution of the strength }\\
+\multicolumn{8}{l}{of interpersonal bahavior. $\kappa$ and $\lambda$ are the concentration}\\
+\multicolumn{8}{l}{and skewness parameters for the distribution of the type}\\
+\multicolumn{8}{l}{of interpersonal behavior.}\\
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{\label{tab:estCLGPNM}Results, cross-validation mean and standard deviation, for the GPN-SSN model}
+\centering
+\begin{tabular}[t]{lllllll}
+\toprule
+\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
+\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
+  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
+\midrule
+$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.11) & 1.75 (0.09) & 2.50 (0.11)\\
+$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.34 (0.06) & 1.10 (0.05) & 1.57 (0.06)\\
+$\beta_{0_y}$ & 0.33 (0.01) & 0.30 (0.30) & 0.36 (0.01) & 0.33 (0.01) & 0.30 (0.01) & 0.36 (0.01)\\
+\addlinespace
+$\beta_{1_s^{I}}$ & 0.09 (0.01) & 0.05 (0.01) & 0.13 (0.01) & 0.60 (0.06) & 0.33 (0.05) & 0.90 (0.06)\\
+$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.04) & 0.66 (0.04)\\
+$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
+\addlinespace
+$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.44 (0.15) & 1.72 (0.07) & 3.46 (0.14)\\
+$\sum_{s_{2,2}}$ & 0.02 (0.00) & 0.02 (0.00) & 0.03 (0.00) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
+$\sum_{y_{3,3}}$ & 0.03 (0.00) & 0.02 (0.02) & 0.04 (0.00) & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00)\\
+$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.08 (0.06) & -0.20 (0.06) & 0.34 (0.06)\\
+$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.00) & 0.32 (0.01)\\
+$\sum_{sy_{2,3}}$ & 0.01 (0.00) & 0.01 (0.01) & 0.02 (0.00) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
+$\lambda$ & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01) & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01)\\
+\bottomrule
+\multicolumn{7}{l}{Note: $\beta_{0_s^{I}}$, $\beta_{0_s^{II}}$ and $\beta_{0_y}$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{at the average self-efficacy. $\beta_{1_s^{I}}$, $\beta_{1_s^{II}}$ and $\beta_{1_y}$ inform us about the effect of self-efficacy }\\
+\multicolumn{7}{l}{on the type and strength of interpersonal behavior. $\sum_{s_{1,1}}$, $\sum_{s_{1,2}}$, $\sum_{s_{2,2}}$, $\sum_{y_{3,3}}$, $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ }\\
+\multicolumn{7}{l}{are elements of the variance-covariance matrix of which $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ inform us about}\\
+\multicolumn{7}{l}{the dependence between the type and strength of interpersonal behavior.}\\
+\multicolumn{7}{l}{$\lambda$ is the skewness parameter of the distribution of the strengths of interpersonal behavior.}\\
+\end{tabular}
+\end{table}
+
+\subsubsection{Average type and strength of interpersonal behavior}
+
+The parameters \(\gamma_0\) in the CL-PN, CL-GPN and Abe-Ley model and
+the parameter \(\beta_{0_y}\) in the GPN-SSN model inform us about the
+strength of interpersonal behavior at the average self-efficacy. For the
+CL-PN, CL-GPN and GPN-SSN model the parameters are estimated at 0.38,
+0.37 and 0.30 respectively and are a direct prediction of the strength
+of interpersonal behavior at the average self-efficacy. The estimate for
+the GPN-SSN model is notably lower and likely to be caused by its skewed
+distribution for the strengths of interpersonal behavior. In the Abe-Ley
+model, \(\gamma_0\) influences the shape parameter of the distribution
+of the strength of interpersonal behavior and does not directly estimate
+the average strength. Instead we can use the survival function to say
+something about the probability of having a certain strength of
+interpersonal behavior. Figure \ref{reglineweib} shows this function for
+several values of self-efficacy. We look at the survival function at
+average values of self-efficacy. Note that this function is the average
+of all survival functions for observations that fall within 1 standard
+deviation of the mean. The survival function indicates that the
+probability of having a low strength of interpersonal behavior is higher
+than having a high strength. We however can not make any direct
+statement about the estimated strength using the Abe-Ley model.\newline
+\indent The parameters \(\beta_0^{I}\), \(\beta_0^{II}\), \(\beta_0\),
+\(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) inform us about the
+type of interpersonal behavior at the average self-efficacy for the
+CL-PN, CL-GPN, Abe-Ley and GPN-SSN model respectively. For the CL-PN,
+CL-GPN and GPN-SSN model we need to combine the estimates for the
+underlying bivariate components \(\{I, II\}\) into one circular estimate
+using the double arctangent
+function\footnote{\(\mbox{atan2}(\beta_0^{II}, \beta_0^{I})\) or
+\(\mbox{atan2}(\beta_{0_{s^{II}}}, \beta_{0_{s^{I}}})\)}. Table
+\ref{tab:means} shows that these circular estimates are similar for the
+three models at 32.29\(^\circ\), 33.70\(^\circ\) and 35.53\(^\circ\). In
+the Abe-Ley model the type of interpersonal behavior at the average
+self-efficacy is estimated at 0.36 radians or 20.63\(^\circ\).
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
+\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
+\label{reglineweib}
+\end{figure}
+
+\begin{table}
+\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean (at SE = 0) in the CL-PN, CL-GPN and GPN-SSN models}
 \centering
@@
   \midrule
-Cl-PN & 32.29 & 24.81 & 39.71\\
+CL-PN & 32.29 & 24.81 & 39.71\\
 CL-GPN & 33.70 & 26.72 & 41.15\\
-GPN-SSN & 35.30 & 28.31 & 43.10\\
+GPN-SSN & 35.53 & 28.40 & 43.30\\
 \bottomrule
-\multicolumn{4}{l}{\textsuperscript{a} Note that these means are based on}\\
-\multicolumn{4}{l}{the posterior predictive}\\
-\multicolumn{4}{l}{distribution for the intercepts}\\
-\multicolumn{4}{l}{following (Wang \& Gelfand, 2013).}\\
+\multicolumn{4}{l}{Note that these means are based on}\\
+\multicolumn{4}{l}{their posterior predictive distribution }\\
+\multicolumn{4}{l}{following (Wang and Gelfand, 2013)}\\
 \end{tabular}
@@
 
-First recall the regression equations predicting the linear outcome
-\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i\]
-\noindent and circular outcome
-\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
-  \mu_{i}^{I}  \\
-\mu_{i}^{II}
- \end{pmatrix}=\begin{pmatrix}
-  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
-  \beta_0^{II} + \beta_1^{II}\text{SE}_i
- \end{pmatrix}.\] \noindent  For both models, \(\hat{y}\) is the
-predicted strength of interpersonal behavior, \(\hat{\boldsymbol{\mu}}\)
-is the predicted mean vector of the type of interpersonal behavior and
-the \(\gamma\)'s and \(\beta\)'s are intercepts and regression
-coefficients.\newline \indent Table \ref{tab:estCLGPN} shows the results
-for the modified CL-PN and CL-GPN models fit to the teacher dataset. The
-estimates in this table are the averages and standard deviations of the
-10 cross-validation estimates. We show both the estimated posterior mode
-and the 95\% highest posterior density (HPD) interval for each
-parameter. The posterior estimates for the \(\beta\)'s and \(\gamma\)'s
-are quite similar for both models. We also see this similarity when we
-look at the predicted circular and linear means. The predicted linear
-mean is equal to \(\gamma_0\) which is equal to 0.38 and 0.37 in the
-CL-PN and CL-GPN models, respectively. The predicted circular means can
-be computed from \(\beta_0^{I}\) and \(\beta_0^{II}\) using a double
-arctangent function \(\mbox{atan2}(\beta_0^{II}, \beta_0^{I})\), see
-\eqref{PredVal}. If we do this at each iteration of the MCMC sampler we
-get the posterior distribution of the circular mean in both models.
-Table \ref{tab:means} shows the posterior mode of the estimated circular
-means, which are very similar (32.29\(^\circ\) and
-33.54\(^\circ\)).\newline
-\indent Next we investigate the effect of self-efficacy. In the CL-PN
-model we can transform the regression parameters on the two components
-\(I\) and \(II\) to one circular regression coefficient \(b_c\) using
-the methods described by Cremers, Mulder, \& Klugkist (2018b). The
-dashed line in Figure \ref{regline} shows the predicted regression line
-from the CL-PN model. Here \(b_c\) is the slope of this line at the
-inflection point (black square) and its posterior mode is estimated at
-1.67 (-24.66,
-29.33)\footnote{Note that this is a linear approximation to the
-circular regression line representing the slope at a specific point. Therefore
-it is possible for the HPD interval to be wider than $2\pi$. In this case the
-interval is much wider and covers 0, indicating there is no evidence to reject
-the null-hypothesis of no effect.}. Even though the HPD of \(b_c\)
-includes 0, indicating there is no evidence to reject the null-hypotheis
-of no effect, we continue with its interpretation for eductional
-purposes. The interpretation of \(b_c\) is that at the inflection point
-an increase of 1 unit in self-efficacy leads to an increase of
-\(1.67*(180/\pi) = 95.68^\circ\) in the type of interpersonal
-behavior..However, as we can see in Figure \ref{regline} the inflection
-point lies almost outside the range of the actual data. Instead of
-looking at the inflection point we might compute the slope of the
-regression line at the average self-efficacy. This parameter we call the
-slope at the mean (\(SAM\)) (Cremers et al., 2018b) and it is estimated
-at 0.02 (0.01; 0.04) for our data. This means that at the average
-self-efficacy an increase of 1 unit only leads to an increase of
-\(0.02*(180/\pi) = 1.15^\circ\) in the type of interpersonal behavior.
-The HPD of the \(SAM\) does not include 0 which means that the effect at
-the average self-efficacy can be distinguished from 0.\newline
-\indent In the CL-GPN model we cannot compute circular regression
-coefficients such as the \(b_c\) and \(SAM\) computed above due to the
-fact that not only the mean vector \(\boldsymbol{\mu}\) but also the
-covariance matrix \(\boldsymbol{\Sigma}\) influences the predicted value
-on the circle. Instead, we will compute posterior predictive
+\subsubsection{The effect of self-efficacy}
+
+The parameters \(\gamma_1\) in the CL-PN, CL-GPN, Abe-Ley model and
+\(\beta_{1_y}\) in the GPN-SSN model inform us about the effect of
+self-efficacy on the strength of interpersonal behavior. For the CL-PN,
+CL-GPN and GPN-SSN model the parameters are estimated at 0.03, 0.03 and
+0.09 respectively and are a direct estimate of the effect of
+self-efficacy on the strength of interpersonal behavior, \emph{i.e.} an
+increase of 1 unit in self-efficacy leads to an increase of 0.09 units
+in the strength of interpersonal behavior according to the GPN-SSN
+model. These estimates are however quite small and only different from 0
+(the HPD interval does not contain 0) in the GPN-SSN model. It is hard
+to say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to
+base our conclusions on. The model CL-GPN and CL-PN fit the linear
+outcome best according to the model fit in Table \ref{tab:ModelFit}. In
+these models the linear outcome has a symmetric distribution whereas in
+the GPN-SSN the distribution of the linear outcome is skewed. However,
+the effect of self-efficacy is different from 0 only in the GPN-SSN
+model which does not seem to match with its lower model fit.\newline
+\indent In the Abe-Ley model, \(\gamma_1\) influences the shape
+parameter of the distribution of the strength of interpersonal behavior
+and does not directly estimate the effect of self-efficacy. Instead we
+can use the survival function to say something about the probability of
+having a certain strength of interpersonal behavior for different values
+of self-efficacy. Figure \ref{reglineweib} shows this function for low,
+average and high values of self-efficacy (as defined in Figure
+\ref{dataplot}). This function indicates that the effect of
+self-efficacy on the strength of interpersonal behavior is not linear.
+The probability of having a higher strength of interpersonal behavior is
+highest for low self-efficacy and lowest for average
+self-efficacy.\newline \indent The parameters \(\beta_1^{I}\),
+\(\beta_1^{II}\), \(\beta_1\), \(\beta_{1_{s^{I}}}\) and
+\(\beta_{1_{s^{II}}}\) inform us about the effect of self-efficacy on
+the type of interpersonal behavior in the CL-PN, CL-GPN, Abe-Ley and
+GPN-SSN model respectively. For the CL-PN and Abe-Ley model we have
+drawn the circular regression lines for this effect in Figure
+\ref{regline} (see the description of the CL-PN and CL-GPN models). For
+the CL-PN model the inflection point is indicated with a square in
+Figure \ref{regline}. The inflection point for the Abe-Ley model falls
+outside the bounds of the plot and is therefore not displayed. The slope
+at the inflection point, \(b_c\), for the CL-PN model is computed by
+using methods from Cremers et al. (2018b) and is equal to 1.67 (-24.66,
+29.33)\footnote{Note that this is a linear approximation
+to the circular regression line representing the slope at a specific point.
+Therefore it is possible for the HPD interval to be wider than $2\pi$. In this
+case the interval is much wider and covers 0, indicating there is no evidence
+for an effect.} The parameter \(\beta_1\) is the slope at the inflection
+point for the Abe-Ley model and is equal to -0.03. Even though these
+slopes are quite different, the regression lines in Figure \ref{regline}
+are quite similar in the data range. Both the regression line of the
+Abe-Ley model and the CL-PN model show slopes that are not very steep in
+the range of the data indicating that the effect of self-efficacy on the
+type of interpersonal behavior is not large. \newline \indent In the
+CL-GPN and GPN-SSN model we cannot compute circular regression
+coefficients due to the fact that not only the mean vector of the GPN
+distribution but also the covariance matrix influences the predicted
+value on the circle. Instead, we will compute posterior predictive
 distributions for the predicted circular outcome of individuals scoring
@@
 \(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\).
-Note that we display the modes and HPD intervals for the posterior
-predictive distributions on the interval \([0^\circ, 360^\circ)\). Note
-that \(44.49^\circ = 404.49^\circ\) due to the periodicity of a circular
-variable. The posterior mode estimate of \(215.74^\circ\) thus lies
-within its HPD interval \((147.36^\circ, \: 44.49^\circ)\).The HPD
-intervals of the three posterior predictive distributions overlap. Had
-they not overlapped we could have concluded that as the self-efficacy
-increases, the score of the teacher on the IPC moves counterclockwise.
-\newline
-\indent The effect of self-efficacy on the strength of interpersonal
-behavior is quantified by \(\gamma_1\) in both the CL-PN and CL-GPN
-models. This parameter can be interpreted in the same way as in a usual
-regression model. The HPD interval of \(\gamma_1\) includes 0 which
-means that there is not enough evidence for an effect of self-efficacy
-on the strength of interpersonal behavior. Note that the lower bounds of
-the HPD interval of this parameter are however very close to
-zero.\newline
-\indent  The relation between the circular and linear outcome, that is,
-between type of interpersonal behavior and strength of interpersonal
-behavior, is described by the parameters \(\gamma_{\cos}\) and
-\(\gamma_{\sin}\). The HPD interval of \(\gamma_{\cos}\) does not
-include 0 for both the CL-PN and CL-GPN models, meaning that the cosine
-component of the type of interpersonal behavior has an effect on the
-strength of interpersonal behavior. In the teacher data the sine and
-cosine components have a substantive meaning. In this case the Communion
-(cosine) component of the IPC positively effects the strength of a
-teachers' type of interpersonal behavior, in plain words: teachers
-exhibiting interpersonal behavior types with higher communion scores
-(e.g., `helpful' and `understanding' in Figure 2) are stronger in their
-behavior.
+\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\)
+for the CL-GPN model. Note that we display the modes and HPD intervals
+for the posterior predictive distributions on the interval
+\([0^\circ, 360^\circ)\) and that \(44.49^\circ = 404.49^\circ\) due to
+the periodicity of a circular variable. The posterior mode estimate of
+\(215.74^\circ\) thus lies within its HPD interval
+\((147.36^\circ, \: 44.49^\circ)\). For the GPN-SSN model the modes and
+95\% HPD intervals of the posterior predictive distributiona are
+\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.12^\circ, \: 72.02^\circ)\),
+\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
+\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\).
+For both the CL-GPN and GPN-SSN model the HPD intervals of the mode of
+the posterior predictive intervals of individuals scoring the minimum,
+median and maximum self-efficacy overlap. This indicates that the effect
+of self-efficacy, if there is any, on the type of interpersonal behavior
+a teacher shows is not expected to be strong. Had the HPD intervals not
+overlapped we could have concluded that as the self-efficacy increases,
+the score of the teacher on the IPC moves counterclockwise.\newline
 
@@
 
-\subsubsection{The modified Abe-Ley model} \label{resAL}
-
-\begin{table}
-
-\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation, for the modified Abe-Ley model}
-\centering
-\begin{tabular}[t]{ll}
-\toprule
-\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
-\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
-$\eta_0$ & 0.36 (0.02)\\
-$\eta_1$ & -0.03 (0.01)\\
-$\nu_0$ & 1.17 (0.02)\\
-$\nu_1$ & 0.04 (0.02)\\
-$\alpha$ & 3.66 (0.12)\\
-$\kappa$ & 1.51 (0.08)\\
-$\lambda$ & 0.70 (0.05)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-First recall the regression equations predicting the linear outcome
-\[\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i)\]
-
-\noindent and circular outcome
-\[\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).\]
-
-\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type of
-interpersonal behavior and \(\hat{\beta}\) is the predicted scale
-parameter of the strength of interpersonal behavior, while the
-\(\eta\)'s and \(\nu\)'s are regression parameters.\newline
-\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley
-model fit to the teacher dataset. The estimates in this table are the
-averages and standard deviations of the 10 cross-validation estimates.
-The estimate of the circular mean at an average self-efficacy,
-\(\eta_0\), equals 0.36 radians or 20.62\(^\circ\). For the Abe-Ley
-model we can also investigate the effect of self-efficacy on the type of
-interpersonal behavior. The solid line in Figure \ref{regline} shows
-this effect. Unlike for the regression line of the CL-PN model, the
-inflection point of the regression line of the Abe-Ley model lies
-outside the x-range of the figure. \newline
-\indent For the Abe-Ley model the conditional distribution for the
-linear outcome is Weibull. This means that we can use methods from
-survival analsis to interpret the effect of self-efficacy. In survival
-analysis a `survival' function is used in which time is plotted against
-the probability of survival of subjects suffering from a specific
-medical condition. In our data however we plot the strength on the IPC
-against the probability of a teacher having such a strength. This
-probability is computed using the `survival-function'
-\(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}})\)
-with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
-\ref{reglineweib}, we plot the survival function for the minimum, median
-and maximum value of self-efficacy. We conclude that stronger
-interpersonal behaviors are less probable. We also see that the relation
-between self-efficacy and the strength on the IPC is not linear; the
-probability of having a stronger interpersonal behavior is higher for
-both the minimum and maximum self-efficacy compared to a median
-self-efficacy score. Note however that the circular outcome also
-influences the survival function. The relation between the type of
-interpersonal behavior and the strength of interpersonal behavior may
-thus influence the shape of the survival function.\newline
-
-\begin{figure}
-\centering
-\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
-\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
-\label{reglineweib}
-\end{figure}
-
-\subsubsection{The modified GPN-SSN model} \label{resGPNSSN}
-
-\begin{table}
-\caption{\label{tab:estCLGPNM}Results, cross-validation mean and standard deviation, for the GPN-SSN model}
-\centering
-\begin{tabular}[t]{lllllll}
-\toprule
-\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
-\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
-  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
-\midrule
-$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.10) & 1.75 (0.09) & 2.50 (0.11)\\
-$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.33 (0.07) & 1.10 (0.05) & 1.57 (0.06)\\
-$\beta_{0_y}$ & 0.33 (0.01) & 0.30 (0.30) & 0.36 (0.01) & 0.33 (0.01) & 0.30 (0.01) & 0.36 (0.01)\\
-$\beta_{1_s^{I}}$ & 0.09 (0.01) & 0.05 (0.01) & 0.13 (0.01) & 0.60 (0.06) & 0.33 (0.05) & 0.90 (0.06)\\
-$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.03) & 0.66 (0.04)\\
-\addlinespace
-$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
-$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.43 (0.14) & 1.72 (0.07) & 3.46 (0.13)\\
-$\sum_{s_{2,2}}$ & 0.02 (0.00) & 0.02 (0.00) & 0.03 (0.00) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
-$\sum_{y_{3,3}}$ & 0.03 (0.00) & 0.02 (0.02) & 0.04 (0.00) & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00)\\
-$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.07 (0.06) & -0.20 (0.06) & 0.35 (0.06)\\
-\addlinespace
-$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.01) & 0.32 (0.01)\\
-$\sum_{sy_{2,3}}$ & 0.01 (0.00) & 0.01 (0.01) & 0.02 (0.00) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
-$\lambda$ & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01) & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-First recall the regression equation predicting the circular and linear
-outcome:
-
-\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +\boldsymbol{\beta}_1\text{SE}_i,\]
-
-\noindent where
-\(\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t\),
-\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
-and
-\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
-The parameters \(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\),
-\(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) are the intercepts and
-regression coefficients for the circular outcome and \(\beta_{0_y}\) and
-\(\beta_{1_y}\) are the intercept and regression coefficient for the
-linear outcome.\newline
-\indent Table \ref{tab:estCLGPNM} shows the results for the modified
-GPN-SSN model fit to the teacher dataset. The estimates in this table
-are the averages and standard deviations of the 10 cross-validation
-estimates. We show both the estimated posterior mode and the 95\%
-highest posterior density (HPD) interval for each parameter. The
-predicted circular means can be computed from \(\beta_{0_{s^{I}}}\) and
-\(\beta_{0_{s^{II}}}\) in a similar fashion as for the CL-PN and CL-GPN
-models. Table \ref{tab:means} shows the posterior mode of the estimated
-circular mean, which equals 35.30\(^\circ\) and hence is very similar to
-those of the CL-PN and CL-GPN models.\newline
-\indent For the same reason as in the CL-GPN model we cannot compute
-circular regression coefficients for the effect of self-efficacy on the
-type of interpersonal behavior such as the \(b_c\) and \(SAM\). Instead,
-we will again compute posterior predictive distributions for the
-predicted circular outcome of individuals scoring the minimum, maximum
-and median self-efficacy. The modes and 95\% HPD intervals of these
-posterior predictive distributions are
-\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.11^\circ, \: 72.02^\circ)\),
-\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\).
-On a circle the HPD intervals of the three posterior predictive
-distributions overlap. Had they not overlapped we could have concluded
-that as the self-efficacy increases, the score of the teacher on the IPC
-moves counterclockwise. The effect of self-efficacy on the strength of
-interpersonal behavior is quantified by \(\beta_{1_y}\), and we learn
-that for a 1 unit increase in self-efficacy the strength of
-interpersonal behavior increases by 0.09. The average strength is
-quantified by \(\beta_{0_y}\) and equals 0.33.\newline
-\indent To investigate the association between the linear and circular
-outcome we look at the covariances between the linear outcome and the
+\subsubsection{Dependence between type and strength of interpersonal behavior.}
+
+The relation between the type and strength of interpersonal behavior in
+the CL-PN and CL-GPN model, is described by the parameters
+\(\gamma_{\cos}\) and \(\gamma_{\sin}\). The HPD interval of
+\(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
+models, meaning that the cosine component of the type of interpersonal
+behavior has an effect on the strength of interpersonal
+behavior.\newline \indent In the teacher data the sine and cosine
+components have a substantive meaning. This is illustrated in Figure
+\ref{QTI}. In a unit circle the horizontal axis (Communion) represents
+the cosine of and the vertical axis (Agency) represents the sine of an
+angle. For the teacher data this means that the Communion (cosine)
+dimension of the IPC positively effects the strength of a teachers' type
+of interpersonal behavior, in plain words: teachers exhibiting
+interpersonal behavior types with higher communion scores (e.g.,
+\enquote{helpful} and \enquote{understanding} in Figure 2) are stronger
+in their interpersonal behavior.\newline  \indent In the GPN-SSN model
+the dependence between the type and strengths of interpersonal behavior
+is modelled through the covariances between the linear outcome and the
 sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
@@
 \(\sum_{sy_{1,3}} = 0.23\), are different from zero, but the one of the
-cosine component is larger. This means the correlation with the
-Communion component is larger and that teachers scoring both high on
-Communion and Agency show stronger behavior. This is a slightly
-different conclusion from the one in the CL-PN and CL-GPN models.
+cosine component, and thus the correlation with the Communion dimension,
+is larger. This mean that teachers scoring both high on Communion and
+Agency show stronger behavior. Together with the results from the CL-PN
+and CL-GPN models in the previous paragraph this translates to the
+conclusion that teachers with the strongest interpersonal behavior have
+a type of interpersonal behavior between 0\(^\circ\) and 90\(^\circ\).
+To get these scores on the circle both the Agency and the Communion
+score of a Teacher have to be positive (see \eqref{PredVal}). This
+corresponds to the pattern observed in the teacher data in Figure
+\ref{QTI}. At a strength of 0.4 and up we see that the scores on the
+circle range on average between 0\(^\circ\) and 100\(^\circ\).
 
@@
 
-In this section we will assess the overall fit of the cylindrical models
-to the data via the PLSL criterion described in Section \ref{Modelfit}.
-Table \ref{tab:ModelFit} shows the values of this criterion for the
-linear and circular outcomes of the four models.\newline
-\indent The CL-PN and CL-GPN models have the best out-of-sample
-predictive performance for the linear outcome. They show roughly the
-same performance because they model the linear outcome in the same way.
-Only the value of \(r\) in \eqref{circlinlink} differs. We should note
-that even though the predictive performance of the Abe-Ley model for the
-linear outcome is worst on average, the standard deviation of the
-cross-validation estimates is rather large. This means that in some
-samples, the Abe-Ley model shows a lower PLSL value than the average of
-25.49\newline
-\indent The Abe-Ley model has the best out-of-sample predictive
-performance for the circular outcome. This would suggest that for the
-circular variable a slightly skewed distribution fits best. However,
-both the GPN-SSN and the CL-GPN models fit much worse even though the
-distribution for the circular outcome in these models can also take a
-skewed shape. It should be noted that the standard deviation of the
-cross-validation estimates is rather large for the Abe-Ley and the
-CL-GPN model. It is possible that these large standard deviations for
-the PLSL are caused by the fact that they are computed for a relatively
-small sample size, but this does not explain why the PLSL has a large
-standard deviation for only a few cylindrical models and not for all.
+Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the
+linear and circular outcomes of the four cylindrical models fit to the
+teacher data.\newline \indent The CL-PN and CL-GPN models have the best
+out-of-sample predictive performance for the linear outcome. They show
+roughly the same performance because they model the linear outcome in
+the same way. We should note that even though the predictive performance
+of the Abe-Ley model for the linear outcome is worst on average, the
+standard deviation of the cross-validation estimates is rather large.
+This means that in some samples, the Abe-Ley model shows a lower PLSL
+value than the average of 25.49\newline \indent The Abe-Ley model has
+the best out-of-sample predictive performance for the circular outcome.
+This would suggest that for the circular variable a slightly skewed
+distribution fits best. However, both the GPN-SSN and the CL-GPN models
+fit much worse even though the distribution for the circular outcome in
+these models can also take a skewed shape. It should be noted that the
+standard deviation of the cross-validation estimates is rather large for
+the Abe-Ley and the CL-GPN model. It is possible that these large
+standard deviations for the PLSL are caused by the fact that they are
+computed for a relatively small sample size, but this does not explain
+why the PLSL has a large standard deviation for only a few cylindrical
+models and not for all.\newline \indent In this situation, where one
+model fits the linear outcome best and another one fits the circular
+outcome best, it is hard to determine which model we should choose. In
+this case the results for the CL-PN /CL-GPN and Abe-Ley model are quite
+different regarding the effect of self-efficacy on the linear outcome
+(strength of interpersonal behavior). Because the Abe-Ley fit for the
+linear part is worst we would choose to trust the results for the CL-PN
+and CL-GPN model here. For the circular part however the results of the
+CL-PN/CL-GPN model do not differ as much from the Abe-Ley model and we
+reach the same conclusion for both models, namely that the effect of
+self-efficacy on type of interpersonal behavior is not very strong.
+Therefore we would prefer the CL-PN/CL-GPN models in this case because
+where it matters in terms of interpretation (the linear part) they show
+better fit.
 
@@
 methods can be used to analyze a dataset on the interpersonal behavior
-of teachers. In this final section we will comment on the differences
-between these models, the results from the analysis of the teacher data
-and how cylindrical models improve the analysis of such cylindrical
-data.\newline
-\indent In terms of interpretability, the CL-PN and Abe-Ley models
-perform best. In the CL-GPN and GPN-SSN models the interpretation of the
-parameters of the circular outcome component is not straightforward, if
-at all possible. This is caused by the fact that in addition to the mean
-vector the covariance matrix of the GPN distribution affects the
-location of the circular data, making it difficult to compute regression
-coefficients on the circle. Wang \& Gelfand (2013) state that Monte
-Carlo integration can be used to compute a circular mean and variance
-for the GPN distribution. In future research, this solution might be
-applied to the methods of Cremers et al. (2018b) in order to compute
-circular coefficients for GPN models.\newline
-\indent In terms of flexibility the GPN-SSN model scores best. Multiple
-linear and circular outcomes can be included and we can thus apply the
-model to multivariate cylindrical data. In addition the GPN-SSN, the
-CL-GPN and CL-PN models are extendable to a mixed-effects structure and
-can thus also be fit to longitudinal data (see Nuñez-Antonio \&
-Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al. (2016) for
-hierarchical/mixed-effects models for the PN and GPN distributions
-respectively). For the Abe-Ley model this may also be possible but has
-not been done in previous research for the conditional distribution of
-its circular outcome (sine-skewed von Mises). Concerning asymmetry, both
-the GPN-SSN as well as the Abe-Ley model allow for non-symmetrical
-shapes of the distributions of both the linear and circular outcome,
-while the CL-GPN model permits an asymmetric circular outcome.\newline
-\indent To investigate model fit for the teacher data, we assessed
-out-of-sample predictive performance for both the linear and circular
-outcome. The CL-PN and CL-GPN models have the best fit for the linear
-outcome while the Abe-Ley model fits best the circular outcome.
-Differences in fit, in addition to being a result of different
-distributional assumptions, may also be caused by the way in which the
-relation between the linear and circular outcome is modeled. Whereas in
-both the Abe-Ley and GPN-SSN models the distribution of the linear
-outcome is conditioned on the circular outcome and vice versa, the
-distribution of the circular outcome in the CL-PN and CL-GPN models is
-independent of the linear outcome. In these models the circular outcome
-is regressed onto the linear outcome.\newline
-\indent The four cylindrical models that were modified to the regression
-context in this paper are not the only cylindrical distributions
-available from the literature. Other interesting cylindrical
-distributions have been introduced by Fernández-Durán (2007), Kato \&
-Shimizu (2008) and Sugasawa (2015) (for more references we refer to
-Chapter 2 of Ley \& Verdebout (2017)). In the present study we have
-decided not to include these distributions for reasons of space,
-complexity of the models and ease of implementing a regression
-structure. In future research however it would be interesting to
-investigate other types of cylindrical distributions as well in order to
-compare the interpretability, flexibility and model fit to the models
-developed in the present study.\newline
-\indent We conclude the paper on a general note regarding cylindrical
-models. They offer new insights into data of a cylindrical nature in
-psychology. Concerning the example used in this paper, the advantage of
-cylindrical data analysis is that we were able to analyze all circular
-and linear information in the data simultaneously. In previous research,
-the two components of the interpersonal circumplex (\emph{i.e.}, Agency
-and Communion) were analyzed separately. Such an approach also provides
+of teachers. In this final section we will first comment on what
+researchers can gain by using cylindrical models for the teacher data.
+Subsequently we will comment on the differences between the cylindrical
+models that were introduced in this paper.\newline \indent Concerning
+the teacher data, the advantage of cylindrical data analysis is that we
+were able to analyze the information about the type and strength of
+interpersonal behavior simultaneously. In previous research, the two
+components of the interpersonal circumplex (\emph{i.e.}, Agency and
+Communion) were analyzed separately. Such an approach also provides
 information about the strength of teachers' score on Agency and
 Communion, yet a large portion of information about the combination of
-Agency and Communion, which describes the kind of behavior that is
+Agency and Communion, which describes the type of behavior that is
 observed, gets lost. A first solution to include both dimensions as a
@@
 strength of interpersonal behavior and how these are influenced by
-teachers' self-efficacy in classroom management. The results of the
-present study therefore provide an answer to an often stated problem in
-data analysis of interpersonal circumplex data.
+teachers' self-efficacy in classroom management. Although we do not find
+any strong effects of self-efficacy on either the type or strength of
+behavior, the four cylindrical models do provide a way of analyzing and
+interpreting this effect. This is beneficial for future research in
+which we may want to investigate the effect of further covariates on
+data from the circumplex. Furthermore, in addition to being able to
+assess the influence of covariates, the cylindrical models also provide
+information about the dependence between the type and strength of
+interpersonal behavior. We found that stronger behavior is associated
+with those types of behavior that show higher scores on the Communion
+dimension (and also on the Agency dimension). This translates to
+stronger behavior being shown by teachers whose type of interpersonal
+behavior ranges between 0\(^\circ\) and 90\(^\circ\), the
+\enquote{helpful} and \enquote{directing} subtypes according to Figure
+\ref{QTI}.
+\textcolor{red}{Is dit nieuw/zijn er papers die hier iets over
+zeggen?}\newline \indent As mentioned in the introducion data from the
+interpersonal circumplex is not the only type of cylindrical data that
+occurs in psychology. The methods presented in this paper are also of
+use for research on human navigation and eye-tracking research.
+Furthermore, even though cylindrical models are already used in fields
+outside of psychology, the addition of a regression structure to the
+models is of use in these fields as well. \newline \indent In terms of
+interpretability, the CL-PN and Abe-Ley models perform best out of the
+four cylindrical models. In the CL-GPN and GPN-SSN models the
+interpretation of the parameters of the circular outcome component is
+not straightforward, if at all possible. This is caused by the fact that
+in addition to the mean vector the covariance matrix of the GPN
+distribution affects the location of the circular data, making it
+difficult to compute regression coefficients on the circle. Wang and
+Gelfand (2013) state that Monte Carlo integration can be used to compute
+a circular mean and variance for the GPN distribution. In future
+research, this solution might be applied to the methods of Cremers et
+al. (2018b) in order to compute circular coefficients for GPN
+models.\newline \indent In terms of flexibility the GPN-SSN model scores
+best. Multiple linear and circular outcomes can be included and we can
+thus apply the model to multivariate cylindrical data. In addition the
+GPN-SSN, the CL-GPN and CL-PN models are extendable to a mixed-effects
+structure and can thus also be fit to longitudinal data (see
+Nuñez-Antonio and Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et
+al. (2016) for hierarchical/mixed-effects models for the PN and GPN
+distributions respectively). For the Abe-Ley model this may also be
+possible but has not been done in previous research for the conditional
+distribution of its circular outcome (sine-skewed von Mises). Concerning
+asymmetry, both the GPN-SSN as well as the Abe-Ley model allow for
+non-symmetrical shapes of the distributions of both the linear and
+circular outcome, while the CL-GPN model permits an asymmetric circular
+outcome.\newline \indent The four cylindrical models that were modified
+to the regression context in this paper are not the only cylindrical
+distributions available from the literature. Other interesting
+cylindrical distributions have been introduced by Fernández-Durán
+(2007), Kato and Shimizu (2008) and Sugasawa (2015) (for more references
+we refer to Chapter 2 of Ley and Verdebout (2017)). In the present study
+we have decided not to include these distributions for reasons of space,
+complexity of the models and ease of implementing a regression
+structure. In future research however it would be interesting to
+investigate other types of cylindrical distributions as well in order to
+compare the interpretability, flexibility and model fit to the models
+developed in the present study.\newline
 
 \section*{References}
-\DeclareRobustCommand{\VANDER}[3]{#3}
-
-\hypertarget{ref-abe2017tractable}{}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-abe2017tractable}{}%
 Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
@@
 
-\hypertarget{ref-chrastil2017rotational}{}
+\leavevmode\hypertarget{ref-chrastil2017rotational}{}%
 Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
@@
 
-\hypertarget{ref-Claessens2016side}{}
+\leavevmode\hypertarget{ref-Claessens2016side}{}%
 Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
@@
 
-\hypertarget{ref-Cremers2018Assessing}{}
+\leavevmode\hypertarget{ref-Cremers2018Assessing}{}%
 Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
@@
 
-\hypertarget{ref-CremersMulderKlugkist2017}{}
+\leavevmode\hypertarget{ref-CremersMulderKlugkist2017}{}%
 Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
@@
 
-\hypertarget{ref-fernandez2007models}{}
+\leavevmode\hypertarget{ref-fernandez2007models}{}%
 Fernández-Durán, J. (2007). Models for circular--linear and
@@
 
-\hypertarget{ref-fisher1995statistical}{}
+\leavevmode\hypertarget{ref-fisher1995statistical}{}%
 Fisher, N. I. (1995). \emph{Statistical analysis of circular data}.
@@
 
-\hypertarget{ref-garcia2014test}{}
+\leavevmode\hypertarget{ref-garcia2014test}{}%
 García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
@@
 
-\hypertarget{ref-garcia2013exploring}{}
+\leavevmode\hypertarget{ref-garcia2013exploring}{}%
 García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
@@
 
-\hypertarget{ref-BDA}{}
+\leavevmode\hypertarget{ref-BDA}{}%
 Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
@@
 
-\hypertarget{ref-gneiting2007strictly}{}
+\leavevmode\hypertarget{ref-gneiting2007strictly}{}%
 Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
@@
 
-\hypertarget{ref-gurtman2009exploring}{}
+\leavevmode\hypertarget{ref-gurtman2009exploring}{}%
 Gurtman, M. B. (2009). Exploring personality with the interpersonal
@@
 
-\hypertarget{ref-hernandez2016general}{}
+\leavevmode\hypertarget{ref-gurtman2011reasoning}{}%
+Gurtman, M. B. (2011). Handbook of interpersonal psychology. In L. M.
+Horowitz \& S. Strack (Eds.), (pp. 299--311). New York: Wiley.
+
+\leavevmode\hypertarget{ref-hernandez2016general}{}%
 Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
@@
 
-\hypertarget{ref-horowitz2010handbook}{}
+\leavevmode\hypertarget{ref-horowitz2010handbook}{}%
 Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
@@
 
-\hypertarget{ref-jammalamadaka2001topics}{}
+\leavevmode\hypertarget{ref-jammalamadaka2001topics}{}%
 Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
@@
 
-\hypertarget{ref-johnson1978some}{}
+\leavevmode\hypertarget{ref-johnson1978some}{}%
 Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
@@
 
-\hypertarget{ref-kato2008dependent}{}
+\leavevmode\hypertarget{ref-kato2008dependent}{}%
 Kato, S., \& Shimizu, K. (2008). Dependent models for observations which
@@
 
-\hypertarget{ref-lagona2015hidden}{}
+\leavevmode\hypertarget{ref-lagona2015hidden}{}%
 Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
@@
 
-\hypertarget{ref-ley2017modern}{}
+\leavevmode\hypertarget{ref-leary1957}{}%
+Leary, T. (1957). \emph{An interpersonal diagnosis of personality}. New
+York: Ronald Press Company.
+
+\leavevmode\hypertarget{ref-ley2017modern}{}%
 Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
@@
 
-\hypertarget{ref-mardia2000directional}{}
+\leavevmode\hypertarget{ref-locke2007selfefficacy}{}%
+Locke, K. D., \& Sadler, P. (2007). Self-efficacy, values, and
+complementarity in dyadic interactions: Integrating interpersonal and
+social-cognitive theory. \emph{Personality and Social Psychology
+Bulletin}, \emph{33}, 94--109.
+doi:\href{https://doi.org/10.1177/0146167206293375}{10.1177/0146167206293375}
+
+\leavevmode\hypertarget{ref-mardia2000directional}{}%
 Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
@@
 
-\hypertarget{ref-mardia1978model}{}
+\leavevmode\hypertarget{ref-mardia1978model}{}%
 Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical
@@
 
-\hypertarget{ref-mastrantonio2018joint}{}
+\leavevmode\hypertarget{ref-mastrantonio2018joint}{}%
 Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
@@
 
-\hypertarget{ref-mastrantonio2015bayesian}{}
+\leavevmode\hypertarget{ref-mastrantonio2015bayesian}{}%
 Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
@@
 
-\hypertarget{ref-nunez2014bayesian}{}
+\leavevmode\hypertarget{ref-nunez2014bayesian}{}%
 Nuñez-Antonio, G., \& Gutiérrez-Peña, E. (2014). A Bayesian model for
@@
 
-\hypertarget{ref-nunez2011bayesian}{}
+\leavevmode\hypertarget{ref-nunez2011bayesian}{}%
 Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
@@
 
-\hypertarget{ref-olmos2012extension}{}
-Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
-extension of the half-normal distribution. \emph{Statistical Papers},
-\emph{53}(4), 875--886.
-doi:\href{https://doi.org/10.1007/s00362-011-0391-4}{10.1007/s00362-011-0391-4}
-
-\hypertarget{ref-pennings2018interpersonal}{}
+\leavevmode\hypertarget{ref-pennings2017phd}{}%
+Pennings, H. J. M. (2017a). \emph{Interpersonal dynamics in
+teacher-student interactions and relationships} (PhD thesis). Enschede:
+Ipskamp drukkers.
+
+\leavevmode\hypertarget{ref-pennings2017complexity}{}%
+Pennings, H. J. M. (2017b). Using a complexity approach to study the
+interpersonal dynamics in teacher-student interactions: A case study of
+two teachers. \emph{Complicity: An International Journal of Complexity
+and Education}, \emph{14}(2), 88--103.
+
+\leavevmode\hypertarget{ref-pennings2018interpersonal}{}%
 Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C.,
@@
 
-\hypertarget{ref-rayner200935th}{}
+\leavevmode\hypertarget{ref-plutchik1997general}{}%
+Plutchik, R. (1997). Circumplex models of personality and emotions. In
+C. Plutchik Robert. (Ed.), (pp. 17--45). Washington, DC, US: American
+Psychological Association.
+
+\leavevmode\hypertarget{ref-rayner200935th}{}%
 Rayner, K. (2009). The 35th Sir Frederick Bartlett lecture: Eye
@@
 
-\hypertarget{ref-sahu2003new}{}
+\leavevmode\hypertarget{ref-russell1980circumplex}{}%
+Russell, J. A. (1980). A circumplex model of affect. \emph{Journal of
+Personality and Social Psychology}, \emph{39}(6), 1161--1178.
+doi:\href{https://doi.org/10.1037/h0077714}{10.1037/h0077714}
+
+\leavevmode\hypertarget{ref-sahu2003new}{}%
 Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
@@
 
-\hypertarget{ref-sugasawa2015flexible}{}
+\leavevmode\hypertarget{ref-schwartz1992values}{}%
+Schwartz, S. H. (1992). Advances in experimental social psychology. In
+M. Zanna (Ed.), (Vol. 25, pp. 1--65). San Diego, CA: Academic Press.
+
+\leavevmode\hypertarget{ref-sugasawa2015flexible}{}%
 Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the
@@
 
-\hypertarget{ref-wang2012directional}{}
+\leavevmode\hypertarget{ref-wang2012directional}{}%
 Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
@@
 
-\hypertarget{refs}{}
-\hypertarget{ref-vanderWant2015role}{}
+\leavevmode\hypertarget{ref-vanderWant2015role}{}%
 \VANDER{Want}{Van der}{van der} Want, A. C. (2015). \emph{Teachers'
@@
 
-\hypertarget{ref-wubbels2006interpersonal}{}
+\leavevmode\hypertarget{ref-want2018selfefficacy}{}%
+\VANDER{Want}{Van der}{van der} Want, A. C., Den Brok, P., Beijaard, D.,
+Brekelmans, M., Claessens, L. C., \& Pennings, H. J. M. (n.d.). The
+relation between teachers' interpersonal role identity and their
+self-efficacy, burnout and work engagement. \emph{Professional
+Development in Education}.
+doi:\href{https://doi.org/10.1080/19415257.2018.1511453}{10.1080/19415257.2018.1511453}
+
+\leavevmode\hypertarget{ref-wiggins1996history}{}%
+Wiggins, J. (1996). An informal history of the interpersonal circumplex
+tradition. \emph{Journal of Personality Assessment}, \emph{66}(2),
+217--233.
+
+\leavevmode\hypertarget{ref-wright2009integrating}{}%
+Wright, A. G., Pincus, A. L., Conroy, D. E., \& Hilsenroth, M. J.
+(2009). Integrating methods to optimize circumplex description and
+comparison of groups. \emph{Journal of Personality Assessment},
+\emph{91}(4), 311--322.
+doi:\href{https://doi.org/10.1080/00223890902935696}{10.1080/00223890902935696}
+
+\leavevmode\hypertarget{ref-wubbels2006interpersonal}{}%
 Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
-An interpersonal perspective on classroom management in secondary
-classrooms in the Netherlands. In C. Evertson \& C. S. Weinstein (Eds.),
-\emph{Handbook of classroom management: Research, practice, and
-contemporary issues} (pp. 1161--1191). Malwah, NJ: Lawrence Erlbaum
-Associates.
+Handbook of classroom management: Research practice and contemporary
+issues. In C. Evertson \& C. S. Weinstein (Eds.), \emph{Handbook of
+classroom management: Research, practice, and contemporary issues} (pp.
+1161--1191). Malwah, NJ: Lawrence Erlbaum Associates.
 

### 2019-01-28 Manuscript/Plots/circregline.tex
Update manuscript

% Created by tikzDevice version 0.11 on 2019-01-14 18:24:38
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,162.86) --
	( 65.83,162.75) --
	( 69.16,162.64) --
	( 72.48,162.53) --
	( 75.81,162.41) --
	( 79.13,162.29) --
	( 82.46,162.17) --
	( 85.79,162.05) --
	( 89.11,161.92) --
	( 92.44,161.79) --
	( 95.77,161.66) --
	( 99.09,161.52) --
	(102.42,161.39) --
	(105.74,161.25) --
	(109.07,161.10) --
	(112.40,160.96) --
	(115.72,160.81) --
	(119.05,160.65) --
	(122.37,160.50) --
	(125.70,160.34) --
	(129.03,160.17) --
	(132.35,160.00) --
	(135.68,159.83) --
	(139.01,159.66) --
	(142.33,159.48) --
	(145.66,159.29) --
	(148.98,159.11) --
	(152.31,158.92) --
	(155.64,158.72) --
	(158.96,158.52) --
	(162.29,158.31) --
	(165.61,158.10) --
	(168.94,157.89) --
	(172.27,157.67) --
	(175.59,157.44) --
	(178.92,157.21) --
	(182.24,156.97) --
	(185.57,156.73) --
	(188.90,156.48) --
	(192.22,156.23) --
	(195.55,155.97) --
	(198.88,155.71) --
	(202.20,155.44) --
	(205.53,155.16) --
	(208.85,154.88) --
	(212.18,154.59) --
	(215.51,154.29) --
	(218.83,153.99) --
	(222.16,153.68) --
	(225.48,153.37) --
	(228.81,153.05) --
	(232.14,152.72) --
	(235.46,152.38) --
	(238.79,152.04) --
	(242.11,151.70) --
	(245.44,151.34) --
	(248.77,150.98) --
	(252.09,150.62) --
	(255.42,150.24) --
	(258.75,149.87) --
	(262.07,149.48) --
	(265.40,149.09) --
	(268.72,148.70) --
	(272.05,148.30) --
	(275.38,147.89) --
	(278.70,147.48) --
	(282.03,147.07) --
	(285.35,146.65) --
	(288.68,146.23) --
	(292.01,145.80) --
	(295.33,145.37) --
	(298.66,144.94) --
	(301.98,144.51) --
	(305.31,144.08) --
	(308.64,143.64) --
	(311.96,143.20) --
	(315.29,142.77) --
	(318.61,142.33) --
	(321.94,141.90) --
	(325.27,141.46) --
	(328.59,141.03) --
	(331.92,140.60) --
	(335.25,140.18) --
	(338.57,139.75) --
	(341.90,139.33) --
	(345.22,138.91) --
	(348.55,138.50) --
	(351.88,138.09) --
	(355.20,137.69) --
	(358.53,137.29) --
	(361.85,136.89) --
	(365.18,136.50) --
	(368.51,136.12) --
	(371.83,135.74) --
	(375.16,135.37) --
	(378.48,135.01) --
	(381.81,134.65) --
	(385.14,134.30) --
	(388.46,133.95) --
	(391.79,133.61) --
	(395.12,133.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {Predictor};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {Circular Outcome};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.77, 61.20) -- ( 95.77, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.03, 61.20) -- (129.03, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.29, 61.20) -- (162.29, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.55, 61.20) -- (195.55, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.81, 61.20) -- (228.81, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.07, 61.20) -- (262.07, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.33, 61.20) -- (295.33, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.59, 61.20) -- (328.59, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.85, 61.20) -- (361.85, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 65.95, 49.20) {-5};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 99.21, 49.20) {-4};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (132.47, 49.20) {-3};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (165.73, 49.20) {-2};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (198.99, 49.20) {-1};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (232.25, 49.20) {0};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (265.51, 49.20) {1};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (298.78, 49.20) {2};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (332.04, 49.20) {3};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (365.30, 49.20) {4};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at (398.56, 49.20) {5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 73.08) -- ( 49.20,191.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 73.08) -- ( 43.20, 73.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 92.88) -- ( 43.20, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,112.67) -- ( 43.20,112.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,132.47) -- ( 43.20,132.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,152.27) -- ( 43.20,152.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,172.07) -- ( 43.20,172.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,191.87) -- ( 43.20,191.87);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20, 69.64) {-180};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20, 89.43) {-120};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,109.23) {-60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,129.03) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,148.83) {60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,168.62) {120};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,188.42) {180};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.02,158.22) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.85,159.69) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.61,151.61) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.98,148.39) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.79,155.75) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.86,157.39) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.16,155.69) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.66,171.40) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.04,158.89) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,150.31) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.55,161.39) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.63,160.68) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.52,155.42) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.11,155.35) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.48,157.22) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.64,163.99) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.89,164.74) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (165.20,164.22) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.49,164.42) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.99,160.22) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.12,163.21) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.64,161.85) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.92,171.78) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.25,165.01) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.43,151.21) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.92,167.18) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.59,162.38) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.33,161.26) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.60,150.13) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.14,166.79) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.94,161.45) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.22,157.60) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.57,153.31) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.86,169.09) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.65,150.41) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.15,172.02) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.67,150.12) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.20,143.33) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.89,157.54) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.66,159.68) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.36,163.30) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.12,157.27) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.57,175.94) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.36,159.37) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.14,158.59) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.39,158.49) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.67,146.33) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.06,163.79) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.81,161.49) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.25,166.02) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.90,165.72) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.59,157.86) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.12,162.13) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.59,158.84) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.66,155.93) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.84,178.08) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.54,155.29) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,147.04) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.85,151.23) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.94,158.72) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.33,159.32) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.88,160.11) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.29,167.10) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.90,170.31) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (183.37,163.10) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.59,154.67) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.97,165.31) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.72,166.89) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.41,159.43) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.55,174.39) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.62,152.08) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.11,154.69) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.82,144.67) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.42,154.56) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.54,161.75) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.53,156.02) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.38,158.24) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.53,162.99) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.19,151.16) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.74,168.56) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.69,137.52) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.75,150.38) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.77,160.14) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.15,158.52) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.03,161.22) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.77,164.82) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.56,159.90) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.54,156.59) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.41,153.00) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.75,142.23) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.42,150.06) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.58,157.89) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.28,151.60) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.36,159.44) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.36,163.61) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.85,161.55) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.09,164.98) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.39,146.50) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.11,160.63) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.66,161.53) circle (  1.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.22,151.12) rectangle (227.72,155.62);
\end{scope}
\end{tikzpicture}

\end{document}

### 2019-01-28 Manuscript/Plots/dataplot.tex
Update manuscript

--- 2018-07-02 Manuscript/Plots/dataplot.tex
@@
-% Created by tikzDevice version 0.11 on 2018-07-02 13:49:21
+% Created by tikzDevice version 0.11 on 2019-01-11 18:28:29
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (180.78,184.51) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.44,124.69) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.74,137.28) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.12,176.35) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.66,196.39) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12,196.40) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12,137.73) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.43,184.51) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.16,142.64) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.57,196.39) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (300.32,152.77) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.60,102.73) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.19,147.01) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.97,142.83) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.65,144.99) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.34,118.50) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.42,169.19) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.48,141.73) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.48,102.73) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.38,141.15) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.22,165.95) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.28,141.82) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.93,142.83) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.14,140.40) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.46,141.89) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.70,137.38) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.02,142.94) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.66,159.30) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11,146.53) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.37,148.62) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.75,141.10) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.43,147.05) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.15,118.50) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.02,141.70) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.52,151.58) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.92,158.12) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.84,144.85) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.87,141.70) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.27,153.34) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.76,150.56) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.88,142.14) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.84,140.49) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.54,152.49) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.26,140.05) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.52, 81.92) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.08,152.67) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.61, 81.92) circle (  2.25);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (124.33,154.75) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.17,144.16) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.57,143.29) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.76, 67.58) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.47,142.73) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.69,142.68) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.27,134.29) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.30, 91.85) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (216.19,168.30) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.40,130.36) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.80,143.75) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.59,181.23) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.48,144.16) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.57,139.37) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.81, 67.58) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.52,144.62) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.72,146.45) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.23,120.54) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.32,126.33) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.63,138.45) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.75,145.53) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.10,129.29) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.13,144.94) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.60,109.16) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.64,140.58) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.19,130.05) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.61,143.46) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.97,136.21) circle (  2.25);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.81,142.68) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.69,150.83) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.60,144.82) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.73, 91.85) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.57,145.06) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.02,152.93) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.91,140.50) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.97,136.97) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.97,143.78) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.23,133.68) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.56,146.95) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.41,132.80) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.22,152.07) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.25,142.45) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.31,169.02) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.34,140.18) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.72,137.19) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.96,130.36) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.40, 97.42) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.24,163.45) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.79,189.67) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.06,142.56) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.51, 77.91) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.26,181.23) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.12,147.47) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.20,136.51) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.21,139.37) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.13,145.17) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.60,138.81) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.48,152.66) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.86,142.00) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.37,144.48) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.01,141.17) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.61,147.04) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (297.18,134.55) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.04,137.00) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 78.13) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.82,140.18) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.81,144.62) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.90,149.59) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.74,147.67) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.65,149.14) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.88,120.54) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.38,138.45) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.32,142.69) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.63,153.78) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.51,135.76) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.31,129.29) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.59,138.14) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.83,109.16) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.55,155.42) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.83,143.47) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.63,156.34) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.48,136.09) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.44,147.25) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.65,180.17) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.56,156.05) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.32,146.74) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.44,151.99) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.48,130.94) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.20,164.64) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.19,157.43) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (335.89,140.67) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.89,159.57) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.22,139.35) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.95,144.27) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.21,153.02) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.28,151.62) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.36,130.05) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.85,136.21) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.12,143.35) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.83,133.88) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.86,145.10) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.61,143.21) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.15,141.54) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.05,138.82) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.50,143.52) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.10,141.05) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.67,137.53) circle (  2.25);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.51,138.43) circle (  2.25);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.65,143.52) circle (  2.25);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.75, 61.20) -- (395.39, 61.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.22, 61.20) -- (403.43, 61.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.75, 61.20) -- ( 70.75, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.22, 61.20) -- ( 74.22, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.13, 61.20) -- (117.13, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.06, 61.20) -- (140.06, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.51, 61.20) -- (163.51, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.90, 61.20) -- (205.90, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.88, 61.20) -- (209.88, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.75, 61.20) -- (271.75, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.26, 61.20) -- (256.26, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.59, 61.20) -- (337.59, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.64, 61.20) -- (302.64, 55.20);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.43, 61.20) -- (403.43, 55.20);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (349.01, 61.20) -- (349.01, 55.20);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 74.22, 39.60) {0.1};
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.39, 61.20) -- (395.39, 55.20);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (140.06, 39.60) {0.2};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 70.75, 39.60) {0.1};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (205.90, 39.60) {0.3};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (117.13, 39.60) {0.2};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (271.75, 39.60) {0.4};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (163.51, 39.60) {0.3};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (337.59, 39.60) {0.5};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (209.88, 39.60) {0.4};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (256.26, 39.60) {0.5};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (302.64, 39.60) {0.6};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (349.01, 39.60) {0.7};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.39, 39.60) {0.8};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (403.43, 39.60) {0.6};
 
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {IPC location};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {IPC};
+\end{scope}
+\begin{scope}
+\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
+\definecolor{drawColor}{RGB}{0,0,0}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (307.75,165.69) --
+	(310.78,160.44) --
+	(304.72,160.44) --
+	(307.75,165.69);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.23,156.95) --
+	(282.26,151.70) --
+	(276.20,151.70) --
+	(279.23,156.95);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (311.37,146.41) --
+	(314.40,141.16) --
+	(308.34,141.16) --
+	(311.37,146.41);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.67,147.56) --
+	(433.62,142.44);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62,142.31) --
+	(427.64,142.31) --
+	(430.67,147.56);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.81,136.83) --
+	(233.84,131.58) --
+	(227.78,131.58) --
+	(230.81,136.83);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.06,136.63) --
+	(220.09,131.39) --
+	(214.03,131.39) --
+	(217.06,136.63);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.44,149.83) --
+	(366.47,144.58) --
+	(360.41,144.58) --
+	(363.44,149.83);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.18,153.72) --
+	(344.21,148.47) --
+	(338.15,148.47) --
+	(341.18,153.72);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.02,150.26) --
+	(293.05,145.01) --
+	(286.99,145.01) --
+	(290.02,150.26);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.60,143.72) --
+	(390.63,138.47) --
+	(384.57,138.47) --
+	(387.60,143.72);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.74,147.88) --
+	(421.77,142.63) --
+	(415.71,142.63) --
+	(418.74,147.88);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.53,148.36) --
+	(406.56,143.11) --
+	(400.50,143.11) --
+	(403.53,148.36);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.84,140.78) --
+	(180.87,135.53) --
+	(174.81,135.53) --
+	(177.84,140.78);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.24,146.14) --
+	(397.27,140.89) --
+	(391.21,140.89) --
+	(394.24,146.14);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (400.15,156.27) --
+	(403.18,151.02) --
+	(397.12,151.02) --
+	(400.15,156.27);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.82,150.51) --
+	(334.85,145.26) --
+	(328.79,145.26) --
+	(331.82,150.51);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.13,148.48) --
+	(368.16,143.24) --
+	(362.10,143.24) --
+	(365.13,148.48);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.12,172.69) --
+	(291.15,167.44) --
+	(285.09,167.44) --
+	(288.12,172.69);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.28,145.23) --
+	(396.31,139.98) --
+	(390.25,139.98) --
+	(393.28,145.23);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.18,144.65) --
+	(227.21,139.40) --
+	(221.15,139.40) --
+	(224.18,144.65);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.43,169.45) --
+	(218.46,164.20) --
+	(212.40,164.20) --
+	(215.43,169.45);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (297.87,145.31) --
+	(300.90,140.07) --
+	(294.84,140.07) --
+	(297.87,145.31);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.53,140.88) --
+	(335.56,135.63) --
+	(329.50,135.63) --
+	(332.53,140.88);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.19,146.44) --
+	(279.22,141.20) --
+	(273.16,141.20) --
+	(276.19,146.44);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.12,162.80) --
+	(280.15,157.56) --
+	(274.09,157.56) --
+	(277.12,162.80);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (196.83,150.02) --
+	(199.86,144.78) --
+	(193.80,144.78) --
+	(196.83,150.02);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.84,152.12) --
+	(347.87,146.87) --
+	(341.81,146.87) --
+	(344.84,152.12);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.46,150.55) --
+	(285.49,145.30) --
+	(279.43,145.30) --
+	(282.46,150.55);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.91,145.20) --
+	(301.94,139.96) --
+	(295.88,139.96) --
+	(298.91,145.20);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.76,155.08) --
+	(282.79,149.83) --
+	(276.73,149.83) --
+	(279.76,155.08);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.67,161.61) --
+	(179.70,156.37) --
+	(173.64,156.37) --
+	(176.67,161.61);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.10,145.20) --
+	(330.13,139.95) --
+	(324.07,139.95) --
+	(327.10,145.20);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.43,156.84) --
+	(140.46,151.59) --
+	(134.39,151.59) --
+	(137.43,156.84);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.48,154.06) --
+	(267.51,148.81) --
+	(261.45,148.81) --
+	(264.48,154.06);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.90,145.64) --
+	(227.93,140.39) --
+	(221.87,140.39) --
+	(224.90,145.64);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (405.14,143.98) --
+	(408.17,138.74) --
+	(402.11,138.74) --
+	(405.14,143.98);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.56,155.98) --
+	(339.59,150.74) --
+	(333.53,150.74) --
+	(336.56,155.98);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.03,143.55) --
+	(315.06,138.30) --
+	(309.00,138.30) --
+	(312.03,143.55);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.78,146.79) --
+	(380.81,141.54) --
+	(374.75,141.54) --
+	(377.78,146.79);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (216.93,137.79) --
+	(219.96,132.54) --
+	(213.90,132.54) --
+	(216.93,137.79);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.70,171.80) --
+	(283.73,166.55) --
+	(277.67,166.55) --
+	(280.70,171.80);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.15,147.25) --
+	(398.18,142.00) --
+	(392.12,142.00) --
+	(395.15,147.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.08,149.95) --
+	(114.11,144.70) --
+	(108.05,144.70) --
+	(111.08,149.95);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.58,129.83) --
+	(103.61,124.58) --
+	( 97.55,124.58) --
+	(100.58,129.83);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.78,149.03) --
+	(376.81,143.79) --
+	(370.75,143.79) --
+	(373.78,149.03);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.05,148.44) --
+	(373.08,143.19) --
+	(367.02,143.19) --
+	(370.05,148.44);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (375.00,146.96) --
+	(378.03,141.71) --
+	(371.97,141.71) --
+	(375.00,146.96);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.59,154.33) --
+	(325.62,149.08) --
+	(319.56,149.08) --
+	(322.59,154.33);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.84,148.32) --
+	(424.87,143.07) --
+	(418.81,143.07) --
+	(421.84,148.32);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.19,148.56) --
+	(223.22,143.31) --
+	(217.16,143.31) --
+	(220.19,148.56);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.92,156.43) --
+	(301.95,151.18) --
+	(295.89,151.18) --
+	(298.92,156.43);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.18,144.00) --
+	(418.21,138.75) --
+	(412.15,138.75) --
+	(415.18,144.00);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.84,140.47) --
+	(372.87,135.22) --
+	(366.81,135.22) --
+	(369.84,140.47);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.40,147.27) --
+	(327.43,142.03) --
+	(321.37,142.03) --
+	(324.40,147.27);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.94,137.18) --
+	(253.97,131.93) --
+	(247.91,131.93) --
+	(250.94,137.18);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.83,150.45) --
+	(326.86,145.20) --
+	(320.80,145.20) --
+	(323.83,150.45);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.32,155.57) --
+	(238.35,150.32) --
+	(232.29,150.32) --
+	(235.32,155.57);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (296.40,145.95) --
+	(299.43,140.70) --
+	(293.37,140.70) --
+	(296.40,145.95);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.22,172.52) --
+	(295.25,167.27) --
+	(289.19,167.27) --
+	(292.22,172.52);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.60,143.68) --
+	(333.63,138.43) --
+	(327.57,138.43) --
+	(330.60,143.68);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.62,140.69) --
+	(396.65,135.44) --
+	(390.59,135.44) --
+	(393.62,140.69);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.33,100.91) --
+	(207.36, 95.67) --
+	(201.30, 95.67) --
+	(204.33,100.91);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.65,166.95) --
+	(188.68,161.70) --
+	(182.62,161.70) --
+	(185.65,166.95);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.10,193.17) --
+	(381.13,187.92) --
+	(375.07,187.92) --
+	(378.10,193.17);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.89,146.06) --
+	(223.92,140.81) --
+	(217.86,140.81) --
+	(220.89,146.06);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (123.56, 81.40) --
+	(126.59, 76.16) --
+	(120.53, 76.16) --
+	(123.56, 81.40);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.61,150.97) --
+	(327.64,145.72) --
+	(321.58,145.72) --
+	(324.61,150.97);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.54,140.00) --
+	(357.57,134.76) --
+	(351.51,134.76) --
+	(354.54,140.00);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.67,148.67) --
+	(344.70,143.42) --
+	(338.64,143.42) --
+	(341.67,148.67);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.26,142.31) --
+	(426.29,137.06) --
+	(420.23,137.06) --
+	(423.26,142.31);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.10,156.16) --
+	(240.13,150.91) --
+	(234.07,150.91) --
+	(237.10,156.16);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.69,138.05) --
+	(398.72,132.80) --
+	(392.66,132.80) --
+	(395.69,138.05);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.01, 81.63) --
+	(108.04, 76.38) --
+	(101.98, 76.38) --
+	(105.01, 81.63);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.12,143.68) --
+	(337.15,138.43) --
+	(331.09,138.43) --
+	(334.12,143.68);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.80,153.09) --
+	(247.83,147.84) --
+	(241.77,147.84) --
+	(244.80,153.09);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.71,151.17) --
+	(271.74,145.92) --
+	(265.68,145.92) --
+	(268.71,151.17);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (294.13,152.64) --
+	(297.16,147.39) --
+	(291.10,147.39) --
+	(294.13,152.64);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (289.41,146.19) --
+	(292.44,140.94) --
+	(286.38,140.94) --
+	(289.41,146.19);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.59,157.28) --
+	(173.62,152.03) --
+	(167.56,152.03) --
+	(170.59,157.28);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.72,141.64) --
+	(258.75,136.39) --
+	(252.69,136.39) --
+	(255.72,141.64);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.29,158.92) --
+	(203.32,153.67) --
+	(197.26,153.67) --
+	(200.29,158.92);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.31,159.84) --
+	(125.34,154.59) --
+	(119.28,154.59) --
+	(122.31,159.84);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.48,139.59) --
+	(109.51,134.34) --
+	(103.45,134.34) --
+	(106.48,139.59);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.80,150.75) --
+	(208.83,145.50) --
+	(202.77,145.50) --
+	(205.80,150.75);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.94,150.23) --
+	(344.97,144.99) --
+	(338.91,144.99) --
+	(341.94,150.23);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (285.37,134.44) --
+	(288.40,129.19) --
+	(282.34,129.19) --
+	(285.37,134.44);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.38,168.14) --
+	(316.41,162.89) --
+	(310.35,162.89) --
+	(313.38,168.14);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.95,160.93) --
+	(243.98,155.68) --
+	(237.92,155.68) --
+	(240.95,160.93);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.50,163.07) --
+	(270.53,157.82) --
+	(264.47,157.82) --
+	(267.50,163.07);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.30,142.85) --
+	(309.33,137.60) --
+	(303.27,137.60) --
+	(306.30,142.85);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62,142.52) --
+	(430.67,142.52) --
+	(433.62,147.64);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.28,155.12) --
+	(302.31,149.87) --
+	(296.25,149.87) --
+	(299.28,155.12);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (287.70,146.84) --
+	(290.73,141.60) --
+	(284.67,141.60) --
+	(287.70,146.84);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.71,137.38) --
+	(176.74,132.13) --
+	(170.68,132.13) --
+	(173.71,137.38);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.23,148.60) --
+	(283.26,143.35) --
+	(277.20,143.35) --
+	(280.23,148.60);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (375.00,146.71) --
+	(378.03,141.46) --
+	(371.97,141.46) --
+	(375.00,146.71);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (272.12,145.04) --
+	(275.15,139.79) --
+	(269.09,139.79) --
+	(272.12,145.04);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.89,144.55) --
+	(277.92,139.31) --
+	(271.86,139.31) --
+	(274.89,144.55);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.87,140.40) -- (285.23,140.40);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,137.22) -- (282.05,143.58);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.39,144.85) -- (223.75,144.85);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.57,141.67) -- (220.57,148.04);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.61,152.67) -- (243.98,152.67);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.80,149.49) -- (240.80,155.85);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.10,154.75) -- (153.47,154.75);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.28,151.57) -- (150.28,157.94);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.62,140.58) -- (346.99,140.58);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.80,137.40) -- (343.80,143.76);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.37,132.80) -- (380.74,132.80);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.55,129.61) -- (377.55,135.98);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.84,144.48) -- (433.21,144.48);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.03,141.30) -- (430.03,147.66);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.33,147.04) -- (342.69,147.04);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.51,143.86) -- (339.51,150.22);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.87,143.47) -- (330.23,143.47);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.05,140.29) -- (327.05,146.65);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (202.93,180.17) -- (209.30,180.17);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.11,176.99) -- (206.11,183.35);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.89,156.05) -- (216.26,156.05);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.08,152.87) -- (213.08,159.24);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.80,151.99) -- (135.17,151.99);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.98,148.81) -- (131.98,155.17);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.96,153.02) -- (285.33,153.02);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.15,149.84) -- (282.15,156.20);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.69,137.53) -- (391.05,137.53);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.87,134.35) -- (387.87,140.71);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.67, 92.98) circle (  2.25);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.67, 84.47) --
+	(316.70, 79.23) --
+	(310.64, 79.23) --
+	(313.67, 84.47);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.49, 68.98) -- (316.85, 68.98);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.67, 65.79) -- (313.67, 72.16);
+
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (322.67, 89.53) {low SE};
+
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (322.67, 77.53) {average SE};
+
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (322.67, 65.53) {high SE};
 \end{scope}

### 2019-01-28 Manuscript/Plots/plotprojecting.tex
Update manuscript

% Created by tikzDevice version 0.11 on 2019-01-11 12:07:34
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.70,157.66) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.71,147.24) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.38,157.59) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.80,158.39) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.60,130.97) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.64,143.95) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.34,137.33) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (102.34,184.94) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (124.92,132.59) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.92,131.10) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.85, 74.95) -- ( 77.05, 74.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.95, 72.85) -- ( 74.95, 77.05);

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (108.40, 74.95) --
	(108.34, 77.05) --
	(108.14, 79.14) --
	(107.81, 81.22) --
	(107.35, 83.27) --
	(106.77, 85.29) --
	(106.06, 87.26) --
	(105.22, 89.19) --
	(104.27, 91.07) --
	(103.20, 92.87) --
	(102.02, 94.61) --
	(100.73, 96.27) --
	( 99.34, 97.85) --
	( 97.85, 99.34) --
	( 96.27,100.73) --
	( 94.61,102.02) --
	( 92.87,103.20) --
	( 91.07,104.27) --
	( 89.19,105.22) --
	( 87.26,106.06) --
	( 85.29,106.77) --
	( 83.27,107.35) --
	( 81.22,107.81) --
	( 79.14,108.14) --
	( 77.05,108.34) --
	( 74.95,108.41) --
	( 72.85,108.34) --
	( 70.75,108.14) --
	( 68.68,107.81) --
	( 66.63,107.35) --
	( 64.61,106.77) --
	( 62.63,106.06) --
	( 60.70,105.22) --
	( 58.83,104.27) --
	( 57.02,103.20) --
	( 55.28,102.02) --
	( 53.62,100.73) --
	( 52.04, 99.34) --
	( 50.56, 97.85) --
	( 49.17, 96.27) --
	( 47.88, 94.61) --
	( 46.70, 92.87) --
	( 45.63, 91.07) --
	( 44.67, 89.19) --
	( 43.84, 87.26) --
	( 43.13, 85.29) --
	( 42.54, 83.27) --
	( 42.08, 81.22) --
	( 41.75, 79.14) --
	( 41.55, 77.05) --
	( 41.49, 74.95) --
	( 41.55, 72.85) --
	( 41.75, 70.75) --
	( 42.08, 68.68) --
	( 42.54, 66.63) --
	( 43.13, 64.61) --
	( 43.84, 62.63) --
	( 44.67, 60.70) --
	( 45.63, 58.83) --
	( 46.70, 57.02) --
	( 47.88, 55.28) --
	( 49.17, 53.62) --
	( 50.56, 52.04) --
	( 52.04, 50.56) --
	( 53.62, 49.17) --
	( 55.28, 47.88) --
	( 57.02, 46.70) --
	( 58.83, 45.63) --
	( 60.70, 44.67) --
	( 62.63, 43.84) --
	( 64.61, 43.13) --
	( 66.63, 42.54) --
	( 68.68, 42.08) --
	( 70.75, 41.75) --
	( 72.85, 41.55) --
	( 74.95, 41.49) --
	( 77.05, 41.55) --
	( 79.14, 41.75) --
	( 81.22, 42.08) --
	( 83.27, 42.54) --
	( 85.29, 43.13) --
	( 87.26, 43.84) --
	( 89.19, 44.67) --
	( 91.07, 45.63) --
	( 92.87, 46.70) --
	( 94.61, 47.88) --
	( 96.27, 49.17) --
	( 97.85, 50.56) --
	( 99.34, 52.04) --
	(100.73, 53.62) --
	(102.02, 55.28) --
	(103.20, 57.02) --
	(104.27, 58.83) --
	(105.22, 60.70) --
	(106.06, 62.63) --
	(106.77, 64.61) --
	(107.35, 66.63) --
	(107.81, 68.68) --
	(108.14, 70.75) --
	(108.34, 72.85) --
	(108.40, 74.95);
\end{scope}
\begin{scope}
\path[clip] (216.81,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.51,157.66) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.52,147.24) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (326.19,157.59) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.61,158.39) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (368.41,130.97) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.45,143.95) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (362.15,137.33) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.15,184.94) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.73,132.59) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.73,131.10) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (216.81,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (289.66, 74.95) -- (293.86, 74.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.76, 72.85) -- (291.76, 77.05);

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (325.21, 74.95) --
	(325.15, 77.05) --
	(324.95, 79.14) --
	(324.62, 81.22) --
	(324.16, 83.27) --
	(323.58, 85.29) --
	(322.87, 87.26) --
	(322.03, 89.19) --
	(321.08, 91.07) --
	(320.01, 92.87) --
	(318.83, 94.61) --
	(317.54, 96.27) --
	(316.15, 97.85) --
	(314.66, 99.34) --
	(313.08,100.73) --
	(311.42,102.02) --
	(309.68,103.20) --
	(307.88,104.27) --
	(306.00,105.22) --
	(304.07,106.06) --
	(302.10,106.77) --
	(300.08,107.35) --
	(298.03,107.81) --
	(295.95,108.14) --
	(293.86,108.34) --
	(291.76,108.41) --
	(289.66,108.34) --
	(287.56,108.14) --
	(285.49,107.81) --
	(283.44,107.35) --
	(281.42,106.77) --
	(279.44,106.06) --
	(277.51,105.22) --
	(275.64,104.27) --
	(273.83,103.20) --
	(272.09,102.02) --
	(270.43,100.73) --
	(268.85, 99.34) --
	(267.37, 97.85) --
	(265.98, 96.27) --
	(264.69, 94.61) --
	(263.51, 92.87) --
	(262.44, 91.07) --
	(261.48, 89.19) --
	(260.65, 87.26) --
	(259.94, 85.29) --
	(259.35, 83.27) --
	(258.89, 81.22) --
	(258.56, 79.14) --
	(258.36, 77.05) --
	(258.30, 74.95) --
	(258.36, 72.85) --
	(258.56, 70.75) --
	(258.89, 68.68) --
	(259.35, 66.63) --
	(259.94, 64.61) --
	(260.65, 62.63) --
	(261.48, 60.70) --
	(262.44, 58.83) --
	(263.51, 57.02) --
	(264.69, 55.28) --
	(265.98, 53.62) --
	(267.37, 52.04) --
	(268.85, 50.56) --
	(270.43, 49.17) --
	(272.09, 47.88) --
	(273.83, 46.70) --
	(275.64, 45.63) --
	(277.51, 44.67) --
	(279.44, 43.84) --
	(281.42, 43.13) --
	(283.44, 42.54) --
	(285.49, 42.08) --
	(287.56, 41.75) --
	(289.66, 41.55) --
	(291.76, 41.49) --
	(293.86, 41.55) --
	(295.95, 41.75) --
	(298.03, 42.08) --
	(300.08, 42.54) --
	(302.10, 43.13) --
	(304.07, 43.84) --
	(306.00, 44.67) --
	(307.88, 45.63) --
	(309.68, 46.70) --
	(311.42, 47.88) --
	(313.08, 49.17) --
	(314.66, 50.56) --
	(316.15, 52.04) --
	(317.54, 53.62) --
	(318.83, 55.28) --
	(320.01, 57.02) --
	(321.08, 58.83) --
	(322.03, 60.70) --
	(322.87, 62.63) --
	(323.58, 64.61) --
	(324.16, 66.63) --
	(324.62, 68.68) --
	(324.95, 70.75) --
	(325.15, 72.85) --
	(325.21, 74.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.69,102.54) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.58, 96.22) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.62,105.83) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.42,103.36) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.77, 94.69) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.34, 90.57) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.80, 97.14) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.84,107.41) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.67,100.23) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.91, 94.50) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.69,102.54) -- (348.51,157.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.58, 96.22) -- (379.52,147.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.62,105.83) -- (326.19,157.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.42,103.36) -- (343.61,158.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.77, 94.69) -- (368.41,130.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.34, 90.57) -- (422.45,143.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.80, 97.14) -- (362.15,137.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.84,107.41) -- (319.15,184.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.67,100.23) -- (341.73,132.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.91, 94.50) -- (369.73,131.10);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (310.69,102.54) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (317.58, 96.22) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (304.62,105.83) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (309.42,103.36) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (318.77, 94.69) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (321.34, 90.57) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (316.80, 97.14) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (299.84,107.41) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (313.67,100.23) -- (291.76, 74.95);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (318.91, 94.50) -- (291.76, 74.95);
\end{scope}
\begin{scope}
\path[clip] (433.62,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.50,102.54) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.39, 96.22) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.43,105.83) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.23,103.36) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.58, 94.69) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.15, 90.57) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (533.61, 97.14) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (516.65,107.41) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (530.48,100.23) circle (  1.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.72, 94.50) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (433.62,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.47, 74.95) -- (510.67, 74.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (508.57, 72.85) -- (508.57, 77.05);

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (542.02, 74.95) --
	(541.96, 77.05) --
	(541.76, 79.14) --
	(541.43, 81.22) --
	(540.97, 83.27) --
	(540.39, 85.29) --
	(539.68, 87.26) --
	(538.84, 89.19) --
	(537.89, 91.07) --
	(536.82, 92.87) --
	(535.64, 94.61) --
	(534.35, 96.27) --
	(532.96, 97.85) --
	(531.47, 99.34) --
	(529.89,100.73) --
	(528.23,102.02) --
	(526.49,103.20) --
	(524.69,104.27) --
	(522.81,105.22) --
	(520.88,106.06) --
	(518.91,106.77) --
	(516.89,107.35) --
	(514.84,107.81) --
	(512.76,108.14) --
	(510.67,108.34) --
	(508.57,108.41) --
	(506.47,108.34) --
	(504.37,108.14) --
	(502.30,107.81) --
	(500.25,107.35) --
	(498.23,106.77) --
	(496.25,106.06) --
	(494.32,105.22) --
	(492.45,104.27) --
	(490.64,103.20) --
	(488.90,102.02) --
	(487.24,100.73) --
	(485.66, 99.34) --
	(484.18, 97.85) --
	(482.79, 96.27) --
	(481.50, 94.61) --
	(480.32, 92.87) --
	(479.25, 91.07) --
	(478.29, 89.19) --
	(477.46, 87.26) --
	(476.75, 85.29) --
	(476.16, 83.27) --
	(475.70, 81.22) --
	(475.37, 79.14) --
	(475.17, 77.05) --
	(475.11, 74.95) --
	(475.17, 72.85) --
	(475.37, 70.75) --
	(475.70, 68.68) --
	(476.16, 66.63) --
	(476.75, 64.61) --
	(477.46, 62.63) --
	(478.29, 60.70) --
	(479.25, 58.83) --
	(480.32, 57.02) --
	(481.50, 55.28) --
	(482.79, 53.62) --
	(484.18, 52.04) --
	(485.66, 50.56) --
	(487.24, 49.17) --
	(488.90, 47.88) --
	(490.64, 46.70) --
	(492.45, 45.63) --
	(494.32, 44.67) --
	(496.25, 43.84) --
	(498.23, 43.13) --
	(500.25, 42.54) --
	(502.30, 42.08) --
	(504.37, 41.75) --
	(506.47, 41.55) --
	(508.57, 41.49) --
	(510.67, 41.55) --
	(512.76, 41.75) --
	(514.84, 42.08) --
	(516.89, 42.54) --
	(518.91, 43.13) --
	(520.88, 43.84) --
	(522.81, 44.67) --
	(524.69, 45.63) --
	(526.49, 46.70) --
	(528.23, 47.88) --
	(529.89, 49.17) --
	(531.47, 50.56) --
	(532.96, 52.04) --
	(534.35, 53.62) --
	(535.64, 55.28) --
	(536.82, 57.02) --
	(537.89, 58.83) --
	(538.84, 60.70) --
	(539.68, 62.63) --
	(540.39, 64.61) --
	(540.97, 66.63) --
	(541.43, 68.68) --
	(541.76, 70.75) --
	(541.96, 72.85) --
	(542.02, 74.95);
\end{scope}
\end{tikzpicture}

\end{document}

### 2019-01-28 Manuscript/Plots/survivaldiffSE.tex
Update manuscript

@@
-% Created by tikzDevice version 0.11 on 2018-09-18 14:43:22
+% Created by tikzDevice version 0.11 on 2019-01-11 16:36:22
 % !TEX encoding = UTF-8 Unicode
@@
 \path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,198.38) --
-	( 95.77,197.84) --
-	(112.40,196.49) --
-	(129.03,194.01) --
-	(145.66,190.14) --
-	(162.29,184.70) --
-	(178.92,177.62) --
-	(195.55,168.93) --
-	(212.18,158.83) --
-	(228.81,147.63) --
-	(245.44,135.76) --
-	(262.07,123.72) --
-	(278.70,112.05) --
-	(295.33,101.23) --
-	(311.96, 91.68) --
-	(328.59, 83.63) --
-	(345.22, 77.20) --
-	(361.85, 72.33) --
-	(378.49, 68.84) --
+	( 79.14,197.09) --
+	( 95.77,194.00) --
+	(112.40,189.53) --
+	(129.03,183.87) --
+	(145.66,177.12) --
+	(162.29,169.41) --
+	(178.92,160.85) --
+	(195.55,151.61) --
+	(212.18,141.85) --
+	(228.81,131.78) --
+	(245.44,121.67) --
+	(262.07,111.79) --
+	(278.70,102.46) --
+	(295.33, 93.93) --
+	(311.96, 86.45) --
+	(328.59, 80.16) --
+	(345.22, 75.11) --
+	(361.85, 71.25) --
+	(378.49, 68.43) --
 	(395.12, 66.48);
@@
 \path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,196.04) --
-	( 95.77,190.46) --
-	(112.40,182.63) --
-	(129.03,173.19) --
-	(145.66,162.73) --
-	(162.29,151.74) --
-	(178.92,140.68) --
-	(195.55,129.91) --
-	(212.18,119.75) --
-	(228.81,110.39) --
-	(245.44,101.97) --
-	(262.07, 94.57) --
-	(278.70, 88.20) --
-	(295.33, 82.82) --
-	(311.96, 78.36) --
-	(328.59, 74.74) --
-	(345.22, 71.84) --
-	(361.85, 69.57) --
-	(378.49, 67.81) --
+	( 79.14,196.69) --
+	( 95.77,192.37) --
+	(112.40,186.05) --
+	(129.03,178.13) --
+	(145.66,169.03) --
+	(162.29,159.10) --
+	(178.92,148.73) --
+	(195.55,138.24) --
+	(212.18,127.94) --
+	(228.81,118.07) --
+	(245.44,108.86) --
+	(262.07,100.46) --
+	(278.70, 92.98) --
+	(295.33, 86.50) --
+	(311.96, 81.02) --
+	(328.59, 76.51) --
+	(345.22, 72.91) --
+	(361.85, 70.12) --
+	(378.49, 68.02) --
 	(395.12, 66.48);
@@
 \path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 62.50,198.47) --
-	( 79.14,198.08) --
-	( 95.77,196.53) --
-	(112.40,193.46) --
-	(129.03,188.72) --
-	(145.66,182.27) --
-	(162.29,174.23) --
-	(178.92,164.82) --
-	(195.55,154.36) --
-	(212.18,143.26) --
-	(228.81,131.93) --
-	(245.44,120.83) --
-	(262.07,110.32) --
-	(278.70,100.74) --
-	(295.33, 92.29) --
-	(311.96, 85.12) --
-	(328.59, 79.23) --
-	(345.22, 74.56) --
-	(361.85, 70.98) --
-	(378.49, 68.35) --
-	(395.12, 66.47);
+	( 79.14,197.17) --
+	( 95.77,193.69) --
+	(112.40,188.31) --
+	(129.03,181.27) --
+	(145.66,172.87) --
+	(162.29,163.43) --
+	(178.92,153.30) --
+	(195.55,142.80) --
+	(212.18,132.28) --
+	(228.81,122.04) --
+	(245.44,112.32) --
+	(262.07,103.35) --
+	(278.70, 95.29) --
+	(295.33, 88.25) --
+	(311.96, 82.28) --
+	(328.59, 77.36) --
+	(345.22, 73.43) --
+	(361.85, 70.39) --
+	(378.49, 68.13) --
+	(395.12, 66.48);
 
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,183.02) {min(SE)};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,183.02) {low SE};
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {median(SE)};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {average SE};
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,159.02) {max(SE)};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,159.02) {high SE};
 \end{scope}

### 2019-01-28 Manuscript/Plots/whycirc.tex
Update manuscript

% Created by tikzDevice version 0.11 on 2019-01-11 12:07:28
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,289.08);
\begin{scope}
\path[clip] ( 72.00, 72.00) rectangle (324.61,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.91,116.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.59,126.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.77,124.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.96,110.91) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.11,162.36) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.51,157.79) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.82,113.20) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.36,177.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.61,128.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.59,151.72) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.35,139.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.58,162.59) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.32,154.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.52,135.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.61,175.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.70,161.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.96,124.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.51,161.61) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.35,140.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (123.60,154.28) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.71, 72.00) -- (318.91, 72.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.71, 72.00) -- ( 77.71, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.91, 72.00) -- (117.91, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.11, 72.00) -- (158.11, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.31, 72.00) -- (198.31, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.51, 72.00) -- (238.51, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.71, 72.00) -- (278.71, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.91, 72.00) -- (318.91, 66.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 77.71, 50.40) {-2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (117.91, 50.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (158.11, 50.40) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (198.31, 50.40) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (238.51, 50.40) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (278.71, 50.40) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (318.91, 50.40) {10};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,100.14) -- ( 72.00,260.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,100.14) -- ( 66.00,100.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,140.34) -- ( 66.00,140.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,180.54) -- ( 66.00,180.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,220.74) -- ( 66.00,220.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.00,260.94) -- ( 66.00,260.94);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 60.00, 96.01) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 60.00,136.21) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 60.00,176.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 60.00,216.61) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 60.00,256.81) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00, 12.00) rectangle (324.61,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.50] at (198.31, 14.40) {Communion};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.50] at ( 21.60,180.54) {Agency};
\end{scope}
\begin{scope}
\path[clip] ( 72.00, 72.00) rectangle (324.61,289.08);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (154.73,136.97) --
	(161.48,136.97) --
	(161.48,143.72) --
	(154.73,143.72) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.01,100.14) --
	(137.97,101.40) --
	(137.85,102.66) --
	(137.65,103.91) --
	(137.38,105.14) --
	(137.02,106.35) --
	(136.60,107.54) --
	(136.09,108.70) --
	(135.52,109.82) --
	(134.88,110.91) --
	(134.17,111.95) --
	(133.39,112.95) --
	(132.56,113.90) --
	(131.67,114.79) --
	(130.72,115.63) --
	(129.72,116.40) --
	(128.68,117.11) --
	(127.59,117.75) --
	(126.47,118.33) --
	(125.31,118.83) --
	(124.12,119.26) --
	(122.91,119.61) --
	(121.67,119.88) --
	(120.43,120.08) --
	(119.17,120.20) --
	(117.91,120.24) --
	(116.65,120.20) --
	(115.39,120.08) --
	(114.14,119.88) --
	(112.91,119.61) --
	(111.70,119.26) --
	(110.51,118.83) --
	(109.35,118.33) --
	(108.22,117.75) --
	(107.14,117.11) --
	(106.09,116.40) --
	(105.10,115.63) --
	(104.15,114.79) --
	(103.26,113.90) --
	(102.42,112.95) --
	(101.65,111.95) --
	(100.94,110.91) --
	(100.29,109.82) --
	( 99.72,108.70) --
	( 99.22,107.54) --
	( 98.79,106.35) --
	( 98.44,105.14) --
	( 98.16,103.91) --
	( 97.97,102.66) --
	( 97.85,101.40) --
	( 97.81,100.14) --
	( 97.85, 98.88) --
	( 97.97, 97.62) --
	( 98.16, 96.37) --
	( 98.44, 95.14) --
	( 98.79, 93.93) --
	( 99.22, 92.74) --
	( 99.72, 91.58) --
	(100.29, 90.46) --
	(100.94, 89.37) --
	(101.65, 88.33) --
	(102.42, 87.33) --
	(103.26, 86.38) --
	(104.15, 85.49) --
	(105.10, 84.65) --
	(106.09, 83.88) --
	(107.14, 83.17) --
	(108.22, 82.53) --
	(109.35, 81.95) --
	(110.51, 81.45) --
	(111.70, 81.02) --
	(112.91, 80.67) --
	(114.14, 80.40) --
	(115.39, 80.20) --
	(116.65, 80.08) --
	(117.91, 80.04) --
	(119.17, 80.08) --
	(120.43, 80.20) --
	(121.67, 80.40) --
	(122.91, 80.67) --
	(124.12, 81.02) --
	(125.31, 81.45) --
	(126.47, 81.95) --
	(127.59, 82.53) --
	(128.68, 83.17) --
	(129.72, 83.88) --
	(130.72, 84.65) --
	(131.67, 85.49) --
	(132.56, 86.38) --
	(133.39, 87.33) --
	(134.17, 88.33) --
	(134.88, 89.37) --
	(135.52, 90.46) --
	(136.09, 91.58) --
	(136.60, 92.74) --
	(137.02, 93.93) --
	(137.38, 95.14) --
	(137.65, 96.37) --
	(137.85, 97.62) --
	(137.97, 98.88) --
	(138.01,100.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.00,118.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.05,118.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.33,117.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.49,119.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.01,114.46) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.54,112.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.68,118.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.82,111.01) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.50,107.77) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.44,115.01) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.59,117.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.61,104.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.26,116.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.48,111.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.81,111.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.99, 99.41) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.31,117.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.57,113.88) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.79,118.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.90,102.24) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (396.61, 72.00) rectangle (649.23,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.72,156.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (533.41,166.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.59,164.69) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.77,151.11) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.92,202.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.33,197.99) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.63,153.40) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (532.18,217.30) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.43,168.86) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40,191.92) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.16,180.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.40,202.79) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.13,195.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (507.33,176.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.42,215.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.52,201.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.77,164.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (540.33,201.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (565.17,180.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (488.41,194.48) circle (  2.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.32, 72.00) -- (643.52, 72.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.32, 72.00) -- (402.32, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.52, 72.00) -- (442.52, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.72, 72.00) -- (482.72, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.92, 72.00) -- (522.92, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (563.12, 72.00) -- (563.12, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (603.32, 72.00) -- (603.32, 66.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (643.52, 72.00) -- (643.52, 66.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (402.32, 50.40) {-2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (442.52, 50.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (482.72, 50.40) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (522.92, 50.40) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (563.12, 50.40) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (603.32, 50.40) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (643.52, 50.40) {10};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,100.14) -- (396.61,260.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,100.14) -- (390.61,100.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,140.34) -- (390.61,140.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,180.54) -- (390.61,180.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,220.74) -- (390.61,220.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (396.61,260.94) -- (390.61,260.94);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at (384.61, 96.01) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at (384.61,136.21) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at (384.61,176.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at (384.61,216.61) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.20] at (384.61,256.81) {8};
\end{scope}
\begin{scope}
\path[clip] (324.61, 12.00) rectangle (649.23,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.50] at (522.92, 14.40) {Communion};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.50] at (346.21,180.54) {Agency};
\end{scope}
\begin{scope}
\path[clip] (396.61, 72.00) rectangle (649.23,289.08);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (519.55,177.16) --
	(526.30,177.16) --
	(526.30,183.91) --
	(519.55,183.91) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.62,100.14) --
	(462.58,101.40) --
	(462.46,102.66) --
	(462.27,103.91) --
	(461.99,105.14) --
	(461.64,106.35) --
	(461.21,107.54) --
	(460.71,108.70) --
	(460.14,109.82) --
	(459.49,110.91) --
	(458.78,111.95) --
	(458.01,112.95) --
	(457.17,113.90) --
	(456.28,114.79) --
	(455.33,115.63) --
	(454.34,116.40) --
	(453.29,117.11) --
	(452.21,117.75) --
	(451.08,118.33) --
	(449.92,118.83) --
	(448.73,119.26) --
	(447.52,119.61) --
	(446.29,119.88) --
	(445.04,120.08) --
	(443.78,120.20) --
	(442.52,120.24) --
	(441.26,120.20) --
	(440.00,120.08) --
	(438.76,119.88) --
	(437.52,119.61) --
	(436.31,119.26) --
	(435.12,118.83) --
	(433.96,118.33) --
	(432.84,117.75) --
	(431.75,117.11) --
	(430.71,116.40) --
	(429.71,115.63) --
	(428.76,114.79) --
	(427.87,113.90) --
	(427.04,112.95) --
	(426.26,111.95) --
	(425.55,110.91) --
	(424.91,109.82) --
	(424.34,108.70) --
	(423.83,107.54) --
	(423.41,106.35) --
	(423.05,105.14) --
	(422.78,103.91) --
	(422.58,102.66) --
	(422.46,101.40) --
	(422.42,100.14) --
	(422.46, 98.88) --
	(422.58, 97.62) --
	(422.78, 96.37) --
	(423.05, 95.14) --
	(423.41, 93.93) --
	(423.83, 92.74) --
	(424.34, 91.58) --
	(424.91, 90.46) --
	(425.55, 89.37) --
	(426.26, 88.33) --
	(427.04, 87.33) --
	(427.87, 86.38) --
	(428.76, 85.49) --
	(429.71, 84.65) --
	(430.71, 83.88) --
	(431.75, 83.17) --
	(432.84, 82.53) --
	(433.96, 81.95) --
	(435.12, 81.45) --
	(436.31, 81.02) --
	(437.52, 80.67) --
	(438.76, 80.40) --
	(440.00, 80.20) --
	(441.26, 80.08) --
	(442.52, 80.04) --
	(443.78, 80.08) --
	(445.04, 80.20) --
	(446.29, 80.40) --
	(447.52, 80.67) --
	(448.73, 81.02) --
	(449.92, 81.45) --
	(451.08, 81.95) --
	(452.21, 82.53) --
	(453.29, 83.17) --
	(454.34, 83.88) --
	(455.33, 84.65) --
	(456.28, 85.49) --
	(457.17, 86.38) --
	(458.01, 87.33) --
	(458.78, 88.33) --
	(459.49, 89.37) --
	(460.14, 90.46) --
	(460.71, 91.58) --
	(461.21, 92.74) --
	(461.64, 93.93) --
	(461.99, 95.14) --
	(462.27, 96.37) --
	(462.46, 97.62) --
	(462.58, 98.88) --
	(462.62,100.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.76,116.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.35,116.40) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.58,116.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.75,117.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.67,114.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (457.56,113.48) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.02,115.88) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.49,112.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.56,112.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.35,114.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.80,116.05) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.39,109.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.62,115.39) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (457.82,113.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.47,112.38) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.34,107.20) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.11,115.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (457.01,114.07) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.58,116.92) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,108.93) circle (  2.25);
\end{scope}
\end{tikzpicture}

\end{document}

### 2019-01-28 Manuscript/SupplementaryMaterial.Rmd
Update manuscript

@@
 
-In this Supplementary Material we outline the MCMC procedures to fit the cylindrical regression models.
+In this Supplementary Material we give a more detailed description of the cylindrical regression models and outline the MCMC procedures to fit them.
 R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
@@
 
+\section{Four cylindrical regression models}\label{Models}
+
+\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}
+
+Following @mastrantonio2015bayesian we consider in this section two models where
+the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
+and $q$ covariates is specified as
+\begin{equation}\label{circlinlink}
+Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
+\end{equation}
+\noindent where the random variable \(R\geq0\) will be introduced below, the
+error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
+intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
+covariates. In both of these models the
+conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
+by
+\begin{equation}\label{ycondtheta}
+f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
+\end{equation}
+\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
+\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
+\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
+\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
+in its location part.\newline
+\indent For the circular outcome we assume either a projected normal (PN) or a general
+projected normal (GPN) distribution. These distributions arise from the radial
+projection of a distribution defined on the plane onto the circle. The relation
+between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
+outcome \(\Theta\) is defined as follows
+\begin{equation}\label{projection}
+\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
+\end{equation}
+\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
+the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
+assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
+$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1
+\end{bmatrix}\), and \(\xi,\tau \in (-\infty, +\infty)\) (as in
+@hernandez2016general). This leads to the
+circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
+will now detail how we modify both cylindrical distributions to also incorporate
+covariates for the circular part.
+
+\subsubsection{The modified CL-PN distribution}
+
+Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
+PN distribution equals
+\begin{equation}\label{pnreg}
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
+\end{equation}
+\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
+\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
+\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
+the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
+\(n\) is the sample size, are generated independently from the distribution with
+density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
+is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
+where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
+contains the covariate values and the value 1 to estimate an intercept and
+\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
+the regression coefficients and intercepts.
+
+\subsubsection{The modified CL-GPN distribution}
+
+Following @wang2012directional and @hernandez2016general the joint density of
+\(R\) and \(\Theta\) for the GPN distribution equals
+\begin{equation}\label{gpnreg}
+f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
+\end{equation}
+\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
+\xi^2 & \xi\\ \xi & 1 \end{bmatrix}\). In a regression setup the outcomes
+\(\theta_i\) and \(r_i\) for each individual are generated independently from
+\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
+defined in the same way via covariates as for the modified CL-PN distribution.
+
+\subsubsection{Parameter estimation}
+
+Both cylindrical models introduced here are estimated using Markov Chain Monte
+Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
+@hernandez2016general for the regression of the circular outcome. A detailed
+description of the Bayesian estimation and MCMC samplers can be found in the
+Supplementary Material.
+
+\subsection{The modified Abe-Ley model}\label{WeiSSVM}
+
+This model is an extension of the cylindrical model introduced in
+@abe2017tractable to the regression context. The joint density of \(\Theta\) and
+\(Y\), in this model defined only on the positive real half-line \([0, +
+\infty)\), reads
+\begin{equation}\label{WeiSSVMdensity}
+f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
+                 (1 +\lambda\sin(\theta - \mu))
+                 y^{\alpha-1}
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
+\end{equation}
+\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
+\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
+\(\kappa\) also regulating the circular-linear dependence. Our modification
+occurs at the level of the linear scale parameter \(\nu>0\) and circular
+location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
+covariates: \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
+\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
+parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
+\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
+\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\beta_0 \in [0,
+2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
+regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
+\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
+vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
+is a vector of predictor values for the prediction of \(\theta\). In a
+regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
+generated independently from the modified
+density \eqref{WeiSSVMdensity}.\newline
+\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
+\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
+\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
+distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
+with location parameter \(\mu\) and concentration parameter \((\nu
+y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
+\begin{align}\label{WeiSSVMLikelihood}
+l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
+   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
+   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
+   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
+   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
+\end{align}
+\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
+the maximum likelihood (ML) estimates for the parameters of the model.
+
+\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}
+
+This model is an extension of the cylindrical model introduced by
+@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
+  outcomes and \(w\) independent linear outcomes. The circular outcomes
+  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
+  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
+  distribution.
+The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes.
+The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
+  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
+  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
+  so-called augmented representation (as in \eqref{projection} and
+  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
+  multivariate skew normal distribution. Following @mastrantonio2018joint the
+  linear outcomes are represented as
+\[\boldsymbol{Y} = \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
+\noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear outcome
+\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
+is a \(w \times w\) diagonal matrix with diagonal elements
+\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
+\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
+$w$-dimensional half normal distribution [@olmos2012extension], and
+\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
+that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
+is normally distributed with mean \(\boldsymbol{M}_y +
+\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
+\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
+\boldsymbol{D}^t)^t\) is defined as:
+\begin{equation}\label{YDjoint}
+f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
+\end{equation}
+\noindent where
+\(\phi_\ell(\cdot|\boldsymbol{M}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
+the \(\ell\)-dimensional normal density with mean vector
+\(\boldsymbol{M}_\ell\) and
+covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
+modelling the augmented representations of \(\boldsymbol{\Theta}\) and
+\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
+distribution. The joint density of the model is then represented by:
+\begin{equation}\label{YDThetarjoint} 
+f(\boldsymbol{\theta}, \boldsymbol{r},
+\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
+\boldsymbol{y}^t)^t \mid \boldsymbol{M} + (\boldsymbol{0}_{2m}^t, ({\rm
+diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
+\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
+1}^{m}r_j, 
+\end{equation}
+\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
+r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{M} =
+(\boldsymbol{M}_s^t, \boldsymbol{M}_y^t)^t\) and \(\boldsymbol{\Sigma} =
+\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
+\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
+The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
+of and covariances between the augmented representations of the circular outcome
+and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
+augmented representations of the circular outcome and the linear outcome.
+\newline
+\indent In our regression extension we have \(i = 1, \dots, n\) observations of
+\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
+the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{M}_i =
+\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
+\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
+is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
+and the $g$ covariate values. \newline
+\indent We estimate the model using MCMC methods. A detailed description of these
+methods is given in the Supplementary Material.
+
+\section{Model fit}
+
+\indent We use the loglikelihoods of the following conditional densities for the
+computation of the PLSL in the teacher data:
+
+\begin{itemize}
+\item For the modified CL-PN model:
+
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
+
+\item For the modified CL-GPN model:
+
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
+
+\item For the modified Abe-Ley model:
+
+$y_i \mid \theta_i, \nu_i, \mu_i, \kappa, \alpha \sim W\left(\nu_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.
+
+$\theta_i \mid y_i, \nu_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\nu_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.
+
+\item For the modified joint projected and skew normal model:
+
+$y_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(M_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{M}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$
+
+$\theta_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{M}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - M_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$
+
+where $SSN$ is the skew normal distribution.
+\end{itemize}
+
+\section{MCMC procedures}
 \subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}
@@
 \item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
-\item Prior for $\rho$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
+\item Prior for $\xi$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
 \item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
@@
 
-\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$ for $j \in \{0,1\}$, $\rho = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$.
+\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$ for $j \in \{0,1\}$, $\xi = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}$.
 Also set starting values $r_i = 1$ in the training and holdout set.
@@
 $$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
-\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$ for $j \in \{0,1\}$, $\rho$ and $\tau$ for the training dataset from their conditional posteriors:
+\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$ for $j \in \{0,1\}$, $\xi$ and $\tau$ for the training dataset from their conditional posteriors:
 
@@
 \item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
-\item Posterior for $\rho$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = (\boldsymbol{\beta}^{I})^t\boldsymbol{z}_i$ and $\mu_i^{II} = (\boldsymbol{\beta}^{II})^t\boldsymbol{z}_i$.
-\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{I} + \rho(s^{II}_{i} - \mu_i^{II})\})^2 + \beta_0$
+\item Posterior for $\xi$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = (\boldsymbol{\beta}^{I})^t\boldsymbol{z}_i$ and $\mu_i^{II} = (\boldsymbol{\beta}^{II})^t\boldsymbol{z}_i$.
+\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{I} + \xi(s^{II}_{i} - \mu_i^{II})\})^2 + \beta_0$
 \end{itemize}
@@
 
-\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\rho$ and $\tau$ for the training dataset.
+\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\xi$ and $\tau$ for the training dataset.
 \item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We visually assess convergence using traceplots.
@@
 \end{enumerate}
+
+\section*{References}
+<div id="refs"></div>

### 2019-01-28 Manuscript/SupplementaryMaterial.tex
Update manuscript

@@
 \setlength\parindent{24pt}
-\usepackage{setspace}\doublespacing
+\usepackage{setspace}
+
+\doublespacing
 
@@
 
-In this Supplementary Material we outline the MCMC procedures to fit the
-cylindrical regression models. R-code for the MCMC sampler and the
-analysis of the teacher data can be found here:
+In this Supplementary Material we give a more detailed description of
+the cylindrical regression models and outline the MCMC procedures to fit
+them. R-code for the MCMC sampler and the analysis of the teacher data
+can be found here:
 \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
@@
 
+\section{Four cylindrical regression models}\label{Models}
+
+\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}
+
+Following Mastrantonio, Maruotti, \& Jona-Lasinio (2015) we consider in
+this section two models where the relation between
+\(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\) and \(q\)
+covariates is specified as \begin{equation}\label{circlinlink}
+Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
+\end{equation} \noindent where the random variable \(R\geq0\) will be
+introduced below, the error term \(\epsilon \sim N(0, \sigma^2)\) with
+variance \(\sigma^2>0\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
+the intercept and regression coefficients and \(x_1, \dots, x_q\) are
+the \(q\) covariates. In both of these models the conditional
+distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given by
+\begin{equation}\label{ycondtheta}
+f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
+\end{equation} \noindent where
+\(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
+\(r \geq 0\). The linear outcome thus has a normal distribution
+conditional on \(\Theta\) and \(R\) and contains already linear
+covariates \(x_1, \dots, x_q\) in its location part.\newline \indent For
+the circular outcome we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise
+from the radial projection of a distribution defined on the plane onto
+the circle. The relation between a bivariate vector \(\boldsymbol{S}\)
+in the plane and the circular outcome \(\Theta\) is defined as follows
+\begin{equation}\label{projection}
+\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
+\end{equation} \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\),
+the Euclidean norm of the bivariate vector \(\boldsymbol{S}\). In the PN
+distribution we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the
+GPN we assume
+\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
+\(\boldsymbol{\mu} \in \mathbb{R}^2\),
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\),
+and \(\xi,\tau \in (-\infty, +\infty)\) (as in Hernandez-Stumpfhauser,
+Breidt, \& Woerd (2016)). This leads to the circular-linear PN (CL-PN)
+and circular-linear GPN (CL-GPN) distributions. We will now detail how
+we modify both cylindrical distributions to also incorporate covariates
+for the circular part.
+
+\subsubsection{The modified CL-PN distribution}
+
+Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
+density of \(\Theta\) and \(R\) for the PN distribution equals
+\begin{equation}\label{pnreg}
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
+\end{equation} \noindent where
+\(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
+and \(r\) is the same as in \eqref{circlinlink} and \eqref{ycondtheta}
+and is defined in \eqref{projection}. In a regression setup the outcomes
+\(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where \(n\) is
+the sample size, are generated independently from the distribution with
+density \eqref{pnreg}. The mean vector
+\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
+\(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
+vector \(\boldsymbol{z}_i\) is a vector of dimension \(p + 1\) that
+contains the covariate values and the value 1 to estimate an intercept
+and
+\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
+contains the regression coefficients and intercepts.
+
+\subsubsection{The modified CL-GPN distribution}
+
+Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
+(2016) the joint density of \(R\) and \(\Theta\) for the GPN
+distribution equals \begin{equation}\label{gpnreg}
+f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
+\end{equation} \noindent where we recall that
+\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\).
+In a regression setup the outcomes \(\theta_i\) and \(r_i\) for each
+individual are generated independently from \eqref{gpnreg}. The mean
+vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the same
+way via covariates as for the modified CL-PN distribution.
+
+\subsubsection{Parameter estimation}
+
+Both cylindrical models introduced here are estimated using Markov Chain
+Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
+Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
+regression of the circular outcome. A detailed description of the
+Bayesian estimation and MCMC samplers can be found in the Supplementary
+Material.
+
+\subsection{The modified Abe-Ley model}\label{WeiSSVM}
+
+This model is an extension of the cylindrical model introduced in Abe \&
+Ley (2017) to the regression context. The joint density of \(\Theta\)
+and \(Y\), in this model defined only on the positive real half-line
+\([0, + \infty)\), reads \begin{equation}\label{WeiSSVMdensity}
+f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
+                 (1 +\lambda\sin(\theta - \mu))
+                 y^{\alpha-1}
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
+\end{equation} \noindent where \(\alpha > 0\) is a linear shape
+parameter, \(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular
+concentration and skewness parameters with \(\kappa\) also regulating
+the circular-linear dependence. Our modification occurs at the level of
+the linear scale parameter \(\nu>0\) and circular location parameter
+\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
+\(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
+\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
+The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
+coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
+\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
+parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
+\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
+\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
+where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
+of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
+is a vector of predictor values for the prediction of \(\theta\). In a
+regression setup the outcome vector \((\theta_i, y_i)^t\) for each
+individual is generated independently from the modified density
+\eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
+conditional distribution of \(Y\) given \(\Theta=\theta\) is a Weibull
+distribution with shape \(\alpha\) and scale
+\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the
+conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
+von Mises distribution with location parameter \(\mu\) and concentration
+parameter \((\nu y)^\alpha\tanh(\kappa)\). The log-likelihood for this
+model equals \begin{align}\label{WeiSSVMLikelihood}
+l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
+   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
+   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
+   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
+   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
+\end{align} \noindent We can use numerical optimization (Nelder-Mead) to
+find solutions for the maximum likelihood (ML) estimates for the
+parameters of the model.
+
+\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}
+
+This model is an extension of the cylindrical model introduced by
+Mastrantonio (2018) to the regression context. Both models contain \(m\)
+independent circular outcomes and \(w\) independent linear outcomes. The
+circular outcomes
+\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\)
+are modelled together by a multivariate GPN distribution. The joint
+distribution of \(\boldsymbol{\Theta}\) and \(\boldsymbol{R}\) can thus
+be modeled as the product of (\ref{gpnreg}) for each of the \(m\)
+circular outcomes. The linear outcomes
+\(\boldsymbol{Y} = (\boldsymbol{Y}_1,  \dots, \boldsymbol{Y}_w)\) are
+modelled together by a multivariate skew normal distribution (Sahu, Dey,
+\& Branco, 2003). Because the GPN distribution is modelled using a
+so-called augmented representation (as in \eqref{projection} and
+\eqref{gpnreg}) it is convenient to use a similar tactic for modelling
+the multivariate skew normal distribution. Following Mastrantonio (2018)
+the linear outcomes are represented as
+\[\boldsymbol{Y} = \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
+\noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear
+outcome \(\boldsymbol{Y}\),
+\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
+\(w \times w\) diagonal matrix with diagonal elements
+\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
+\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
+\(w\)-dimensional half normal distribution (Olmos, Varela, Gómez, \&
+Bolfarine, 2012), and
+\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\).
+This means that, conditional on the auxiliary data \(\boldsymbol{D}\),
+\(\boldsymbol{Y}\) is normally distributed with mean
+\(\boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and covariance
+matrix \(\boldsymbol{\Sigma}_y\). The joint density for
+\((\boldsymbol{Y}^t, \boldsymbol{D}^t)^t\) is defined as:
+\begin{equation}\label{YDjoint}
+f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
+\end{equation} \noindent where
+\(\phi_\ell(\cdot|\boldsymbol{M}_\ell,\boldsymbol{\Sigma}_\ell)\) stands
+for the \(\ell\)-dimensional normal density with mean vector
+\(\boldsymbol{M}_\ell\) and covariance \(\boldsymbol{\Sigma}_\ell\). As
+in Mastrantonio (2018) dependence between the linear and circular
+outcome is created by modelling the augmented representations of
+\(\boldsymbol{\Theta}\) and \(\boldsymbol{Y}\) together in a \(2m + w\)
+dimensional normal distribution. The joint density of the model is then
+represented by: \begin{equation}\label{YDThetarjoint} 
+f(\boldsymbol{\theta}, \boldsymbol{r},
+\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
+\boldsymbol{y}^t)^t \mid \boldsymbol{M} + (\boldsymbol{0}_{2m}^t, ({\rm
+diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
+\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
+1}^{m}r_j, 
+\end{equation} \noindent where
+\(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots, r_m(\cos(\theta_m), \sin(\theta_m)))^t\),
+the mean vector
+\(\boldsymbol{M} = (\boldsymbol{M}_s^t, \boldsymbol{M}_y^t)^t\) and
+\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
+The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
+variances of and covariances between the augmented representations of
+the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
+contains covariances between the augmented representations of the
+circular outcome and the linear outcome. \newline \indent In our
+regression extension we have \(i = 1, \dots, n\) observations of \(m\)
+circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean
+in the density in \eqref{YDThetarjoint} then becomes
+\(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
+\(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with
+regression coefficients and intercepts and \(\boldsymbol{x}_i\) is a
+\(g + 1\) dimensional vector containing the value 1 to estimate an
+intercept and the \(g\) covariate values. \newline \indent We estimate
+the model using MCMC methods. A detailed description of these methods is
+given in the Supplementary Material.
+
+\section{Model fit}
+
+\indent We use the loglikelihoods of the following conditional densities
+for the computation of the PLSL in the teacher data:
+
+\begin{itemize}
+\item For the modified CL-PN model:
+
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
+
+\item For the modified CL-GPN model:
+
+$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
+
+\item For the modified Abe-Ley model:
+
+$y_i \mid \theta_i, \nu_i, \mu_i, \kappa, \alpha \sim W\left(\nu_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.
+
+$\theta_i \mid y_i, \nu_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\nu_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.
+
+\item For the modified joint projected and skew normal model:
+
+$y_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(M_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{M}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$
+
+$\theta_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{M}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - M_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$
+
+where $SSN$ is the skew normal distribution.
+\end{itemize}
+
+\section{MCMC procedures}
 \subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}
@@
 \newpage
-
 \subsection{Bayesian Model and MCMC procedure for the modified CL-GPN model}\label{A2}
@@
 \newpage
-
 \subsection{Bayesian Model and MCMC procedure for the modified GPN-SSN model}\label{A3}
@@
 
+\section*{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-abe2017tractable}{}%
+Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
+for cylindrical data, with applications. \emph{Econometrics and
+Statistics}, \emph{4}, 91--104.
+doi:\href{https://doi.org/10.1016/j.ecosta.2016.04.001}{10.1016/j.ecosta.2016.04.001}
+
+\leavevmode\hypertarget{ref-hernandez2016general}{}%
+Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
+(2016). The general projected normal distribution of arbitrary
+dimension: Modeling and Bayesian inference. \emph{Bayesian Analysis},
+\emph{12}(1), 113--133.
+doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}
+
+\leavevmode\hypertarget{ref-mastrantonio2018joint}{}%
+Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
+distribution for poly-cylindrical data. \emph{Journal of Multivariate
+Analysis}, \emph{165}, 14--26.
+doi:\href{https://doi.org/10.1016/j.jmva.2017.11.006}{10.1016/j.jmva.2017.11.006}
+
+\leavevmode\hypertarget{ref-mastrantonio2015bayesian}{}%
+Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
+hidden Markov modelling using circular-linear general projected normal
+distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
+doi:\href{https://doi.org/10.1002/env.2326}{10.1002/env.2326}
+
+\leavevmode\hypertarget{ref-nunez2011bayesian}{}%
+Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
+Bayesian regression model for circular data based on the projected
+normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
+185--201.
+doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
+
+\leavevmode\hypertarget{ref-olmos2012extension}{}%
+Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
+extension of the half-normal distribution. \emph{Statistical Papers},
+\emph{53}(4), 875--886.
+doi:\href{https://doi.org/10.1007/s00362-011-0391-4}{10.1007/s00362-011-0391-4}
+
+\leavevmode\hypertarget{ref-sahu2003new}{}%
+Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
+multivariate skew distributions with applications to Bayesian regression
+models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
+doi:\href{https://doi.org/10.2307/3316064}{10.2307/3316064}
+
+\leavevmode\hypertarget{ref-wang2012directional}{}%
+Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
+general projected normal distribution. \emph{Statistical Methodology},
+\emph{10}(1), 113--127.
+doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}
+
 \end{document}

### 2019-01-28 Manuscript/R-code/Abe-Ley optimization.R
Annotate R-code

@@ -1,7 +1,23 @@
+# Abe-Ley optimization
+#
+# This script contains helper functions for optimization of the Abe-Ley model for cylindrical data
+#
+
+
+## define helper functions ##
+
+# input function for optimization
+
+# input
+# par = vector of starting values for the parameters
+# data = dataframe containing the circular (theta) and linear outcome  (y)
+
+# output = loglikelihood
+
 func.regII <- function(par, data){
   
-  eta <- par[1:2]
-  nu <- par[3:4]
+  beta <- par[1:2]
+  gamma <- par[3:4]
   kappa <- par[5]
   alpha <- par[6]
   lambda <- par[7]
@@ -13,10 +29,10 @@ func.regII <- function(par, data){
   n <- length(data$y)
   
   p1 <- n*(log(alpha) - log(2*pi*cosh(kappa)))
-  p2 <- alpha * sum(log(exp(x%*%nu)))
-  p3 <- sum(log(1 + lambda*sin(theta - eta[1] + 2*atan(z[,2]*eta[2]))))
+  p2 <- alpha * sum(log(exp(x%*%gamma)))
+  p3 <- sum(log(1 + lambda*sin(theta - beta[1] + 2*atan(z[,2]*beta[2]))))
   p4 <- (alpha - 1) * sum(log(y))
-  p5 <- sum((exp(x%*%nu)*y)^alpha*(1 - tanh(kappa)*cos(theta - eta[1] +  2*atan(z[,2]*eta[2]))))
+  p5 <- sum((exp(x%*%gamma)*y)^alpha*(1 - tanh(kappa)*cos(theta - beta[1] +  2*atan(z[,2]*beta[2]))))
   
   
   p1+p2+p3+p4-p5
@@ -24,11 +40,17 @@ func.regII <- function(par, data){
   
 }
 
+## function to compute the linear loglikelihood ##
+
+# input
+# par = vector of parameter values
+# data = dataframe containing the circular (theta) and linear outcome  (y)
+
 func.reg.cond.lin <- function(par, data){
   
   
-  eta <- par[1:2]
-  nu <- par[3:4]
+  beta <- par[1:2]
+  gamma <- par[3:4]
   kappa <- par[5]
   alpha <- par[6]
   lambda <- par[7]
[…]

### 2019-01-28 Manuscript/R-code/Posterior Sampling CL-GPN.R

@@ -1,16 +1,31 @@
-#Functions for posterior distributions
+# Posterior Sampling CL-GPN
+#
+# This script contains functions to sample from the posterior of the CL-GPN model for cylindrical data
+#
 
-#Sampling Gamma and Sigma, the regression coefficients and residual variance of 
-#the regression on the linear outcome y.
 
-#N = sample size
-#X = design matrix
-#y = linear outcome
-#sig = residual variance (starting value or value from previous iteration)
-#p.gamma.mu = prior mean vector for the regression coefficients
-#p.gamma.prec = prior precision for the regression coefficients
-#p.sig.shape = prior shape parameter for the residual variance
-#p.sig.scale = prior scale parameter for the residual variance
+
+
+
+##### define helper functions #####
+
+
+## Functions to sample from posterior distributions ##
+
+
+# Sample Gamma and Sigma, the regression coefficients and residual variance of 
+# the regression on the linear outcome y.
+
+# input
+
+# N = sample size
+# X = design matrix
+# y = linear outcome
+# sig = residual variance (starting value / previous iteration)
+# p.gamma.mu = prior mean vector for the regression coefficients
+# p.gamma.prec = prior precision for the regression coefficients
+# p.sig.shape = prior shape parameter for the residual variance
+# p.sig.scale = prior scale parameter for the residual variance
 
 post.gamma.sig <- function(N, X, y, sig, p.gamma.mu, p.gamma.prec, p.sig.shape, p.sig.scale){
   
@@ -29,8 +44,19 @@ post.gamma.sig <- function(N, X, y, sig, p.gamma.mu, p.gamma.prec, p.sig.shape, 
   
 }
 
-#Sampling rho, element of sigma, the variance-covariance matrix of the circular 
-#outcome
+# Sample rho, element of sigma, the variance-covariance matrix of the circular 
+# outcome
+
+# input
+
+# Y = augmented circular outcome
+# Z = design matrix
+# Beta.I = regression coefficients for the first component (starting value / previous iteration)
+# Beta.II = regression coefficients for the second component (starting value / previous iteration)
+# tau = element of sigma (starting value / previous iteration)
+# p.rho.mu = prior mean rho
+# p.rho.var = prior variance rho
+
 
 post.rho <- function(Y, Z, Beta.I, Beta.II, tau, p.rho.mu, p.rho.var){
   
[…]

### 2019-01-28 Manuscript/R-code/Posterior Sampling CL-PN.R

@@ -1,16 +1,30 @@
-#Functions for posterior distributions
+# Posterior Sampling CL-PN
+#
+# This script contains functions to sample from the posterior of the CL-PN model for cylindrical data
+#
 
-#Sampling Gamma and Sigma, the regression coefficients and residual variance of 
-#the regression on the linear outcome y.
 
-#N = sample size
-#X = design matrix
-#y = linear outcome
-#sigma = residual variance (starting value or value from previous iteration)
-#p.gamma.mu = prior mean vector for the regression coefficients
-#p.gamma.prec = prior precision for the regression coefficients
-#p.sigma.shape = prior shape parameter for the residual variance
-#p.sigma.scale = prior scale parameter for the residual variance
+
+
+
+##### define helper functions #####
+
+
+## Functions to sample from posterior distributions ##
+
+# Sample Gamma and Sigma, the regression coefficients and residual variance of 
+# the regression on the linear outcome y.
+
+# input
+
+# N = sample size
+# X = design matrix
+# y = linear outcome
+# sig = residual variance (starting value / previous iteration)
+# p.gamma.mu = prior mean vector for the regression coefficients
+# p.gamma.prec = prior precision for the regression coefficients
+# p.sig.shape = prior shape parameter for the residual variance
+# p.sig.scale = prior scale parameter for the residual variance
 
 post.gamma.sigma <- function(N, X, y, sigma, p.gamma.mu, p.gamma.prec, p.sigma.shape, p.sigma.scale){
   
@@ -30,13 +44,15 @@ post.gamma.sigma <- function(N, X, y, sigma, p.gamma.mu, p.gamma.prec, p.sigma.s
   
 }
 
-#Sampling Beta, the regression coefficients of one component (I, II) of the 
-#regression on the circular outcome theta.
+# Sample Beta, the regression coefficients of one component (I, II) of the 
+# regression on the circular outcome theta.
 
-#Z = design matrix
-#Y = (cos(theta), sin(theta))*r
-#p.beta.mu = prior mean for the regression coefficients
-#p.beta.prec = prior precision for the regression coefficients
+# input
+
+# Z = design matrix
+# Y = (cos(theta), sin(theta))*r (augmented circular outcome)
+# p.beta.mu = prior mean for the regression coefficients
+# p.beta.prec = prior precision for the regression coefficients
 
 
 post.beta <- function(Z, Y, p.beta.mu, p.beta.prec){
[…]

### 2019-01-28 Manuscript/R-code/Posterior Sampling Joint GPN-SSN.R

@@ -1,14 +1,37 @@
+# Posterior Sampling Joint GPN-SSN
+#
+# This script contains functions to sample from the posterior of the joint GPN-SSN model for cylindrical data
+#
+
+
+
+
+
+##### load required packages #####
+
 require(MCMCpack)
 require(truncnorm)
-#require(LaplacesDemon)
 require(bayesm)
 require(mvtnorm)
 
-#Constraint matrix function (for 1 circular and 1 linear variable)
+
+
+
+
+##### define helper functions #####
+
+
+## Helper function to compute the constrained matrix for 1 circular and 1 linear variable ##
+
+# var = the element of the matrix that should be constrained to 1
 
 cmat <- function(var){diag(3)*(c(sqrt(var), sqrt(var), 1))}
 
-#sample from matrix normal (adapted from LaplacesDemon, function is.symmetric.matrix does not work properly; floating point problems...)
+
+## Sample from matrix normal distribution ##
+
+# This function is adapted from LaplacesDemon, 
+# the function is.symmetric.matrix from LaplacesDemon does not work properly due to floating point problems.
 
 rmatrixnorm <- function (M, U, V) 
 {
@@ -35,18 +58,19 @@ rmatrixnorm <- function (M, U, V)
   return(X)
 }
 
-#Functions for posterior distributions
 
-#Sample Beta, a matrix of regression coefficients for the circular and linear 
-#variables
+## Functions to sample from posterior distributions ##
 
-#Y = the circular and linear outcomes
-#X = a design matrix
-#Beta = Beta matrix (previous iteration/starting value)
-#Kappa = Kappa matrix (previous iteration/starting value)
-#Sigma  = Sigma matrix (previous iteration/starting value)
-#p.Beta = 
-#p.Kappa = 
+# Sample Beta, regression coefficient matrix
+
+# input:
+
+# Y = the circular and linear outcomes
+# X = a design matrix (covariate values)
+# Beta = Beta matrix (previous iteration/starting value)
+# Sigma  = Sigma matrix (previous iteration/starting value)
+# p.Kappa = prior scale matrix
+# p.Beta = prior location matrix normal distribution
 
 post.Beta <- function(Y, X, Beta, Sigma, p.Beta, p.Kappa){
 
@@ -59,6 +83,18 @@ return(Beta)
 
 }
 
[…]

### 2019-01-28 Manuscript/R-code/Sample Abe-Ley.R

@@ -1,4 +1,16 @@
-#simulate from wrapped cauchy distribution
+# Sample Abe-Ley 
+#
+# This script contains helper functions for sampling from the Abe-Ley model for cylindrical data
+#
+
+
+## simulate from wrapped cauchy distribution ##
+
+# input
+
+# n = sample size
+# mu = circular mean
+# rho = circular concentration
 
 rwcauchy <- function(n, mu, rho){
   
@@ -14,8 +26,14 @@ rwcauchy <- function(n, mu, rho){
   
 }
 
-#simulate from abe-ley distribution
-rweiSSVM <- function(n, beta = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
+## simulate from Abe-Ley distribution ##
+
+# input
+
+# n = sample size
+# nu, alpha, mu, kappa, lambda = parameters Abe-Ley distribution
+
+rweiSSVM <- function(n, nu = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
   
   theta <- c()
   x <- c()
@@ -33,11 +51,11 @@ rweiSSVM <- function(n, beta = 1, alpha = 1, mu = pi, kappa = 1, lambda = 0){
       theta[i] <- -theta.1
     }
     
-    shape <- beta * (1-tanh(kappa)*cos(theta-mu))^(1/alpha)
-    #compute scale parameter for different parametrization,
-    #shape = (scale_parWeibull in R)^(-alpha) = 1/(scale_parWeibull in R)^alpha dus 
-    #scale_parWeibull in R = kth root(1/shape) dus
-    #scale_parWeibull in R = (1/b)^(1/alpha)
+    shape <- nu * (1-tanh(kappa)*cos(theta-mu))^(1/alpha)
+    # We compute the scale parameter for a different parametrization,
+    # shape = (scale_parWeibull in R)^(-alpha) = 1/(scale_parWeibull in R)^alpha so,
+    # scale_parWeibull in R = kth root(1/shape) so,
+    # scale_parWeibull in R = (1/b)^(1/alpha)
     scale <- (1/shape)^(1/alpha) 
     
     x[i] <- rweibull(1, shape = alpha, scale = scale)

### 2019-01-29 Manuscript/R-code/Posterior Sampling CL-GPN.R
Update code annotation

@@ -44,7 +44,7 @@ post.gamma.sig <- function(N, X, y, sig, p.gamma.mu, p.gamma.prec, p.sig.shape, 
   
 }
 
-# Sample rho, element of sigma, the variance-covariance matrix of the circular 
+# Sample xi, element of sigma, the variance-covariance matrix of the circular 
 # outcome
 
 # input
@@ -54,19 +54,19 @@ post.gamma.sig <- function(N, X, y, sig, p.gamma.mu, p.gamma.prec, p.sig.shape, 
 # Beta.I = regression coefficients for the first component (starting value / previous iteration)
 # Beta.II = regression coefficients for the second component (starting value / previous iteration)
 # tau = element of sigma (starting value / previous iteration)
-# p.rho.mu = prior mean rho
-# p.rho.var = prior variance rho
+# p.xi.mu = prior mean xi
+# p.xi.var = prior variance xi
 
 
-post.rho <- function(Y, Z, Beta.I, Beta.II, tau, p.rho.mu, p.rho.var){
+post.xi <- function(Y, Z, Beta.I, Beta.II, tau, p.xi.mu, p.xi.var){
   
-  num <- tau^(-1) * t(Y[,1] - Z%*%Beta.I)%*%(Y[,2] - Z%*%Beta.II) + p.rho.mu*p.rho.var^(-1)
-  den <- tau^(-1) * t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II) + p.rho.var^(-1)
+  num <- tau^(-1) * t(Y[,1] - Z%*%Beta.I)%*%(Y[,2] - Z%*%Beta.II) + p.xi.mu*p.xi.var^(-1)
+  den <- tau^(-1) * t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II) + p.xi.var^(-1)
   
-  mu.rho.post <- num/den
-  var.rho.post <- 1/(tau^(-1)*(t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II)) + p.rho.var^(-1))
+  mu.xi.post <- num/den
+  var.xi.post <- 1/(tau^(-1)*(t(Y[,2] - Z%*%Beta.II)%*%(Y[,2] - Z%*%Beta.II)) + p.xi.var^(-1))
   
-  rho <- rnorm(1, mu.rho.post, var.rho.post)
[…]

### 2019-01-29 Manuscript/ManuscriptMarkdownCV.Rmd
Update manuscript

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
+    address       : ""
     email         : "joliencremers@gmail.com"
@@
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
 
@@
 sciences data with a circular and linear outcome more commonly occur in
-meteorology meteorology [@garcia2013exploring], ecology [@garcia2014test] or
+meteorology [@garcia2013exploring], ecology [@garcia2014test] or
 marine research [@lagona2015hidden]
 \newline
-\indent Up until now researchers studying cylindrical data had to rely on the linear
-statistical methods to analyze their research results. Focusing on only the
-circular variable or only on the linear variable. However, lately more and more
+\indent Up until now researchers studying cylindrical data had to rely on linear
+statistical methods to analyze their research results. However, lately more and more
 of these researchers acknowledge that linear methods are not sufficient and call
@@
 \newline
-\indent The aim in the present paper is twofold. Firstly, we intend to fill this gap in
-the literature on cylindrical models by adapting four existing cylindrical
-models in such a way that they include a regression of both the linear and
-circular component of a cylindrical variable onto a set of covariates. From now
-on we will therefore refer to the components of the cylindrical variable as
-outcome components. Secondly, we will show how using these models can benefit
-the analysis of circumplex data and cylindrical data in psychology in general.
-More specifically we will show these benefits for interpersonal teacher data
-from the field of educational psychology. Apart from modelling the dependence
-between the linear and circular component of a cylindrical variable we would
-also like to predict the two components from a set of covariates in a regression
-model. In our analysis, we will show how this can be done in four different ways
-using the adapted cylindrical models.
+\indent The aim in the present paper is twofold. Firstly, we we intend to fill
+the above mentioned gap in the literature by showing the use of cylindrical
+models can benefit the analysis of circumplex data and cylindrical data in
+psychology in general. More specifically we will show these benefits for
+interpersonal teacher data from the field of educational psychology. Apart from
+modelling the dependence between the linear and circular component of a
+cylindrical variable we would also like to predict the two components from a set
+of covariates in a regression model. Our second aim therefore is to adapt
+several existing cylindrical models in such a way that they include a regression
+of both the linear and circular component of a cylindrical variable onto a set
+of covariates. From now on we will therefore refer to the components of the
+cylindrical variable as outcome components. These adapted circular models are
+then used  to analyse the teacher data.
 
@@
 
-Thus, data that consist of a variable and a circular variable is called
+Data that consist of a variable and a circular variable is called
 cylindrical data. A circular variable is different from the linear variable in
@@
 dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
-this space are characterized by different teacher behavior, e.g. 'helpful' or
-'uncertain'. This two-dimensional space is called the interpersonal
-circle/circumplex (IPC). The IPC is ``a continuous order with no beginning or
-end'' [@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering
-and the IPC is therefore often called the interpersonal circumplex. The ordering
+(vertical) and Communion (horizontal), see Figure \ref{QTI}. This space is
+called the interpersonal circle/circumplex (IPC) and different parts of this
+space are characterized by different teacher behavior, e.g. 'helpful' or
+'uncertain'. The IPC is ``a continuous order with no beginning or end''
+[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
+the IPC is therefore often called the interpersonal circumplex. The ordering
 also implies that scores on the IPC could be viewed as a circular variable. This
@@
 the origin in circumplex data depends on the scaling of the Agency and Communion
-scores; their respective 0 scores form the origin. Agency and Communion are
-measured on a scale from 1 to 5 and for analysis purposes they are later
-converted to scale ranging from -1 to 1. Scaling is however only considered an
-issue in those instances where cylindrical data is derived from measurements in
-bivariate space.}, where \(A\) represents a
-score on the Agency dimension and \(C\) represents a score on the Communion
-dimension. Note that when placing a unit circle on Figure \ref{QTI} we see that
-the Agency dimension is related to the sine of the circular score and the
-Communion dimension is related to the cosine of the circular score.
+scores. Agency and Communion are measured on a scale from 1 to 5 and for
+analysis purposes they are later converted to scale ranging from -1 to 1.  Their
+respective 0 scores form the origin. Scaling is however only considered an issue
+in those instances where cylindrical data is derived from measurements in
+bivariate space.}, where \(A\) represents a score on the Agency dimension and
+\(C\) represents a score on the Communion dimension. Note that when placing a
+unit circle on Figure \ref{QTI} we see that the Agency dimension is related to
+the sine of the circular score and the Communion dimension is related to the
+cosine of the circular score.
 \begin{equation}\label{PredVal}
@@
 in line with its theoretical definition and it allows us to analyse the blend of
-the two dimensions Agency and Communion instead of both dimensions separately
-providing us with new insights compared to a separate analysis of the two
-dimensions that is standard in the literature (see e.g. @pennings2018interpersonal, @wright2009integrating or @wubbels2006interpersonal). There is however one main drawback: when two-dimensional
-data are converted to the circle we lose some information, namely the length of
-the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm
-\(\mid\mid (A, C)^t \mid\mid\). This length represents the strength of the
-interpersonal behavior a teacher shows towards his/her students. In a
-cylindrical model this strength (the linear outcome) can be modeled together
-with the type of interpersonal behavior of a teacher (the circular outcome).
-This leads to an improved analysis of interpersonal circumplex data, over either
-analyzing the two dimensions separately or using a circular model, because we
-take all information, circular and linear, into account. In the next section we
-introduce several cylindrical models that can be used to analyze the teacher
-data. First however we will provide descriptives for the teacher data.
+the two dimensions Agency and Communion instead of both dimensions separately.
+This provides us with new insights compared to a separate analysis of the two
+dimensions that is standard in the literature (see e.g.
+@pennings2018interpersonal, @wright2009integrating or
+@wubbels2006interpersonal). There is however one main drawback: when
+two-dimensional data are converted to the circle we lose some information,
+namely the length of the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its
+Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
+strength of the interpersonal behavior a teacher shows towards his/her students.
+In a cylindrical model this strength (the linear outcome) can be modeled
+together with the type of interpersonal behavior of a teacher (the circular
+outcome). This leads to an improved analysis of interpersonal circumplex data,
+over either analyzing the two dimensions separately or using a circular model,
+because we take all information, circular and linear, into account. In the next
+section we introduce several cylindrical models that can be used to analyze the
+teacher data. First however we will provide descriptives for the teacher data.
 
@@
 consider one measurement, the first occasion (2010) and largest class if data
-for multiple classes were available. This results in a sample of 151 teachers. In addition to the type of interpersonal
-behavior (IPC), the circular outcome, and the strength of interpersonal behavior
-(IPC strength), the linear outcome, a teachers' self-efficacy (\verb|SE|)
-concerning classroom management is used as covariate in the analysis.
-In previous research, in psychology and education it had been shown that higher self-efficacy is related to the quality of interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
-After listwise
-deletion of missings ($3$ in total, only for the self-efficacy) we have a sample
-of 148 teachers. Table \ref{Tableteacherdescriptives} shows descriptives for the
-dataset. For the circular variable IPC we show sample estimates for the circular
-mean $\bar{\theta}$ and concentration $\hat{\rho}$. The circular concentration
-lies between 0, meaning the data is not concentrated at all \emph{i.e.} spread
-over the entire circle, and 1, meaning all data is concentrated at a single
-point on the circle The population values of these parameters are usually, and
-also in this paper, referred to as $\mu$ (circular location) and $\kappa$
-(circular concentration). For the linear variables (strength IPC and SE) we show
-sample estimates of the linear mean and standard deviation. Figure
-\ref{dataplot} is a scatterplot showing the relation between the linear and
-circular outcome of the teacher data for teachers with low SE (below 1 sd below
-the mean), average SE (between 1 sd below and 1 sd above the mean) and high SE
-(above 1 sd above the mean).
+for multiple classes were available. This results in a sample of 151 teachers.
+In addition to the type of interpersonal behavior (IPC), the circular outcome,
+and the strength of interpersonal behavior (IPC strength), the linear outcome, a
+teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
+covariate in the analysis. In previous research, in psychology and education it
+has been shown that higher self-efficacy is related to the quality of
+interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
+After listwise deletion of missings ($3$ in total, only for the self-efficacy)
+we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
+descriptives for the dataset. For the circular variable IPC we show sample
+estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.
+The circular concentration lies between 0, meaning the data is not concentrated
+at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
+concentrated at a single point on the circle The population values of these
+parameters are usually, and also in this paper, referred to as $\mu$ (circular
+location) and $\kappa$ (circular concentration). For the linear variables
+(strength IPC and SE) we show sample estimates of the linear mean and standard
+deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
+the linear and circular outcome of the teacher data for teachers with low SE
+(below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
+mean) and high SE (above 1 sd above the mean).
 
@@
 ways of modelling the the linear and circular outcome and thereby illustrate a
-wider range of cylindrical models available in the literature. All four
+wider range of cylindrical models available in the literature. The
 cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
@@
 four cylindrical  models are introduced separately in the subsections below.
-However, to provide a more succinct overview and comparison of the four models,
+However, to provide a more succinct overview and comparison of the four models, Table
 \ref{TableModels} gives an overview of the similarities and differences between
@@
 iterations of the MCMC models to 20,000 and subtracted a burn-in of 5000 to
-reach convergence (the geweke diagnostics show absolute z-scores over 1.96 in
+reach convergence (the Geweke diagnostics show absolute z-scores over 1.96 in
 6\% of the estimated parameters). Note that we choose the same number of
@@
 beneficial for future research in which we may want to investigate the effect of
-further covariates on data from the circumplex. Furthermore, in addition to
-being able to assess the influence of covariates, the cylindrical models also
-provide information about the dependence between the type and strength of
-interpersonal behavior. We found that stronger behavior is associated with those
-types of behavior that show higher scores on the Communion dimension (and also
-on the Agency dimension). This translates to stronger behavior being shown by
-teachers whose type of interpersonal behavior ranges between 0$^\circ$ and
-90$^\circ$, the 'helpful' and 'directing' subtypes according to Figure
-\ref{QTI}. \textcolor{red}{Is dit nieuw/zijn er papers die hier iets over
-zeggen?}\newline
-\indent As mentioned in the introducion data from the interpersonal circumplex
+further covariates on data from the circumplex. \newline
+\indent Furthermore, in addition to being able to assess the influence of
+covariates, the cylindrical models also provide information about the dependence
+between the type and strength of interpersonal behavior. We found that stronger
+behavior is associated with higher scores on the Communion and in some models
+also the Agency dimension. This implies that teachers whose type of
+interpersonal behavior ranges between 0$^\circ$ and 90$^\circ$, the 'helpful'
+and 'directing' subtypes are stronger in their behavior than teachers of the
+other subtypes.\newline
+\indent As mentioned in the introduction, data from the interpersonal circumplex
 is not the only type of cylindrical data that occurs in psychology. The methods

### 2019-01-29 Manuscript/ManuscriptMarkdownCV.tex
Update manuscript

@@
 
-\authornote{Add complete departmental affiliations for each
-author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
+\authornote{
 
 Correspondence concerning this article should be addressed to Jolien
-Cremers, Postal address. E-mail:
+Cremers, . E-mail:
 \href{mailto:joliencremers@gmail.com}{\nolinkurl{joliencremers@gmail.com}}}
@@
 the social sciences data with a circular and linear outcome more
-commonly occur in meteorology meteorology (García-Portugués, Crujeiras,
-\& González-Manteiga, 2013), ecology (García-Portugués, Barros,
-Crujeiras, González-Manteiga, \& Pereira, 2014) or marine research
-(Lagona, Picone, Maruotti, \& Cosoli, 2015) \newline \indent Up until
-now researchers studying cylindrical data had to rely on the linear
-statistical methods to analyze their research results. Focusing on only
-the circular variable or only on the linear variable. However, lately
-more and more of these researchers acknowledge that linear methods are
-not sufficient and call for new methods (Gurtman, 2011; Pennings, 2017b;
-Wright, Pincus, Conroy, \& Hilsenroth, 2009) that take into account both
-the circular and the linear component of these data. \newline
-\indent The aim in the present paper is twofold. Firstly, we intend to
-fill this gap in the literature on cylindrical models by adapting four
-existing cylindrical models in such a way that they include a regression
-of both the linear and circular component of a cylindrical variable onto
-a set of covariates. From now on we will therefore refer to the
-components of the cylindrical variable as outcome components. Secondly,
-we will show how using these models can benefit the analysis of
-circumplex data and cylindrical data in psychology in general. More
-specifically we will show these benefits for interpersonal teacher data
-from the field of educational psychology. Apart from modelling the
-dependence between the linear and circular component of a cylindrical
-variable we would also like to predict the two components from a set of
-covariates in a regression model. In our analysis, we will show how this
-can be done in four different ways using the adapted cylindrical models.
+commonly occur in meteorology (García-Portugués, Crujeiras, \&
+González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
+González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
+Maruotti, \& Cosoli, 2015) \newline \indent Up until now researchers
+studying cylindrical data had to rely on linear statistical methods to
+analyze their research results. However, lately more and more of these
+researchers acknowledge that linear methods are not sufficient and call
+for new methods (Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy,
+\& Hilsenroth, 2009) that take into account both the circular and the
+linear component of these data. \newline \indent The aim in the present
+paper is twofold. Firstly, we we intend to fill the above mentioned gap
+in the literature by showing the use of cylindrical models can benefit
+the analysis of circumplex data and cylindrical data in psychology in
+general. More specifically we will show these benefits for interpersonal
+teacher data from the field of educational psychology. Apart from
+modelling the dependence between the linear and circular component of a
+cylindrical variable we would also like to predict the two components
+from a set of covariates in a regression model. Our second aim therefore
+is to adapt several existing cylindrical models in such a way that they
+include a regression of both the linear and circular component of a
+cylindrical variable onto a set of covariates. From now on we will
+therefore refer to the components of the cylindrical variable as outcome
+components. These adapted circular models are then used to analyse the
+teacher data.
 
@@
 
-Thus, data that consist of a variable and a circular variable is called
+Data that consist of a variable and a circular variable is called
 cylindrical data. A circular variable is different from the linear
@@
 be placed in a two-dimensional space formed by Agency (vertical) and
-Communion (horizontal), see Figure \ref{QTI}. Different parts of this
-space are characterized by different teacher behavior,
-e.g.~\enquote{helpful} or \enquote{uncertain}. This two-dimensional
-space is called the interpersonal circle/circumplex (IPC). The IPC is
-``a continuous order with no beginning or end'' (Gurtman, 2009, p. 2).
-We call such ordering a circumplex ordering and the IPC is therefore
-often called the interpersonal circumplex. The ordering also implies
-that scores on the IPC could be viewed as a circular variable. This
-circular variable represents the type of interpersonal behavior that a
-teacher shows towards his/her students.\newline \indent Cremers et al.
-(2018a) explain the circular nature of the IPC data and analyze them as
-such using a circular regression model. The two dimension scores, Agency
-and Communion, can be converted to a circular score using the
-two-argument arctangent function in
-\eqref{PredVal}\footnote{The selection of
+Communion (horizontal), see Figure \ref{QTI}. This space is called the
+interpersonal circle/circumplex (IPC) and different parts of this space
+are characterized by different teacher behavior, e.g.~\enquote{helpful}
+or \enquote{uncertain}. The IPC is ``a continuous order with no
+beginning or end'' (Gurtman, 2009, p. 2). We call such ordering a
+circumplex ordering and the IPC is therefore often called the
+interpersonal circumplex. The ordering also implies that scores on the
+IPC could be viewed as a circular variable. This circular variable
+represents the type of interpersonal behavior that a teacher shows
+towards his/her students.\newline \indent Cremers et al. (2018a) explain
+the circular nature of the IPC data and analyze them as such using a
+circular regression model. The two dimension scores, Agency and
+Communion, can be converted to a circular score using the two-argument
+arctangent function in \eqref{PredVal}\footnote{The selection of
 the origin in circumplex data depends on the scaling of the Agency and Communion
-scores; their respective 0 scores form the origin. Agency and Communion are
-measured on a scale from 1 to 5 and for analysis purposes they are later
-converted to scale ranging from -1 to 1. Scaling is however only considered an
-issue in those instances where cylindrical data is derived from measurements in
+scores. Agency and Communion are measured on a scale from 1 to 5 and for
+analysis purposes they are later converted to scale ranging from -1 to 1.  Their
+respective 0 scores form the origin. Scaling is however only considered an issue
+in those instances where cylindrical data is derived from measurements in
 bivariate space.}, where \(A\) represents a score on the Agency
@@
 analyse the blend of the two dimensions Agency and Communion instead of
-both dimensions separately providing us with new insights compared to a
-separate analysis of the two dimensions that is standard in the
-literature (see e.g.~Pennings et al. (2018), Wright et al. (2009) or
+both dimensions separately. This provides us with new insights compared
+to a separate analysis of the two dimensions that is standard in the
+literature (see e.g. Pennings et al. (2018), Wright et al. (2009) or
 Wubbels et al. (2006)). There is however one main drawback: when
@@
 covariate in the analysis. In previous research, in psychology and
-education it had been shown that higher self-efficacy is related to the
+education it has been shown that higher self-efficacy is related to the
 quality of interpersonal interactions (Locke \& Sadler, 2007;
-\VANDER{Want}{Van der}{van der} Want et al., n.d.). After listwise
+\VANDER{Want}{Van der}{van der} Want et al., 2018). After listwise
 deletion of missings (\(3\) in total, only for the self-efficacy) we
@@
 IPC and SE) we show sample estimates of the linear mean and standard
-deviation. Figure \ref{dataplot} is a scatterplot showing the relation
-between the linear and circular outcome of the teacher data for teachers
-with low SE (below 1 sd below the mean), average SE (between 1 sd below
-and 1 sd above the mean) and high SE (above 1 sd above the mean).
+deviation (sd). Figure \ref{dataplot} is a scatterplot showing the
+relation between the linear and circular outcome of the teacher data for
+teachers with low SE (below 1 sd below the mean), average SE (between 1
+sd below and 1 sd above the mean) and high SE (above 1 sd above the
+mean).
 
@@
 circular outcome and thereby illustrate a wider range of cylindrical
-models available in the literature. All four cylindrical models contain
-a set of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\)
+models available in the literature. The cylindrical models contain a set
+of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\)
 predictors \(\boldsymbol{z} = z_1, \dots, z_q\) for the linear and
@@
 separately in the subsections below. However, to provide a more succinct
-overview and comparison of the four models, \ref{TableModels} gives an
-overview of the similarities and differences between the models.
+overview and comparison of the four models, Table \ref{TableModels}
+gives an overview of the similarities and differences between the
+models.
 
@@
 which we may want to investigate the effect of further covariates on
-data from the circumplex. Furthermore, in addition to being able to
-assess the influence of covariates, the cylindrical models also provide
-information about the dependence between the type and strength of
-interpersonal behavior. We found that stronger behavior is associated
-with those types of behavior that show higher scores on the Communion
-dimension (and also on the Agency dimension). This translates to
-stronger behavior being shown by teachers whose type of interpersonal
-behavior ranges between 0\(^\circ\) and 90\(^\circ\), the
-\enquote{helpful} and \enquote{directing} subtypes according to Figure
-\ref{QTI}.
-\textcolor{red}{Is dit nieuw/zijn er papers die hier iets over
-zeggen?}\newline \indent As mentioned in the introducion data from the
-interpersonal circumplex is not the only type of cylindrical data that
-occurs in psychology. The methods presented in this paper are also of
-use for research on human navigation and eye-tracking research.
-Furthermore, even though cylindrical models are already used in fields
-outside of psychology, the addition of a regression structure to the
-models is of use in these fields as well. \newline \indent In terms of
-interpretability, the CL-PN and Abe-Ley models perform best out of the
-four cylindrical models. In the CL-GPN and GPN-SSN models the
-interpretation of the parameters of the circular outcome component is
-not straightforward, if at all possible. This is caused by the fact that
-in addition to the mean vector the covariance matrix of the GPN
-distribution affects the location of the circular data, making it
-difficult to compute regression coefficients on the circle. Wang and
-Gelfand (2013) state that Monte Carlo integration can be used to compute
-a circular mean and variance for the GPN distribution. In future
-research, this solution might be applied to the methods of Cremers et
-al. (2018b) in order to compute circular coefficients for GPN
-models.\newline \indent In terms of flexibility the GPN-SSN model scores
-best. Multiple linear and circular outcomes can be included and we can
-thus apply the model to multivariate cylindrical data. In addition the
-GPN-SSN, the CL-GPN and CL-PN models are extendable to a mixed-effects
-structure and can thus also be fit to longitudinal data (see
-Nuñez-Antonio and Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et
-al. (2016) for hierarchical/mixed-effects models for the PN and GPN
-distributions respectively). For the Abe-Ley model this may also be
-possible but has not been done in previous research for the conditional
-distribution of its circular outcome (sine-skewed von Mises). Concerning
-asymmetry, both the GPN-SSN as well as the Abe-Ley model allow for
-non-symmetrical shapes of the distributions of both the linear and
-circular outcome, while the CL-GPN model permits an asymmetric circular
-outcome.\newline \indent The four cylindrical models that were modified
-to the regression context in this paper are not the only cylindrical
-distributions available from the literature. Other interesting
-cylindrical distributions have been introduced by Fernández-Durán
-(2007), Kato and Shimizu (2008) and Sugasawa (2015) (for more references
-we refer to Chapter 2 of Ley and Verdebout (2017)). In the present study
-we have decided not to include these distributions for reasons of space,
-complexity of the models and ease of implementing a regression
-structure. In future research however it would be interesting to
-investigate other types of cylindrical distributions as well in order to
-compare the interpretability, flexibility and model fit to the models
-developed in the present study.\newline
+data from the circumplex. \newline \indent Furthermore, in addition to
+being able to assess the influence of covariates, the cylindrical models
+also provide information about the dependence between the type and
+strength of interpersonal behavior. We found that stronger behavior is
+associated with higher scores on the Communion and in some models also
+the Agency dimension. This implies that teachers whose type of
+interpersonal behavior ranges between 0\(^\circ\) and 90\(^\circ\), the
+\enquote{helpful} and \enquote{directing} subtypes are stronger in their
+behavior than teachers of the other subtypes.\newline \indent As
+mentioned in the introduction, data from the interpersonal circumplex is
+not the only type of cylindrical data that occurs in psychology. The
+methods presented in this paper are also of use for research on human
+navigation and eye-tracking research. Furthermore, even though
+cylindrical models are already used in fields outside of psychology, the
+addition of a regression structure to the models is of use in these
+fields as well. \newline \indent In terms of interpretability, the CL-PN
+and Abe-Ley models perform best out of the four cylindrical models. In
+the CL-GPN and GPN-SSN models the interpretation of the parameters of
+the circular outcome component is not straightforward, if at all
+possible. This is caused by the fact that in addition to the mean vector
+the covariance matrix of the GPN distribution affects the location of
+the circular data, making it difficult to compute regression
+coefficients on the circle. Wang and Gelfand (2013) state that Monte
+Carlo integration can be used to compute a circular mean and variance
+for the GPN distribution. In future research, this solution might be
+applied to the methods of Cremers et al. (2018b) in order to compute
+circular coefficients for GPN models.\newline \indent In terms of
+flexibility the GPN-SSN model scores best. Multiple linear and circular
+outcomes can be included and we can thus apply the model to multivariate
+cylindrical data. In addition the GPN-SSN, the CL-GPN and CL-PN models
+are extendable to a mixed-effects structure and can thus also be fit to
+longitudinal data (see Nuñez-Antonio and Gutiérrez-Peña (2014) and
+Hernandez-Stumpfhauser et al. (2016) for hierarchical/mixed-effects
+models for the PN and GPN distributions respectively). For the Abe-Ley
+model this may also be possible but has not been done in previous
+research for the conditional distribution of its circular outcome
+(sine-skewed von Mises). Concerning asymmetry, both the GPN-SSN as well
+as the Abe-Ley model allow for non-symmetrical shapes of the
+distributions of both the linear and circular outcome, while the CL-GPN
+model permits an asymmetric circular outcome.\newline \indent The four
+cylindrical models that were modified to the regression context in this
+paper are not the only cylindrical distributions available from the
+literature. Other interesting cylindrical distributions have been
+introduced by Fernández-Durán (2007), Kato and Shimizu (2008) and
+Sugasawa (2015) (for more references we refer to Chapter 2 of Ley and
+Verdebout (2017)). In the present study we have decided not to include
+these distributions for reasons of space, complexity of the models and
+ease of implementing a regression structure. In future research however
+it would be interesting to investigate other types of cylindrical
+distributions as well in order to compare the interpretability,
+flexibility and model fit to the models developed in the present
+study.\newline
 
@@
 \VANDER{Want}{Van der}{van der} Want, A. C., Den Brok, P., Beijaard, D.,
-Brekelmans, M., Claessens, L. C., \& Pennings, H. J. M. (n.d.). The
+Brekelmans, M., Claessens, L. C., \& Pennings, H. J. M. (2018). The
 relation between teachers' interpersonal role identity and their
 self-efficacy, burnout and work engagement. \emph{Professional
-Development in Education}.
+Development in Education}, Advance Online Publication.
 doi:\href{https://doi.org/10.1080/19415257.2018.1511453}{10.1080/19415257.2018.1511453}

### 2019-02-02 Manuscript/Manuscript.Rmd
R-code update

--- 2019-01-29 Manuscript/ManuscriptMarkdownCV.Rmd
@@
 keywords          : "cylindrical data, regression, interpersonal behavior"
-wordcount         : "X"
+#wordcount         : "X"
 
@@
 #kable(PLSL, "latex", booktabs = T, escape = FALSE,
-#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four cylindrical models")%>%
+#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in #the four cylindrical models")%>%
 #add_header_above(c("Model" = 1, "Circular" = 2, "Linear" = 2))
@@
 CL-PN & 82.96 & (9.47) & -17.65 & (3.70)\\
-CL-GPN & 86.05 & (16.63) & -18.30 & (3.00)\\
+CL-GPN & 78.21 & (14.53) & -18.30 & (3.00)\\
 Abe-Ley & 31.97 & (22.07) & 25.49 & (17.46)\\

### 2019-02-02 Manuscript/R-code/Posterior Sampling CL-GPN.R

@@ -190,13 +190,13 @@ llik.circ <- function(theta, Z, beta.I, beta.II, r, tau, Sigma){
   mu <- cbind(mub1, mub2)
   u <- cbind(cos(theta), sin(theta))
   
-  ll <- 0
+  ll <- length(theta)*(log(1) - log(2*pi*sqrt(tau)))
   
   for(i in 1: length(theta)){
     
     Dbd <- cbind(r[i]*cos(theta[i]) - mub1[i], r[i]*sin(theta[i]) - mub2[i])
     
-    ll <- ll + log(r[i]) + log(1) - log(2*pi+sqrt(tau)) - (Dbd%*%solve(Sigma)%*%t(Dbd))/(2*tau)
+    ll <- ll + log(r[i]) - (Dbd%*%solve(Sigma)%*%t(Dbd))/(2*tau)
 
   }
   
@@ -240,13 +240,6 @@ llik.circ <- function(theta, Z, beta.I, beta.II, r, tau, Sigma){
 # theta_pred.max = posterior predictive values for theta in the holdout set (using the maximum covariate value)
 # theta_pred.mean = posterior predictive values for theta in the holdout set (using the mean covariate value)
 # theta_pred.median = posterior predictive values for theta in the holdout set (using the median covariate value)
-
-return(list("Gamma"= Gamma, "BI" = BI, "BII" = BII, "Sig" = Sig, "Sigma" = Sigma,
-            "ll.circ" = ll.circ, "ll.lin" = ll.lin,
-            "theta_pred" = theta_pred, "y_pred" = y_pred,
-            "theta_pred.hold" = theta_pred.hold, "y_pred.hold" = y_pred.hold,
-            "theta_pred.min" = theta_pred_min,  "theta_pred.max" = theta_pred_max,
-            "theta_pred.mean" = theta_pred_mean, "theta_pred.meadian" = theta_pred_median))
 
 CLGPN <- function(theta, y, X, Z, its, p, theta.hold, y.hold, X.hold, Z.hold){
   

### 2019-02-02 Manuscript/R-code/Posterior Sampling CL-PN.R

@@ -121,13 +121,13 @@ llik.circ <- function(theta, ZI, ZII, beta.I, beta.II, r){
   mub2 <- ZII%*%beta.II
   
   mu <- cbind(mub1, mub2)
-  ll <- 0
+  ll <- length(theta) * (log(1) - log(2*pi))
   
   for(i in 1: length(theta)){
     
     Dbd <- cos(theta[i])*mub1[i] + sin(theta[i])*mub2[i]
     
-    ll <- ll + log(1) - log(2*pi) + (-t(mu[i,])%*%mu[i,])/2 + (-r[i]^2 + 2*r[i]*Dbd)/2
+    ll <- ll + (-t(mu[i,])%*%mu[i,])/2 + 0.5*(-r[i]^2 + 2*r[i]*Dbd)
     
   }
  

### 2019-02-02 Manuscript/Manuscript.tex
Manuscript update

--- 2019-01-29 Manuscript/ManuscriptMarkdownCV.tex
@@
 \textsuperscript{1} Department of Methodology and Statistics, Utrecht University\\\textsuperscript{2} TNO\\\textsuperscript{3} Department of Education, Utrecht University\\\textsuperscript{4} Department of Applied Mathematics, Computer Science and Statistics, Ghent University}
-\keywords{cylindrical data, regression, interpersonal behavior\newline\indent Word count: X}
+\keywords{cylindrical data, regression, interpersonal behavior}
 \usepackage{csquotes}
@@
 the number of iterations of the MCMC models to 20,000 and subtracted a
-burn-in of 5000 to reach convergence (the geweke diagnostics show
+burn-in of 5000 to reach convergence (the Geweke diagnostics show
 absolute z-scores over 1.96 in 6\% of the estimated parameters). Note

### 2019-02-02 Manuscript/SupplementaryMaterial.Rmd
Manuscript update

@@
 \begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\mu^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(u^t\mu) \right]\},
 \end{equation}
@@
 \begin{equation}\label{gpnreg}
-f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
+f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2\tau^2}\right],
 \end{equation}
@@
 
-\indent We use the loglikelihoods of the following conditional densities for the
+\indent We use the following loglikelihoods for the
 computation of the PLSL in the teacher data:
@@
 
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.
+$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$
+
+$$l(\theta) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
+where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
 
@@
 
-$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.
+$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$
+
+$$l(\theta) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
+where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
 
@@
 
-$y_i \mid \theta_i, \nu_i, \mu_i, \kappa, \alpha \sim W\left(\nu_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.
+$$l(y) = \log\alpha + \sum\log h_i^\alpha + \sum\log y_i^{\alpha - 1} - \sum(h_iy_i)^\alpha$$
 
-$\theta_i \mid y_i, \nu_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\nu_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.
+where $h_i = \exp\{\hat{y}_i\}\{1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i)\}^{1/\alpha}$, $\hat{y}_i = \gamma_0 + \gamma_1\text{SE}_i$ and $\hat{\theta}_i = \beta_0 +  2\tan^{-1}(\beta_1\text{SE}_i))$ .
 
-\item For the modified joint projected and skew normal model:
+$$l(\theta) = \log(1) - \sum\log 2\pi I_0(c_i) + \sum\log\{1 + \lambda \sin(\theta_i - \hat{\theta}_i\} + \sum c_i\cos(\theta_i - \hat{\theta}_i)$$
+
+where $c_i = y_i^{\alpha}\exp(\hat{y}_i)^{\alpha}\tanh\kappa$, and $I_{0}$ is a Bessel function of order $0$.
+
+\item For the modified joint projected and skew normal model we take the loglikelihoods of the following distributions:
 
@@
 where $SSN$ is the skew normal distribution.
+Computationally this comes down to taking the log of the density values for a univariate and multivariate normal distribution (with mean and variance specified as above) for the linear and circular outcome respectively.
+
 \end{itemize}

### 2019-02-02 Manuscript/SupplementaryMaterial.tex
Manuscript update

@@
 \begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\mu^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(u^t\mu) \right]\},
 \end{equation} \noindent where
@@
 distribution equals \begin{equation}\label{gpnreg}
-f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
+f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2\tau^2}\right],
 \end{equation} \noindent where we recall that
@@
 
-\indent We use the loglikelihoods of the following conditional densities
-for the computation of the PLSL in the teacher data:
+\indent We use the following loglikelihoods for the computation of the
+PLSL in the teacher data:
 

### 2019-02-02 Manuscript/Manuscript.Rmd
Manuscript Update

@@
   - \DeclareRobustCommand{\VANDER}[3]{#2} 
+  - \DeclareRobustCommand{\VAN}[3]{#2}
+  - \DeclareRobustCommand{\DEN}[3]{#2}
   - \usepackage{multirow}

### 2019-02-02 Manuscript/Manuscript.tex
Manuscript Update

@@
 \DeclareRobustCommand{\VANDER}[3]{#2}
+\DeclareRobustCommand{\VAN}[3]{#2}
+\DeclareRobustCommand{\DEN}[3]{#2}
 \usepackage{multirow}
@@
 circle/circumplex (Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans,
-Brok, \& Tartwijk, 2006), the circumplex of affect (Russell, 1980), the
-circumplex of human emotion (Plutchik, 1997) or the model of human
-values (Schwartz, 1992). \newline \indent Also, many of the more recent
-types of data that are studied in psychology are cylindrical. For
-example, research on human navigation uses data where distance (i.e.,
-linear component) and direction (i.e., the linear component) are of
-interest (Chrastil \& Warren, 2017) or in eye-tracking, the saccade data
-also consist of both the direction (i.e., the circular variable) and the
-duration (i.e., the linear variable)(e.g., Rayner (2009)). Apart from
-the social sciences data with a circular and linear outcome more
-commonly occur in meteorology (García-Portugués, Crujeiras, \&
-González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
-González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
-Maruotti, \& Cosoli, 2015) \newline \indent Up until now researchers
-studying cylindrical data had to rely on linear statistical methods to
-analyze their research results. However, lately more and more of these
-researchers acknowledge that linear methods are not sufficient and call
-for new methods (Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy,
-\& Hilsenroth, 2009) that take into account both the circular and the
-linear component of these data. \newline \indent The aim in the present
-paper is twofold. Firstly, we we intend to fill the above mentioned gap
-in the literature by showing the use of cylindrical models can benefit
-the analysis of circumplex data and cylindrical data in psychology in
-general. More specifically we will show these benefits for interpersonal
-teacher data from the field of educational psychology. Apart from
-modelling the dependence between the linear and circular component of a
-cylindrical variable we would also like to predict the two components
-from a set of covariates in a regression model. Our second aim therefore
-is to adapt several existing cylindrical models in such a way that they
-include a regression of both the linear and circular component of a
-cylindrical variable onto a set of covariates. From now on we will
-therefore refer to the components of the cylindrical variable as outcome
-components. These adapted circular models are then used to analyse the
-teacher data.
+\DEN{Brok}{Den}{den} Brok, \& \VAN{Tartwijk}{Van}{van} Tartwijk, 2006),
+the circumplex of affect (Russell, 1980), the circumplex of human
+emotion (Plutchik, 1997) or the model of human values (Schwartz, 1992).
+\newline \indent Also, many of the more recent types of data that are
+studied in psychology are cylindrical. For example, research on human
+navigation uses data where distance (i.e., linear component) and
+direction (i.e., the linear component) are of interest (Chrastil \&
+Warren, 2017) or in eye-tracking, the saccade data also consist of both
+the direction (i.e., the circular variable) and the duration (i.e., the
+linear variable)(e.g., Rayner (2009)). Apart from the social sciences
+data with a circular and linear outcome more commonly occur in
+meteorology (García-Portugués, Crujeiras, \& González-Manteiga, 2013),
+ecology (García-Portugués, Barros, Crujeiras, González-Manteiga, \&
+Pereira, 2014) or marine research (Lagona, Picone, Maruotti, \& Cosoli,
+2015) \newline \indent Up until now researchers studying cylindrical
+data had to rely on linear statistical methods to analyze their research
+results. However, lately more and more of these researchers acknowledge
+that linear methods are not sufficient and call for new methods
+(Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy, \& Hilsenroth,
+2009) that take into account both the circular and the linear component
+of these data. \newline \indent The aim in the present paper is twofold.
+Firstly, we we intend to fill the above mentioned gap in the literature
+by showing the use of cylindrical models can benefit the analysis of
+circumplex data and cylindrical data in psychology in general. More
+specifically we will show these benefits for interpersonal teacher data
+from the field of educational psychology. Apart from modelling the
+dependence between the linear and circular component of a cylindrical
+variable we would also like to predict the two components from a set of
+covariates in a regression model. Our second aim therefore is to adapt
+several existing cylindrical models in such a way that they include a
+regression of both the linear and circular component of a cylindrical
+variable onto a set of covariates. From now on we will therefore refer
+to the components of the cylindrical variable as outcome components.
+These adapted circular models are then used to analyse the teacher data.
 
@@
 
+\DeclareRobustCommand{\VANDER}[3]{#3}
+\DeclareRobustCommand{\VAN}[3]{#3}
+\DeclareRobustCommand{\DEN}[3]{#3}
+
 \hypertarget{refs}{}
@@
 Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C.,
-\VANDER{Want}{Van der}{van der} Want, A. C., \& Tartwijk, J. van.
-(2018). Interpersonal adaptation in teacher-student interaction.
-\emph{Learning and Instruction}, \emph{55}, 41--57.
+\VANDER{Want}{Van der}{van der} Want, A. C., \& \VAN{Tartwijk}{Van}{van}
+Tartwijk, J. (2018). Interpersonal adaptation in teacher-student
+interaction. \emph{Learning and Instruction}, \emph{55}, 41--57.
 doi:\href{https://doi.org/10.1016/j.learninstruc.2017.09.005}{10.1016/j.learninstruc.2017.09.005}
@@
 \leavevmode\hypertarget{ref-want2018selfefficacy}{}%
-\VANDER{Want}{Van der}{van der} Want, A. C., Den Brok, P., Beijaard, D.,
-Brekelmans, M., Claessens, L. C., \& Pennings, H. J. M. (2018). The
-relation between teachers' interpersonal role identity and their
-self-efficacy, burnout and work engagement. \emph{Professional
+\VANDER{Want}{Van der}{van der} Want, A. C., \DEN{Brok}{Den}{den} Brok,
+P., Beijaard, D., Brekelmans, M., Claessens, L. C., \& Pennings, H. J.
+M. (2018). The relation between teachers' interpersonal role identity
+and their self-efficacy, burnout and work engagement. \emph{Professional
 Development in Education}, Advance Online Publication.
@@
 \leavevmode\hypertarget{ref-wubbels2006interpersonal}{}%
-Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
-Handbook of classroom management: Research practice and contemporary
-issues. In C. Evertson \& C. S. Weinstein (Eds.), \emph{Handbook of
-classroom management: Research, practice, and contemporary issues} (pp.
-1161--1191). Malwah, NJ: Lawrence Erlbaum Associates.
+Wubbels, T., Brekelmans, M., \DEN{Brok}{Den}{den} Brok, P., \&
+\VAN{Tartwijk}{Van}{van} Tartwijk, J. (2006). Handbook of classroom
+management: Research practice and contemporary issues. In C. Evertson \&
+C. S. Weinstein (Eds.), \emph{Handbook of classroom management:
+Research, practice, and contemporary issues} (pp. 1161--1191). Malwah,
+NJ: Lawrence Erlbaum Associates.
 

### 2019-03-03 Manuscript/Manuscript.Rmd
Manuscript Update

@@
 as vectors, with a directional measure (i.e., the circular component) and a
-measure indicating the magnitude (i.e., linear component). Many established
+measure indicating the magnitude (i.e., the linear component). Many established
 models in psychology are often referred to as circular or circumplex models, but
@@
 cylindrical. For example, research on human navigation uses data where distance
-(i.e., linear component) and direction (i.e., the linear component) are of
+(i.e., linear component) and direction (i.e., the circular component) are of
 interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
 consist of both the direction (i.e., the circular variable) and the duration
-(i.e., the linear variable)(e.g., @rayner200935th). Apart from the social
-sciences data with a circular and linear outcome more commonly occur in
+(i.e., the linear variable) (e.g., @rayner200935th). Apart from the social
+sciences, data with a circular and linear outcome more commonly occur in
 meteorology [@garcia2013exploring], ecology [@garcia2014test] or
@@
 \newline
-\indent The aim in the present paper is twofold. Firstly, we we intend to fill
-the above mentioned gap in the literature by showing the use of cylindrical
+\indent The aim in the present paper is twofold. Firstly, we intend to fill
+the above mentioned gap in the literature by showing that the use of cylindrical
 models can benefit the analysis of circumplex data and cylindrical data in
@@
 of covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. These adapted circular models are
+cylindrical variable as outcome components. These adapted cylindrical models are
 then used  to analyse the teacher data.
@@
 
-Data that consist of a variable and a circular variable is called
+Data that consist of a linear variable and a circular variable is called
 cylindrical data. A circular variable is different from the linear variable in
@@
 scores, Agency and Communion, can be converted to a circular score using the
-two-argument arctangent function in \eqref{PredVal}\footnote{The selection of
+two-argument arctangent function in \eqref{PredVal}, where \(A\) represents a score on the Agency dimension and
+\(C\) represents a score on the Communion dimension \footnote{The selection of
 the origin in circumplex data depends on the scaling of the Agency and Communion
@@
 in those instances where cylindrical data is derived from measurements in
-bivariate space.}, where \(A\) represents a score on the Agency dimension and
-\(C\) represents a score on the Communion dimension. Note that when placing a
+bivariate space.}. Note that when placing a
 unit circle on Figure \ref{QTI} we see that the Agency dimension is related to
@@
 at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
-concentrated at a single point on the circle The population values of these
+concentrated at a single point on the circle. The population values of these
 parameters are usually, and also in this paper, referred to as $\mu$ (circular
@@
 regression structure could be incorporated. But also because they show different
-ways of modelling the the linear and circular outcome and thereby illustrate a
+ways of modelling the linear and circular outcome and thereby illustrate a
 wider range of cylindrical models available in the literature. The
 cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
-x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_q$ for the linear and
+x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_p$ for the linear and
 circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
@@
 $\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
-$\:\:$Shape & symmetric, & assymetric, & assymetric, & assymetric, \\
+$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
             & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
@@
 $\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
-$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(0, + \infty)$\\
-$\:\:$Shape & symmetric, & symmetric, & assymetric, & assymetric, \\
+$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
+$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
             & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
@@
 in the middle plot. We call $\boldsymbol{S}$ the augmented representation of the
-circular outcome, it is a variable that in contrast to \(\Theta\) is not
+circular outcome. It is a variable that in contrast to \(\Theta\) is not
 observed and thus considered latent or auxiliary. This then means that we do not
 model \(\Theta\) directly but indirectly through $\boldsymbol{S}$. \newline
-\indent For both the PN and GPN distribution the circular location parameter
+\indent For both the PN and GPN distributions the circular location parameter
 \(\mu\in [0, 2\pi)\) is modeled as \(\hat{\mu}_i =
@@
 matrix and the circular predicted values should be computed using numerical
-integretion or Monte Carlo methods because a closed form expression for the mean
+integration or Monte Carlo methods because a closed form expression for the mean
 direction is not available.} where $\boldsymbol{\beta}^{I}$ and
@@
 the regression line starts flattening off (indicated with a square in Figure
-\ref{circregline}). By default, the parameters from the PN and GPN model do not
+\ref{circregline}). By default, the parameters from the PN and GPN models do not
 directly describe this inflection point. For the PN distribution however,
@@
 for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2
-& \xi\\ \xi & 1 \end{bmatrix}\), allowing for multimodality and
-assymetry/skewness.\newline
+& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
+asymetry/skewness.\newline
 \indent For the teacher dataset the predicted linear outcome, strengths of
@@
 the effect of the sine and cosine of the type of behavior which are related to
-the scores on the Agency an Communion dimensions respectively.\newline
+the scores on the Agency and Communion dimensions respectively.\newline
 \indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
@@
   \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to.
-The square indicates the inflection point of the regression live.}
+The square indicates the inflection point of the regression line.}
   \label{circregline}
@@
 $\gamma_0$ as well as the effect of self-efficacy on type and strength of
-teacher behavior through the parameters, $\beta_{1}$ and $\gamma_1$.\newline
+teacher behavior through the parameters $\beta_{1}$ and $\gamma_1$.\newline
 \indent We can use numerical optimization (Nelder-Mead) to find solutions for
@@
   \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
-  distribution [@sahu2003new]. Thus also in this model the linear outcome is
-  defined constrained to be positive. Because the GPN distribution is modelled
+  distribution [@sahu2003new]. Because the GPN distribution is modelled
   using a so-called augmented representation (see also the description of the
@@
   modelling the multivariate skew normal distribution. As in
-@mastrantonio2018joint dependence between the linear and circular outcome is
+@mastrantonio2018joint, dependence between the linear and circular outcome is
   created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
@@
 \beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
-The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
+The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
 linear outcomes in the GPN-SSN model are
@@
 $\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
-on type and strength of teacher behavior through the parameters,
+on type and strength of teacher behavior through the parameters
 $\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
 because the type and strength of interpersonal behavior are modelled together
-using a multivariate normal distribution we can through its variance-covariance
-matrix also assess the dependence between the type and strength of interpersonal
+using a multivariate normal distribution we can, through its variance-covariance
+matrix, also assess the dependence between the type and strength of interpersonal
 behavior. \newline
@@
 6\% of the estimated parameters). Note that we choose the same number of
-iterations for all three models models estimated using MCMC prodedures to make
+iterations for all three models estimated using MCMC procedures to make
 their comparison via the PLSL as fair as possible. Lastly, the predictor
@@
 the results for the four cylindrical models that were fit to the teacher data.
-For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN we show
+For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN, we show
 descriptives of the posterior of the estimated parameters (posterior mode and
@@
 \multicolumn{7}{l}{type and strength of interpersonal behavior. $\gamma_{cos}$ and $\gamma_{sin}$ inform us about the dependence}\\
-\multicolumn{7}{l}{ between the type and strength of interpersonal behavior. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are are}\\
+\multicolumn{7}{l}{ between the type and strength of interpersonal behavior. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are}\\
 \multicolumn{7}{l}{elements of the variance-covariance matrix of the type of interpersonal behavior in the}\\
@@
 Mean & 0.36 & -0.03 & 1.17 & 0.04 & 3.66 & 1.51 & 0.70 \\
-SD & 0.02) & 0.01 & 0.02 & 0.02 & 0.12 & 0.08 & 0.05\\
+SD & 0.02 & 0.01 & 0.02 & 0.02 & 0.12 & 0.08 & 0.05\\
 \bottomrule
@@
 \multicolumn{7}{l}{on the type and strength of interpersonal behavior. $\sum_{s_{1,1}}$, $\sum_{s_{1,2}}$, $\sum_{s_{2,2}}$, $\sum_{y_{3,3}}$, $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ }\\
-\multicolumn{7}{l}{are elements of the variance-covariance matrix of which $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ inform us about}\\
+\multicolumn{7}{l}{are elements of the variance-covariance matrix of which $\sum_{sy_{1,3}}$ and $\sum_{sy_{2,3}}$ inform us about}\\
 \multicolumn{7}{l}{the dependence between the type and strength of interpersonal behavior.}\\
@@
 interpersonal behavior at the average self-efficacy. For the CL-PN, CL-GPN and
-GPN-SSN model the parameters are estimated at 0.38, 0.37 and 0.30 respectively
+GPN-SSN models the parameters are estimated at 0.38, 0.37 and 0.33 respectively
 and are a direct prediction of the strength of interpersonal behavior at the
@@
 
-The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley model and $\beta_{1_y}$
+The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley models and $\beta_{1_y}$
 in the GPN-SSN model inform us about the effect of self-efficacy on the strength
@@
 say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to base our
-conclusions on. The model CL-GPN and CL-PN fit the linear outcome best according
+conclusions on. The models CL-GPN and CL-PN fit the linear outcome best according
 to the model fit in Table \ref{tab:ModelFit}. In these models the linear outcome has a
@@
 self-efficacy on the type of interpersonal behavior in the CL-PN, CL-GPN,
-Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley model we have
+Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley models we have
 drawn the circular regression lines for this effect in Figure \ref{regline} (see
-the description of the CL-PN and CL-GPN models). For the CL-PN model the
-inflection point is indicated with a square in Figure \ref{regline}. The
-inflection point for the Abe-Ley model falls outside the bounds of the plot and
-is therefore not displayed. The slope at the inflection point, $b_c$, for the
-CL-PN model is computed by using methods from @CremersMulderKlugkist2017 and is
-equal to 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation
-to the circular regression line representing the slope at a specific point.
-Therefore it is possible for the HPD interval to be wider than $2\pi$. In this
-case the interval is much wider and covers 0, indicating there is no evidence
-for an effect.} The parameter $\beta_1$ is the slope at the inflection point for
-the Abe-Ley model and is equal to -0.03. Even though these slopes are quite
+the description of the CL-PN and CL-GPN models for a detailed explanation of
+circular regression lines). For the CL-PN model the inflection point is
+indicated with a square in Figure \ref{regline}. The inflection point for the
+Abe-Ley model falls outside the bounds of the plot and is therefore not
+displayed. The slope at the inflection point, $b_c$, for the CL-PN model is
+computed by using methods from @CremersMulderKlugkist2017 and is equal to 1.67
+(-24.66, 29.33)\footnote{Note that this is a linear approximation to the
+circular regression line representing the slope at a specific point. Therefore
+it is possible for the HPD interval to be wider than $2\pi$. In this case the
+interval is much wider and covers 0, indicating there is no evidence for an
+effect.}. The parameter $\beta_1$ is the slope at the inflection point for the
+Abe-Ley model and is equal to -0.03. Even though these slopes are
 different, the regression lines in Figure \ref{regline} are quite similar in the
@@
 \newline
-\indent In the CL-GPN and GPN-SSN model we cannot compute circular regression
+\indent In the CL-GPN and GPN-SSN models we cannot compute circular regression
 coefficients due to the fact that not only the mean vector of the GPN
@@
 44.49^\circ)$. For the GPN-SSN model the modes and 95\% HPD intervals of the
-posterior predictive distributiona are \(\hat{\theta}_{SE_{min}} = 206.87^\circ
+posterior predictive distributions are \(\hat{\theta}_{SE_{min}} = 206.87^\circ
 (117.12^\circ, \: 72.02^\circ)\), \(\hat{\theta}_{SE_{median}} = 24.68^\circ
@@
 The relation between the type and strength of interpersonal behavior in the
-CL-PN and CL-GPN model, is described by the parameters $\gamma_{\cos}$ and
+CL-PN and CL-GPN models is described by the parameters $\gamma_{\cos}$ and
 $\gamma_{\sin}$. The HPD interval of \(\gamma_{\cos}\) does not include 0 for
@@
 meaning. This is illustrated in Figure \ref{QTI}. In a unit circle the
-horizontal axis (Communion) represents the cosine of and the vertical axis
+horizontal axis (Communion) represents the cosine and the vertical axis
 (Agency) represents the sine of an angle. For the teacher data this means that
@@
 component, and thus the correlation with the Communion dimension, is larger.
-This mean that teachers scoring both high on Communion and Agency show stronger
+This means that teachers scoring both high on Communion and Agency show stronger
 behavior. Together with the results from the CL-PN and CL-GPN models in the
@@
 cross-validation estimates is rather large. This means that in some samples, the
-Abe-Ley model shows a lower PLSL value than the average of 25.49\newline
+Abe-Ley model shows a lower PLSL value than the average of 25.49.\newline
 \indent The Abe-Ley model has the best out-of-sample predictive performance for
@@
 linear part is worst we would choose to trust the results for the CL-PN and
-CL-GPN model here. For the circular part however the results of the CL-PN/CL-GPN
-model do not differ as much from the Abe-Ley model and we reach the same
+CL-GPN models here. For the circular part however the results of the CL-PN/CL-GPN
+models do not differ as much from the Abe-Ley model and we reach the same
 conclusion for both models, namely that the effect of self-efficacy on type of

### 2019-03-03 Manuscript/Manuscript.tex
Manuscript Update

@@
   \title{Regression models for Cylindrical data in Psychology}
-    \author{Jolien Cremers\textsuperscript{1}, Helena J.M.
-Pennings\textsuperscript{2,3}, \& Christophe Ley\textsuperscript{4}}
+    \author{Jolien Cremers\textsuperscript{1}, Helena J.M. Pennings\textsuperscript{2,3}, \& Christophe Ley\textsuperscript{4}}
     \date{}
@@
 
-Correspondence concerning this article should be addressed to Jolien
-Cremers, . E-mail:
-\href{mailto:joliencremers@gmail.com}{\nolinkurl{joliencremers@gmail.com}}}
+Correspondence concerning this article should be addressed to Jolien Cremers, . E-mail: \href{mailto:joliencremers@gmail.com}{\nolinkurl{joliencremers@gmail.com}}}
 
@@
 
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-In social sciences the use of cylindrical data is very common. Such data
-consist of a linear and a circular component. Gurtman (2011) refers to
-such data as vectors, with a directional measure (i.e., the circular
-component) and a measure indicating the magnitude (i.e., linear
-component). Many established models in psychology are often referred to
-as circular or circumplex models, but those models are cylindrical.
-Examples of such cylindrical models are the interpersonal
-circle/circumplex (Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans,
-\DEN{Brok}{Den}{den} Brok, \& \VAN{Tartwijk}{Van}{van} Tartwijk, 2006),
-the circumplex of affect (Russell, 1980), the circumplex of human
-emotion (Plutchik, 1997) or the model of human values (Schwartz, 1992).
-\newline \indent Also, many of the more recent types of data that are
-studied in psychology are cylindrical. For example, research on human
-navigation uses data where distance (i.e., linear component) and
-direction (i.e., the linear component) are of interest (Chrastil \&
-Warren, 2017) or in eye-tracking, the saccade data also consist of both
-the direction (i.e., the circular variable) and the duration (i.e., the
-linear variable)(e.g., Rayner (2009)). Apart from the social sciences
-data with a circular and linear outcome more commonly occur in
-meteorology (García-Portugués, Crujeiras, \& González-Manteiga, 2013),
-ecology (García-Portugués, Barros, Crujeiras, González-Manteiga, \&
-Pereira, 2014) or marine research (Lagona, Picone, Maruotti, \& Cosoli,
-2015) \newline \indent Up until now researchers studying cylindrical
-data had to rely on linear statistical methods to analyze their research
-results. However, lately more and more of these researchers acknowledge
-that linear methods are not sufficient and call for new methods
-(Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy, \& Hilsenroth,
-2009) that take into account both the circular and the linear component
-of these data. \newline \indent The aim in the present paper is twofold.
-Firstly, we we intend to fill the above mentioned gap in the literature
-by showing the use of cylindrical models can benefit the analysis of
-circumplex data and cylindrical data in psychology in general. More
-specifically we will show these benefits for interpersonal teacher data
-from the field of educational psychology. Apart from modelling the
-dependence between the linear and circular component of a cylindrical
-variable we would also like to predict the two components from a set of
-covariates in a regression model. Our second aim therefore is to adapt
-several existing cylindrical models in such a way that they include a
-regression of both the linear and circular component of a cylindrical
-variable onto a set of covariates. From now on we will therefore refer
-to the components of the cylindrical variable as outcome components.
-These adapted circular models are then used to analyse the teacher data.
+In social sciences the use of cylindrical data is very common. Such data consist
+of a linear and a circular component. Gurtman (2011) refers to such data
+as vectors, with a directional measure (i.e., the circular component) and a
+measure indicating the magnitude (i.e., the linear component). Many established
+models in psychology are often referred to as circular or circumplex models, but
+those models are cylindrical. Examples of such cylindrical models are the
+interpersonal circle/circumplex
+(Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans, \DEN{Brok}{Den}{den} Brok, \& \VAN{Tartwijk}{Van}{van} Tartwijk, 2006), the circumplex of
+affect (Russell, 1980), the circumplex of human emotion
+(Plutchik, 1997) or the model of human values (Schwartz, 1992).
+\newline
+\indent Also, many of the more recent types of data that are studied in psychology are
+cylindrical. For example, research on human navigation uses data where distance
+(i.e., linear component) and direction (i.e., the circular component) are of
+interest (Chrastil \& Warren, 2017) or in eye-tracking, the saccade data also
+consist of both the direction (i.e., the circular variable) and the duration
+(i.e., the linear variable) (e.g., Rayner (2009)). Apart from the social
+sciences, data with a circular and linear outcome more commonly occur in
+meteorology (García-Portugués, Crujeiras, \& González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras, González-Manteiga, \& Pereira, 2014) or
+marine research (Lagona, Picone, Maruotti, \& Cosoli, 2015)
+\newline
+\indent Up until now researchers studying cylindrical data had to rely on linear
+statistical methods to analyze their research results. However, lately more and more
+of these researchers acknowledge that linear methods are not sufficient and call
+for new methods (Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy, \& Hilsenroth, 2009) that take into account both the circular and the linear
+component of these data.
+\newline
+\indent The aim in the present paper is twofold. Firstly, we intend to fill
+the above mentioned gap in the literature by showing that the use of cylindrical
+models can benefit the analysis of circumplex data and cylindrical data in
+psychology in general. More specifically we will show these benefits for
+interpersonal teacher data from the field of educational psychology. Apart from
+modelling the dependence between the linear and circular component of a
+cylindrical variable we would also like to predict the two components from a set
+of covariates in a regression model. Our second aim therefore is to adapt
+several existing cylindrical models in such a way that they include a regression
+of both the linear and circular component of a cylindrical variable onto a set
+of covariates. From now on we will therefore refer to the components of the
+cylindrical variable as outcome components. These adapted cylindrical models are
+then used to analyse the teacher data.
 
@@
 
-Data that consist of a variable and a circular variable is called
-cylindrical data. A circular variable is different from the linear
-variable in the sense that it is measured on a different scale. Figure
-\ref{circline} shows the difference between a circular scale (right) and
-a linear scale (left). The most important difference is that on a
-circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are
-connected and in fact represent the same number while on a linear scale
-the two ends, \(-\infty\) and \(\infty\), are not connected and
-consequently the values 0\(^\circ\) and 360\(^\circ\) are located on
-different places on the scale. Both circular data and cylindrical data
-require special analysis methods due to this periodicity in the scale of
-a circular variable (see e.g.~Fisher (1995) for an introduction to
-circular data and Mardia and Jupp (2000), Jammalamadaka and Sengupta
-(2001) and Ley and Verdebout (2017) for a more elaborate
-overview).\newline \indent In the literature, several methods have been
-put forward to model the relation between the linear and circular
-component of a cylindrical variable. Some of these are based on
-regressing the linear component onto the circular component using the
-following type of relation: \[y = \beta_0 +
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is
-the linear component and \(\theta\) the circular component (Johnson \&
-Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \&
-Jona-Lasinio, 2015). Others model the relation in a different way,
-e.g.~by specifying a multivariate model for several linear and circular
-variables and modelling their covariance matrix (Mastrantonio, 2018) or
-by proposing a joint cylindrical distribution. For example, Abe and Ley
-(2017) introduce a cylindrical distribution based on a Weibull
-distribution for the linear component and a sine-skewed von Mises
-distribution for the circular component and link these through their
-respective shape and concentration parameters. However, none of the
-methods that have been proposed thus far include additional covariates
-onto which both the circular and linear component are regressed.
+Data that consist of a linear variable and a circular variable is called
+cylindrical data. A circular variable is different from the linear variable in
+the sense that it is measured on a different scale. Figure \ref{circline} shows
+the difference between a circular scale (right) and a linear scale (left). The
+most important difference is that on a circular scale the datapoints 0\(^\circ\)
+and 360\(^\circ\) are connected and in fact represent the same number while on a
+linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
+consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
+places on the scale. Both circular data and cylindrical data require special
+analysis methods due to this periodicity in the scale of a circular variable
+(see e.g.~Fisher (1995) for an introduction to circular data and
+Mardia and Jupp (2000), Jammalamadaka and Sengupta (2001) and Ley and Verdebout (2017) for a more
+elaborate overview).\newline
+\indent In the literature, several methods have been put forward to model the
+relation between the linear and circular component of a cylindrical variable.
+Some of these are based on regressing the linear component onto the circular
+component using the following type of relation: \[y = \beta_0 +
+\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
+linear component and \(\theta\) the circular component (Johnson \& Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \& Jona-Lasinio, 2015). Others model the relation in a
+different way, e.g.~by specifying a multivariate model for several linear and
+circular variables and modelling their covariance matrix
+(Mastrantonio, 2018) or by proposing a joint cylindrical distribution. For
+example, Abe and Ley (2017) introduce a cylindrical distribution based on a
+Weibull distribution for the linear component and a sine-skewed von Mises
+distribution for the circular component and link these through their respective
+shape and concentration parameters. However, none of the methods that have been
+proposed thus far include additional covariates onto which both the circular and
+linear component are regressed.
 
@@
 
-The motivating example for this article comes from the field of
-educational psychology and was collected for the studies on classroom
-climate of \VANDER{Want}{Van der}{van der} Want (2015), Claessens (2016)
-and Pennings (2017a). An indicator of the quality of the classroom
-climate is the students' perception of their teachers' interpersonal
-behavior. These interpersonal perceptions, both in educational
-psychology as well as in other areas of psychology, can be measured
-using circumplex measurement instruments (see Horowitz and Strack (2011)
-for an overview of many such instruments).\newline \indent The
-circumplex data used in this paper are measured using the Questionnaire
-on Teacher Interaction (QTI) (Wubbels et al., 2006) which is one such
-circumplex measurement instrument. The QTI is designed to measure
-student perceptions of their teachers' interpersonal behavior and
-contains items that load on two interpersonal dimensions: Agency and
-Communion. Agency refers to the degree of power or control a teacher
-exerts in interaction with his/her students. Communion refers to the
-degree of friendliness or affiliation a teacher conveys in interaction
-with his/her students. The loadings on the two dimensions of the QTI can
-be placed in a two-dimensional space formed by Agency (vertical) and
-Communion (horizontal), see Figure \ref{QTI}. This space is called the
-interpersonal circle/circumplex (IPC) and different parts of this space
-are characterized by different teacher behavior, e.g.~\enquote{helpful}
-or \enquote{uncertain}. The IPC is ``a continuous order with no
-beginning or end'' (Gurtman, 2009, p. 2). We call such ordering a
-circumplex ordering and the IPC is therefore often called the
-interpersonal circumplex. The ordering also implies that scores on the
-IPC could be viewed as a circular variable. This circular variable
-represents the type of interpersonal behavior that a teacher shows
-towards his/her students.\newline \indent Cremers et al. (2018a) explain
-the circular nature of the IPC data and analyze them as such using a
-circular regression model. The two dimension scores, Agency and
-Communion, can be converted to a circular score using the two-argument
-arctangent function in \eqref{PredVal}\footnote{The selection of
+The motivating example for this article comes from the field of educational
+psychology and was collected for the studies on classroom climate of
+\VANDER{Want}{Van der}{van der} Want (2015), Claessens (2016) and Pennings (2017a). An
+indicator of the quality of the classroom climate is the students' perception of
+their teachers' interpersonal behavior. These interpersonal perceptions, both in
+educational psychology as well as in other areas of psychology, can be measured
+using circumplex measurement instruments (see Horowitz and Strack (2011) for an
+overview of many such instruments).\newline
+\indent The circumplex data used in this paper are measured using the
+Questionnaire on Teacher Interaction (QTI) (Wubbels et al., 2006) which is
+one such circumplex measurement instrument. The QTI is designed to measure
+student perceptions of their teachers' interpersonal behavior and contains items
+that load on two interpersonal dimensions: Agency and Communion. Agency refers
+to the degree of power or control a teacher exerts in interaction with his/her
+students. Communion refers to the degree of friendliness or affiliation a
+teacher conveys in interaction with his/her students. The loadings on the two
+dimensions of the QTI can be placed in a two-dimensional space formed by Agency
+(vertical) and Communion (horizontal), see Figure \ref{QTI}. This space is
+called the interpersonal circle/circumplex (IPC) and different parts of this
+space are characterized by different teacher behavior, e.g.~\enquote{helpful} or
+\enquote{uncertain}. The IPC is ``a continuous order with no beginning or end''
+(Gurtman, 2009, p. 2). We call such ordering a circumplex ordering and
+the IPC is therefore often called the interpersonal circumplex. The ordering
+also implies that scores on the IPC could be viewed as a circular variable. This
+circular variable represents the type of interpersonal behavior that a teacher
+shows towards his/her students.\newline
+\indent Cremers et al. (2018a) explain the circular nature of the IPC data and
+analyze them as such using a circular regression model. The two dimension
+scores, Agency and Communion, can be converted to a circular score using the
+two-argument arctangent function in \eqref{PredVal}, where \(A\) represents a score on the Agency dimension and
+\(C\) represents a score on the Communion dimension \footnote{The selection of
 the origin in circumplex data depends on the scaling of the Agency and Communion
@@
 in those instances where cylindrical data is derived from measurements in
-bivariate space.}, where \(A\) represents a score on the Agency
-dimension and \(C\) represents a score on the Communion dimension. Note
-that when placing a unit circle on Figure \ref{QTI} we see that the
-Agency dimension is related to the sine of the circular score and the
-Communion dimension is related to the cosine of the circular score.
+bivariate space.}. Note that when placing a
+unit circle on Figure \ref{QTI} we see that the Agency dimension is related to
+the sine of the circular score and the Communion dimension is related to the
+cosine of the circular score.
 \begin{equation}\label{PredVal}
@@
 \right.
-\end{equation} The resulting circular variable \(\theta\) can then be
-modelled and takes values in the interval \([0, 2\pi)\) or
-\([0^\circ, 360^\circ)\). Note that the round brackets mean that
-\(2\pi\) and \(360^\circ\) are not included in the interval since these
-represent the same value as 0 as a result of periodicity. \newline
-\indent A circular analysis of circumplex data has several benefits: it
-is more in line with its theoretical definition and it allows us to
-analyse the blend of the two dimensions Agency and Communion instead of
-both dimensions separately. This provides us with new insights compared
-to a separate analysis of the two dimensions that is standard in the
-literature (see e.g. Pennings et al. (2018), Wright et al. (2009) or
+\end{equation}
+The resulting circular variable \(\theta\) can then be modelled and takes values
+in the interval \([0, 2\pi)\) or \([0^\circ, 360^\circ)\). Note that the round
+brackets mean that \(2\pi\) and \(360^\circ\) are not included in the interval since
+these represent the same value as 0 as a result of periodicity. \newline
+\indent A circular analysis of circumplex data has several benefits: it is more
+in line with its theoretical definition and it allows us to analyse the blend of
+the two dimensions Agency and Communion instead of both dimensions separately.
+This provides us with new insights compared to a separate analysis of the two
+dimensions that is standard in the literature (see e.g.
+Pennings et al. (2018), Wright et al. (2009) or
 Wubbels et al. (2006)). There is however one main drawback: when
-two-dimensional data are converted to the circle we lose some
-information, namely the length of the two-dimensional vector
-\((A, C)^t\), \emph{i.e.}, its Euclidean norm
-\(\mid\mid (A, C)^t \mid\mid\). This length represents the strength of
-the interpersonal behavior a teacher shows towards his/her students. In
-a cylindrical model this strength (the linear outcome) can be modeled
-together with the type of interpersonal behavior of a teacher (the
-circular outcome). This leads to an improved analysis of interpersonal
-circumplex data, over either analyzing the two dimensions separately or
-using a circular model, because we take all information, circular and
-linear, into account. In the next section we introduce several
-cylindrical models that can be used to analyze the teacher data. First
-however we will provide descriptives for the teacher data.
+two-dimensional data are converted to the circle we lose some information,
+namely the length of the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its
+Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
+strength of the interpersonal behavior a teacher shows towards his/her students.
+In a cylindrical model this strength (the linear outcome) can be modeled
+together with the type of interpersonal behavior of a teacher (the circular
+outcome). This leads to an improved analysis of interpersonal circumplex data,
+over either analyzing the two dimensions separately or using a circular model,
+because we take all information, circular and linear, into account. In the next
+section we introduce several cylindrical models that can be used to analyze the
+teacher data. First however we will provide descriptives for the teacher data.
 
@@
 
-The teacher data was collected between 2010 and 2015 and contains
-several repeated measures on the IPC of 161 teachers. Measurements were
-obtained using the QTI and taken in different years and classes. For
-this paper we only consider one measurement, the first occasion (2010)
-and largest class if data for multiple classes were available. This
-results in a sample of 151 teachers. In addition to the type of
-interpersonal behavior (IPC), the circular outcome, and the strength of
-interpersonal behavior (IPC strength), the linear outcome, a teachers'
-self-efficacy (\verb|SE|) concerning classroom management is used as
-covariate in the analysis. In previous research, in psychology and
-education it has been shown that higher self-efficacy is related to the
-quality of interpersonal interactions (Locke \& Sadler, 2007;
-\VANDER{Want}{Van der}{van der} Want et al., 2018). After listwise
-deletion of missings (\(3\) in total, only for the self-efficacy) we
-have a sample of 148 teachers. Table \ref{Tableteacherdescriptives}
-shows descriptives for the dataset. For the circular variable IPC we
-show sample estimates for the circular mean \(\bar{\theta}\) and
-concentration \(\hat{\rho}\). The circular concentration lies between 0,
-meaning the data is not concentrated at all \emph{i.e.} spread over the
-entire circle, and 1, meaning all data is concentrated at a single point
-on the circle The population values of these parameters are usually, and
-also in this paper, referred to as \(\mu\) (circular location) and
-\(\kappa\) (circular concentration). For the linear variables (strength
-IPC and SE) we show sample estimates of the linear mean and standard
-deviation (sd). Figure \ref{dataplot} is a scatterplot showing the
-relation between the linear and circular outcome of the teacher data for
-teachers with low SE (below 1 sd below the mean), average SE (between 1
-sd below and 1 sd above the mean) and high SE (above 1 sd above the
-mean).
+The teacher data was collected between 2010 and 2015 and contains several
+repeated measures on the IPC of 161 teachers. Measurements were obtained using
+the QTI and taken in different years and classes. For this paper we only
+consider one measurement, the first occasion (2010) and largest class if data
+for multiple classes were available. This results in a sample of 151 teachers.
+In addition to the type of interpersonal behavior (IPC), the circular outcome,
+and the strength of interpersonal behavior (IPC strength), the linear outcome, a
+teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
+covariate in the analysis. In previous research, in psychology and education it
+has been shown that higher self-efficacy is related to the quality of
+interpersonal interactions (Locke \& Sadler, 2007; \VANDER{Want}{Van der}{van der} Want et al., 2018).
+After listwise deletion of missings (\(3\) in total, only for the self-efficacy)
+we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
+descriptives for the dataset. For the circular variable IPC we show sample
+estimates for the circular mean \(\bar{\theta}\) and concentration \(\hat{\rho}\).
+The circular concentration lies between 0, meaning the data is not concentrated
+at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
+concentrated at a single point on the circle. The population values of these
+parameters are usually, and also in this paper, referred to as \(\mu\) (circular
+location) and \(\kappa\) (circular concentration). For the linear variables
+(strength IPC and SE) we show sample estimates of the linear mean and standard
+deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
+the linear and circular outcome of the teacher data for teachers with low SE
+(below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
+mean) and high SE (above 1 sd above the mean).
 
@@
 
-One of the goals of this paper is to show the benefits of cylindrical
-methods for the analysis of circumplex data and cylindrical data in
-psychology in general. To do so we decide to focus on four cylindrical
-models. The models were selected for their relatively low complexity and
-the ease with which a regression structure could be incorporated. But
-also because they show different ways of modelling the the linear and
-circular outcome and thereby illustrate a wider range of cylindrical
-models available in the literature. The cylindrical models contain a set
-of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\)
-predictors \(\boldsymbol{z} = z_1, \dots, z_q\) for the linear and
-circular outcomes, \(Y\) and \(\Theta\), respectively. The first two
-models are based on a construction by Mastrantonio et al. (2015), while
-the other models are extensions of the models from Abe and Ley (2017)
-and Mastrantonio (2018). The four cylindrical models are introduced
-separately in the subsections below. However, to provide a more succinct
-overview and comparison of the four models, Table \ref{TableModels}
-gives an overview of the similarities and differences between the
-models.
+One of the goals of this paper is to show the benefits of cylindrical methods
+for the analysis of circumplex data and cylindrical data in psychology in
+general. To do so we decide to focus on four cylindrical models. The models were
+selected for their relatively low complexity and the ease with which a
+regression structure could be incorporated. But also because they show different
+ways of modelling the linear and circular outcome and thereby illustrate a
+wider range of cylindrical models available in the literature. The
+cylindrical models contain a set of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\) predictors \(\boldsymbol{z} = z_1, \dots, z_p\) for the linear and
+circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
+based on a construction by Mastrantonio et al. (2015), while the other models are
+extensions of the models from Abe and Ley (2017) and Mastrantonio (2018). The
+four cylindrical models are introduced separately in the subsections below.
+However, to provide a more succinct overview and comparison of the four models, Table
+\ref{TableModels} gives an overview of the similarities and differences between
+the models.
 
@@
 
-Following Mastrantonio et al. (2015) we consider two models where the
-relation between \(\Theta \in [0, 2\pi)\) and
-\(Y\in (-\infty, + \infty)\) and \(q\) covariates is specified as
+Following Mastrantonio et al. (2015) we consider two models where
+the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
+and \(q\) covariates is specified as
 \begin{equation}\label{circlinlink}
 Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
-\end{equation} \noindent where the random variable \(R\geq0\) will be
-introduced below, the error term \(\epsilon \sim N(0, \sigma^2)\) with
-variance \(\sigma^2>0\),
-\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
-the intercept and regression coefficients and \(x_1, \dots, x_q\) are
-the \(q\) covariates for the prediction of the linear outcome. We thus
-assume a normal distribution for the linear outcome.\newline \indent For
-the circular outcome we assume either a projected normal (PN) or a
-general projected normal (GPN) distribution. These distributions arise
-from a projection of a distribution defined in bivariate space onto the
-circle. Figure \ref{projection} represents this projection. In the left
-plot of Figure \ref{projection} we see datapoints from the bivariate
-variable \(\boldsymbol{S}\) that in the middle plot are projected to
-form the circular outcome \(\Theta\) in the right plot. Mathematically
-the relation between \(\boldsymbol{S}\) and \(\Theta\) is defined as
-follows \begin{equation}\label{projection}
+\end{equation}
+\noindent where the random variable \(R\geq0\) will be introduced below, the
+error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
+\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
+intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
+covariates for the prediction of the linear outcome. We thus assume a normal
+distribution for the linear outcome.\newline
+\indent For the circular outcome we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise from a
+projection of a distribution defined in bivariate space onto the circle. Figure
+\ref{projection} represents this projection. In the left plot of Figure
+\ref{projection} we see datapoints from the bivariate variable
+\(\boldsymbol{S}\) that in the middle plot are projected to form the circular
+outcome \(\Theta\) in the right plot. Mathematically the relation between
+\(\boldsymbol{S}\) and \(\Theta\) is defined as follows
+\begin{equation}\label{projection}
 \boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
-\end{equation} \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\),
-the Euclidean norm of \(\boldsymbol{S}\); the lines connecting the
-bivariate datapoints to the origin in the middle plot. We call
-\(\boldsymbol{S}\) the augmented representation of the circular outcome,
-it is a variable that in contrast to \(\Theta\) is not observed and thus
-considered latent or auxiliary. This then means that we do not model
-\(\Theta\) directly but indirectly through \(\boldsymbol{S}\). \newline
-\indent For both the PN and GPN distribution the circular location
-parameter \(\mu\in [0, 2\pi)\) is modeled as
-\(\hat{\mu}_i = \mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) = \mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i, \boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
+\end{equation}
+\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
+\(\boldsymbol{S}\); the lines connecting the bivariate datapoints to the origin
+in the middle plot. We call \(\boldsymbol{S}\) the augmented representation of the
+circular outcome. It is a variable that in contrast to \(\Theta\) is not
+observed and thus considered latent or auxiliary. This then means that we do not
+model \(\Theta\) directly but indirectly through \(\boldsymbol{S}\). \newline
+\indent For both the PN and GPN distributions the circular location parameter
+\(\mu\in [0, 2\pi)\) is modeled as \(\hat{\mu}_i = \mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) = \mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i, \boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
 model the circular location parameter also depends on the variance-covariance
 matrix and the circular predicted values should be computed using numerical
-integretion or Monte Carlo methods because a closed form expression for the mean
+integration or Monte Carlo methods because a closed form expression for the mean
 direction is not available.} where \(\boldsymbol{\beta}^{I}\) and
 \(\boldsymbol{\beta}^{II}\) are vectors with intercepts and regression
-coefficients for the prediction of \(\boldsymbol{S}\) and
-\(\boldsymbol{z}_i\) is a vector with predictor values for each
-individual \(i \in 1, \dots, n\) where \(n\) is the sample size. Note
-that as a result of the augmented representation of the circular outcome
-we have two sets of regression coefficients and intercepts, one for each
-bivariate component of \(\boldsymbol{S}\). This leads to problems when
-we want to interpret the effect of a covariate on the circle. A circular
-regression line is shown in Figure \ref{circregline}, with covariate
-values on the x-axis and the predicted circular outcome on the y-axis.
-As can be seen it is of a non-linear character meaning that the effect
-of a covariate is different at different values of the covariate. A
-circular regression line is usually described by the slope at the
-inflection point, the point at which the slope of the regression line
-starts flattening off (indicated with a square in Figure
-\ref{circregline}). By default, the parameters from the PN and GPN model
-do not directly describe this inflection point. For the PN distribution
-however, Cremers et al. (2018b) solved this interpretation problem and
-introduce new circular regression coefficients. They introduce a new
-parameter \(b_c\) that describes the slope at the inflection point of
-the regression line. For the GPN distrbution the interpretation problem
-however remains.\newline \indent The main difference between the PN and
-GPN distribution lies in the definition of their covariance matrix. For
-the PN distribution this is an identity matrix, causing the distribution
-to be unimodal and symmetric, whereas for the GPN distribution
-\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\),
-allowing for multimodality and assymetry/skewness.\newline \indent For
-the teacher dataset the predicted linear outcome, strengths of
+coefficients for the prediction of \(\boldsymbol{S}\) and \(\boldsymbol{z}_i\) is a
+vector with predictor values for each individual \(i \in 1, \dots, n\) where \(n\)
+is the sample size. Note that as a result of the augmented representation of the
+circular outcome we have two sets of regression coefficients and intercepts, one
+for each bivariate component of \(\boldsymbol{S}\). This leads to problems when we
+want to interpret the effect of a covariate on the circle. A circular regression
+line is shown in Figure \ref{circregline}, with covariate values on the x-axis
+and the predicted circular outcome on the y-axis. As can be seen it is of a
+non-linear character meaning that the effect of a covariate is different at
+different values of the covariate. A circular regression line is usually
+described by the slope at the inflection point, the point at which the slope of
+the regression line starts flattening off (indicated with a square in Figure
+\ref{circregline}). By default, the parameters from the PN and GPN models do not
+directly describe this inflection point. For the PN distribution however,
+Cremers et al. (2018b) solved this interpretation problem and introduce new
+circular regression coefficients. They introduce a new parameter \(b_c\) that
+describes the slope at the inflection point of the regression line. For the
+GPN distrbution the interpretation problem however remains.\newline
+\indent The main difference between the PN and GPN distribution lies in the
+definition of their covariance matrix. For the PN distribution this is an
+identity matrix, causing the distribution to be unimodal and symmetric, whereas
+for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
+asymetry/skewness.\newline
+\indent For the teacher dataset the predicted linear outcome, strengths of
 interpersonal behavior, in the CL-PN and CL-GPN model is the following:
 \[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
-\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted
-circular outcome, type of interpersonal behavior, equals:
-\[\hat{\theta}_i = \mu_i =
+\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted circular
+outcome, type of interpersonal behavior, equals: \[\hat{\theta}_i = \mu_i =
 \mbox{atan2} (\beta_0^{II} + \beta_1^{II}\text{SE}_i, \beta_0^{I} +
-\beta_1^{I}\text{SE}_i).\] \noindent where \(\text{SE}_i\) is the
-self-efficacy score of one individual. The CL-PN and CL-GPN models thus
-allow us to assess the average type and strength of interpersonal
-behavior through the parameters \(\beta_{0}^{I}\), \(\beta_{0}^{II}\)
-and \(\gamma_0\) as well as the effect of self-efficacy on type and
-strength of teacher behavior through the parameters, \(\beta_{1}^{I}\),
-\(\beta_{1}^{II}\) and \(\gamma_1\). In addition, because the type and
-strength of interpersonal behavior are modelled together via the
-regression in \eqref{circlinlink} we can assess the effect of the type
-of interpersonal behavior on the strength through the parameters
-\(\gamma_{sin}\) and \(\gamma_{cos}\). In the teacher data these are the
-regression coefficients for the effect of the sine and cosine of the
-type of behavior which are related to the scores on the Agency an
-Communion dimensions respectively.\newline \indent Both the CL-PN and
-CL-GPN models are estimated using Markov Chain Monte Carlo (MCMC)
-methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011),
-Wang and Gelfand (2013) and Hernandez-Stumpfhauser, Breidt, and Woerd
-(2016) for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in
-the Supplementary Material.
+\beta_1^{I}\text{SE}_i).\] \noindent where \(\text{SE}_i\) is the self-efficacy
+score of one individual. The CL-PN and CL-GPN models thus allow us to assess the
+average type and strength of interpersonal behavior through the parameters
+\(\beta_{0}^{I}\), \(\beta_{0}^{II}\) and \(\gamma_0\) as well as the effect of
+self-efficacy on type and strength of teacher behavior through the parameters,
+\(\beta_{1}^{I}\), \(\beta_{1}^{II}\) and \(\gamma_1\). In addition, because the type
+and strength of interpersonal behavior are modelled together via the regression
+in \eqref{circlinlink} we can assess the effect of the type of interpersonal
+behavior on the strength through the parameters \(\gamma_{sin}\) and
+\(\gamma_{cos}\). In the teacher data these are the regression coefficients for
+the effect of the sine and cosine of the type of behavior which are related to
+the scores on the Agency and Communion dimensions respectively.\newline
+\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
+Carlo (MCMC) methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011), Wang and Gelfand (2013) and
+Hernandez-Stumpfhauser, Breidt, and Woerd (2016) for the regression of the circular outcome. A detailed
+description of the Bayesian estimation and MCMC samplers can be found in the
+Supplementary Material.
 
@@
 
-This model is an extension of the cylindrical model introduced in Abe
-and Ley (2017) to the regression context. It concerns a combination of a
-Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
-\(\alpha\), for the linear outcome and a sine-skewed von Mises
-distribution, with location parameter \(\mu\in [0, 2\pi)\),
-concentration parameter \(\kappa>0\) and skewness
-\(\lambda \in [-1,1]\), for the circular outcome. In contrast to the
-CL-PN and CL-GPN models, the linear outcome \(Y\) is in this model
-defined only on the positive real half-line \([0, + \infty)\) and thus
-can not be negative.\newline \indent In this model we predict the linear
-scale parameter and circular location parameter, both of which we can
-express in terms of covariates:
-\(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
-\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
-The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
-coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
-\(y\) where \(j = 0, \dots, q\) and \(\gamma_0\) is the intercept. The
-parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
-\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
-\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
-where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
-of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
-is a vector of predictor values for the prediction of
-\(\theta\).\newline \indent For the teacher data, the predicted values
-for the circular outcome in the Abe-Ley model are:
-\[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
-\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear
-outcome. The conditional distribution for the linear outcome is Weibull,
-meaning that we can use methods from survival analsis to interpret the
-effect of a predictor. In survival analysis a \enquote{survival}
-function is used in which time is plotted against the probability of
-survival of subjects suffering from a specific medical condition. In the
-teacher data we can thus compute the probability of a teacher having a
-specific strength on the IPC. This probability is computed using the
+This model is an extension of the cylindrical model introduced in
+Abe and Ley (2017) to the regression context. It concerns a combination
+of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
+\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
+with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
+\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
+contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
+model defined only on the positive real half-line \([0, + \infty)\) and thus can
+not be negative.\newline
+\indent In this model we predict the linear scale parameter and circular
+location parameter, both of which we can express in terms of
+covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
+\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
+parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
+\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0, \dots, q\) and \(\gamma_0\) is the intercept.
+The parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
+regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
+\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
+vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
+is a vector of predictor values for the prediction of \(\theta\).\newline
+\indent For the teacher data, the predicted values for the circular outcome in
+the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
+\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
+The conditional distribution for the linear outcome is Weibull, meaning that we
+can use methods from survival analsis to interpret the effect of a predictor. In
+survival analysis a \enquote{survival} function is used in which time is plotted against
+the probability of survival of subjects suffering from a specific medical
+condition. In the teacher data we can thus compute the probability of a teacher
+having a specific strength on the IPC. This probability is computed using the
 \enquote{survival-function} defined as \[\exp(-\alpha
 y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
-with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the
-survival function we also see that the circular concentration parameter
-\(\kappa\) and the linear shape parameter \(\alpha\) regulate the
-circular-linear dependence in the Abe-Ley model. The Abe-Ley model thus
-allows us to assess the average type and strength of interpersonal
-behavior through the parameters \(\beta_{0}\) and \(\gamma_0\) as well
-as the effect of self-efficacy on type and strength of teacher behavior
-through the parameters, \(\beta_{1}\) and \(\gamma_1\).\newline
-\indent We can use numerical optimization (Nelder-Mead) to find
-solutions for the maximum likelihood (ML) estimates for the parameters
-of the model.
+with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
+function we also see that the circular concentration parameter \(\kappa\) and
+the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
+the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
+and strength of interpersonal behavior through the parameters \(\beta_{0}\) and
+\(\gamma_0\) as well as the effect of self-efficacy on type and strength of
+teacher behavior through the parameters \(\beta_{1}\) and \(\gamma_1\).\newline
+\indent We can use numerical optimization (Nelder-Mead) to find solutions for
+the maximum likelihood (ML) estimates for the parameters of the model.
 
@@
 independent circular outcomes and \(w\) independent linear outcomes. The
-circular outcomes
-\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\)
-are modelled together by a multivariate GPN distribution. The linear
-outcomes
-\(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,  \boldsymbol{Y}_w)\) are
-modelled together by a multivariate skew normal distribution (Sahu, Dey,
-\& Branco, 2003). Thus also in this model the linear outcome is defined
-constrained to be positive. Because the GPN distribution is modelled
-using a so-called augmented representation (see also the description of
-the CL-PN and CL-GPN models) it is convenient to use a similar tactic
-for modelling the multivariate skew normal distribution. As in
-Mastrantonio (2018) dependence between the linear and circular outcome
-is created by modelling the augmented representations of
-\(\boldsymbol{\Theta}\) and \(\boldsymbol{Y}\) together in a \(2m + w\)
-dimensional normal distribution.\newline \indent This means that we have
-a shared mean vector and variance-covariance matrix for the linear and
-circular outcome(s), much like having multiple outcomes in a MANOVA
-(multivariate analysis of variance) model. In our regression extension
-of the GPN-SSN model we have \(i = 1, \dots, n\) observations of \(m\)
-circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean
-vector then becomes
-\(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
-\(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with
-regression coefficients and intercepts and \(\boldsymbol{x}_i\) is a
-\(g + 1\) dimensional vector containing the value 1 to estimate an
-intercept and the \(g\) covariate values. This means that in contrast to
-the other three models, we have to use the same set of predictors for
-the circular and linear outcome.\newline \indent For the teacher data,
-\(\boldsymbol{B} = \begin{bmatrix} \beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
-The predicted
-circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.}
-and linear outcomes in the GPN-SSN model are
+circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
+distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
+distribution (Sahu, Dey, \& Branco, 2003). Because the GPN distribution is modelled
+using a so-called augmented representation (see also the description of the
+CL-PN and CL-GPN models) it is convenient to use a similar tactic for
+modelling the multivariate skew normal distribution. As in
+Mastrantonio (2018), dependence between the linear and circular outcome is
+created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
+and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
+distribution.\newline
+\indent This means that we have a shared mean vector and variance-covariance
+matrix for the linear and circular outcome(s), much like having multiple
+outcomes in a MANOVA (multivariate analysis of variance) model. In our
+regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
+observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
+covariates.
+The mean vector then becomes \(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with regression coefficients and intercepts and
+\(\boldsymbol{x}_i\) is a \(g + 1\) dimensional vector containing the value 1 to
+estimate an intercept and the \(g\) covariate values. This means that in contrast
+to the other three models, we have to use the same set of predictors for the
+circular and linear outcome.\newline
+\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix} \beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
+The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
+linear outcomes in the GPN-SSN model are
 \[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
@@
 
-\noindent The GPN-SSN model thus allows us to assess the average type
-and strength of interpersonal behavior through the parameters
-\(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\) and \(\beta_{0_{y}}\) as
-well as the effect of self-efficacy on type and strength of teacher
-behavior through the parameters, \(\beta_{1_{s^{I}}}\),
-\(\beta_{1_{s^{II}}}\) and \(\beta_{1_{y}}\). In addition, because the
-type and strength of interpersonal behavior are modelled together using
-a multivariate normal distribution we can through its
-variance-covariance matrix also assess the dependence between the type
-and strength of interpersonal behavior. \newline \indent We estimate the
-model using MCMC methods. A detailed description of these methods is
-given in the Supplementary Material.
+\noindent The GPN-SSN model thus allows us to assess the average type and
+strength of interpersonal behavior through the parameters \(\beta_{0_{s^{I}}}\),
+\(\beta_{0_{s^{II}}}\) and \(\beta_{0_{y}}\) as well as the effect of self-efficacy
+on type and strength of teacher behavior through the parameters
+\(\beta_{1_{s^{I}}}\), \(\beta_{1_{s^{II}}}\) and \(\beta_{1_{y}}\). In addition,
+because the type and strength of interpersonal behavior are modelled together
+using a multivariate normal distribution we can, through its variance-covariance
+matrix, also assess the dependence between the type and strength of interpersonal
+behavior. \newline
+\indent We estimate the model using MCMC methods. A detailed description of these
+methods is given in the Supplementary Material.
 
@@
 
-For the four cylindrical models we focus on their out-of-sample
-predictive performance to determine the fit of the model. To do so we
-use k-fold cross-validation and split our data into 10 folds. Each of
-these folds (10 \(\%\) of the sample) is used once as a holdout set and
-9 times as part of a training set. The analysis will thus be performed
-10 times, each time on a different training set.\newline \indent A
-proper criterion to compare out-of-sample predictive performance is the
-Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007). The
-lower the value of this criterion, the better the predictive performance
-of the model. Using ML estimates this criterion can be computed as
-follows: \begin{equation}\label{PLSLML}
+For the four cylindrical models we focus on their out-of-sample predictive
+performance to determine the fit of the model. To do so we use k-fold
+cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
+of the sample) is used once as a holdout set and 9 times as part of a training
+set. The analysis will thus be performed 10 times, each time on a different
+training set.\newline
+\indent A proper criterion to compare out-of-sample predictive performance is
+the Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007). The lower the
+value of this criterion, the better the predictive performance of the model.
+Using ML estimates this criterion can be computed as follows:
+\begin{equation}\label{PLSLML}
 PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
-\end{equation} \noindent where \(l\) is the model likelihood, \(M\) is
-the sample size of the holdout set, \(x_i\) is the \(i^{th}\) datapoint
-from the holdout set and \(\hat{\boldsymbol{\vartheta}}\) are the ML
-estimates of the model parameters. Using posterior samples the criterion
-is similar to the log pointwise predictive density (lppd) (Gelman et
-al., 2014, p. 169) and can be computed as:
+\end{equation}
+\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
+holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
+\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
+Using posterior samples the criterion is similar to the log pointwise predictive
+density (lppd) (Gelman et al., 2014, p. 169) and can be computed as:
 \begin{equation}\label{PLSLBayes}
 PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
-\end{equation} \noindent where \(B\) is the amount of posterior samples
-and \(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
-model parameters for the \(j^{th}\) iteration. Because the joint density
-and thus also the likelihood for the modified GPN-SSN model is not
-available in closed form (Mastrantonio, 2018) we compute the PLSL for
-the circular and linear outcome separately for all models. Note that
-although we fit the CL-PN, CL-GPN and GPN-SSN models using Bayesian
-statistics, we do not take prior information into account when assessing
-model fit with the PLSL. According to Gelman et al. (2014) this is not
-necessary since we are assessing the fit of a model to data, the holdout
-set, only. They argue that the prior in such case is only of interest
-for estimating the parameters of the model but not for determining the
-predictive accuracy.\newline \indent For each of the four cylindrical
-models and for each of the 10 cross-validation analyses we can then
-compute a PLSL for the circular and linear outcome by using the
-conditional log-likelihoods of the respective outcome (see Supplementary
-Material for a definition of the loglikelihoods). To evaluate the
+\end{equation}
+\noindent where \(B\) is the amount of posterior samples and
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
+parameters for the \(j^{th}\) iteration. Because the joint density and thus also
+the likelihood for the modified GPN-SSN model is not
+available in closed form (Mastrantonio, 2018) we compute the PLSL for the
+circular and linear outcome separately for all models. Note that although we fit
+the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
+prior information into account when assessing model fit with the PLSL. According
+to Gelman et al. (2014) this is not necessary since we are assessing the fit of a model
+to data, the holdout set, only. They argue that the prior in such case is only
+of interest for estimating the parameters of the model but not for determining
+the predictive accuracy.\newline
+\indent For each of the four cylindrical models and for each of the 10
+cross-validation analyses we can then compute a PLSL for the circular and linear
+outcome by using the conditional log-likelihoods of the respective outcome (see
+Supplementary Material for a definition of the loglikelihoods). To evaluate the
 predictive performance we average across the PLSL criteria of the
-cross-validation analyses. We also assess the cross-validation
-variability by means of the standard deviations of the PLSL criteria.
+cross-validation analyses. We also assess the cross-validation variability by
+means of the standard deviations of the PLSL criteria.
 
@@
 In this section we analyze the teacher data with the help of the four
-cylindrical models from the previous section. We will present the
-results, posterior estimates and their interpretation for all four
-models.
+cylindrical models from the previous section. We will present the results,
+posterior estimates and their interpretation for all four models.
 
@@
 
-In the Supplementary Material we have described the starting values for
-the MCMC procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it
-remains to specify the starting values for the maximum likelihood based
-Abe-Ley model:
-\(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
+In the Supplementary Material we have described the starting values for the MCMC
+procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it remains to specify
+the starting values for the maximum likelihood based Abe-Ley model: \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
 The initial number of iterations for the three MCMC samplers was set to
-2000. After convergence checks via traceplots we concluded that some of
-the parameters of the GPN-SSN model did not converge. Therefore we set
-the number of iterations of the MCMC models to 20,000 and subtracted a
-burn-in of 5000 to reach convergence (the Geweke diagnostics show
-absolute z-scores over 1.96 in 6\% of the estimated parameters). Note
-that we choose the same number of iterations for all three models models
-estimated using MCMC prodedures to make their comparison via the PLSL as
-fair as possible. Lastly, the predictor \verb|SE| was centered before
-inclusion in the analysis as this allows the intercepts to bear the
-classical meaning of average behavior.\newline \indent Tables
-\ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show the
-results for the four cylindrical models that were fit to the teacher
-data. For the models estimated using MCMC methods, CL-PN, CL-GPN and
-GPN-SSN we show descriptives of the posterior of the estimated
-parameters (posterior mode and lower and upper bound of the 95\% highest
-posterior density (HPD) interval). For the Abe-Ley model we show the
-maximum likelihood estimates of the parameters. To compare the results
-of the four models we focus on the following aspects: the estimated
-average scores (intercept) on the type and strength of interpersonal
+2000. After convergence checks via traceplots we concluded that some of the
+parameters of the GPN-SSN model did not converge. Therefore we set the number of
+iterations of the MCMC models to 20,000 and subtracted a burn-in of 5000 to
+reach convergence (the Geweke diagnostics show absolute z-scores over 1.96 in
+6\% of the estimated parameters). Note that we choose the same number of
+iterations for all three models estimated using MCMC procedures to make
+their comparison via the PLSL as fair as possible. Lastly, the predictor
+\verb|SE| was centered before inclusion in the analysis as this allows the
+intercepts to bear the classical meaning of average behavior.\newline
+\indent Tables \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show
+the results for the four cylindrical models that were fit to the teacher data.
+For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN, we show
+descriptives of the posterior of the estimated parameters (posterior mode and
+lower and upper bound of the 95\% highest posterior density (HPD) interval). For
+the Abe-Ley model we show the maximum likelihood estimates of the parameters. To
+compare the results of the four models we focus on the following aspects: the
+estimated average scores (intercept) on the type and strength of interpersonal
 behavior (1), the effect of self-efficacy on the type and strength of
-interpersonal behavior (2), the dependence between the type and strength
-of interpersonal behavior (3) and the model fit (4).
+interpersonal behavior (2), the dependence between the type and strength of
+interpersonal behavior (3) and the model fit (4).
 
@@
 
-The parameters \(\gamma_0\) in the CL-PN, CL-GPN and Abe-Ley model and
-the parameter \(\beta_{0_y}\) in the GPN-SSN model inform us about the
-strength of interpersonal behavior at the average self-efficacy. For the
-CL-PN, CL-GPN and GPN-SSN model the parameters are estimated at 0.38,
-0.37 and 0.30 respectively and are a direct prediction of the strength
-of interpersonal behavior at the average self-efficacy. The estimate for
-the GPN-SSN model is notably lower and likely to be caused by its skewed
-distribution for the strengths of interpersonal behavior. In the Abe-Ley
-model, \(\gamma_0\) influences the shape parameter of the distribution
-of the strength of interpersonal behavior and does not directly estimate
-the average strength. Instead we can use the survival function to say
-something about the probability of having a certain strength of
-interpersonal behavior. Figure \ref{reglineweib} shows this function for
-several values of self-efficacy. We look at the survival function at
-average values of self-efficacy. Note that this function is the average
-of all survival functions for observations that fall within 1 standard
-deviation of the mean. The survival function indicates that the
-probability of having a low strength of interpersonal behavior is higher
-than having a high strength. We however can not make any direct
-statement about the estimated strength using the Abe-Ley model.\newline
+The parameters \(\gamma_0\) in the CL-PN, CL-GPN and Abe-Ley model and the
+parameter \(\beta_{0_y}\) in the GPN-SSN model inform us about the strength of
+interpersonal behavior at the average self-efficacy. For the CL-PN, CL-GPN and
+GPN-SSN models the parameters are estimated at 0.38, 0.37 and 0.33 respectively
+and are a direct prediction of the strength of interpersonal behavior at the
+average self-efficacy. The estimate for the GPN-SSN model is notably lower and
+likely to be caused by its skewed distribution for the strengths of
+interpersonal behavior. In the Abe-Ley model, \(\gamma_0\) influences the shape
+parameter of the distribution of the strength of interpersonal behavior and does
+not directly estimate the average strength. Instead we can use the survival
+function to say something about the probability of having a certain strength of
+interpersonal behavior. Figure \ref{reglineweib} shows this function for several
+values of self-efficacy. We look at the survival function at average values of
+self-efficacy. Note that this function is the average of all survival functions
+for observations that fall within 1 standard deviation of the mean. The survival
+function indicates that the probability of having a low strength of
+interpersonal behavior is higher than having a high strength. We however can not
+make any direct statement about the estimated strength using the Abe-Ley
+model.\newline
 \indent The parameters \(\beta_0^{I}\), \(\beta_0^{II}\), \(\beta_0\),
-\(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) inform us about the
-type of interpersonal behavior at the average self-efficacy for the
-CL-PN, CL-GPN, Abe-Ley and GPN-SSN model respectively. For the CL-PN,
-CL-GPN and GPN-SSN model we need to combine the estimates for the
-underlying bivariate components \(\{I, II\}\) into one circular estimate
-using the double arctangent
+\(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) inform us about the type of
+interpersonal behavior at the average self-efficacy for the CL-PN, CL-GPN,
+Abe-Ley and GPN-SSN model respectively. For the CL-PN, CL-GPN and GPN-SSN model
+we need to combine the estimates for the underlying bivariate components \(\{I, II\}\) into one circular estimate using the double arctangent
 function\footnote{\(\mbox{atan2}(\beta_0^{II}, \beta_0^{I})\) or
-\(\mbox{atan2}(\beta_{0_{s^{II}}}, \beta_{0_{s^{I}}})\)}. Table
-\ref{tab:means} shows that these circular estimates are similar for the
-three models at 32.29\(^\circ\), 33.70\(^\circ\) and 35.53\(^\circ\). In
-the Abe-Ley model the type of interpersonal behavior at the average
-self-efficacy is estimated at 0.36 radians or 20.63\(^\circ\).
+\(\mbox{atan2}(\beta_{0_{s^{II}}}, \beta_{0_{s^{I}}})\)}.
+Table \ref{tab:means} shows that these circular estimates are similar for the
+three models at 32.29\(^\circ\), 33.70\(^\circ\) and 35.53\(^\circ\). In the Abe-Ley
+model the type of interpersonal behavior at the average self-efficacy is
+estimated at 0.36 radians or 20.63\(^\circ\).
 
@@
 
-The parameters \(\gamma_1\) in the CL-PN, CL-GPN, Abe-Ley model and
-\(\beta_{1_y}\) in the GPN-SSN model inform us about the effect of
-self-efficacy on the strength of interpersonal behavior. For the CL-PN,
-CL-GPN and GPN-SSN model the parameters are estimated at 0.03, 0.03 and
-0.09 respectively and are a direct estimate of the effect of
-self-efficacy on the strength of interpersonal behavior, \emph{i.e.} an
-increase of 1 unit in self-efficacy leads to an increase of 0.09 units
-in the strength of interpersonal behavior according to the GPN-SSN
-model. These estimates are however quite small and only different from 0
-(the HPD interval does not contain 0) in the GPN-SSN model. It is hard
-to say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to
-base our conclusions on. The model CL-GPN and CL-PN fit the linear
-outcome best according to the model fit in Table \ref{tab:ModelFit}. In
-these models the linear outcome has a symmetric distribution whereas in
-the GPN-SSN the distribution of the linear outcome is skewed. However,
-the effect of self-efficacy is different from 0 only in the GPN-SSN
-model which does not seem to match with its lower model fit.\newline
-\indent In the Abe-Ley model, \(\gamma_1\) influences the shape
-parameter of the distribution of the strength of interpersonal behavior
-and does not directly estimate the effect of self-efficacy. Instead we
-can use the survival function to say something about the probability of
-having a certain strength of interpersonal behavior for different values
-of self-efficacy. Figure \ref{reglineweib} shows this function for low,
-average and high values of self-efficacy (as defined in Figure
-\ref{dataplot}). This function indicates that the effect of
-self-efficacy on the strength of interpersonal behavior is not linear.
-The probability of having a higher strength of interpersonal behavior is
-highest for low self-efficacy and lowest for average
-self-efficacy.\newline \indent The parameters \(\beta_1^{I}\),
-\(\beta_1^{II}\), \(\beta_1\), \(\beta_{1_{s^{I}}}\) and
-\(\beta_{1_{s^{II}}}\) inform us about the effect of self-efficacy on
-the type of interpersonal behavior in the CL-PN, CL-GPN, Abe-Ley and
-GPN-SSN model respectively. For the CL-PN and Abe-Ley model we have
-drawn the circular regression lines for this effect in Figure
-\ref{regline} (see the description of the CL-PN and CL-GPN models). For
-the CL-PN model the inflection point is indicated with a square in
-Figure \ref{regline}. The inflection point for the Abe-Ley model falls
-outside the bounds of the plot and is therefore not displayed. The slope
-at the inflection point, \(b_c\), for the CL-PN model is computed by
-using methods from Cremers et al. (2018b) and is equal to 1.67 (-24.66,
-29.33)\footnote{Note that this is a linear approximation
-to the circular regression line representing the slope at a specific point.
-Therefore it is possible for the HPD interval to be wider than $2\pi$. In this
-case the interval is much wider and covers 0, indicating there is no evidence
-for an effect.} The parameter \(\beta_1\) is the slope at the inflection
-point for the Abe-Ley model and is equal to -0.03. Even though these
-slopes are quite different, the regression lines in Figure \ref{regline}
-are quite similar in the data range. Both the regression line of the
-Abe-Ley model and the CL-PN model show slopes that are not very steep in
-the range of the data indicating that the effect of self-efficacy on the
-type of interpersonal behavior is not large. \newline \indent In the
-CL-GPN and GPN-SSN model we cannot compute circular regression
+The parameters \(\gamma_1\) in the CL-PN, CL-GPN, Abe-Ley models and \(\beta_{1_y}\)
+in the GPN-SSN model inform us about the effect of self-efficacy on the strength
+of interpersonal behavior. For the CL-PN, CL-GPN and GPN-SSN model the
+parameters are estimated at 0.03, 0.03 and 0.09 respectively and are a direct
+estimate of the effect of self-efficacy on the strength of interpersonal
+behavior, \emph{i.e.} an increase of 1 unit in self-efficacy leads to an
+increase of 0.09 units in the strength of interpersonal behavior according to
+the GPN-SSN model. These estimates are however quite small and only different
+from 0 (the HPD interval does not contain 0) in the GPN-SSN model. It is hard to
+say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to base our
+conclusions on. The models CL-GPN and CL-PN fit the linear outcome best according
+to the model fit in Table \ref{tab:ModelFit}. In these models the linear outcome has a
+symmetric distribution whereas in the GPN-SSN the distribution of the linear
+outcome is skewed. However, the effect of self-efficacy is different from 0 only
+in the GPN-SSN model which does not seem to match with its lower model fit.\newline
+\indent In the Abe-Ley model, \(\gamma_1\) influences the shape parameter of the distribution of
+the strength of interpersonal behavior and does not directly estimate the effect
+of self-efficacy. Instead we can use the survival function to say something
+about the probability of having a certain strength of interpersonal behavior for
+different values of self-efficacy. Figure \ref{reglineweib} shows this function
+for low, average and high values of self-efficacy (as defined in Figure
+\ref{dataplot}). This function indicates that the effect of self-efficacy on the
+strength of interpersonal behavior is not linear. The probability of having a
+higher strength of interpersonal behavior is highest for low self-efficacy and
+lowest for average self-efficacy.\newline
+\indent The parameters \(\beta_1^{I}\), \(\beta_1^{II}\), \(\beta_1\),
+\(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) inform us about the effect of
+self-efficacy on the type of interpersonal behavior in the CL-PN, CL-GPN,
+Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley models we have
+drawn the circular regression lines for this effect in Figure \ref{regline} (see
+the description of the CL-PN and CL-GPN models for a detailed explanation of
+circular regression lines). For the CL-PN model the inflection point is
+indicated with a square in Figure \ref{regline}. The inflection point for the
+Abe-Ley model falls outside the bounds of the plot and is therefore not
+displayed. The slope at the inflection point, \(b_c\), for the CL-PN model is
+computed by using methods from Cremers et al. (2018b) and is equal to 1.67
+(-24.66, 29.33)\footnote{Note that this is a linear approximation to the
+circular regression line representing the slope at a specific point. Therefore
+it is possible for the HPD interval to be wider than $2\pi$. In this case the
+interval is much wider and covers 0, indicating there is no evidence for an
+effect.}. The parameter \(\beta_1\) is the slope at the inflection point for the
+Abe-Ley model and is equal to -0.03. Even though these slopes are
+different, the regression lines in Figure \ref{regline} are quite similar in the
+data range. Both the regression line of the Abe-Ley model and the CL-PN model
+show slopes that are not very steep in the range of the data indicating that the
+effect of self-efficacy on the type of interpersonal behavior is not large.
+\newline
+\indent In the CL-GPN and GPN-SSN models we cannot compute circular regression
 coefficients due to the fact that not only the mean vector of the GPN
-distribution but also the covariance matrix influences the predicted
-value on the circle. Instead, we will compute posterior predictive
-distributions for the predicted circular outcome of individuals scoring
-the minimum, maximum and median self-efficacy. The modes and 95\% HPD
-intervals of these posterior predictive distributions are
-\(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\),
-\(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\)
-for the CL-GPN model. Note that we display the modes and HPD intervals
-for the posterior predictive distributions on the interval
-\([0^\circ, 360^\circ)\) and that \(44.49^\circ = 404.49^\circ\) due to
-the periodicity of a circular variable. The posterior mode estimate of
-\(215.74^\circ\) thus lies within its HPD interval
-\((147.36^\circ, \: 44.49^\circ)\). For the GPN-SSN model the modes and
-95\% HPD intervals of the posterior predictive distributiona are
-\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.12^\circ, \: 72.02^\circ)\),
-\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
-\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\).
-For both the CL-GPN and GPN-SSN model the HPD intervals of the mode of
-the posterior predictive intervals of individuals scoring the minimum,
-median and maximum self-efficacy overlap. This indicates that the effect
-of self-efficacy, if there is any, on the type of interpersonal behavior
-a teacher shows is not expected to be strong. Had the HPD intervals not
-overlapped we could have concluded that as the self-efficacy increases,
-the score of the teacher on the IPC moves counterclockwise.\newline
+distribution but also the covariance matrix influences the predicted value on
+the circle. Instead, we will compute posterior predictive distributions for the
+predicted circular outcome of individuals scoring the minimum, maximum and
+median self-efficacy. The modes and 95\% HPD intervals of these posterior
+predictive distributions are \(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\), \(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\), \(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\) for the CL-GPN model.
+Note that we display the
+modes and HPD intervals for the posterior predictive distributions on the
+interval \([0^\circ, 360^\circ)\) and that \(44.49^\circ = 404.49^\circ\) due to the
+periodicity of a circular variable. The posterior mode estimate of
+\(215.74^\circ\) thus lies within its HPD interval \((147.36^\circ, \: 44.49^\circ)\). For the GPN-SSN model the modes and 95\% HPD intervals of the
+posterior predictive distributions are \(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.12^\circ, \: 72.02^\circ)\), \(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\), \(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\).
+For both the CL-GPN and GPN-SSN model the HPD
+intervals of the mode of the posterior predictive intervals of individuals
+scoring the minimum, median and maximum self-efficacy overlap. This indicates
+that the effect of self-efficacy, if there is any, on the type of interpersonal
+behavior a teacher shows is not expected to be strong. Had the HPD intervals not
+overlapped we could have concluded that as the self-efficacy increases, the
+score of the teacher on the IPC moves counterclockwise.\newline
 
@@
 
-The relation between the type and strength of interpersonal behavior in
-the CL-PN and CL-GPN model, is described by the parameters
-\(\gamma_{\cos}\) and \(\gamma_{\sin}\). The HPD interval of
-\(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
-models, meaning that the cosine component of the type of interpersonal
-behavior has an effect on the strength of interpersonal
-behavior.\newline \indent In the teacher data the sine and cosine
-components have a substantive meaning. This is illustrated in Figure
-\ref{QTI}. In a unit circle the horizontal axis (Communion) represents
-the cosine of and the vertical axis (Agency) represents the sine of an
-angle. For the teacher data this means that the Communion (cosine)
-dimension of the IPC positively effects the strength of a teachers' type
-of interpersonal behavior, in plain words: teachers exhibiting
-interpersonal behavior types with higher communion scores (e.g.,
-\enquote{helpful} and \enquote{understanding} in Figure 2) are stronger
-in their interpersonal behavior.\newline  \indent In the GPN-SSN model
-the dependence between the type and strengths of interpersonal behavior
-is modelled through the covariances between the linear outcome and the
-sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
+The relation between the type and strength of interpersonal behavior in the
+CL-PN and CL-GPN models is described by the parameters \(\gamma_{\cos}\) and
+\(\gamma_{\sin}\). The HPD interval of \(\gamma_{\cos}\) does not include 0 for
+both the CL-PN and CL-GPN models, meaning that the cosine component of the type
+of interpersonal behavior has an effect on the strength of interpersonal
+behavior.\newline
+\indent In the teacher data the sine and cosine components have a substantive
+meaning. This is illustrated in Figure \ref{QTI}. In a unit circle the
+horizontal axis (Communion) represents the cosine and the vertical axis
+(Agency) represents the sine of an angle. For the teacher data this means that
+the Communion (cosine) dimension of the IPC positively effects the strength of a
+teachers' type of interpersonal behavior, in plain words: teachers exhibiting
+interpersonal behavior types with higher communion scores (e.g., \enquote{helpful} and
+\enquote{understanding} in Figure 2) are stronger in their interpersonal behavior.\newline 
+\indent In the GPN-SSN model the dependence between the type and strengths of
+interpersonal behavior is modelled through the covariances between the linear
+outcome and the sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
 \(\sum_{sy_{1,3}}\). Both covariances, \(\sum_{sy_{2,3}} = 0.09\) and
-\(\sum_{sy_{1,3}} = 0.23\), are different from zero, but the one of the
-cosine component, and thus the correlation with the Communion dimension,
-is larger. This mean that teachers scoring both high on Communion and
-Agency show stronger behavior. Together with the results from the CL-PN
-and CL-GPN models in the previous paragraph this translates to the
-conclusion that teachers with the strongest interpersonal behavior have
-a type of interpersonal behavior between 0\(^\circ\) and 90\(^\circ\).
-To get these scores on the circle both the Agency and the Communion
-score of a Teacher have to be positive (see \eqref{PredVal}). This
-corresponds to the pattern observed in the teacher data in Figure
-\ref{QTI}. At a strength of 0.4 and up we see that the scores on the
-circle range on average between 0\(^\circ\) and 100\(^\circ\).
+\(\sum_{sy_{1,3}} = 0.23\), are different from zero, but the one of the cosine
+component, and thus the correlation with the Communion dimension, is larger.
+This means that teachers scoring both high on Communion and Agency show stronger
+behavior. Together with the results from the CL-PN and CL-GPN models in the
+previous paragraph this translates to the conclusion that teachers with the
+strongest interpersonal behavior have a type of interpersonal behavior between
+0\(^\circ\) and 90\(^\circ\). To get these scores on the circle both the Agency and
+the Communion score of a Teacher have to be positive (see \eqref{PredVal}). This
+corresponds to the pattern observed in the teacher data in Figure \ref{QTI}. At
+a strength of 0.4 and up we see that the scores on the circle range on average
+between 0\(^\circ\) and 100\(^\circ\).
 
@@
 
-Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the
-linear and circular outcomes of the four cylindrical models fit to the
-teacher data.\newline \indent The CL-PN and CL-GPN models have the best
-out-of-sample predictive performance for the linear outcome. They show
-roughly the same performance because they model the linear outcome in
-the same way. We should note that even though the predictive performance
-of the Abe-Ley model for the linear outcome is worst on average, the
-standard deviation of the cross-validation estimates is rather large.
-This means that in some samples, the Abe-Ley model shows a lower PLSL
-value than the average of 25.49\newline \indent The Abe-Ley model has
-the best out-of-sample predictive performance for the circular outcome.
-This would suggest that for the circular variable a slightly skewed
-distribution fits best. However, both the GPN-SSN and the CL-GPN models
-fit much worse even though the distribution for the circular outcome in
-these models can also take a skewed shape. It should be noted that the
-standard deviation of the cross-validation estimates is rather large for
-the Abe-Ley and the CL-GPN model. It is possible that these large
-standard deviations for the PLSL are caused by the fact that they are
-computed for a relatively small sample size, but this does not explain
-why the PLSL has a large standard deviation for only a few cylindrical
-models and not for all.\newline \indent In this situation, where one
-model fits the linear outcome best and another one fits the circular
-outcome best, it is hard to determine which model we should choose. In
-this case the results for the CL-PN /CL-GPN and Abe-Ley model are quite
-different regarding the effect of self-efficacy on the linear outcome
-(strength of interpersonal behavior). Because the Abe-Ley fit for the
-linear part is worst we would choose to trust the results for the CL-PN
-and CL-GPN model here. For the circular part however the results of the
-CL-PN/CL-GPN model do not differ as much from the Abe-Ley model and we
-reach the same conclusion for both models, namely that the effect of
-self-efficacy on type of interpersonal behavior is not very strong.
-Therefore we would prefer the CL-PN/CL-GPN models in this case because
-where it matters in terms of interpretation (the linear part) they show
-better fit.
+Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the linear
+and circular outcomes of the four cylindrical models fit to the teacher
+data.\newline \indent The CL-PN and CL-GPN models have the best out-of-sample
+predictive performance for the linear outcome. They show roughly the same
+performance because they model the linear outcome in the same way. We should
+note that even though the predictive performance of the Abe-Ley model for the
+linear outcome is worst on average, the standard deviation of the
+cross-validation estimates is rather large. This means that in some samples, the
+Abe-Ley model shows a lower PLSL value than the average of 25.49.\newline
+\indent The Abe-Ley model has the best out-of-sample predictive performance for
+the circular outcome. This would suggest that for the circular variable a
+slightly skewed distribution fits best. However, both the GPN-SSN and the
+CL-GPN models fit much worse even though the distribution for the circular
+outcome in these models can also take a skewed shape. It should be noted that
+the standard deviation of the cross-validation estimates is rather large for
+the Abe-Ley and the CL-GPN model. It is possible that these large standard
+deviations for the PLSL are caused by the fact that they are computed for a
+relatively small sample size, but this does not explain why the PLSL has a large
+standard deviation for only a few cylindrical models and not for all.\newline
+\indent In this situation, where one model fits the linear outcome best and
+another one fits the circular outcome best, it is hard to determine which model
+we should choose. In this case the results for the CL-PN /CL-GPN and Abe-Ley
+model are quite different regarding the effect of self-efficacy on the linear
+outcome (strength of interpersonal behavior). Because the Abe-Ley fit for the
+linear part is worst we would choose to trust the results for the CL-PN and
+CL-GPN models here. For the circular part however the results of the CL-PN/CL-GPN
+models do not differ as much from the Abe-Ley model and we reach the same
+conclusion for both models, namely that the effect of self-efficacy on type of
+interpersonal behavior is not very strong. Therefore we would prefer the
+CL-PN/CL-GPN models in this case because where it matters in terms of
+interpretation (the linear part) they show better fit.
 
@@
 
-In this paper we modified four models for cylindrical data in such a way
-that they include a regression of both the linear and circular outcome
-onto a set of covariates. Subsequently we have shown how these four
-methods can be used to analyze a dataset on the interpersonal behavior
-of teachers. In this final section we will first comment on what
-researchers can gain by using cylindrical models for the teacher data.
-Subsequently we will comment on the differences between the cylindrical
-models that were introduced in this paper.\newline \indent Concerning
-the teacher data, the advantage of cylindrical data analysis is that we
-were able to analyze the information about the type and strength of
-interpersonal behavior simultaneously. In previous research, the two
-components of the interpersonal circumplex (\emph{i.e.}, Agency and
-Communion) were analyzed separately. Such an approach also provides
-information about the strength of teachers' score on Agency and
-Communion, yet a large portion of information about the combination of
-Agency and Communion, which describes the type of behavior that is
-observed, gets lost. A first solution to include both dimensions as a
-circular variable in data analysis was described by Cremers et al.
-(2018a). A downside of that analysis was that information about the
-strength of the specific type of interpersonal behavior could not be
-retained. In the present study, we have shown how using cylindrical
-models can simultaneously model the information about the type of and
-strength of interpersonal behavior and how these are influenced by
-teachers' self-efficacy in classroom management. Although we do not find
-any strong effects of self-efficacy on either the type or strength of
-behavior, the four cylindrical models do provide a way of analyzing and
-interpreting this effect. This is beneficial for future research in
-which we may want to investigate the effect of further covariates on
-data from the circumplex. \newline \indent Furthermore, in addition to
-being able to assess the influence of covariates, the cylindrical models
-also provide information about the dependence between the type and
-strength of interpersonal behavior. We found that stronger behavior is
-associated with higher scores on the Communion and in some models also
-the Agency dimension. This implies that teachers whose type of
-interpersonal behavior ranges between 0\(^\circ\) and 90\(^\circ\), the
-\enquote{helpful} and \enquote{directing} subtypes are stronger in their
-behavior than teachers of the other subtypes.\newline \indent As
-mentioned in the introduction, data from the interpersonal circumplex is
-not the only type of cylindrical data that occurs in psychology. The
-methods presented in this paper are also of use for research on human
-navigation and eye-tracking research. Furthermore, even though
-cylindrical models are already used in fields outside of psychology, the
-addition of a regression structure to the models is of use in these
-fields as well. \newline \indent In terms of interpretability, the CL-PN
-and Abe-Ley models perform best out of the four cylindrical models. In
-the CL-GPN and GPN-SSN models the interpretation of the parameters of
-the circular outcome component is not straightforward, if at all
-possible. This is caused by the fact that in addition to the mean vector
-the covariance matrix of the GPN distribution affects the location of
-the circular data, making it difficult to compute regression
-coefficients on the circle. Wang and Gelfand (2013) state that Monte
-Carlo integration can be used to compute a circular mean and variance
-for the GPN distribution. In future research, this solution might be
-applied to the methods of Cremers et al. (2018b) in order to compute
-circular coefficients for GPN models.\newline \indent In terms of
-flexibility the GPN-SSN model scores best. Multiple linear and circular
-outcomes can be included and we can thus apply the model to multivariate
-cylindrical data. In addition the GPN-SSN, the CL-GPN and CL-PN models
-are extendable to a mixed-effects structure and can thus also be fit to
-longitudinal data (see Nuñez-Antonio and Gutiérrez-Peña (2014) and
-Hernandez-Stumpfhauser et al. (2016) for hierarchical/mixed-effects
-models for the PN and GPN distributions respectively). For the Abe-Ley
-model this may also be possible but has not been done in previous
-research for the conditional distribution of its circular outcome
-(sine-skewed von Mises). Concerning asymmetry, both the GPN-SSN as well
-as the Abe-Ley model allow for non-symmetrical shapes of the
-distributions of both the linear and circular outcome, while the CL-GPN
-model permits an asymmetric circular outcome.\newline \indent The four
-cylindrical models that were modified to the regression context in this
-paper are not the only cylindrical distributions available from the
-literature. Other interesting cylindrical distributions have been
-introduced by Fernández-Durán (2007), Kato and Shimizu (2008) and
-Sugasawa (2015) (for more references we refer to Chapter 2 of Ley and
-Verdebout (2017)). In the present study we have decided not to include
-these distributions for reasons of space, complexity of the models and
-ease of implementing a regression structure. In future research however
-it would be interesting to investigate other types of cylindrical
-distributions as well in order to compare the interpretability,
-flexibility and model fit to the models developed in the present
-study.\newline
+In this paper we modified four models for cylindrical data in such a way that
+they include a regression of both the linear and circular outcome onto a set of
+covariates. Subsequently we have shown how these four methods can be used to
+analyze a dataset on the interpersonal behavior of teachers. In this final
+section we will first comment on what researchers can gain by using cylindrical
+models for the teacher data. Subsequently we will comment on the differences
+between the cylindrical models that were introduced in this paper.\newline
+\indent Concerning the teacher data, the advantage of cylindrical data analysis
+is that we were able to analyze the information about the type and strength of
+interpersonal behavior simultaneously. In previous research, the two components
+of the interpersonal circumplex (\emph{i.e.}, Agency and Communion) were
+analyzed separately. Such an approach also provides information about the
+strength of teachers' score on Agency and Communion, yet a large portion of
+information about the combination of Agency and Communion, which describes the
+type of behavior that is observed, gets lost. A first solution to include both
+dimensions as a circular variable in data analysis was described by
+Cremers et al. (2018a). A downside of that analysis was that information about
+the strength of the specific type of interpersonal behavior could not be
+retained. In the present study, we have shown how using cylindrical models can
+simultaneously model the information about the type of and strength of
+interpersonal behavior and how these are influenced by teachers' self-efficacy
+in classroom management. Although we do not find any strong effects of
+self-efficacy on either the type or strength of behavior, the four cylindrical
+models do provide a way of analyzing and interpreting this effect. This is
+beneficial for future research in which we may want to investigate the effect of
+further covariates on data from the circumplex. \newline
+\indent Furthermore, in addition to being able to assess the influence of
+covariates, the cylindrical models also provide information about the dependence
+between the type and strength of interpersonal behavior. We found that stronger
+behavior is associated with higher scores on the Communion and in some models
+also the Agency dimension. This implies that teachers whose type of
+interpersonal behavior ranges between 0\(^\circ\) and 90\(^\circ\), the \enquote{helpful}
+and \enquote{directing} subtypes are stronger in their behavior than teachers of the
+other subtypes.\newline
+\indent As mentioned in the introduction, data from the interpersonal circumplex
+is not the only type of cylindrical data that occurs in psychology. The methods
+presented in this paper are also of use for research on human navigation and
+eye-tracking research. Furthermore, even though cylindrical models are already
+used in fields outside of psychology, the addition of a regression structure to
+the models is of use in these fields as well. \newline
+\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
+best out of the four cylindrical models. In the
+CL-GPN and GPN-SSN models the interpretation of the parameters of the
+circular outcome component is not straightforward, if at all possible.
+This is caused by the fact that in addition to the mean vector the
+covariance matrix of the GPN distribution affects the location of the
+circular data, making it difficult to compute regression coefficients on
+the circle. Wang and Gelfand (2013) state that Monte Carlo integration
+can be used to compute a circular mean and variance for the GPN
+distribution. In future research, this solution might be applied to the
+methods of Cremers et al. (2018b) in order to compute circular
+coefficients for GPN models.\newline
+\indent In terms of flexibility the GPN-SSN model scores best. Multiple
+linear and circular outcomes can be included and we can thus apply the
+model to multivariate cylindrical data. In addition the GPN-SSN, the
+CL-GPN and CL-PN models are extendable to a mixed-effects structure and
+can thus also be fit to longitudinal data (see Nuñez-Antonio and Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al. (2016) for
+hierarchical/mixed-effects models for the PN and GPN distributions
+respectively). For the Abe-Ley model this may also be possible but has
+not been done in previous research for the conditional distribution of
+its circular outcome (sine-skewed von
+Mises). Concerning asymmetry, both the
+GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
+the distributions of both the linear and circular outcome, while the CL-GPN
+model permits an asymmetric circular outcome.\newline
+\indent The four cylindrical models that were modified to the regression
+context in this paper are not the only cylindrical distributions
+available from the literature. Other interesting cylindrical distributions have been introduced by
+Fernández-Durán (2007), Kato and Shimizu (2008) and Sugasawa (2015) (for more references we
+refer to Chapter 2 of Ley and Verdebout (2017)). In the present study we
+have decided not to include these distributions for reasons of space,
+complexity of the models and ease of implementing a regression
+structure. In future research however it would be interesting to
+investigate other types of cylindrical distributions as well in order to
+compare the interpretability, flexibility and model fit to the models
+developed in the present study.\newline
 
 \section*{References}
-
-\DeclareRobustCommand{\VANDER}[3]{#3}
-\DeclareRobustCommand{\VAN}[3]{#3}
-\DeclareRobustCommand{\DEN}[3]{#3}
 
@@
 \leavevmode\hypertarget{ref-abe2017tractable}{}%
-Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
-for cylindrical data, with applications. \emph{Econometrics and
-Statistics}, \emph{4}, 91--104.
+Abe, T., \& Ley, C.
+(2017).
+A tractable, parsimonious and flexible model for cylindrical data, with applications.
+\emph{Econometrics and Statistics}, \emph{4}, 91--104.
 doi:\href{https://doi.org/10.1016/j.ecosta.2016.04.001}{10.1016/j.ecosta.2016.04.001}
@@
 \leavevmode\hypertarget{ref-chrastil2017rotational}{}%
-Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
-integration: Encoding and execution errors in angle reproduction.
+Chrastil, E.
+R., \& Warren, W.
+H.
+(2017).
+Rotational error in path integration: Encoding and execution errors in angle reproduction.
 \emph{Experimental Brain Research}, \emph{235}(6), 1885--1897.
@@
 \leavevmode\hypertarget{ref-Claessens2016side}{}%
-Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
-Teachers' perceptions of teacher--student relationships} (PhD thesis).
+Claessens, L. C. (2016). \emph{Be on my side i'll be on your side : Teachers' perceptions of teacher--student relationships} (PhD thesis).
 
 \leavevmode\hypertarget{ref-Cremers2018Assessing}{}%
-Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
-Bayesian embedding approach to circular regression models.
-\emph{Methodology}, \emph{14}(2), 69--81.
+Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a Bayesian embedding approach to circular regression models. \emph{Methodology}, \emph{14}(2), 69--81.
 
 \leavevmode\hypertarget{ref-CremersMulderKlugkist2017}{}%
-Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
-interpretation of regression coefficients. \emph{British Journal of
-Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
+Cremers, J., Mulder, K.
+T., \& Klugkist, I.
+(2018b).
+Circular interpretation of regression coefficients.
+\emph{British Journal of Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
 doi:\href{https://doi.org/10.1111/bmsp.12108}{10.1111/bmsp.12108}
@@
 \leavevmode\hypertarget{ref-fernandez2007models}{}%
-Fernández-Durán, J. (2007). Models for circular--linear and
-circular--circular data constructed from circular distributions based on
-nonnegative trigonometric sums. \emph{Biometrics}, \emph{63}(2),
-579--585.
+Fernández-Durán, J.
+(2007).
+Models for circular--linear and circular--circular data constructed from circular distributions based on nonnegative trigonometric sums.
+\emph{Biometrics}, \emph{63}(2), 579--585.
 doi:\href{https://doi.org/10.1111/j.1541-0420.2006.00716.x}{10.1111/j.1541-0420.2006.00716.x}
@@
 \leavevmode\hypertarget{ref-fisher1995statistical}{}%
-Fisher, N. I. (1995). \emph{Statistical analysis of circular data}.
-Cambridge: Cambridge University Press.
+Fisher, N. I. (1995). \emph{Statistical analysis of circular data}. Cambridge: Cambridge University Press.
 
 \leavevmode\hypertarget{ref-garcia2014test}{}%
-García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
-González-Manteiga, W., \& Pereira, J. (2014). A test for
-directional-linear independence, with applications to wildfire
-orientation and size. \emph{Stochastic Environmental Research and Risk
-Assessment}, \emph{28}(5), 1261--1275.
+García-Portugués, E., Barros, A.
+M., Crujeiras, R.
+M., González-Manteiga, W., \& Pereira, J.
+(2014).
+A test for directional-linear independence, with applications to wildfire orientation and size.
+\emph{Stochastic Environmental Research and Risk Assessment}, \emph{28}(5), 1261--1275.
 doi:\href{https://doi.org/10.1007/s00477-013-0819-6}{10.1007/s00477-013-0819-6}
@@
 \leavevmode\hypertarget{ref-garcia2013exploring}{}%
-García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
-Exploring wind direction and SO2 concentration by circular--linear
-density estimation. \emph{Stochastic Environmental Research and Risk
-Assessment}, \emph{27}(5), 1055--1067.
+García-Portugués, E., Crujeiras, R.
+M., \& González-Manteiga, W.
+(2013).
+Exploring wind direction and SO2 concentration by circular--linear density estimation.
+\emph{Stochastic Environmental Research and Risk Assessment}, \emph{27}(5), 1055--1067.
 doi:\href{https://doi.org/10.1007/s00477-012-0642-5}{10.1007/s00477-012-0642-5}
@@
 \leavevmode\hypertarget{ref-BDA}{}%
-Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
-(2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman
-\& Hall/CRC.
+Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D. (2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman \& Hall/CRC.
 
 \leavevmode\hypertarget{ref-gneiting2007strictly}{}%
-Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
-prediction, and estimation. \emph{Journal of the American Statistical
-Association}, \emph{102}(477), 359--378.
+Gneiting, T., \& Raftery, A.
+E.
+(2007).
+Strictly proper scoring rules, prediction, and estimation.
+\emph{Journal of the American Statistical Association}, \emph{102}(477), 359--378.
 doi:\href{https://doi.org/10.1198/016214506000001437}{10.1198/016214506000001437}
@@
 \leavevmode\hypertarget{ref-gurtman2009exploring}{}%
-Gurtman, M. B. (2009). Exploring personality with the interpersonal
-circumplex. \emph{Social and Personality Psychology Compass},
-\emph{3}(4), 601--619.
+Gurtman, M.
+B.
+(2009).
+Exploring personality with the interpersonal circumplex.
+\emph{Social and Personality Psychology Compass}, \emph{3}(4), 601--619.
 doi:\href{https://doi.org/10.1111/j.1751-9004.2009.00172.x}{10.1111/j.1751-9004.2009.00172.x}
@@
 \leavevmode\hypertarget{ref-gurtman2011reasoning}{}%
-Gurtman, M. B. (2011). Handbook of interpersonal psychology. In L. M.
-Horowitz \& S. Strack (Eds.), (pp. 299--311). New York: Wiley.
+Gurtman, M. B. (2011). Handbook of interpersonal psychology. In L. M. Horowitz \& S. Strack (Eds.), (pp. 299--311). New York: Wiley.
 
 \leavevmode\hypertarget{ref-hernandez2016general}{}%
-Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
-(2016). The general projected normal distribution of arbitrary
-dimension: Modeling and Bayesian inference. \emph{Bayesian Analysis},
-\emph{12}(1), 113--133.
+Hernandez-Stumpfhauser, D., Breidt, F.
+J., \& Woerd, M.
+J.
+van der.
+(2016).
+The general projected normal distribution of arbitrary dimension: Modeling and Bayesian inference.
+\emph{Bayesian Analysis}, \emph{12}(1), 113--133.
 doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}
@@
 \leavevmode\hypertarget{ref-horowitz2010handbook}{}%
-Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
-psychology: Theory, research, assessment, and therapeutic
-interventions}. Hoboken, NJ: John Wiley \& Sons.
+Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal psychology: Theory, research, assessment, and therapeutic interventions}. Hoboken, NJ: John Wiley \& Sons.
 
 \leavevmode\hypertarget{ref-jammalamadaka2001topics}{}%
-Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
-statistics} (Vol. 5). World Scientific.
+Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular statistics} (Vol. 5). World Scientific.
 
 \leavevmode\hypertarget{ref-johnson1978some}{}%
-Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
-distributions and related regression models. \emph{Journal of the
-American Statistical Association}, \emph{73}(363), 602--606.
+Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear distributions and related regression models. \emph{Journal of the American Statistical Association}, \emph{73}(363), 602--606.
 
 \leavevmode\hypertarget{ref-kato2008dependent}{}%
-Kato, S., \& Shimizu, K. (2008). Dependent models for observations which
-include angular ones. \emph{Journal of Statistical Planning and
-Inference}, \emph{138}(11), 3538--3549.
+Kato, S., \& Shimizu, K.
+(2008).
+Dependent models for observations which include angular ones.
+\emph{Journal of Statistical Planning and Inference}, \emph{138}(11), 3538--3549.
 doi:\href{https://doi.org/10.1016/j.jspi.2006.12.009}{10.1016/j.jspi.2006.12.009}
@@
 \leavevmode\hypertarget{ref-lagona2015hidden}{}%
-Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
-Markov approach to the analysis of space--time environmental data with
-linear and circular components. \emph{Stochastic Environmental Research
-and Risk Assessment}, \emph{29}(2), 397--409.
+Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S.
+(2015).
+A hidden Markov approach to the analysis of space--time environmental data with linear and circular components.
+\emph{Stochastic Environmental Research and Risk Assessment}, \emph{29}(2), 397--409.
 doi:\href{https://doi.org/10.1007/s00477-014-0919-y}{10.1007/s00477-014-0919-y}
@@
 \leavevmode\hypertarget{ref-leary1957}{}%
-Leary, T. (1957). \emph{An interpersonal diagnosis of personality}. New
-York: Ronald Press Company.
+Leary, T. (1957). \emph{An interpersonal diagnosis of personality}. New York: Ronald Press Company.
 
 \leavevmode\hypertarget{ref-ley2017modern}{}%
-Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
-Chapman \& Hall/CRC Press. Boca Raton, FL.
+Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}. Chapman \& Hall/CRC Press. Boca Raton, FL.
 
 \leavevmode\hypertarget{ref-locke2007selfefficacy}{}%
-Locke, K. D., \& Sadler, P. (2007). Self-efficacy, values, and
-complementarity in dyadic interactions: Integrating interpersonal and
-social-cognitive theory. \emph{Personality and Social Psychology
-Bulletin}, \emph{33}, 94--109.
+Locke, K.
+D., \& Sadler, P.
+(2007).
+Self-efficacy, values, and complementarity in dyadic interactions: Integrating interpersonal and social-cognitive theory.
+\emph{Personality and Social Psychology Bulletin}, \emph{33}, 94--109.
 doi:\href{https://doi.org/10.1177/0146167206293375}{10.1177/0146167206293375}
@@
 \leavevmode\hypertarget{ref-mardia2000directional}{}%
-Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
-(Vol. 494). Chichester, England: Wiley.
+Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics} (Vol. 494). Chichester, England: Wiley.
 
 \leavevmode\hypertarget{ref-mardia1978model}{}%
-Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical
-variables with applications. \emph{Journal of the Royal Statistical
-Society Series B (Methodological)}, \emph{40}(2), 229--233.
+Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical variables with applications. \emph{Journal of the Royal Statistical Society Series B (Methodological)}, \emph{40}(2), 229--233.
 
 \leavevmode\hypertarget{ref-mastrantonio2018joint}{}%
-Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
-distribution for poly-cylindrical data. \emph{Journal of Multivariate
-Analysis}, \emph{165}, 14--26.
+Mastrantonio, G.
+(2018).
+The joint projected normal and skew-normal: A distribution for poly-cylindrical data.
+\emph{Journal of Multivariate Analysis}, \emph{165}, 14--26.
 doi:\href{https://doi.org/10.1016/j.jmva.2017.11.006}{10.1016/j.jmva.2017.11.006}
@@
 \leavevmode\hypertarget{ref-mastrantonio2015bayesian}{}%
-Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
-hidden Markov modelling using circular-linear general projected normal
-distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
+Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G.
+(2015).
+Bayesian hidden Markov modelling using circular-linear general projected normal distribution.
+\emph{Environmetrics}, \emph{26}(2), 145--158.
 doi:\href{https://doi.org/10.1002/env.2326}{10.1002/env.2326}
@@
 \leavevmode\hypertarget{ref-nunez2014bayesian}{}%
-Nuñez-Antonio, G., \& Gutiérrez-Peña, E. (2014). A Bayesian model for
-longitudinal circular data based on the projected normal distribution.
+Nuñez-Antonio, G., \& Gutiérrez-Peña, E.
+(2014).
+A Bayesian model for longitudinal circular data based on the projected normal distribution.
 \emph{Computational Statistics \& Data Analysis}, \emph{71}, 506--519.
@@
 \leavevmode\hypertarget{ref-nunez2011bayesian}{}%
-Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
-Bayesian regression model for circular data based on the projected
-normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
-185--201.
+Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G.
+(2011).
+A Bayesian regression model for circular data based on the projected normal distribution.
+\emph{Statistical Modelling}, \emph{11}(3), 185--201.
 doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
@@
 \leavevmode\hypertarget{ref-pennings2017phd}{}%
-Pennings, H. J. M. (2017a). \emph{Interpersonal dynamics in
-teacher-student interactions and relationships} (PhD thesis). Enschede:
-Ipskamp drukkers.
+Pennings, H. J. M. (2017a). \emph{Interpersonal dynamics in teacher-student interactions and relationships} (PhD thesis). Enschede: Ipskamp drukkers.
 
 \leavevmode\hypertarget{ref-pennings2017complexity}{}%
-Pennings, H. J. M. (2017b). Using a complexity approach to study the
-interpersonal dynamics in teacher-student interactions: A case study of
-two teachers. \emph{Complicity: An International Journal of Complexity
-and Education}, \emph{14}(2), 88--103.
+Pennings, H.
+J.
+M.
+(2017b).
+Using a complexity approach to study the interpersonal dynamics in teacher-student interactions: A case study of two teachers.
+\emph{Complicity: An International Journal of Complexity and Education}, \emph{14}(2), 88--103.
 
 \leavevmode\hypertarget{ref-pennings2018interpersonal}{}%
-Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C.,
-\VANDER{Want}{Van der}{van der} Want, A. C., \& \VAN{Tartwijk}{Van}{van}
-Tartwijk, J. (2018). Interpersonal adaptation in teacher-student
-interaction. \emph{Learning and Instruction}, \emph{55}, 41--57.
+Pennings, H.
+J.
+M., Brekelmans, M., Sadler, P., Claessens, L.
+C., \VANDER{Want}{Van der}{van der} Want, A.
+C., \& \VAN{Tartwijk}{Van}{van} Tartwijk, J.
+(2018).
+Interpersonal adaptation in teacher-student interaction.
+\emph{Learning and Instruction}, \emph{55}, 41--57.
 doi:\href{https://doi.org/10.1016/j.learninstruc.2017.09.005}{10.1016/j.learninstruc.2017.09.005}
@@
 \leavevmode\hypertarget{ref-plutchik1997general}{}%
-Plutchik, R. (1997). Circumplex models of personality and emotions. In
-C. Plutchik Robert. (Ed.), (pp. 17--45). Washington, DC, US: American
-Psychological Association.
+Plutchik, R. (1997). Circumplex models of personality and emotions. In C. Plutchik Robert. (Ed.), (pp. 17--45). Washington, DC, US: American Psychological Association.
 
 \leavevmode\hypertarget{ref-rayner200935th}{}%
-Rayner, K. (2009). The 35th Sir Frederick Bartlett lecture: Eye
-movements and attention in reading, scene perception, and visual search.
-\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8),
-1457--1506.
+Rayner, K.
+(2009).
+The 35th Sir Frederick Bartlett lecture: Eye movements and attention in reading, scene perception, and visual search.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8), 1457--1506.
 doi:\href{https://doi.org/10.1080/17470210902816461}{10.1080/17470210902816461}
@@
 \leavevmode\hypertarget{ref-russell1980circumplex}{}%
-Russell, J. A. (1980). A circumplex model of affect. \emph{Journal of
-Personality and Social Psychology}, \emph{39}(6), 1161--1178.
-doi:\href{https://doi.org/10.1037/h0077714}{10.1037/h0077714}
+Russell, J. A. (1980). A circumplex model of affect. \emph{Journal of Personality and Social Psychology}, \emph{39}(6), 1161--1178. doi:\href{https://doi.org/10.1037/h0077714}{10.1037/h0077714}
 
 \leavevmode\hypertarget{ref-sahu2003new}{}%
-Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
-multivariate skew distributions with applications to Bayesian regression
-models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
+Sahu, S.
+K., Dey, D.
+K., \& Branco, M.
+D.
+(2003).
+A new class of multivariate skew distributions with applications to Bayesian regression models.
+\emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
 doi:\href{https://doi.org/10.2307/3316064}{10.2307/3316064}
@@
 \leavevmode\hypertarget{ref-schwartz1992values}{}%
-Schwartz, S. H. (1992). Advances in experimental social psychology. In
-M. Zanna (Ed.), (Vol. 25, pp. 1--65). San Diego, CA: Academic Press.
+Schwartz, S. H. (1992). Advances in experimental social psychology. In M. Zanna (Ed.), (Vol. 25, pp. 1--65). San Diego, CA: Academic Press.
 
 \leavevmode\hypertarget{ref-sugasawa2015flexible}{}%
-Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the
-cylinder}. Retrieved from
-\href{arXiv:\%201501.06332v2}{arXiv: 1501.06332v2}
+Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the cylinder}. Retrieved from \href{arXiv:\%201501.06332v2}{arXiv: 1501.06332v2}
 
 \leavevmode\hypertarget{ref-wang2012directional}{}%
-Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
-general projected normal distribution. \emph{Statistical Methodology},
-\emph{10}(1), 113--127.
+Wang, F., \& Gelfand, A.
+E.
+(2013).
+Directional data analysis under the general projected normal distribution.
+\emph{Statistical Methodology}, \emph{10}(1), 113--127.
 doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}
@@
 \leavevmode\hypertarget{ref-vanderWant2015role}{}%
-\VANDER{Want}{Van der}{van der} Want, A. C. (2015). \emph{Teachers'
-interpersonal role identity.} (PhD thesis).
+\VANDER{Want}{Van der}{van der} Want, A. C. (2015). \emph{Teachers' interpersonal role identity.} (PhD thesis).
 
 \leavevmode\hypertarget{ref-want2018selfefficacy}{}%
-\VANDER{Want}{Van der}{van der} Want, A. C., \DEN{Brok}{Den}{den} Brok,
-P., Beijaard, D., Brekelmans, M., Claessens, L. C., \& Pennings, H. J.
-M. (2018). The relation between teachers' interpersonal role identity
-and their self-efficacy, burnout and work engagement. \emph{Professional
-Development in Education}, Advance Online Publication.
+\VANDER{Want}{Van der}{van der} Want, A.
+C., \DEN{Brok}{Den}{den} Brok, P., Beijaard, D., Brekelmans, M., Claessens, L.
+C., \& Pennings, H.
+J.
+M.
+(2018).
+The relation between teachers' interpersonal role identity and their self-efficacy, burnout and work engagement.
+\emph{Professional Development in Education}, Advance Online Publication.
 doi:\href{https://doi.org/10.1080/19415257.2018.1511453}{10.1080/19415257.2018.1511453}
@@
 \leavevmode\hypertarget{ref-wiggins1996history}{}%
-Wiggins, J. (1996). An informal history of the interpersonal circumplex
-tradition. \emph{Journal of Personality Assessment}, \emph{66}(2),
-217--233.
+Wiggins, J. (1996). An informal history of the interpersonal circumplex tradition. \emph{Journal of Personality Assessment}, \emph{66}(2), 217--233.
 
 \leavevmode\hypertarget{ref-wright2009integrating}{}%
-Wright, A. G., Pincus, A. L., Conroy, D. E., \& Hilsenroth, M. J.
-(2009). Integrating methods to optimize circumplex description and
-comparison of groups. \emph{Journal of Personality Assessment},
-\emph{91}(4), 311--322.
+Wright, A.
+G., Pincus, A.
+L., Conroy, D.
+E., \& Hilsenroth, M.
+J.
+(2009).
+Integrating methods to optimize circumplex description and comparison of groups.
+\emph{Journal of Personality Assessment}, \emph{91}(4), 311--322.
 doi:\href{https://doi.org/10.1080/00223890902935696}{10.1080/00223890902935696}
@@
 \leavevmode\hypertarget{ref-wubbels2006interpersonal}{}%
-Wubbels, T., Brekelmans, M., \DEN{Brok}{Den}{den} Brok, P., \&
-\VAN{Tartwijk}{Van}{van} Tartwijk, J. (2006). Handbook of classroom
-management: Research practice and contemporary issues. In C. Evertson \&
-C. S. Weinstein (Eds.), \emph{Handbook of classroom management:
-Research, practice, and contemporary issues} (pp. 1161--1191). Malwah,
-NJ: Lawrence Erlbaum Associates.
+Wubbels, T., Brekelmans, M., \DEN{Brok}{Den}{den} Brok, P., \& \VAN{Tartwijk}{Van}{van} Tartwijk, J.
+(2006).
+Handbook of classroom management: Research practice and contemporary issues.
+In C.
+Evertson \& C.
+S.
+Weinstein (Eds.), \emph{Handbook of classroom management: Research, practice, and contemporary issues} (pp.
+1161--1191).
+Malwah, NJ: Lawrence Erlbaum Associates.
 

### 2019-03-03 Manuscript/SupplementaryMaterial.Rmd
Manuscript Update

@@
 Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in the
-Supplementary Material.
+@hernandez2016general for the regression of the circular outcome.
 
@@
    &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
-   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
+   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
    +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
-   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
+   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
 \end{align}
@@
 and the $g$ covariate values. \newline
-\indent We estimate the model using MCMC methods. A detailed description of these
-methods is given in the Supplementary Material.
 
@@
 
-\indent We use the following loglikelihoods for the
+\indent We use the following (conditional) loglikelihoods for the
 computation of the PLSL in the teacher data:
@@
 
-$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+$$l(y \mid \theta, r) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
 \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$
 
-$$l(\theta) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
+$$l(\theta, r) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
 where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
@@
 
-$$l(y) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
+$$l(y \mid \theta, r) = \log(1) - \log(\sqrt{2\pi\sigma^2}) + \sum(\hat{y}_i-(\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
 \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i))^2/2\sigma^2$$
@@
 
-$$l(y) = \log\alpha + \sum\log h_i^\alpha + \sum\log y_i^{\alpha - 1} - \sum(h_iy_i)^\alpha$$
+$$l(y \mid \theta) = \log\alpha + \sum\log h_i^\alpha + \sum\log y_i^{\alpha - 1} - \sum(h_iy_i)^\alpha$$
 
-where $h_i = \exp\{\hat{y}_i\}\{1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i)\}^{1/\alpha}$, $\hat{y}_i = \gamma_0 + \gamma_1\text{SE}_i$ and $\hat{\theta}_i = \beta_0 +  2\tan^{-1}(\beta_1\text{SE}_i))$ .
+where $h_i = \exp(\hat{y}_i)\{1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i)\}^{1/\alpha}$, $\hat{y}_i = \gamma_0 + \gamma_1\text{SE}_i$ and $\hat{\theta}_i = \beta_0 +  2\tan^{-1}(\beta_1\text{SE}_i))$ .
 
-$$l(\theta) = \log(1) - \sum\log 2\pi I_0(c_i) + \sum\log\{1 + \lambda \sin(\theta_i - \hat{\theta}_i\} + \sum c_i\cos(\theta_i - \hat{\theta}_i)$$
+$$l(\theta \mid y) = \log(1) - \sum\log 2\pi I_0(c_i) + \sum\log\{1 + \lambda \sin(\theta_i - \hat{\theta}_i\} + \sum c_i\cos(\theta_i - \hat{\theta}_i)$$
 
-where $c_i = y_i^{\alpha}\exp(\hat{y}_i)^{\alpha}\tanh\kappa$, and $I_{0}$ is a Bessel function of order $0$.
+where $c_i = y_i^{\alpha}\exp(\hat{y}_i)^{\alpha}\tanh\kappa$, and $I_{0}$ is a modified Bessel function of order $0$.
 
@@
 \end{enumerate}
-
+\newpage
 \section*{References}

### 2019-03-03 Manuscript/SupplementaryMaterial.tex
Manuscript Update

@@
 Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
-regression of the circular outcome. A detailed description of the
-Bayesian estimation and MCMC samplers can be found in the Supplementary
-Material.
+regression of the circular outcome.
 
@@
    &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
-   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
+   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
    +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
-   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
+   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
 \end{align} \noindent We can use numerical optimization (Nelder-Mead) to
@@
 \(g + 1\) dimensional vector containing the value 1 to estimate an
-intercept and the \(g\) covariate values. \newline \indent We estimate
-the model using MCMC methods. A detailed description of these methods is
-given in the Supplementary Material.
+intercept and the \(g\) covariate values. \newline
 
@@
 
-\indent We use the following loglikelihoods for the computation of the
-PLSL in the teacher data:
+\indent We use the following (conditional) loglikelihoods for the
+computation of the PLSL in the teacher data:
 

### 2019-03-06 Manuscript/Manuscript.Rmd
Manuscript Update

@@
   - \usepackage{subcaption}
+  - \raggedbottom
    #% set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)
@@
 ```
-
-\section{Introduction}\label{Introduction}
-
+\vspace{-0.75cm}
 In social sciences the use of cylindrical data is very common. Such data consist
@@
 cylindrical. For example, research on human navigation uses data where distance
-(i.e., linear component) and direction (i.e., the circular component) are of
+(i.e., the linear component) and direction (i.e., the circular component) are of
 interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
@@
 
-\subsection{Cylindrical data}
-
+\vspace{-0.75cm}
+\subsection{Cylindrical Data}
+\vspace{-0.75cm}
 Data that consist of a linear variable and a circular variable is called
@@
 
+\vspace{-0.75cm}
 \section{Teacher Data}\label{Example}
-
+\vspace{-0.75cm}
 \begin{figure}
@@
 
+\vspace{-0.75cm}
 \subsection{Data Description}\label{DataDescriptives}
-
+\vspace{-0.75cm}
 ```{r loaddata, return = FALSE, echo = FALSE, results = FALSE}
@@
 
+\vspace{-0.75cm}
 \section{Four Cylindrical Regression Models}\label{Models}
-
+\vspace{-0.75cm}
 One of the goals of this paper is to show the benefits of cylindrical methods
@@
 
+\vspace{-0.75cm}
 \subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
-
+\vspace{-0.75cm}
 Following @mastrantonio2015bayesian we consider two models where
@@
 & \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
-asymetry/skewness.\newline
+asymmetry/skewness.\newline
 \indent For the teacher dataset the predicted linear outcome, strengths of
@@
 
+\vspace{-0.75cm}
 \subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
-
+\vspace{-0.75cm}
 This model is an extension of the cylindrical model introduced in
@@
 
+\vspace{-0.75cm}
 \subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
-
+\vspace{-0.75cm}
 This model is an extension of the cylindrical model introduced by
@@
 matrix, also assess the dependence between the type and strength of interpersonal
-behavior. \newline
+behavior. \newpage
 \indent We estimate the model using MCMC methods. A detailed description of these
@@
 
+\vspace{-0.75cm}
 \subsection{Model Fit Criterion}\label{Modelfit}
-
+\vspace{-0.75cm}
 For the four cylindrical models we focus on their out-of-sample predictive
@@
 
-\section{Data Analysis}\label{DataAnalysis}
-
+\vspace{-0.75cm}
+\section{Results}\label{DataAnalysis}
+\vspace{-0.75cm}
 In this section we analyze the teacher data with the help of the four
@@
 
-\subsection{Results \& Analysis}\label{DataResults}
-
+\vspace{-0.75cm}
+\subsection{Analysis}\label{DataResults}
+\vspace{-0.75cm}
 In the Supplementary Material we have described the starting values for the MCMC
@@
 
+\vspace{-0.5cm}
 \subsubsection{Average type and strength of interpersonal behavior}
@@
 
+\vspace{-0.5cm}
 \subsubsection{The effect of self-efficacy}
-
 The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley models and $\beta_{1_y}$
@@
 overlapped we could have concluded that as the self-efficacy increases, the
-score of the teacher on the IPC moves counterclockwise.\newline
+score of the teacher on the IPC moves counterclockwise.
 
@@
 \end{figure}
-
+\vspace{-0.5cm}
 \subsubsection{Dependence between type and strength of interpersonal behavior.}
-
 The relation between the type and strength of interpersonal behavior in the
@@
 
+\vspace{-0.5cm}
 \subsubsection{Model fit}
-
 Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the linear
@@
 
+\vspace{-0.75cm}
 \section{Discussion}\label{Discussion}
-
+\vspace{-0.75cm}
 In this paper we modified four models for cylindrical data in such a way that
@@
 compare the interpretability, flexibility and model fit to the models
-developed in the present study.\newline
-
-\section*{References}
+developed in the present study.
+
+\newpage
+\vspace{-0.75cm}
+\section*{{\normalfont References}}
+\vspace{-0.75cm}
 <div id="refs"></div>

### 2019-03-06 Manuscript/Manuscript.tex
Manuscript Update

@@
 \usepackage{subcaption}
+\raggedbottom
 
@@
 
-\section{Introduction}\label{Introduction}
+\vspace{-0.75cm}
 
@@
 cylindrical. For example, research on human navigation uses data where distance
-(i.e., linear component) and direction (i.e., the circular component) are of
+(i.e., the linear component) and direction (i.e., the circular component) are of
 interest (Chrastil \& Warren, 2017) or in eye-tracking, the saccade data also
@@
 
-\subsection{Cylindrical data}
+\vspace{-0.75cm}
+\subsection{Cylindrical Data}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.75cm}
 \subsection{Data Description}\label{DataDescriptives}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.75cm}
 \section{Four Cylindrical Regression Models}\label{Models}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.75cm}
 \subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+\vspace{-0.75cm}
 
@@
 for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
-asymetry/skewness.\newline
+asymmetry/skewness.\newline
 \indent For the teacher dataset the predicted linear outcome, strengths of
@@
 
+\vspace{-0.75cm}
 \subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.75cm}
 \subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.75cm}
 \subsection{Model Fit Criterion}\label{Modelfit}
+\vspace{-0.75cm}
 
@@
 
-\section{Data Analysis}\label{DataAnalysis}
+\vspace{-0.75cm}
+\section{Results}\label{DataAnalysis}
+\vspace{-0.75cm}
 
@@
 
-\subsection{Results \& Analysis}\label{DataResults}
+\vspace{-0.75cm}
+\subsection{Analysis}\label{DataResults}
+\vspace{-0.75cm}
 
@@
 
+\vspace{-0.5cm}
 \subsubsection{The effect of self-efficacy}
@@
 \end{figure}
-
+\vspace{-0.5cm}
 \subsubsection{Dependence between type and strength of interpersonal behavior.}
@@
 
+\vspace{-0.5cm}
 \subsubsection{Model fit}
@@
 
+\vspace{-0.75cm}
 \section{Discussion}\label{Discussion}
+\vspace{-0.75cm}
 
@@
 compare the interpretability, flexibility and model fit to the models
-developed in the present study.\newline
-
-\section*{References}
+developed in the present study.
+
+\newpage
+\vspace{-0.75cm}
+\section*{{\normalfont References}}
+\vspace{-0.75cm}
 

### 2019-03-08 Manuscript/Manuscript.tex
Manuscript update
Manuscript Update

@@
-\documentclass[man]{apa6}
+\documentclass[man,mask]{apa6}
 \usepackage{lmodern}
@@
 \vspace{0.5cm}
-\textsuperscript{1} Department of Methodology and Statistics, Utrecht University\\\textsuperscript{2} TNO\\\textsuperscript{3} Department of Education, Utrecht University\\\textsuperscript{4} Department of Applied Mathematics, Computer Science and Statistics, Ghent University}
+\textsuperscript{1} Department of Biostatistics, University of Copenhagen\\\textsuperscript{2} TNO\\\textsuperscript{3} Department of Education, Utrecht University\\\textsuperscript{4} Department of Applied Mathematics, Computer Science and Statistics, Ghent University}
 \keywords{cylindrical data, regression, interpersonal behavior}
@@
 
-Correspondence concerning this article should be addressed to Jolien Cremers, . E-mail: \href{mailto:joliencremers@gmail.com}{\nolinkurl{joliencremers@gmail.com}}}
+Correspondence concerning this article should be addressed to Jolien Cremers, . E-mail: \href{mailto:jcre@sund.ku.dk}{\nolinkurl{jcre@sund.ku.dk}}}
 
@@
 Carlo (MCMC) methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011), Wang and Gelfand (2013) and
-Hernandez-Stumpfhauser, Breidt, and Woerd (2016) for the regression of the circular outcome. A detailed
+Hernandez-Stumpfhauser, Breidt, and \VANDER{Woerd}{Van der}{van der} Woerd (2016) for the regression of the circular outcome. A detailed
 description of the Bayesian estimation and MCMC samplers can be found in the
@@
 Hernandez-Stumpfhauser, D., Breidt, F.
-J., \& Woerd, M.
+J., \& \VANDER{Woerd}{Van der}{van der} Woerd, M.
 J.
-van der.
 (2016).

### 2019-03-08 Manuscript/SupplementaryMaterial.tex
Supplementary Material Update
Manuscript Update

@@
 \doublespacing
+\DeclareRobustCommand{\VANDER}[3]{#2}
+\DeclareRobustCommand{\VAN}[3]{#2}
+\DeclareRobustCommand{\DEN}[3]{#2}
 
@@
 and \(\xi,\tau \in (-\infty, +\infty)\) (as in Hernandez-Stumpfhauser,
-Breidt, \& Woerd (2016)). This leads to the circular-linear PN (CL-PN)
-and circular-linear GPN (CL-GPN) distributions. We will now detail how
-we modify both cylindrical distributions to also incorporate covariates
-for the circular part.
+Breidt, \& \VANDER{Woerd}{Van der}{van der} Woerd (2016)). This leads to
+the circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN)
+distributions. We will now detail how we modify both cylindrical
+distributions to also incorporate covariates for the circular part.
 
@@
 
-$$l(\theta) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
+$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
 where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
@@
 \leavevmode\hypertarget{ref-hernandez2016general}{}%
-Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
-(2016). The general projected normal distribution of arbitrary
-dimension: Modeling and Bayesian inference. \emph{Bayesian Analysis},
-\emph{12}(1), 113--133.
+Hernandez-Stumpfhauser, D., Breidt, F. J., \&
+\VANDER{Woerd}{Van der}{van der} Woerd, M. J. (2016). The general
+projected normal distribution of arbitrary dimension: Modeling and
+Bayesian inference. \emph{Bayesian Analysis}, \emph{12}(1), 113--133.
 doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}

### 2019-07-26 Manuscript/Manuscript.Rmd
Manuscript update
Corrections Christophe

@@
     address       : ""
-    email         : "joliencremers@gmail.com"
+    email         : "jcre@sund.ku.dk"
   - name          : "Helena J.M. Pennings"
@@
   - id            : "1"
-    institution   : "Department of Methodology and Statistics, Utrecht University"
+    institution   : "Department of Biostatistics, University of Copenhagen"
   - id            : "2"
@@
   movement), eye-tracking research (direction and length of saccades) and
-  data from an interpersonal circumplex (type and strength of
-  interpersonal behavior). In this paper we adapt four models for
+  data from an interpersonal circumplex (location and intensity on the IPC). 
+  In this paper we adapt four models for
   cylindrical data to include a regression of the circular and linear
@@
 
+floatsintext      : no
 figurelist        : no
@@
 linenumbers       : yes
-mask              : no
+mask              : yes
 draft             : no
@@
 ```
-\vspace{-0.75cm}
-In social sciences the use of cylindrical data is very common. Such data consist
-of a linear and a circular component. @gurtman2011reasoning refers to such data
-as vectors, with a directional measure (i.e., the circular component) and a
-measure indicating the magnitude (i.e., the linear component). Many established
-models in psychology are often referred to as circular or circumplex models, but
-those models are cylindrical. Examples of such cylindrical models are the
-interpersonal circle/circumplex
+In the social sciences the use of cylindrical data is very common. Such data
+consist of a linear and a circular component. @gurtman2011reasoning refers to
+such data as vectors, with a directional measure (i.e., the circular component)
+and a measure indicating the magnitude (i.e., the linear component). Many
+established models in psychology are often referred to as circular or circumplex
+models, but those models are cylindrical. Examples of such cylindrical models
+are the interpersonal circle/circumplex
 [@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
@@
 \newline
-\indent Also, many of the more recent types of data that are studied in psychology are
-cylindrical. For example, research on human navigation uses data where distance
-(i.e., the linear component) and direction (i.e., the circular component) are of
-interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
-consist of both the direction (i.e., the circular variable) and the duration
-(i.e., the linear variable) (e.g., @rayner200935th). Apart from the social
-sciences, data with a circular and linear outcome more commonly occur in
-meteorology [@garcia2013exploring], ecology [@garcia2014test] or
-marine research [@lagona2015hidden]
-\newline
+\indent Also, many of the more recent types of data that are studied in
+psychology are cylindrical. For example, research on human navigation uses data
+where distance (i.e., the linear component) and direction (i.e., the circular
+component) are of interest [@chrastil2017rotational] or in eye-tracking, the
+saccade data also consist of both the direction (i.e., the circular variable)
+and the duration (i.e., the linear variable) (e.g., @rayner200935th). Apart from
+the social sciences, data with a circular and linear component more commonly
+occur in meteorology [@garcia2013exploring], ecology [@garcia2014test] or marine
+research [@lagona2015hidden] \newline
 \indent Up until now researchers studying cylindrical data had to rely on linear
-statistical methods to analyze their research results. However, lately more and more
-of these researchers acknowledge that linear methods are not sufficient and call
-for new methods [@gurtman2011reasoning; @pennings2017complexity;
+statistical methods to analyze their research results. However, lately more and
+more of these researchers acknowledge that linear methods are not sufficient and
+call for new methods [@gurtman2011reasoning; @pennings2017complexity;
 @wright2009integrating] that take into account both the circular and the linear
@@
 \newline
-\indent The aim in the present paper is twofold. Firstly, we intend to fill
-the above mentioned gap in the literature by showing that the use of cylindrical
+\indent The aim in the present paper is twofold. Firstly, we intend to fill the
+above mentioned gap in the literature by showing that the use of cylindrical
 models can benefit the analysis of circumplex data and cylindrical data in
@@
 of both the linear and circular component of a cylindrical variable onto a set
-of covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. These adapted cylindrical models are
-then used  to analyse the teacher data.
-
-\vspace{-0.75cm}
-\subsection{Cylindrical Data}
-\vspace{-0.75cm}
-Data that consist of a linear variable and a circular variable is called
-cylindrical data. A circular variable is different from the linear variable in
+of covariates. These adapted cylindrical models are then used  to analyse the
+teacher data.
+
+\section{Modelling Framework}
+Data that consist of a linear variable and a circular variable are called
+cylindrical data. The circular variable is different from the linear variable in
 the sense that it is measured on a different scale. Figure \ref{circline} shows
@@
 @mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
-elaborate overview).\newline
-\indent In the literature, several methods have been put forward to model the
-relation between the linear and circular component of a cylindrical variable.
+elaborate overview).
+
+\hfil [Figure \ref{circline} about here] \hfil
+
+\indent From now on we will refer to the circular random variable as $\Theta$,
+the circular component, and the linear random variable as $Y$, the linear
+component, with realizations $\theta_i$ and $y_i$ for each measurement $i=1,
+\dots, n$ where $n$ is the sample size. The circular component is measured in
+radians and consequently takes values in the interval $[0, 2\pi)$. Note that the
+round brackets mean that $2\pi$ is not included in the interval since this
+represents the same value as 0 as a result of periodicity. Radians can be
+converted to degrees using the relation $1 \:\text{rad} =
+1*\frac{180^\circ}{\pi}$. A circular variable is usually described using its
+first two moments, circular location or mean $\mu$ and mean resultant length
+$\rho$\footnote{In the von Mises distribution, a common distribution for a
+circular outcome, we have a concentration parameter $\kappa$ that is related to
+$\rho$ as $\rho = A_1(\kappa)$, where $A_1{\kappa} = I_1(\kappa)/I_0(\kappa)$
+and $I_0()$ and $I_1()$ are modified Bessel functions of order 0 and 1,
+respectively.}. The sample values of these two parameters are referred to as
+$\bar{\theta}$ and $\hat{\rho}$. The mean resultant length lies between 0,
+meaning the data is not concentrated at all \emph{i.e.} spread over the entire
+circle, and 1, meaning all data is concentrated at a single point on the circle.
+The linear component is measured on the real line and can take values on
+$(-\infty, \infty)$ or we can constrain it to the positive real line $(0,
+\infty)$. Depending on which distributional assumptions are used we describe the
+distribution of the linear component using a mean $\mu$ and variance $\sigma^2$
+or a scale $\nu$ and shape $\alpha$ parameter. When skewed distributions are
+used for either the circular or linear component we may also include a skewness
+parameter $\lambda$.\newline
+\indent In a cylindrical framework we model the location $\mu_c$ of the circular
+variable and the mean $\mu_l$ or scale $\nu$ of the linear variable. Note that
+to prevent confusion we distinguish between the circular location and linear
+mean using subscripts. In this paper we use covariates to predict the circular
+and linear component using the following general type of prediction equations
+for a cylindrical model with $q$ covariates for the linear component and $p$
+covariates for the circular component:
+\begin{equation}\label{circpredgen}
+\hat{\theta_i} = g(\beta_0 + \beta_1z_1 + \dots + \beta_qz_p),
+\end{equation}
+\begin{equation}\label{linpredgen}
+\hat{y_i} = h(\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q),
+\end{equation}
+where $g()$ and $h()$ are link functions, $\boldsymbol{\beta} = (\beta_0, \dots,
+\beta_p)$ is a vector of regression coefficients for the circular variable and
+$\boldsymbol{\gamma} = (\gamma_0, \dots, \gamma_q)$ is a vector of regression
+coefficients for the linear variable. The link functions that are chosen depend
+on the type of distribution we choose for the linear and circular component. In
+this paper we use either an identity link or an exponential function (when
+modelling a scale parameter) for the linear component. For the circular
+component the link functions are specific to circular data and include
+arctangent functions. The specific type of link function and prediction
+equations used will be introduced in the respective descriptions of the models
+used in this paper.\newline
+\indent In addition to modelling the circular and linear component of a
+cylindrical variable separately using covariates we also model the relation
+between them. In the literature, several methods have been put forward to do so.
 Some of these are based on regressing the linear component onto the circular
-component using the following type of relation: \[y = \beta_0 +
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
-linear component and \(\theta\) the circular component [@mardia1978model;
-@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
-different way, e.g. by specifying a multivariate model for several linear and
-circular variables and modelling their covariance matrix
-[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
-example, @abe2017tractable introduce a cylindrical distribution based on a
-Weibull distribution for the linear component and a sine-skewed von Mises
-distribution for the circular component and link these through their respective
-shape and concentration parameters. However, none of the methods that have been
-proposed thus far include additional covariates onto which both the circular and
-linear component are regressed.
+component by including $\sin(\theta_i)$ and $\cos(\theta_i)$ into the prediction
+equation in \eqref{linpredgen} [@mardia1978model;@johnson1978some;
+@mastrantonio2015bayesian]. The linear component is thus modeled using the sine
+and cosine of the circular component. Others model the relation in a different
+way, e.g. by specifying a multivariate model for the linear and circular
+variable and modelling their covariance matrix [@mastrantonio2018joint] or by
+proposing a joint cylindrical distribution. For example, @abe2017tractable
+introduce a cylindrical distribution based on a Weibull distribution for the
+linear component and a sine-skewed von Mises distribution for the circular
+component and link these through their respective shape and concentration
+parameters.
 
@@
 
-\vspace{-0.75cm}
+\subsection{Four Cylindrical Regression Models}\label{Models}
+One of the goals of this paper is to show the benefits of cylindrical methods
+for the analysis of circumplex data and cylindrical data in psychology in
+general. To do so we focus on four cylindrical models. The models were selected
+for their relatively low complexity and the ease with which a regression
+structure could be incorporated. But also because they show different ways of
+modelling the linear and circular component and thereby illustrate a wider range
+of cylindrical models available in the literature. As outlined in the previous
+section, the cylindrical models contain a set of $q$ predictors $\boldsymbol{x}
+= x_1, \dots, x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_p$ for the
+linear and circular components, \(Y\) and \(\Theta\), respectively. The first
+two models are based on a construction by @mastrantonio2015bayesian, while the
+other models are extensions of the models from @abe2017tractable and
+@mastrantonio2018joint. The four cylindrical  models are introduced separately
+in the subsections below. However, to provide a more succinct overview and
+comparison of the four models, Table \ref{TableModels} gives an overview of the
+similarities and differences between the models.
+
+\begin{sidewaystable}[h]
+\centering
+\caption{Comparison of the four cylindrical regression models} 
+\begin{tabular}{lllll}
+  \noalign{\smallskip}\hline\noalign{\smallskip}
+\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
+$\Theta$ & &&&\\
+$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
+$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
+$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
+            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
+$Y$& &&&\\
+$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
+$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
+$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
+            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
+$\Theta$-$Y$ dependence &                                   &                                   & & \\
+                        & $Y$ regressed on                  & $Y$ regressed on                  & circular concentration $\kappa$  & covariance \\
+                        & $\sin(\Theta)$ and $\cos(\Theta)$ & $\sin(\Theta)$ and $\cos(\Theta)$ & and linear scale $\alpha$ & matrix\\\hline
+\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
+\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
+
+\end{tabular}
+\label{TableModels}
+\end{sidewaystable}
+
+\hfil [Table \ref{TableModels} about here] \hfil
+
+\subsubsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+Following @mastrantonio2015bayesian we consider two models where the prediction
+equation for the linear component is specified by combining \eqref{linpredgen},
+with an identity link function and including $\sin(\theta_i)$ and
+$\cos(\theta_i)$ as follows:
+\begin{equation}\label{linpredCLPNCLGPN}
+\hat{y_i} = \gamma_0 + \gamma_{cos}*\cos(\theta_i)*r_i + \gamma_{sin}*\sin(\theta_i)*r_i + \gamma_1*x_1 + \dots + \gamma_q*x_q,
+\end{equation}
+\noindent where $r_i$ is a realization of the unobserved random variable
+\(R\geq0\) that will be introduced below, \(\gamma_0, \gamma_{cos},
+\gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the intercept and regression
+coefficients and \(x_1, \dots, x_q\) are the \(q\) covariates for the prediction
+of the linear component. We assume a normal $N(\mu_l, \sigma^2)$ distribution
+for the linear component.\newline
+\indent For the circular component we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise from a
+projection of a distribution defined in bivariate space onto the circle. Figure
+\ref{projection} represents this projection. In the left plot of Figure
+\ref{projection} we see realizations $(s_i^{I}, s_i^{II})$ from the bivariate
+normal variable \(\boldsymbol{S}\) that in the middle plot are projected to form
+the circular component \(\Theta\) in the right plot. Mathematically the relation
+between \(\boldsymbol{S}\) and \(\Theta\) is defined as follows:
+\begin{equation}\label{projection_eq}
+\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
+\end{equation}
+\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
+\(\boldsymbol{S}\), that is represented by the lines connecting the bivariate
+datapoints to the origin in the middle plot. We call $\boldsymbol{S}$ the
+augmented representation of the circular component. It is a variable that in
+contrast to \(\Theta\) is not observed and thus considered latent or auxiliary.
+This then means that we do not model \(\Theta\) directly but indirectly through
+$\boldsymbol{S}$.
+
+\hfil [Figure \ref{projection} about here] \hfil
+
+\indent For both the PN and GPN distributions the circular location parameter
+\(\mu_c\in [0, 2\pi)\) is modeled.\footnote{Note that for the CL-GPN
+model the circular location parameter also depends on the variance-covariance
+matrix and the circular predicted values should be computed using numerical
+integration or Monte Carlo methods because a closed form expression for the mean
+direction is not available.} The prediction equation for the circular component is specified by using a double arctangent link function in \eqref{circpredgen} as follows:
+\begin{equation}\label{circpredCLPNCLGPN}
+\hat{\theta_i} = \text{atan2}(\boldsymbol{\beta^{II}}\boldsymbol{z}_i, \boldsymbol{\beta^{I}}\boldsymbol{z}_i)
+\end{equation}
+\noindent where $\boldsymbol{\beta}^{I} = (\beta_0^{I},
+\beta_1^{I}, \dots, \beta_p^{I})$ and
+$\beta^{II} = (\beta_0^{II},
+\beta_1^{II}, \dots, \beta_p^{II})$ are vectors with
+intercepts and regression coefficients for the prediction of $S^{I}$ and
+$S^{II}$ and $\boldsymbol{z}_i$ is a vector with predictor values for each
+individual $i \in 1, \dots, n$ where $n$ is the sample size. Note that as a
+result of the augmented representation of the circular component we have two sets
+of regression coefficients and intercepts, in contrast to a single set in
+\eqref{circpredgen}. This leads to problems when we want to interpret the effect
+of a covariate on the circle. A circular regression line for simulated data is shown in Figure
+\ref{circregline}, with covariate values on the x-axis and the predicted
+circular component on the y-axis. As can be seen it is of a non-linear character
+meaning that the effect of a covariate is different at different values of the
+covariate. A circular regression line is usually described by the slope at the
+inflection point, the point at which the slope of the regression line starts
+flattening off (indicated with a square in Figure \ref{circregline}). By
+default, the parameters from the PN and GPN models do not directly describe this
+inflection point. For the PN distribution however, @CremersMulderKlugkist2017
+solved this interpretation problem. They introduce a new parameter $b_c$ that describes the slope at
+the inflection point of the regression line. For the GPN distribution the
+interpretation problem however remains.
+
+\hfil [Figure \ref{circregline} about here] \hfil
+
+\indent The main difference between the PN and GPN distribution lies in the
+definition of their covariance matrix. For the PN distribution this is an
+identity matrix, causing the distribution to be unimodal and symmetric, whereas
+for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2
+& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\),
+allowing for multimodality and asymmetry/skewness as illustrated in Figure \ref{plotGPN}.
+
+\hfil [Figure \ref{plotGPN} about here] \hfil
+
+\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
+Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
+@hernandez2016general for the regression of the circular component. A detailed
+description of the Bayesian estimation and MCMC samplers can be found in the
+Supplementary Material.
+
+```{r plotprojecting, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+library(MASS)
+library(plotrix)
+
+set.seed(10200)
+
+Y <- mvrnorm(10, c(2,2), (diag(2)))
+
+theta <- atan2(Y[,2], Y[,1])
+
+tikz("Plots/plotprojecting.tex", standAlone =TRUE, height = 3, width = 9, pointsize = 12, engine = "pdftex")
+
+par(mfrow = c(1,3), oma =c(0,0,0,0), mar = c(0,0,0,0))
+
+plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+points(x = cos(theta), y = sin(theta))
+segments(x0 = cos(theta), x1 = Y[,1], y0 = sin(theta), y1 = Y[,2])
+segments(x0 = cos(theta), x1 = 0, y0 = sin(theta), y1 = 0, lty = 2)
+
+plot(x = cos(theta), y = sin(theta), ylim = c(-2, 4), xlim = c(-2, 4),
+     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
+
+points(0,0, pch = 3)
+draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
+
+dev.off()
+
+tools::texi2dvi("Plots/plotprojecting.tex",pdf=TRUE)
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
+\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
+The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norms of the bivariate datapoints, realizations of the random variable $R$.}
+\label{projection}
+\end{figure}
+
+```{r circregline, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+
+tikz("Plots/circregline.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
+
+set.seed(101)
+
+Xreal <- rnorm(100, -2, 1.5)
+
+CompI  <- 1.1 + 2.5*Xreal + rnorm(100, 0.1,1)
+CompII <- 2.1 + -0.4*Xreal + rnorm(100, 0.1,1)
+
+Outreal <- atan2(CompII, CompI)
+
+X <- seq(-5, 5, 0.1)
+
+predCompI  <- 1.1 + 2.5*X
+predCompII <- 2.1 + -0.4*X
+
+PredCirc <- atan2(predCompII, predCompI)
+
+axval <- ((1.1*2.5)+(2.1*-0.4))/(1.1^2 + 2.1^2)
+acval <- atan2(2.1 + -0.4*axval, 1.1 + 2.5*axval)
+
+plot(X, PredCirc*(180/pi), ylim = c(-200,200), type = "l", xlab = "Predictor",
+     ylab = "Circular Outcome", bty = 'n', yaxt = "n", xaxt = "n")
+axis(1, xaxp = c(-5, 0, 5), at = c(seq(-5, 5, by = 1)),
+     labels = c(seq(-5, 5, by = 1)), las = 2)
+axis(2, xaxp = c(-180, 0,180), at = c(seq(-180, 180, by = 60)),
+     labels = c(seq(-180, 180, by = 60)), las = 2)
+points(Xreal, Outreal*(180/pi), cex=0.5)
+points(axval,
+       acval*(180/pi),
+       pch = 0)
+
+dev.off()
+tools::texi2dvi("Plots/circregline.tex",pdf=TRUE)
+
+```
+
+\begin{figure}[]
+  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
+  \caption{Circular regression line for the relation between a covariate and a circular component with the simulated data the regression line was fit to.
+The square indicates the inflection point of the regression line.}
+  \label{circregline}
+\end{figure}
+
+```{r plotGPN, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
+
+require(MASS)
+require(circular)
+
+y1 = mvrnorm(1000, c(2,2), matrix(c(1^2 + 0^2, 0, 0, 1), nrow = 2, ncol = 2))
+y2 = mvrnorm(1000, c(2,2), matrix(c(5^2 + 0.9^2, 0.9, 0.9, 1), nrow = 2, ncol = 2))
+y3 = mvrnorm(1000, c(2,2), matrix(c(10^2 + -0.5^2, -0.5, -0.5, 1), nrow = 2, ncol = 2))
+
+theta_1 = atan2(y1[,2], y1[,1])
+theta_2 = atan2(y2[,2], y2[,1])
+theta_3 = atan2(y3[,2], y3[,1])
+
+tikz("Plots/plotGPN.tex", standAlone =TRUE, height = 3, width = 9, pointsize = 12, engine = "pdftex")
+
+par(mfrow=c(1,3),
+    oma=c(0,0,0,0),
+    mar=c(0,0,0,0))
+
+plot(as.circular(theta_1), stack = TRUE,
+     main = " ", ylab = " ", xlab = " ", shrink = 1.3, axes = FALSE)
+axis.circular(at = 0, labels = "0")
+plot(as.circular(theta_2), stack = TRUE,
+     main = " ", ylab = " ", xlab = " ", shrink = 1.3, axes = FALSE)
+axis.circular(at = 0, labels = "0")
+plot(as.circular(theta_3), stack = TRUE,
+     main = " ", ylab = " ", xlab = " ", shrink = 1.3, axes = FALSE)
+axis.circular(at = 0, labels = "0")
+
+#Close the device
+dev.off()
+
+# Compile the tex file
+tools::texi2dvi("Plots/plotGPN.tex",pdf=TRUE)
+
+```
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotGPN.pdf}
+\caption{Plot showing 1000 samples from three GPN distributions with mean vector $(2,2)$ and $\tau = 1,5,10$ and $\xi = 0, 0.9, -0.5$ from left to right}
+\label{plotGPN}
+\end{figure}
+
+\subsubsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+This model is an extension of the cylindrical model introduced in
+@abe2017tractable to the regression context. The circular dependence between the
+  linear and circular component, $Y$ and $\Theta$, is defined through a joint
+  density function as follows:
+
+\begin{equation}\label{WeiSSVMdensity}
+f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
+                 (1 +\lambda\sin(\theta - \mu_c))
+                 y^{\alpha-1}
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu_c))],
+\end{equation}
+
+\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
+\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters
+respectively. The Abe-Ley density thus concerns a combination of a Weibull
+distribution, with scale parameter \(\nu>0\) and shape parameter \(\alpha\), for
+the linear component and a sine-skewed von Mises distribution, with location
+parameter \(\mu_c\in [0, 2\pi)\), concentration parameter \(\kappa>0\) and
+skewness \(\lambda \in [-1,1]\), for the circular component. In contrast to the
+CL-PN and CL-GPN models, the linear component \(Y\) is in this model defined
+only on the positive real half-line \([0, + \infty)\) and thus can not be
+negative. Our modification of the Abe-Ley model occurs at the level of the
+linear scale parameter \(\nu>0\) and circular location parameter \(\mu_c\in [0,
+2\pi)\), both of which we will model using covariates.\newline
+\indent The prediction equation for the circular component equals:
+\begin{equation}\label{circpredcAL}
+\hat{\theta}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{\beta}\boldsymbol{z}_i)
+\end{equation}
+Note that we have taken out the intercept $\beta_0$ from the link function $g() =
+2\tan^{-1}$ in \eqref{circpredgen}. The vector $\boldsymbol{\beta}$ thus does
+not contain an intercept. We do not directly predict the linear component. The
+conditional distribution for the linear component is Weibull, meaning that we can
+use methods from survival analsis to interpret the effect of a predictor. In
+survival analysis a 'survival' function is used in which time is plotted against
+the probability of survival of subjects suffering from a specific medical
+condition. This probability is computed using the 'survival-function' defined as
+\begin{equation}\label{survivalfunc}
+\exp(-\alpha
+y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i))^{1/\alpha}}),
+\end{equation}
+\noindent with \(\hat{\nu}_i = \exp(\boldsymbol{\gamma}\boldsymbol{x}_i)\). Note that $\boldsymbol{x}_i$ also includes a 1 to be able to estimate the intercept $\gamma_0$. From
+the survival function we also see that the circular concentration parameter
+\(\kappa\) and the linear shape parameter \(\alpha\) regulate the
+circular-linear dependence in the Abe-Ley model.
+In the exponent or power of the linear outcome, the prediction error on the circle, $\cos(\theta_i - \hat{\theta_i})$, is multiplied by $\tanh{\kappa}$, and raised to the power $1-\alpha$.\newline
+\indent We can use numerical optimization (Nelder-Mead) to find solutions for
+the maximum likelihood (ML) estimates for the parameters of the model.
+
+\subsubsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
+This model is an extension of the cylindrical model introduced by
+@mastrantonio2018joint to the regression context. Although the model may contain
+several circular and linear components we will restrict ourselves to one
+circular and one linear component in this paper. The circular component is
+modelled by a GPN distribution, as in the CL-GPN model, while the linear component is modelled by a skew
+normal distribution [@sahu2003new]. Because the GPN distribution is modelled
+using a so-called augmented representation (see \eqref{projection_eq}) it is
+convenient to use a similar tactic for modelling the skew normal distribution.
+As in @mastrantonio2018joint, dependence between the linear and circular
+component is created by modelling the augmented representation of \(\Theta\) and
+\(Y\) together in a 3 dimensional normal distribution. The joint density of the model is then represented by:
+\begin{equation}\label{YDThetarjoint} 
+f(\theta, r, y, d) = 2\phi_{3}((\boldsymbol{s}^t,
+y^t)^t \mid \boldsymbol{M} + (0,0,\lambda d)^t, \boldsymbol{\Sigma})
+\phi(d \mid 0, 1)r, 
+\end{equation}
+\noindent where $\phi$ represents the normal probability density function, \(\boldsymbol{s} = (r(\cos(\theta), \sin(\theta)))^t\), $d$ is
+the augmented representation of the linear component, $\boldsymbol{M} =
+\boldsymbol{B}^t\boldsymbol{X}$, $\boldsymbol{B}_{(g + 1) \times 3}$ is
+defined below, $\boldsymbol{X}$ is a design matrix and $\boldsymbol{\Sigma} =
+\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
+\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )$ a
+variance-covariance matrix. The matrices \(\boldsymbol{\Sigma}_s\) and \(\boldsymbol{\Sigma}_y\) are the
+covariance matrices for the variances of and covariances between the augmented
+representations of the circular and linear component respectively.
+Note that because we only have one circular and one linear component in this paper, \(\boldsymbol{\Sigma}_s\) is a two by two matrix and \(\boldsymbol{\Sigma}_y\) is a scalar representing the variance of the augmented linear component.
+The matrix
+\(\boldsymbol{\Sigma}_{sy}\) contains covariances between the augmented
+representations of the circular and linear component. The matrix with
+regression coefficients and intercepts, $\boldsymbol{B}$ is defined as follows:
+\begin{equation}\label{regmatGPNSSN}
+\boldsymbol{B} = \begin{bmatrix}
+\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\
+\beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}}\\
+\vdots & \vdots & \vdots \\
+\beta_{g_{s^{I}}} & \beta_{g_{s^{II}}} & \beta_{g_{y}} \end{bmatrix}.
+\end{equation} 
+\noindent Note that in contrast to the previous three models where we use $p$
+and $q$ covariates for the circular and linear component respectively, we use $g
+= p = q$ covariates for both components in the GPN-SSN model. Unlike the other
+models, where $\mu_l$ and $\mu_c$ are modelled separately, the GPN-SSN thus has
+a shared mean vector $\boldsymbol{M}$ and variance-covariance matrix
+$\boldsymbol{\Sigma}$ for the linear and circular component, much like having
+multiple outcomes in a MANOVA (multivariate analysis of variance) model. This
+implies that the covariances in the variance-covariance matrix describe the
+dependence between the circular and linear component.\newline
+\indent The prediction equation for the circular component\footnote{Note that for the  GPN-SSN model the predicted circular component also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} is similar to \eqref{circpredCLPNCLGPN} and equal to:
+\begin{equation}\label{circpredGPNSSN}
+\hat{\theta}_i = \text{atan2}(\boldsymbol{\beta^{II}}\boldsymbol{z}_i, \boldsymbol{\beta^{I}}\boldsymbol{z}_i),
+\end{equation}
+\noindent where $\boldsymbol{\beta^I} = (\beta_{0_{s^{I}}}, \beta_{1_{s^{I}}}, \dots, \beta_{g_{s^{I}}})$ and $\boldsymbol{\beta^{II}} = (\beta_{0_{s^{II}}}, \beta_{1_{s^{II}}}, \dots, \beta_{g_{s^{II}}})$.
+For the linear component we subsititute $h()$ in \eqref{linpredgen} for an identity link and $\boldsymbol{\gamma}$ for $\boldsymbol{\beta}$:
+\begin{equation}\label{linpredGPNSSN}
+\hat{y}_i = \boldsymbol{\beta}\boldsymbol{x}_i,
+\end{equation}
+\noindent where $\boldsymbol{x}_i = \boldsymbol{z}_i$ and $\boldsymbol{\beta} = (\beta_{0_{y}}, \beta_{1_{y}}, \dots, \beta_{g_{y}})$.\newline
+\indent We estimate the model using MCMC methods. A detailed description of these
+methods is given in the Supplementary Material.
+
+\subsection{Model Fit Criterion}\label{Modelfit}
+For the four cylindrical models we focus on their out-of-sample predictive
+performance to determine the fit of the model. To prevent possible problems
+concerning overestimation, we use k-fold cross-validation and split our data
+into 10 folds. Each of these folds (10 \(\%\) of the sample) is used once as a
+holdout set and 9 times as part of a training set. The analysis will thus be
+performed 10 times, each time on a different training set.\newline
+\indent A proper criterion to compare out-of-sample predictive performance is
+the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
+value of this criterion, the better the predictive performance of the model.
+Because the joint density and thus also the likelihood for the modified GPN-SSN
+model is not available in closed form [@mastrantonio2018joint] we compute the
+PLSL for the circular and linear component separately for all models. Using ML
+estimates this criterion can be computed as follows for the circular and linear
+component:
+\begin{equation}\label{PLSLMLcirc}
+PLSL_c = -2 \sum_{i = 1}^{M}\log l(\theta_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
+\end{equation}
+\begin{equation}\label{PLSLMLlin}
+PLSL_l = -2 \sum_{i = 1}^{M}\log l(y_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
+\end{equation}
+\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
+holdout set, \(y_i\) and $\theta_i$ are the \(i^{th}\) datapoints from the
+holdout set and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the
+model parameters. Using posterior samples the criterion is similar to the log
+pointwise predictive density (lppd) [@BDA, p. 169] and can be computed for the
+circular and linear component as:
+\begin{equation}\label{PLSLBayescirc}
+PLSL_c = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(\theta_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
+\end{equation}
+\begin{equation}\label{PLSLBayeslin}
+PLSL_l = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(y_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
+\end{equation}
+\noindent where \(B\) is the amount of posterior samples and
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
+parameters for the \(j^{th}\) iteration. Note that although we fit the CL-PN,
+CL-GPN and GPN-SSN models using Bayesian statistics, we do not take prior
+information into account when assessing model fit with the PLSL. According to
+@BDA this is not necessary since we are assessing the fit of a model to data,
+the holdout set, only. They argue that the prior in such case is only of
+interest for estimating the parameters of the model but not for determining the
+predictive accuracy.\newline
+\indent For each of the four cylindrical models and for each of the 10
+cross-validation analyses we can then compute a PLSL for the circular and linear
+component by using the conditional log-likelihoods of the respective component (see
+Supplementary Material for a definition of the loglikelihoods). To evaluate the
+predictive performance we average across the PLSL criteria of the
+cross-validation analyses. We also assess the cross-validation variability by
+means of the standard deviations of the PLSL criteria.
+
 \section{Teacher Data}\label{Example}
-\vspace{-0.75cm}
 \begin{figure}
@@
 also implies that scores on the IPC could be viewed as a circular variable. This
-circular variable represents the type of interpersonal behavior that a teacher
-shows towards his/her students.\newline
+circular variable represents location on the IPC of the interpersonal behavior
+that a teacher shows towards his/her students.
+
+\hfil [Figure \ref{QTI} about here] \hfil
+
 \indent @Cremers2018Assessing explain the circular nature of the IPC data and
@@
 scores, Agency and Communion, can be converted to a circular score using the
-two-argument arctangent function in \eqref{PredVal}, where \(A\) represents a score on the Agency dimension and
-\(C\) represents a score on the Communion dimension \footnote{The selection of
-the origin in circumplex data depends on the scaling of the Agency and Communion
-scores. Agency and Communion are measured on a scale from 1 to 5 and for
-analysis purposes they are later converted to scale ranging from -1 to 1.  Their
-respective 0 scores form the origin. Scaling is however only considered an issue
-in those instances where cylindrical data is derived from measurements in
-bivariate space.}. Note that when placing a
-unit circle on Figure \ref{QTI} we see that the Agency dimension is related to
-the sine of the circular score and the Communion dimension is related to the
-cosine of the circular score.
+two-argument arctangent function in \eqref{PredVal}, where \(A\) represents a
+score on the Agency dimension and \(C\) represents a score on the Communion
+dimension\footnote{The selection of the origin in circumplex data depends on
+the scaling of the Agency and Communion scores. Agency and Communion are
+measured on a scale from 1 to 5 and for analysis purposes they are later
+converted to a scale ranging from -1 to 1.  Their respective 0 scores form the
+origin.
+Although the scaling influences the average score and spread on the IPC and the intensity, the difference between the individual measurements will be retained (albeit them being different in size).
+With regards to the intensity, scaling has the same effect as in a standard linear regression.
+With regards to the location on the IPC (the circular component) rescaling to -1,1 affects the size of the circular (as well as bivariate linear) regression coefficients.
+This has a positive effect if the original Agency and Communion scores are far from the origin in bivariate space.
+In that case the locations on the IPC are very concentrated before scaling (small variance).
+Such data are hard to estimate using circular models as the circular regression coefficients may be very small (high risk of bias, see Cremers, Mainhard, Klugkist (2018a)).
+Scaling will thus improve estimation.
+Finally note that scaling is only considered an issue in those instances where cylindrical data is derived from measurements in bivariate space.}.
+Note that
+when placing a unit circle on Figure \ref{QTI} we see that the Agency dimension
+is related to the sine of the circular score and the Communion dimension is
+related to the cosine of the circular score.
 \begin{equation}\label{PredVal}
@@
 The resulting circular variable \(\theta\) can then be modelled and takes values
-in the interval $[0, 2\pi)$ or $[0^\circ, 360^\circ)$. Note that the round
-brackets mean that $2\pi$ and $360^\circ$ are not included in the interval since
-these represent the same value as 0 as a result of periodicity. \newline
+in $[0, 2\pi)$.\newline
 \indent A circular analysis of circumplex data has several benefits: it is more
@@
 This provides us with new insights compared to a separate analysis of the two
-dimensions that is standard in the literature (see e.g.
+dimensions that is standard in the literature (see, e.g.,
 @pennings2018interpersonal, @wright2009integrating or
@@
 Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
-strength of the interpersonal behavior a teacher shows towards his/her students.
-In a cylindrical model this strength (the linear outcome) can be modeled
-together with the type of interpersonal behavior of a teacher (the circular
-outcome). This leads to an improved analysis of interpersonal circumplex data,
-over either analyzing the two dimensions separately or using a circular model,
-because we take all information, circular and linear, into account. In the next
-section we introduce several cylindrical models that can be used to analyze the
-teacher data. First however we will provide descriptives for the teacher data.
-
-\vspace{-0.75cm}
+intensity of the interpersonal behavior a teacher shows towards his/her
+students. In a cylindrical model this intensity (the linear component) can be
+modeled together with the location of interpersonal behavior of a teacher on the
+IPC (the circular component). This leads to an improved analysis of interpersonal
+circumplex data, over either analyzing the two dimensions separately or using a
+circular model, because we take all information, circular and linear, into
+account. In the next section we introduce several cylindrical models that can be
+used to analyze the teacher data. First however we will provide descriptives for
+the teacher data and conduct a 'standard' analysis for these data that we can compare the results from the cylindrical models to.
+
 \subsection{Data Description}\label{DataDescriptives}
-\vspace{-0.75cm}
 ```{r loaddata, return = FALSE, echo = FALSE, results = FALSE}
@@
 plot(SE_low$y, SE_low$theta*(180/pi), 
-     ylab = "IPC", xlab = "IPC strength",
+     ylab = "IPC", xlab = "IPC intensity",
      bty = "n", ylim = c(-180, 180))
@@
 for multiple classes were available. This results in a sample of 151 teachers.
-In addition to the type of interpersonal behavior (IPC), the circular outcome,
-and the strength of interpersonal behavior (IPC strength), the linear outcome, a
-teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
-covariate in the analysis. In previous research, in psychology and education it
-has been shown that higher self-efficacy is related to the quality of
-interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
+The data includes the location of interpersonal behavior on the IPC (IPC), the
+circular component, and the intensity of interpersonal behavior (IPC intensity),
+the linear component. It also includes teachers' self-efficacy (\verb|SE|)
+concerning classroom management as covariate that is used to model the IPC and
+IPC intensity. This means that $\boldsymbol{x}_i = \boldsymbol{z}_i = (1,\:
+\text{SE}_i)$ in all cylindrical models except for the Abe-Ley model where
+$\boldsymbol{z}_i = \text{SE}_i$. In previous research, in psychology and
+education it has been shown that higher self-efficacy is related to the quality
+of interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
 After listwise deletion of missings ($3$ in total, only for the self-efficacy)
-we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
-descriptives for the dataset. For the circular variable IPC we show sample
-estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.
-The circular concentration lies between 0, meaning the data is not concentrated
-at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
-concentrated at a single point on the circle. The population values of these
-parameters are usually, and also in this paper, referred to as $\mu$ (circular
-location) and $\kappa$ (circular concentration). For the linear variables
-(strength IPC and SE) we show sample estimates of the linear mean and standard
-deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
-the linear and circular outcome of the teacher data for teachers with low SE
+we have a sample of 148 teachers. Note that we remove the missings here for
+simplicity, however, in general and especially when the number of missings is
+larger, the influence of listwise deletion should be investigated and necessary
+precaution (e.g. imputation of missing values) should be taken. Table
+\ref{Tableteacherdescriptives} shows descriptives for the dataset. Note that we
+also show the scores on the Agency and Communion component of the IPC before
+transformation to a circular score. For the circular variable IPC we show sample
+estimates for the circular mean $\bar{\theta}$ and mean resultant length
+$\hat{\rho}$. For the linear variables (Agency, Communion, IPC intensity and the
+covariate SE) we show sample estimates of the linear mean and standard deviation
+(sd). Figure \ref{dataplot} is a scatterplot showing the relation between the
+linear and circular component of the teacher data for teachers with low SE
 (below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
@@
 Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
+Agency & 0.19 & 0.16 & -0.33 - 0.49 & Linear\\
+Communion & 0.30 & 0.24 & -0.58 - 0.77 & Linear\\
 IPC &33.22$^\circ$& 0.76 & - & Circular\\
-strength IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
+intensity IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
 SE & 5.04 & 1.00 & 1.5 - 7.0 & Linear\\
@@
 \multicolumn{5}{l}{Note: For the circular variable IPC we show sample }\\
-\multicolumn{5}{l}{estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.}\\
+\multicolumn{5}{l}{estimates for the circular mean $\bar{\theta}$ and mean resultant length $\hat{\rho}$.}\\
 \multicolumn{5}{l}{For the linear variable we show the sample mean,}\\
@@
 
-\vspace{-0.75cm}
-\section{Four Cylindrical Regression Models}\label{Models}
-\vspace{-0.75cm}
-One of the goals of this paper is to show the benefits of cylindrical methods
-for the analysis of circumplex data and cylindrical data in psychology in
-general. To do so we decide to focus on four cylindrical models. The models were
-selected for their relatively low complexity and the ease with which a
-regression structure could be incorporated. But also because they show different
-ways of modelling the linear and circular outcome and thereby illustrate a
-wider range of cylindrical models available in the literature. The
-cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
-x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_p$ for the linear and
-circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
-based on a construction by @mastrantonio2015bayesian, while the other models are
-extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
-four cylindrical  models are introduced separately in the subsections below.
-However, to provide a more succinct overview and comparison of the four models, Table
-\ref{TableModels} gives an overview of the similarities and differences between
-the models.
-
-\begin{sidewaystable}[h]
+\hfil [Table \ref{Tableteacherdescriptives} about here] \hfil
+
+\hfil [Figure \ref{dataplot} about here] \hfil
+
+\subsection{Standard Analysis}\label{StandardAnalysis}
+
+To perform a standard analysis of the teacher data we fit two linear regression
+models to the Agency and Communion scores of the teachers separately. In these
+regression models we incorporate an intercept and the covariate self-efficacy.
+Coefficients and standard errors from the models are shown in Table
+\ref{Tableteacherstandardanalysis}. From the results we conclude that the effect
+of self-efficacy is significant for both the Agency and Communion component. An
+increase of 1 unit on self-efficacy leads to a 0.07 increase in Agency and a
+0.09 increase in Communion. Note that in this setup we can only model the effect
+of self-efficacy on the intensity (size of the score) on the Agency and
+Communion component. We can not quantify the effect on the location and
+intensity on the IPC nor can we quantify their dependence.
+
+\hfil [Table \ref{Tableteacherstandardanalysis} about here] \hfil
+
+```{r standardanalysis, cache = TRUE, echo = FALSE, results = FALSE}
+A_res <- lm(lAgency ~ SEc, data = Dat)
+C_res <- lm(lcommunion ~ SEc, data = Dat)
+
+summary(A_res)
+summary(C_res)
+
+summary(Dat$lAgency)
+summary(Dat$lcommunion)
+sd(Dat$lAgency)
+sd(Dat$lcommunion)
+```
+
+\begin{table}[h]
 \centering
-\caption{Comparison of the four cylindrical regression models} 
-\begin{tabular}{lllll}
+\caption{Regression coefficients and standard errors for the standard analysis of the teacher dataset.} 
+\begin{tabular}{lrr}
   \noalign{\smallskip}\hline\noalign{\smallskip}
-\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
-$\Theta$ & &&&\\
-$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
-$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
-$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
-            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
-$Y$& &&&\\
-$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
-$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
-$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
-            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
-$\Theta$-$Y$ dependence &                                   &                                   & & \\
-                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
-                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
-\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
-\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
-
+  & Agency & Communion \\\hline\noalign{\smallskip}
+Intercept & 0.19 (0.01) & 0.30 (0.02) \\
+SE & 0.07 (0.01) & 0.09 (0.02) \\ \hline
+\multicolumn{3}{l}{Note: The intercepts indicate the average }\\
+\multicolumn{3}{l}{Agency and Communion. The coefficient for }\\
+\multicolumn{3}{l}{SE indicates the effect of self-efficacy }\\
+\multicolumn{3}{l}{on Agency and Communion.}
 \end{tabular}
-\label{TableModels}
-\end{sidewaystable}
-
-\vspace{-0.75cm}
-\subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
-\vspace{-0.75cm}
-Following @mastrantonio2015bayesian we consider two models where
-the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
-and $q$ covariates is specified as
-\begin{equation}\label{circlinlink}
-Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
-\end{equation}
-\noindent where the random variable \(R\geq0\) will be introduced below, the
-error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
-\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
-intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
-covariates for the prediction of the linear outcome. We thus assume a normal
-distribution for the linear outcome.\newline
-\indent For the circular outcome we assume either a projected normal (PN) or a
-general projected normal (GPN) distribution. These distributions arise from a
-projection of a distribution defined in bivariate space onto the circle. Figure
-\ref{projection} represents this projection. In the left plot of Figure
-\ref{projection} we see datapoints from the bivariate variable
-\(\boldsymbol{S}\) that in the middle plot are projected to form the circular
-outcome \(\Theta\) in the right plot. Mathematically the relation between
-\(\boldsymbol{S}\) and \(\Theta\) is defined as follows
-\begin{equation}\label{projection}
-\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
-\end{equation}
-\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
-\(\boldsymbol{S}\); the lines connecting the bivariate datapoints to the origin
-in the middle plot. We call $\boldsymbol{S}$ the augmented representation of the
-circular outcome. It is a variable that in contrast to \(\Theta\) is not
-observed and thus considered latent or auxiliary. This then means that we do not
-model \(\Theta\) directly but indirectly through $\boldsymbol{S}$. \newline
-\indent For both the PN and GPN distributions the circular location parameter
-\(\mu\in [0, 2\pi)\) is modeled as \(\hat{\mu}_i =
-\mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) =
-\mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i,
-\boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
-model the circular location parameter also depends on the variance-covariance
-matrix and the circular predicted values should be computed using numerical
-integration or Monte Carlo methods because a closed form expression for the mean
-direction is not available.} where $\boldsymbol{\beta}^{I}$ and
-$\boldsymbol{\beta}^{II}$ are vectors with intercepts and regression
-coefficients for the prediction of $\boldsymbol{S}$ and $\boldsymbol{z}_i$ is a
-vector with predictor values for each individual $i \in 1, \dots, n$ where $n$
-is the sample size. Note that as a result of the augmented representation of the
-circular outcome we have two sets of regression coefficients and intercepts, one
-for each bivariate component of $\boldsymbol{S}$. This leads to problems when we
-want to interpret the effect of a covariate on the circle. A circular regression
-line is shown in Figure \ref{circregline}, with covariate values on the x-axis
-and the predicted circular outcome on the y-axis. As can be seen it is of a
-non-linear character meaning that the effect of a covariate is different at
-different values of the covariate. A circular regression line is usually
-described by the slope at the inflection point, the point at which the slope of
-the regression line starts flattening off (indicated with a square in Figure
-\ref{circregline}). By default, the parameters from the PN and GPN models do not
-directly describe this inflection point. For the PN distribution however,
-@CremersMulderKlugkist2017 solved this interpretation problem and introduce new
-  circular regression coefficients. They introduce a new parameter $b_c$ that
-  describes the slope at the inflection point of the regression line. For the
-  GPN distrbution the interpretation problem however remains.\newline
-\indent The main difference between the PN and GPN distribution lies in the
-definition of their covariance matrix. For the PN distribution this is an
-identity matrix, causing the distribution to be unimodal and symmetric, whereas
-for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2
-& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
-asymmetry/skewness.\newline
-\indent For the teacher dataset the predicted linear outcome, strengths of
-interpersonal behavior, in the CL-PN and CL-GPN model is the following:
-\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
-\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted circular
-outcome, type of interpersonal behavior, equals: \[\hat{\theta}_i = \mu_i =
-\mbox{atan2} (\beta_0^{II} + \beta_1^{II}\text{SE}_i, \beta_0^{I} +
-\beta_1^{I}\text{SE}_i).\] \noindent where $\text{SE}_i$ is the self-efficacy
-score of one individual. The CL-PN and CL-GPN models thus allow us to assess the
-average type and strength of interpersonal behavior through the parameters
-$\beta_{0}^{I}$, $\beta_{0}^{II}$ and $\gamma_0$ as well as the effect of
-self-efficacy on type and strength of teacher behavior through the parameters,
-$\beta_{1}^{I}$, $\beta_{1}^{II}$ and $\gamma_1$. In addition, because the type
-and strength of interpersonal behavior are modelled together via the regression
-in \eqref{circlinlink} we can assess the effect of the type of interpersonal
-behavior on the strength through the parameters $\gamma_{sin}$ and
-$\gamma_{cos}$. In the teacher data these are the regression coefficients for
-the effect of the sine and cosine of the type of behavior which are related to
-the scores on the Agency and Communion dimensions respectively.\newline
-\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
-Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome. A detailed
-  description of the Bayesian estimation and MCMC samplers can be found in the
-  Supplementary Material.
-
-```{r plotprojecting, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-library(MASS)
-library(plotrix)
-
-set.seed(10200)
-
-Y <- mvrnorm(10, c(2,2), (diag(2)))
-
-theta <- atan2(Y[,2], Y[,1])
-
-tikz("Plots/plotprojecting.tex", standAlone =TRUE, height = 3, width = 9, pointsize = 12, engine = "pdftex")
-
-par(mfrow = c(1,3), oma =c(0,0,0,0), mar = c(0,0,0,0))
-
-plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
-     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
-
-points(0,0, pch = 3)
-draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
-
-plot(x = Y[,1], y = Y[,2], ylim = c(-2, 4), xlim = c(-2, 4),
-     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
-
-points(0,0, pch = 3)
-draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
-
-points(x = cos(theta), y = sin(theta))
-segments(x0 = cos(theta), x1 = Y[,1], y0 = sin(theta), y1 = Y[,2])
-segments(x0 = cos(theta), x1 = 0, y0 = sin(theta), y1 = 0, lty = 2)
-
-plot(x = cos(theta), y = sin(theta), ylim = c(-2, 4), xlim = c(-2, 4),
-     ylab = "", xlab ="", bty=  "n", xaxt = "n", yaxt ="n", asp = 1)
-
-points(0,0, pch = 3)
-draw.circle(0, 0, 1, nv = 100, border = NULL, col = NA, lty = 1, lwd = 2)
-
-dev.off()
-
-tools::texi2dvi("Plots/plotprojecting.tex",pdf=TRUE)
-
-```
-
-\begin{figure}
-\centering
-\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
-\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
-The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norm of the bivariate datapoints, the random variable $R$.}
-\label{projection}
-\end{figure}
-
-```{r circregline, cache = TRUE, warning = FALSE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
-
-tikz("Plots/circregline.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")
-
-set.seed(101)
-
-Xreal <- rnorm(100, -2, 1.5)
-
-CompI  <- 1 + 0.3*Xreal + rnorm(100, 0.1,1)
-CompII <- 2 + -0.4*Xreal + rnorm(100, 0.1,1)
-
-Outreal <- atan2(CompII, CompI)
-
-X <- seq(-5, 5, 0.1)
-
-predCompI  <- 1.1 + 0.25*X
-predCompII <- 2.1 + -0.4*X
-
-PredCirc <- atan2(predCompII, predCompI)
-
-axval <- ((1.1*0.25)+(2.1*-0.4))/(1.1^2 + 2.1^2)
-acval <- atan2(2.1 + -0.4*axval, 1.1 + 0.25*axval)
-
-plot(X, PredCirc*(180/pi), ylim = c(-200,200), type = "l", xlab = "Predictor",
-     ylab = "Circular Outcome", bty = 'n', yaxt = "n", xaxt = "n")
-axis(1, xaxp = c(-5, 0, 5), at = c(seq(-5, 5, by = 1)),
-     labels = c(seq(-5, 5, by = 1)), las = 2)
-axis(2, xaxp = c(-180, 0,180), at = c(seq(-180, 180, by = 60)),
-     labels = c(seq(-180, 180, by = 60)), las = 2)
-points(Xreal, Outreal*(180/pi), cex=0.5)
-points(axval,
-       acval*(180/pi),
-       pch = 0)
-
-dev.off()
-tools::texi2dvi("Plots/circregline.tex",pdf=TRUE)
-
-```
-
-\begin{figure}[]
-  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
-  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to.
-The square indicates the inflection point of the regression line.}
-  \label{circregline}
-\end{figure}
-
-\vspace{-0.75cm}
-\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
-\vspace{-0.75cm}
-This model is an extension of the cylindrical model introduced in
-@abe2017tractable to the regression context. It concerns a combination
-of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
-\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
-with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
-\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
-contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
-model defined only on the positive real half-line \([0, + \infty)\) and thus can
-not be negative.\newline
-\indent In this model we predict the linear scale parameter and circular
-location parameter, both of which we can express in terms of
-covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
-\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
-parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
-\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
-\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
-2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
-regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
-\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
-vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
-is a vector of predictor values for the prediction of \(\theta\).\newline
-\indent For the teacher data, the predicted values for the circular outcome in
-the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
-\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
-The conditional distribution for the linear outcome is Weibull, meaning that we
-can use methods from survival analsis to interpret the effect of a predictor. In
-survival analysis a 'survival' function is used in which time is plotted against
-the probability of survival of subjects suffering from a specific medical
-condition. In the teacher data we can thus compute the probability of a teacher
-having a specific strength on the IPC. This probability is computed using the
-'survival-function' defined as  \[\exp(-\alpha
-y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
-with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
-function we also see that the circular concentration parameter \(\kappa\) and
-the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
-the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
-and strength of interpersonal behavior through the parameters $\beta_{0}$ and
-$\gamma_0$ as well as the effect of self-efficacy on type and strength of
-teacher behavior through the parameters $\beta_{1}$ and $\gamma_1$.\newline
-\indent We can use numerical optimization (Nelder-Mead) to find solutions for
-the maximum likelihood (ML) estimates for the parameters of the model.
-
-\vspace{-0.75cm}
-\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
-\vspace{-0.75cm}
-This model is an extension of the cylindrical model introduced by
-@mastrantonio2018joint to the regression context. The model contains \(m\)
-  independent circular outcomes and \(w\) independent linear outcomes. The
-  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
-  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
-  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
-  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
-  distribution [@sahu2003new]. Because the GPN distribution is modelled
-  using a so-called augmented representation (see also the description of the
-  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
-  modelling the multivariate skew normal distribution. As in
-@mastrantonio2018joint, dependence between the linear and circular outcome is
-  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
-  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
-  distribution.\newline
-\indent This means that we have a shared mean vector and variance-covariance
-matrix for the linear and circular outcome(s), much like having multiple
-outcomes in a MANOVA (multivariate analysis of variance) model. In our
-regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
-observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
-covariates. The mean vector then becomes \(\boldsymbol{M}_i =
-\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
-\times (2m + w)\) matrix with regression coefficients and intercepts and
-$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
-estimate an intercept and the $g$ covariate values. This means that in contrast
-to the other three models, we have to use the same set of predictors for the
-circular and linear outcome.\newline
-\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
-\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
-\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
-The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
-linear outcomes in the GPN-SSN model are
-\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
-\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
-\beta_{1_{s^{I}}}\text{SE}_i),\]
-
-\noindent and
-
-\[\hat{y_i} = \beta_{0_{y_i}} +
-\beta_{1_{y_i}}\text{SE}_i.\]
-
-\noindent The GPN-SSN model thus allows us to assess the average type and
-strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
-$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
-on type and strength of teacher behavior through the parameters
-$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
-because the type and strength of interpersonal behavior are modelled together
-using a multivariate normal distribution we can, through its variance-covariance
-matrix, also assess the dependence between the type and strength of interpersonal
-behavior. \newpage
-\indent We estimate the model using MCMC methods. A detailed description of these
-methods is given in the Supplementary Material.
-
-\vspace{-0.75cm}
-\subsection{Model Fit Criterion}\label{Modelfit}
-\vspace{-0.75cm}
-For the four cylindrical models we focus on their out-of-sample predictive
-performance to determine the fit of the model. To do so we use k-fold
-cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
-of the sample) is used once as a holdout set and 9 times as part of a training
-set. The analysis will thus be performed 10 times, each time on a different
-training set.\newline
-\indent A proper criterion to compare out-of-sample predictive performance is
-the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
-value of this criterion, the better the predictive performance of the model.
-Using ML estimates this criterion can be computed as follows:
-\begin{equation}\label{PLSLML}
-PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
-\end{equation}
-\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
-holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
-\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
-Using posterior samples the criterion is similar to the log pointwise predictive
-density (lppd) [@BDA, p. 169] and can be computed as:
-\begin{equation}\label{PLSLBayes}
-PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
-\end{equation}
-\noindent where \(B\) is the amount of posterior samples and
-\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
-parameters for the \(j^{th}\) iteration. Because the joint density and thus also
-the likelihood for the modified GPN-SSN model is not
-available in closed form [@mastrantonio2018joint] we compute the PLSL for the
-circular and linear outcome separately for all models. Note that although we fit
-the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
-prior information into account when assessing model fit with the PLSL. According
-to @BDA this is not necessary since we are assessing the fit of a model
-to data, the holdout set, only. They argue that the prior in such case is only
-of interest for estimating the parameters of the model but not for determining
-the predictive accuracy.\newline
-\indent For each of the four cylindrical models and for each of the 10
-cross-validation analyses we can then compute a PLSL for the circular and linear
-outcome by using the conditional log-likelihoods of the respective outcome (see
-Supplementary Material for a definition of the loglikelihoods). To evaluate the
-predictive performance we average across the PLSL criteria of the
-cross-validation analyses. We also assess the cross-validation variability by
-means of the standard deviations of the PLSL criteria.
-
-\vspace{-0.75cm}
+\label{Tableteacherstandardanalysis}
+\end{table}
+
 \section{Results}\label{DataAnalysis}
-\vspace{-0.75cm}
 In this section we analyze the teacher data with the help of the four
@@
 
-\vspace{-0.75cm}
 \subsection{Analysis}\label{DataResults}
-\vspace{-0.75cm}
 In the Supplementary Material we have described the starting values for the MCMC
@@
 compare the results of the four models we focus on the following aspects: the
-estimated average scores (intercept) on the type and strength of interpersonal
-behavior (1), the effect of self-efficacy on the type and strength of
-interpersonal behavior (2), the dependence between the type and strength of
-interpersonal behavior (3) and the model fit (4).
+estimated average scores (intercept) on the location and intensity on the IPC (1), the effect of self-efficacy on the location and intensity (2), the dependence between the location and intensity (3) and the model fit (4).
+
+\hfil [Table \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} about here] \hfil
 
@@
 \bottomrule
-\multicolumn{7}{l}{Note: $\beta_0^{I}$, $\beta_0^{II}$ and $\gamma_0$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{Note: $\beta_0^{I}$, $\beta_0^{II}$ and $\gamma_0$ inform us about the location and intensity on the IPC}\\
 \multicolumn{7}{l}{at the average self-efficacy. $\beta_1^{I}$, $\beta_1^{II}$ and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
-\multicolumn{7}{l}{type and strength of interpersonal behavior. $\gamma_{cos}$ and $\gamma_{sin}$ inform us about the dependence}\\
-\multicolumn{7}{l}{ between the type and strength of interpersonal behavior. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are}\\
-\multicolumn{7}{l}{elements of the variance-covariance matrix of the type of interpersonal behavior in the}\\
-\multicolumn{7}{l}{CL-GPN model and $\sigma$ is the error standard deviation of the strength of interpersonal behavior.}\\
+\multicolumn{7}{l}{location and intensity on the IPC. $\gamma_{cos}$ and $\gamma_{sin}$ inform us about the dependence}\\
+\multicolumn{7}{l}{ between the location and intensity on the IPC. $\sum_{1,1}$, $\sum_{1,2}$ and $\sum_{2,2}$ are}\\
+\multicolumn{7}{l}{elements of the variance-covariance matrix of the location on the IPC in the}\\
+\multicolumn{7}{l}{CL-GPN model and $\sigma$ is the error standard deviation of the intensity of interpersonal behavior.}\\
 
@@
 \bottomrule
-\multicolumn{8}{l}{Note: $\beta_0$  and $\gamma_0$ inform us about the type and strength }\\
-\multicolumn{8}{l}{of interpersonal behavior at the average self-efficacy. $\beta_1$ }\\
+\multicolumn{8}{l}{Note: $\beta_0$  and $\gamma_0$ inform us about the location and intensity }\\
+\multicolumn{8}{l}{on the IPC at the average self-efficacy. $\beta_1$ }\\
 \multicolumn{8}{l}{and $\gamma_1$ inform us about the effect of self-efficacy on the}\\
-\multicolumn{8}{l}{type and strength of interpersonal behavior. $\alpha$ is the } \\
-\multicolumn{8}{l}{shape parameter of the distribution of the strength }\\
-\multicolumn{8}{l}{of interpersonal bahavior. $\kappa$ and $\lambda$ are the concentration}\\
-\multicolumn{8}{l}{and skewness parameters for the distribution of the type}\\
-\multicolumn{8}{l}{of interpersonal behavior.}\\
+\multicolumn{8}{l}{location and intensity on the IPC. $\alpha$ is the } \\
+\multicolumn{8}{l}{shape parameter of the distribution of the location }\\
+\multicolumn{8}{l}{on the IPC. $\kappa$ and $\lambda$ respectively are the concentration}\\
+\multicolumn{8}{l}{and skewness parameters for the distribution of the location}\\
+\multicolumn{8}{l}{on the IPC.}\\
 \end{tabular}
@@
 \bottomrule
-\multicolumn{7}{l}{Note: $\beta_{0_s^{I}}$, $\beta_{0_s^{II}}$ and $\beta_{0_y}$ inform us about the type and strength of interpersonal behavior}\\
+\multicolumn{7}{l}{Note: $\beta_{0_s^{I}}$, $\beta_{0_s^{II}}$ and $\beta_{0_y}$ inform us about the location and intensity on the IPC}\\
 \multicolumn{7}{l}{at the average self-efficacy. $\beta_{1_s^{I}}$, $\beta_{1_s^{II}}$ and $\beta_{1_y}$ inform us about the effect of self-efficacy }\\
-\multicolumn{7}{l}{on the type and strength of interpersonal behavior. $\sum_{s_{1,1}}$, $\sum_{s_{1,2}}$, $\sum_{s_{2,2}}$, $\sum_{y_{3,3}}$, $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ }\\
+\multicolumn{7}{l}{on the location and intensity on the IPC. $\sum_{s_{1,1}}$, $\sum_{s_{1,2}}$, $\sum_{s_{2,2}}$, $\sum_{y_{3,3}}$, $\sum_{sy_{1,3}}$,  and $\sum_{sy_{2,3}}$ }\\
 \multicolumn{7}{l}{are elements of the variance-covariance matrix of which $\sum_{sy_{1,3}}$ and $\sum_{sy_{2,3}}$ inform us about}\\
-\multicolumn{7}{l}{the dependence between the type and strength of interpersonal behavior.}\\
-\multicolumn{7}{l}{$\lambda$ is the skewness parameter of the distribution of the strengths of interpersonal behavior.}\\
+\multicolumn{7}{l}{the dependence between the location and intensity on the IPC.}\\
+\multicolumn{7}{l}{$\lambda$ is the skewness parameter of the distribution of the intensity of interpersonal behavior.}\\
 \end{tabular}
@@
 
-\vspace{-0.5cm}
-\subsubsection{Average type and strength of interpersonal behavior}
+\subsubsection{Average location and intensity on the IPC}
 The parameters $\gamma_0$ in the CL-PN, CL-GPN and Abe-Ley model and the
-parameter $\beta_{0_y}$ in the GPN-SSN model inform us about the strength of
+parameter $\beta_{0_y}$ in the GPN-SSN model inform us about the intensity of
 interpersonal behavior at the average self-efficacy. For the CL-PN, CL-GPN and
 GPN-SSN models the parameters are estimated at 0.38, 0.37 and 0.33 respectively
-and are a direct prediction of the strength of interpersonal behavior at the
+and are a direct prediction of the intensity of interpersonal behavior at the
 average self-efficacy. The estimate for the GPN-SSN model is notably lower and
-likely to be caused by its skewed distribution for the strengths of
+likely to be caused by its skewed distribution for the intensity of
 interpersonal behavior. In the Abe-Ley model, $\gamma_0$ influences the shape
-parameter of the distribution of the strength of interpersonal behavior and does
-not directly estimate the average strength. Instead we can use the survival
-function to say something about the probability of having a certain strength of
+parameter of the distribution of the intensity of interpersonal behavior and does
+not directly estimate the average intensity. Instead we can use the survival
+function to say something about the probability of having a certain intensity of
 interpersonal behavior. Figure \ref{reglineweib} shows this function for several
@@
 for observations that fall within 1 standard deviation of the mean. The survival
-function indicates that the probability of having a low strength of
-interpersonal behavior is higher than having a high strength. We however can not
-make any direct statement about the estimated strength using the Abe-Ley
-model.\newline
+function indicates that the probability of having a low intensity of
+interpersonal behavior is higher than having a high intensity. We however can not
+make any direct statement about the estimated intensity using the Abe-Ley
+model.
+
+\hfil [Figure \ref{reglineweib} about here] \hfil
+
 \indent The parameters $\beta_0^{I}$, $\beta_0^{II}$, $\beta_0$,
-$\beta_{0_{s^{I}}}$ and $\beta_{0_{s^{II}}}$ inform us about the type of
-interpersonal behavior at the average self-efficacy for the CL-PN, CL-GPN,
+$\beta_{0_{s^{I}}}$ and $\beta_{0_{s^{II}}}$ inform us about the location on the IPC at the average self-efficacy for the CL-PN, CL-GPN,
 Abe-Ley and GPN-SSN model respectively. For the CL-PN, CL-GPN and GPN-SSN model
@@
 three models at 32.29$^\circ$, 33.70$^\circ$ and 35.53$^\circ$. In the Abe-Ley
-model the type of interpersonal behavior at the average self-efficacy is
+model the location on the IPC at the average self-efficacy is
 estimated at 0.36 radians or 20.63$^\circ$.
+
+\hfil [Table \ref{tab:means} about here] \hfil
 
@@
 
-plot(y, s[1,], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
+plot(y, s[1,], type = "l", ylab = "P(IPC intensity)", xlab = "IPC intensity", bty = "n")
 points(y, s[2,], type = "l", lty = 2)
@@
 \includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
-\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
+\caption{Plot showing the probability of having a particular intensity of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
 \label{reglineweib}
@@
 
-\vspace{-0.5cm}
 \subsubsection{The effect of self-efficacy}
 The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley models and $\beta_{1_y}$
-in the GPN-SSN model inform us about the effect of self-efficacy on the strength
+in the GPN-SSN model inform us about the effect of self-efficacy on the intensity
 of interpersonal behavior. For the CL-PN, CL-GPN and GPN-SSN model the
 parameters are estimated at 0.03, 0.03 and 0.09 respectively and are a direct
-estimate of the effect of self-efficacy on the strength of interpersonal
+estimate of the effect of self-efficacy on the intensity of interpersonal
 behavior, \emph{i.e.} an increase of 1 unit in self-efficacy leads to an
-increase of 0.09 units in the strength of interpersonal behavior according to
+increase of 0.09 units in the intensity of interpersonal behavior according to
 the GPN-SSN model. These estimates are however quite small and only different
@@
 say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to base our
-conclusions on. The models CL-GPN and CL-PN fit the linear outcome best according
-to the model fit in Table \ref{tab:ModelFit}. In these models the linear outcome has a
+conclusions on. The models CL-GPN and CL-PN fit the linear component best according
+to the model fit in Table \ref{tab:ModelFit}. In these models the linear component has a
 symmetric distribution whereas in the GPN-SSN the distribution of the linear
-outcome is skewed. However, the effect of self-efficacy is different from 0 only
-in the GPN-SSN model which does not seem to match with its lower model fit.\newline
+component is skewed. However, the effect of self-efficacy is different from 0 only
+in the GPN-SSN model which does not seem to match with its lower model fit.
+
+\hfil [Table \ref{tab:ModelFit} about here] \hfil
+
 \indent In the Abe-Ley model, $\gamma_1$ influences the shape parameter of the distribution of
-the strength of interpersonal behavior and does not directly estimate the effect
+the intensity of interpersonal behavior and does not directly estimate the effect
 of self-efficacy. Instead we can use the survival function to say something
-about the probability of having a certain strength of interpersonal behavior for
+about the probability of having a certain intensity of interpersonal behavior for
 different values of self-efficacy. Figure \ref{reglineweib} shows this function
@@
 \ref{dataplot}). This function indicates that the effect of self-efficacy on the
-strength of interpersonal behavior is not linear. The probability of having a
-higher strength of interpersonal behavior is highest for low self-efficacy and
+intensity of interpersonal behavior is not linear. The probability of having a
+higher intensity of interpersonal behavior is highest for low self-efficacy and
 lowest for average self-efficacy.\newline
@@
 $\beta_{1_{s^{I}}}$ and $\beta_{1_{s^{II}}}$ inform us about the effect of
-self-efficacy on the type of interpersonal behavior in the CL-PN, CL-GPN,
+self-efficacy on the location on the IPC in the CL-PN, CL-GPN,
 Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley models we have
@@
 Abe-Ley model and is equal to -0.03. Even though these slopes are
-different, the regression lines in Figure \ref{regline} are quite similar in the
+different\footnote{Note that the difference in regression lines (and thus inflection point) seems to be influenced by an outlier with a low self-efficacy and IPC value of approximately 160$^{\circ}$ = -200$^{\circ}$.}, the regression lines in Figure \ref{regline} are quite similar in the
 data range. Both the regression line of the Abe-Ley model and the CL-PN model
 show slopes that are not very steep in the range of the data indicating that the
-effect of self-efficacy on the type of interpersonal behavior is not large.
-\newline
+effect of self-efficacy on the location on the IPC is not large.
+
+\hfil [Figure \ref{regline} about here] \hfil
+
 \indent In the CL-GPN and GPN-SSN models we cannot compute circular regression
@@
 the circle. Instead, we will compute posterior predictive distributions for the
-predicted circular outcome of individuals scoring the minimum, maximum and
+predicted circular component of individuals scoring the minimum, maximum and
 median self-efficacy. The modes and 95\% HPD intervals of these posterior
@@
 scoring the minimum, median and maximum self-efficacy overlap. This indicates
-that the effect of self-efficacy, if there is any, on the type of interpersonal
-behavior a teacher shows is not expected to be strong. Had the HPD intervals not
-overlapped we could have concluded that as the self-efficacy increases, the
-score of the teacher on the IPC moves counterclockwise.
+that the effect of self-efficacy, if there is any, on the location on the IPC shows is not expected to be strong. Had the HPD intervals not
+overlapped we could have concluded that compared to teachers with a lower self-efficacy increases, the
+score on the IPC moves counterclockwise.
 
@@
 \end{figure}
-\vspace{-0.5cm}
-\subsubsection{Dependence between type and strength of interpersonal behavior.}
-The relation between the type and strength of interpersonal behavior in the
+
+\subsubsection{Dependence between location and intensity on the IPC}
+The relation between the location and intensity on the IPC in the
 CL-PN and CL-GPN models is described by the parameters $\gamma_{\cos}$ and
 $\gamma_{\sin}$. The HPD interval of \(\gamma_{\cos}\) does not include 0 for
-both the CL-PN and CL-GPN models, meaning that the cosine component of the type
-of interpersonal behavior has an effect on the strength of interpersonal
-behavior.\newline
+both the CL-PN and CL-GPN models, meaning that the cosine component of the location has an effect on the intensity.\newline
 \indent In the teacher data the sine and cosine components have a substantive
@@
 (Agency) represents the sine of an angle. For the teacher data this means that
-the Communion (cosine) dimension of the IPC positively effects the strength of a
-teachers' type of interpersonal behavior, in plain words: teachers exhibiting
+the Communion (cosine) dimension of the IPC positively effects the intensity of a
+teachers' interpersonal behavior, in plain words: teachers exhibiting
 interpersonal behavior types with higher communion scores (e.g., 'helpful' and
 'understanding' in Figure 2) are stronger in their interpersonal behavior.\newline 
-\indent In the GPN-SSN model the dependence between the type and strengths of
-interpersonal behavior is modelled through the covariances between the linear
-outcome and the sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
+\indent In the GPN-SSN model the dependence between the location and intensity on the IPC is modelled through the covariances between the linear
+component and the sine and cosine of the circular component \(\sum_{sy_{2,3}}\) and
 \(\sum_{sy_{1,3}}\). Both covariances, \(\sum_{sy_{2,3}} = 0.09\) and
@@
 behavior. Together with the results from the CL-PN and CL-GPN models in the
-previous paragraph this translates to the conclusion that teachers with the
-strongest interpersonal behavior have a type of interpersonal behavior between
+previous paragraph this translates to the conclusion that teachers with higher
+intensity scores have a location on the IPC that lies between
 0$^\circ$ and 90$^\circ$. To get these scores on the circle both the Agency and
@@
 corresponds to the pattern observed in the teacher data in Figure \ref{QTI}. At
-a strength of 0.4 and up we see that the scores on the circle range on average
+an intensity of 0.4 and up we see that the scores on the circle range on average
 between 0$^\circ$ and 100$^\circ$.
@@
 
-\vspace{-0.5cm}
 \subsubsection{Model fit}
-Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the linear
-and circular outcomes of the four cylindrical models fit to the teacher
+Table \ref{tab:ModelFit} shows the values of the PLSL criterion for the linear
+and circular components of the four cylindrical models fit to the teacher
 data.\newline \indent The CL-PN and CL-GPN models have the best out-of-sample
-predictive performance for the linear outcome. They show roughly the same
-performance because they model the linear outcome in the same way. We should
+predictive performance for the linear component. They show roughly the same
+performance because they model the linear component in the same way. We should
 note that even though the predictive performance of the Abe-Ley model for the
-linear outcome is worst on average, the standard deviation of the
+linear component is worst on average, the standard deviation of the
 cross-validation estimates is rather large. This means that in some samples, the
@@
 \indent The Abe-Ley model has the best out-of-sample predictive performance for
-the circular outcome. This would suggest that for the circular variable a
+the circular component. This would suggest that for the circular variable a
 slightly skewed distribution fits best. However, both the GPN-SSN  and the
 CL-GPN models fit much worse even though the distribution for the circular
-outcome in these models can also take a skewed shape. It should be noted that
+component in these models can also take a skewed shape. It should be noted that
 the  standard deviation of the cross-validation estimates is rather large for
@@
 standard deviation for only a few cylindrical models and not for all.\newline
-\indent In this situation, where one model fits the linear outcome best and
-another one fits the circular outcome best, it is hard to determine which model
+\indent In this situation, where one model fits the linear component best and
+another one fits the circular component best, it is hard to determine which model
 we should choose. In this case the results for the CL-PN /CL-GPN and Abe-Ley
 model are quite different regarding the effect of self-efficacy on the linear
-outcome (strength of interpersonal behavior). Because the Abe-Ley fit for the
+component (intensity of interpersonal behavior). Because the Abe-Ley fit for the
 linear part is worst we would choose to trust the results for the CL-PN and
@@
 models do not differ as much from the Abe-Ley model and we reach the same
-conclusion for both models, namely that the effect of self-efficacy on type of
-interpersonal behavior is not very strong. Therefore we would prefer the
+conclusion for both models, namely that the effect of self-efficacy on location on the IPC is not very strong. Therefore we would prefer the
 CL-PN/CL-GPN models in this case because where it matters in terms of
@@
 
-\caption{\label{tab:ModelFit}PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four cylindrical models}
+\caption{\label{tab:ModelFit}PLSL criteria, cross-validation mean and standard deviation, for the circular and linear component in the four cylindrical models}
 \centering
@@
 
-\vspace{-0.75cm}
 \section{Discussion}\label{Discussion}
-\vspace{-0.75cm}
 In this paper we modified four models for cylindrical data in such a way that
-they include a regression of both the linear and circular outcome onto a set of
+they include a regression of both the linear and circular component onto a set of
 covariates. Subsequently we have shown how these four methods can be used to
@@
 \indent Concerning the teacher data, the advantage of cylindrical data analysis
-is that we were able to analyze the information about the type and strength of
-interpersonal behavior simultaneously. In previous research, the two components
-of the interpersonal circumplex (\emph{i.e.}, Agency and Communion) were
-analyzed separately. Such an approach also provides information about the
-strength of teachers' score on Agency and Communion, yet a large portion of
-information about the combination of Agency and Communion, which describes the
-type of behavior that is observed, gets lost. A first solution to include both
-dimensions as a circular variable in data analysis was described by
+is that we were able to analyze the information about the location and intensity
+on the IPC simultaneously. In previous research, the two components of the
+interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
+separately. Such an approach only provides information about the intensity of a
+teachers' score on Agency and Communion and the information about the
+combination of Agency and Communion, which describes the location on the IPC,
+gets lost as was observed in the Standard Analysis section. A first solution to
+include both dimensions as a circular variable in data analysis was described by
 @Cremers2018Assessing. A downside of that analysis was that information about
-the strength of the specific type of interpersonal behavior could not be
-retained. In the present study, we have shown how using cylindrical models can
-simultaneously model the information about the type of and strength of
-interpersonal behavior and how these are influenced by teachers' self-efficacy
-in classroom management. Although we do not find any strong effects of
-self-efficacy on either the type or strength of behavior, the four cylindrical
-models do provide a way of analyzing and interpreting this effect. This is
-beneficial for future research in which we may want to investigate the effect of
-further covariates on data from the circumplex. \newline
+the intensity could not be retained. In the present study, we have shown how
+using cylindrical models can simultaneously model the information about the
+location and intensity on the IPC and how these are influenced by teachers'
+self-efficacy in classroom management. Although we do not find any strong
+effects of self-efficacy on either the location and intensity\footnote{similar
+to the standard analysis}, the four cylindrical models do provide a way of
+analyzing and interpreting this effect. This is beneficial for future research
+in which we may want to investigate the effect of further covariates on data
+from the circumplex. \newline
 \indent Furthermore, in addition to being able to assess the influence of
 covariates, the cylindrical models also provide information about the dependence
-between the type and strength of interpersonal behavior. We found that stronger
-behavior is associated with higher scores on the Communion and in some models
-also the Agency dimension. This implies that teachers whose type of
-interpersonal behavior ranges between 0$^\circ$ and 90$^\circ$, the 'helpful'
-and 'directing' subtypes are stronger in their behavior than teachers of the
-other subtypes.\newline
+between the location and intensity on the IPC. This is a second advantage over
+the standard analysis in which it is impossible to quantify the relation between
+intensity and location. We found that stronger behavior is associated with
+higher scores on the Communion and in some models also the Agency dimension.
+This implies that teachers whose interpersonal behavior ranges between
+0$^\circ$ and 90$^\circ$ on the IPC, the 'helpful' and 'directing' subtypes, are
+stronger in their behavior than teachers of the other subtypes.\newline
 \indent As mentioned in the introduction, data from the interpersonal circumplex
@@
 the models is of use in these fields as well. \newline
-\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
-best out of the four cylindrical models.  In the
-CL-GPN and GPN-SSN models the interpretation of the parameters of the
-circular outcome component is not straightforward, if at all possible.
-This is caused by the fact that in addition to the mean vector the
-covariance matrix of the GPN distribution affects the location of the
-circular data, making it difficult to compute regression coefficients on
-the circle. @wang2012directional state that Monte Carlo integration
-can be used to compute a circular mean and variance for the GPN
-distribution. In future research, this solution might be applied to the
-methods of @CremersMulderKlugkist2017 in order to compute circular
-coefficients for GPN models.\newline
-\indent In terms of flexibility the GPN-SSN model scores best. Multiple
-linear and circular outcomes can be included and we can thus apply the
-model to multivariate cylindrical data. In addition the GPN-SSN, the
-CL-GPN and CL-PN models are extendable to a mixed-effects structure and
-can thus also be fit to longitudinal data (see @nunez2014bayesian and @hernandez2016general for
+\indent Out of the four cylindrical models investigated in this paper, the
+results from CL-PN and Abe-Ley models are most straightforward to interpret. In
+the CL-GPN and GPN-SSN models the interpretation of the parameters of the
+circular outcome component is more complex, if at all possible. This is caused
+by the fact that in addition to the mean vector the covariance matrix of the GPN
+distribution affects the location of the circular data, making it difficult to
+compute regression coefficients on the circle. @wang2012directional state that
+Monte Carlo integration can be used to compute a circular mean and variance for
+the GPN distribution. In future research, this solution might be applied to the
+methods of @CremersMulderKlugkist2017 in order to compute circular coefficients
+for GPN models.\newline
+\indent The GPN-SSN model is more flexible compared to the other three
+cylindrical models that were investigated in this paper. Multiple linear and
+circular components can be included and we can thus apply the model to
+multivariate cylindrical data. In addition the GPN-SSN, the CL-GPN and CL-PN
+models are extendable to a mixed-effects structure and can thus also be fit to
+longitudinal data (see @nunez2014bayesian and @hernandez2016general for
 hierarchical/mixed-effects models for the PN and GPN distributions
-respectively). For the Abe-Ley model this may also be possible but has
-not been done in previous research for the conditional distribution of
-its circular outcome (sine-skewed von
-Mises). Concerning asymmetry, both the
-GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
-the distributions of both the linear and circular outcome, while the CL-GPN
-model permits an asymmetric  circular outcome.\newline
+respectively). For the Abe-Ley model this may also be possible but has not been
+done in previous research for the conditional distribution of its circular
+component (sine-skewed von Mises). Concerning asymmetry, both the GPN-SSN as
+well as the Abe-Ley model allow for non-symmetrical shapes of the distributions
+of both the linear and circular component, while the CL-GPN model permits an
+asymmetric  circular component.\newline
 \indent The four cylindrical models that were modified to the regression
@@
 \newpage
-\vspace{-0.75cm}
 \section*{{\normalfont References}}
-\vspace{-0.75cm}
 <div id="refs"></div>

### 2019-07-26 Manuscript/Manuscript.tex
Corrections Christophe

@@
 
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{lineno}
@@
 movement), eye-tracking research (direction and length of saccades) and
-data from an interpersonal circumplex (type and strength of
-interpersonal behavior). In this paper we adapt four models for
+data from an interpersonal circumplex (location and intensity on the IPC).
+In this paper we adapt four models for
 cylindrical data to include a regression of the circular and linear
@@
 
-\vspace{-0.75cm}
-
-In social sciences the use of cylindrical data is very common. Such data consist
-of a linear and a circular component. Gurtman (2011) refers to such data
-as vectors, with a directional measure (i.e., the circular component) and a
-measure indicating the magnitude (i.e., the linear component). Many established
-models in psychology are often referred to as circular or circumplex models, but
-those models are cylindrical. Examples of such cylindrical models are the
-interpersonal circle/circumplex
+In the social sciences the use of cylindrical data is very common. Such data
+consist of a linear and a circular component. Gurtman (2011) refers to
+such data as vectors, with a directional measure (i.e., the circular component)
+and a measure indicating the magnitude (i.e., the linear component). Many
+established models in psychology are often referred to as circular or circumplex
+models, but those models are cylindrical. Examples of such cylindrical models
+are the interpersonal circle/circumplex
 (Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans, \DEN{Brok}{Den}{den} Brok, \& \VAN{Tartwijk}{Van}{van} Tartwijk, 2006), the circumplex of
@@
 \newline
-\indent Also, many of the more recent types of data that are studied in psychology are
-cylindrical. For example, research on human navigation uses data where distance
-(i.e., the linear component) and direction (i.e., the circular component) are of
-interest (Chrastil \& Warren, 2017) or in eye-tracking, the saccade data also
-consist of both the direction (i.e., the circular variable) and the duration
-(i.e., the linear variable) (e.g., Rayner (2009)). Apart from the social
-sciences, data with a circular and linear outcome more commonly occur in
-meteorology (García-Portugués, Crujeiras, \& González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras, González-Manteiga, \& Pereira, 2014) or
-marine research (Lagona, Picone, Maruotti, \& Cosoli, 2015)
-\newline
+\indent Also, many of the more recent types of data that are studied in
+psychology are cylindrical. For example, research on human navigation uses data
+where distance (i.e., the linear component) and direction (i.e., the circular
+component) are of interest (Chrastil \& Warren, 2017) or in eye-tracking, the
+saccade data also consist of both the direction (i.e., the circular variable)
+and the duration (i.e., the linear variable) (e.g., Rayner (2009)). Apart from
+the social sciences, data with a circular and linear component more commonly
+occur in meteorology (García-Portugués, Crujeiras, \& González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras, González-Manteiga, \& Pereira, 2014) or marine
+research (Lagona, Picone, Maruotti, \& Cosoli, 2015) \newline
 \indent Up until now researchers studying cylindrical data had to rely on linear
-statistical methods to analyze their research results. However, lately more and more
-of these researchers acknowledge that linear methods are not sufficient and call
-for new methods (Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy, \& Hilsenroth, 2009) that take into account both the circular and the linear
+statistical methods to analyze their research results. However, lately more and
+more of these researchers acknowledge that linear methods are not sufficient and
+call for new methods (Gurtman, 2011; Pennings, 2017b; Wright, Pincus, Conroy, \& Hilsenroth, 2009) that take into account both the circular and the linear
 component of these data.
 \newline
-\indent The aim in the present paper is twofold. Firstly, we intend to fill
-the above mentioned gap in the literature by showing that the use of cylindrical
+\indent The aim in the present paper is twofold. Firstly, we intend to fill the
+above mentioned gap in the literature by showing that the use of cylindrical
 models can benefit the analysis of circumplex data and cylindrical data in
@@
 of both the linear and circular component of a cylindrical variable onto a set
-of covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. These adapted cylindrical models are
-then used to analyse the teacher data.
-
-\vspace{-0.75cm}
-\subsection{Cylindrical Data}
-\vspace{-0.75cm}
-
-Data that consist of a linear variable and a circular variable is called
-cylindrical data. A circular variable is different from the linear variable in
+of covariates. These adapted cylindrical models are then used to analyse the
+teacher data.
+
+\section{Modelling Framework}
+
+Data that consist of a linear variable and a circular variable are called
+cylindrical data. The circular variable is different from the linear variable in
 the sense that it is measured on a different scale. Figure \ref{circline} shows
@@
 Mardia and Jupp (2000), Jammalamadaka and Sengupta (2001) and Ley and Verdebout (2017) for a more
-elaborate overview).\newline
-\indent In the literature, several methods have been put forward to model the
-relation between the linear and circular component of a cylindrical variable.
+elaborate overview).
+
+\hfil [Figure \ref{circline} about here] \hfil
+
+\indent From now on we will refer to the circular random variable as \(\Theta\),
+the circular component, and the linear random variable as \(Y\), the linear
+component, with realizations \(\theta_i\) and \(y_i\) for each measurement \(i=1, \dots, n\) where \(n\) is the sample size. The circular component is measured in
+radians and consequently takes values in the interval \([0, 2\pi)\). Note that the
+round brackets mean that \(2\pi\) is not included in the interval since this
+represents the same value as 0 as a result of periodicity. Radians can be
+converted to degrees using the relation \(1 \:\text{rad} = 1*\frac{180^\circ}{\pi}\). A circular variable is usually described using its
+first two moments, circular location or mean \(\mu\) and mean resultant length
+\(\rho\)\footnote{In the von Mises distribution, a common distribution for a
+circular outcome, we have a concentration parameter $\kappa$ that is related to
+$\rho$ as $\rho = A_1(\kappa)$, where $A_1{\kappa} = I_1(\kappa)/I_0(\kappa)$
+and $I_0()$ and $I_1()$ are modified Bessel functions of order 0 and 1,
+respectively.}. The sample values of these two parameters are referred to as
+\(\bar{\theta}\) and \(\hat{\rho}\). The mean resultant length lies between 0,
+meaning the data is not concentrated at all \emph{i.e.} spread over the entire
+circle, and 1, meaning all data is concentrated at a single point on the circle.
+The linear component is measured on the real line and can take values on
+\((-\infty, \infty)\) or we can constrain it to the positive real line \((0, \infty)\). Depending on which distributional assumptions are used we describe the
+distribution of the linear component using a mean \(\mu\) and variance \(\sigma^2\)
+or a scale \(\nu\) and shape \(\alpha\) parameter. When skewed distributions are
+used for either the circular or linear component we may also include a skewness
+parameter \(\lambda\).\newline
+\indent In a cylindrical framework we model the location \(\mu_c\) of the circular
+variable and the mean \(\mu_l\) or scale \(\nu\) of the linear variable. Note that
+to prevent confusion we distinguish between the circular location and linear
+mean using subscripts. In this paper we use covariates to predict the circular
+and linear component using the following general type of prediction equations
+for a cylindrical model with \(q\) covariates for the linear component and \(p\)
+covariates for the circular component:
+\begin{equation}\label{circpredgen}
+\hat{\theta_i} = g(\beta_0 + \beta_1z_1 + \dots + \beta_qz_p),
+\end{equation}
+\begin{equation}\label{linpredgen}
+\hat{y_i} = h(\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q),
+\end{equation}
+where \(g()\) and \(h()\) are link functions, \(\boldsymbol{\beta} = (\beta_0, \dots, \beta_p)\) is a vector of regression coefficients for the circular variable and
+\(\boldsymbol{\gamma} = (\gamma_0, \dots, \gamma_q)\) is a vector of regression
+coefficients for the linear variable. The link functions that are chosen depend
+on the type of distribution we choose for the linear and circular component. In
+this paper we use either an identity link or an exponential function (when
+modelling a scale parameter) for the linear component. For the circular
+component the link functions are specific to circular data and include
+arctangent functions. The specific type of link function and prediction
+equations used will be introduced in the respective descriptions of the models
+used in this paper.\newline
+\indent In addition to modelling the circular and linear component of a
+cylindrical variable separately using covariates we also model the relation
+between them. In the literature, several methods have been put forward to do so.
 Some of these are based on regressing the linear component onto the circular
-component using the following type of relation: \[y = \beta_0 +
-\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
-linear component and \(\theta\) the circular component (Johnson \& Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \& Jona-Lasinio, 2015). Others model the relation in a
-different way, e.g.~by specifying a multivariate model for several linear and
-circular variables and modelling their covariance matrix
-(Mastrantonio, 2018) or by proposing a joint cylindrical distribution. For
-example, Abe and Ley (2017) introduce a cylindrical distribution based on a
-Weibull distribution for the linear component and a sine-skewed von Mises
-distribution for the circular component and link these through their respective
-shape and concentration parameters. However, none of the methods that have been
-proposed thus far include additional covariates onto which both the circular and
-linear component are regressed.
+component by including \(\sin(\theta_i)\) and \(\cos(\theta_i)\) into the prediction
+equation in \eqref{linpredgen} (Johnson \& Wehrly, 1978; Mardia \& Sutton, 1978; Mastrantonio, Maruotti, \& Jona-Lasinio, 2015). The linear component is thus modeled using the sine
+and cosine of the circular component. Others model the relation in a different
+way, e.g.~by specifying a multivariate model for the linear and circular
+variable and modelling their covariance matrix (Mastrantonio, 2018) or by
+proposing a joint cylindrical distribution. For example, Abe and Ley (2017)
+introduce a cylindrical distribution based on a Weibull distribution for the
+linear component and a sine-skewed von Mises distribution for the circular
+component and link these through their respective shape and concentration
+parameters.
 
@@
 
-\vspace{-0.75cm}
+\subsection{Four Cylindrical Regression Models}\label{Models}
+
+One of the goals of this paper is to show the benefits of cylindrical methods
+for the analysis of circumplex data and cylindrical data in psychology in
+general. To do so we focus on four cylindrical models. The models were selected
+for their relatively low complexity and the ease with which a regression
+structure could be incorporated. But also because they show different ways of
+modelling the linear and circular component and thereby illustrate a wider range
+of cylindrical models available in the literature. As outlined in the previous
+section, the cylindrical models contain a set of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\) predictors \(\boldsymbol{z} = z_1, \dots, z_p\) for the
+linear and circular components, \(Y\) and \(\Theta\), respectively. The first
+two models are based on a construction by Mastrantonio et al. (2015), while the
+other models are extensions of the models from Abe and Ley (2017) and
+Mastrantonio (2018). The four cylindrical models are introduced separately
+in the subsections below. However, to provide a more succinct overview and
+comparison of the four models, Table \ref{TableModels} gives an overview of the
+similarities and differences between the models.
+
+\begin{sidewaystable}[h]
+\centering
+\caption{Comparison of the four cylindrical regression models} 
+\begin{tabular}{lllll}
+  \noalign{\smallskip}\hline\noalign{\smallskip}
+\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
+$\Theta$ & &&&\\
+$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
+$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
+$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
+            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
+$Y$& &&&\\
+$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
+$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
+$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
+            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
+$\Theta$-$Y$ dependence &                                   &                                   & & \\
+                        & $Y$ regressed on                  & $Y$ regressed on                  & circular concentration $\kappa$  & covariance \\
+                        & $\sin(\Theta)$ and $\cos(\Theta)$ & $\sin(\Theta)$ and $\cos(\Theta)$ & and linear scale $\alpha$ & matrix\\\hline
+\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
+\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
+
+\end{tabular}
+\label{TableModels}
+\end{sidewaystable}
+
+\hfil [Table \ref{TableModels} about here] \hfil
+
+\subsubsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+
+Following Mastrantonio et al. (2015) we consider two models where the prediction
+equation for the linear component is specified by combining \eqref{linpredgen},
+with an identity link function and including \(\sin(\theta_i)\) and
+\(\cos(\theta_i)\) as follows:
+\begin{equation}\label{linpredCLPNCLGPN}
+\hat{y_i} = \gamma_0 + \gamma_{cos}*\cos(\theta_i)*r_i + \gamma_{sin}*\sin(\theta_i)*r_i + \gamma_1*x_1 + \dots + \gamma_q*x_q,
+\end{equation}
+\noindent where \(r_i\) is a realization of the unobserved random variable
+\(R\geq0\) that will be introduced below, \(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the intercept and regression
+coefficients and \(x_1, \dots, x_q\) are the \(q\) covariates for the prediction
+of the linear component. We assume a normal \(N(\mu_l, \sigma^2)\) distribution
+for the linear component.\newline
+\indent For the circular component we assume either a projected normal (PN) or a
+general projected normal (GPN) distribution. These distributions arise from a
+projection of a distribution defined in bivariate space onto the circle. Figure
+\ref{projection} represents this projection. In the left plot of Figure
+\ref{projection} we see realizations \((s_i^{I}, s_i^{II})\) from the bivariate
+normal variable \(\boldsymbol{S}\) that in the middle plot are projected to form
+the circular component \(\Theta\) in the right plot. Mathematically the relation
+between \(\boldsymbol{S}\) and \(\Theta\) is defined as follows:
+\begin{equation}\label{projection_eq}
+\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
+\end{equation}
+\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
+\(\boldsymbol{S}\), that is represented by the lines connecting the bivariate
+datapoints to the origin in the middle plot. We call \(\boldsymbol{S}\) the
+augmented representation of the circular component. It is a variable that in
+contrast to \(\Theta\) is not observed and thus considered latent or auxiliary.
+This then means that we do not model \(\Theta\) directly but indirectly through
+\(\boldsymbol{S}\).
+
+\hfil [Figure \ref{projection} about here] \hfil
+
+\indent For both the PN and GPN distributions the circular location parameter
+\(\mu_c\in [0, 2\pi)\) is modeled.\footnote{Note that for the CL-GPN
+model the circular location parameter also depends on the variance-covariance
+matrix and the circular predicted values should be computed using numerical
+integration or Monte Carlo methods because a closed form expression for the mean
+direction is not available.} The prediction equation for the circular component is specified by using a double arctangent link function in \eqref{circpredgen} as follows:
+\begin{equation}\label{circpredCLPNCLGPN}
+\hat{\theta_i} = \text{atan2}(\boldsymbol{\beta^{II}}\boldsymbol{z}_i, \boldsymbol{\beta^{I}}\boldsymbol{z}_i)
+\end{equation}
+\noindent where \(\boldsymbol{\beta}^{I} = (\beta_0^{I}, \beta_1^{I}, \dots, \beta_p^{I})\) and
+\(\beta^{II} = (\beta_0^{II}, \beta_1^{II}, \dots, \beta_p^{II})\) are vectors with
+intercepts and regression coefficients for the prediction of \(S^{I}\) and
+\(S^{II}\) and \(\boldsymbol{z}_i\) is a vector with predictor values for each
+individual \(i \in 1, \dots, n\) where \(n\) is the sample size. Note that as a
+result of the augmented representation of the circular component we have two sets
+of regression coefficients and intercepts, in contrast to a single set in
+\eqref{circpredgen}. This leads to problems when we want to interpret the effect
+of a covariate on the circle. A circular regression line for simulated data is shown in Figure
+\ref{circregline}, with covariate values on the x-axis and the predicted
+circular component on the y-axis. As can be seen it is of a non-linear character
+meaning that the effect of a covariate is different at different values of the
+covariate. A circular regression line is usually described by the slope at the
+inflection point, the point at which the slope of the regression line starts
+flattening off (indicated with a square in Figure \ref{circregline}). By
+default, the parameters from the PN and GPN models do not directly describe this
+inflection point. For the PN distribution however, Cremers et al. (2018b)
+solved this interpretation problem. They introduce a new parameter \(b_c\) that describes the slope at
+the inflection point of the regression line. For the GPN distribution the
+interpretation problem however remains.
+
+\hfil [Figure \ref{circregline} about here] \hfil
+
+\indent The main difference between the PN and GPN distribution lies in the
+definition of their covariance matrix. For the PN distribution this is an
+identity matrix, causing the distribution to be unimodal and symmetric, whereas
+for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\),
+allowing for multimodality and asymmetry/skewness as illustrated in Figure \ref{plotGPN}.
+
+\hfil [Figure \ref{plotGPN} about here] \hfil
+
+\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
+Carlo (MCMC) methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011), Wang and Gelfand (2013) and
+Hernandez-Stumpfhauser, Breidt, and \VANDER{Woerd}{Van der}{van der} Woerd (2016) for the regression of the circular component. A detailed
+description of the Bayesian estimation and MCMC samplers can be found in the
+Supplementary Material.
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
+\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
+The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norms of the bivariate datapoints, realizations of the random variable $R$.}
+\label{projection}
+\end{figure}
+
+\begin{figure}[]
+  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
+  \caption{Circular regression line for the relation between a covariate and a circular component with the simulated data the regression line was fit to.
+The square indicates the inflection point of the regression line.}
+  \label{circregline}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width = \textwidth]{Plots/plotGPN.pdf}
+\caption{Plot showing 1000 samples from three GPN distributions with mean vector $(2,2)$ and $\tau = 1,5,10$ and $\xi = 0, 0.9, -0.5$ from left to right}
+\label{plotGPN}
+\end{figure}
+
+\subsubsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+
+This model is an extension of the cylindrical model introduced in
+Abe and Ley (2017) to the regression context. The circular dependence between the
+linear and circular component, \(Y\) and \(\Theta\), is defined through a joint
+density function as follows:
+
+\begin{equation}\label{WeiSSVMdensity}
+f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
+                 (1 +\lambda\sin(\theta - \mu_c))
+                 y^{\alpha-1}
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu_c))],
+\end{equation}
+
+\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
+\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters
+respectively. The Abe-Ley density thus concerns a combination of a Weibull
+distribution, with scale parameter \(\nu>0\) and shape parameter \(\alpha\), for
+the linear component and a sine-skewed von Mises distribution, with location
+parameter \(\mu_c\in [0, 2\pi)\), concentration parameter \(\kappa>0\) and
+skewness \(\lambda \in [-1,1]\), for the circular component. In contrast to the
+CL-PN and CL-GPN models, the linear component \(Y\) is in this model defined
+only on the positive real half-line \([0, + \infty)\) and thus can not be
+negative. Our modification of the Abe-Ley model occurs at the level of the
+linear scale parameter \(\nu>0\) and circular location parameter \(\mu_c\in [0, 2\pi)\), both of which we will model using covariates.\newline
+\indent The prediction equation for the circular component equals:
+\begin{equation}\label{circpredcAL}
+\hat{\theta}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{\beta}\boldsymbol{z}_i)
+\end{equation}
+Note that we have taken out the intercept \(\beta_0\) from the link function \(g() = 2\tan^{-1}\) in \eqref{circpredgen}. The vector \(\boldsymbol{\beta}\) thus does
+not contain an intercept. We do not directly predict the linear component. The
+conditional distribution for the linear component is Weibull, meaning that we can
+use methods from survival analsis to interpret the effect of a predictor. In
+survival analysis a \enquote{survival} function is used in which time is plotted against
+the probability of survival of subjects suffering from a specific medical
+condition. This probability is computed using the \enquote{survival-function} defined as
+\begin{equation}\label{survivalfunc}
+\exp(-\alpha
+y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\theta}_i))^{1/\alpha}}),
+\end{equation}
+\noindent with \(\hat{\nu}_i = \exp(\boldsymbol{\gamma}\boldsymbol{x}_i)\). Note that \(\boldsymbol{x}_i\) also includes a 1 to be able to estimate the intercept \(\gamma_0\). From
+the survival function we also see that the circular concentration parameter
+\(\kappa\) and the linear shape parameter \(\alpha\) regulate the
+circular-linear dependence in the Abe-Ley model.
+In the exponent or power of the linear outcome, the prediction error on the circle, \(\cos(\theta_i - \hat{\theta_i})\), is multiplied by \(\tanh{\kappa}\), and raised to the power \(1-\alpha\).\newline
+\indent We can use numerical optimization (Nelder-Mead) to find solutions for
+the maximum likelihood (ML) estimates for the parameters of the model.
+
+\subsubsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
+
+This model is an extension of the cylindrical model introduced by
+Mastrantonio (2018) to the regression context. Although the model may contain
+several circular and linear components we will restrict ourselves to one
+circular and one linear component in this paper. The circular component is
+modelled by a GPN distribution, as in the CL-GPN model, while the linear component is modelled by a skew
+normal distribution (Sahu, Dey, \& Branco, 2003). Because the GPN distribution is modelled
+using a so-called augmented representation (see \eqref{projection_eq}) it is
+convenient to use a similar tactic for modelling the skew normal distribution.
+As in Mastrantonio (2018), dependence between the linear and circular
+component is created by modelling the augmented representation of \(\Theta\) and
+\(Y\) together in a 3 dimensional normal distribution. The joint density of the model is then represented by:
+\begin{equation}\label{YDThetarjoint} 
+f(\theta, r, y, d) = 2\phi_{3}((\boldsymbol{s}^t,
+y^t)^t \mid \boldsymbol{M} + (0,0,\lambda d)^t, \boldsymbol{\Sigma})
+\phi(d \mid 0, 1)r, 
+\end{equation}
+\noindent where \(\phi\) represents the normal probability density function, \(\boldsymbol{s} = (r(\cos(\theta), \sin(\theta)))^t\), \(d\) is
+the augmented representation of the linear component, \(\boldsymbol{M} = \boldsymbol{B}^t\boldsymbol{X}\), \(\boldsymbol{B}_{(g + 1) \times 3}\) is
+defined below, \(\boldsymbol{X}\) is a design matrix and \(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\) a
+variance-covariance matrix. The matrices \(\boldsymbol{\Sigma}_s\) and \(\boldsymbol{\Sigma}_y\) are the
+covariance matrices for the variances of and covariances between the augmented
+representations of the circular and linear component respectively.
+Note that because we only have one circular and one linear component in this paper, \(\boldsymbol{\Sigma}_s\) is a two by two matrix and \(\boldsymbol{\Sigma}_y\) is a scalar representing the variance of the augmented linear component.
+The matrix
+\(\boldsymbol{\Sigma}_{sy}\) contains covariances between the augmented
+representations of the circular and linear component. The matrix with
+regression coefficients and intercepts, \(\boldsymbol{B}\) is defined as follows:
+\begin{equation}\label{regmatGPNSSN}
+\boldsymbol{B} = \begin{bmatrix}
+\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\
+\beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}}\\
+\vdots & \vdots & \vdots \\
+\beta_{g_{s^{I}}} & \beta_{g_{s^{II}}} & \beta_{g_{y}} \end{bmatrix}.
+\end{equation}
+\noindent Note that in contrast to the previous three models where we use \(p\)
+and \(q\) covariates for the circular and linear component respectively, we use \(g = p = q\) covariates for both components in the GPN-SSN model. Unlike the other
+models, where \(\mu_l\) and \(\mu_c\) are modelled separately, the GPN-SSN thus has
+a shared mean vector \(\boldsymbol{M}\) and variance-covariance matrix
+\(\boldsymbol{\Sigma}\) for the linear and circular component, much like having
+multiple outcomes in a MANOVA (multivariate analysis of variance) model. This
+implies that the covariances in the variance-covariance matrix describe the
+dependence between the circular and linear component.\newline
+\indent The prediction equation for the circular component\footnote{Note that for the  GPN-SSN model the predicted circular component also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} is similar to \eqref{circpredCLPNCLGPN} and equal to:
+\begin{equation}\label{circpredGPNSSN}
+\hat{\theta}_i = \text{atan2}(\boldsymbol{\beta^{II}}\boldsymbol{z}_i, \boldsymbol{\beta^{I}}\boldsymbol{z}_i),
+\end{equation}
+\noindent where \(\boldsymbol{\beta^I} = (\beta_{0_{s^{I}}}, \beta_{1_{s^{I}}}, \dots, \beta_{g_{s^{I}}})\) and \(\boldsymbol{\beta^{II}} = (\beta_{0_{s^{II}}}, \beta_{1_{s^{II}}}, \dots, \beta_{g_{s^{II}}})\).
+For the linear component we subsititute \(h()\) in \eqref{linpredgen} for an identity link and \(\boldsymbol{\gamma}\) for \(\boldsymbol{\beta}\):
+\begin{equation}\label{linpredGPNSSN}
+\hat{y}_i = \boldsymbol{\beta}\boldsymbol{x}_i,
+\end{equation}
+\noindent where \(\boldsymbol{x}_i = \boldsymbol{z}_i\) and \(\boldsymbol{\beta} = (\beta_{0_{y}}, \beta_{1_{y}}, \dots, \beta_{g_{y}})\).\newline
+\indent We estimate the model using MCMC methods. A detailed description of these
+methods is given in the Supplementary Material.
+
+\subsection{Model Fit Criterion}\label{Modelfit}
+
+For the four cylindrical models we focus on their out-of-sample predictive
+performance to determine the fit of the model. To prevent possible problems
+concerning overestimation, we use k-fold cross-validation and split our data
+into 10 folds. Each of these folds (10 \(\%\) of the sample) is used once as a
+holdout set and 9 times as part of a training set. The analysis will thus be
+performed 10 times, each time on a different training set.\newline
+\indent A proper criterion to compare out-of-sample predictive performance is
+the Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007). The lower the
+value of this criterion, the better the predictive performance of the model.
+Because the joint density and thus also the likelihood for the modified GPN-SSN
+model is not available in closed form (Mastrantonio, 2018) we compute the
+PLSL for the circular and linear component separately for all models. Using ML
+estimates this criterion can be computed as follows for the circular and linear
+component:
+\begin{equation}\label{PLSLMLcirc}
+PLSL_c = -2 \sum_{i = 1}^{M}\log l(\theta_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
+\end{equation}
+\begin{equation}\label{PLSLMLlin}
+PLSL_l = -2 \sum_{i = 1}^{M}\log l(y_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
+\end{equation}
+\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
+holdout set, \(y_i\) and \(\theta_i\) are the \(i^{th}\) datapoints from the
+holdout set and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the
+model parameters. Using posterior samples the criterion is similar to the log
+pointwise predictive density (lppd) (Gelman et al., 2014, p. 169) and can be computed for the
+circular and linear component as:
+\begin{equation}\label{PLSLBayescirc}
+PLSL_c = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(\theta_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
+\end{equation}
+\begin{equation}\label{PLSLBayeslin}
+PLSL_l = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(y_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
+\end{equation}
+\noindent where \(B\) is the amount of posterior samples and
+\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
+parameters for the \(j^{th}\) iteration. Note that although we fit the CL-PN,
+CL-GPN and GPN-SSN models using Bayesian statistics, we do not take prior
+information into account when assessing model fit with the PLSL. According to
+Gelman et al. (2014) this is not necessary since we are assessing the fit of a model to data,
+the holdout set, only. They argue that the prior in such case is only of
+interest for estimating the parameters of the model but not for determining the
+predictive accuracy.\newline
+\indent For each of the four cylindrical models and for each of the 10
+cross-validation analyses we can then compute a PLSL for the circular and linear
+component by using the conditional log-likelihoods of the respective component (see
+Supplementary Material for a definition of the loglikelihoods). To evaluate the
+predictive performance we average across the PLSL criteria of the
+cross-validation analyses. We also assess the cross-validation variability by
+means of the standard deviations of the PLSL criteria.
+
 \section{Teacher Data}\label{Example}
-\vspace{-0.75cm}
 \begin{figure}
@@
 also implies that scores on the IPC could be viewed as a circular variable. This
-circular variable represents the type of interpersonal behavior that a teacher
-shows towards his/her students.\newline
+circular variable represents location on the IPC of the interpersonal behavior
+that a teacher shows towards his/her students.
+
+\hfil [Figure \ref{QTI} about here] \hfil
+
 \indent Cremers et al. (2018a) explain the circular nature of the IPC data and
@@
 The resulting circular variable \(\theta\) can then be modelled and takes values
-in the interval \([0, 2\pi)\) or \([0^\circ, 360^\circ)\). Note that the round
-brackets mean that \(2\pi\) and \(360^\circ\) are not included in the interval since
-these represent the same value as 0 as a result of periodicity. \newline
+in \([0, 2\pi)\).\newline
 \indent A circular analysis of circumplex data has several benefits: it is more
@@
 This provides us with new insights compared to a separate analysis of the two
-dimensions that is standard in the literature (see e.g.
+dimensions that is standard in the literature (see, e.g.,
 Pennings et al. (2018), Wright et al. (2009) or
@@
 Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
-strength of the interpersonal behavior a teacher shows towards his/her students.
-In a cylindrical model this strength (the linear outcome) can be modeled
-together with the type of interpersonal behavior of a teacher (the circular
-outcome). This leads to an improved analysis of interpersonal circumplex data,
-over either analyzing the two dimensions separately or using a circular model,
-because we take all information, circular and linear, into account. In the next
-section we introduce several cylindrical models that can be used to analyze the
-teacher data. First however we will provide descriptives for the teacher data.
-
-\vspace{-0.75cm}
+intensity of the interpersonal behavior a teacher shows towards his/her
+students. In a cylindrical model this intensity (the linear component) can be
+modeled together with the location of interpersonal behavior of a teacher on the
+IPC (the circular component). This leads to an improved analysis of interpersonal
+circumplex data, over either analyzing the two dimensions separately or using a
+circular model, because we take all information, circular and linear, into
+account. In the next section we introduce several cylindrical models that can be
+used to analyze the teacher data. First however we will provide descriptives for
+the teacher data and conduct a \enquote{standard} analysis for these data that we can compare the results from the cylindrical models to.
+
 \subsection{Data Description}\label{DataDescriptives}
-\vspace{-0.75cm}
 
@@
 for multiple classes were available. This results in a sample of 151 teachers.
-In addition to the type of interpersonal behavior (IPC), the circular outcome,
-and the strength of interpersonal behavior (IPC strength), the linear outcome, a
-teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
-covariate in the analysis. In previous research, in psychology and education it
-has been shown that higher self-efficacy is related to the quality of
-interpersonal interactions (Locke \& Sadler, 2007; \VANDER{Want}{Van der}{van der} Want et al., 2018).
+The data includes the location of interpersonal behavior on the IPC (IPC), the
+circular component, and the intensity of interpersonal behavior (IPC intensity),
+the linear component. It also includes teachers' self-efficacy (\verb|SE|)
+concerning classroom management as covariate that is used to model the IPC and
+IPC intensity. This means that \(\boldsymbol{x}_i = \boldsymbol{z}_i = (1,\: \text{SE}_i)\) in all cylindrical models except for the Abe-Ley model where
+\(\boldsymbol{z}_i = \text{SE}_i\). In previous research, in psychology and
+education it has been shown that higher self-efficacy is related to the quality
+of interpersonal interactions (Locke \& Sadler, 2007; \VANDER{Want}{Van der}{van der} Want et al., 2018).
 After listwise deletion of missings (\(3\) in total, only for the self-efficacy)
-we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
-descriptives for the dataset. For the circular variable IPC we show sample
-estimates for the circular mean \(\bar{\theta}\) and concentration \(\hat{\rho}\).
-The circular concentration lies between 0, meaning the data is not concentrated
-at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
-concentrated at a single point on the circle. The population values of these
-parameters are usually, and also in this paper, referred to as \(\mu\) (circular
-location) and \(\kappa\) (circular concentration). For the linear variables
-(strength IPC and SE) we show sample estimates of the linear mean and standard
-deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
-the linear and circular outcome of the teacher data for teachers with low SE
+we have a sample of 148 teachers. Note that we remove the missings here for
+simplicity, however, in general and especially when the number of missings is
+larger, the influence of listwise deletion should be investigated and necessary
+precaution (e.g.~imputation of missing values) should be taken. Table
+\ref{Tableteacherdescriptives} shows descriptives for the dataset. Note that we
+also show the scores on the Agency and Communion component of the IPC before
+transformation to a circular score. For the circular variable IPC we show sample
+estimates for the circular mean \(\bar{\theta}\) and mean resultant length
+\(\hat{\rho}\). For the linear variables (Agency, Communion, IPC intensity and the
+covariate SE) we show sample estimates of the linear mean and standard deviation
+(sd). Figure \ref{dataplot} is a scatterplot showing the relation between the
+linear and circular component of the teacher data for teachers with low SE
 (below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
@@
 
-\vspace{-0.75cm}
-\section{Four Cylindrical Regression Models}\label{Models}
-\vspace{-0.75cm}
-
-One of the goals of this paper is to show the benefits of cylindrical methods
-for the analysis of circumplex data and cylindrical data in psychology in
-general. To do so we decide to focus on four cylindrical models. The models were
-selected for their relatively low complexity and the ease with which a
-regression structure could be incorporated. But also because they show different
-ways of modelling the linear and circular outcome and thereby illustrate a
-wider range of cylindrical models available in the literature. The
-cylindrical models contain a set of \(q\) predictors \(\boldsymbol{x} = x_1, \dots, x_q\) and \(p\) predictors \(\boldsymbol{z} = z_1, \dots, z_p\) for the linear and
-circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
-based on a construction by Mastrantonio et al. (2015), while the other models are
-extensions of the models from Abe and Ley (2017) and Mastrantonio (2018). The
-four cylindrical models are introduced separately in the subsections below.
-However, to provide a more succinct overview and comparison of the four models, Table
-\ref{TableModels} gives an overview of the similarities and differences between
-the models.
-
-\begin{sidewaystable}[h]
+\hfil [Table \ref{Tableteacherdescriptives} about here] \hfil
+
+\hfil [Figure \ref{dataplot} about here] \hfil
+
+\subsection{Standard Analysis}\label{StandardAnalysis}
+
+To perform a standard analysis of the teacher data we fit two linear regression
+models to the Agency and Communion scores of the teachers separately. In these
+regression models we incorporate an intercept and the covariate self-efficacy.
+Coefficients and standard errors from the models are shown in Table
+\ref{Tableteacherstandardanalysis}. From the results we conclude that the effect
+of self-efficacy is significant for both the Agency and Communion component. An
+increase of 1 unit on self-efficacy leads to a 0.07 increase in Agency and a
+0.09 increase in Communion. Note that in this setup we can only model the effect
+of self-efficacy on the intensity (size of the score) on the Agency and
+Communion component. We can not quantify the effect on the location and
+intensity on the IPC nor can we quantify their dependence.
+
+\hfil [Table \ref{Tableteacherstandardanalysis} about here] \hfil
+
+\begin{table}[h]
 \centering
-\caption{Comparison of the four cylindrical regression models} 
-\begin{tabular}{lllll}
+\caption{Regression coefficients and standard errors for the standard analysis of the teacher dataset.} 
+\begin{tabular}{lrr}
   \noalign{\smallskip}\hline\noalign{\smallskip}
-\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
-$\Theta$ & &&&\\
-$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
-$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
-$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
-            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
-$Y$& &&&\\
-$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
-$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
-$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
-            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
-$\Theta$-$Y$ dependence &                                   &                                   & & \\
-                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
-                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
-\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
-\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\
-
+  & Agency & Communion \\\hline\noalign{\smallskip}
+Intercept & 0.19 (0.01) & 0.30 (0.02) \\
+SE & 0.07 (0.01) & 0.09 (0.02) \\ \hline
+\multicolumn{3}{l}{Note: The intercepts indicate the average }\\
+\multicolumn{3}{l}{Agency and Communion. The coefficient for }\\
+\multicolumn{3}{l}{SE indicates the effect of self-efficacy }\\
+\multicolumn{3}{l}{on Agency and Communion.}
 \end{tabular}
-\label{TableModels}
-\end{sidewaystable}
-
-\vspace{-0.75cm}
-\subsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
-\vspace{-0.75cm}
-
-Following Mastrantonio et al. (2015) we consider two models where
-the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
-and \(q\) covariates is specified as
-\begin{equation}\label{circlinlink}
-Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
-\end{equation}
-\noindent where the random variable \(R\geq0\) will be introduced below, the
-error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
-\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
-intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
-covariates for the prediction of the linear outcome. We thus assume a normal
-distribution for the linear outcome.\newline
-\indent For the circular outcome we assume either a projected normal (PN) or a
-general projected normal (GPN) distribution. These distributions arise from a
-projection of a distribution defined in bivariate space onto the circle. Figure
-\ref{projection} represents this projection. In the left plot of Figure
-\ref{projection} we see datapoints from the bivariate variable
-\(\boldsymbol{S}\) that in the middle plot are projected to form the circular
-outcome \(\Theta\) in the right plot. Mathematically the relation between
-\(\boldsymbol{S}\) and \(\Theta\) is defined as follows
-\begin{equation}\label{projection}
-\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
-\end{equation}
-\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
-\(\boldsymbol{S}\); the lines connecting the bivariate datapoints to the origin
-in the middle plot. We call \(\boldsymbol{S}\) the augmented representation of the
-circular outcome. It is a variable that in contrast to \(\Theta\) is not
-observed and thus considered latent or auxiliary. This then means that we do not
-model \(\Theta\) directly but indirectly through \(\boldsymbol{S}\). \newline
-\indent For both the PN and GPN distributions the circular location parameter
-\(\mu\in [0, 2\pi)\) is modeled as \(\hat{\mu}_i = \mbox{atan2}(\hat{\mu}_i^{II}, \hat{\mu}_i^{I}) = \mbox{atan2}(\boldsymbol{\beta}^{II}\boldsymbol{z}_i, \boldsymbol{\beta}^{I}\boldsymbol{z}_i)\)\footnote{Note that for the CL-GPN
-model the circular location parameter also depends on the variance-covariance
-matrix and the circular predicted values should be computed using numerical
-integration or Monte Carlo methods because a closed form expression for the mean
-direction is not available.} where \(\boldsymbol{\beta}^{I}\) and
-\(\boldsymbol{\beta}^{II}\) are vectors with intercepts and regression
-coefficients for the prediction of \(\boldsymbol{S}\) and \(\boldsymbol{z}_i\) is a
-vector with predictor values for each individual \(i \in 1, \dots, n\) where \(n\)
-is the sample size. Note that as a result of the augmented representation of the
-circular outcome we have two sets of regression coefficients and intercepts, one
-for each bivariate component of \(\boldsymbol{S}\). This leads to problems when we
-want to interpret the effect of a covariate on the circle. A circular regression
-line is shown in Figure \ref{circregline}, with covariate values on the x-axis
-and the predicted circular outcome on the y-axis. As can be seen it is of a
-non-linear character meaning that the effect of a covariate is different at
-different values of the covariate. A circular regression line is usually
-described by the slope at the inflection point, the point at which the slope of
-the regression line starts flattening off (indicated with a square in Figure
-\ref{circregline}). By default, the parameters from the PN and GPN models do not
-directly describe this inflection point. For the PN distribution however,
-Cremers et al. (2018b) solved this interpretation problem and introduce new
-circular regression coefficients. They introduce a new parameter \(b_c\) that
-describes the slope at the inflection point of the regression line. For the
-GPN distrbution the interpretation problem however remains.\newline
-\indent The main difference between the PN and GPN distribution lies in the
-definition of their covariance matrix. For the PN distribution this is an
-identity matrix, causing the distribution to be unimodal and symmetric, whereas
-for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
-asymmetry/skewness.\newline
-\indent For the teacher dataset the predicted linear outcome, strengths of
-interpersonal behavior, in the CL-PN and CL-GPN model is the following:
-\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i +
-\gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i.\] The predicted circular
-outcome, type of interpersonal behavior, equals: \[\hat{\theta}_i = \mu_i =
-\mbox{atan2} (\beta_0^{II} + \beta_1^{II}\text{SE}_i, \beta_0^{I} +
-\beta_1^{I}\text{SE}_i).\] \noindent where \(\text{SE}_i\) is the self-efficacy
-score of one individual. The CL-PN and CL-GPN models thus allow us to assess the
-average type and strength of interpersonal behavior through the parameters
-\(\beta_{0}^{I}\), \(\beta_{0}^{II}\) and \(\gamma_0\) as well as the effect of
-self-efficacy on type and strength of teacher behavior through the parameters,
-\(\beta_{1}^{I}\), \(\beta_{1}^{II}\) and \(\gamma_1\). In addition, because the type
-and strength of interpersonal behavior are modelled together via the regression
-in \eqref{circlinlink} we can assess the effect of the type of interpersonal
-behavior on the strength through the parameters \(\gamma_{sin}\) and
-\(\gamma_{cos}\). In the teacher data these are the regression coefficients for
-the effect of the sine and cosine of the type of behavior which are related to
-the scores on the Agency and Communion dimensions respectively.\newline
-\indent Both the CL-PN and CL-GPN models are estimated using Markov Chain Monte
-Carlo (MCMC) methods based on Nuñez-Antonio, Gutiérrez-Peña, and Escarela (2011), Wang and Gelfand (2013) and
-Hernandez-Stumpfhauser, Breidt, and \VANDER{Woerd}{Van der}{van der} Woerd (2016) for the regression of the circular outcome. A detailed
-description of the Bayesian estimation and MCMC samplers can be found in the
-Supplementary Material.
-
-\begin{figure}
-\centering
-\includegraphics[width = \textwidth]{Plots/plotprojecting.pdf}
-\caption{Plot showing the projection of datapoints in bivariate space, $\boldsymbol{S}$, (left) to the circle (right).
-The lines connecting the bivariate datapoints to the circular datapoints represent the euclidean norm of the bivariate datapoints, the random variable $R$.}
-\label{projection}
-\end{figure}
-
-\begin{figure}[]
-  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
-  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to.
-The square indicates the inflection point of the regression line.}
-  \label{circregline}
-\end{figure}
-
-\vspace{-0.75cm}
-\subsection{The Modified Abe-Ley Model}\label{WeiSSVM}
-\vspace{-0.75cm}
-
-This model is an extension of the cylindrical model introduced in
-Abe and Ley (2017) to the regression context. It concerns a combination
-of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
-\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
-with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
-\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
-contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
-model defined only on the positive real half-line \([0, + \infty)\) and thus can
-not be negative.\newline
-\indent In this model we predict the linear scale parameter and circular
-location parameter, both of which we can express in terms of
-covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
-\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
-parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
-\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0, \dots, q\) and \(\gamma_0\) is the intercept.
-The parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
-regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
-\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
-vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
-is a vector of predictor values for the prediction of \(\theta\).\newline
-\indent For the teacher data, the predicted values for the circular outcome in
-the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
-\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
-The conditional distribution for the linear outcome is Weibull, meaning that we
-can use methods from survival analsis to interpret the effect of a predictor. In
-survival analysis a \enquote{survival} function is used in which time is plotted against
-the probability of survival of subjects suffering from a specific medical
-condition. In the teacher data we can thus compute the probability of a teacher
-having a specific strength on the IPC. This probability is computed using the
-\enquote{survival-function} defined as \[\exp(-\alpha
-y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
-with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
-function we also see that the circular concentration parameter \(\kappa\) and
-the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
-the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
-and strength of interpersonal behavior through the parameters \(\beta_{0}\) and
-\(\gamma_0\) as well as the effect of self-efficacy on type and strength of
-teacher behavior through the parameters \(\beta_{1}\) and \(\gamma_1\).\newline
-\indent We can use numerical optimization (Nelder-Mead) to find solutions for
-the maximum likelihood (ML) estimates for the parameters of the model.
-
-\vspace{-0.75cm}
-\subsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
-\vspace{-0.75cm}
-
-This model is an extension of the cylindrical model introduced by
-Mastrantonio (2018) to the regression context. The model contains \(m\)
-independent circular outcomes and \(w\) independent linear outcomes. The
-circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
-distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
-distribution (Sahu, Dey, \& Branco, 2003). Because the GPN distribution is modelled
-using a so-called augmented representation (see also the description of the
-CL-PN and CL-GPN models) it is convenient to use a similar tactic for
-modelling the multivariate skew normal distribution. As in
-Mastrantonio (2018), dependence between the linear and circular outcome is
-created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
-and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
-distribution.\newline
-\indent This means that we have a shared mean vector and variance-covariance
-matrix for the linear and circular outcome(s), much like having multiple
-outcomes in a MANOVA (multivariate analysis of variance) model. In our
-regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
-observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
-covariates.
-The mean vector then becomes \(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1) \times (2m + w)\) matrix with regression coefficients and intercepts and
-\(\boldsymbol{x}_i\) is a \(g + 1\) dimensional vector containing the value 1 to
-estimate an intercept and the \(g\) covariate values. This means that in contrast
-to the other three models, we have to use the same set of predictors for the
-circular and linear outcome.\newline
-\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix} \beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} & \beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\).
-The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
-linear outcomes in the GPN-SSN model are
-\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
-\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
-\beta_{1_{s^{I}}}\text{SE}_i),\]
-
-\noindent and
-
-\[\hat{y_i} = \beta_{0_{y_i}} +
-\beta_{1_{y_i}}\text{SE}_i.\]
-
-\noindent The GPN-SSN model thus allows us to assess the average type and
-strength of interpersonal behavior through the parameters \(\beta_{0_{s^{I}}}\),
-\(\beta_{0_{s^{II}}}\) and \(\beta_{0_{y}}\) as well as the effect of self-efficacy
-on type and strength of teacher behavior through the parameters
-\(\beta_{1_{s^{I}}}\), \(\beta_{1_{s^{II}}}\) and \(\beta_{1_{y}}\). In addition,
-because the type and strength of interpersonal behavior are modelled together
-using a multivariate normal distribution we can, through its variance-covariance
-matrix, also assess the dependence between the type and strength of interpersonal
-behavior. \newpage
-\indent We estimate the model using MCMC methods. A detailed description of these
-methods is given in the Supplementary Material.
-
-\vspace{-0.75cm}
-\subsection{Model Fit Criterion}\label{Modelfit}
-\vspace{-0.75cm}
-
-For the four cylindrical models we focus on their out-of-sample predictive
-performance to determine the fit of the model. To do so we use k-fold
-cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
-of the sample) is used once as a holdout set and 9 times as part of a training
-set. The analysis will thus be performed 10 times, each time on a different
-training set.\newline
-\indent A proper criterion to compare out-of-sample predictive performance is
-the Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007). The lower the
-value of this criterion, the better the predictive performance of the model.
-Using ML estimates this criterion can be computed as follows:
-\begin{equation}\label{PLSLML}
-PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
-\end{equation}
-\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
-holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
-\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
-Using posterior samples the criterion is similar to the log pointwise predictive
-density (lppd) (Gelman et al., 2014, p. 169) and can be computed as:
-\begin{equation}\label{PLSLBayes}
-PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
-\end{equation}
-\noindent where \(B\) is the amount of posterior samples and
-\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
-parameters for the \(j^{th}\) iteration. Because the joint density and thus also
-the likelihood for the modified GPN-SSN model is not
-available in closed form (Mastrantonio, 2018) we compute the PLSL for the
-circular and linear outcome separately for all models. Note that although we fit
-the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
-prior information into account when assessing model fit with the PLSL. According
-to Gelman et al. (2014) this is not necessary since we are assessing the fit of a model
-to data, the holdout set, only. They argue that the prior in such case is only
-of interest for estimating the parameters of the model but not for determining
-the predictive accuracy.\newline
-\indent For each of the four cylindrical models and for each of the 10
-cross-validation analyses we can then compute a PLSL for the circular and linear
-outcome by using the conditional log-likelihoods of the respective outcome (see
-Supplementary Material for a definition of the loglikelihoods). To evaluate the
-predictive performance we average across the PLSL criteria of the
-cross-validation analyses. We also assess the cross-validation variability by
-means of the standard deviations of the PLSL criteria.
-
-\vspace{-0.75cm}
+\label{Tableteacherstandardanalysis}
+\end{table}
+
 \section{Results}\label{DataAnalysis}
-\vspace{-0.75cm}
 
@@
 
-\vspace{-0.75cm}
 \subsection{Analysis}\label{DataResults}
-\vspace{-0.75cm}
 
@@
 
-\vspace{-0.5cm}
-\subsubsection{Average type and strength of interpersonal behavior}
+\subsubsection{Average location and intensity on the IPC}
 
 The parameters \(\gamma_0\) in the CL-PN, CL-GPN and Abe-Ley model and the
-parameter \(\beta_{0_y}\) in the GPN-SSN model inform us about the strength of
+parameter \(\beta_{0_y}\) in the GPN-SSN model inform us about the intensity of
 interpersonal behavior at the average self-efficacy. For the CL-PN, CL-GPN and
 GPN-SSN models the parameters are estimated at 0.38, 0.37 and 0.33 respectively
-and are a direct prediction of the strength of interpersonal behavior at the
+and are a direct prediction of the intensity of interpersonal behavior at the
 average self-efficacy. The estimate for the GPN-SSN model is notably lower and
-likely to be caused by its skewed distribution for the strengths of
+likely to be caused by its skewed distribution for the intensity of
 interpersonal behavior. In the Abe-Ley model, \(\gamma_0\) influences the shape
-parameter of the distribution of the strength of interpersonal behavior and does
-not directly estimate the average strength. Instead we can use the survival
-function to say something about the probability of having a certain strength of
+parameter of the distribution of the intensity of interpersonal behavior and does
+not directly estimate the average intensity. Instead we can use the survival
+function to say something about the probability of having a certain intensity of
 interpersonal behavior. Figure \ref{reglineweib} shows this function for several
@@
 for observations that fall within 1 standard deviation of the mean. The survival
-function indicates that the probability of having a low strength of
-interpersonal behavior is higher than having a high strength. We however can not
-make any direct statement about the estimated strength using the Abe-Ley
-model.\newline
+function indicates that the probability of having a low intensity of
+interpersonal behavior is higher than having a high intensity. We however can not
+make any direct statement about the estimated intensity using the Abe-Ley
+model.
+
+\hfil [Figure \ref{reglineweib} about here] \hfil
+
 \indent The parameters \(\beta_0^{I}\), \(\beta_0^{II}\), \(\beta_0\),
-\(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) inform us about the type of
-interpersonal behavior at the average self-efficacy for the CL-PN, CL-GPN,
+\(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) inform us about the location on the IPC at the average self-efficacy for the CL-PN, CL-GPN,
 Abe-Ley and GPN-SSN model respectively. For the CL-PN, CL-GPN and GPN-SSN model
@@
 three models at 32.29\(^\circ\), 33.70\(^\circ\) and 35.53\(^\circ\). In the Abe-Ley
-model the type of interpersonal behavior at the average self-efficacy is
+model the location on the IPC at the average self-efficacy is
 estimated at 0.36 radians or 20.63\(^\circ\).
+
+\hfil [Table \ref{tab:means} about here] \hfil
 
@@
 
-\vspace{-0.5cm}
 \subsubsection{The effect of self-efficacy}
@@
 The parameters \(\gamma_1\) in the CL-PN, CL-GPN, Abe-Ley models and \(\beta_{1_y}\)
-in the GPN-SSN model inform us about the effect of self-efficacy on the strength
+in the GPN-SSN model inform us about the effect of self-efficacy on the intensity
 of interpersonal behavior. For the CL-PN, CL-GPN and GPN-SSN model the
 parameters are estimated at 0.03, 0.03 and 0.09 respectively and are a direct
-estimate of the effect of self-efficacy on the strength of interpersonal
+estimate of the effect of self-efficacy on the intensity of interpersonal
 behavior, \emph{i.e.} an increase of 1 unit in self-efficacy leads to an
-increase of 0.09 units in the strength of interpersonal behavior according to
+increase of 0.09 units in the intensity of interpersonal behavior according to
 the GPN-SSN model. These estimates are however quite small and only different
@@
 say which of the three models, CL-PN, CL-GPN or GPN-SSN, to use to base our
-conclusions on. The models CL-GPN and CL-PN fit the linear outcome best according
-to the model fit in Table \ref{tab:ModelFit}. In these models the linear outcome has a
+conclusions on. The models CL-GPN and CL-PN fit the linear component best according
+to the model fit in Table \ref{tab:ModelFit}. In these models the linear component has a
 symmetric distribution whereas in the GPN-SSN the distribution of the linear
-outcome is skewed. However, the effect of self-efficacy is different from 0 only
-in the GPN-SSN model which does not seem to match with its lower model fit.\newline
+component is skewed. However, the effect of self-efficacy is different from 0 only
+in the GPN-SSN model which does not seem to match with its lower model fit.
+
+\hfil [Table \ref{tab:ModelFit} about here] \hfil
+
 \indent In the Abe-Ley model, \(\gamma_1\) influences the shape parameter of the distribution of
-the strength of interpersonal behavior and does not directly estimate the effect
+the intensity of interpersonal behavior and does not directly estimate the effect
 of self-efficacy. Instead we can use the survival function to say something
-about the probability of having a certain strength of interpersonal behavior for
+about the probability of having a certain intensity of interpersonal behavior for
 different values of self-efficacy. Figure \ref{reglineweib} shows this function
@@
 \(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) inform us about the effect of
-self-efficacy on the type of interpersonal behavior in the CL-PN, CL-GPN,
+self-efficacy on the location on the IPC in the CL-PN, CL-GPN,
 Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley models we have
@@
 \end{figure}
-\vspace{-0.5cm}
-\subsubsection{Dependence between type and strength of interpersonal behavior.}
-
-The relation between the type and strength of interpersonal behavior in the
+
+\subsubsection{Dependence between location and intensity on the IPC}
+
+The relation between the location and intensity on the IPC in the
 CL-PN and CL-GPN models is described by the parameters \(\gamma_{\cos}\) and
 \(\gamma_{\sin}\). The HPD interval of \(\gamma_{\cos}\) does not include 0 for
-both the CL-PN and CL-GPN models, meaning that the cosine component of the type
-of interpersonal behavior has an effect on the strength of interpersonal
-behavior.\newline
+both the CL-PN and CL-GPN models, meaning that the cosine component of the location has an effect on the intensity.\newline
 \indent In the teacher data the sine and cosine components have a substantive
@@
 (Agency) represents the sine of an angle. For the teacher data this means that
-the Communion (cosine) dimension of the IPC positively effects the strength of a
-teachers' type of interpersonal behavior, in plain words: teachers exhibiting
+the Communion (cosine) dimension of the IPC positively effects the intensity of a
+teachers' interpersonal behavior, in plain words: teachers exhibiting
 interpersonal behavior types with higher communion scores (e.g., \enquote{helpful} and
 \enquote{understanding} in Figure 2) are stronger in their interpersonal behavior.\newline 
-\indent In the GPN-SSN model the dependence between the type and strengths of
-interpersonal behavior is modelled through the covariances between the linear
-outcome and the sine and cosine of the circular outcome \(\sum_{sy_{2,3}}\) and
+\indent In the GPN-SSN model the dependence between the location and intensity on the IPC is modelled through the covariances between the linear
+component and the sine and cosine of the circular component \(\sum_{sy_{2,3}}\) and
 \(\sum_{sy_{1,3}}\). Both covariances, \(\sum_{sy_{2,3}} = 0.09\) and
@@
 behavior. Together with the results from the CL-PN and CL-GPN models in the
-previous paragraph this translates to the conclusion that teachers with the
-strongest interpersonal behavior have a type of interpersonal behavior between
+previous paragraph this translates to the conclusion that teachers with higher
+intensity scores have a location on the IPC that lies between
 0\(^\circ\) and 90\(^\circ\). To get these scores on the circle both the Agency and
@@
 corresponds to the pattern observed in the teacher data in Figure \ref{QTI}. At
-a strength of 0.4 and up we see that the scores on the circle range on average
+an intensity of 0.4 and up we see that the scores on the circle range on average
 between 0\(^\circ\) and 100\(^\circ\).
 
-\vspace{-0.5cm}
 \subsubsection{Model fit}
 
-Table \ref{tab:ModelFit} shows the values fof the PLSL criterion for the linear
-and circular outcomes of the four cylindrical models fit to the teacher
+Table \ref{tab:ModelFit} shows the values of the PLSL criterion for the linear
+and circular components of the four cylindrical models fit to the teacher
 data.\newline \indent The CL-PN and CL-GPN models have the best out-of-sample
-predictive performance for the linear outcome. They show roughly the same
-performance because they model the linear outcome in the same way. We should
+predictive performance for the linear component. They show roughly the same
+performance because they model the linear component in the same way. We should
 note that even though the predictive performance of the Abe-Ley model for the
-linear outcome is worst on average, the standard deviation of the
+linear component is worst on average, the standard deviation of the
 cross-validation estimates is rather large. This means that in some samples, the
@@
 \indent The Abe-Ley model has the best out-of-sample predictive performance for
-the circular outcome. This would suggest that for the circular variable a
+the circular component. This would suggest that for the circular variable a
 slightly skewed distribution fits best. However, both the GPN-SSN and the
 CL-GPN models fit much worse even though the distribution for the circular
-outcome in these models can also take a skewed shape. It should be noted that
+component in these models can also take a skewed shape. It should be noted that
 the standard deviation of the cross-validation estimates is rather large for
@@
 
-\vspace{-0.75cm}
 \section{Discussion}\label{Discussion}
-\vspace{-0.75cm}
 
 In this paper we modified four models for cylindrical data in such a way that
-they include a regression of both the linear and circular outcome onto a set of
+they include a regression of both the linear and circular component onto a set of
 covariates. Subsequently we have shown how these four methods can be used to
@@
 \indent Concerning the teacher data, the advantage of cylindrical data analysis
-is that we were able to analyze the information about the type and strength of
-interpersonal behavior simultaneously. In previous research, the two components
-of the interpersonal circumplex (\emph{i.e.}, Agency and Communion) were
-analyzed separately. Such an approach also provides information about the
-strength of teachers' score on Agency and Communion, yet a large portion of
-information about the combination of Agency and Communion, which describes the
-type of behavior that is observed, gets lost. A first solution to include both
-dimensions as a circular variable in data analysis was described by
+is that we were able to analyze the information about the location and intensity
+on the IPC simultaneously. In previous research, the two components of the
+interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
+separately. Such an approach only provides information about the intensity of a
+teachers' score on Agency and Communion and the information about the
+combination of Agency and Communion, which describes the location on the IPC,
+gets lost as was observed in the Standard Analysis section. A first solution to
+include both dimensions as a circular variable in data analysis was described by
 Cremers et al. (2018a). A downside of that analysis was that information about
-the strength of the specific type of interpersonal behavior could not be
-retained. In the present study, we have shown how using cylindrical models can
-simultaneously model the information about the type of and strength of
-interpersonal behavior and how these are influenced by teachers' self-efficacy
-in classroom management. Although we do not find any strong effects of
-self-efficacy on either the type or strength of behavior, the four cylindrical
-models do provide a way of analyzing and interpreting this effect. This is
-beneficial for future research in which we may want to investigate the effect of
-further covariates on data from the circumplex. \newline
+the intensity could not be retained. In the present study, we have shown how
+using cylindrical models can simultaneously model the information about the
+location and intensity on the IPC and how these are influenced by teachers'
+self-efficacy in classroom management. Although we do not find any strong
+effects of self-efficacy on either the location and intensity\footnote{similar
+to the standard analysis}, the four cylindrical models do provide a way of
+analyzing and interpreting this effect. This is beneficial for future research
+in which we may want to investigate the effect of further covariates on data
+from the circumplex. \newline
 \indent Furthermore, in addition to being able to assess the influence of
 covariates, the cylindrical models also provide information about the dependence
-between the type and strength of interpersonal behavior. We found that stronger
-behavior is associated with higher scores on the Communion and in some models
-also the Agency dimension. This implies that teachers whose type of
-interpersonal behavior ranges between 0\(^\circ\) and 90\(^\circ\), the \enquote{helpful}
-and \enquote{directing} subtypes are stronger in their behavior than teachers of the
-other subtypes.\newline
+between the location and intensity on the IPC. This is a second advantage over
+the standard analysis in which it is impossible to quantify the relation between
+intensity and location. We found that stronger behavior is associated with
+higher scores on the Communion and in some models also the Agency dimension.
+This implies that teachers whose interpersonal behavior ranges between
+0\(^\circ\) and 90\(^\circ\) on the IPC, the \enquote{helpful} and \enquote{directing} subtypes, are
+stronger in their behavior than teachers of the other subtypes.\newline
 \indent As mentioned in the introduction, data from the interpersonal circumplex
@@
 the models is of use in these fields as well. \newline
-\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
-best out of the four cylindrical models. In the
-CL-GPN and GPN-SSN models the interpretation of the parameters of the
-circular outcome component is not straightforward, if at all possible.
-This is caused by the fact that in addition to the mean vector the
-covariance matrix of the GPN distribution affects the location of the
-circular data, making it difficult to compute regression coefficients on
-the circle. Wang and Gelfand (2013) state that Monte Carlo integration
-can be used to compute a circular mean and variance for the GPN
-distribution. In future research, this solution might be applied to the
-methods of Cremers et al. (2018b) in order to compute circular
-coefficients for GPN models.\newline
-\indent In terms of flexibility the GPN-SSN model scores best. Multiple
-linear and circular outcomes can be included and we can thus apply the
-model to multivariate cylindrical data. In addition the GPN-SSN, the
-CL-GPN and CL-PN models are extendable to a mixed-effects structure and
-can thus also be fit to longitudinal data (see Nuñez-Antonio and Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al. (2016) for
+\indent Out of the four cylindrical models investigated in this paper, the
+results from CL-PN and Abe-Ley models are most straightforward to interpret. In
+the CL-GPN and GPN-SSN models the interpretation of the parameters of the
+circular outcome component is more complex, if at all possible. This is caused
+by the fact that in addition to the mean vector the covariance matrix of the GPN
+distribution affects the location of the circular data, making it difficult to
+compute regression coefficients on the circle. Wang and Gelfand (2013) state that
+Monte Carlo integration can be used to compute a circular mean and variance for
+the GPN distribution. In future research, this solution might be applied to the
+methods of Cremers et al. (2018b) in order to compute circular coefficients
+for GPN models.\newline
+\indent The GPN-SSN model is more flexible compared to the other three
+cylindrical models that were investigated in this paper. Multiple linear and
+circular components can be included and we can thus apply the model to
+multivariate cylindrical data. In addition the GPN-SSN, the CL-GPN and CL-PN
+models are extendable to a mixed-effects structure and can thus also be fit to
+longitudinal data (see Nuñez-Antonio and Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al. (2016) for
 hierarchical/mixed-effects models for the PN and GPN distributions
-respectively). For the Abe-Ley model this may also be possible but has
-not been done in previous research for the conditional distribution of
-its circular outcome (sine-skewed von
-Mises). Concerning asymmetry, both the
-GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
-the distributions of both the linear and circular outcome, while the CL-GPN
-model permits an asymmetric circular outcome.\newline
+respectively). For the Abe-Ley model this may also be possible but has not been
+done in previous research for the conditional distribution of its circular
+component (sine-skewed von Mises). Concerning asymmetry, both the GPN-SSN as
+well as the Abe-Ley model allow for non-symmetrical shapes of the distributions
+of both the linear and circular component, while the CL-GPN model permits an
+asymmetric circular component.\newline
 \indent The four cylindrical models that were modified to the regression
@@
 \newpage
-\vspace{-0.75cm}
 \section*{{\normalfont References}}
-\vspace{-0.75cm}
-
-\DeclareRobustCommand{\VANDER}[3]{#3}
-\DeclareRobustCommand{\VAN}[3]{#3}
-\DeclareRobustCommand{\DEN}[3]{#3}
 

### 2019-07-26 Manuscript/Manuscript R1/Manuscript.Rmd
Update old version

--- 2019-01-29 Manuscript/ManuscriptMarkdownCV.Rmd
@@
 cylindrical. For example, research on human navigation uses data where distance
-(i.e., linear component) and direction (i.e., the linear component) are of
+(i.e., the linear component) and direction (i.e., the circular component) are of
 interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
 consist of both the direction (i.e., the circular variable) and the duration
-(i.e., the linear variable)(e.g., @rayner200935th). Apart from the social
-sciences data with a circular and linear outcome more commonly occur in
+(i.e., the linear variable) (e.g., @rayner200935th). Apart from the social
+sciences, data with a circular and linear outcome more commonly occur in
 meteorology [@garcia2013exploring], ecology [@garcia2014test] or
@@
 of covariates. From now on we will therefore refer to the components of the
-cylindrical variable as outcome components. These adapted circular models are
+cylindrical variable as outcome components. These adapted cylindrical models are
 then used  to analyse the teacher data.
 
-\subsection{Cylindrical data}
-
-Data that consist of a variable and a circular variable is called
+\vspace{-0.75cm}
+\subsection{Cylindrical Data}
+\vspace{-0.75cm}
+Data that consist of a linear variable and a circular variable is called
 cylindrical data. A circular variable is different from the linear variable in
@@
 for the GPN distribution \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2
-& \xi\\ \xi & 1 \end{bmatrix}\), allowing for multimodality and
-assymetry/skewness.\newline
+& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
+asymmetry/skewness.\newline
 \indent For the teacher dataset the predicted linear outcome, strengths of
@@
 $\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
-on type and strength of teacher behavior through the parameters,
+on type and strength of teacher behavior through the parameters
 $\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
 because the type and strength of interpersonal behavior are modelled together
-using a multivariate normal distribution we can through its variance-covariance
-matrix also assess the dependence between the type and strength of interpersonal
-behavior. \newline
+using a multivariate normal distribution we can, through its variance-covariance
+matrix, also assess the dependence between the type and strength of interpersonal
+behavior. \newpage
 \indent We estimate the model using MCMC methods. A detailed description of these
@@
 
+\vspace{-0.5cm}
 \subsubsection{The effect of self-efficacy}
-
-The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley model and $\beta_{1_y}$
+The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley models and $\beta_{1_y}$
 in the GPN-SSN model inform us about the effect of self-efficacy on the strength
@@
 \end{figure}
-
+\vspace{-0.5cm}
 \subsubsection{Dependence between type and strength of interpersonal behavior.}
-
 The relation between the type and strength of interpersonal behavior in the
-CL-PN and CL-GPN model, is described by the parameters $\gamma_{\cos}$ and
+CL-PN and CL-GPN models is described by the parameters $\gamma_{\cos}$ and
 $\gamma_{\sin}$. The HPD interval of \(\gamma_{\cos}\) does not include 0 for

### 2019-07-26 Manuscript/Manuscript R1/SupplementaryMaterial.Rmd
Update old version

--- 2019-03-03 Manuscript/SupplementaryMaterial.Rmd
@@
   - \usepackage{setspace}\doublespacing
+  - \DeclareRobustCommand{\VANDER}[3]{#2} 
+  - \DeclareRobustCommand{\VAN}[3]{#2}
+  - \DeclareRobustCommand{\DEN}[3]{#2}
 documentclass: article

### 2019-07-26 Manuscript/SupplementaryMaterial.Rmd
Supplementary Material Update
Update supplementary material

@@
 R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
-Note that the dimensions of the objects (design matrices, mean vectors, etc.) are those that were used in the analysis of the teacher data where we have 1 circular outcome, 1 linear outcome and estimate an intercept and regression coefficient for the covariate self-efficacy.
-Note that for the regression of the linear component in the CL-PN and CL-GPN models we also have the sine and cosine of the circular outcome in the regression equation, this makes the vector with regression coefficients, $\boldsymbol{\gamma}$, four-dimensional.
+Note that the dimensions of the objects (design matrices, mean vectors, etc.) are those that were used in the analysis of the teacher data where we have 1 circular component, 1 linear component and estimate an intercept and regression coefficient for the covariate self-efficacy.
+Note that for the regression of the linear component in the CL-PN and CL-GPN models we also have the sine and cosine of the circular component in the regression equation, this makes the vector with regression coefficients, $\boldsymbol{\gamma}$, four-dimensional.
 
@@
 Following @mastrantonio2015bayesian we consider in this section two models where
-the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
-and $q$ covariates is specified as
-\begin{equation}\label{circlinlink}
-Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
+the prediction equation for the linear component is specified as
+\begin{equation}\label{linpredCLPNCLGPN}
+\hat{y_i} = \gamma_0 + \gamma_{cos}*\cos(\theta_i)*r_i + \gamma_{sin}*\sin(\theta_i)*r_i + \gamma_1*x_1 + \dots + \gamma_q*x_q,
 \end{equation}
-\noindent where the random variable \(R\geq0\) will be introduced below, the
-error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
+\noindent where $r_i$ is a realization of the unobserved the random variable
+\(R\geq0\) that will be introduced below,
 \(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
@@
 \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
-\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
+\(r \geq 0\). The linear component thus has a normal distribution conditional on
 \(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
 in its location part.\newline
-\indent For the circular outcome we assume either a projected normal (PN) or a general
+\indent For the circular component we assume either a projected normal (PN) or a general
 projected normal (GPN) distribution. These distributions arise from the radial
@@
 between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
-outcome \(\Theta\) is defined as follows
+component \(\Theta\) is defined as follows
 \begin{equation}\label{projection}
@@
 \begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\mu^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(u^t\mu) \right]\},
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\boldsymbol{\mu}^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(\boldsymbol{u}^t\boldsymbol{\mu}) \right]\},
 \end{equation}
 \noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
-\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
-\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
-the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
+\end{bmatrix}\) and $r$ is defined in \eqref{projection}. In a regression setup
+the outcome components \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
 \(n\) is the sample size, are generated independently from the distribution with
@@
 \noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
-\xi^2 & \xi\\ \xi & 1 \end{bmatrix}\). In a regression setup the outcomes
+\xi^2 & \xi\\ \xi & 1 \end{bmatrix}\). In a regression setup the outcome components
 \(\theta_i\) and \(r_i\) for each individual are generated independently from
@@
 Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
-@hernandez2016general for the regression of the circular outcome.
+@hernandez2016general for the regression of the circular component.
 
@@
 f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
-                 (1 +\lambda\sin(\theta - \mu))
+                 (1 +\lambda\sin(\theta - \mu_c))
                  y^{\alpha-1}
-                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu_c))],
 \end{equation}
@@
 occurs at the level of the linear scale parameter \(\nu>0\) and circular
-location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
+location parameter \(\mu_c\in [0, 2\pi)\), both of which we express in terms of
 covariates: \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
-\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
+\(\mu_{c,i} = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
 parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
@@
 is a vector of predictor values for the prediction of \(\theta\). In a
-regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
+regression setup the outcome component vector \((\theta_i, y_i)^t\) for each individual is
 generated independently from the modified
@@
 \(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
-\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
+\(\nu(1-\tanh(\kappa)\cos(\theta - \mu_c))^{1/\alpha}\) and the conditional
 distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
-with location parameter \(\mu\) and concentration parameter \((\nu
+with location parameter \(\mu_c\) and concentration parameter \((\nu
 y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
@@
 @mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
-  outcomes and \(w\) independent linear outcomes. The circular outcomes
+  components and \(w\) independent linear components. The circular components
   \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
@@
   distribution.
-The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes.
-The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
+The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular components.
+The linear components \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
   \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
@@
   multivariate skew normal distribution. Following @mastrantonio2018joint the
-  linear outcomes are represented as
+  linear components are represented as
 \[\boldsymbol{Y} = \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
-\noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear outcome
+\noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear component
 \(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
@@
 \(\boldsymbol{M}_\ell\) and
-covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
+covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular component is created by
 modelling the augmented representations of \(\boldsymbol{\Theta}\) and
@@
 The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
-of and covariances between the augmented representations of the circular outcome
+of and covariances between the augmented representations of the circular component
 and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
-augmented representations of the circular outcome and the linear outcome.
+augmented representations of the circular component and the linear component.
 \newline
 \indent In our regression extension we have \(i = 1, \dots, n\) observations of
-\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
+\(m\) circular components, \(w\) linear components and \(g\) covariates. The mean in
 the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{M}_i =
@@
 
-$$l(\theta, r) = \log(1) - \log(2\pi) + \sum -0.5\hat{\mu}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\mu}_i)$$
-where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
+$$l(\theta, r) = \log(1) - \log(2\pi) + \sum -0.5\hat{\boldsymbol{\mu}}_i^2 - 0.5(r_i^2 - 2r_iu_i^t\hat{\boldsymbol{\mu}}_i)$$
+where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\boldsymbol{\mu}}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
 
@@
 
-$$l(\theta) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
-where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
+$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\boldsymbol{\mu}}_i\Sigma^{-1}(u_i^t\hat{\boldsymbol{\mu}}_i)^t)/2\tau^2$$
+where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\boldsymbol{\mu}}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
 
@@
 where $SSN$ is the skew normal distribution.
-Computationally this comes down to taking the log of the density values for a univariate and multivariate normal distribution (with mean and variance specified as above) for the linear and circular outcome respectively.
+Computationally this comes down to taking the log of the density values for a univariate and multivariate normal distribution (with mean and variance specified as above) for the linear and circular component respectively.
 
@@
 \begin{enumerate}
-\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k_{n \times 2}$ (for $k \in \{I,II\}$) and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set.
+\item Split the data, with the circular component $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear component $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k_{n \times 2}$ (for $k \in \{I,II\}$) and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear component respectively, in a training (90\%) and holdout (10\%) set.
 \item Define the prior parameters for the training set. In this paper we use:
@@
 Also set starting values $r_i = 1$ in the training and holdout set.
-\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
+\item Compute the latent bivariate scores $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular component for the holdout and training dataset as follows:
 $$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i)\\  r_i\sin (\theta_i)\end{bmatrix}.$$
@@
 \end{itemize}
-\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset.
+\item Compute the PLSL for the circular and linear component on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset.
 \item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We assess convergence visually using traceplots.
@@
 \begin{enumerate}
-\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}_{n \times 2}$  and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set.
+\item Split the data, with the circular component $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear component $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}_{n \times 2}$  and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear component respectively, in a training (90\%) and holdout (10\%) set.
 \item Define the prior parameters for the training set. In this paper we use:
@@
 Also set starting values $r_i = 1$ in the training and holdout set.
-\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
+\item Compute the latent bivariate scores $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular component for the holdout and training dataset as follows:
 $$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
@@
 
-\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\xi$ and $\tau$ for the training dataset.
+\item Compute the PLSL for the circular and linear component on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\xi$ and $\tau$ for the training dataset.
 \item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We visually assess convergence using traceplots.
@@
 \begin{enumerate}
-\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set.
-Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation.
-It can however be generalized to a situation with multiple circular and linear outcomes without too much effort.
+\item Split the data, with the circular component $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear component $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set.
+Note that in this paper we have only one circular component and one linear component and the MCMC procedure outlined here is specified for this situation.
+It can however be generalized to a situation with multiple circular and linear components without too much effort.
 \item Define the prior parameters for the training set.
-Since we have only one circular outcome, one linear outcome and one covariate, we have $m = 1$, $w = 1$ and $g = 1$.
+Since we have only one circular component, one linear component and one covariate, we have $m = 1$, $w = 1$ and $g = 1$.
 In this paper we use the following priors:
@@
 Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set.
-\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
+\item Compute the latent bivariate scores $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular component for the holdout and training dataset as follows:
 $$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
-\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:
+\item Compute the latent scores $\tilde{y}_i$ underlying the linear component for the holdout and training dataset as follows:
 $$\tilde{y}_i = \lambda d_i. $$
@@
 The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i^t, y_i)^t \sim N_{2m+w}(\boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, \lambda d_i)^t, \boldsymbol{\Sigma})$.
-Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can in this model (where the variance-covariance matrix of the circular outcome is not constrained in the estimation procedure) simply define the $r_i$ as the Euclidean norm of the bivariate datapoints.
+Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can in this model (where the variance-covariance matrix of the circular component is not constrained in the estimation procedure) simply define the $r_i$ as the Euclidean norm of the bivariate datapoints.
 However, for didactic purposes we continue with the explanation of the sampling procedure.
@@
 
-\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.
+\item Compute the PLSL for the circular and linear component on the holdout set using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.
 

### 2019-07-26 Manuscript/SupplementaryMaterial.tex
Update supplementary material

@@
 etc.) are those that were used in the analysis of the teacher data where
-we have 1 circular outcome, 1 linear outcome and estimate an intercept
-and regression coefficient for the covariate self-efficacy. Note that
-for the regression of the linear component in the CL-PN and CL-GPN
-models we also have the sine and cosine of the circular outcome in the
-regression equation, this makes the vector with regression coefficients,
-\(\boldsymbol{\gamma}\), four-dimensional.
+we have 1 circular component, 1 linear component and estimate an
+intercept and regression coefficient for the covariate self-efficacy.
+Note that for the regression of the linear component in the CL-PN and
+CL-GPN models we also have the sine and cosine of the circular component
+in the regression equation, this makes the vector with regression
+coefficients, \(\boldsymbol{\gamma}\), four-dimensional.
 
@@
 Following Mastrantonio, Maruotti, \& Jona-Lasinio (2015) we consider in
-this section two models where the relation between
-\(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\) and \(q\)
-covariates is specified as \begin{equation}\label{circlinlink}
-Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
-\end{equation} \noindent where the random variable \(R\geq0\) will be
-introduced below, the error term \(\epsilon \sim N(0, \sigma^2)\) with
-variance \(\sigma^2>0\),
+this section two models where the prediction equation for the linear
+component is specified as \begin{equation}\label{linpredCLPNCLGPN}
+\hat{y_i} = \gamma_0 + \gamma_{cos}*\cos(\theta_i)*r_i + \gamma_{sin}*\sin(\theta_i)*r_i + \gamma_1*x_1 + \dots + \gamma_q*x_q,
+\end{equation} \noindent where \(r_i\) is a realization of the
+unobserved the random variable \(R\geq0\) that will be introduced below,
 \(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
@@
 \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
-\(r \geq 0\). The linear outcome thus has a normal distribution
+\(r \geq 0\). The linear component thus has a normal distribution
 conditional on \(\Theta\) and \(R\) and contains already linear
 covariates \(x_1, \dots, x_q\) in its location part.\newline \indent For
-the circular outcome we assume either a projected normal (PN) or a
+the circular component we assume either a projected normal (PN) or a
 general projected normal (GPN) distribution. These distributions arise
@@
 the circle. The relation between a bivariate vector \(\boldsymbol{S}\)
-in the plane and the circular outcome \(\Theta\) is defined as follows
+in the plane and the circular component \(\Theta\) is defined as follows
 \begin{equation}\label{projection}
@@
 \begin{equation}\label{pnreg}
-f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\mu^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(u^t\mu) \right]\},
+f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\boldsymbol{\mu}^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(\boldsymbol{u}^t\boldsymbol{\mu}) \right]\},
 \end{equation} \noindent where
 \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
-and \(r\) is the same as in \eqref{circlinlink} and \eqref{ycondtheta}
-and is defined in \eqref{projection}. In a regression setup the outcomes
-\(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where \(n\) is
-the sample size, are generated independently from the distribution with
-density \eqref{pnreg}. The mean vector
-\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
+and \(r\) is defined in \eqref{projection}. In a regression setup the
+outcome components \(\theta_i,r_i\) for each individual
+\(i = 1, \dots, n\), where \(n\) is the sample size, are generated
+independently from the distribution with density \eqref{pnreg}. The mean
+vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
 \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
@@
 \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\).
-In a regression setup the outcomes \(\theta_i\) and \(r_i\) for each
-individual are generated independently from \eqref{gpnreg}. The mean
-vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the same
-way via covariates as for the modified CL-PN distribution.
+In a regression setup the outcome components \(\theta_i\) and \(r_i\)
+for each individual are generated independently from \eqref{gpnreg}. The
+mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the
+same way via covariates as for the modified CL-PN distribution.
 
@@
 Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
-regression of the circular outcome.
+regression of the circular component.
 
@@
 f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
-                 (1 +\lambda\sin(\theta - \mu))
+                 (1 +\lambda\sin(\theta - \mu_c))
                  y^{\alpha-1}
-                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
+                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu_c))],
 \end{equation} \noindent where \(\alpha > 0\) is a linear shape
@@
 the linear scale parameter \(\nu>0\) and circular location parameter
-\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
+\(\mu_c\in [0, 2\pi)\), both of which we express in terms of covariates:
 \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
-\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
+\(\mu_{c,i} = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
 The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
@@
 is a vector of predictor values for the prediction of \(\theta\). In a
-regression setup the outcome vector \((\theta_i, y_i)^t\) for each
-individual is generated independently from the modified density
+regression setup the outcome component vector \((\theta_i, y_i)^t\) for
+each individual is generated independently from the modified density
 \eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
@@
 distribution with shape \(\alpha\) and scale
-\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the
+\(\nu(1-\tanh(\kappa)\cos(\theta - \mu_c))^{1/\alpha}\) and the
 conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
-von Mises distribution with location parameter \(\mu\) and concentration
-parameter \((\nu y)^\alpha\tanh(\kappa)\). The log-likelihood for this
-model equals \begin{align}\label{WeiSSVMLikelihood}
+von Mises distribution with location parameter \(\mu_c\) and
+concentration parameter \((\nu y)^\alpha\tanh(\kappa)\). The
+log-likelihood for this model equals
+\begin{align}\label{WeiSSVMLikelihood}
 l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
@@
 Mastrantonio (2018) to the regression context. Both models contain \(m\)
-independent circular outcomes and \(w\) independent linear outcomes. The
-circular outcomes
+independent circular components and \(w\) independent linear components.
+The circular components
 \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,  \boldsymbol{\Theta}_m)\)
@@
 be modeled as the product of (\ref{gpnreg}) for each of the \(m\)
-circular outcomes. The linear outcomes
+circular components. The linear components
 \(\boldsymbol{Y} = (\boldsymbol{Y}_1,  \dots, \boldsymbol{Y}_w)\) are
@@
 the multivariate skew normal distribution. Following Mastrantonio (2018)
-the linear outcomes are represented as
+the linear components are represented as
 \[\boldsymbol{Y} = \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
 \noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear
-outcome \(\boldsymbol{Y}\),
+component \(\boldsymbol{Y}\),
 \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
@@
 in Mastrantonio (2018) dependence between the linear and circular
-outcome is created by modelling the augmented representations of
+component is created by modelling the augmented representations of
 \(\boldsymbol{\Theta}\) and \(\boldsymbol{Y}\) together in a \(2m + w\)
@@
 variances of and covariances between the augmented representations of
-the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
+the circular component and the matrix \(\boldsymbol{\Sigma}_{sy}\)
 contains covariances between the augmented representations of the
-circular outcome and the linear outcome. \newline \indent In our
+circular component and the linear component. \newline \indent In our
 regression extension we have \(i = 1, \dots, n\) observations of \(m\)
-circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean
-in the density in \eqref{YDThetarjoint} then becomes
+circular components, \(w\) linear components and \(g\) covariates. The
+mean in the density in \eqref{YDThetarjoint} then becomes
 \(\boldsymbol{M}_i = \boldsymbol{B}^t\boldsymbol{x}_i\) where
@@
 
-$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
-where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
+$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\boldsymbol{\mu}}_i\Sigma^{-1}(u_i^t\hat{\boldsymbol{\mu}}_i)^t)/2\tau^2$$
+where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\boldsymbol{\mu}}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.
 
@@
 \section*{References}
-
-\DeclareRobustCommand{\VANDER}[3]{#3}
-\DeclareRobustCommand{\VAN}[3]{#3}
-\DeclareRobustCommand{\DEN}[3]{#3}
 

### 2019-07-26 Manuscript/Plots/circregline.tex
Manuscript update
Update plots

@@
-% Created by tikzDevice version 0.11 on 2019-01-14 18:24:38
+% Created by tikzDevice version 0.12 on 2019-06-18 13:58:30
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,162.86) --
-	( 65.83,162.75) --
-	( 69.16,162.64) --
-	( 72.48,162.53) --
-	( 75.81,162.41) --
-	( 79.13,162.29) --
-	( 82.46,162.17) --
-	( 85.79,162.05) --
-	( 89.11,161.92) --
-	( 92.44,161.79) --
-	( 95.77,161.66) --
-	( 99.09,161.52) --
-	(102.42,161.39) --
-	(105.74,161.25) --
-	(109.07,161.10) --
-	(112.40,160.96) --
-	(115.72,160.81) --
-	(119.05,160.65) --
-	(122.37,160.50) --
-	(125.70,160.34) --
-	(129.03,160.17) --
-	(132.35,160.00) --
-	(135.68,159.83) --
-	(139.01,159.66) --
-	(142.33,159.48) --
-	(145.66,159.29) --
-	(148.98,159.11) --
-	(152.31,158.92) --
-	(155.64,158.72) --
-	(158.96,158.52) --
-	(162.29,158.31) --
-	(165.61,158.10) --
-	(168.94,157.89) --
-	(172.27,157.67) --
-	(175.59,157.44) --
-	(178.92,157.21) --
-	(182.24,156.97) --
-	(185.57,156.73) --
-	(188.90,156.48) --
-	(192.22,156.23) --
-	(195.55,155.97) --
-	(198.88,155.71) --
-	(202.20,155.44) --
-	(205.53,155.16) --
-	(208.85,154.88) --
-	(212.18,154.59) --
-	(215.51,154.29) --
-	(218.83,153.99) --
-	(222.16,153.68) --
-	(225.48,153.37) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,185.34) --
+	( 65.83,185.26) --
+	( 69.16,185.19) --
+	( 72.48,185.10) --
+	( 75.81,185.02) --
+	( 79.13,184.93) --
+	( 82.46,184.84) --
+	( 85.79,184.74) --
+	( 89.11,184.64) --
+	( 92.44,184.53) --
+	( 95.77,184.42) --
+	( 99.09,184.30) --
+	(102.42,184.17) --
+	(105.74,184.04) --
+	(109.07,183.90) --
+	(112.40,183.75) --
+	(115.72,183.60) --
+	(119.05,183.43) --
+	(122.37,183.25) --
+	(125.70,183.07) --
+	(129.03,182.87) --
+	(132.35,182.65) --
+	(135.68,182.42) --
+	(139.01,182.17) --
+	(142.33,181.91) --
+	(145.66,181.62) --
+	(148.98,181.31) --
+	(152.31,180.98) --
+	(155.64,180.61) --
+	(158.96,180.21) --
+	(162.29,179.78) --
+	(165.61,179.30) --
+	(168.94,178.78) --
+	(172.27,178.20) --
+	(175.59,177.55) --
+	(178.92,176.84) --
+	(182.24,176.05) --
+	(185.57,175.16) --
+	(188.90,174.17) --
+	(192.22,173.06) --
+	(195.55,171.82) --
+	(198.88,170.44) --
+	(202.20,168.90) --
+	(205.53,167.21) --
+	(208.85,165.37) --
+	(212.18,163.40) --
+	(215.51,161.33) --
+	(218.83,159.21) --
+	(222.16,157.09) --
+	(225.48,155.02) --
 	(228.81,153.05) --
-	(232.14,152.72) --
-	(235.46,152.38) --
-	(238.79,152.04) --
-	(242.11,151.70) --
-	(245.44,151.34) --
-	(248.77,150.98) --
-	(252.09,150.62) --
-	(255.42,150.24) --
-	(258.75,149.87) --
-	(262.07,149.48) --
-	(265.40,149.09) --
-	(268.72,148.70) --
-	(272.05,148.30) --
-	(275.38,147.89) --
-	(278.70,147.48) --
-	(282.03,147.07) --
-	(285.35,146.65) --
-	(288.68,146.23) --
-	(292.01,145.80) --
-	(295.33,145.37) --
-	(298.66,144.94) --
-	(301.98,144.51) --
-	(305.31,144.08) --
-	(308.64,143.64) --
-	(311.96,143.20) --
-	(315.29,142.77) --
-	(318.61,142.33) --
-	(321.94,141.90) --
-	(325.27,141.46) --
-	(328.59,141.03) --
-	(331.92,140.60) --
-	(335.25,140.18) --
-	(338.57,139.75) --
-	(341.90,139.33) --
-	(345.22,138.91) --
-	(348.55,138.50) --
-	(351.88,138.09) --
-	(355.20,137.69) --
-	(358.53,137.29) --
-	(361.85,136.89) --
-	(365.18,136.50) --
-	(368.51,136.12) --
-	(371.83,135.74) --
-	(375.16,135.37) --
-	(378.48,135.01) --
-	(381.81,134.65) --
-	(385.14,134.30) --
-	(388.46,133.95) --
-	(391.79,133.61) --
-	(395.12,133.28);
+	(232.14,151.20) --
+	(235.46,149.50) --
+	(238.79,147.96) --
+	(242.11,146.57) --
+	(245.44,145.33) --
+	(248.77,144.21) --
+	(252.09,143.22) --
+	(255.42,142.33) --
+	(258.75,141.53) --
+	(262.07,140.81) --
+	(265.40,140.17) --
+	(268.72,139.59) --
+	(272.05,139.06) --
+	(275.38,138.58) --
+	(278.70,138.14) --
+	(282.03,137.74) --
+	(285.35,137.38) --
+	(288.68,137.04) --
+	(292.01,136.73) --
+	(295.33,136.44) --
+	(298.66,136.18) --
+	(301.98,135.93) --
+	(305.31,135.70) --
+	(308.64,135.48) --
+	(311.96,135.28) --
+	(315.29,135.09) --
+	(318.61,134.92) --
+	(321.94,134.75) --
+	(325.27,134.59) --
+	(328.59,134.44) --
+	(331.92,134.30) --
+	(335.25,134.17) --
+	(338.57,134.05) --
+	(341.90,133.93) --
+	(345.22,133.81) --
+	(348.55,133.71) --
+	(351.88,133.60) --
+	(355.20,133.51) --
+	(358.53,133.41) --
+	(361.85,133.32) --
+	(365.18,133.24) --
+	(368.51,133.16) --
+	(371.83,133.08) --
+	(375.16,133.00) --
+	(378.48,132.93) --
+	(381.81,132.86) --
+	(385.14,132.80) --
+	(388.46,132.73) --
+	(391.79,132.67) --
+	(395.12,132.61);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.02,158.22) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.02,181.14) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.85,159.69) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.85,182.30) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.61,151.61) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.61,177.89) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.98,148.39) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.98,175.77) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.79,155.75) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.79,170.46) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.86,157.39) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.86,159.98) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.16,155.69) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.16,171.69) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.66,171.40) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.66,182.72) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.04,158.89) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.04,167.90) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,150.31) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,176.94) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.55,161.39) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.55,171.92) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.63,160.68) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.63,188.28) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.52,155.42) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.52,153.26) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.11,155.35) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.11,182.96) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.48,157.22) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.48,175.85) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.64,163.99) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.64,180.05) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.89,164.74) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.89,183.01) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (165.20,164.22) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (165.20,180.92) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.49,164.42) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.49,182.67) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.99,160.22) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.99,182.74) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.12,163.21) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.12,178.44) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.64,161.85) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.64,173.37) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.92,171.78) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.92,183.94) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.25,165.01) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.25,183.47) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.43,151.21) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.43,167.35) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.92,167.18) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.92,185.52) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.59,162.38) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.59,173.86) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.33,161.26) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (156.33,180.40) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.60,150.13) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.60,164.35) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.14,166.79) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.14,180.48) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.94,161.45) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.94,169.03) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.22,157.60) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.22,173.59) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.57,153.31) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.57,163.99) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.86,169.09) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.86,184.49) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.65,150.41) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.65,152.49) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.15,172.02) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.15,185.40) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.67,150.12) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.67,169.72) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.20,143.33) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.20,149.37) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.89,157.54) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.89,181.79) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.66,159.68) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.66,176.92) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.36,163.30) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.36,175.75) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.12,157.27) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.12,173.04) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.57,175.94) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.57,188.78) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.36,159.37) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.36,180.78) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.14,158.59) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.14,183.16) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.39,158.49) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.39,178.95) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.67,146.33) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.67,176.97) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.06,163.79) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.06,183.13) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.81,161.49) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.81,179.99) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.25,166.02) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.25,186.11) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.90,165.72) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.90,184.75) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.59,157.86) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.59,181.45) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.12,162.13) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (191.12,173.55) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.59,158.84) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.59,184.34) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.66,155.93) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (199.66,177.20) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.84,178.08) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.84,187.31) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.54,155.29) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.54,176.65) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,147.04) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,149.23) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.85,151.23) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.85,157.31) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.94,158.72) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.94,181.61) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.33,159.32) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.33,181.55) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.88,160.11) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.88,186.63) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.29,167.10) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.29,181.09) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.90,170.31) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.90,183.49) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (183.37,163.10) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (183.37,172.89) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.59,154.67) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.59,170.03) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.97,165.31) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.97,183.03) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.72,166.89) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.72,181.12) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.41,159.43) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.41,178.76) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.55,174.39) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.55,185.90) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.62,152.08) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.62,149.31) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.11,154.69) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.11,148.13) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.82,144.67) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.82,148.51) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.42,154.56) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.42,176.11) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.54,161.75) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.54,186.58) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.53,156.02) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.53,181.81) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.38,158.24) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.38,175.85) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.53,162.99) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.53,186.10) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.19,151.16) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.19,173.93) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.74,168.56) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.74,186.43) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.69,137.52) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.69,135.77) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.75,150.38) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.75,185.95) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.77,160.14) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.77,180.96) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.15,158.52) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.15,179.11) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.03,161.22) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.03,182.70) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.77,164.82) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.77,184.21) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.56,159.90) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.56,180.51) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.54,156.59) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.54,155.92) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.41,153.00) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.41,177.08) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.75,142.23) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.75,140.77) circle (  1.13);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.58,157.89) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.58,174.58) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.28,151.60) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.28,179.11) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.36,159.44) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.36,181.34) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.36,163.61) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.36,177.73) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.85,161.55) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (122.85,181.60) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.09,164.98) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.09,183.50) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.39,146.50) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.39,146.22) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.11,160.63) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.11,177.33) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.66,161.53) circle (  1.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.66,183.51) circle (  1.13);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.22,151.12) rectangle (227.72,155.62);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.86,145.14) rectangle (242.36,149.64);
 \end{scope}

### 2019-07-26 Manuscript/Plots/survivaldiffSE.tex
Manuscript update
Update plots

@@
-% Created by tikzDevice version 0.11 on 2019-01-11 16:36:22
+% Created by tikzDevice version 0.12 on 2019-06-13 11:45:21
 % !TEX encoding = UTF-8 Unicode
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {strength IPC};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {IPC intensity};
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {P(strength IPC)};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {P(IPC intensity)};
 \end{scope}

### 2019-08-12 Manuscript/Manuscript.Rmd
Update manuscript + comments heleen

@@
 ```
-In the social sciences the use of cylindrical data is very common. Such data
-consist of a linear and a circular component. @gurtman2011reasoning refers to
-such data as vectors, with a directional measure (i.e., the circular component)
-and a measure indicating the magnitude (i.e., the linear component). Many
-established models in psychology are often referred to as circular or circumplex
-models, but those models are cylindrical. Examples of such cylindrical models
-are the interpersonal circle/circumplex
+In the social sciences the use of cylindrical data is very common. Cylindrical
+data consist of a linear and a circular component. @gurtman2011reasoning refers
+to cylindrical data as vectors, with a directional measure (i.e., the circular
+component) and a measure indicating the magnitude (i.e., the linear component).
+Many established models in psychology are often referred to as circular or
+circumplex models, but those models are cylindrical. Examples of such
+cylindrical models are the interpersonal circle/circumplex
 [@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
@@
 
-\hfil [Figure \ref{circline} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{circline} about here] \hfil
 
@@
 1*\frac{180^\circ}{\pi}$. A circular variable is usually described using its
-first two moments, circular location or mean $\mu$ and mean resultant length
-$\rho$\footnote{In the von Mises distribution, a common distribution for a
+first two moments; the circular location or mean $\mu$ (first moment) and mean resultant length
+$\rho$ (second moment)\footnote{In the von Mises distribution, a common distribution for a
 circular outcome, we have a concentration parameter $\kappa$ that is related to
@@
 
-\hfil [Table \ref{TableModels} about here] \hfil
-
-\subsubsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+\hfil \hspace{2cm}[Table \ref{TableModels} about here] \hfil
+
+\subsubsection{The modified Circular-Linear Projected Normal (CL-PN) and modified Circular-Linear General Projected Normal (CL-GPN) models}\label{CL-(G)PN}
 Following @mastrantonio2015bayesian we consider two models where the prediction
@@
 
-\hfil [Figure \ref{projection} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{projection} about here] \hfil
 
@@
 
-\hfil [Figure \ref{circregline} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{circregline} about here] \hfil
 
@@
 
-\hfil [Figure \ref{plotGPN} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{plotGPN} about here] \hfil
 
@@
 
-\subsubsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+\subsubsection{The modified Abe-Ley model}\label{WeiSSVM}
 This model is an extension of the cylindrical model introduced in
@@
 
-\subsubsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
+\subsubsection{The modified Joint Projected and Skew Normal model (GPN-SSN)}\label{CL-GPN_multivariate}
 This model is an extension of the cylindrical model introduced by
@@
 \centering
-\includegraphics[width = 0.8\textwidth]{Plots/IPC-T.png}
+\includegraphics[width = 0.8\textwidth]{Plots/IPC-T2.png}
 \caption{The interpersonal circle for teachers (IPC-T). The words presented in
@@
 student perceptions of their teachers' interpersonal behavior and contains items
-that load on two interpersonal dimensions: Agency and Communion. Agency refers
-to the degree of power or control a teacher exerts in interaction with his/her
-students. Communion refers to the degree of friendliness or affiliation a
-teacher conveys in interaction with his/her students. The loadings on the two
-dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. This space is
-called the interpersonal circle/circumplex (IPC) and different parts of this
-space are characterized by different teacher behavior, e.g. 'helpful' or
-'uncertain'. The IPC is ``a continuous order with no beginning or end''
-[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
-the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular variable. This
-circular variable represents location on the IPC of the interpersonal behavior
-that a teacher shows towards his/her students.
-
-\hfil [Figure \ref{QTI} about here] \hfil
-
-\indent @Cremers2018Assessing explain the circular nature of the IPC data and
-analyze them as such using a circular regression model. The two dimension
+that load on two interpersonal dimensions: Agency and Communion. The item scores
+on the QTI are measured on a scale from 1 to 5 and are later converted to a
+scale ranging from -1 to 1. Subsequently they are converted to scores on Agency
+and Communion (see also footnote 2 of @wubbels2005two for more details on this
+procedure). Agency refers to the degree of power or control a teacher exerts in
+interaction with his/her students. Communion refers to the degree of
+friendliness or affiliation a teacher conveys in interaction with his/her
+students.  The loadings on the two dimensions of the QTI can be placed in a
+two-dimensional space formed by Agency (vertical) and Communion (horizontal),
+see Figure \ref{QTI}. This space is called the interpersonal circle/circumplex
+(IPC) and different parts of the IPC are characterized by different teacher
+behaviors, e.g. 'helpful' or 'uncertain'. The IPC is ``a continuous order with no
+beginning or end'' [@gurtman2009exploring, p. 2]. We call such ordering a
+circumplex ordering and the IPC is therefore often called the interpersonal
+circumplex. The circumplex ordering also implies that scores on the IPC could be viewed as
+a circular variable. This circular variable represents location on the IPC of
+the interpersonal behavior that a teacher shows towards his/her students.
+
+\hfil \hspace{2cm} [Figure \ref{QTI} about here] \hfil
+
+\indent @Cremers2018Assessing  and @cremers2019circular explain the circular nature of the IPC data and analyze them as such using a circular regression and mixed-effects model respectively.
+The two dimension
 scores, Agency and Communion, can be converted to a circular score using the
@@
 dimension\footnote{The selection of the origin in circumplex data depends on
-the scaling of the Agency and Communion scores. Agency and Communion are
-measured on a scale from 1 to 5 and for analysis purposes they are later
-converted to a scale ranging from -1 to 1.  Their respective 0 scores form the
+the scaling of the Agency and Communion scores. Their respective 0 scores form the
 origin.
@@
 @wubbels2006interpersonal). There is however one main drawback: when
-two-dimensional data are converted to the circle we lose some information,
-namely the length of the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its
-Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
-intensity of the interpersonal behavior a teacher shows towards his/her
-students. In a cylindrical model this intensity (the linear component) can be
-modeled together with the location of interpersonal behavior of a teacher on the
-IPC (the circular component). This leads to an improved analysis of interpersonal
-circumplex data, over either analyzing the two dimensions separately or using a
-circular model, because we take all information, circular and linear, into
-account. In the next section we introduce several cylindrical models that can be
-used to analyze the teacher data. First however we will provide descriptives for
-the teacher data and conduct a 'standard' analysis for these data that we can compare the results from the cylindrical models to.
+two-dimensional data are converted to the location on the circle we lose some
+information, namely the length of the two-dimensional vector \((A, C)^t\),
+\emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length
+represents the intensity of the interpersonal behavior a teacher shows towards
+his/her students. In a cylindrical model this intensity (the linear component)
+can be modeled together with the location of interpersonal behavior of a teacher
+on the IPC (the circular component). This leads to an improved analysis of
+interpersonal circumplex data, over either analyzing the two dimensions
+separately or using a circular model, because we take all information, circular
+and linear, into account. In the next section we introduce several cylindrical
+models that can be used to analyze the teacher data. First however we will
+provide descriptives for the teacher data and conduct a 'standard' analysis for
+these data that we can compare the results from the cylindrical models to.
 
@@
 of interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
-After listwise deletion of missings ($3$ in total, only for the self-efficacy)
-we have a sample of 148 teachers. Note that we remove the missings here for
-simplicity, however, in general and especially when the number of missings is
-larger, the influence of listwise deletion should be investigated and necessary
-precaution (e.g. imputation of missing values) should be taken. Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset. Note that we
-also show the scores on the Agency and Communion component of the IPC before
-transformation to a circular score. For the circular variable IPC we show sample
-estimates for the circular mean $\bar{\theta}$ and mean resultant length
+After listwise deletion of missings ($3$ in total, only for self-efficacy) we
+have a sample of 148 teachers. We used listwise deletion of the missings in this
+paper for simplicity. However, in general and especially when the number of
+missings is larger, the influence of listwise deletion should be investigated
+and necessary precaution (e.g. imputation of missing values) should be taken.
+Table \ref{Tableteacherdescriptives} shows descriptives for the dataset. Note
+that we also show the scores on the Agency and Communion component of the IPC
+before transformation to a circular score. For the circular variable IPC we show
+sample estimates for the circular mean $\bar{\theta}$ and mean resultant length
 $\hat{\rho}$. For the linear variables (Agency, Communion, IPC intensity and the
@@
 
-\hfil [Table \ref{Tableteacherdescriptives} about here] \hfil
-
-\hfil [Figure \ref{dataplot} about here] \hfil
+\hfil \hspace{2cm} [Table \ref{Tableteacherdescriptives} about here] \hfil
+
+\hfil \hspace{2cm} [Figure \ref{dataplot} about here] \hfil
 
@@
 0.09 increase in Communion. Note that in this setup we can only model the effect
-of self-efficacy on the intensity (size of the score) on the Agency and
+of self-efficacy on the intensity, the size of the score, of the Agency and
 Communion component. We can not quantify the effect on the location and
@@
 
-\hfil [Table \ref{Tableteacherstandardanalysis} about here] \hfil
+\hfil \hspace{2cm} [Table \ref{Tableteacherstandardanalysis} about here] \hfil
 
@@
 
-\hfil [Table \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} about here] \hfil
+\hfil \hspace{2cm} [Table \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} about here] \hfil
 
@@
 
-\hfil [Figure \ref{reglineweib} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{reglineweib} about here] \hfil
 
@@
 
-\hfil [Table \ref{tab:means} about here] \hfil
+\hfil \hspace{2cm} [Table \ref{tab:means} about here] \hfil
 
@@
 
-\hfil [Table \ref{tab:ModelFit} about here] \hfil
+\hfil \hspace{2cm} [Table \ref{tab:ModelFit} about here] \hfil
 
@@
 
-\hfil [Figure \ref{regline} about here] \hfil
+\hfil \hspace{2cm} [Figure \ref{regline} about here] \hfil
 
@@
 separately. Such an approach only provides information about the intensity of a
-teachers' score on Agency and Communion and the information about the
+teachers' score on Agency and Communion separately and the information about the
 combination of Agency and Communion, which describes the location on the IPC,
@@
 presented in this paper are also of use for research on human navigation and
-eye-tracking research. Furthermore, even though cylindrical models are already
-used in fields outside of psychology, the addition of a regression structure to
-the models is of use in these fields as well. \newline
+eye-tracking research. For example, in a study by @pannasch2008duration, the relation between
+saccade amplitudes and fixation duration is investigated in different
+experimental settings. The saccade amplitude is a circular outcome that measures
+the turning angle of the eye between fixations on different focal points, e.g.,
+an individual first looks at one object in a picture and then moves focus to
+another object. The outcome variables of interest are thus cylindrical in nature
+(saccade amplitude is a circular outcome and fixation duration a linear one).
+More importantly however, the researchers are interested in both the relation
+between the circular and linear outcome and how covariates (different
+experimental settings) affect the outcomes. These research questions can be
+answered using the cylindrical models introduced in this paper. Furthermore,
+even though cylindrical models are already used in fields outside of psychology,
+the addition of a regression structure to the models is of use in these fields
+as well. \newline
 \indent Out of the four cylindrical models investigated in this paper, the
@@
 
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+
 \newpage

### 2019-08-12 Manuscript/Manuscript.tex
Update manuscript + comments heleen

@@
 
-In the social sciences the use of cylindrical data is very common. Such data
-consist of a linear and a circular component. Gurtman (2011) refers to
-such data as vectors, with a directional measure (i.e., the circular component)
-and a measure indicating the magnitude (i.e., the linear component). Many
-established models in psychology are often referred to as circular or circumplex
-models, but those models are cylindrical. Examples of such cylindrical models
-are the interpersonal circle/circumplex
+In the social sciences the use of cylindrical data is very common. Cylindrical
+data consist of a linear and a circular component. Gurtman (2011) refers
+to cylindrical data as vectors, with a directional measure (i.e., the circular
+component) and a measure indicating the magnitude (i.e., the linear component).
+Many established models in psychology are often referred to as circular or
+circumplex models, but those models are cylindrical. Examples of such
+cylindrical models are the interpersonal circle/circumplex
 (Leary, 1957; Wiggins, 1996; Wubbels, Brekelmans, \DEN{Brok}{Den}{den} Brok, \& \VAN{Tartwijk}{Van}{van} Tartwijk, 2006), the circumplex of
@@
 
-\hfil [Figure \ref{circline} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{circline} about here{]} \hfil
 
@@
 converted to degrees using the relation \(1 \:\text{rad} = 1*\frac{180^\circ}{\pi}\). A circular variable is usually described using its
-first two moments, circular location or mean \(\mu\) and mean resultant length
-\(\rho\)\footnote{In the von Mises distribution, a common distribution for a
+first two moments; the circular location or mean \(\mu\) (first moment) and mean resultant length
+\(\rho\) (second moment)\footnote{In the von Mises distribution, a common distribution for a
 circular outcome, we have a concentration parameter $\kappa$ that is related to
@@
 
-\hfil [Table \ref{TableModels} about here] \hfil
-
-\subsubsection{The Modified Circular-Linear Projected Normal (CL-PN) and Modified Circular-Linear General Projected Normal (CL-GPN) Models}\label{CL-(G)PN}
+\hfil \hspace{2cm}{[}Table \ref{TableModels} about here{]} \hfil
+
+\subsubsection{The modified Circular-Linear Projected Normal (CL-PN) and modified Circular-Linear General Projected Normal (CL-GPN) models}\label{CL-(G)PN}
 
@@
 
-\hfil [Figure \ref{projection} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{projection} about here{]} \hfil
 
@@
 
-\hfil [Figure \ref{circregline} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{circregline} about here{]} \hfil
 
@@
 
-\hfil [Figure \ref{plotGPN} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{plotGPN} about here{]} \hfil
 
@@
 
-\subsubsection{The Modified Abe-Ley Model}\label{WeiSSVM}
+\subsubsection{The modified Abe-Ley model}\label{WeiSSVM}
 
@@
 
-\subsubsection{Modified Joint Projected and Skew Normal Model (GPN-SSN)}\label{CL-GPN_multivariate}
+\subsubsection{The modified Joint Projected and Skew Normal model (GPN-SSN)}\label{CL-GPN_multivariate}
 
@@
 student perceptions of their teachers' interpersonal behavior and contains items
-that load on two interpersonal dimensions: Agency and Communion. Agency refers
-to the degree of power or control a teacher exerts in interaction with his/her
-students. Communion refers to the degree of friendliness or affiliation a
-teacher conveys in interaction with his/her students. The loadings on the two
-dimensions of the QTI can be placed in a two-dimensional space formed by Agency
-(vertical) and Communion (horizontal), see Figure \ref{QTI}. This space is
-called the interpersonal circle/circumplex (IPC) and different parts of this
-space are characterized by different teacher behavior, e.g.~\enquote{helpful} or
-\enquote{uncertain}. The IPC is ``a continuous order with no beginning or end''
-(Gurtman, 2009, p. 2). We call such ordering a circumplex ordering and
-the IPC is therefore often called the interpersonal circumplex. The ordering
-also implies that scores on the IPC could be viewed as a circular variable. This
-circular variable represents location on the IPC of the interpersonal behavior
-that a teacher shows towards his/her students.
-
-\hfil [Figure \ref{QTI} about here] \hfil
-
-\indent Cremers et al. (2018a) explain the circular nature of the IPC data and
-analyze them as such using a circular regression model. The two dimension
+that load on two interpersonal dimensions: Agency and Communion. The item scores
+on the QTI are measured on a scale from 1 to 5 and are later converted to a
+scale ranging from -1 to 1. Subsequently they are converted to scores on Agency
+and Communion (see also footnote 2 of Wubbels and Brekelmans (2005) for more details on this
+procedure). Agency refers to the degree of power or control a teacher exerts in
+interaction with his/her students. Communion refers to the degree of
+friendliness or affiliation a teacher conveys in interaction with his/her
+students. The loadings on the two dimensions of the QTI can be placed in a
+two-dimensional space formed by Agency (vertical) and Communion (horizontal),
+see Figure \ref{QTI}. This space is called the interpersonal circle/circumplex
+(IPC) and different parts of the IPC are characterized by different teacher
+behaviors, e.g.~\enquote{helpful} or \enquote{uncertain}. The IPC is ``a continuous order with no
+beginning or end'' (Gurtman, 2009, p. 2). We call such ordering a
+circumplex ordering and the IPC is therefore often called the interpersonal
+circumplex. The circumplex ordering also implies that scores on the IPC could be viewed as
+a circular variable. This circular variable represents location on the IPC of
+the interpersonal behavior that a teacher shows towards his/her students.
+
+\hfil \hspace{2cm} {[}Figure \ref{QTI} about here{]} \hfil
+
+\indent Cremers et al.
+(2018a) and Cremers, Pennings, Mainhard, and Klugkist (2019) explain the circular nature of the IPC data and analyze them as such using a circular regression and mixed-effects model respectively.
+The two dimension
 scores, Agency and Communion, can be converted to a circular score using the
@@
 Wubbels et al. (2006)). There is however one main drawback: when
-two-dimensional data are converted to the circle we lose some information,
-namely the length of the two-dimensional vector \((A, C)^t\), \emph{i.e.}, its
-Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length represents the
-intensity of the interpersonal behavior a teacher shows towards his/her
-students. In a cylindrical model this intensity (the linear component) can be
-modeled together with the location of interpersonal behavior of a teacher on the
-IPC (the circular component). This leads to an improved analysis of interpersonal
-circumplex data, over either analyzing the two dimensions separately or using a
-circular model, because we take all information, circular and linear, into
-account. In the next section we introduce several cylindrical models that can be
-used to analyze the teacher data. First however we will provide descriptives for
-the teacher data and conduct a \enquote{standard} analysis for these data that we can compare the results from the cylindrical models to.
+two-dimensional data are converted to the location on the circle we lose some
+information, namely the length of the two-dimensional vector \((A, C)^t\),
+\emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\). This length
+represents the intensity of the interpersonal behavior a teacher shows towards
+his/her students. In a cylindrical model this intensity (the linear component)
+can be modeled together with the location of interpersonal behavior of a teacher
+on the IPC (the circular component). This leads to an improved analysis of
+interpersonal circumplex data, over either analyzing the two dimensions
+separately or using a circular model, because we take all information, circular
+and linear, into account. In the next section we introduce several cylindrical
+models that can be used to analyze the teacher data. First however we will
+provide descriptives for the teacher data and conduct a \enquote{standard} analysis for
+these data that we can compare the results from the cylindrical models to.
 
@@
 of interpersonal interactions (Locke \& Sadler, 2007; \VANDER{Want}{Van der}{van der} Want et al., 2018).
-After listwise deletion of missings (\(3\) in total, only for the self-efficacy)
-we have a sample of 148 teachers. Note that we remove the missings here for
-simplicity, however, in general and especially when the number of missings is
-larger, the influence of listwise deletion should be investigated and necessary
-precaution (e.g.~imputation of missing values) should be taken. Table
-\ref{Tableteacherdescriptives} shows descriptives for the dataset. Note that we
-also show the scores on the Agency and Communion component of the IPC before
-transformation to a circular score. For the circular variable IPC we show sample
-estimates for the circular mean \(\bar{\theta}\) and mean resultant length
+After listwise deletion of missings (\(3\) in total, only for self-efficacy) we
+have a sample of 148 teachers. We used listwise deletion of the missings in this
+paper for simplicity. However, in general and especially when the number of
+missings is larger, the influence of listwise deletion should be investigated
+and necessary precaution (e.g.~imputation of missing values) should be taken.
+Table \ref{Tableteacherdescriptives} shows descriptives for the dataset. Note
+that we also show the scores on the Agency and Communion component of the IPC
+before transformation to a circular score. For the circular variable IPC we show
+sample estimates for the circular mean \(\bar{\theta}\) and mean resultant length
 \(\hat{\rho}\). For the linear variables (Agency, Communion, IPC intensity and the
@@
 
-\hfil [Table \ref{Tableteacherdescriptives} about here] \hfil
-
-\hfil [Figure \ref{dataplot} about here] \hfil
+\hfil \hspace{2cm} {[}Table \ref{Tableteacherdescriptives} about here{]} \hfil
+
+\hfil \hspace{2cm} {[}Figure \ref{dataplot} about here{]} \hfil
 
@@
 
-\hfil [Table \ref{Tableteacherstandardanalysis} about here] \hfil
+\hfil \hspace{2cm} {[}Table \ref{Tableteacherstandardanalysis} about here{]} \hfil
 
@@
 
-\hfil [Table \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} about here] \hfil
+\hfil \hspace{2cm} {[}Table \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} about here{]} \hfil
 
@@
 
-\hfil [Figure \ref{reglineweib} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{reglineweib} about here{]} \hfil
 
@@
 
-\hfil [Table \ref{tab:means} about here] \hfil
+\hfil \hspace{2cm} {[}Table \ref{tab:means} about here{]} \hfil
 
@@
 
-\hfil [Table \ref{tab:ModelFit} about here] \hfil
+\hfil \hspace{2cm} {[}Table \ref{tab:ModelFit} about here{]} \hfil
 
@@
 
-\hfil [Figure \ref{regline} about here] \hfil
+\hfil \hspace{2cm} {[}Figure \ref{regline} about here{]} \hfil
 
@@
 presented in this paper are also of use for research on human navigation and
-eye-tracking research. Furthermore, even though cylindrical models are already
-used in fields outside of psychology, the addition of a regression structure to
-the models is of use in these fields as well. \newline
+eye-tracking research. For example, in a study by Pannasch, Helmert, Roth, Herbold, and Walter (2008), the relation between
+saccade amplitudes and fixation duration is investigated in different
+experimental settings. The saccade amplitude is a circular outcome that measures
+the turning angle of the eye between fixations on different focal points, e.g.,
+an individual first looks at one object in a picture and then moves focus to
+another object. The outcome variables of interest are thus cylindrical in nature
+(saccade amplitude is a circular outcome and fixation duration a linear one).
+More importantly however, the researchers are interested in both the relation
+between the circular and linear outcome and how covariates (different
+experimental settings) affect the outcomes. These research questions can be
+answered using the cylindrical models introduced in this paper. Furthermore,
+even though cylindrical models are already used in fields outside of psychology,
+the addition of a regression structure to the models is of use in these fields
+as well. \newline
 \indent Out of the four cylindrical models investigated in this paper, the
@@
 
+\DeclareRobustCommand{\VANDER}[3]{#3} 
+\DeclareRobustCommand{\VAN}[3]{#3}
+\DeclareRobustCommand{\DEN}[3]{#3}
+
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+
 \newpage
@@
 
+\leavevmode\hypertarget{ref-cremers2019circular}{}%
+Cremers, J., Pennings, H.
+J.
+M., Mainhard, T., \& Klugkist, I.
+(2019).
+Circular modelling of circumplex measurements for interpersonal behavior.
+\emph{Assessment}, OnlineFirst.
+doi:\href{https://doi.org/10.1177/1073191119858407}{10.1177/1073191119858407}
+
 \leavevmode\hypertarget{ref-fernandez2007models}{}%
@@
 doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}
+
+\leavevmode\hypertarget{ref-pannasch2008duration}{}%
+Pannasch, S., Helmert, J.
+R., Roth, K., Herbold, A.-K., \& Walter, H.
+(2008).
+Visual fixation durations and saccade amplitudes: Shifting relationship in a variety of conditions.
+\emph{Journal of Eye Movement Research}, \emph{2}(2).
+doi:\href{https://doi.org/10.16910/jemr.2.2.4}{10.16910/jemr.2.2.4}
 
@@
 
+\leavevmode\hypertarget{ref-wubbels2005two}{}%
+Wubbels, T., \& Brekelmans, M.
+(2005).
+Two decades of research on teacher--student relationships in class.
+\emph{International Journal of Educational Research}, \emph{43}(1), 6--24.
+doi:\href{https://doi.org/10.1016/j.ijer.2006.03.003}{10.1016/j.ijer.2006.03.003}
+
 \leavevmode\hypertarget{ref-wubbels2006interpersonal}{}%
@@
 (2006).
-Handbook of classroom management: Research practice and contemporary issues.
+An interpersonal perspective on classroom management in secondary classrooms in the Netherlands.
 In C.

### 2019-08-12 Manuscript/Titlepage.Rmd
Update manuscript + comments heleen

---
title: "Regression models for Cylindrical data in Psychology"
author: |
  | Jolien Cremers^[Corresponding author: jcre@sund.ku.dk]$^{\:\:1}$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
  | $^1$Department of Biostatistics, Copenhagen University
  | $^2$TNO
  | $^3$Department of Education, Utrecht University
  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 12pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section*{Acknowledgements}
We thank Irene Klugkist for insightful remarks and comments on the present
paper. Jolien Cremers acknowledges financial support from a Vidi grant awarded
to I. Klugkist from the Dutch Organization for Scientific Research (NWO
452‐12‐010).
